### 2024-04-22 paper.qmd
Initial commit

---
title: "market-justice-preferences"
author: "Equipo EDUMER"
format: html
editor: visual
bibliography: "input/bib/just-ineq-merit.bib"
---

# The socialization of meritocracy and market justice preferences at school

## Introduction

Since its origins, educational institutions have been related to the idea of social mobility and access to better opportunities. Therefore, the consistent evidence of the high level of social reproduction at the school level represents a threat to the promise of education and a meritocratic system [@bourdieu_reproduction_1990]. A large part of the research in this field at an international level has addressed the extent to which the social origin of students affects their academic results and their life opportunities [@vonhippel_test_2019], confirming that schools have severe difficulties in closing the gaps of origin. Besides this socioeconomic perspective on school opportunities, recent research has addressed to what extent inequalities in the school context are also influencing students' perceptions, beliefs, and attitudes: Are social inequalities perceived at the school context? Are they rejected by the students, particularly those who are worst-off in socioeconomic terms? Or, Is there evidence at the school level that social inequalities are tolerated and even justified? [@batruch_belief_2022; @wiederkehr_belief_2015].

In the present paper, we deal with the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system. In particular, we focus on market justice preferences (Lane, 1986), which refer to the preferences for distributing public goods (as health and education) based on criteria such as competition and payment capacity. Although from a rational point of view it could be expected an opposition to market justice by the underprivileged majority,  we argue that in a social environment characterized by a promotion of meritocratic ideals - as the school system - would lead to the opposite: a larger market justice preferences.

Given that the school environment has an important focus on performance, achievement and acknowledgment, meritocracy has been one of the principal concepts used for understanding and even for justifying performance differences among students. Meritocracy is a distributive system based on the belief that people should be rewarded and promoted based on their abilities, knowledge, and achievements (Young, 1958). It is often seen as a way to create equal opportunities and fairness, as individuals can rise to positions of power and influence based on their own merit rather than their background or connections. However, some argue that meritocracy can actually lead to tolerating or even justifying social inequalities, as it can create a hierarchy where those who already have resources and advantages are more likely to succeed. In this regard, a great deal of academic research about meritocracy delves into the assessment of to what extent rewards and privileges in society are related to merit, emphasizing the so-called unfulfillable promise of meritocracy (Mijs, 2016). Complementing this agenda, a second and emerging research area deals with subjective aspects of meritocracy, such as perceptions and beliefs.

\[definición de market justice, welfare marketization, (Lindh)\]

\[contexto Chileno market justice, dar cifras/porcentajes de privatización de salud, educación y pensiones; asociar al tema de la calidad (lo público es peor)\]

The perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society (Duru-Bellat & Tenret, 2012; Castillo et al., 2019). This perception can vary greatly depending on individual experiences, social, economic, and cultural background. Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work. In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality (Lampert, Mijs, etc). Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger justification of social inequalities, as individual achievement would be seen as rewarded and social policies as less necessary (Batruch et al, 2021).

Most of the research that has related meritocratic beliefs to inequality justification so far has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies. Regarding schools, the way in which they deal with unequal conditions of origin has been linked to the \_hidden curriculum\_ (Chafel, 1997, https://doi.org/10.1002/J.2379-3988.1997.TB00004.X), whereby students learn about distributive norms in society and mechanisms of justification of social differences. Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population \[\@batruch_belief_2022;\@wiederkehr_belief_2015\], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger justification of social inequalities, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary (Batruch et al, 2021).

\~

\[introducir distinción sobre percepción de meritocracia a nivel escolar y a nivel de la sociedad\]

\[resaltar 3 focos de innovación: market justice preferences at school level, relación con meritocracia, y esquema de meritocracia en la escuela y en la sociedad\]

The present paper deals with the association between the perception of meritocracy and the justification of social inequalities, with two main focuses. Firstly, it assesses the justification of social inequalities in social policy domains, such as health, pensions, and education. We argue that individuals who perceive more meritocracy would be more willing to justify better services in these domains for those with higher incomes.  Secondly, we focus on the student-age population as we point out that it is possible to track down the origin of meritocratic beliefs (and their consequences) to early socialization processes. To this regard, we take into account the family and the school as two main socialization agencies that play a significant role in the socialization of cultural beliefs by transmitting cultural norms, values, and expectations to young people.

## Justification of inequality

Research on social stratification beliefs, which explore individual perceptions of who deserves what and why (Kluegel & Smith, 1987), highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness. The influence of ideologies (Wegener & Liebig, 1995) and cultural schemas (Homan et al., 2017) is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations. While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature (Castillo, 2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),  there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles. This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades (Centeno & Cohen, 2012; Harvey, 2007), affecting areas such as pensions, health services, and education. The justification of social inequality based on marketype criteria has been conceptualized as the individuals’ adherence to one specific justice evaluation, the market justice, that is, affording legitimacy to the allocation of goods and services based on prices and individuals’ ability to pay (Boltanski & Chiapello- 2007; Lane 1986; Streeck 2012).

Robert E. Lane proposed the underpinnings of market justice, which he differentiated from political justice. For him, “it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets” (1986, p. 384). Following the theory of relative deprivation –a social phenomenon arising when individuals cannot afford what most others in their environment can (Merton 1950)– Lane notes that, in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions. Although empirical research has shown that, contrary to Lane’s observation, relative deprivation, even in market settings, produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action such as protests and revolt (Greitemeyer and Sagioglou 2016; Mishra and Carleton 2015; Smith et al. 2012; Power 2018; Power, Madsen, and Morton 2020), the conceptualization of market justice has been since then closely coupled to the merit principle for allocating outcomes: the claim that the unequal levels of well-being individuals enjoy ought to be, to some extent, a function of their talents and efforts, regardless of their needs or membership, the two latter being the realm of political justice and its closely coupled principles of need –allocating outcomes to those who require them most– and equality –allocating the same outcome to everyone–, which have been at the center of welfarism (see Wilson 2003). The underlying notion of market justice also resonates in its application to welfare regimes, providing a framework for understanding the varied global approaches to managing social services.

The management of social services manifests in varying approaches across nations, with substantial differences in funding and delivery methods (Jensen, 2008; Stoy, 2014). Nordic countries, for example, predominantly employ public agencies to produce and provide social services, funding these through collective taxation and offering them in kind to the majority of citizens. This system prioritizes political justice, placing it above market mechanisms in accessing services. In contrast, other countries rely more heavily on for-profit entities and private funding, where service distribution depends mainly on individual financial capacity of paying user fees, highlighting the influence of market justice in service allocation. The trend toward marketization of welfare services has been growing since the 1980s (Salamon, 1993), and this shift is increasingly evident even in countries where market solutions have traditionally had a minor role in social policy (Sivesind, 2017).

The question arises whether adults and children justify unequal access to welfare services based on market justice principles. Influenced by theories of policy feedback, which suggest that social welfare policies can reinforce (positive feedback) or undermine (negative feedback) previous policy trajectories (Fernandez and Jaime-Castillo 2013; Pierson 2000; Weaver 2010), citizens’ beliefs about market justice are likely also shaped by the institutional and social contexts they encounter. Indeed, the justification of inequality in access to essential services like education and health, based on one’s ability to pay, shows significant variation across countries, as demonstrated by international surveys, although they do not usually include children in their samples. For instance, the 2019 International Social Survey Program (ISSP) provides insights into how adults perceive inequality in accessing welfare services. Figure 1 illustrates the variation in agreement levels regarding whether it is just for individuals with higher incomes to purchase better education or healthcare. Notably, in 18 of the 29 countries analyzed, there is a greater justification for market inequality in healthcare access than in education. Nevertheless, the general sentiment typically ranges from 'somewhat unjust' to 'neither just nor unjust, with mixed feelings.

```{r}
#| fig-cap: "Average of market justice preferences by country"
```

While contemporary policy developments have increasingly embraced the marketization of social welfare in areas such as education, pensions, and health services, questions about public acceptance of these changes persist. Lindh’s (2015) analysis of ISSP 2009 data from 17 OECD countries reveals a general lack of support for market-based distribution of social services, suggesting widespread disapproval of market stratification of essential services. This finding is corroborated by Soler-Martínez and colleagues’ (2023) research from Latinobarómetro 2020 across 18 Latin American countries, where concerns about health and education access predominated over income inequality. These results indicate that reforms toward welfare marketization are typically driven by elite political decisions rather than grassroots demand.

Despite high-income inequality and limited social mobility in Latin America, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region (Bucca, 2016; Chong et al., 2008; Torche, 2014; Salgado and Castillo, 2023). The reliance on private welfare providers and widespread user fees (Molyneux, 2008) adds complexity to this context. Yet, research on children’s justification of market-based inequalities in accessing welfare services remains limited, especially in Latin America, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.

## Meritocracy

The concept of meritocracy frequently appears nowadays when analyzing cultural determinants of social inequalities. In general, it is mentioned as a value associated with justice, as it would link efforts and talents with rewards in an equitable manner. This normative sense is quite far from its original formulation by Michael Young (1958) in the satirical novel "The Rise of Meritocracy", where it ironically represented a mechanism for reproducing the inequalities of origin. The meritocratic ideal had remained relatively unchallenged until a series of recent publications turned into its potential consequences for maintaining social inequality. Perhaps one of the most recent sources in this line is Michael Sandel's "The Tyranny of Merit", where he strongly questions the implications of carrying out the principle of merit in societies that do not guarantee equal opportunities and that generate a feeling of scarce recognition and appreciation of those who receive lesser rewards: "In society's eyes, and perhaps also their own, their work no longer signified a valued contribution to the common good." (Sandel, pp).

Empirical research on meritocracy has increased along with the philosophical-normative discussion on meritocracy in recent years. Particularly from a sociological perspective, meritocracy has been used in research on social mobility to characterize societies with low mobility that threaten the meritocratic ideal (Goldthorpe). More recently, sociology and social psychology research has attended to the subjective aspects vis-a-vis beliefs in meritocracy. The label of beliefs in this realm covers a series of areas, such as attitudes, perceptions, and preferences (Castillo et al), whereby most of the link this subjective dimensions to individual socio-structural factors and  context-level determinants. For instance, some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs (Mijs 2021), and how larger inequality affects meritocratic beliefs (Morris et al 2022). Based on these findings, a research agenda has been reinforced on the legitimizing role of meritocracy, in line with previous studies using the concept of a just world (Lerner, Dalbert) and the theory of system justification (Jost & Major).

How do meritocratic beliefs legitimize inequalities? Empirical studies have used experiments and surveys to address this question. For instance, the evidence suggests that just world beliefs correlate negatively with support for redistributive compensation systems (Frank et al., 2015). Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed (Evans & Kelley, 2018). Almås and colleagues (2020) found that in a relatively unequal society (the United States), the highly educated accept inequality significantly more than the less educated because they perceive inequality as justifiable owing to differences in productivity (i.e., merit), whereas in a relatively equal society (Norway), the less educated accept inequality more, but not significantly more than the highly educated because meritocratic values are less prevalent. Barr and Miller (2020) also addressed this triple interaction between the level of inequality in a society, the individual level of education, and the perceived origin of the disparity (either by luck or effort) to determine the extent to which inequality is accepted. They found that the interaction's mechanism varies depending on the compared societies. Finally, García-Sánchez and colleagues (2020), using data for 41 countries from the International Social Survey Programme (ISSP), found that the perceived size of the income gap correlated positively with support for progressive taxation. Still, this association was weaker among those who endorsed meritocratic and equal opportunity beliefs. In the same line, experimental research by Durante & Putterman (2009) subjects support less redistribution when the initial distribution is determined according to task performance.

Research about meritocratic beliefs at school age is rather scarce, leaving a wide research gap as schools are one of the primary socialization institutions where achievement based on merit explains success (Erivwo et al 2021). Nevertheless, it is possible to find some initial works focused on the area of distributive justice at school that are closely related to meritocratic beliefs, as the ones by Resh and Sabbagh (quotes). Using justice in grade obtention as a measure of distributive justice and meritocracy (Sabbagh et al 2006), they find for instance that a larger sense of distributive justice about grades is associated to higher socio-economic status (Resh 2010), have a positive effect on liberal democratic orientation and on trust in people and in formal institutions (Resh & Sabbagh 2013), and tend to refrain from violence and to engage to a greater extent in extra-curricular school activity and community volunteering (Resh & Sabbagh 2017).

## Children's judgments of inequality, schools, and family background

Research indicates various socialization practices at families and schools during childhood and adolescence that impact dispositional and behavioral tendencies concerning justifications of inequality in adult life. The differences in economic understanding across age groups are consistent with research on cognitive development (Choudhury et al., 2006). Adolescents with mature socio-cognitive abilities tend to express stronger preferences for fairness than infants and children (Wynn et al., 2017). As children grow older, they become more likely to behave fairly, with their early-emerging strict egalitarianism being replaced by an increasing endorsement of fairness principles and engagement in collaborative activities (Huppert et al., 2019; McAuliffe et al., 2017). In these activities, their fairness views consider individual contributions, merits, and circumstances (Almås et al., 2010; Huppert et al., 2019; Sigelman & Waitzman, 1991). Engelmann and Tomasello (2019) claim that children’s sense of fairness emerges at three years old, and we can observe it in collaborative activities, where they accept inequality if the procedure gives everyone an equal chance. Therefore, children at this age respond to unequal distributions based on interpersonal concerns, as they already demand equal respect. In any case, between 3 and 8 years of age, inequitable and anti-meritorious allocations are evaluated more negatively, but equitable and meritorious allocations are not evaluated more positively (Elenbaas, 2019).

Some research shows that the social environment in which children develop, such as family and school, is associated with their prosocial behaviors by playing an essential role in the transmission of equity norms (Schunk & Zipperle, 2023; Kosse et al., 2020). In fact, schools contribute to institutionalizing and reproducing inequality by promoting values, norms, practices, and languages familiar to higher-class families because the dominant group’s culture shapes educational institutions (Bourdieu & Passeron, 1990). Middle- and upper-class students are better equipped to face academic challenges and are more familiar with academic expectations (Mikus et al., 2019). Such familiarity represents cultural capital in educational contexts because higher-status students come to school ready to meet these expectations and reap the benefits (Jack, 2016; Khan, 2011). Conversely, lower-status children lacking cultural capital must catch up while experiencing inequitable comparisons (Goudeau & Croizet, 2017). Additionally, academic achievement is treated as the outcome of dispositional factors (e.g., pupils’ efforts and talents or lack of them) rather than the result of differential access to critical resources. Due to the meritocratic frame schools encourage, both low- and high-status individuals believe that students’ success or failure is not due to their family background but rather to differences in efforts and talents (Darnon et al., 2018). In this sense, we believe that the perception of meritocracy can influence students' judgments about market justice preferences. Furthermore, we believe there is a difference between students' perceptions of meritocracy based on their own experience in school and what they perceive in society at large. Consistently, our first two hypotheses are:

H1a: Students who perceive that there is more meritocracy at school will show larger market justice preferences

H1b: Students who perceive that there is more meritocracy in society will show larger market justice preferences

Family background and family socialization practices also contribute to children’s and adolescent’s market justice preferences. For example, Almås and colleagues (2017) found that adolescents from low-socioeconomic-status families are likelier to have an egalitarian fairness view and consider an equal distribution as fair in a situation with unequal merits. The authors speculate that differences in socialization practices across status groups might bring about, to a great extent, the fairness views of children and adolescents because social status seems to interact with these evaluations (e.g., Hvidberg et al., 2023).

The classic work of Kohn showed that middle-class parents value the expression of internal states and emotions, such as self-control, curiosity, happiness, and consideration, while working-class parents promote deference, obedience, and conformity to authority (Kohn, 1963; Kohn & Schooler, 1969). Although parents from all social backgrounds encourage individualism in their children, this shared norm translates into different forms in high and low social classes (1999). Acemoglu (2021) claimed that the values families impart to their children interact with social mobility. Because obedience is a valuable characteristic for employers, in low-wage and social mobility environments, low-income families impart values of obedience to their children to prevent disadvantaging them in labor markets. On the one hand, children from privileged families are socialized to adopt a clear conception of individualism that highlights their internal states, independence, and idiosyncrasies. In contrast, children from disadvantaged families are socialized to support a more balanced view of individualism that considers personal characteristics as resources to overcome collective impediments on the path to upward mobility (Iacoviello & Lorenzi-Cioldi, 2019). In this way, we believe that there are differences in the socialization of values according to socioeconomic differences that could influence market justice preferences and, therefore, we propose the following hypothesis:

H2: Students from families of higher social status will show larger market justice preferences.

Recent empirical research has demonstrated that the institutional design of schools, coupled with the meritocratic ideology it fosters, significantly influences children’s and adolescents’ views on inequality and deservingness. For example, Jonsson and Beach’s (2015) study revealed that higher-status adolescents in Sweden tend to perpetuate social class stereotypes while describing the vocational and academic tracks. Academic track students are depicted as wealthy, intelligent, ambitious, and diligent, while vocational track students are characterized as poor, unambitious, unintelligent, and lackadaisical. These stereotypes help individuals maintain a sense of superiority over others and legitimize the prevailing social hierarchies and economic disparities (Jost & Burgess, 2000).

H3: Students from schools of higher social status will show larger market justice preferences.

H4: Students from schools with higher average levels of academic achievement will show larger market justice preferences.

\[interacciones\]

H5: The perception of meritocracy in school and society will moderate the effect of family social status on market justice preferences.

H6: The perception of meritocracy in school and society will moderate the effect of school status on market justice preferences.

H7: The perception of meritocracy in school and society will moderate the effect of school academic achievement on market justice preferences.

### Summary of hypotheses

```{r}
#| fig-cap: "Summary of hypotheses"
```

## Methods

### Data

The main data source for the analysis is the First Study of Citizenship Education in Chile, carried out by the Education Quality Agency of the Ministry of Education. The application date was November 9, 2017. The target population of this study is Eighth-grade students from 242 schools. In the data, there are 8,589 students and 6,770 parents. This database was analyzed with the R package “ResponsePatterns” to detect possible repetitive and “careless” response patterns and thus contribute to a higher quality of research data (Gottfried, et al 2022). Responses from 171 students and 79 parents were removed, which, when merged, gave a total of 6,270 valid cases.

The analysis of school variables includes data from the Ministry of Education's SIMCE 2017 database. This database contains information at the school level, such as the administrative dependency, its socioeconomic classification, and the achievement scores obtained in the mathematics and language census tests. It is available for free use on the MINEDUC \[web page\].

After eliminating missing cases, the final sample used in the analysis was based on 5,047 students and parents of 231 schools for the dependent variable of access to social services.

### Variables

**Dependent variables**

This study has three dependent variables related to the justification of social inequality in specific policy domains. The first asks whether access to social services should be conditional on income, i,.e., “It is just that in Chile people who can pay have a better education for their children”. Students rated their preferences using the following responses: “strongly disagree”, “Disagree”, “Agree”, and ”strongly agree”. An average index is built with these items (Cronbach's Alpha = 0,86). (Apéndice: items en español y su correspondiente traducción al inglés)

Table 1 shows the items used, their response categories, and their frequencies.

```{r}
#| tbl-cap: "Dependent variables"
```

**Independent variables**

For the primary independent variable, the perception of meritocracy, five items address the perception of rewards according to effort and intelligence at the school and societal levels. At the school level, students answer whether “Intelligence is important to get good grades” and “Effort is important to get good grades”. At the societal level, students respond to the following questions: “In Chile, people are rewarded for their efforts”, “In Chile, people get what they deserve”, and “In Chile, people are rewarded for their intelligence and skills”. Each item was answered on a four-point scale ranging from “Strongly disagree” to “Strongly agree”.

The rest of the independent variables are divided into individual and school levels. At the individual level, family socioeconomic status was measured by the parents' highest educational level and the number of books at home. Likewise, an index of access to technology includes the number of computers, tablets, and cell phones at home, as well as whether there is an Internet connection. Table 2 shows the items used, their response categories, and their frequency.

```{r}
#| tbl-cap: "Independent variables"
```

The school-level variables are the administrative dependency of the school, the socioeconomic classification made by the Ministry of Education, the level of performance in the SIMCE test of the school, and the proportion of parents with university or postgraduate degrees. Table 3 shows the items used, response categories, and frequency.

```{r}
#| tbl-cap: "School context variables"
```

### Methods

The data has a hierarchical structure of students nested in schools, so the model estimation is performed in a multilevel (random effects) framework. This modeling approach lets us correctly estimate individual and contextual effects in a single model. We estimate cumulative link mixed models for the ordinal dependent variables, whereas we use linear mixed effects models for the average index of inequality justification.

The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open ScorrectlyOSF), the access to the document is available at this \[link\](https://doi.org/10.17605/OSF.IO/UFSDV). The statistical analysis of this research was performed using the free software R version 4.1.3.

## Analysis

### Descriptive analysis

```{r}
#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
```

Figure 3 shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram en figure X). On the left, we observe the social meritocracy diagrams, while on the right the school meritocracy diagrams are shown. For the three variables of perception of meritocracy in society the relationship is clear, since the average of market justice preferences increases the more there is agreement that people are rewarded for their effort, merit, and talent. This relationship needs to be clarified in the case of the variables of perception of meritocracy at school. To the extent that there is more agreement that the perception of talent is essential for obtaining good grades, the average of market justice preferences increases, but this relationship is not as clear as with the variables of meritocracy in society. In addition, the graph does not show a clear trend in the relationship between the perception that effort is essential to obtain good grades and market justice preferences.

### 2024-04-22 paper.tex
Initial commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={market-justice-preferences},
  pdfauthor={Equipo EDUMER},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{market-justice-preferences}
\author{Equipo EDUMER}
\date{}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden]}{\end{tcolorbox}}\fi

\hypertarget{the-socialization-of-meritocracy-and-market-justice-preferences-at-school}{%
\section{The socialization of meritocracy and market justice preferences
at
school}\label{the-socialization-of-meritocracy-and-market-justice-preferences-at-school}}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Since its origins, educational institutions have been related to the
idea of social mobility and access to better opportunities. Therefore,
the consistent evidence of the high level of social reproduction at the
school level represents a threat to the promise of education and a
meritocratic system {[}@bourdieu\_reproduction\_1990{]}. A large part of
the research in this field at an international level has addressed the
extent to which the social origin of students affects their academic
results and their life opportunities {[}@vonhippel\_test\_2019{]},
confirming that schools have severe difficulties in closing the gaps of
origin. Besides this socioeconomic perspective on school opportunities,
recent research has addressed to what extent inequalities in the school
context are also influencing students' perceptions, beliefs, and
attitudes: Are social inequalities perceived at the school context? Are
they rejected by the students, particularly those who are worst-off in
socioeconomic terms? Or, Is there evidence at the school level that
social inequalities are tolerated and even justified?
{[}@batruch\_belief\_2022; @wiederkehr\_belief\_2015{]}.~

\end{document}

### 2024-04-23 paper.qmd
update references
faltan aun

@@
 author: "Equipo EDUMER"
-format: html
+format: pdf
 editor: visual
 bibliography: "input/bib/just-ineq-merit.bib"
+link-citations: true
 ---
@@
 Given that the school environment has an important focus on performance, achievement and acknowledgment, meritocracy has been one of the principal concepts used for understanding and even for justifying performance differences among students.
-Meritocracy is a distributive system based on the belief that people should be rewarded and promoted based on their abilities, knowledge, and achievements (Young, 1958).
+Meritocracy is a distributive system based on the belief that people should be rewarded and promoted based on their abilities, knowledge, and achievements [@young_rise_1958].
 It is often seen as a way to create equal opportunities and fairness, as individuals can rise to positions of power and influence based on their own merit rather than their background or connections.
 However, some argue that meritocracy can actually lead to tolerating or even justifying social inequalities, as it can create a hierarchy where those who already have resources and advantages are more likely to succeed.
-In this regard, a great deal of academic research about meritocracy delves into the assessment of to what extent rewards and privileges in society are related to merit, emphasizing the so-called unfulfillable promise of meritocracy (Mijs, 2016).
+In this regard, a great deal of academic research about meritocracy delves into the assessment of to what extent rewards and privileges in society are related to merit, emphasizing the so-called unfulfillable promise of meritocracy [@mijs_stratified_2016].
 Complementing this agenda, a second and emerging research area deals with subjective aspects of meritocracy, such as perceptions and beliefs.
@@
 
-The perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society (Duru-Bellat & Tenret, 2012; Castillo et al., 2019).
+The perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society [@duru-bellat_who_2012; @castillo_meritocracia_2019].
 This perception can vary greatly depending on individual experiences, social, economic, and cultural background.
 Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work.
-In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality (Lampert, Mijs, etc).
-Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger justification of social inequalities, as individual achievement would be seen as rewarded and social policies as less necessary (Batruch et al, 2021).
+In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2021a].
+Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger justification of social inequalities, as individual achievement would be seen as rewarded and social policies as less necessary [@batruch_belief_2022].
 
 Most of the research that has related meritocratic beliefs to inequality justification so far has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies.
-Regarding schools, the way in which they deal with unequal conditions of origin has been linked to the \_hidden curriculum\_ (Chafel, 1997, https://doi.org/10.1002/J.2379-3988.1997.TB00004.X), whereby students learn about distributive norms in society and mechanisms of justification of social differences.
-Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population \[\@batruch_belief_2022;\@wiederkehr_belief_2015\], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger justification of social inequalities, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary (Batruch et al, 2021).
+Regarding schools, the way in which they deal with unequal conditions of origin has been linked to the *hidden curriculum* [@chafel_schooling_1997], whereby students learn about distributive norms in society and mechanisms of justification of social differences.
+Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population [@batruch_belief_2022; @wiederkehr_belief_2015], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger justification of social inequalities, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary [@batruch_belief_2022].
 
@@
 
-Research on social stratification beliefs, which explore individual perceptions of who deserves what and why (Kluegel & Smith, 1987), highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness.
-The influence of ideologies (Wegener & Liebig, 1995) and cultural schemas (Homan et al., 2017) is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations.
-While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature (Castillo, 2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),  there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles.
+Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegelBeliefsInequalityAmericans1987], highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness.
+The influence of ideologies [@wegenerDominantIdeologiesVariation1995] and cultural schemas [@homanBeingBecomingPoor2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations.
+While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castilloLegitimacyInequalityHighly2011; ] (Castillo, 2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),  there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles.
 This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades (Centeno & Cohen, 2012; Harvey, 2007), affecting areas such as pensions, health services, and education.
@@
 The question arises whether adults and children justify unequal access to welfare services based on market justice principles.
-Influenced by theories of policy feedback, which suggest that social welfare policies can reinforce (positive feedback) or undermine (negative feedback) previous policy trajectories (Fernandez and Jaime-Castillo 2013; Pierson 2000; Weaver 2010), citizens’ beliefs about market justice are likely also shaped by the institutional and social contexts they encounter.
+Influenced by theories of policy feedback, which suggest that social welfare policies can reinforce (positive feedback) or undermine (negative feedback) previous policy trajectories [@fernandez_positive_2013] (Fernandez and Jaime-Castillo 2013; Pierson 2000; Weaver 2010), citizens’ beliefs about market justice are likely also shaped by the institutional and social contexts they encounter.
 Indeed, the justification of inequality in access to essential services like education and health, based on one’s ability to pay, shows significant variation across countries, as demonstrated by international surveys, although they do not usually include children in their samples.
@@
 While contemporary policy developments have increasingly embraced the marketization of social welfare in areas such as education, pensions, and health services, questions about public acceptance of these changes persist.
-Lindh’s (2015) analysis of ISSP 2009 data from 17 OECD countries reveals a general lack of support for market-based distribution of social services, suggesting widespread disapproval of market stratification of essential services.
+@lindh_public_2015 analysis of ISSP 2009 data from 17 OECD countries reveals a general lack of support for market-based distribution of social services, suggesting widespread disapproval of market stratification of essential services.
 This finding is corroborated by Soler-Martínez and colleagues’ (2023) research from Latinobarómetro 2020 across 18 Latin American countries, where concerns about health and education access predominated over income inequality.
@@
 
-Despite high-income inequality and limited social mobility in Latin America, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region (Bucca, 2016; Chong et al., 2008; Torche, 2014; Salgado and Castillo, 2023).
+Despite high-income inequality and limited social mobility in Latin America, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@buccaMeritBlameUnequal2016; @chongMysteryDiscriminationLatin2008; @torcheIntergenerationalMobilityInequality2014; @salgadoInequalityStratificationLatin2023].
 The reliance on private welfare providers and widespread user fees (Molyneux, 2008) adds complexity to this context.
@@
 In general, it is mentioned as a value associated with justice, as it would link efforts and talents with rewards in an equitable manner.
-This normative sense is quite far from its original formulation by Michael Young (1958) in the satirical novel "The Rise of Meritocracy", where it ironically represented a mechanism for reproducing the inequalities of origin.
+This normative sense is quite far from its original formulation by @young_rise_1958 in the satirical novel "The Rise of Meritocracy", where it ironically represented a mechanism for reproducing the inequalities of origin.
 The meritocratic ideal had remained relatively unchallenged until a series of recent publications turned into its potential consequences for maintaining social inequality.
-Perhaps one of the most recent sources in this line is Michael Sandel's "The Tyranny of Merit", where he strongly questions the implications of carrying out the principle of merit in societies that do not guarantee equal opportunities and that generate a feeling of scarce recognition and appreciation of those who receive lesser rewards: "In society's eyes, and perhaps also their own, their work no longer signified a valued contribution to the common good." (Sandel, pp).
+Perhaps one of the most recent sources in this line is Michael Sandel's "The Tyranny of Merit", where he strongly questions the implications of carrying out the principle of merit in societies that do not guarantee equal opportunities and that generate a feeling of scarce recognition and appreciation of those who receive lesser rewards: "In society's eyes, and perhaps also their own, their work no longer signified a valued contribution to the common good." [@sandel_tyranny_2020, pp.].
 
 Empirical research on meritocracy has increased along with the philosophical-normative discussion on meritocracy in recent years.
-Particularly from a sociological perspective, meritocracy has been used in research on social mobility to characterize societies with low mobility that threaten the meritocratic ideal (Goldthorpe).
+Particularly from a sociological perspective, meritocracy has been used in research on social mobility to characterize societies with low mobility that threaten the meritocratic ideal [@goldthorpe_myth_2003].
 More recently, sociology and social psychology research has attended to the subjective aspects vis-a-vis beliefs in meritocracy.
 The label of beliefs in this realm covers a series of areas, such as attitudes, perceptions, and preferences (Castillo et al), whereby most of the link this subjective dimensions to individual socio-structural factors and  context-level determinants.
-For instance, some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs (Mijs 2021), and how larger inequality affects meritocratic beliefs (Morris et al 2022).
+For instance, some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2021a], and how larger inequality affects meritocratic beliefs (Morris et al 2022).
 Based on these findings, a research agenda has been reinforced on the legitimizing role of meritocracy, in line with previous studies using the concept of a just world (Lerner, Dalbert) and the theory of system justification (Jost & Major).
@@
 Empirical studies have used experiments and surveys to address this question.
-For instance, the evidence suggests that just world beliefs correlate negatively with support for redistributive compensation systems (Frank et al., 2015).
-Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed (Evans & Kelley, 2018).
-Almås and colleagues (2020) found that in a relatively unequal society (the United States), the highly educated accept inequality significantly more than the less educated because they perceive inequality as justifiable owing to differences in productivity (i.e., merit), whereas in a relatively equal society (Norway), the less educated accept inequality more, but not significantly more than the highly educated because meritocratic values are less prevalent.
-Barr and Miller (2020) also addressed this triple interaction between the level of inequality in a society, the individual level of education, and the perceived origin of the disparity (either by luck or effort) to determine the extent to which inequality is accepted.
+For instance, the evidence suggests that just world beliefs correlate negatively with support for redistributive compensation systems [@frank_performance_2015].
+Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed [@evans_strong_2018].
+@almas_cutthroat_2020 found that in a relatively unequal society (the United States), the highly educated accept inequality significantly more than the less educated because they perceive inequality as justifiable owing to differences in productivity (i.e., merit), whereas in a relatively equal society (Norway), the less educated accept inequality more, but not significantly more than the highly educated because meritocratic values are less prevalent.
+@barr_effect_2020 also addressed this triple interaction between the level of inequality in a society, the individual level of education, and the perceived origin of the disparity (either by luck or effort) to determine the extent to which inequality is accepted.
 They found that the interaction's mechanism varies depending on the compared societies.
-Finally, García-Sánchez and colleagues (2020), using data for 41 countries from the International Social Survey Programme (ISSP), found that the perceived size of the income gap correlated positively with support for progressive taxation.
+Finally, @garcia-sanchez_attitudes_2020, using data for 41 countries from the International Social Survey Programme (ISSP), found that the perceived size of the income gap correlated positively with support for progressive taxation.
 Still, this association was weaker among those who endorsed meritocratic and equal opportunity beliefs.
@@
 Nevertheless, it is possible to find some initial works focused on the area of distributive justice at school that are closely related to meritocratic beliefs, as the ones by Resh and Sabbagh (quotes).
-Using justice in grade obtention as a measure of distributive justice and meritocracy (Sabbagh et al 2006), they find for instance that a larger sense of distributive justice about grades is associated to higher socio-economic status (Resh 2010), have a positive effect on liberal democratic orientation and on trust in people and in formal institutions (Resh & Sabbagh 2013), and tend to refrain from violence and to engage to a greater extent in extra-curricular school activity and community volunteering (Resh & Sabbagh 2017).
+Using justice in grade obtention as a measure of distributive justice and meritocracy (Sabbagh et al 2006), they find for instance that a larger sense of distributive justice about grades is associated to higher socio-economic status [@resh_sense_2010], have a positive effect on liberal democratic orientation and on trust in people and in formal institutions (Resh & Sabbagh 2013), and tend to refrain from violence and to engage to a greater extent in extra-curricular school activity and community volunteering [@resh_sense_2017].
 

### 2024-04-23 01-introduction.qmd
restructurar repo

--- 2024-04-23 paper.qmd
@@
 ---
-title: "market-justice-preferences"
+title: "01-introduction"
 author: "Equipo EDUMER"
@@
 editor: visual
-bibliography: "input/bib/just-ineq-merit.bib"
-link-citations: true
 ---
-
-# The socialization of meritocracy and market justice preferences at school
 
@@
 To this regard, we take into account the family and the school as two main socialization agencies that play a significant role in the socialization of cultural beliefs by transmitting cultural norms, values, and expectations to young people.
-
-## Justification of inequality
-
-Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegelBeliefsInequalityAmericans1987], highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness.
-The influence of ideologies [@wegenerDominantIdeologiesVariation1995] and cultural schemas [@homanBeingBecomingPoor2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations.
-While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castilloLegitimacyInequalityHighly2011; ] (Castillo, 2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),  there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles.
-This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades (Centeno & Cohen, 2012; Harvey, 2007), affecting areas such as pensions, health services, and education.
-The justification of social inequality based on marketype criteria has been conceptualized as the individuals’ adherence to one specific justice evaluation, the market justice, that is, affording legitimacy to the allocation of goods and services based on prices and individuals’ ability to pay (Boltanski & Chiapello- 2007; Lane 1986; Streeck 2012).
-
-Robert E.
-Lane proposed the underpinnings of market justice, which he differentiated from political justice.
-For him, “it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets” (1986, p.
-384).
-Following the theory of relative deprivation –a social phenomenon arising when individuals cannot afford what most others in their environment can (Merton 1950)– Lane notes that, in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions.
-Although empirical research has shown that, contrary to Lane’s observation, relative deprivation, even in market settings, produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action such as protests and revolt (Greitemeyer and Sagioglou 2016; Mishra and Carleton 2015; Smith et al.
-2012; Power 2018; Power, Madsen, and Morton 2020), the conceptualization of market justice has been since then closely coupled to the merit principle for allocating outcomes: the claim that the unequal levels of well-being individuals enjoy ought to be, to some extent, a function of their talents and efforts, regardless of their needs or membership, the two latter being the realm of political justice and its closely coupled principles of need –allocating outcomes to those who require them most– and equality –allocating the same outcome to everyone–, which have been at the center of welfarism (see Wilson 2003).
-The underlying notion of market justice also resonates in its application to welfare regimes, providing a framework for understanding the varied global approaches to managing social services.
-
-The management of social services manifests in varying approaches across nations, with substantial differences in funding and delivery methods (Jensen, 2008; Stoy, 2014).
-Nordic countries, for example, predominantly employ public agencies to produce and provide social services, funding these through collective taxation and offering them in kind to the majority of citizens.
-This system prioritizes political justice, placing it above market mechanisms in accessing services.
-In contrast, other countries rely more heavily on for-profit entities and private funding, where service distribution depends mainly on individual financial capacity of paying user fees, highlighting the influence of market justice in service allocation.
-The trend toward marketization of welfare services has been growing since the 1980s (Salamon, 1993), and this shift is increasingly evident even in countries where market solutions have traditionally had a minor role in social policy (Sivesind, 2017).
-
-The question arises whether adults and children justify unequal access to welfare services based on market justice principles.
-Influenced by theories of policy feedback, which suggest that social welfare policies can reinforce (positive feedback) or undermine (negative feedback) previous policy trajectories [@fernandez_positive_2013] (Fernandez and Jaime-Castillo 2013; Pierson 2000; Weaver 2010), citizens’ beliefs about market justice are likely also shaped by the institutional and social contexts they encounter.
-Indeed, the justification of inequality in access to essential services like education and health, based on one’s ability to pay, shows significant variation across countries, as demonstrated by international surveys, although they do not usually include children in their samples.
-For instance, the 2019 International Social Survey Program (ISSP) provides insights into how adults perceive inequality in accessing welfare services.
-Figure 1 illustrates the variation in agreement levels regarding whether it is just for individuals with higher incomes to purchase better education or healthcare.
-Notably, in 18 of the 29 countries analyzed, there is a greater justification for market inequality in healthcare access than in education.
-Nevertheless, the general sentiment typically ranges from 'somewhat unjust' to 'neither just nor unjust, with mixed feelings.
-
-```{r}
-#| fig-cap: "Average of market justice preferences by country"
-```
-
-While contemporary policy developments have increasingly embraced the marketization of social welfare in areas such as education, pensions, and health services, questions about public acceptance of these changes persist.
-@lindh_public_2015 analysis of ISSP 2009 data from 17 OECD countries reveals a general lack of support for market-based distribution of social services, suggesting widespread disapproval of market stratification of essential services.
-This finding is corroborated by Soler-Martínez and colleagues’ (2023) research from Latinobarómetro 2020 across 18 Latin American countries, where concerns about health and education access predominated over income inequality.
-These results indicate that reforms toward welfare marketization are typically driven by elite political decisions rather than grassroots demand.
-
-Despite high-income inequality and limited social mobility in Latin America, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@buccaMeritBlameUnequal2016; @chongMysteryDiscriminationLatin2008; @torcheIntergenerationalMobilityInequality2014; @salgadoInequalityStratificationLatin2023].
-The reliance on private welfare providers and widespread user fees (Molyneux, 2008) adds complexity to this context.
-Yet, research on children’s justification of market-based inequalities in accessing welfare services remains limited, especially in Latin America, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.
-
-## Meritocracy
-
-The concept of meritocracy frequently appears nowadays when analyzing cultural determinants of social inequalities.
-In general, it is mentioned as a value associated with justice, as it would link efforts and talents with rewards in an equitable manner.
-This normative sense is quite far from its original formulation by @young_rise_1958 in the satirical novel "The Rise of Meritocracy", where it ironically represented a mechanism for reproducing the inequalities of origin.
-The meritocratic ideal had remained relatively unchallenged until a series of recent publications turned into its potential consequences for maintaining social inequality.
-Perhaps one of the most recent sources in this line is Michael Sandel's "The Tyranny of Merit", where he strongly questions the implications of carrying out the principle of merit in societies that do not guarantee equal opportunities and that generate a feeling of scarce recognition and appreciation of those who receive lesser rewards: "In society's eyes, and perhaps also their own, their work no longer signified a valued contribution to the common good." [@sandel_tyranny_2020, pp.].
-
-Empirical research on meritocracy has increased along with the philosophical-normative discussion on meritocracy in recent years.
-Particularly from a sociological perspective, meritocracy has been used in research on social mobility to characterize societies with low mobility that threaten the meritocratic ideal [@goldthorpe_myth_2003].
-More recently, sociology and social psychology research has attended to the subjective aspects vis-a-vis beliefs in meritocracy.
-The label of beliefs in this realm covers a series of areas, such as attitudes, perceptions, and preferences (Castillo et al), whereby most of the link this subjective dimensions to individual socio-structural factors and  context-level determinants.
-For instance, some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2021a], and how larger inequality affects meritocratic beliefs (Morris et al 2022).
-Based on these findings, a research agenda has been reinforced on the legitimizing role of meritocracy, in line with previous studies using the concept of a just world (Lerner, Dalbert) and the theory of system justification (Jost & Major).
-
-How do meritocratic beliefs legitimize inequalities?
-Empirical studies have used experiments and surveys to address this question.
-For instance, the evidence suggests that just world beliefs correlate negatively with support for redistributive compensation systems [@frank_performance_2015].
-Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed [@evans_strong_2018].
-@almas_cutthroat_2020 found that in a relatively unequal society (the United States), the highly educated accept inequality significantly more than the less educated because they perceive inequality as justifiable owing to differences in productivity (i.e., merit), whereas in a relatively equal society (Norway), the less educated accept inequality more, but not significantly more than the highly educated because meritocratic values are less prevalent.
-@barr_effect_2020 also addressed this triple interaction between the level of inequality in a society, the individual level of education, and the perceived origin of the disparity (either by luck or effort) to determine the extent to which inequality is accepted.
-They found that the interaction's mechanism varies depending on the compared societies.
-Finally, @garcia-sanchez_attitudes_2020, using data for 41 countries from the International Social Survey Programme (ISSP), found that the perceived size of the income gap correlated positively with support for progressive taxation.
-Still, this association was weaker among those who endorsed meritocratic and equal opportunity beliefs.
-In the same line, experimental research by Durante & Putterman (2009) subjects support less redistribution when the initial distribution is determined according to task performance.
-
-Research about meritocratic beliefs at school age is rather scarce, leaving a wide research gap as schools are one of the primary socialization institutions where achievement based on merit explains success (Erivwo et al 2021).
-Nevertheless, it is possible to find some initial works focused on the area of distributive justice at school that are closely related to meritocratic beliefs, as the ones by Resh and Sabbagh (quotes).
-Using justice in grade obtention as a measure of distributive justice and meritocracy (Sabbagh et al 2006), they find for instance that a larger sense of distributive justice about grades is associated to higher socio-economic status [@resh_sense_2010], have a positive effect on liberal democratic orientation and on trust in people and in formal institutions (Resh & Sabbagh 2013), and tend to refrain from violence and to engage to a greater extent in extra-curricular school activity and community volunteering [@resh_sense_2017].
-
-## Children's judgments of inequality, schools, and family background
-
-Research indicates various socialization practices at families and schools during childhood and adolescence that impact dispositional and behavioral tendencies concerning justifications of inequality in adult life.
-The differences in economic understanding across age groups are consistent with research on cognitive development (Choudhury et al., 2006).
-Adolescents with mature socio-cognitive abilities tend to express stronger preferences for fairness than infants and children (Wynn et al., 2017).
-As children grow older, they become more likely to behave fairly, with their early-emerging strict egalitarianism being replaced by an increasing endorsement of fairness principles and engagement in collaborative activities (Huppert et al., 2019; McAuliffe et al., 2017).
-In these activities, their fairness views consider individual contributions, merits, and circumstances (Almås et al., 2010; Huppert et al., 2019; Sigelman & Waitzman, 1991).
-Engelmann and Tomasello (2019) claim that children’s sense of fairness emerges at three years old, and we can observe it in collaborative activities, where they accept inequality if the procedure gives everyone an equal chance.
-Therefore, children at this age respond to unequal distributions based on interpersonal concerns, as they already demand equal respect.
-In any case, between 3 and 8 years of age, inequitable and anti-meritorious allocations are evaluated more negatively, but equitable and meritorious allocations are not evaluated more positively (Elenbaas, 2019).
-
-Some research shows that the social environment in which children develop, such as family and school, is associated with their prosocial behaviors by playing an essential role in the transmission of equity norms (Schunk & Zipperle, 2023; Kosse et al., 2020).
-In fact, schools contribute to institutionalizing and reproducing inequality by promoting values, norms, practices, and languages familiar to higher-class families because the dominant group’s culture shapes educational institutions (Bourdieu & Passeron, 1990).
-Middle- and upper-class students are better equipped to face academic challenges and are more familiar with academic expectations (Mikus et al., 2019).
-Such familiarity represents cultural capital in educational contexts because higher-status students come to school ready to meet these expectations and reap the benefits (Jack, 2016; Khan, 2011).
-Conversely, lower-status children lacking cultural capital must catch up while experiencing inequitable comparisons (Goudeau & Croizet, 2017).
-Additionally, academic achievement is treated as the outcome of dispositional factors (e.g., pupils’ efforts and talents or lack of them) rather than the result of differential access to critical resources.
-Due to the meritocratic frame schools encourage, both low- and high-status individuals believe that students’ success or failure is not due to their family background but rather to differences in efforts and talents (Darnon et al., 2018).
-In this sense, we believe that the perception of meritocracy can influence students' judgments about market justice preferences.
-Furthermore, we believe there is a difference between students' perceptions of meritocracy based on their own experience in school and what they perceive in society at large.
-Consistently, our first two hypotheses are:
-
-H1a: Students who perceive that there is more meritocracy at school will show larger market justice preferences
-
-H1b: Students who perceive that there is more meritocracy in society will show larger market justice preferences
-
-Family background and family socialization practices also contribute to children’s and adolescent’s market justice preferences.
-For example, Almås and colleagues (2017) found that adolescents from low-socioeconomic-status families are likelier to have an egalitarian fairness view and consider an equal distribution as fair in a situation with unequal merits.
-The authors speculate that differences in socialization practices across status groups might bring about, to a great extent, the fairness views of children and adolescents because social status seems to interact with these evaluations (e.g., Hvidberg et al., 2023).
-
-The classic work of Kohn showed that middle-class parents value the expression of internal states and emotions, such as self-control, curiosity, happiness, and consideration, while working-class parents promote deference, obedience, and conformity to authority (Kohn, 1963; Kohn & Schooler, 1969).
-Although parents from all social backgrounds encourage individualism in their children, this shared norm translates into different forms in high and low social classes (1999).
-Acemoglu (2021) claimed that the values families impart to their children interact with social mobility.
-Because obedience is a valuable characteristic for employers, in low-wage and social mobility environments, low-income families impart values of obedience to their children to prevent disadvantaging them in labor markets.
-On the one hand, children from privileged families are socialized to adopt a clear conception of individualism that highlights their internal states, independence, and idiosyncrasies.
-In contrast, children from disadvantaged families are socialized to support a more balanced view of individualism that considers personal characteristics as resources to overcome collective impediments on the path to upward mobility (Iacoviello & Lorenzi-Cioldi, 2019).
-In this way, we believe that there are differences in the socialization of values according to socioeconomic differences that could influence market justice preferences and, therefore, we propose the following hypothesis:
-
-H2: Students from families of higher social status will show larger market justice preferences.
-
-Recent empirical research has demonstrated that the institutional design of schools, coupled with the meritocratic ideology it fosters, significantly influences children’s and adolescents’ views on inequality and deservingness.
-For example, Jonsson and Beach’s (2015) study revealed that higher-status adolescents in Sweden tend to perpetuate social class stereotypes while describing the vocational and academic tracks.
-Academic track students are depicted as wealthy, intelligent, ambitious, and diligent, while vocational track students are characterized as poor, unambitious, unintelligent, and lackadaisical.
-These stereotypes help individuals maintain a sense of superiority over others and legitimize the prevailing social hierarchies and economic disparities (Jost & Burgess, 2000).
-
-H3: Students from schools of higher social status will show larger market justice preferences.
-
-H4: Students from schools with higher average levels of academic achievement will show larger market justice preferences.
-
-\[interacciones\]
-
-H5: The perception of meritocracy in school and society will moderate the effect of family social status on market justice preferences.
-
-H6: The perception of meritocracy in school and society will moderate the effect of school status on market justice preferences.
-
-H7: The perception of meritocracy in school and society will moderate the effect of school academic achievement on market justice preferences.
-
-### Summary of hypotheses
-
-```{r}
-#| fig-cap: "Summary of hypotheses"
-```
-
-## Methods
-
-### Data
-
-The main data source for the analysis is the First Study of Citizenship Education in Chile, carried out by the Education Quality Agency of the Ministry of Education.
-The application date was November 9, 2017.
-The target population of this study is Eighth-grade students from 242 schools.
-In the data, there are 8,589 students and 6,770 parents.
-This database was analyzed with the R package “ResponsePatterns” to detect possible repetitive and “careless” response patterns and thus contribute to a higher quality of research data (Gottfried, et al 2022).
-Responses from 171 students and 79 parents were removed, which, when merged, gave a total of 6,270 valid cases.
-
-The analysis of school variables includes data from the Ministry of Education's SIMCE 2017 database.
-This database contains information at the school level, such as the administrative dependency, its socioeconomic classification, and the achievement scores obtained in the mathematics and language census tests.
-It is available for free use on the MINEDUC \[web page\].
-
-After eliminating missing cases, the final sample used in the analysis was based on 5,047 students and parents of 231 schools for the dependent variable of access to social services.
-
-### Variables
-
-**Dependent variables**
-
-This study has three dependent variables related to the justification of social inequality in specific policy domains.
-The first asks whether access to social services should be conditional on income, i,.e., “It is just that in Chile people who can pay have a better education for their children”.
-Students rated their preferences using the following responses: “strongly disagree”, “Disagree”, “Agree”, and ”strongly agree”.
-An average index is built with these items (Cronbach's Alpha = 0,86).
-(Apéndice: items en español y su correspondiente traducción al inglés)
-
-Table 1 shows the items used, their response categories, and their frequencies.
-
-```{r}
-#| tbl-cap: "Dependent variables"
-```
-
-**Independent variables**
-
-For the primary independent variable, the perception of meritocracy, five items address the perception of rewards according to effort and intelligence at the school and societal levels.
-At the school level, students answer whether “Intelligence is important to get good grades” and “Effort is important to get good grades”.
-At the societal level, students respond to the following questions: “In Chile, people are rewarded for their efforts”, “In Chile, people get what they deserve”, and “In Chile, people are rewarded for their intelligence and skills”.
-Each item was answered on a four-point scale ranging from “Strongly disagree” to “Strongly agree”.
-
-The rest of the independent variables are divided into individual and school levels.
-At the individual level, family socioeconomic status was measured by the parents' highest educational level and the number of books at home.
-Likewise, an index of access to technology includes the number of computers, tablets, and cell phones at home, as well as whether there is an Internet connection.
-Table 2 shows the items used, their response categories, and their frequency.
-
-```{r}
-#| tbl-cap: "Independent variables"
-```
-
-The school-level variables are the administrative dependency of the school, the socioeconomic classification made by the Ministry of Education, the level of performance in the SIMCE test of the school, and the proportion of parents with university or postgraduate degrees.
-Table 3 shows the items used, response categories, and frequency.
-
-```{r}
-#| tbl-cap: "School context variables"
-```
-
-### Methods
-
-The data has a hierarchical structure of students nested in schools, so the model estimation is performed in a multilevel (random effects) framework.
-This modeling approach lets us correctly estimate individual and contextual effects in a single model.
-We estimate cumulative link mixed models for the ordinal dependent variables, whereas we use linear mixed effects models for the average index of inequality justification.
-
-The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open ScorrectlyOSF), the access to the document is available at this \[link\](https://doi.org/10.17605/OSF.IO/UFSDV).
-The statistical analysis of this research was performed using the free software R version 4.1.3.
-
-## Analysis
-
-### Descriptive analysis
-
-```{r}
-#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
-```
-
-Figure 3 shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram en figure X).
-On the left, we observe the social meritocracy diagrams, while on the right the school meritocracy diagrams are shown.
-For the three variables of perception of meritocracy in society the relationship is clear, since the average of market justice preferences increases the more there is agreement that people are rewarded for their effort, merit, and talent.
-This relationship needs to be clarified in the case of the variables of perception of meritocracy at school.
-To the extent that there is more agreement that the perception of talent is essential for obtaining good grades, the average of market justice preferences increases, but this relationship is not as clear as with the variables of meritocracy in society.
-In addition, the graph does not show a clear trend in the relationship between the perception that effort is essential to obtain good grades and market justice preferences.

### 2024-04-23 02-antecedentes.qmd
restructurar repo

--- 2024-04-23 paper.qmd
@@
 ---
-title: "market-justice-preferences"
+title: "02-antecedentes"
 author: "Equipo EDUMER"
@@
 editor: visual
-bibliography: "input/bib/just-ineq-merit.bib"
-link-citations: true
 ---
-
-# The socialization of meritocracy and market justice preferences at school
-
-## Introduction
-
-Since its origins, educational institutions have been related to the idea of social mobility and access to better opportunities.
-Therefore, the consistent evidence of the high level of social reproduction at the school level represents a threat to the promise of education and a meritocratic system [@bourdieu_reproduction_1990].
-A large part of the research in this field at an international level has addressed the extent to which the social origin of students affects their academic results and their life opportunities [@vonhippel_test_2019], confirming that schools have severe difficulties in closing the gaps of origin.
-Besides this socioeconomic perspective on school opportunities, recent research has addressed to what extent inequalities in the school context are also influencing students' perceptions, beliefs, and attitudes: Are social inequalities perceived at the school context?
-Are they rejected by the students, particularly those who are worst-off in socioeconomic terms?
-Or, Is there evidence at the school level that social inequalities are tolerated and even justified?
-[@batruch_belief_2022; @wiederkehr_belief_2015].
-
-In the present paper, we deal with the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system.
-In particular, we focus on market justice preferences (Lane, 1986), which refer to the preferences for distributing public goods (as health and education) based on criteria such as competition and payment capacity.
-Although from a rational point of view it could be expected an opposition to market justice by the underprivileged majority,  we argue that in a social environment characterized by a promotion of meritocratic ideals - as the school system - would lead to the opposite: a larger market justice preferences.
-
-Given that the school environment has an important focus on performance, achievement and acknowledgment, meritocracy has been one of the principal concepts used for understanding and even for justifying performance differences among students.
-Meritocracy is a distributive system based on the belief that people should be rewarded and promoted based on their abilities, knowledge, and achievements [@young_rise_1958].
-It is often seen as a way to create equal opportunities and fairness, as individuals can rise to positions of power and influence based on their own merit rather than their background or connections.
-However, some argue that meritocracy can actually lead to tolerating or even justifying social inequalities, as it can create a hierarchy where those who already have resources and advantages are more likely to succeed.
-In this regard, a great deal of academic research about meritocracy delves into the assessment of to what extent rewards and privileges in society are related to merit, emphasizing the so-called unfulfillable promise of meritocracy [@mijs_stratified_2016].
-Complementing this agenda, a second and emerging research area deals with subjective aspects of meritocracy, such as perceptions and beliefs.
-
-\[definición de market justice, welfare marketization, (Lindh)\]
-
-\[contexto Chileno market justice, dar cifras/porcentajes de privatización de salud, educación y pensiones; asociar al tema de la calidad (lo público es peor)\]
-
-The perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society [@duru-bellat_who_2012; @castillo_meritocracia_2019].
-This perception can vary greatly depending on individual experiences, social, economic, and cultural background.
-Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work.
-In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2021a].
-Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger justification of social inequalities, as individual achievement would be seen as rewarded and social policies as less necessary [@batruch_belief_2022].
-
-Most of the research that has related meritocratic beliefs to inequality justification so far has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies.
-Regarding schools, the way in which they deal with unequal conditions of origin has been linked to the *hidden curriculum* [@chafel_schooling_1997], whereby students learn about distributive norms in society and mechanisms of justification of social differences.
-Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population [@batruch_belief_2022; @wiederkehr_belief_2015], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger justification of social inequalities, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary [@batruch_belief_2022].
-
-\~
-
-\[introducir distinción sobre percepción de meritocracia a nivel escolar y a nivel de la sociedad\]
-
-\[resaltar 3 focos de innovación: market justice preferences at school level, relación con meritocracia, y esquema de meritocracia en la escuela y en la sociedad\]
-
-The present paper deals with the association between the perception of meritocracy and the justification of social inequalities, with two main focuses.
-Firstly, it assesses the justification of social inequalities in social policy domains, such as health, pensions, and education.
-We argue that individuals who perceive more meritocracy would be more willing to justify better services in these domains for those with higher incomes.  Secondly, we focus on the student-age population as we point out that it is possible to track down the origin of meritocratic beliefs (and their consequences) to early socialization processes.
-To this regard, we take into account the family and the school as two main socialization agencies that play a significant role in the socialization of cultural beliefs by transmitting cultural norms, values, and expectations to young people.
 
@@
 
-## Methods
-
-### Data
-
-The main data source for the analysis is the First Study of Citizenship Education in Chile, carried out by the Education Quality Agency of the Ministry of Education.
-The application date was November 9, 2017.
-The target population of this study is Eighth-grade students from 242 schools.
-In the data, there are 8,589 students and 6,770 parents.
-This database was analyzed with the R package “ResponsePatterns” to detect possible repetitive and “careless” response patterns and thus contribute to a higher quality of research data (Gottfried, et al 2022).
-Responses from 171 students and 79 parents were removed, which, when merged, gave a total of 6,270 valid cases.
-
-The analysis of school variables includes data from the Ministry of Education's SIMCE 2017 database.
-This database contains information at the school level, such as the administrative dependency, its socioeconomic classification, and the achievement scores obtained in the mathematics and language census tests.
-It is available for free use on the MINEDUC \[web page\].
-
-After eliminating missing cases, the final sample used in the analysis was based on 5,047 students and parents of 231 schools for the dependent variable of access to social services.
-
-### Variables
-
-**Dependent variables**
-
-This study has three dependent variables related to the justification of social inequality in specific policy domains.
-The first asks whether access to social services should be conditional on income, i,.e., “It is just that in Chile people who can pay have a better education for their children”.
-Students rated their preferences using the following responses: “strongly disagree”, “Disagree”, “Agree”, and ”strongly agree”.
-An average index is built with these items (Cronbach's Alpha = 0,86).
-(Apéndice: items en español y su correspondiente traducción al inglés)
-
-Table 1 shows the items used, their response categories, and their frequencies.
-
-```{r}
-#| tbl-cap: "Dependent variables"
-```
-
-**Independent variables**
-
-For the primary independent variable, the perception of meritocracy, five items address the perception of rewards according to effort and intelligence at the school and societal levels.
-At the school level, students answer whether “Intelligence is important to get good grades” and “Effort is important to get good grades”.
-At the societal level, students respond to the following questions: “In Chile, people are rewarded for their efforts”, “In Chile, people get what they deserve”, and “In Chile, people are rewarded for their intelligence and skills”.
-Each item was answered on a four-point scale ranging from “Strongly disagree” to “Strongly agree”.
-
-The rest of the independent variables are divided into individual and school levels.
-At the individual level, family socioeconomic status was measured by the parents' highest educational level and the number of books at home.
-Likewise, an index of access to technology includes the number of computers, tablets, and cell phones at home, as well as whether there is an Internet connection.
-Table 2 shows the items used, their response categories, and their frequency.
-
-```{r}
-#| tbl-cap: "Independent variables"
-```
-
-The school-level variables are the administrative dependency of the school, the socioeconomic classification made by the Ministry of Education, the level of performance in the SIMCE test of the school, and the proportion of parents with university or postgraduate degrees.
-Table 3 shows the items used, response categories, and frequency.
-
-```{r}
-#| tbl-cap: "School context variables"
-```
-
-### Methods
-
-The data has a hierarchical structure of students nested in schools, so the model estimation is performed in a multilevel (random effects) framework.
-This modeling approach lets us correctly estimate individual and contextual effects in a single model.
-We estimate cumulative link mixed models for the ordinal dependent variables, whereas we use linear mixed effects models for the average index of inequality justification.
-
-The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open ScorrectlyOSF), the access to the document is available at this \[link\](https://doi.org/10.17605/OSF.IO/UFSDV).
-The statistical analysis of this research was performed using the free software R version 4.1.3.
-
-## Analysis
-
-### Descriptive analysis
-
-```{r}
-#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
-```
-
-Figure 3 shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram en figure X).
-On the left, we observe the social meritocracy diagrams, while on the right the school meritocracy diagrams are shown.
-For the three variables of perception of meritocracy in society the relationship is clear, since the average of market justice preferences increases the more there is agreement that people are rewarded for their effort, merit, and talent.
-This relationship needs to be clarified in the case of the variables of perception of meritocracy at school.
-To the extent that there is more agreement that the perception of talent is essential for obtaining good grades, the average of market justice preferences increases, but this relationship is not as clear as with the variables of meritocracy in society.
-In addition, the graph does not show a clear trend in the relationship between the perception that effort is essential to obtain good grades and market justice preferences.
+##

### 2024-04-23 03-methods.qmd
restructurar repo

--- 2024-04-22 paper.qmd
@@
 ---
-title: "market-justice-preferences"
-author: "Equipo EDUMER"
-format: html
+title: "03-methods"
+author: "Equipo Edumer"
+format: pdf
 editor: visual
-bibliography: "input/bib/just-ineq-merit.bib"
 ---
-
-# The socialization of meritocracy and market justice preferences at school
-
-## Introduction
-
-Since its origins, educational institutions have been related to the idea of social mobility and access to better opportunities.
-Therefore, the consistent evidence of the high level of social reproduction at the school level represents a threat to the promise of education and a meritocratic system [@bourdieu_reproduction_1990].
-A large part of the research in this field at an international level has addressed the extent to which the social origin of students affects their academic results and their life opportunities [@vonhippel_test_2019], confirming that schools have severe difficulties in closing the gaps of origin.
-Besides this socioeconomic perspective on school opportunities, recent research has addressed to what extent inequalities in the school context are also influencing students' perceptions, beliefs, and attitudes: Are social inequalities perceived at the school context?
-Are they rejected by the students, particularly those who are worst-off in socioeconomic terms?
-Or, Is there evidence at the school level that social inequalities are tolerated and even justified?
-[@batruch_belief_2022; @wiederkehr_belief_2015].
-
-In the present paper, we deal with the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system.
-In particular, we focus on market justice preferences (Lane, 1986), which refer to the preferences for distributing public goods (as health and education) based on criteria such as competition and payment capacity.
-Although from a rational point of view it could be expected an opposition to market justice by the underprivileged majority,  we argue that in a social environment characterized by a promotion of meritocratic ideals - as the school system - would lead to the opposite: a larger market justice preferences.
-
-Given that the school environment has an important focus on performance, achievement and acknowledgment, meritocracy has been one of the principal concepts used for understanding and even for justifying performance differences among students.
-Meritocracy is a distributive system based on the belief that people should be rewarded and promoted based on their abilities, knowledge, and achievements (Young, 1958).
-It is often seen as a way to create equal opportunities and fairness, as individuals can rise to positions of power and influence based on their own merit rather than their background or connections.
-However, some argue that meritocracy can actually lead to tolerating or even justifying social inequalities, as it can create a hierarchy where those who already have resources and advantages are more likely to succeed.
-In this regard, a great deal of academic research about meritocracy delves into the assessment of to what extent rewards and privileges in society are related to merit, emphasizing the so-called unfulfillable promise of meritocracy (Mijs, 2016).
-Complementing this agenda, a second and emerging research area deals with subjective aspects of meritocracy, such as perceptions and beliefs.
-
-\[definición de market justice, welfare marketization, (Lindh)\]
-
-\[contexto Chileno market justice, dar cifras/porcentajes de privatización de salud, educación y pensiones; asociar al tema de la calidad (lo público es peor)\]
-
-The perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society (Duru-Bellat & Tenret, 2012; Castillo et al., 2019).
-This perception can vary greatly depending on individual experiences, social, economic, and cultural background.
-Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work.
-In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality (Lampert, Mijs, etc).
-Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger justification of social inequalities, as individual achievement would be seen as rewarded and social policies as less necessary (Batruch et al, 2021).
-
-Most of the research that has related meritocratic beliefs to inequality justification so far has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies.
-Regarding schools, the way in which they deal with unequal conditions of origin has been linked to the \_hidden curriculum\_ (Chafel, 1997, https://doi.org/10.1002/J.2379-3988.1997.TB00004.X), whereby students learn about distributive norms in society and mechanisms of justification of social differences.
-Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population \[\@batruch_belief_2022;\@wiederkehr_belief_2015\], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger justification of social inequalities, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary (Batruch et al, 2021).
-
-\~
-
-\[introducir distinción sobre percepción de meritocracia a nivel escolar y a nivel de la sociedad\]
-
-\[resaltar 3 focos de innovación: market justice preferences at school level, relación con meritocracia, y esquema de meritocracia en la escuela y en la sociedad\]
-
-The present paper deals with the association between the perception of meritocracy and the justification of social inequalities, with two main focuses.
-Firstly, it assesses the justification of social inequalities in social policy domains, such as health, pensions, and education.
-We argue that individuals who perceive more meritocracy would be more willing to justify better services in these domains for those with higher incomes.  Secondly, we focus on the student-age population as we point out that it is possible to track down the origin of meritocratic beliefs (and their consequences) to early socialization processes.
-To this regard, we take into account the family and the school as two main socialization agencies that play a significant role in the socialization of cultural beliefs by transmitting cultural norms, values, and expectations to young people.
-
-## Justification of inequality
-
-Research on social stratification beliefs, which explore individual perceptions of who deserves what and why (Kluegel & Smith, 1987), highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness.
-The influence of ideologies (Wegener & Liebig, 1995) and cultural schemas (Homan et al., 2017) is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations.
-While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature (Castillo, 2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),  there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles.
-This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades (Centeno & Cohen, 2012; Harvey, 2007), affecting areas such as pensions, health services, and education.
-The justification of social inequality based on marketype criteria has been conceptualized as the individuals’ adherence to one specific justice evaluation, the market justice, that is, affording legitimacy to the allocation of goods and services based on prices and individuals’ ability to pay (Boltanski & Chiapello- 2007; Lane 1986; Streeck 2012).
-
-Robert E.
-Lane proposed the underpinnings of market justice, which he differentiated from political justice.
-For him, “it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets” (1986, p.
-384).
-Following the theory of relative deprivation –a social phenomenon arising when individuals cannot afford what most others in their environment can (Merton 1950)– Lane notes that, in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions.
-Although empirical research has shown that, contrary to Lane’s observation, relative deprivation, even in market settings, produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action such as protests and revolt (Greitemeyer and Sagioglou 2016; Mishra and Carleton 2015; Smith et al.
-2012; Power 2018; Power, Madsen, and Morton 2020), the conceptualization of market justice has been since then closely coupled to the merit principle for allocating outcomes: the claim that the unequal levels of well-being individuals enjoy ought to be, to some extent, a function of their talents and efforts, regardless of their needs or membership, the two latter being the realm of political justice and its closely coupled principles of need –allocating outcomes to those who require them most– and equality –allocating the same outcome to everyone–, which have been at the center of welfarism (see Wilson 2003).
-The underlying notion of market justice also resonates in its application to welfare regimes, providing a framework for understanding the varied global approaches to managing social services.
-
-The management of social services manifests in varying approaches across nations, with substantial differences in funding and delivery methods (Jensen, 2008; Stoy, 2014).
-Nordic countries, for example, predominantly employ public agencies to produce and provide social services, funding these through collective taxation and offering them in kind to the majority of citizens.
-This system prioritizes political justice, placing it above market mechanisms in accessing services.
-In contrast, other countries rely more heavily on for-profit entities and private funding, where service distribution depends mainly on individual financial capacity of paying user fees, highlighting the influence of market justice in service allocation.
-The trend toward marketization of welfare services has been growing since the 1980s (Salamon, 1993), and this shift is increasingly evident even in countries where market solutions have traditionally had a minor role in social policy (Sivesind, 2017).
-
-The question arises whether adults and children justify unequal access to welfare services based on market justice principles.
-Influenced by theories of policy feedback, which suggest that social welfare policies can reinforce (positive feedback) or undermine (negative feedback) previous policy trajectories (Fernandez and Jaime-Castillo 2013; Pierson 2000; Weaver 2010), citizens’ beliefs about market justice are likely also shaped by the institutional and social contexts they encounter.
-Indeed, the justification of inequality in access to essential services like education and health, based on one’s ability to pay, shows significant variation across countries, as demonstrated by international surveys, although they do not usually include children in their samples.
-For instance, the 2019 International Social Survey Program (ISSP) provides insights into how adults perceive inequality in accessing welfare services.
-Figure 1 illustrates the variation in agreement levels regarding whether it is just for individuals with higher incomes to purchase better education or healthcare.
-Notably, in 18 of the 29 countries analyzed, there is a greater justification for market inequality in healthcare access than in education.
-Nevertheless, the general sentiment typically ranges from 'somewhat unjust' to 'neither just nor unjust, with mixed feelings.
-
-```{r}
-#| fig-cap: "Average of market justice preferences by country"
-```
-
-While contemporary policy developments have increasingly embraced the marketization of social welfare in areas such as education, pensions, and health services, questions about public acceptance of these changes persist.
-Lindh’s (2015) analysis of ISSP 2009 data from 17 OECD countries reveals a general lack of support for market-based distribution of social services, suggesting widespread disapproval of market stratification of essential services.
-This finding is corroborated by Soler-Martínez and colleagues’ (2023) research from Latinobarómetro 2020 across 18 Latin American countries, where concerns about health and education access predominated over income inequality.
-These results indicate that reforms toward welfare marketization are typically driven by elite political decisions rather than grassroots demand.
-
-Despite high-income inequality and limited social mobility in Latin America, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region (Bucca, 2016; Chong et al., 2008; Torche, 2014; Salgado and Castillo, 2023).
-The reliance on private welfare providers and widespread user fees (Molyneux, 2008) adds complexity to this context.
-Yet, research on children’s justification of market-based inequalities in accessing welfare services remains limited, especially in Latin America, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.
-
-## Meritocracy
-
-The concept of meritocracy frequently appears nowadays when analyzing cultural determinants of social inequalities.
-In general, it is mentioned as a value associated with justice, as it would link efforts and talents with rewards in an equitable manner.
-This normative sense is quite far from its original formulation by Michael Young (1958) in the satirical novel "The Rise of Meritocracy", where it ironically represented a mechanism for reproducing the inequalities of origin.
-The meritocratic ideal had remained relatively unchallenged until a series of recent publications turned into its potential consequences for maintaining social inequality.
-Perhaps one of the most recent sources in this line is Michael Sandel's "The Tyranny of Merit", where he strongly questions the implications of carrying out the principle of merit in societies that do not guarantee equal opportunities and that generate a feeling of scarce recognition and appreciation of those who receive lesser rewards: "In society's eyes, and perhaps also their own, their work no longer signified a valued contribution to the common good." (Sandel, pp).
-
-Empirical research on meritocracy has increased along with the philosophical-normative discussion on meritocracy in recent years.
-Particularly from a sociological perspective, meritocracy has been used in research on social mobility to characterize societies with low mobility that threaten the meritocratic ideal (Goldthorpe).
-More recently, sociology and social psychology research has attended to the subjective aspects vis-a-vis beliefs in meritocracy.
-The label of beliefs in this realm covers a series of areas, such as attitudes, perceptions, and preferences (Castillo et al), whereby most of the link this subjective dimensions to individual socio-structural factors and  context-level determinants.
-For instance, some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs (Mijs 2021), and how larger inequality affects meritocratic beliefs (Morris et al 2022).
-Based on these findings, a research agenda has been reinforced on the legitimizing role of meritocracy, in line with previous studies using the concept of a just world (Lerner, Dalbert) and the theory of system justification (Jost & Major).
-
-How do meritocratic beliefs legitimize inequalities?
-Empirical studies have used experiments and surveys to address this question.
-For instance, the evidence suggests that just world beliefs correlate negatively with support for redistributive compensation systems (Frank et al., 2015).
-Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed (Evans & Kelley, 2018).
-Almås and colleagues (2020) found that in a relatively unequal society (the United States), the highly educated accept inequality significantly more than the less educated because they perceive inequality as justifiable owing to differences in productivity (i.e., merit), whereas in a relatively equal society (Norway), the less educated accept inequality more, but not significantly more than the highly educated because meritocratic values are less prevalent.
-Barr and Miller (2020) also addressed this triple interaction between the level of inequality in a society, the individual level of education, and the perceived origin of the disparity (either by luck or effort) to determine the extent to which inequality is accepted.
-They found that the interaction's mechanism varies depending on the compared societies.
-Finally, García-Sánchez and colleagues (2020), using data for 41 countries from the International Social Survey Programme (ISSP), found that the perceived size of the income gap correlated positively with support for progressive taxation.
-Still, this association was weaker among those who endorsed meritocratic and equal opportunity beliefs.
-In the same line, experimental research by Durante & Putterman (2009) subjects support less redistribution when the initial distribution is determined according to task performance.
-
-Research about meritocratic beliefs at school age is rather scarce, leaving a wide research gap as schools are one of the primary socialization institutions where achievement based on merit explains success (Erivwo et al 2021).
-Nevertheless, it is possible to find some initial works focused on the area of distributive justice at school that are closely related to meritocratic beliefs, as the ones by Resh and Sabbagh (quotes).
-Using justice in grade obtention as a measure of distributive justice and meritocracy (Sabbagh et al 2006), they find for instance that a larger sense of distributive justice about grades is associated to higher socio-economic status (Resh 2010), have a positive effect on liberal democratic orientation and on trust in people and in formal institutions (Resh & Sabbagh 2013), and tend to refrain from violence and to engage to a greater extent in extra-curricular school activity and community volunteering (Resh & Sabbagh 2017).
-
-## Children's judgments of inequality, schools, and family background
-
-Research indicates various socialization practices at families and schools during childhood and adolescence that impact dispositional and behavioral tendencies concerning justifications of inequality in adult life.
-The differences in economic understanding across age groups are consistent with research on cognitive development (Choudhury et al., 2006).
-Adolescents with mature socio-cognitive abilities tend to express stronger preferences for fairness than infants and children (Wynn et al., 2017).
-As children grow older, they become more likely to behave fairly, with their early-emerging strict egalitarianism being replaced by an increasing endorsement of fairness principles and engagement in collaborative activities (Huppert et al., 2019; McAuliffe et al., 2017).
-In these activities, their fairness views consider individual contributions, merits, and circumstances (Almås et al., 2010; Huppert et al., 2019; Sigelman & Waitzman, 1991).
-Engelmann and Tomasello (2019) claim that children’s sense of fairness emerges at three years old, and we can observe it in collaborative activities, where they accept inequality if the procedure gives everyone an equal chance.
-Therefore, children at this age respond to unequal distributions based on interpersonal concerns, as they already demand equal respect.
-In any case, between 3 and 8 years of age, inequitable and anti-meritorious allocations are evaluated more negatively, but equitable and meritorious allocations are not evaluated more positively (Elenbaas, 2019).
-
-Some research shows that the social environment in which children develop, such as family and school, is associated with their prosocial behaviors by playing an essential role in the transmission of equity norms (Schunk & Zipperle, 2023; Kosse et al., 2020).
-In fact, schools contribute to institutionalizing and reproducing inequality by promoting values, norms, practices, and languages familiar to higher-class families because the dominant group’s culture shapes educational institutions (Bourdieu & Passeron, 1990).
-Middle- and upper-class students are better equipped to face academic challenges and are more familiar with academic expectations (Mikus et al., 2019).
-Such familiarity represents cultural capital in educational contexts because higher-status students come to school ready to meet these expectations and reap the benefits (Jack, 2016; Khan, 2011).
-Conversely, lower-status children lacking cultural capital must catch up while experiencing inequitable comparisons (Goudeau & Croizet, 2017).
-Additionally, academic achievement is treated as the outcome of dispositional factors (e.g., pupils’ efforts and talents or lack of them) rather than the result of differential access to critical resources.
-Due to the meritocratic frame schools encourage, both low- and high-status individuals believe that students’ success or failure is not due to their family background but rather to differences in efforts and talents (Darnon et al., 2018).
-In this sense, we believe that the perception of meritocracy can influence students' judgments about market justice preferences.
-Furthermore, we believe there is a difference between students' perceptions of meritocracy based on their own experience in school and what they perceive in society at large.
-Consistently, our first two hypotheses are:
-
-H1a: Students who perceive that there is more meritocracy at school will show larger market justice preferences
-
-H1b: Students who perceive that there is more meritocracy in society will show larger market justice preferences
-
-Family background and family socialization practices also contribute to children’s and adolescent’s market justice preferences.
-For example, Almås and colleagues (2017) found that adolescents from low-socioeconomic-status families are likelier to have an egalitarian fairness view and consider an equal distribution as fair in a situation with unequal merits.
-The authors speculate that differences in socialization practices across status groups might bring about, to a great extent, the fairness views of children and adolescents because social status seems to interact with these evaluations (e.g., Hvidberg et al., 2023).
-
-The classic work of Kohn showed that middle-class parents value the expression of internal states and emotions, such as self-control, curiosity, happiness, and consideration, while working-class parents promote deference, obedience, and conformity to authority (Kohn, 1963; Kohn & Schooler, 1969).
-Although parents from all social backgrounds encourage individualism in their children, this shared norm translates into different forms in high and low social classes (1999).
-Acemoglu (2021) claimed that the values families impart to their children interact with social mobility.
-Because obedience is a valuable characteristic for employers, in low-wage and social mobility environments, low-income families impart values of obedience to their children to prevent disadvantaging them in labor markets.
-On the one hand, children from privileged families are socialized to adopt a clear conception of individualism that highlights their internal states, independence, and idiosyncrasies.
-In contrast, children from disadvantaged families are socialized to support a more balanced view of individualism that considers personal characteristics as resources to overcome collective impediments on the path to upward mobility (Iacoviello & Lorenzi-Cioldi, 2019).
-In this way, we believe that there are differences in the socialization of values according to socioeconomic differences that could influence market justice preferences and, therefore, we propose the following hypothesis:
-
-H2: Students from families of higher social status will show larger market justice preferences.
-
-Recent empirical research has demonstrated that the institutional design of schools, coupled with the meritocratic ideology it fosters, significantly influences children’s and adolescents’ views on inequality and deservingness.
-For example, Jonsson and Beach’s (2015) study revealed that higher-status adolescents in Sweden tend to perpetuate social class stereotypes while describing the vocational and academic tracks.
-Academic track students are depicted as wealthy, intelligent, ambitious, and diligent, while vocational track students are characterized as poor, unambitious, unintelligent, and lackadaisical.
-These stereotypes help individuals maintain a sense of superiority over others and legitimize the prevailing social hierarchies and economic disparities (Jost & Burgess, 2000).
-
-H3: Students from schools of higher social status will show larger market justice preferences.
-
-H4: Students from schools with higher average levels of academic achievement will show larger market justice preferences.
-
-\[interacciones\]
-
-H5: The perception of meritocracy in school and society will moderate the effect of family social status on market justice preferences.
-
-H6: The perception of meritocracy in school and society will moderate the effect of school status on market justice preferences.
-
-H7: The perception of meritocracy in school and society will moderate the effect of school academic achievement on market justice preferences.
-
-### Summary of hypotheses
-
-```{r}
-#| fig-cap: "Summary of hypotheses"
-```
 
@@
 The statistical analysis of this research was performed using the free software R version 4.1.3.
-
-## Analysis
-
-### Descriptive analysis
-
-```{r}
-#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
-```
-
-Figure 3 shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram en figure X).
-On the left, we observe the social meritocracy diagrams, while on the right the school meritocracy diagrams are shown.
-For the three variables of perception of meritocracy in society the relationship is clear, since the average of market justice preferences increases the more there is agreement that people are rewarded for their effort, merit, and talent.
-This relationship needs to be clarified in the case of the variables of perception of meritocracy at school.
-To the extent that there is more agreement that the perception of talent is essential for obtaining good grades, the average of market justice preferences increases, but this relationship is not as clear as with the variables of meritocracy in society.
-In addition, the graph does not show a clear trend in the relationship between the perception that effort is essential to obtain good grades and market justice preferences.

### 2024-04-23 04-analisis.qmd
restructurar repo

---
title: "04-analisis"
author: "Equipo Edumer"
format: pdf
editor: visual
---

## Analysis

### Descriptive analysis

```{r}
#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
```

Figure 3 shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram en figure X). On the left, we observe the social meritocracy diagrams, while on the right the school meritocracy diagrams are shown. For the three variables of perception of meritocracy in society the relationship is clear, since the average of market justice preferences increases the more there is agreement that people are rewarded for their effort, merit, and talent. This relationship needs to be clarified in the case of the variables of perception of meritocracy at school. To the extent that there is more agreement that the perception of talent is essential for obtaining good grades, the average of market justice preferences increases, but this relationship is not as clear as with the variables of meritocracy in society. In addition, the graph does not show a clear trend in the relationship between the perception that effort is essential to obtain good grades and market justice preferences.

### Cumulative link mixed models for the Justification of inequality in education, health and pensions

Three Cumulative link mixed models were estimated for the ordinal dependent variables of justice in differential access to pensions, education, and health according to income. Figure 4 shows the estimation of this regression model containing all the variables used in the study for the three dependent variables separately. However, this figure shows only the effect of meritocracy variables on society and school as independent variables. Complete models are available in the appendix.

```{r}
#| fig-cap: "Odds-ratios of justification of education, pensions and health by social and school meritocracy"
```

Figure 4 shows that for the three dependent variables of justice in differential access to pensions, education, and health, the trend is similar. In the context of school meritocracy, the effects are mixed. As school talent increases, the justification for differentiated access to pensions, education, and health increases; on the contrary, as the school effort variable increases, the justification for differentiated access to pensions, education, and health decreases, keeping the rest of the independent variables constant. Regarding the variables of social meritocracy, the three variables of talent, effort, and deservingness show that as these increase, the justification for differentiated access to pensions, education, and health also increases.

### Multilevel regression models for market justice preferences

```{r}
#| tbl-cap: "Individual effects of multilevel regression models"
```

Table 4 shows the results of the multilevel estimation for justice market preferences. For this variable, the intraclass correlation obtained shows that the variation between schools corresponds to 4% of the variation of students’ preferences. This means that there is low variance between schools and therefore limits the possibilities of finding effects at the aggregate level.

Model 1 introduces social meritocratic variables: effort (whether efforts are rewarded), deservingness (people get what they deserve), and social talent (intelligence and skills are rewarded in society). In line with our hypotheses, the perception of a meritocratic society is positively related to the justification of inequality. Model 2 shows a mixed picture: while those perceiving that talent is rewarded also justify the inequality, the perception that effort is rewarded at school is negatively related to justification of inequality. Family background variables in Model 3 reveal that education and technology access are not related to justification of inequality, whereby we observe a negative impact of family cultural capital as measured by the number of books at home. While school socio-structural variables added in Model 4 show no significant effects, average achievement scores in the SIMCE test depict a negative relationship with the dependent variable, meaning that students that attend schools with better achievement scores on average justify less inequality.

In relation to model fit, when comparing the deviance with a model without predictors (null model), all the models have a statistically significant difference, with model 5 having the lowest deviance. According to Raudenbush and Bryk's (2002) estimate of R2, the level 1 variance of model 5 is 0.11 and the level 2 variance is 0.72. The total variance of model 5 according to Snijders and Bosker (2012) is 0.14.

```{r}
#| tbl-cap: "Contextual effects of multilevel regression models"
```

### Interactions effects

```{r}
#| tbl-cap: "Interactions effects"
```

The interaction terms in Table 6 suggest that students’ family background moderates the relationship between their meritocratic perceptions in Chile and their justification of inequality. The direction of these effects confirms our initial prediction. Thus, in line with our Hypothesis 3, the relationship between social effort and justification of inequality becomes less positive for those students whose parents achieved university or postgraduate education. That is, lower-status students (measured by parental education) justify more inequality when they adhere more strongly to deservingnessmeritocratic perceptions in Chile. We depict this moderating effect in Figure 1. This result confirms the enlightenment thesis (see above). The same moderating effect is observed for students’ family cultural capital: the relationship between effort (but not deservingness nor social talent) and justification of inequality becomes less positive for students with more than 25 books at home. Interestingly, family cultural capital also moderates the relationship between meritocratic beliefs at school (as measured by students’ belief that effort is important to get good grades) and justification of inequality.

At the school level, we also observe moderating effects of school status and students’ meritocratic beliefs over their justification of inequality. Thus, in line with our Hypothesis 5, high-status schools justify less inequality when, on average, their students have a greater perception of meritocracy in Chile, as measured by the three indicators we used (i.e., effort, deservingness, and talent). Finally, as we stated in Hypothesis 7, lowhigh-achieving schools justify more inequality when their students have, on average, stronger meritocratic beliefs in Chile. In Figure 25 we depict this moderating effect of school achievement for the relationship between deservingness and justification of inequality. We did not observe these moderating effects at the school level for students’ meritocratic beliefs in the school (i.e., the idea that effort or talent are important to get good grades).

```{r}
#| fig-cap: "Interaction between deservingness and justification of inequality by school SIMCE achievement"
```

### 2024-04-23 05-conclusiones.qmd
restructurar repo

---
title: "05-conclusiones"
author: "Equipo EDUMER"
format: pdf
editor: visual
---

##

-   Efecto ilustrador de la educación? (esfuerzo -\> redistribución)

### 2024-04-23 06-anexos.qmd
restructurar repo

---
title: "06-anexos"
author: "Equipo EDUMER"
format: pdf
editor: visual
---

```{r}
#| tbl-cap: "Translation of inequality justification items"
```

|                                                                                                                    |                                                                                                                                              |
|--------------------------------------------------------------------------------------------------------------------|----------------------------------------------------------------------------------------------------------------------------------------------|
| **English translation**                                                                                            | **Original Spanish**                                                                                                                         |
| It is just that in Chile people who cna pay have a better education for their children                             | Es justo que en Chile las personas que puedan pagar tengan una mejor educación para sus hijos.                                               |
| It is just that in Chile people with higher incomes can have better pensions than people with low incomes          | Es justo que en Chile las personas con mayores ingresos puedan tener mejores pensiones que las personas de ingresos más bajos                |
| It is just that in Chile people with higher incomes can access better health services than people with low incomes | Es justo que en Chile las personas con mayores ingresos puedan acceder a una mejor atención de salud que las personas con ingresos más bajos |

```{r}
#| fig-cap: "Access to better health services"
```

```{r}
#| fig-cap: "Access to better education"
```

```{r}
#| fig-cap: "Access to better pensions"
```

### 2024-04-23 paper.qmd
restructurar repo

@@
 
-## Introduction
-
-Since its origins, educational institutions have been related to the idea of social mobility and access to better opportunities.
-Therefore, the consistent evidence of the high level of social reproduction at the school level represents a threat to the promise of education and a meritocratic system [@bourdieu_reproduction_1990].
-A large part of the research in this field at an international level has addressed the extent to which the social origin of students affects their academic results and their life opportunities [@vonhippel_test_2019], confirming that schools have severe difficulties in closing the gaps of origin.
-Besides this socioeconomic perspective on school opportunities, recent research has addressed to what extent inequalities in the school context are also influencing students' perceptions, beliefs, and attitudes: Are social inequalities perceived at the school context?
-Are they rejected by the students, particularly those who are worst-off in socioeconomic terms?
-Or, Is there evidence at the school level that social inequalities are tolerated and even justified?
-[@batruch_belief_2022; @wiederkehr_belief_2015].
-
-In the present paper, we deal with the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system.
-In particular, we focus on market justice preferences (Lane, 1986), which refer to the preferences for distributing public goods (as health and education) based on criteria such as competition and payment capacity.
-Although from a rational point of view it could be expected an opposition to market justice by the underprivileged majority,  we argue that in a social environment characterized by a promotion of meritocratic ideals - as the school system - would lead to the opposite: a larger market justice preferences.
-
-Given that the school environment has an important focus on performance, achievement and acknowledgment, meritocracy has been one of the principal concepts used for understanding and even for justifying performance differences among students.
-Meritocracy is a distributive system based on the belief that people should be rewarded and promoted based on their abilities, knowledge, and achievements [@young_rise_1958].
-It is often seen as a way to create equal opportunities and fairness, as individuals can rise to positions of power and influence based on their own merit rather than their background or connections.
-However, some argue that meritocracy can actually lead to tolerating or even justifying social inequalities, as it can create a hierarchy where those who already have resources and advantages are more likely to succeed.
-In this regard, a great deal of academic research about meritocracy delves into the assessment of to what extent rewards and privileges in society are related to merit, emphasizing the so-called unfulfillable promise of meritocracy [@mijs_stratified_2016].
-Complementing this agenda, a second and emerging research area deals with subjective aspects of meritocracy, such as perceptions and beliefs.
-
-\[definición de market justice, welfare marketization, (Lindh)\]
-
-\[contexto Chileno market justice, dar cifras/porcentajes de privatización de salud, educación y pensiones; asociar al tema de la calidad (lo público es peor)\]
-
-The perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society [@duru-bellat_who_2012; @castillo_meritocracia_2019].
-This perception can vary greatly depending on individual experiences, social, economic, and cultural background.
-Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work.
-In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2021a].
-Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger justification of social inequalities, as individual achievement would be seen as rewarded and social policies as less necessary [@batruch_belief_2022].
-
-Most of the research that has related meritocratic beliefs to inequality justification so far has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies.
-Regarding schools, the way in which they deal with unequal conditions of origin has been linked to the *hidden curriculum* [@chafel_schooling_1997], whereby students learn about distributive norms in society and mechanisms of justification of social differences.
-Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population [@batruch_belief_2022; @wiederkehr_belief_2015], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger justification of social inequalities, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary [@batruch_belief_2022].
-
-\~
-
-\[introducir distinción sobre percepción de meritocracia a nivel escolar y a nivel de la sociedad\]
-
-\[resaltar 3 focos de innovación: market justice preferences at school level, relación con meritocracia, y esquema de meritocracia en la escuela y en la sociedad\]
-
-The present paper deals with the association between the perception of meritocracy and the justification of social inequalities, with two main focuses.
-Firstly, it assesses the justification of social inequalities in social policy domains, such as health, pensions, and education.
-We argue that individuals who perceive more meritocracy would be more willing to justify better services in these domains for those with higher incomes.  Secondly, we focus on the student-age population as we point out that it is possible to track down the origin of meritocratic beliefs (and their consequences) to early socialization processes.
-To this regard, we take into account the family and the school as two main socialization agencies that play a significant role in the socialization of cultural beliefs by transmitting cultural norms, values, and expectations to young people.
-
-## Justification of inequality
-
-Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegelBeliefsInequalityAmericans1987], highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness.
-The influence of ideologies [@wegenerDominantIdeologiesVariation1995] and cultural schemas [@homanBeingBecomingPoor2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations.
-While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castilloLegitimacyInequalityHighly2011; ] (Castillo, 2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),  there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles.
-This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades (Centeno & Cohen, 2012; Harvey, 2007), affecting areas such as pensions, health services, and education.
-The justification of social inequality based on marketype criteria has been conceptualized as the individuals’ adherence to one specific justice evaluation, the market justice, that is, affording legitimacy to the allocation of goods and services based on prices and individuals’ ability to pay (Boltanski & Chiapello- 2007; Lane 1986; Streeck 2012).
-
-Robert E.
-Lane proposed the underpinnings of market justice, which he differentiated from political justice.
-For him, “it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets” (1986, p.
-384).
-Following the theory of relative deprivation –a social phenomenon arising when individuals cannot afford what most others in their environment can (Merton 1950)– Lane notes that, in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions.
-Although empirical research has shown that, contrary to Lane’s observation, relative deprivation, even in market settings, produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action such as protests and revolt (Greitemeyer and Sagioglou 2016; Mishra and Carleton 2015; Smith et al.
-2012; Power 2018; Power, Madsen, and Morton 2020), the conceptualization of market justice has been since then closely coupled to the merit principle for allocating outcomes: the claim that the unequal levels of well-being individuals enjoy ought to be, to some extent, a function of their talents and efforts, regardless of their needs or membership, the two latter being the realm of political justice and its closely coupled principles of need –allocating outcomes to those who require them most– and equality –allocating the same outcome to everyone–, which have been at the center of welfarism (see Wilson 2003).
-The underlying notion of market justice also resonates in its application to welfare regimes, providing a framework for understanding the varied global approaches to managing social services.
-
-The management of social services manifests in varying approaches across nations, with substantial differences in funding and delivery methods (Jensen, 2008; Stoy, 2014).
-Nordic countries, for example, predominantly employ public agencies to produce and provide social services, funding these through collective taxation and offering them in kind to the majority of citizens.
-This system prioritizes political justice, placing it above market mechanisms in accessing services.
-In contrast, other countries rely more heavily on for-profit entities and private funding, where service distribution depends mainly on individual financial capacity of paying user fees, highlighting the influence of market justice in service allocation.
-The trend toward marketization of welfare services has been growing since the 1980s (Salamon, 1993), and this shift is increasingly evident even in countries where market solutions have traditionally had a minor role in social policy (Sivesind, 2017).
-
-The question arises whether adults and children justify unequal access to welfare services based on market justice principles.
-Influenced by theories of policy feedback, which suggest that social welfare policies can reinforce (positive feedback) or undermine (negative feedback) previous policy trajectories [@fernandez_positive_2013] (Fernandez and Jaime-Castillo 2013; Pierson 2000; Weaver 2010), citizens’ beliefs about market justice are likely also shaped by the institutional and social contexts they encounter.
-Indeed, the justification of inequality in access to essential services like education and health, based on one’s ability to pay, shows significant variation across countries, as demonstrated by international surveys, although they do not usually include children in their samples.
-For instance, the 2019 International Social Survey Program (ISSP) provides insights into how adults perceive inequality in accessing welfare services.
-Figure 1 illustrates the variation in agreement levels regarding whether it is just for individuals with higher incomes to purchase better education or healthcare.
-Notably, in 18 of the 29 countries analyzed, there is a greater justification for market inequality in healthcare access than in education.
-Nevertheless, the general sentiment typically ranges from 'somewhat unjust' to 'neither just nor unjust, with mixed feelings.
-
-```{r}
-#| fig-cap: "Average of market justice preferences by country"
-```
-
-While contemporary policy developments have increasingly embraced the marketization of social welfare in areas such as education, pensions, and health services, questions about public acceptance of these changes persist.
-@lindh_public_2015 analysis of ISSP 2009 data from 17 OECD countries reveals a general lack of support for market-based distribution of social services, suggesting widespread disapproval of market stratification of essential services.
-This finding is corroborated by Soler-Martínez and colleagues’ (2023) research from Latinobarómetro 2020 across 18 Latin American countries, where concerns about health and education access predominated over income inequality.
-These results indicate that reforms toward welfare marketization are typically driven by elite political decisions rather than grassroots demand.
-
-Despite high-income inequality and limited social mobility in Latin America, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@buccaMeritBlameUnequal2016; @chongMysteryDiscriminationLatin2008; @torcheIntergenerationalMobilityInequality2014; @salgadoInequalityStratificationLatin2023].
-The reliance on private welfare providers and widespread user fees (Molyneux, 2008) adds complexity to this context.
-Yet, research on children’s justification of market-based inequalities in accessing welfare services remains limited, especially in Latin America, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.
-
-## Meritocracy
-
-The concept of meritocracy frequently appears nowadays when analyzing cultural determinants of social inequalities.
-In general, it is mentioned as a value associated with justice, as it would link efforts and talents with rewards in an equitable manner.
-This normative sense is quite far from its original formulation by @young_rise_1958 in the satirical novel "The Rise of Meritocracy", where it ironically represented a mechanism for reproducing the inequalities of origin.
-The meritocratic ideal had remained relatively unchallenged until a series of recent publications turned into its potential consequences for maintaining social inequality.
-Perhaps one of the most recent sources in this line is Michael Sandel's "The Tyranny of Merit", where he strongly questions the implications of carrying out the principle of merit in societies that do not guarantee equal opportunities and that generate a feeling of scarce recognition and appreciation of those who receive lesser rewards: "In society's eyes, and perhaps also their own, their work no longer signified a valued contribution to the common good." [@sandel_tyranny_2020, pp.].
-
-Empirical research on meritocracy has increased along with the philosophical-normative discussion on meritocracy in recent years.
-Particularly from a sociological perspective, meritocracy has been used in research on social mobility to characterize societies with low mobility that threaten the meritocratic ideal [@goldthorpe_myth_2003].
-More recently, sociology and social psychology research has attended to the subjective aspects vis-a-vis beliefs in meritocracy.
-The label of beliefs in this realm covers a series of areas, such as attitudes, perceptions, and preferences (Castillo et al), whereby most of the link this subjective dimensions to individual socio-structural factors and  context-level determinants.
-For instance, some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2021a], and how larger inequality affects meritocratic beliefs (Morris et al 2022).
-Based on these findings, a research agenda has been reinforced on the legitimizing role of meritocracy, in line with previous studies using the concept of a just world (Lerner, Dalbert) and the theory of system justification (Jost & Major).
-
-How do meritocratic beliefs legitimize inequalities?
-Empirical studies have used experiments and surveys to address this question.
-For instance, the evidence suggests that just world beliefs correlate negatively with support for redistributive compensation systems [@frank_performance_2015].
-Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed [@evans_strong_2018].
-@almas_cutthroat_2020 found that in a relatively unequal society (the United States), the highly educated accept inequality significantly more than the less educated because they perceive inequality as justifiable owing to differences in productivity (i.e., merit), whereas in a relatively equal society (Norway), the less educated accept inequality more, but not significantly more than the highly educated because meritocratic values are less prevalent.
-@barr_effect_2020 also addressed this triple interaction between the level of inequality in a society, the individual level of education, and the perceived origin of the disparity (either by luck or effort) to determine the extent to which inequality is accepted.
-They found that the interaction's mechanism varies depending on the compared societies.
-Finally, @garcia-sanchez_attitudes_2020, using data for 41 countries from the International Social Survey Programme (ISSP), found that the perceived size of the income gap correlated positively with support for progressive taxation.
-Still, this association was weaker among those who endorsed meritocratic and equal opportunity beliefs.
-In the same line, experimental research by Durante & Putterman (2009) subjects support less redistribution when the initial distribution is determined according to task performance.
-
-Research about meritocratic beliefs at school age is rather scarce, leaving a wide research gap as schools are one of the primary socialization institutions where achievement based on merit explains success (Erivwo et al 2021).
-Nevertheless, it is possible to find some initial works focused on the area of distributive justice at school that are closely related to meritocratic beliefs, as the ones by Resh and Sabbagh (quotes).
-Using justice in grade obtention as a measure of distributive justice and meritocracy (Sabbagh et al 2006), they find for instance that a larger sense of distributive justice about grades is associated to higher socio-economic status [@resh_sense_2010], have a positive effect on liberal democratic orientation and on trust in people and in formal institutions (Resh & Sabbagh 2013), and tend to refrain from violence and to engage to a greater extent in extra-curricular school activity and community volunteering [@resh_sense_2017].
-
-## Children's judgments of inequality, schools, and family background
-
-Research indicates various socialization practices at families and schools during childhood and adolescence that impact dispositional and behavioral tendencies concerning justifications of inequality in adult life.
-The differences in economic understanding across age groups are consistent with research on cognitive development (Choudhury et al., 2006).
-Adolescents with mature socio-cognitive abilities tend to express stronger preferences for fairness than infants and children (Wynn et al., 2017).
-As children grow older, they become more likely to behave fairly, with their early-emerging strict egalitarianism being replaced by an increasing endorsement of fairness principles and engagement in collaborative activities (Huppert et al., 2019; McAuliffe et al., 2017).
-In these activities, their fairness views consider individual contributions, merits, and circumstances (Almås et al., 2010; Huppert et al., 2019; Sigelman & Waitzman, 1991).
-Engelmann and Tomasello (2019) claim that children’s sense of fairness emerges at three years old, and we can observe it in collaborative activities, where they accept inequality if the procedure gives everyone an equal chance.
-Therefore, children at this age respond to unequal distributions based on interpersonal concerns, as they already demand equal respect.
-In any case, between 3 and 8 years of age, inequitable and anti-meritorious allocations are evaluated more negatively, but equitable and meritorious allocations are not evaluated more positively (Elenbaas, 2019).
-
-Some research shows that the social environment in which children develop, such as family and school, is associated with their prosocial behaviors by playing an essential role in the transmission of equity norms (Schunk & Zipperle, 2023; Kosse et al., 2020).
-In fact, schools contribute to institutionalizing and reproducing inequality by promoting values, norms, practices, and languages familiar to higher-class families because the dominant group’s culture shapes educational institutions (Bourdieu & Passeron, 1990).
-Middle- and upper-class students are better equipped to face academic challenges and are more familiar with academic expectations (Mikus et al., 2019).
-Such familiarity represents cultural capital in educational contexts because higher-status students come to school ready to meet these expectations and reap the benefits (Jack, 2016; Khan, 2011).
-Conversely, lower-status children lacking cultural capital must catch up while experiencing inequitable comparisons (Goudeau & Croizet, 2017).
-Additionally, academic achievement is treated as the outcome of dispositional factors (e.g., pupils’ efforts and talents or lack of them) rather than the result of differential access to critical resources.
-Due to the meritocratic frame schools encourage, both low- and high-status individuals believe that students’ success or failure is not due to their family background but rather to differences in efforts and talents (Darnon et al., 2018).
-In this sense, we believe that the perception of meritocracy can influence students' judgments about market justice preferences.
-Furthermore, we believe there is a difference between students' perceptions of meritocracy based on their own experience in school and what they perceive in society at large.
-Consistently, our first two hypotheses are:
-
-H1a: Students who perceive that there is more meritocracy at school will show larger market justice preferences
-
-H1b: Students who perceive that there is more meritocracy in society will show larger market justice preferences
-
-Family background and family socialization practices also contribute to children’s and adolescent’s market justice preferences.
-For example, Almås and colleagues (2017) found that adolescents from low-socioeconomic-status families are likelier to have an egalitarian fairness view and consider an equal distribution as fair in a situation with unequal merits.
-The authors speculate that differences in socialization practices across status groups might bring about, to a great extent, the fairness views of children and adolescents because social status seems to interact with these evaluations (e.g., Hvidberg et al., 2023).
-
-The classic work of Kohn showed that middle-class parents value the expression of internal states and emotions, such as self-control, curiosity, happiness, and consideration, while working-class parents promote deference, obedience, and conformity to authority (Kohn, 1963; Kohn & Schooler, 1969).
-Although parents from all social backgrounds encourage individualism in their children, this shared norm translates into different forms in high and low social classes (1999).
-Acemoglu (2021) claimed that the values families impart to their children interact with social mobility.
-Because obedience is a valuable characteristic for employers, in low-wage and social mobility environments, low-income families impart values of obedience to their children to prevent disadvantaging them in labor markets.
-On the one hand, children from privileged families are socialized to adopt a clear conception of individualism that highlights their internal states, independence, and idiosyncrasies.
-In contrast, children from disadvantaged families are socialized to support a more balanced view of individualism that considers personal characteristics as resources to overcome collective impediments on the path to upward mobility (Iacoviello & Lorenzi-Cioldi, 2019).
-In this way, we believe that there are differences in the socialization of values according to socioeconomic differences that could influence market justice preferences and, therefore, we propose the following hypothesis:
-
-H2: Students from families of higher social status will show larger market justice preferences.
-
-Recent empirical research has demonstrated that the institutional design of schools, coupled with the meritocratic ideology it fosters, significantly influences children’s and adolescents’ views on inequality and deservingness.
-For example, Jonsson and Beach’s (2015) study revealed that higher-status adolescents in Sweden tend to perpetuate social class stereotypes while describing the vocational and academic tracks.
-Academic track students are depicted as wealthy, intelligent, ambitious, and diligent, while vocational track students are characterized as poor, unambitious, unintelligent, and lackadaisical.
-These stereotypes help individuals maintain a sense of superiority over others and legitimize the prevailing social hierarchies and economic disparities (Jost & Burgess, 2000).
-
-H3: Students from schools of higher social status will show larger market justice preferences.
-
-H4: Students from schools with higher average levels of academic achievement will show larger market justice preferences.
-
-\[interacciones\]
-
-H5: The perception of meritocracy in school and society will moderate the effect of family social status on market justice preferences.
-
-H6: The perception of meritocracy in school and society will moderate the effect of school status on market justice preferences.
-
-H7: The perception of meritocracy in school and society will moderate the effect of school academic achievement on market justice preferences.
-
-### Summary of hypotheses
-
-```{r}
-#| fig-cap: "Summary of hypotheses"
-```
-
-## Methods
-
-### Data
-
-The main data source for the analysis is the First Study of Citizenship Education in Chile, carried out by the Education Quality Agency of the Ministry of Education.
-The application date was November 9, 2017.
-The target population of this study is Eighth-grade students from 242 schools.
-In the data, there are 8,589 students and 6,770 parents.
-This database was analyzed with the R package “ResponsePatterns” to detect possible repetitive and “careless” response patterns and thus contribute to a higher quality of research data (Gottfried, et al 2022).
-Responses from 171 students and 79 parents were removed, which, when merged, gave a total of 6,270 valid cases.
-
-The analysis of school variables includes data from the Ministry of Education's SIMCE 2017 database.
-This database contains information at the school level, such as the administrative dependency, its socioeconomic classification, and the achievement scores obtained in the mathematics and language census tests.
-It is available for free use on the MINEDUC \[web page\].
-
-After eliminating missing cases, the final sample used in the analysis was based on 5,047 students and parents of 231 schools for the dependent variable of access to social services.
-
-### Variables
-
-**Dependent variables**
-
-This study has three dependent variables related to the justification of social inequality in specific policy domains.
-The first asks whether access to social services should be conditional on income, i,.e., “It is just that in Chile people who can pay have a better education for their children”.
-Students rated their preferences using the following responses: “strongly disagree”, “Disagree”, “Agree”, and ”strongly agree”.
-An average index is built with these items (Cronbach's Alpha = 0,86).
-(Apéndice: items en español y su correspondiente traducción al inglés)
-
-Table 1 shows the items used, their response categories, and their frequencies.
-
-```{r}
-#| tbl-cap: "Dependent variables"
-```
-
-**Independent variables**
-
-For the primary independent variable, the perception of meritocracy, five items address the perception of rewards according to effort and intelligence at the school and societal levels.
-At the school level, students answer whether “Intelligence is important to get good grades” and “Effort is important to get good grades”.
-At the societal level, students respond to the following questions: “In Chile, people are rewarded for their efforts”, “In Chile, people get what they deserve”, and “In Chile, people are rewarded for their intelligence and skills”.
-Each item was answered on a four-point scale ranging from “Strongly disagree” to “Strongly agree”.
-
-The rest of the independent variables are divided into individual and school levels.
-At the individual level, family socioeconomic status was measured by the parents' highest educational level and the number of books at home.
-Likewise, an index of access to technology includes the number of computers, tablets, and cell phones at home, as well as whether there is an Internet connection.
-Table 2 shows the items used, their response categories, and their frequency.
-
-```{r}
-#| tbl-cap: "Independent variables"
-```
-
-The school-level variables are the administrative dependency of the school, the socioeconomic classification made by the Ministry of Education, the level of performance in the SIMCE test of the school, and the proportion of parents with university or postgraduate degrees.
-Table 3 shows the items used, response categories, and frequency.
-
-```{r}
-#| tbl-cap: "School context variables"
-```
-
-### Methods
-
-The data has a hierarchical structure of students nested in schools, so the model estimation is performed in a multilevel (random effects) framework.
-This modeling approach lets us correctly estimate individual and contextual effects in a single model.
-We estimate cumulative link mixed models for the ordinal dependent variables, whereas we use linear mixed effects models for the average index of inequality justification.
-
-The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open ScorrectlyOSF), the access to the document is available at this \[link\](https://doi.org/10.17605/OSF.IO/UFSDV).
-The statistical analysis of this research was performed using the free software R version 4.1.3.
-
-## Analysis
-
-### Descriptive analysis
-
-```{r}
-#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
-```
-
-Figure 3 shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram en figure X).
-On the left, we observe the social meritocracy diagrams, while on the right the school meritocracy diagrams are shown.
-For the three variables of perception of meritocracy in society the relationship is clear, since the average of market justice preferences increases the more there is agreement that people are rewarded for their effort, merit, and talent.
-This relationship needs to be clarified in the case of the variables of perception of meritocracy at school.
-To the extent that there is more agreement that the perception of talent is essential for obtaining good grades, the average of market justice preferences increases, but this relationship is not as clear as with the variables of meritocracy in society.
-In addition, the graph does not show a clear trend in the relationship between the perception that effort is essential to obtain good grades and market justice preferences.
+##

### 2024-04-24 01-introduction.qmd
render paper

@@
 ---
-title: "01-introduction"
 author: "Equipo EDUMER"
@@
 Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work.
-In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2021a].
+In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2021].
 Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger justification of social inequalities, as individual achievement would be seen as rewarded and social policies as less necessary [@batruch_belief_2022].

### 2024-04-24 02-antecedentes.qmd
render paper

@@
 ---
-title: "02-antecedentes"
 author: "Equipo EDUMER"
@@
 
-Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegelBeliefsInequalityAmericans1987], highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness.
-The influence of ideologies [@wegenerDominantIdeologiesVariation1995] and cultural schemas [@homanBeingBecomingPoor2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations.
-While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castilloLegitimacyInequalityHighly2011; ] (Castillo, 2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),  there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles.
+Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegel_beliefs_1987], highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness.
+The influence of ideologies [@wegener_dominant_1995] and cultural schemas [@homan_being_2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations.
+While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; ] (Castillo, 2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),  there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles.
 This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades (Centeno & Cohen, 2012; Harvey, 2007), affecting areas such as pensions, health services, and education.

### 2024-04-24 paper.tex
render paper

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
 \fi
@@
 \makeatother
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
+  \fi
+  % set entry spacing
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
 \KOMAoption{captions}{tableheading}
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
 \makeatletter
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
 \@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
 \makeatother
@@
 \fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \hypersetup{
-  pdftitle={market-justice-preferences},
   pdfauthor={Equipo EDUMER},
@@
 
-\title{market-justice-preferences}
-\author{Equipo EDUMER}
+\author{}
 \date{}
@@
 \begin{document}
-\maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden]}{\end{tcolorbox}}\fi
 
-\hypertarget{the-socialization-of-meritocracy-and-market-justice-preferences-at-school}{%
 \section{The socialization of meritocracy and market justice preferences
 at
-school}\label{the-socialization-of-meritocracy-and-market-justice-preferences-at-school}}
+school}\label{the-socialization-of-meritocracy-and-market-justice-preferences-at-school}
 
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+\subsection{Introduction}\label{introduction}
 
@@
 school level represents a threat to the promise of education and a
-meritocratic system {[}@bourdieu\_reproduction\_1990{]}. A large part of
-the research in this field at an international level has addressed the
-extent to which the social origin of students affects their academic
-results and their life opportunities {[}@vonhippel\_test\_2019{]},
+meritocratic system (\citeproc{ref-bourdieu_reproduction_1990}{Bourdieu
+and Passeron 1990}). A large part of the research in this field at an
+international level has addressed the extent to which the social origin
+of students affects their academic results and their life opportunities
+(\citeproc{ref-vonhippel_test_2019}{Von Hippel and Hamrock 2019}),
 confirming that schools have severe difficulties in closing the gaps of
@@
 social inequalities are tolerated and even justified?
-{[}@batruch\_belief\_2022; @wiederkehr\_belief\_2015{]}.~
+(\citeproc{ref-batruch_belief_2022}{Batruch et al. 2022};
+\citeproc{ref-wiederkehr_belief_2015}{Wiederkehr et al. 2015}).~
+
+In the present paper, we deal with the justification of social
+inequalities by eighth-grade students in Chile, a country characterized
+by a highly stratified educational system. In particular, we focus on
+market justice preferences (Lane, 1986), which refer to the preferences
+for distributing public goods (as health and education) based on
+criteria such as competition and payment capacity. Although from a
+rational point of view it could be expected an opposition to market
+justice by the underprivileged majority,~ we argue that in a social
+environment characterized by a promotion of meritocratic ideals - as the
+school system - would lead to the opposite: a larger market justice
+preferences.~
+
+Given that the school environment has an important focus on performance,
+achievement and acknowledgment, meritocracy has been one of the
+principal concepts used for understanding and even for justifying
+performance differences among students. Meritocracy is a distributive
+system based on the belief that people should be rewarded and promoted
+based on their abilities, knowledge, and achievements
+(\citeproc{ref-young_rise_1958}{Young 1958}). It is often seen as a way
+to create equal opportunities and fairness, as individuals can rise to
+positions of power and influence based on their own merit rather than
+their background or connections. However, some argue that meritocracy
+can actually lead to tolerating or even justifying social inequalities,
+as it can create a hierarchy where those who already have resources and
+advantages are more likely to succeed. In this regard, a great deal of
+academic research about meritocracy delves into the assessment of to
+what extent rewards and privileges in society are related to merit,
+emphasizing the so-called unfulfillable promise of meritocracy
+(\citeproc{ref-mijs_stratified_2016}{J. Mijs 2016}). Complementing this
+agenda, a second and emerging research area deals with subjective
+aspects of meritocracy, such as perceptions and beliefs.~
+
+{[}definición de market justice, welfare marketization, (Lindh){]}
+
+{[}contexto Chileno market justice, dar cifras/porcentajes de
+privatización de salud, educación y pensiones; asociar al tema de la
+calidad (lo público es peor){]}
+
+The perception of meritocracy refers to how individuals view and
+understand the concept of meritocracy in their own society
+(\citeproc{ref-duru-bellat_who_2012}{Duru-Bellat and Tenret 2012};
+\citeproc{ref-castillo_meritocracia_2019}{Juan C. Castillo et al.
+2019}). This perception can vary greatly depending on individual
+experiences, social, economic, and cultural background. Some people may
+see meritocracy as a fair and just system that allows anyone to succeed
+based on their abilities and hard work. In contrast, others may view it
+as a myth or a cover for existing power dynamics and inequality, serving
+to maintain and even reinforce inequality
+(\citeproc{ref-lampert_meritocratic_2013}{Lampert 2013};
+\citeproc{ref-mijs_paradox_2021}{J. J. B. Mijs 2021}). Based in this
+last perspective, we argue that individuals with a higher perception of
+meritocracy will show a larger justification of social inequalities, as
+individual achievement would be seen as rewarded and social policies as
+less necessary (\citeproc{ref-batruch_belief_2022}{Batruch et al.
+2022}).
+
+Most of the research that has related meritocratic beliefs to inequality
+justification so far has only considered adults, leaving aside the study
+of how beliefs in this field develop at student age as well as the
+impact of the school context and the family as the main socialization
+agencies. Regarding schools, the way in which they deal with unequal
+conditions of origin has been linked to the \emph{hidden curriculum}
+(\citeproc{ref-chafel_schooling_1997}{Chafel 1997}), whereby students
+learn about distributive norms in society and mechanisms of
+justification of social differences. Based on recent studies that relate
+school meritocracy to the justification of economic inequalities in the
+adult population (\citeproc{ref-batruch_belief_2022}{Batruch et al.
+2022}; \citeproc{ref-wiederkehr_belief_2015}{Wiederkehr et al. 2015}),
+the central hypothesis guiding this research is that school-age students
+with a higher perception of meritocracy - both at school and at the
+societal level - will show a larger justification of social
+inequalities, as individual achievement would be seen as appropriately
+rewarded and social mechanisms for correcting inequalities as less
+necessary (\citeproc{ref-batruch_belief_2022}{Batruch et al. 2022}).
+
+\textasciitilde{}
+
+{[}introducir distinción sobre percepción de meritocracia a nivel
+escolar y a nivel de la sociedad{]}
+
+{[}resaltar 3 focos de innovación: market justice preferences at school
+level, relación con meritocracia, y esquema de meritocracia en la
+escuela y en la sociedad{]}
+
+The present paper deals with the association between the perception of
+meritocracy and the justification of social inequalities, with two main
+focuses. Firstly, it assesses the justification of social inequalities
+in social policy domains, such as health, pensions, and education. We
+argue that individuals who perceive more meritocracy would be more
+willing to justify better services in these domains for those with
+higher incomes.~ Secondly, we focus on the student-age population as we
+point out that it is possible to track down the origin of meritocratic
+beliefs (and their consequences) to early socialization processes. To
+this regard, we take into account the family and the school as two main
+socialization agencies that play a significant role in the socialization
+of cultural beliefs by transmitting cultural norms, values, and
+expectations to young people.~
+
+\subsection{Justification of
+inequality}\label{justification-of-inequality}
+
+Research on social stratification beliefs, which explore individual
+perceptions of who deserves what and why
+(\citeproc{ref-kluegel_beliefs_1987}{Kluegel and Smith 1987}),
+highlights that people's explanations and justifications of social
+inequality are closely tied to their judgments of deservingness. The
+influence of ideologies (\citeproc{ref-wegener_dominant_1995}{Wegener
+and Liebig 1995}) and cultural schemas
+(\citeproc{ref-homan_being_2017}{Homan, Valentino, and Weed 2017}) is
+pivotal in shaping these explanations by offering symbolic
+representations that frame societal structures and expectations. While
+significant attention has been paid to wage inequality, income
+distribution, and payment differentials in the literature {[}Juan Carlos
+Castillo (\citeproc{ref-castillo_legitimacy_2011}{2011}); {]} (Castillo,
+2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),~ there has
+been less examination of public beliefs about which life domains should
+be governed by market relations and even less about children's
+acceptance or rejection of these market principles. This oversight is
+notable given the extensive encroachment of market logic into public
+goods, welfare policy, and social services over the past five decades
+(Centeno \& Cohen, 2012; Harvey, 2007), affecting areas such as
+pensions, health services, and education. The justification of social
+inequality based on marketype criteria has been conceptualized as the
+individuals' adherence to one specific justice evaluation, the market
+justice, that is, affording legitimacy to the allocation of goods and
+services based on prices and individuals' ability to pay (Boltanski \&
+Chiapello- 2007; Lane 1986; Streeck 2012).
+
+Robert E. Lane proposed the underpinnings of market justice, which he
+differentiated from political justice. For him, ``it is the genius of
+the market to stimulate wants without at the same time stimulating a
+sense of deserving more than one gets'' (1986, p.~384). Following the
+theory of relative deprivation --a social phenomenon arising when
+individuals cannot afford what most others in their environment can
+(Merton 1950)-- Lane notes that, in market settings, social comparisons
+are more likely to motivate increased effort rather than feelings of
+acute injustice because individuals attribute outcomes to their actions.
+Although empirical research has shown that, contrary to Lane's
+observation, relative deprivation, even in market settings, produces
+feelings of dissatisfaction, anger, and resentment that might motivate
+forms of collective action such as protests and revolt (Greitemeyer and
+Sagioglou 2016; Mishra and Carleton 2015; Smith et al.~2012; Power 2018;
+Power, Madsen, and Morton 2020), the conceptualization of market justice
+has been since then closely coupled to the merit principle for
+allocating outcomes: the claim that the unequal levels of well-being
+individuals enjoy ought to be, to some extent, a function of their
+talents and efforts, regardless of their needs or membership, the two
+latter being the realm of political justice and its closely coupled
+principles of need --allocating outcomes to those who require them
+most-- and equality --allocating the same outcome to everyone--, which
+have been at the center of welfarism (see Wilson 2003). The underlying
+notion of market justice also resonates in its application to welfare
+regimes, providing a framework for understanding the varied global
+approaches to managing social services.
+
+The management of social services manifests in varying approaches across
+nations, with substantial differences in funding and delivery methods
+(Jensen, 2008; Stoy, 2014). Nordic countries, for example, predominantly
+employ public agencies to produce and provide social services, funding
+these through collective taxation and offering them in kind to the
+majority of citizens. This system prioritizes political justice, placing
+it above market mechanisms in accessing services. In contrast, other
+countries rely more heavily on for-profit entities and private funding,
+where service distribution depends mainly on individual financial
+capacity of paying user fees, highlighting the influence of market
+justice in service allocation. The trend toward marketization of welfare
+services has been growing since the 1980s (Salamon, 1993), and this
+shift is increasingly evident even in countries where market solutions
+have traditionally had a minor role in social policy (Sivesind, 2017).
+
+The question arises whether adults and children justify unequal access
+to welfare services based on market justice principles. Influenced by
+theories of policy feedback, which suggest that social welfare policies
+can reinforce (positive feedback) or undermine (negative feedback)
+previous policy trajectories
+(\citeproc{ref-fernandez_positive_2013}{Fernandez and Jaime-Castillo
+2013}) (Fernandez and Jaime-Castillo 2013; Pierson 2000; Weaver 2010),
+citizens' beliefs about market justice are likely also shaped by the
+institutional and social contexts they encounter. Indeed, the
+justification of inequality in access to essential services like
+education and health, based on one's ability to pay, shows significant
+variation across countries, as demonstrated by international surveys,
+although they do not usually include children in their samples. For
+instance, the 2019 International Social Survey Program (ISSP) provides
+insights into how adults perceive inequality in accessing welfare
+services. Figure 1 illustrates the variation in agreement levels
+regarding whether it is just for individuals with higher incomes to
+purchase better education or healthcare. Notably, in 18 of the 29
+countries analyzed, there is a greater justification for market
+inequality in healthcare access than in education. Nevertheless, the
+general sentiment typically ranges from `somewhat unjust' to 'neither
+just nor unjust, with mixed feelings.
+
+While contemporary policy developments have increasingly embraced the
+marketization of social welfare in areas such as education, pensions,
+and health services, questions about public acceptance of these changes
+persist. Lindh (\citeproc{ref-lindh_public_2015}{2015}) analysis of ISSP
+2009 data from 17 OECD countries reveals a general lack of support for
+market-based distribution of social services, suggesting widespread
+disapproval of market stratification of essential services. This finding
+is corroborated by Soler-Martínez and colleagues' (2023) research from
+Latinobarómetro 2020 across 18 Latin American countries, where concerns
+about health and education access predominated over income inequality.
+These results indicate that reforms toward welfare marketization are
+typically driven by elite political decisions rather than grassroots
+demand.~
+
+Despite high-income inequality and limited social mobility in Latin
+America, there is a prevalent belief that individuals are solely
+responsible for their economic outcomes, a view that varies across the
+region
+(\citeproc{ref-buccaMeritBlameUnequal2016}{\textbf{buccaMeritBlameUnequal2016?}};
+\citeproc{ref-chongMysteryDiscriminationLatin2008}{\textbf{chongMysteryDiscriminationLatin2008?}};
+\citeproc{ref-torcheIntergenerationalMobilityInequality2014}{\textbf{torcheIntergenerationalMobilityInequality2014?}};
+\citeproc{ref-salgadoInequalityStratificationLatin2023}{\textbf{salgadoInequalityStratificationLatin2023?}}).
+The reliance on private welfare providers and widespread user fees
+(Molyneux, 2008) adds complexity to this context. Yet, research on
+children's justification of market-based inequalities in accessing
+welfare services remains limited, especially in Latin America,
+highlighting a significant gap in understanding how younger generations
+view market-based access to welfare and whether these views are
+associated with their meritocratic beliefs.~
+
+\subsection{Meritocracy}\label{meritocracy}
+
+The concept of meritocracy frequently appears nowadays when analyzing
+cultural determinants of social inequalities. In general, it is
+mentioned as a value associated with justice, as it would link efforts
+and talents with rewards in an equitable manner. This normative sense is
+quite far from its original formulation by Young
+(\citeproc{ref-young_rise_1958}{1958}) in the satirical novel ``The Rise
+of Meritocracy'', where it ironically represented a mechanism for
+reproducing the inequalities of origin. The meritocratic ideal had
+remained relatively unchallenged until a series of recent publications
+turned into its potential consequences for maintaining social
+inequality. Perhaps one of the most recent sources in this line is
+Michael Sandel's ``The Tyranny of Merit'', where he strongly questions
+the implications of carrying out the principle of merit in societies
+that do not guarantee equal opportunities and that generate a feeling of
+scarce recognition and appreciation of those who receive lesser rewards:
+``In society's eyes, and perhaps also their own, their work no longer
+signified a valued contribution to the common good.''
+(\citeproc{ref-sandel_tyranny_2020}{Sandel 2020}, pp.).
+
+Empirical research on meritocracy has increased along with the
+philosophical-normative discussion on meritocracy in recent years.
+Particularly from a sociological perspective, meritocracy has been used
+in research on social mobility to characterize societies with low
+mobility that threaten the meritocratic ideal
+(\citeproc{ref-goldthorpe_myth_2003}{Goldthorpe 2003}). More recently,
+sociology and social psychology research has attended to the subjective
+aspects vis-a-vis beliefs in meritocracy. The label of beliefs in this
+realm covers a series of areas, such as attitudes, perceptions, and
+preferences (Castillo et al), whereby most of the link this subjective
+dimensions to individual socio-structural factors and~ context-level
+determinants. For instance, some studies have analyzed how those with
+greater privileges believe more in meritocracy (Reynolds \& Chan 2014),
+how greater economic inequality increases meritocratic beliefs
+(\citeproc{ref-mijs_paradox_2021a}{\textbf{mijs\_paradox\_2021a?}}), and
+how larger inequality affects meritocratic beliefs (Morris et al 2022).
+Based on these findings, a research agenda has been reinforced on the
+legitimizing role of meritocracy, in line with previous studies using
+the concept of a just world (Lerner, Dalbert) and the theory of system
+justification (Jost \& Major).
+
+How do meritocratic beliefs legitimize inequalities? Empirical studies
+have used experiments and surveys to address this question. For
+instance, the evidence suggests that just world beliefs correlate
+negatively with support for redistributive compensation systems
+(\citeproc{ref-frank_performance_2015}{Frank, Wertenbroch, and Maddux
+2015}). Conversely, individuals tend to support redistribution when they
+believe that the disadvantaged lack the opportunities to succeed
+(\citeproc{ref-evans_strong_2018}{Evans and Kelley 2018}). Almås,
+Cappelen, and Tungodden (\citeproc{ref-almas_cutthroat_2020}{2020})
+found that in a relatively unequal society (the United States), the
+highly educated accept inequality significantly more than the less
+educated because they perceive inequality as justifiable owing to
+differences in productivity (i.e., merit), whereas in a relatively equal
+society (Norway), the less educated accept inequality more, but not
+significantly more than the highly educated because meritocratic values
+are less prevalent. Barr and Miller
+(\citeproc{ref-barr_effect_2020}{2020}) also addressed this triple
+interaction between the level of inequality in a society, the individual
+level of education, and the perceived origin of the disparity (either by
+luck or effort) to determine the extent to which inequality is accepted.
+They found that the interaction's mechanism varies depending on the
+compared societies. Finally, García-Sánchez et al.
+(\citeproc{ref-garcia-sanchez_attitudes_2020}{2020}), using data for 41
+countries from the International Social Survey Programme (ISSP), found
+that the perceived size of the income gap correlated positively with
+support for progressive taxation. Still, this association was weaker
+among those who endorsed meritocratic and equal opportunity beliefs. In
+the same line, experimental research by Durante \& Putterman (2009)
+subjects support less redistribution when the initial distribution is
+determined according to task performance.
+
+Research about meritocratic beliefs at school age is rather scarce,
+leaving a wide research gap as schools are one of the primary
+socialization institutions where achievement based on merit explains
+success (Erivwo et al 2021). Nevertheless, it is possible to find some
+initial works focused on the area of distributive justice at school that
+are closely related to meritocratic beliefs, as the ones by Resh and
+Sabbagh (quotes). Using justice in grade obtention as a measure of
+distributive justice and meritocracy (Sabbagh et al 2006), they find for
+instance that a larger sense of distributive justice about grades is
+associated to higher socio-economic status
+(\citeproc{ref-resh_sense_2010}{Resh 2010}), have a positive effect on
+liberal democratic orientation and on trust in people and in formal
+institutions (Resh \& Sabbagh 2013), and tend to refrain from violence
+and to engage to a greater extent in extra-curricular school activity
+and community volunteering (\citeproc{ref-resh_sense_2017}{Resh and
+Sabbagh 2017}).
+
+\subsection{Children's judgments of inequality, schools, and family
+background}\label{childrens-judgments-of-inequality-schools-and-family-background}
+
+Research indicates various socialization practices at families and
+schools during childhood and adolescence that impact dispositional and
+behavioral tendencies concerning justifications of inequality in adult
+life. The differences in economic understanding across age groups are
+consistent with research on cognitive development (Choudhury et al.,
+2006). Adolescents with mature socio-cognitive abilities tend to express
+stronger preferences for fairness than infants and children (Wynn et
+al., 2017). As children grow older, they become more likely to behave
+fairly, with their early-emerging strict egalitarianism being replaced
+by an increasing endorsement of fairness principles and engagement in
+collaborative activities (Huppert et al., 2019; McAuliffe et al., 2017).
+In these activities, their fairness views consider individual
+contributions, merits, and circumstances (Almås et al., 2010; Huppert et
+al., 2019; Sigelman \& Waitzman, 1991). Engelmann and Tomasello (2019)
+claim that children's sense of fairness emerges at three years old, and
+we can observe it in collaborative activities, where they accept
+inequality if the procedure gives everyone an equal chance. Therefore,
+children at this age respond to unequal distributions based on
+interpersonal concerns, as they already demand equal respect. In any
+case, between 3 and 8 years of age, inequitable and anti-meritorious
+allocations are evaluated more negatively, but equitable and meritorious
+allocations are not evaluated more positively (Elenbaas, 2019).
+
+Some research shows that the social environment in which children
+develop, such as family and school, is associated with their prosocial
+behaviors by playing an essential role in the transmission of equity
+norms (Schunk \& Zipperle, 2023; Kosse et al., 2020). In fact, schools
+contribute to institutionalizing and reproducing inequality by promoting
+values, norms, practices, and languages familiar to higher-class
+families because the dominant group's culture shapes educational
+institutions (Bourdieu \& Passeron, 1990). Middle- and upper-class
+students are better equipped to face academic challenges and are more
+familiar with academic expectations (Mikus et al., 2019). Such
+familiarity represents cultural capital in educational contexts because
+higher-status students come to school ready to meet these expectations
+and reap the benefits (Jack, 2016; Khan, 2011). Conversely, lower-status
+children lacking cultural capital must catch up while experiencing
+inequitable comparisons (Goudeau \& Croizet, 2017). Additionally,
+academic achievement is treated as the outcome of dispositional factors
+(e.g., pupils' efforts and talents or lack of them) rather than the
+result of differential access to critical resources. Due to the
+meritocratic frame schools encourage, both low- and high-status
+individuals believe that students' success or failure is not due to
+their family background but rather to differences in efforts and talents
+(Darnon et al., 2018). In this sense, we believe that the perception of
+meritocracy can influence students' judgments about market justice
+preferences. Furthermore, we believe there is a difference between
+students' perceptions of meritocracy based on their own experience in
+school and what they perceive in society at large. Consistently, our
+first two hypotheses are:
+
+H1a: Students who perceive that there is more meritocracy at school will
+show larger market justice preferences
+
+H1b: Students who perceive that there is more meritocracy in society
+will show larger market justice preferences
+
+Family background and family socialization practices also contribute to
+children's and adolescent's market justice preferences. For example,
+Almås and colleagues (2017) found that adolescents from
+low-socioeconomic-status families are likelier to have an egalitarian
+fairness view and consider an equal distribution as fair in a situation
+with unequal merits. The authors speculate that differences in
+socialization practices across status groups might bring about, to a
+great extent, the fairness views of children and adolescents because
+social status seems to interact with these evaluations (e.g., Hvidberg
+et al., 2023).~
+
+The classic work of Kohn showed that middle-class parents value the
+expression of internal states and emotions, such as self-control,
+curiosity, happiness, and consideration, while working-class parents
+promote deference, obedience, and conformity to authority (Kohn, 1963;
+Kohn \& Schooler, 1969). Although parents from all social backgrounds
+encourage individualism in their children, this shared norm translates
+into different forms in high and low social classes (1999). Acemoglu
+(2021) claimed that the values families impart to their children
+interact with social mobility. Because obedience is a valuable
+characteristic for employers, in low-wage and social mobility
+environments, low-income families impart values of obedience to their
+children to prevent disadvantaging them in labor markets. On the one
+hand, children from privileged families are socialized to adopt a clear
+conception of individualism that highlights their internal states,
+independence, and idiosyncrasies. In contrast, children from
+disadvantaged families are socialized to support a more balanced view of
+individualism that considers personal characteristics as resources to
+overcome collective impediments on the path to upward mobility
+(Iacoviello \& Lorenzi-Cioldi, 2019). In this way, we believe that there
+are differences in the socialization of values according to
+socioeconomic differences that could influence market justice
+preferences and, therefore, we propose the following hypothesis:
+
+H2: Students from families of higher social status will show larger
+market justice preferences.
+
+Recent empirical research has demonstrated that the institutional design
+of schools, coupled with the meritocratic ideology it fosters,
+significantly influences children's and adolescents' views on inequality
+and deservingness. For example, Jonsson and Beach's (2015) study
+revealed that higher-status adolescents in Sweden tend to perpetuate
+social class stereotypes while describing the vocational and academic
+tracks. Academic track students are depicted as wealthy, intelligent,
+ambitious, and diligent, while vocational track students are
+characterized as poor, unambitious, unintelligent, and lackadaisical.
+These stereotypes help individuals maintain a sense of superiority over
+others and legitimize the prevailing social hierarchies and economic
+disparities (Jost \& Burgess, 2000).
+
+H3: Students from schools of higher social status will show larger
+market justice preferences.
+
+H4: Students from schools with higher average levels of academic
+achievement will show larger market justice preferences.
+
+{[}interacciones{]}
+
+H5: The perception of meritocracy in school and society will moderate
+the effect of family social status on market justice preferences.
+
+H6: The perception of meritocracy in school and society will moderate
+the effect of school status on market justice preferences.
+
+H7: The perception of meritocracy in school and society will moderate
+the effect of school academic achievement on market justice preferences.
+
+\subsubsection{Summary of hypotheses}\label{summary-of-hypotheses}
+
+\subsection{}\label{section}
+
+\subsection{Methods}\label{methods}
+
+\subsubsection{Data}\label{data}
+
+The main data source for the analysis is the First Study of Citizenship
+Education in Chile, carried out by the Education Quality Agency of the
+Ministry of Education. The application date was November 9, 2017. The
+target population of this study is Eighth-grade students from 242
+schools. In the data, there are 8,589 students and 6,770 parents. This
+database was analyzed with the R package ``ResponsePatterns'' to detect
+possible repetitive and ``careless'' response patterns and thus
+contribute to a higher quality of research data (Gottfried, et al 2022).
+Responses from 171 students and 79 parents were removed, which, when
+merged, gave a total of 6,270 valid cases.
+
+The analysis of school variables includes data from the Ministry of
+Education's SIMCE 2017 database. This database contains information at
+the school level, such as the administrative dependency, its
+socioeconomic classification, and the achievement scores obtained in the
+mathematics and language census tests. It is available for free use on
+the MINEDUC {[}web page{]}.
+
+After eliminating missing cases, the final sample used in the analysis
+was based on 5,047 students and parents of 231 schools for the dependent
+variable of access to social services.
+
+\subsubsection{Variables}\label{variables}
+
+\textbf{Dependent variables}
+
+This study has three dependent variables related to the justification of
+social inequality in specific policy domains. The first asks whether
+access to social services should be conditional on income, i,.e., ``It
+is just that in Chile people who can pay have a better education for
+their children''. Students rated their preferences using the following
+responses: ``strongly disagree'', ``Disagree'', ``Agree'', and
+''strongly agree''. An average index is built with these items
+(Cronbach's Alpha = 0,86). (Apéndice: items en español y su
+correspondiente traducción al inglés)
+
+Table 1 shows the items used, their response categories, and their
+frequencies.
+
+\textbf{Independent variables}
+
+For the primary independent variable, the perception of meritocracy,
+five items address the perception of rewards according to effort and
+intelligence at the school and societal levels. At the school level,
+students answer whether ``Intelligence is important to get good grades''
+and ``Effort is important to get good grades''. At the societal level,
+students respond to the following questions: ``In Chile, people are
+rewarded for their efforts'', ``In Chile, people get what they
+deserve'', and ``In Chile, people are rewarded for their intelligence
+and skills''. Each item was answered on a four-point scale ranging from
+``Strongly disagree'' to ``Strongly agree''.
+
+The rest of the independent variables are divided into individual and
+school levels. At the individual level, family socioeconomic status was
+measured by the parents' highest educational level and the number of
+books at home. Likewise, an index of access to technology includes the
+number of computers, tablets, and cell phones at home, as well as
+whether there is an Internet connection. Table 2 shows the items used,
+their response categories, and their frequency.
+
+The school-level variables are the administrative dependency of the
+school, the socioeconomic classification made by the Ministry of
+Education, the level of performance in the SIMCE test of the school, and
+the proportion of parents with university or postgraduate degrees. Table
+3 shows the items used, response categories, and frequency.
+
+\subsubsection{Methods}\label{methods-1}
+
+The data has a hierarchical structure of students nested in schools, so
+the model estimation is performed in a multilevel (random effects)
+framework. This modeling approach lets us correctly estimate individual
+and contextual effects in a single model. We estimate cumulative link
+mixed models for the ordinal dependent variables, whereas we use linear
+mixed effects models for the average index of inequality justification.
+
+The hypotheses of this research were pre-registered in the Open Science
+Framework platform of the Center for Open ScorrectlyOSF), the access to
+the document is available at this
+{[}link{]}(https://doi.org/10.17605/OSF.IO/UFSDV). The statistical
+analysis of this research was performed using the free software R
+version 4.1.3.
+
+\subsection{Analysis}\label{analysis}
+
+\subsubsection{Descriptive analysis}\label{descriptive-analysis}
+
+Figure 3 shows a series of graphs depicting the association between the
+variables of market justice preferences - in education, health, and
+pensions - and the variables of meritocratic perception at school
+(effort and talent) and in society (effort, talent, and deservingness)
+(see conceptual diagram en figure X). On the left, we observe the social
+meritocracy diagrams, while on the right the school meritocracy diagrams
+are shown. For the three variables of perception of meritocracy in
+society the relationship is clear, since the average of market justice
+preferences increases the more there is agreement that people are
+rewarded for their effort, merit, and talent. This relationship needs to
+be clarified in the case of the variables of perception of meritocracy
+at school. To the extent that there is more agreement that the
+perception of talent is essential for obtaining good grades, the average
+of market justice preferences increases, but this relationship is not as
+clear as with the variables of meritocracy in society. In addition, the
+graph does not show a clear trend in the relationship between the
+perception that effort is essential to obtain good grades and market
+justice preferences.
+
+\subsubsection{Cumulative link mixed models for the Justification of
+inequality in education, health and
+pensions}\label{cumulative-link-mixed-models-for-the-justification-of-inequality-in-education-health-and-pensions}
+
+Three Cumulative link mixed models were estimated for the ordinal
+dependent variables of justice in differential access to pensions,
+education, and health according to income. Figure 4 shows the estimation
+of this regression model containing all the variables used in the study
+for the three dependent variables separately. However, this figure shows
+only the effect of meritocracy variables on society and school as
+independent variables. Complete models are available in the appendix.
+
+Figure 4 shows that for the three dependent variables of justice in
+differential access to pensions, education, and health, the trend is
+similar. In the context of school meritocracy, the effects are mixed. As
+school talent increases, the justification for differentiated access to
+pensions, education, and health increases; on the contrary, as the
+school effort variable increases, the justification for differentiated
+access to pensions, education, and health decreases, keeping the rest of
+the independent variables constant. Regarding the variables of social
+meritocracy, the three variables of talent, effort, and deservingness
+show that as these increase, the justification for differentiated access
+to pensions, education, and health also increases.
+
+\subsubsection{Multilevel regression models for market justice
+preferences}\label{multilevel-regression-models-for-market-justice-preferences}
+
+Table 4 shows the results of the multilevel estimation for justice
+market preferences. For this variable, the intraclass correlation
+obtained shows that the variation between schools corresponds to 4\% of
+the variation of students' preferences. This means that there is low
+variance between schools and therefore limits the possibilities of
+finding effects at the aggregate level.
+
+Model 1 introduces social meritocratic variables: effort (whether
+efforts are rewarded), deservingness (people get what they deserve), and
+social talent (intelligence and skills are rewarded in society). In line
+with our hypotheses, the perception of a meritocratic society is
+positively related to the justification of inequality. Model 2 shows a
+mixed picture: while those perceiving that talent is rewarded also
+justify the inequality, the perception that effort is rewarded at school
+is negatively related to justification of inequality. Family background
+variables in Model 3 reveal that education and technology access are not
+related to justification of inequality, whereby we observe a negative
+impact of family cultural capital as measured by the number of books at
+home. While school socio-structural variables added in Model 4 show no
+significant effects, average achievement scores in the SIMCE test depict
+a negative relationship with the dependent variable, meaning that
+students that attend schools with better achievement scores on average
+justify less inequality.~
+
+In relation to model fit, when comparing the deviance with a model
+without predictors (null model), all the models have a statistically
+significant difference, with model 5 having the lowest deviance.
+According to Raudenbush and Bryk's (2002) estimate of R2, the level 1
+variance of model 5 is 0.11 and the level 2 variance is 0.72. The total
+variance of model 5 according to Snijders and Bosker (2012) is 0.14.
+
+\subsubsection{Interactions effects}\label{interactions-effects}
+
+The interaction terms in Table 6 suggest that students' family
+background moderates the relationship between their meritocratic
+perceptions in Chile and their justification of inequality. The
+direction of these effects confirms our initial prediction. Thus, in
+line with our Hypothesis 3, the relationship between social effort and
+justification of inequality becomes less positive for those students
+whose parents achieved university or postgraduate education. That is,
+lower-status students (measured by parental education) justify more
+inequality when they adhere more strongly to deservingnessmeritocratic
+perceptions in Chile. We depict this moderating effect in Figure 1. This
+result confirms the enlightenment thesis (see above). The same
+moderating effect is observed for students' family cultural capital: the
+relationship between effort (but not deservingness nor social talent)
+and justification of inequality becomes less positive for students with
+more than 25 books at home. Interestingly, family cultural capital also
+moderates the relationship between meritocratic beliefs at school (as
+measured by students' belief that effort is important to get good
+grades) and justification of inequality.
+
+At the school level, we also observe moderating effects of school status
+and students' meritocratic beliefs over their justification of
+inequality. Thus, in line with our Hypothesis 5, high-status schools
+justify less inequality when, on average, their students have a greater
+perception of meritocracy in Chile, as measured by the three indicators
+we used (i.e., effort, deservingness, and talent). Finally, as we stated
+in Hypothesis 7, lowhigh-achieving schools justify more inequality when
+their students have, on average, stronger meritocratic beliefs in Chile.
+In Figure 25 we depict this moderating effect of school achievement for
+the relationship between deservingness and justification of inequality.
+We did not observe these moderating effects at the school level for
+students' meritocratic beliefs in the school (i.e., the idea that effort
+or talent are important to get good grades).
+
+\subsection{Conclusion}\label{conclusion}
+
+\begin{itemize}
+\tightlist
+\item
+  Efecto ilustrador de la educación? (esfuerzo -\textgreater{}
+  redistribución)
+\end{itemize}
+
+\subsection{Appendix}\label{appendix}
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.4496}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5504}}@{}}
+\toprule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+\textbf{English translation} & \textbf{Original Spanish} \\
+It is just that in Chile people who cna pay have a better education for
+their children & Es justo que en Chile las personas que puedan pagar
+tengan una mejor educación para sus hijos. \\
+It is just that in Chile people with higher incomes can have better
+pensions than people with low incomes & Es justo que en Chile las
+personas con mayores ingresos puedan tener mejores pensiones que las
+personas de ingresos más bajos \\
+It is just that in Chile people with higher incomes can access better
+health services than people with low incomes & Es justo que en Chile las
+personas con mayores ingresos puedan acceder a una mejor atención de
+salud que las personas con ingresos más bajos \\
+\end{longtable}
+
+\phantomsection\label{refs}
+\begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-almas_cutthroat_2020}
+Almås, Ingvild, Alexander W. Cappelen, and Bertil Tungodden. 2020.
+{``Cutthroat Capitalism Versus Cuddly Socialism: Are Americans More
+Meritocratic and Efficiency-Seeking Than Scandinavians?''} \emph{Journal
+of Political Economy} 128 (5): 1753--88.
+\url{https://doi.org/10.1086/705551}.
+
+\bibitem[\citeproctext]{ref-barr_effect_2020}
+Barr, Abigail, and Luis Miller. 2020. {``The Effect of Education, Income
+Inequality and Merit on Inequality Acceptance.''} \emph{Journal of
+Economic Psychology} 80 (May): 102276.
+\url{https://doi.org/10.1016/j.joep.2020.102276}.
+
+\bibitem[\citeproctext]{ref-batruch_belief_2022}
+Batruch, Anatolia, Jolanda Jetten, Herman Van de Werfhorst, Céline
+Darnon, and Fabrizio Butera. 2022. {``Belief in School Meritocracy and
+the Legitimization of Social and Income Inequality.''} \emph{Social
+Psychological and Personality Science}, August, 194855062211110.
+\url{https://doi.org/10.1177/19485506221111017}.
+
+\bibitem[\citeproctext]{ref-bourdieu_reproduction_1990}
+Bourdieu, Pierre, and Jean Claude Passeron. 1990. \emph{Reproduction in
+Education, Society and Culture}. Second Edition. Sage Publications Ltd.
+
+\bibitem[\citeproctext]{ref-castillo_legitimacy_2011}
+Castillo, Juan Carlos. 2011. {``Legitimacy of Inequality in a Highly
+Unequal Context: Evidence from the Chilean Case.''} \emph{Social Justice
+Research} 24 (4): 314--40.
+\url{https://doi.org/10.1007/s11211-011-0144-5}.
+
+\bibitem[\citeproctext]{ref-castillo_meritocracia_2019}
+Castillo, Juan C., Alex Torres, Jorge Atria, and Luis Maldonado. 2019.
+{``Meritocracia y Desigualdad Económica: Percepciones, Preferencias e
+Implicancias.''} \emph{Revista Internacional de Sociología} 77 (1): 117.
+\url{https://doi.org/10.3989/ris.2019.77.1.17.114}.
+
+\bibitem[\citeproctext]{ref-chafel_schooling_1997}
+Chafel, Judith A. 1997. {``Schooling, the Hidden Curriculum, and
+Children's Conceptions of Poverty.''} \emph{Social Policy Report} 11
+(1): 1--28. \url{https://doi.org/10.1002/j.2379-3988.1997.tb00004.x}.
+
+\bibitem[\citeproctext]{ref-duru-bellat_who_2012}
+Duru-Bellat, Marie, and Elise Tenret. 2012. {``Who's for Meritocracy?
+Individual and Contextual Variations in the Faith.''} \emph{Comparative
+Education Review} 56 (2): 223--47. \url{https://doi.org/10.1086/661290}.
+
+\bibitem[\citeproctext]{ref-evans_strong_2018}
+Evans, M. D. R., and Jonathan Kelley. 2018. {``Strong Welfare States Do
+Not Intensify Public Support for Income Redistribution, but Even Reduce
+It Among the Prosperous: A Multilevel Analysis of Public Opinion in 30
+Countries.''} \emph{Societies} 8 (4): 105.
+\url{https://doi.org/10.3390/soc8040105}.
+
+\bibitem[\citeproctext]{ref-fernandez_positive_2013}
+Fernandez, J. J., and A. M. Jaime-Castillo. 2013. {``Positive or
+Negative Policy Feedbacks? Explaining Popular Attitudes Towards
+Pragmatic Pension Policy Reforms.''} \emph{European Sociological Review}
+29 (4): 803--15. \url{https://doi.org/10.1093/esr/jcs059}.
+
+\bibitem[\citeproctext]{ref-frank_performance_2015}
+Frank, Douglas H., Klaus Wertenbroch, and William W. Maddux. 2015.
+{``Performance Pay or Redistribution? Cultural Differences in Just-World
+Beliefs and Preferences for Wage Inequality.''} \emph{Organizational
+Behavior and Human Decision Processes} 130 (September): 160--70.
+\url{https://doi.org/10.1016/j.obhdp.2015.04.001}.
+
+\bibitem[\citeproctext]{ref-garcia-sanchez_attitudes_2020}
+García-Sánchez, Efraín, Danny Osborne, Guillermo B. Willis, and Rosa
+Rodríguez-Bailón. 2020. {``Attitudes Towards Redistribution and the
+Interplay Between Perceptions and Beliefs about Inequality.''}
+\emph{British Journal of Social Psychology} 59 (1): 111--36.
+\url{https://doi.org/10.1111/bjso.12326}.
+
+\bibitem[\citeproctext]{ref-goldthorpe_myth_2003}
+Goldthorpe, John. 2003. {``The Myth of Education-Based Meritocracy.''}
+\emph{New Economy} 10 (4): 234--39.
+\url{https://doi.org/10.1046/j.1468-0041.2003.00324.x}.
+
+\bibitem[\citeproctext]{ref-homan_being_2017}
+Homan, Patricia, Lauren Valentino, and Emi Weed. 2017. {``Being and
+Becoming Poor: How Cultural Schemas Shape Beliefs about Poverty.''}
+\emph{Social Forces} 95 (3): 1023--48.
+\url{https://doi.org/10.1093/sf/sox007}.
+
+\bibitem[\citeproctext]{ref-kluegel_beliefs_1987}
+Kluegel, James R., and Eliot R. Smith. 1987. \emph{Beliefs about
+Inequality: Americans' Views of What Is and What Ought to Be}. New York:
+Routledge.
+
+\bibitem[\citeproctext]{ref-lampert_meritocratic_2013}
+Lampert, Khen. 2013. \emph{Meritocratic Education and Social
+Worthlessness}. London: Palgrave Macmillan UK.
+\url{https://doi.org/10.1057/9781137324894}.
+
+\bibitem[\citeproctext]{ref-lindh_public_2015}
+Lindh, Arvid. 2015. {``Public Opinion Against Markets? Attitudes Towards
+Market Distribution of Social Services -- a Comparison of 17
+Countries.''} \emph{Social Policy \& Administration} 49 (7): 887--910.
+\url{https://doi.org/10.1111/spol.12105}.
+
+\bibitem[\citeproctext]{ref-mijs_stratified_2016}
+Mijs, Jonathan. 2016. {``Stratified Failure: Educational Stratification
+and Students' Attributions of Their Mathematics Performance in 24
+Countries.''} \emph{Sociology of Education} 89 (2): 137--53.
+\url{https://doi.org/10.1177/0038040716636434}.
+
+\bibitem[\citeproctext]{ref-mijs_paradox_2021}
+Mijs, Jonathan J B. 2021. {``The Paradox of Inequality: Income
+Inequality and Belief in Meritocracy Go Hand in Hand.''}
+\emph{Socio-Economic Review} 19 (1): 7--35.
+\url{https://doi.org/10.1093/ser/mwy051}.
+
+\bibitem[\citeproctext]{ref-resh_sense_2010}
+Resh, Nura. 2010. {``Sense of Justice about Grades in School: Is It
+Stratified Like Academic Achievement?''} \emph{Social Psychology of
+Education} 13 (3): 313--29.
+\url{https://doi.org/10.1007/s11218-010-9117-z}.
+
+\bibitem[\citeproctext]{ref-resh_sense_2017}
+Resh, Nura, and Clara Sabbagh. 2017. {``Sense of Justice in School and
+Civic Behavior.''} \emph{Social Psychology of Education} 20 (2):
+387--409. \url{https://doi.org/10.1007/s11218-017-9375-0}.
+
+\bibitem[\citeproctext]{ref-sandel_tyranny_2020}
+Sandel, Michael J. 2020. \emph{The Tyranny of Merit: What's Become of
+the Common Good?} First edition. New York: {Farrar, Straus and Giroux}.
+
+\bibitem[\citeproctext]{ref-vonhippel_test_2019}
+Von Hippel, Paul, and Caitlin Hamrock. 2019. {``Do Test Score Gaps Grow
+Before, During, or Between the School Years? Measurement Artifacts and
+What We Can Know in Spite of Them.''} \emph{Sociological Science} 6:
+43--80. \url{https://doi.org/10.15195/v6.a3}.
+
+\bibitem[\citeproctext]{ref-wegener_dominant_1995}
+Wegener, Bernd, and Stefan Liebig. 1995. {``Dominant Ideologies and the
+Variation of Distributive Justice Norms: A Comparison of East and West
+Germany, and the United States.''} In \emph{Social Justice and Political
+Change: Public Opinion in Capitalist and Post-Communist States}, edited
+by James R. Kluegel, David S. Mason, and Bernd Wegener, 239--59. New
+York: Walter de Gruyter.
+
+\bibitem[\citeproctext]{ref-wiederkehr_belief_2015}
+Wiederkehr, Virginie, Virginie Bonnot, Silvia Krauth-Gruber, and Céline
+Darnon. 2015. {``Belief in School Meritocracy as a System-Justifying
+Tool for Low Status Students.''} \emph{Frontiers in Psychology} 6.
+\url{https://www.frontiersin.org/articles/10.3389/fpsyg.2015.01053}.
+
+\bibitem[\citeproctext]{ref-young_rise_1958}
+Young, Michael. 1958. \emph{The Rise of the Meritocracy}. New Brunswick,
+N.J., U.S.A: Transaction Publishers.
+
+\end{CSLReferences}
 

### 2024-04-25 01-introduction.qmd
update references

@@
 In the present paper, we deal with the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system.
-In particular, we focus on market justice preferences (Lane, 1986), which refer to the preferences for distributing public goods (as health and education) based on criteria such as competition and payment capacity.
+In particular, we focus on market justice preferences [@lane_market_1986], which refer to the preferences for distributing public goods (as health and education) based on criteria such as competition and payment capacity.
 Although from a rational point of view it could be expected an opposition to market justice by the underprivileged majority,  we argue that in a social environment characterized by a promotion of meritocratic ideals - as the school system - would lead to the opposite: a larger market justice preferences.

### 2024-04-25 02-antecedentes.qmd
update references

@@
 The influence of ideologies [@wegener_dominant_1995] and cultural schemas [@homan_being_2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations.
-While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; ] (Castillo, 2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),  there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles.
-This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades (Centeno & Cohen, 2012; Harvey, 2007), affecting areas such as pensions, health services, and education.
-The justification of social inequality based on marketype criteria has been conceptualized as the individuals’ adherence to one specific justice evaluation, the market justice, that is, affording legitimacy to the allocation of goods and services based on prices and individuals’ ability to pay (Boltanski & Chiapello- 2007; Lane 1986; Streeck 2012).
+While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999 ; @shariff_income_2016],  there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles.
+This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades [@centeno_arc_2012; @harvey_breve_2015], affecting areas such as pensions, health services, and education.
+The justification of social inequality based on marketype criteria has been conceptualized as the individuals’ adherence to one specific justice evaluation, the market justice, that is, affording legitimacy to the allocation of goods and services based on prices and individuals’ ability to pay [@boltanski_new_2005; @lane_market_1986; @streeck_citizens_2012].
 
@@
 Lane proposed the underpinnings of market justice, which he differentiated from political justice.
-For him, “it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets” (1986, p.
-384).
+For him, “it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets” [-@lane_market_1986; p.
+384].
 Following the theory of relative deprivation –a social phenomenon arising when individuals cannot afford what most others in their environment can (Merton 1950)– Lane notes that, in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions.
-Although empirical research has shown that, contrary to Lane’s observation, relative deprivation, even in market settings, produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action such as protests and revolt (Greitemeyer and Sagioglou 2016; Mishra and Carleton 2015; Smith et al.
-2012; Power 2018; Power, Madsen, and Morton 2020), the conceptualization of market justice has been since then closely coupled to the merit principle for allocating outcomes: the claim that the unequal levels of well-being individuals enjoy ought to be, to some extent, a function of their talents and efforts, regardless of their needs or membership, the two latter being the realm of political justice and its closely coupled principles of need –allocating outcomes to those who require them most– and equality –allocating the same outcome to everyone–, which have been at the center of welfarism (see Wilson 2003).
+Although empirical research has shown that, contrary to Lane’s observation, relative deprivation, even in market settings, produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action such as protests and revolt [@greitemeyer_subjective_2016; @mishra_subjective_2015; @smith_relative_2012; @power_deprivationprotest_2018; @power_relative_2020], the conceptualization of market justice has been since then closely coupled to the merit principle for allocating outcomes: the claim that the unequal levels of well-being individuals enjoy ought to be, to some extent, a function of their talents and efforts, regardless of their needs or membership, the two latter being the realm of political justice and its closely coupled principles of need –allocating outcomes to those who require them most– and equality –allocating the same outcome to everyone–, which have been at the center of welfarism (see Wilson 2003).
 The underlying notion of market justice also resonates in its application to welfare regimes, providing a framework for understanding the varied global approaches to managing social services.
 
-The management of social services manifests in varying approaches across nations, with substantial differences in funding and delivery methods (Jensen, 2008; Stoy, 2014).
+The management of social services manifests in varying approaches across nations, with substantial differences in funding and delivery methods [@jensen_worlds_2008; @stoy_worlds_2014].
 Nordic countries, for example, predominantly employ public agencies to produce and provide social services, funding these through collective taxation and offering them in kind to the majority of citizens.
@@
 In contrast, other countries rely more heavily on for-profit entities and private funding, where service distribution depends mainly on individual financial capacity of paying user fees, highlighting the influence of market justice in service allocation.
-The trend toward marketization of welfare services has been growing since the 1980s (Salamon, 1993), and this shift is increasingly evident even in countries where market solutions have traditionally had a minor role in social policy (Sivesind, 2017).
+The trend toward marketization of welfare services has been growing since the 1980s [@salamon_marketization_1993], and this shift is increasingly evident even in countries where market solutions have traditionally had a minor role in social policy [@sivesind_changing_2017].
 
 The question arises whether adults and children justify unequal access to welfare services based on market justice principles.
-Influenced by theories of policy feedback, which suggest that social welfare policies can reinforce (positive feedback) or undermine (negative feedback) previous policy trajectories [@fernandez_positive_2013] (Fernandez and Jaime-Castillo 2013; Pierson 2000; Weaver 2010), citizens’ beliefs about market justice are likely also shaped by the institutional and social contexts they encounter.
+Influenced by theories of policy feedback, which suggest that social welfare policies can reinforce (positive feedback) or undermine (negative feedback) previous policy trajectories [@fernandez_positive_2013; @pierson_increasing_2000; @weaver_paths_2010], citizens’ beliefs about market justice are likely also shaped by the institutional and social contexts they encounter.
 Indeed, the justification of inequality in access to essential services like education and health, based on one’s ability to pay, shows significant variation across countries, as demonstrated by international surveys, although they do not usually include children in their samples.
@@
 @lindh_public_2015 analysis of ISSP 2009 data from 17 OECD countries reveals a general lack of support for market-based distribution of social services, suggesting widespread disapproval of market stratification of essential services.
-This finding is corroborated by Soler-Martínez and colleagues’ (2023) research from Latinobarómetro 2020 across 18 Latin American countries, where concerns about health and education access predominated over income inequality.
+This finding is corroborated by @soler-martinez_concerns_2023 research from Latinobarómetro 2020 across 18 Latin American countries, where concerns about health and education access predominated over income inequality.
 These results indicate that reforms toward welfare marketization are typically driven by elite political decisions rather than grassroots demand.
 
-Despite high-income inequality and limited social mobility in Latin America, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@buccaMeritBlameUnequal2016; @chongMysteryDiscriminationLatin2008; @torcheIntergenerationalMobilityInequality2014; @salgadoInequalityStratificationLatin2023].
-The reliance on private welfare providers and widespread user fees (Molyneux, 2008) adds complexity to this context.
+Despite high-income inequality and limited social mobility in Latin America, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@bucca_merit_2016; @chong_mystery_2008; @torche_intergenerational_2014; @salgado_inequality_2023].
+The reliance on private welfare providers and widespread user fees [@molyneux_neoliberal_2008] adds complexity to this context.
 Yet, research on children’s justification of market-based inequalities in accessing welfare services remains limited, especially in Latin America, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.
@@
 The label of beliefs in this realm covers a series of areas, such as attitudes, perceptions, and preferences (Castillo et al), whereby most of the link this subjective dimensions to individual socio-structural factors and  context-level determinants.
-For instance, some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2021a], and how larger inequality affects meritocratic beliefs (Morris et al 2022).
+For instance, some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2021], and how larger inequality affects meritocratic beliefs [@morris_representing_2022].
 Based on these findings, a research agenda has been reinforced on the legitimizing role of meritocracy, in line with previous studies using the concept of a just world (Lerner, Dalbert) and the theory of system justification (Jost & Major).
@@
 Still, this association was weaker among those who endorsed meritocratic and equal opportunity beliefs.
-In the same line, experimental research by Durante & Putterman (2009) subjects support less redistribution when the initial distribution is determined according to task performance.
+In the same line, experimental research by @durante_preferences_2009 subjects support less redistribution when the initial distribution is determined according to task performance.
 
-Research about meritocratic beliefs at school age is rather scarce, leaving a wide research gap as schools are one of the primary socialization institutions where achievement based on merit explains success (Erivwo et al 2021).
+Research about meritocratic beliefs at school age is rather scarce, leaving a wide research gap as schools are one of the primary socialization institutions where achievement based on merit explains success [@erivwo_meritocracy_2021].
 Nevertheless, it is possible to find some initial works focused on the area of distributive justice at school that are closely related to meritocratic beliefs, as the ones by Resh and Sabbagh (quotes).
-Using justice in grade obtention as a measure of distributive justice and meritocracy (Sabbagh et al 2006), they find for instance that a larger sense of distributive justice about grades is associated to higher socio-economic status [@resh_sense_2010], have a positive effect on liberal democratic orientation and on trust in people and in formal institutions (Resh & Sabbagh 2013), and tend to refrain from violence and to engage to a greater extent in extra-curricular school activity and community volunteering [@resh_sense_2017].
+Using justice in grade obtention as a measure of distributive justice and meritocracy (Sabbagh et al 2006), they find for instance that a larger sense of distributive justice about grades is associated to higher socio-economic status [@resh_sense_2010], have a positive effect on liberal democratic orientation and on trust in people and in formal institutions [@resh_sense_2014], and tend to refrain from violence and to engage to a greater extent in extra-curricular school activity and community volunteering [@resh_sense_2017].
 
@@
 Research indicates various socialization practices at families and schools during childhood and adolescence that impact dispositional and behavioral tendencies concerning justifications of inequality in adult life.
-The differences in economic understanding across age groups are consistent with research on cognitive development (Choudhury et al., 2006).
-Adolescents with mature socio-cognitive abilities tend to express stronger preferences for fairness than infants and children (Wynn et al., 2017).
-As children grow older, they become more likely to behave fairly, with their early-emerging strict egalitarianism being replaced by an increasing endorsement of fairness principles and engagement in collaborative activities (Huppert et al., 2019; McAuliffe et al., 2017).
-In these activities, their fairness views consider individual contributions, merits, and circumstances (Almås et al., 2010; Huppert et al., 2019; Sigelman & Waitzman, 1991).
-Engelmann and Tomasello (2019) claim that children’s sense of fairness emerges at three years old, and we can observe it in collaborative activities, where they accept inequality if the procedure gives everyone an equal chance.
+The differences in economic understanding across age groups are consistent with research on cognitive development [@choudhury_social_2006].
+Adolescents with mature socio-cognitive abilities tend to express stronger preferences for fairness than infants and children [@wynn_not_2018].
+As children grow older, they become more likely to behave fairly, with their early-emerging strict egalitarianism being replaced by an increasing endorsement of fairness principles and engagement in collaborative activities [@huppert_development_2019; @mcauliffe_developmental_2017].
+In these activities, their fairness views consider individual contributions, merits, and circumstances [@almas_fairness_2010; @huppert_development_2019; @sigelman_development_1991].
+@engelmann_children_2019 claim that children’s sense of fairness emerges at three years old, and we can observe it in collaborative activities, where they accept inequality if the procedure gives everyone an equal chance.
 Therefore, children at this age respond to unequal distributions based on interpersonal concerns, as they already demand equal respect.
-In any case, between 3 and 8 years of age, inequitable and anti-meritorious allocations are evaluated more negatively, but equitable and meritorious allocations are not evaluated more positively (Elenbaas, 2019).
+In any case, between 3 and 8 years of age, inequitable and anti-meritorious allocations are evaluated more negatively, but equitable and meritorious allocations are not evaluated more positively [@elenbaas_unfairness_2019].
 
-Some research shows that the social environment in which children develop, such as family and school, is associated with their prosocial behaviors by playing an essential role in the transmission of equity norms (Schunk & Zipperle, 2023; Kosse et al., 2020).
-In fact, schools contribute to institutionalizing and reproducing inequality by promoting values, norms, practices, and languages familiar to higher-class families because the dominant group’s culture shapes educational institutions (Bourdieu & Passeron, 1990).
-Middle- and upper-class students are better equipped to face academic challenges and are more familiar with academic expectations (Mikus et al., 2019).
-Such familiarity represents cultural capital in educational contexts because higher-status students come to school ready to meet these expectations and reap the benefits (Jack, 2016; Khan, 2011).
-Conversely, lower-status children lacking cultural capital must catch up while experiencing inequitable comparisons (Goudeau & Croizet, 2017).
+Some research shows that the social environment in which children develop, such as family and school, is associated with their prosocial behaviors by playing an essential role in the transmission of equity norms [@schunk_fairness_2023; @kosse_prosociality_2020].
+In fact, schools contribute to institutionalizing and reproducing inequality by promoting values, norms, practices, and languages familiar to higher-class families because the dominant group’s culture shapes educational institutions [@bourdieu_reproduction_1990].
+Middle- and upper-class students are better equipped to face academic challenges and are more familiar with academic expectations [@mikus_children_2020].
+Such familiarity represents cultural capital in educational contexts because higher-status students come to school ready to meet these expectations and reap the benefits [@jack_no_2016; @khan_privilege_2011].
+Conversely, lower-status children lacking cultural capital must catch up while experiencing inequitable comparisons [@goudeau_hidden_2017].
 Additionally, academic achievement is treated as the outcome of dispositional factors (e.g., pupils’ efforts and talents or lack of them) rather than the result of differential access to critical resources.
-Due to the meritocratic frame schools encourage, both low- and high-status individuals believe that students’ success or failure is not due to their family background but rather to differences in efforts and talents (Darnon et al., 2018).
+Due to the meritocratic frame schools encourage, both low- and high-status individuals believe that students’ success or failure is not due to their family background but rather to differences in efforts and talents [@darnon_where_2018].
 In this sense, we believe that the perception of meritocracy can influence students' judgments about market justice preferences.
@@
 Family background and family socialization practices also contribute to children’s and adolescent’s market justice preferences.
-For example, Almås and colleagues (2017) found that adolescents from low-socioeconomic-status families are likelier to have an egalitarian fairness view and consider an equal distribution as fair in a situation with unequal merits.
-The authors speculate that differences in socialization practices across status groups might bring about, to a great extent, the fairness views of children and adolescents because social status seems to interact with these evaluations (e.g., Hvidberg et al., 2023).
+For example, @almas_fairness_2017 found that adolescents from low-socioeconomic-status families are likelier to have an egalitarian fairness view and consider an equal distribution as fair in a situation with unequal merits.
+The authors speculate that differences in socialization practices across status groups might bring about, to a great extent, the fairness views of children and adolescents because social status seems to interact with these evaluations [e.g., @hvidberg_social_2023].
 
-The classic work of Kohn showed that middle-class parents value the expression of internal states and emotions, such as self-control, curiosity, happiness, and consideration, while working-class parents promote deference, obedience, and conformity to authority (Kohn, 1963; Kohn & Schooler, 1969).
+The classic work of Kohn showed that middle-class parents value the expression of internal states and emotions, such as self-control, curiosity, happiness, and consideration, while working-class parents promote deference, obedience, and conformity to authority [@kohn_social_1963; @kohn_class_1969].
 Although parents from all social backgrounds encourage individualism in their children, this shared norm translates into different forms in high and low social classes (1999).
-Acemoglu (2021) claimed that the values families impart to their children interact with social mobility.
+@acemoglu_obedience_2021 claimed that the values families impart to their children interact with social mobility.
 Because obedience is a valuable characteristic for employers, in low-wage and social mobility environments, low-income families impart values of obedience to their children to prevent disadvantaging them in labor markets.
 On the one hand, children from privileged families are socialized to adopt a clear conception of individualism that highlights their internal states, independence, and idiosyncrasies.
-In contrast, children from disadvantaged families are socialized to support a more balanced view of individualism that considers personal characteristics as resources to overcome collective impediments on the path to upward mobility (Iacoviello & Lorenzi-Cioldi, 2019).
+In contrast, children from disadvantaged families are socialized to support a more balanced view of individualism that considers personal characteristics as resources to overcome collective impediments on the path to upward mobility [@iacoviello_collectivism_2019].
 In this way, we believe that there are differences in the socialization of values according to socioeconomic differences that could influence market justice preferences and, therefore, we propose the following hypothesis:
@@
 Recent empirical research has demonstrated that the institutional design of schools, coupled with the meritocratic ideology it fosters, significantly influences children’s and adolescents’ views on inequality and deservingness.
-For example, Jonsson and Beach’s (2015) study revealed that higher-status adolescents in Sweden tend to perpetuate social class stereotypes while describing the vocational and academic tracks.
+For example, @jonsson_institutional_2015 study revealed that higher-status adolescents in Sweden tend to perpetuate social class stereotypes while describing the vocational and academic tracks.
 Academic track students are depicted as wealthy, intelligent, ambitious, and diligent, while vocational track students are characterized as poor, unambitious, unintelligent, and lackadaisical.
-These stereotypes help individuals maintain a sense of superiority over others and legitimize the prevailing social hierarchies and economic disparities (Jost & Burgess, 2000).
+These stereotypes help individuals maintain a sense of superiority over others and legitimize the prevailing social hierarchies and economic disparities [@jost_attitudinal_2000]
 

### 2024-04-25 paper.tex
update references

@@
 \makeatother
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
   \fi
   % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 \KOMAoption{captions}{tableheading}
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
 \makeatletter
@@
 \makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
 \makeatother
 \makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
 \makeatother
@@
 \fi
-\usepackage{bookmark}
-
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \begin{document}
-
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+
+\hypertarget{the-socialization-of-meritocracy-and-market-justice-preferences-at-school}{%
 \section{The socialization of meritocracy and market justice preferences
 at
-school}\label{the-socialization-of-meritocracy-and-market-justice-preferences-at-school}
-
-\subsection{Introduction}\label{introduction}
+school}\label{the-socialization-of-meritocracy-and-market-justice-preferences-at-school}}
+
+\hypertarget{introduction}{%
+\subsection{Introduction}\label{introduction}}
 
@@
 school level represents a threat to the promise of education and a
-meritocratic system (\citeproc{ref-bourdieu_reproduction_1990}{Bourdieu
-and Passeron 1990}). A large part of the research in this field at an
+meritocratic system
+(\protect\hyperlink{ref-bourdieu_reproduction_1990}{Bourdieu and
+Passeron 1990}). A large part of the research in this field at an
 international level has addressed the extent to which the social origin
 of students affects their academic results and their life opportunities
-(\citeproc{ref-vonhippel_test_2019}{Von Hippel and Hamrock 2019}),
-confirming that schools have severe difficulties in closing the gaps of
-origin. Besides this socioeconomic perspective on school opportunities,
-recent research has addressed to what extent inequalities in the school
-context are also influencing students' perceptions, beliefs, and
-attitudes: Are social inequalities perceived at the school context? Are
-they rejected by the students, particularly those who are worst-off in
-socioeconomic terms? Or, Is there evidence at the school level that
-social inequalities are tolerated and even justified?
-(\citeproc{ref-batruch_belief_2022}{Batruch et al. 2022};
-\citeproc{ref-wiederkehr_belief_2015}{Wiederkehr et al. 2015}).~
+(\protect\hyperlink{ref-vonhippel_test_2019}{Von Hippel and Hamrock
+2019}), confirming that schools have severe difficulties in closing the
+gaps of origin. Besides this socioeconomic perspective on school
+opportunities, recent research has addressed to what extent inequalities
+in the school context are also influencing students' perceptions,
+beliefs, and attitudes: Are social inequalities perceived at the school
+context? Are they rejected by the students, particularly those who are
+worst-off in socioeconomic terms? Or, Is there evidence at the school
+level that social inequalities are tolerated and even justified?
+(\protect\hyperlink{ref-batruch_belief_2022}{Batruch et al. 2022};
+\protect\hyperlink{ref-wiederkehr_belief_2015}{Wiederkehr et al.
+2015}).~
 
@@
 by a highly stratified educational system. In particular, we focus on
-market justice preferences (Lane, 1986), which refer to the preferences
-for distributing public goods (as health and education) based on
-criteria such as competition and payment capacity. Although from a
-rational point of view it could be expected an opposition to market
-justice by the underprivileged majority,~ we argue that in a social
-environment characterized by a promotion of meritocratic ideals - as the
-school system - would lead to the opposite: a larger market justice
-preferences.~
+market justice preferences
+(\protect\hyperlink{ref-lane_market_1986}{Lane 1986}), which refer to
+the preferences for distributing public goods (as health and education)
+based on criteria such as competition and payment capacity. Although
+from a rational point of view it could be expected an opposition to
+market justice by the underprivileged majority,~ we argue that in a
+social environment characterized by a promotion of meritocratic ideals -
+as the school system - would lead to the opposite: a larger market
+justice preferences.~
 
@@
 based on their abilities, knowledge, and achievements
-(\citeproc{ref-young_rise_1958}{Young 1958}). It is often seen as a way
-to create equal opportunities and fairness, as individuals can rise to
-positions of power and influence based on their own merit rather than
-their background or connections. However, some argue that meritocracy
-can actually lead to tolerating or even justifying social inequalities,
-as it can create a hierarchy where those who already have resources and
-advantages are more likely to succeed. In this regard, a great deal of
-academic research about meritocracy delves into the assessment of to
-what extent rewards and privileges in society are related to merit,
-emphasizing the so-called unfulfillable promise of meritocracy
-(\citeproc{ref-mijs_stratified_2016}{J. Mijs 2016}). Complementing this
-agenda, a second and emerging research area deals with subjective
-aspects of meritocracy, such as perceptions and beliefs.~
+(\protect\hyperlink{ref-young_rise_1958}{Young 1958}). It is often seen
+as a way to create equal opportunities and fairness, as individuals can
+rise to positions of power and influence based on their own merit rather
+than their background or connections. However, some argue that
+meritocracy can actually lead to tolerating or even justifying social
+inequalities, as it can create a hierarchy where those who already have
+resources and advantages are more likely to succeed. In this regard, a
+great deal of academic research about meritocracy delves into the
+assessment of to what extent rewards and privileges in society are
+related to merit, emphasizing the so-called unfulfillable promise of
+meritocracy (\protect\hyperlink{ref-mijs_stratified_2016}{Mijs 2016}).
+Complementing this agenda, a second and emerging research area deals
+with subjective aspects of meritocracy, such as perceptions and
+beliefs.~
 
@@
 understand the concept of meritocracy in their own society
-(\citeproc{ref-duru-bellat_who_2012}{Duru-Bellat and Tenret 2012};
-\citeproc{ref-castillo_meritocracia_2019}{Juan C. Castillo et al.
-2019}). This perception can vary greatly depending on individual
+(\protect\hyperlink{ref-duru-bellat_who_2012}{Duru-Bellat and Tenret
+2012}; \protect\hyperlink{ref-castillo_meritocracia_2019}{Castillo et
+al. 2019}). This perception can vary greatly depending on individual
 experiences, social, economic, and cultural background. Some people may
@@
 to maintain and even reinforce inequality
-(\citeproc{ref-lampert_meritocratic_2013}{Lampert 2013};
-\citeproc{ref-mijs_paradox_2021}{J. J. B. Mijs 2021}). Based in this
+(\protect\hyperlink{ref-lampert_meritocratic_2013}{Lampert 2013};
+\protect\hyperlink{ref-mijs_paradox_2021}{Mijs 2021}). Based in this
 last perspective, we argue that individuals with a higher perception of
@@
 individual achievement would be seen as rewarded and social policies as
-less necessary (\citeproc{ref-batruch_belief_2022}{Batruch et al.
-2022}).
+less necessary (\protect\hyperlink{ref-batruch_belief_2022}{Batruch et
+al. 2022}).
 
@@
 conditions of origin has been linked to the \emph{hidden curriculum}
-(\citeproc{ref-chafel_schooling_1997}{Chafel 1997}), whereby students
-learn about distributive norms in society and mechanisms of
+(\protect\hyperlink{ref-chafel_schooling_1997}{Chafel 1997}), whereby
+students learn about distributive norms in society and mechanisms of
 justification of social differences. Based on recent studies that relate
 school meritocracy to the justification of economic inequalities in the
-adult population (\citeproc{ref-batruch_belief_2022}{Batruch et al.
-2022}; \citeproc{ref-wiederkehr_belief_2015}{Wiederkehr et al. 2015}),
-the central hypothesis guiding this research is that school-age students
-with a higher perception of meritocracy - both at school and at the
-societal level - will show a larger justification of social
-inequalities, as individual achievement would be seen as appropriately
-rewarded and social mechanisms for correcting inequalities as less
-necessary (\citeproc{ref-batruch_belief_2022}{Batruch et al. 2022}).
+adult population (\protect\hyperlink{ref-batruch_belief_2022}{Batruch et
+al. 2022}; \protect\hyperlink{ref-wiederkehr_belief_2015}{Wiederkehr et
+al. 2015}), the central hypothesis guiding this research is that
+school-age students with a higher perception of meritocracy - both at
+school and at the societal level - will show a larger justification of
+social inequalities, as individual achievement would be seen as
+appropriately rewarded and social mechanisms for correcting inequalities
+as less necessary (\protect\hyperlink{ref-batruch_belief_2022}{Batruch
+et al. 2022}).
 
@@
 
+\hypertarget{justification-of-inequality}{%
 \subsection{Justification of
-inequality}\label{justification-of-inequality}
+inequality}\label{justification-of-inequality}}
 
@@
 perceptions of who deserves what and why
-(\citeproc{ref-kluegel_beliefs_1987}{Kluegel and Smith 1987}),
+(\protect\hyperlink{ref-kluegel_beliefs_1987}{Kluegel and Smith 1987}),
 highlights that people's explanations and justifications of social
 inequality are closely tied to their judgments of deservingness. The
-influence of ideologies (\citeproc{ref-wegener_dominant_1995}{Wegener
-and Liebig 1995}) and cultural schemas
-(\citeproc{ref-homan_being_2017}{Homan, Valentino, and Weed 2017}) is
-pivotal in shaping these explanations by offering symbolic
-representations that frame societal structures and expectations. While
-significant attention has been paid to wage inequality, income
-distribution, and payment differentials in the literature {[}Juan Carlos
-Castillo (\citeproc{ref-castillo_legitimacy_2011}{2011}); {]} (Castillo,
-2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),~ there has
-been less examination of public beliefs about which life domains should
-be governed by market relations and even less about children's
-acceptance or rejection of these market principles. This oversight is
-notable given the extensive encroachment of market logic into public
-goods, welfare policy, and social services over the past five decades
-(Centeno \& Cohen, 2012; Harvey, 2007), affecting areas such as
-pensions, health services, and education. The justification of social
-inequality based on marketype criteria has been conceptualized as the
-individuals' adherence to one specific justice evaluation, the market
-justice, that is, affording legitimacy to the allocation of goods and
-services based on prices and individuals' ability to pay (Boltanski \&
-Chiapello- 2007; Lane 1986; Streeck 2012).
+influence of ideologies
+(\protect\hyperlink{ref-wegener_dominant_1995}{Wegener and Liebig 1995})
+and cultural schemas (\protect\hyperlink{ref-homan_being_2017}{Homan,
+Valentino, and Weed 2017}) is pivotal in shaping these explanations by
+offering symbolic representations that frame societal structures and
+expectations. While significant attention has been paid to wage
+inequality, income distribution, and payment differentials in the
+literature {[}Castillo
+(\protect\hyperlink{ref-castillo_legitimacy_2011}{2011}); Evans et al.,
+2010; Jasso (\protect\hyperlink{ref-jasso_how_1999}{1999}) ; Shariff,
+Wiwad, and Aknin
+(\protect\hyperlink{ref-shariff_income_2016}{2016}){]},~ there has been
+less examination of public beliefs about which life domains should be
+governed by market relations and even less about children's acceptance
+or rejection of these market principles. This oversight is notable given
+the extensive encroachment of market logic into public goods, welfare
+policy, and social services over the past five decades
+(\protect\hyperlink{ref-centeno_arc_2012}{Centeno and Cohen 2012};
+\protect\hyperlink{ref-harvey_breve_2015}{Harvey 2015}), affecting areas
+such as pensions, health services, and education. The justification of
+social inequality based on marketype criteria has been conceptualized as
+the individuals' adherence to one specific justice evaluation, the
+market justice, that is, affording legitimacy to the allocation of goods
+and services based on prices and individuals' ability to pay
+(\protect\hyperlink{ref-boltanski_new_2005}{Boltanski and Chiapello
+2005}; \protect\hyperlink{ref-lane_market_1986}{Lane 1986};
+\protect\hyperlink{ref-streeck_citizens_2012}{Streeck 2012}).
 
@@
 the market to stimulate wants without at the same time stimulating a
-sense of deserving more than one gets'' (1986, p.~384). Following the
-theory of relative deprivation --a social phenomenon arising when
+sense of deserving more than one gets'' {[}-Lane
+(\protect\hyperlink{ref-lane_market_1986}{1986}); p.~384{]}. Following
+the theory of relative deprivation --a social phenomenon arising when
 individuals cannot afford what most others in their environment can
@@
 feelings of dissatisfaction, anger, and resentment that might motivate
-forms of collective action such as protests and revolt (Greitemeyer and
-Sagioglou 2016; Mishra and Carleton 2015; Smith et al.~2012; Power 2018;
-Power, Madsen, and Morton 2020), the conceptualization of market justice
-has been since then closely coupled to the merit principle for
+forms of collective action such as protests and revolt
+(\protect\hyperlink{ref-greitemeyer_subjective_2016}{Greitemeyer and
+Sagioglou 2016}; \protect\hyperlink{ref-mishra_subjective_2015}{Mishra
+and Carleton 2015}; \protect\hyperlink{ref-smith_relative_2012}{Smith et
+al. 2012}; \protect\hyperlink{ref-power_deprivationprotest_2018}{Séamus
+A. Power 2018}; \protect\hyperlink{ref-power_relative_2020}{Séamus A.
+Power, Madsen, and Morton 2020}), the conceptualization of market
+justice has been since then closely coupled to the merit principle for
 allocating outcomes: the claim that the unequal levels of well-being
@@
 nations, with substantial differences in funding and delivery methods
-(Jensen, 2008; Stoy, 2014). Nordic countries, for example, predominantly
-employ public agencies to produce and provide social services, funding
-these through collective taxation and offering them in kind to the
-majority of citizens. This system prioritizes political justice, placing
-it above market mechanisms in accessing services. In contrast, other
-countries rely more heavily on for-profit entities and private funding,
-where service distribution depends mainly on individual financial
-capacity of paying user fees, highlighting the influence of market
-justice in service allocation. The trend toward marketization of welfare
-services has been growing since the 1980s (Salamon, 1993), and this
-shift is increasingly evident even in countries where market solutions
-have traditionally had a minor role in social policy (Sivesind, 2017).
+(\protect\hyperlink{ref-jensen_worlds_2008}{Jensen 2008};
+\protect\hyperlink{ref-stoy_worlds_2014}{Stoy 2014}). Nordic countries,
+for example, predominantly employ public agencies to produce and provide
+social services, funding these through collective taxation and offering
+them in kind to the majority of citizens. This system prioritizes
+political justice, placing it above market mechanisms in accessing
+services. In contrast, other countries rely more heavily on for-profit
+entities and private funding, where service distribution depends mainly
+on individual financial capacity of paying user fees, highlighting the
+influence of market justice in service allocation. The trend toward
+marketization of welfare services has been growing since the 1980s
+(\protect\hyperlink{ref-salamon_marketization_1993}{Salamon 1993}), and
+this shift is increasingly evident even in countries where market
+solutions have traditionally had a minor role in social policy
+(\protect\hyperlink{ref-sivesind_changing_2017}{Sivesind 2017}).
 
@@
 previous policy trajectories
-(\citeproc{ref-fernandez_positive_2013}{Fernandez and Jaime-Castillo
-2013}) (Fernandez and Jaime-Castillo 2013; Pierson 2000; Weaver 2010),
-citizens' beliefs about market justice are likely also shaped by the
-institutional and social contexts they encounter. Indeed, the
-justification of inequality in access to essential services like
-education and health, based on one's ability to pay, shows significant
-variation across countries, as demonstrated by international surveys,
-although they do not usually include children in their samples. For
-instance, the 2019 International Social Survey Program (ISSP) provides
-insights into how adults perceive inequality in accessing welfare
-services. Figure 1 illustrates the variation in agreement levels
-regarding whether it is just for individuals with higher incomes to
-purchase better education or healthcare. Notably, in 18 of the 29
-countries analyzed, there is a greater justification for market
-inequality in healthcare access than in education. Nevertheless, the
-general sentiment typically ranges from `somewhat unjust' to 'neither
-just nor unjust, with mixed feelings.
+(\protect\hyperlink{ref-fernandez_positive_2013}{Fernandez and
+Jaime-Castillo 2013};
+\protect\hyperlink{ref-pierson_increasing_2000}{Pierson 2000};
+\protect\hyperlink{ref-weaver_paths_2010}{Weaver 2010}), citizens'
+beliefs about market justice are likely also shaped by the institutional
+and social contexts they encounter. Indeed, the justification of
+inequality in access to essential services like education and health,
+based on one's ability to pay, shows significant variation across
+countries, as demonstrated by international surveys, although they do
+not usually include children in their samples. For instance, the 2019
+International Social Survey Program (ISSP) provides insights into how
+adults perceive inequality in accessing welfare services. Figure 1
+illustrates the variation in agreement levels regarding whether it is
+just for individuals with higher incomes to purchase better education or
+healthcare. Notably, in 18 of the 29 countries analyzed, there is a
+greater justification for market inequality in healthcare access than in
+education. Nevertheless, the general sentiment typically ranges from
+`somewhat unjust' to 'neither just nor unjust, with mixed feelings.
 
@@
 and health services, questions about public acceptance of these changes
-persist. Lindh (\citeproc{ref-lindh_public_2015}{2015}) analysis of ISSP
-2009 data from 17 OECD countries reveals a general lack of support for
-market-based distribution of social services, suggesting widespread
-disapproval of market stratification of essential services. This finding
-is corroborated by Soler-Martínez and colleagues' (2023) research from
-Latinobarómetro 2020 across 18 Latin American countries, where concerns
-about health and education access predominated over income inequality.
-These results indicate that reforms toward welfare marketization are
-typically driven by elite political decisions rather than grassroots
-demand.~
+persist. Lindh (\protect\hyperlink{ref-lindh_public_2015}{2015})
+analysis of ISSP 2009 data from 17 OECD countries reveals a general lack
+of support for market-based distribution of social services, suggesting
+widespread disapproval of market stratification of essential services.
+This finding is corroborated by Soler-Martínez, García-Sánchez, and
+Willis (\protect\hyperlink{ref-soler-martinez_concerns_2023}{2023})
+research from Latinobarómetro 2020 across 18 Latin American countries,
+where concerns about health and education access predominated over
+income inequality. These results indicate that reforms toward welfare
+marketization are typically driven by elite political decisions rather
+than grassroots demand.~
 
@@
 responsible for their economic outcomes, a view that varies across the
-region
-(\citeproc{ref-buccaMeritBlameUnequal2016}{\textbf{buccaMeritBlameUnequal2016?}};
-\citeproc{ref-chongMysteryDiscriminationLatin2008}{\textbf{chongMysteryDiscriminationLatin2008?}};
-\citeproc{ref-torcheIntergenerationalMobilityInequality2014}{\textbf{torcheIntergenerationalMobilityInequality2014?}};
-\citeproc{ref-salgadoInequalityStratificationLatin2023}{\textbf{salgadoInequalityStratificationLatin2023?}}).
-The reliance on private welfare providers and widespread user fees
-(Molyneux, 2008) adds complexity to this context. Yet, research on
-children's justification of market-based inequalities in accessing
-welfare services remains limited, especially in Latin America,
-highlighting a significant gap in understanding how younger generations
-view market-based access to welfare and whether these views are
-associated with their meritocratic beliefs.~
-
-\subsection{Meritocracy}\label{meritocracy}
+region (\protect\hyperlink{ref-bucca_merit_2016}{Bucca 2016};
+\protect\hyperlink{ref-chong_mystery_2008}{Chong and Ñopo 2008};
+\protect\hyperlink{ref-torche_intergenerational_2014}{Torche 2014};
+\protect\hyperlink{ref-salgado_inequality_2023}{Salgado and Castillo
+2023}). The reliance on private welfare providers and widespread user
+fees (\protect\hyperlink{ref-molyneux_neoliberal_2008}{Molyneux 2008})
+adds complexity to this context. Yet, research on children's
+justification of market-based inequalities in accessing welfare services
+remains limited, especially in Latin America, highlighting a significant
+gap in understanding how younger generations view market-based access to
+welfare and whether these views are associated with their meritocratic
+beliefs.~
+
+\hypertarget{meritocracy}{%
+\subsection{Meritocracy}\label{meritocracy}}
 
@@
 quite far from its original formulation by Young
-(\citeproc{ref-young_rise_1958}{1958}) in the satirical novel ``The Rise
-of Meritocracy'', where it ironically represented a mechanism for
-reproducing the inequalities of origin. The meritocratic ideal had
+(\protect\hyperlink{ref-young_rise_1958}{1958}) in the satirical novel
+``The Rise of Meritocracy'', where it ironically represented a mechanism
+for reproducing the inequalities of origin. The meritocratic ideal had
 remained relatively unchallenged until a series of recent publications
@@
 signified a valued contribution to the common good.''
-(\citeproc{ref-sandel_tyranny_2020}{Sandel 2020}, pp.).
+(\protect\hyperlink{ref-sandel_tyranny_2020}{Sandel 2020}, pp.).
 
@@
 mobility that threaten the meritocratic ideal
-(\citeproc{ref-goldthorpe_myth_2003}{Goldthorpe 2003}). More recently,
-sociology and social psychology research has attended to the subjective
-aspects vis-a-vis beliefs in meritocracy. The label of beliefs in this
-realm covers a series of areas, such as attitudes, perceptions, and
-preferences (Castillo et al), whereby most of the link this subjective
-dimensions to individual socio-structural factors and~ context-level
-determinants. For instance, some studies have analyzed how those with
-greater privileges believe more in meritocracy (Reynolds \& Chan 2014),
-how greater economic inequality increases meritocratic beliefs
-(\citeproc{ref-mijs_paradox_2021a}{\textbf{mijs\_paradox\_2021a?}}), and
-how larger inequality affects meritocratic beliefs (Morris et al 2022).
+(\protect\hyperlink{ref-goldthorpe_myth_2003}{Goldthorpe 2003}). More
+recently, sociology and social psychology research has attended to the
+subjective aspects vis-a-vis beliefs in meritocracy. The label of
+beliefs in this realm covers a series of areas, such as attitudes,
+perceptions, and preferences (Castillo et al), whereby most of the link
+this subjective dimensions to individual socio-structural factors and~
+context-level determinants. For instance, some studies have analyzed how
+those with greater privileges believe more in meritocracy (Reynolds \&
+Chan 2014), how greater economic inequality increases meritocratic
+beliefs (\protect\hyperlink{ref-mijs_paradox_2021}{Mijs 2021}), and how
+larger inequality affects meritocratic beliefs
+(\protect\hyperlink{ref-morris_representing_2022}{Morris et al. 2022}).
 Based on these findings, a research agenda has been reinforced on the
@@
 negatively with support for redistributive compensation systems
-(\citeproc{ref-frank_performance_2015}{Frank, Wertenbroch, and Maddux
-2015}). Conversely, individuals tend to support redistribution when they
-believe that the disadvantaged lack the opportunities to succeed
-(\citeproc{ref-evans_strong_2018}{Evans and Kelley 2018}). Almås,
-Cappelen, and Tungodden (\citeproc{ref-almas_cutthroat_2020}{2020})
-found that in a relatively unequal society (the United States), the
-highly educated accept inequality significantly more than the less
-educated because they perceive inequality as justifiable owing to
-differences in productivity (i.e., merit), whereas in a relatively equal
-society (Norway), the less educated accept inequality more, but not
-significantly more than the highly educated because meritocratic values
-are less prevalent. Barr and Miller
-(\citeproc{ref-barr_effect_2020}{2020}) also addressed this triple
-interaction between the level of inequality in a society, the individual
-level of education, and the perceived origin of the disparity (either by
-luck or effort) to determine the extent to which inequality is accepted.
-They found that the interaction's mechanism varies depending on the
-compared societies. Finally, García-Sánchez et al.
-(\citeproc{ref-garcia-sanchez_attitudes_2020}{2020}), using data for 41
-countries from the International Social Survey Programme (ISSP), found
-that the perceived size of the income gap correlated positively with
-support for progressive taxation. Still, this association was weaker
-among those who endorsed meritocratic and equal opportunity beliefs. In
-the same line, experimental research by Durante \& Putterman (2009)
+(\protect\hyperlink{ref-frank_performance_2015}{Frank, Wertenbroch, and
+Maddux 2015}). Conversely, individuals tend to support redistribution
+when they believe that the disadvantaged lack the opportunities to
+succeed (\protect\hyperlink{ref-evans_strong_2018}{Evans and Kelley
+2018}). Almås, Cappelen, and Tungodden
+(\protect\hyperlink{ref-almas_cutthroat_2020}{2020}) found that in a
+relatively unequal society (the United States), the highly educated
+accept inequality significantly more than the less educated because they
+perceive inequality as justifiable owing to differences in productivity
+(i.e., merit), whereas in a relatively equal society (Norway), the less
+educated accept inequality more, but not significantly more than the
+highly educated because meritocratic values are less prevalent. Barr and
+Miller (\protect\hyperlink{ref-barr_effect_2020}{2020}) also addressed
+this triple interaction between the level of inequality in a society,
+the individual level of education, and the perceived origin of the
+disparity (either by luck or effort) to determine the extent to which
+inequality is accepted. They found that the interaction's mechanism
+varies depending on the compared societies. Finally, García-Sánchez et
+al. (\protect\hyperlink{ref-garcia-sanchez_attitudes_2020}{2020}), using
+data for 41 countries from the International Social Survey Programme
+(ISSP), found that the perceived size of the income gap correlated
+positively with support for progressive taxation. Still, this
+association was weaker among those who endorsed meritocratic and equal
+opportunity beliefs. In the same line, experimental research by Durante
+and Putterman (\protect\hyperlink{ref-durante_preferences_2009}{2009})
 subjects support less redistribution when the initial distribution is
@@
 socialization institutions where achievement based on merit explains
-success (Erivwo et al 2021). Nevertheless, it is possible to find some
-initial works focused on the area of distributive justice at school that
-are closely related to meritocratic beliefs, as the ones by Resh and
-Sabbagh (quotes). Using justice in grade obtention as a measure of
-distributive justice and meritocracy (Sabbagh et al 2006), they find for
-instance that a larger sense of distributive justice about grades is
-associated to higher socio-economic status
-(\citeproc{ref-resh_sense_2010}{Resh 2010}), have a positive effect on
-liberal democratic orientation and on trust in people and in formal
-institutions (Resh \& Sabbagh 2013), and tend to refrain from violence
-and to engage to a greater extent in extra-curricular school activity
-and community volunteering (\citeproc{ref-resh_sense_2017}{Resh and
-Sabbagh 2017}).
-
+success (\protect\hyperlink{ref-erivwo_meritocracy_2021}{Erivwo et al.
+2021}). Nevertheless, it is possible to find some initial works focused
+on the area of distributive justice at school that are closely related
+to meritocratic beliefs, as the ones by Resh and Sabbagh (quotes). Using
+justice in grade obtention as a measure of distributive justice and
+meritocracy (Sabbagh et al 2006), they find for instance that a larger
+sense of distributive justice about grades is associated to higher
+socio-economic status (\protect\hyperlink{ref-resh_sense_2010}{Resh
+2010}), have a positive effect on liberal democratic orientation and on
+trust in people and in formal institutions
+(\protect\hyperlink{ref-resh_sense_2014}{Resh and Sabbagh 2014}), and
+tend to refrain from violence and to engage to a greater extent in
+extra-curricular school activity and community volunteering
+(\protect\hyperlink{ref-resh_sense_2017}{Resh and Sabbagh 2017}).
+
+\hypertarget{childrens-judgments-of-inequality-schools-and-family-background}{%
 \subsection{Children's judgments of inequality, schools, and family
-background}\label{childrens-judgments-of-inequality-schools-and-family-background}
+background}\label{childrens-judgments-of-inequality-schools-and-family-background}}
 
@@
 life. The differences in economic understanding across age groups are
-consistent with research on cognitive development (Choudhury et al.,
-2006). Adolescents with mature socio-cognitive abilities tend to express
-stronger preferences for fairness than infants and children (Wynn et
-al., 2017). As children grow older, they become more likely to behave
-fairly, with their early-emerging strict egalitarianism being replaced
-by an increasing endorsement of fairness principles and engagement in
-collaborative activities (Huppert et al., 2019; McAuliffe et al., 2017).
-In these activities, their fairness views consider individual
-contributions, merits, and circumstances (Almås et al., 2010; Huppert et
-al., 2019; Sigelman \& Waitzman, 1991). Engelmann and Tomasello (2019)
-claim that children's sense of fairness emerges at three years old, and
-we can observe it in collaborative activities, where they accept
-inequality if the procedure gives everyone an equal chance. Therefore,
-children at this age respond to unequal distributions based on
-interpersonal concerns, as they already demand equal respect. In any
-case, between 3 and 8 years of age, inequitable and anti-meritorious
-allocations are evaluated more negatively, but equitable and meritorious
-allocations are not evaluated more positively (Elenbaas, 2019).
+consistent with research on cognitive development
+(\protect\hyperlink{ref-choudhury_social_2006}{Choudhury, Blakemore, and
+Charman 2006}). Adolescents with mature socio-cognitive abilities tend
+to express stronger preferences for fairness than infants and children
+(\protect\hyperlink{ref-wynn_not_2018}{Wynn et al. 2018}). As children
+grow older, they become more likely to behave fairly, with their
+early-emerging strict egalitarianism being replaced by an increasing
+endorsement of fairness principles and engagement in collaborative
+activities (\protect\hyperlink{ref-huppert_development_2019}{Huppert et
+al. 2019};
+\protect\hyperlink{ref-mcauliffe_developmental_2017}{McAuliffe et al.
+2017}). In these activities, their fairness views consider individual
+contributions, merits, and circumstances
+(\protect\hyperlink{ref-almas_fairness_2010}{Almås et al. 2010};
+\protect\hyperlink{ref-huppert_development_2019}{Huppert et al. 2019};
+\protect\hyperlink{ref-sigelman_development_1991}{Sigelman and Waitzman
+1991}). Engelmann and Tomasello
+(\protect\hyperlink{ref-engelmann_children_2019}{2019}) claim that
+children's sense of fairness emerges at three years old, and we can
+observe it in collaborative activities, where they accept inequality if
+the procedure gives everyone an equal chance. Therefore, children at
+this age respond to unequal distributions based on interpersonal
+concerns, as they already demand equal respect. In any case, between 3
+and 8 years of age, inequitable and anti-meritorious allocations are
+evaluated more negatively, but equitable and meritorious allocations are
+not evaluated more positively
+(\protect\hyperlink{ref-elenbaas_unfairness_2019}{Elenbaas 2019}).
 
@@
 behaviors by playing an essential role in the transmission of equity
-norms (Schunk \& Zipperle, 2023; Kosse et al., 2020). In fact, schools
-contribute to institutionalizing and reproducing inequality by promoting
-values, norms, practices, and languages familiar to higher-class
-families because the dominant group's culture shapes educational
-institutions (Bourdieu \& Passeron, 1990). Middle- and upper-class
-students are better equipped to face academic challenges and are more
-familiar with academic expectations (Mikus et al., 2019). Such
-familiarity represents cultural capital in educational contexts because
-higher-status students come to school ready to meet these expectations
-and reap the benefits (Jack, 2016; Khan, 2011). Conversely, lower-status
-children lacking cultural capital must catch up while experiencing
-inequitable comparisons (Goudeau \& Croizet, 2017). Additionally,
-academic achievement is treated as the outcome of dispositional factors
-(e.g., pupils' efforts and talents or lack of them) rather than the
-result of differential access to critical resources. Due to the
-meritocratic frame schools encourage, both low- and high-status
-individuals believe that students' success or failure is not due to
-their family background but rather to differences in efforts and talents
-(Darnon et al., 2018). In this sense, we believe that the perception of
-meritocracy can influence students' judgments about market justice
-preferences. Furthermore, we believe there is a difference between
-students' perceptions of meritocracy based on their own experience in
-school and what they perceive in society at large. Consistently, our
-first two hypotheses are:
+norms (\protect\hyperlink{ref-schunk_fairness_2023}{Schunk and Zipperle
+2023}; \protect\hyperlink{ref-kosse_prosociality_2020}{Kosse and Tincani
+2020}). In fact, schools contribute to institutionalizing and
+reproducing inequality by promoting values, norms, practices, and
+languages familiar to higher-class families because the dominant group's
+culture shapes educational institutions
+(\protect\hyperlink{ref-bourdieu_reproduction_1990}{Bourdieu and
+Passeron 1990}). Middle- and upper-class students are better equipped to
+face academic challenges and are more familiar with academic
+expectations (\protect\hyperlink{ref-mikus_children_2020}{Mikus, Tieben,
+and Schober 2020}). Such familiarity represents cultural capital in
+educational contexts because higher-status students come to school ready
+to meet these expectations and reap the benefits
+(\protect\hyperlink{ref-jack_no_2016}{Jack 2016};
+\protect\hyperlink{ref-khan_privilege_2011}{Khan 2011}). Conversely,
+lower-status children lacking cultural capital must catch up while
+experiencing inequitable comparisons
+(\protect\hyperlink{ref-goudeau_hidden_2017}{Goudeau and Croizet 2017}).
+Additionally, academic achievement is treated as the outcome of
+dispositional factors (e.g., pupils' efforts and talents or lack of
+them) rather than the result of differential access to critical
+resources. Due to the meritocratic frame schools encourage, both low-
+and high-status individuals believe that students' success or failure is
+not due to their family background but rather to differences in efforts
+and talents (\protect\hyperlink{ref-darnon_where_2018}{Darnon et al.
+2018}). In this sense, we believe that the perception of meritocracy can
+influence students' judgments about market justice preferences.
+Furthermore, we believe there is a difference between students'
+perceptions of meritocracy based on their own experience in school and
+what they perceive in society at large. Consistently, our first two
+hypotheses are:
 
@@
 children's and adolescent's market justice preferences. For example,
-Almås and colleagues (2017) found that adolescents from
-low-socioeconomic-status families are likelier to have an egalitarian
-fairness view and consider an equal distribution as fair in a situation
-with unequal merits. The authors speculate that differences in
-socialization practices across status groups might bring about, to a
-great extent, the fairness views of children and adolescents because
-social status seems to interact with these evaluations (e.g., Hvidberg
-et al., 2023).~
+Almås et al. (\protect\hyperlink{ref-almas_fairness_2017}{2017}) found
+that adolescents from low-socioeconomic-status families are likelier to
+have an egalitarian fairness view and consider an equal distribution as
+fair in a situation with unequal merits. The authors speculate that
+differences in socialization practices across status groups might bring
+about, to a great extent, the fairness views of children and adolescents
+because social status seems to interact with these evaluations (e.g.,
+\protect\hyperlink{ref-hvidberg_social_2023}{Hvidberg, Kreiner, and
+Stantcheva 2023}).~
 
@@
 curiosity, happiness, and consideration, while working-class parents
-promote deference, obedience, and conformity to authority (Kohn, 1963;
-Kohn \& Schooler, 1969). Although parents from all social backgrounds
-encourage individualism in their children, this shared norm translates
-into different forms in high and low social classes (1999). Acemoglu
-(2021) claimed that the values families impart to their children
-interact with social mobility. Because obedience is a valuable
-characteristic for employers, in low-wage and social mobility
-environments, low-income families impart values of obedience to their
-children to prevent disadvantaging them in labor markets. On the one
-hand, children from privileged families are socialized to adopt a clear
-conception of individualism that highlights their internal states,
-independence, and idiosyncrasies. In contrast, children from
-disadvantaged families are socialized to support a more balanced view of
-individualism that considers personal characteristics as resources to
-overcome collective impediments on the path to upward mobility
-(Iacoviello \& Lorenzi-Cioldi, 2019). In this way, we believe that there
-are differences in the socialization of values according to
-socioeconomic differences that could influence market justice
-preferences and, therefore, we propose the following hypothesis:
+promote deference, obedience, and conformity to authority
+(\protect\hyperlink{ref-kohn_social_1963}{Kohn 1963};
+\protect\hyperlink{ref-kohn_class_1969}{Kohn and Schooler 1969}).
+Although parents from all social backgrounds encourage individualism in
+their children, this shared norm translates into different forms in high
+and low social classes (1999). Acemoglu
+(\protect\hyperlink{ref-acemoglu_obedience_2021}{2021}) claimed that the
+values families impart to their children interact with social mobility.
+Because obedience is a valuable characteristic for employers, in
+low-wage and social mobility environments, low-income families impart
+values of obedience to their children to prevent disadvantaging them in
+labor markets. On the one hand, children from privileged families are
+socialized to adopt a clear conception of individualism that highlights
+their internal states, independence, and idiosyncrasies. In contrast,
+children from disadvantaged families are socialized to support a more
+balanced view of individualism that considers personal characteristics
+as resources to overcome collective impediments on the path to upward
+mobility
+(\protect\hyperlink{ref-iacoviello_collectivism_2019}{Iacoviello and
+Lorenzi-Cioldi 2019}). In this way, we believe that there are
+differences in the socialization of values according to socioeconomic
+differences that could influence market justice preferences and,
+therefore, we propose the following hypothesis:
 
@@
 significantly influences children's and adolescents' views on inequality
-and deservingness. For example, Jonsson and Beach's (2015) study
+and deservingness. For example, Jonsson and Beach
+(\protect\hyperlink{ref-jonsson_institutional_2015}{2015}) study
 revealed that higher-status adolescents in Sweden tend to perpetuate
@@
 others and legitimize the prevailing social hierarchies and economic
-disparities (Jost \& Burgess, 2000).
+disparities (\protect\hyperlink{ref-jost_attitudinal_2000}{Jost and
+Burgess 2000})
 
@@
 
-\subsubsection{Summary of hypotheses}\label{summary-of-hypotheses}
-
-\subsection{}\label{section}
-
-\subsection{Methods}\label{methods}
-
-\subsubsection{Data}\label{data}
+\hypertarget{summary-of-hypotheses}{%
+\subsubsection{Summary of hypotheses}\label{summary-of-hypotheses}}
+
+\hypertarget{section}{%
+\subsection{}\label{section}}
+
+\hypertarget{methods}{%
+\subsection{Methods}\label{methods}}
+
+\hypertarget{data}{%
+\subsubsection{Data}\label{data}}
 
@@
 
-\subsubsection{Variables}\label{variables}
+\hypertarget{variables}{%
+\subsubsection{Variables}\label{variables}}
 
@@
 
-\subsubsection{Methods}\label{methods-1}
+\hypertarget{methods-1}{%
+\subsubsection{Methods}\label{methods-1}}
 
@@
 
-\subsection{Analysis}\label{analysis}
-
-\subsubsection{Descriptive analysis}\label{descriptive-analysis}
+\hypertarget{analysis}{%
+\subsection{Analysis}\label{analysis}}
+
+\hypertarget{descriptive-analysis}{%
+\subsubsection{Descriptive analysis}\label{descriptive-analysis}}
 
@@
 
+\hypertarget{cumulative-link-mixed-models-for-the-justification-of-inequality-in-education-health-and-pensions}{%
 \subsubsection{Cumulative link mixed models for the Justification of
 inequality in education, health and
-pensions}\label{cumulative-link-mixed-models-for-the-justification-of-inequality-in-education-health-and-pensions}
+pensions}\label{cumulative-link-mixed-models-for-the-justification-of-inequality-in-education-health-and-pensions}}
 
@@
 
+\hypertarget{multilevel-regression-models-for-market-justice-preferences}{%
 \subsubsection{Multilevel regression models for market justice
-preferences}\label{multilevel-regression-models-for-market-justice-preferences}
+preferences}\label{multilevel-regression-models-for-market-justice-preferences}}
 
@@
 
-\subsubsection{Interactions effects}\label{interactions-effects}
+\hypertarget{interactions-effects}{%
+\subsubsection{Interactions effects}\label{interactions-effects}}
 
@@
 
-\subsection{Conclusion}\label{conclusion}
+\hypertarget{conclusion}{%
+\subsection{Conclusion}\label{conclusion}}
 
@@
 
-\subsection{Appendix}\label{appendix}
+\hypertarget{appendix}{%
+\subsection{Appendix}\label{appendix}}
 
@@
 
-\phantomsection\label{refs}
+\hypertarget{refs}{}
 \begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-almas_cutthroat_2020}
+\leavevmode\vadjust pre{\hypertarget{ref-acemoglu_obedience_2021}{}}%
+Acemoglu, Daron. 2021. {``Obedience in the {Labor Market} and {Social
+Mobility}: {A Socio-Economic Approach}.''} w29125. Cambridge, MA:
+National Bureau of Economic Research.
+\url{https://doi.org/10.3386/w29125}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-almas_fairness_2017}{}}%
+Almås, Ingvild, Alexander W Cappelen, Kjell G Salvanes, Erik Ø Sørensen,
+and Bertil Tungodden. 2017. {``Fairness and Family Background.''}
+\emph{Politics, Philosophy \& Economics} 16 (2): 117--31.
+\url{https://doi.org/10.1177/1470594X15618966}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-almas_fairness_2010}{}}%
+Almås, Ingvild, Alexander W. Cappelen, Erik Ø. Sørensen, and Bertil
+Tungodden. 2010. {``Fairness and the {Development} of {Inequality
+Acceptance}.''} \emph{Science} 328 (5982): 1176--78.
+\url{https://doi.org/10.1126/science.1187300}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-almas_cutthroat_2020}{}}%
 Almås, Ingvild, Alexander W. Cappelen, and Bertil Tungodden. 2020.
-{``Cutthroat Capitalism Versus Cuddly Socialism: Are Americans More
-Meritocratic and Efficiency-Seeking Than Scandinavians?''} \emph{Journal
-of Political Economy} 128 (5): 1753--88.
+{``Cutthroat {Capitalism} Versus {Cuddly Socialism}: {Are Americans More
+Meritocratic} and {Efficiency-Seeking} Than {Scandinavians}?''}
+\emph{Journal of Political Economy} 128 (5): 1753--88.
 \url{https://doi.org/10.1086/705551}.
 
-\bibitem[\citeproctext]{ref-barr_effect_2020}
+\leavevmode\vadjust pre{\hypertarget{ref-barr_effect_2020}{}}%
 Barr, Abigail, and Luis Miller. 2020. {``The Effect of Education, Income
@@
 
-\bibitem[\citeproctext]{ref-batruch_belief_2022}
+\leavevmode\vadjust pre{\hypertarget{ref-batruch_belief_2022}{}}%
 Batruch, Anatolia, Jolanda Jetten, Herman Van de Werfhorst, Céline
-Darnon, and Fabrizio Butera. 2022. {``Belief in School Meritocracy and
-the Legitimization of Social and Income Inequality.''} \emph{Social
-Psychological and Personality Science}, August, 194855062211110.
-\url{https://doi.org/10.1177/19485506221111017}.
-
-\bibitem[\citeproctext]{ref-bourdieu_reproduction_1990}
+Darnon, and Fabrizio Butera. 2022. {``Belief in {School Meritocracy} and
+the {Legitimization} of {Social} and {Income Inequality}.''}
+\emph{Social Psychological and Personality Science}, August,
+194855062211110. \url{https://doi.org/10.1177/19485506221111017}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-boltanski_new_2005}{}}%
+Boltanski, Luc, and Eve Chiapello. 2005. \emph{The New Spirit of
+Capitalism}. London ; New York: Verso.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bourdieu_reproduction_1990}{}}%
 Bourdieu, Pierre, and Jean Claude Passeron. 1990. \emph{Reproduction in
-Education, Society and Culture}. Second Edition. Sage Publications Ltd.
-
-\bibitem[\citeproctext]{ref-castillo_legitimacy_2011}
-Castillo, Juan Carlos. 2011. {``Legitimacy of Inequality in a Highly
-Unequal Context: Evidence from the Chilean Case.''} \emph{Social Justice
-Research} 24 (4): 314--40.
+{Education}, {Society} and {Culture}}. Second Edition. Sage Publications
+Ltd.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bucca_merit_2016}{}}%
+Bucca, Mauricio. 2016. {``Merit and Blame in Unequal Societies:
+{Explaining Latin Americans}' Beliefs about Wealth and Poverty.''}
+\emph{Research in Social Stratification and Mobility} 44 (June):
+98--112. \url{https://doi.org/10.1016/j.rssm.2016.02.005}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-castillo_legitimacy_2011}{}}%
+Castillo, Juan Carlos. 2011. {``Legitimacy of {Inequality} in a {Highly
+Unequal Context}: {Evidence} from the {Chilean Case}.''} \emph{Social
+Justice Research} 24 (4): 314--40.
 \url{https://doi.org/10.1007/s11211-011-0144-5}.
 
-\bibitem[\citeproctext]{ref-castillo_meritocracia_2019}
-Castillo, Juan C., Alex Torres, Jorge Atria, and Luis Maldonado. 2019.
-{``Meritocracia y Desigualdad Económica: Percepciones, Preferencias e
-Implicancias.''} \emph{Revista Internacional de Sociología} 77 (1): 117.
+\leavevmode\vadjust pre{\hypertarget{ref-castillo_meritocracia_2019}{}}%
+Castillo, Juan Carlos, Alex Torres, Jorge Atria, and Luis Maldonado.
+2019. {``Meritocracia y Desigualdad Econ{ó}mica: {Percepciones},
+Preferencias e Implicancias.''} \emph{Revista Internacional de
+Sociolog{í}a} 77 (1): 117.
 \url{https://doi.org/10.3989/ris.2019.77.1.17.114}.
 
-\bibitem[\citeproctext]{ref-chafel_schooling_1997}
-Chafel, Judith A. 1997. {``Schooling, the Hidden Curriculum, and
-Children's Conceptions of Poverty.''} \emph{Social Policy Report} 11
-(1): 1--28. \url{https://doi.org/10.1002/j.2379-3988.1997.tb00004.x}.
-
-\bibitem[\citeproctext]{ref-duru-bellat_who_2012}
-Duru-Bellat, Marie, and Elise Tenret. 2012. {``Who's for Meritocracy?
-Individual and Contextual Variations in the Faith.''} \emph{Comparative
-Education Review} 56 (2): 223--47. \url{https://doi.org/10.1086/661290}.
-
-\bibitem[\citeproctext]{ref-evans_strong_2018}
-Evans, M. D. R., and Jonathan Kelley. 2018. {``Strong Welfare States Do
-Not Intensify Public Support for Income Redistribution, but Even Reduce
-It Among the Prosperous: A Multilevel Analysis of Public Opinion in 30
-Countries.''} \emph{Societies} 8 (4): 105.
+\leavevmode\vadjust pre{\hypertarget{ref-centeno_arc_2012}{}}%
+Centeno, Miguel A., and Joseph N. Cohen. 2012. {``The {Arc} of
+{Neoliberalism}.''} \emph{Annual Review of Sociology} 38 (1): 317--40.
+\url{https://doi.org/10.1146/annurev-soc-081309-150235}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-chafel_schooling_1997}{}}%
+Chafel, Judith A. 1997. {``Schooling, the {Hidden Curriculum}, and
+{Children}'s {Conceptions} of {Poverty}.''} \emph{Social Policy Report}
+11 (1): 1--28. \url{https://doi.org/10.1002/j.2379-3988.1997.tb00004.x}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-chong_mystery_2008}{}}%
+Chong, Alberto, and Hugo Ñopo. 2008. {``The {Mystery} of
+{Discrimination} in {Latin America}.''} \emph{Econom{í}a} 8 (2):
+79--107. \url{https://doi.org/10.1353/eco.0.0005}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-choudhury_social_2006}{}}%
+Choudhury, Suparna, Sarah-Jayne Blakemore, and Tony Charman. 2006.
+{``Social Cognitive Development During Adolescence.''} \emph{Social
+Cognitive and Affective Neuroscience} 1 (3): 165--74.
+\url{https://doi.org/10.1093/scan/nsl024}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-darnon_where_2018}{}}%
+Darnon, Céline, Virginie Wiederkehr, Benoît Dompnier, and Delphine
+Martinot. 2018. {``{`{Where} There Is a Will, There Is a Way'}: {Belief}
+in School Meritocracy and the Social-Class Achievement Gap.''}
+\emph{British Journal of Social Psychology} 57 (1): 250--62.
+\url{https://doi.org/10.1111/bjso.12214}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-durante_preferences_2009}{}}%
+Durante, Ruben, and Louis G. Putterman. 2009. {``Preferences for
+{Redistribution} and {Perception} of {Fairness}: {An Experimental
+Study}.''} \emph{SSRN Electronic Journal}.
+\url{https://doi.org/10.2139/ssrn.1004573}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-duru-bellat_who_2012}{}}%
+Duru-Bellat, Marie, and Elise Tenret. 2012. {``Who's for {Meritocracy}?
+{Individual} and {Contextual Variations} in the {Faith}.''}
+\emph{Comparative Education Review} 56 (2): 223--47.
+\url{https://doi.org/10.1086/661290}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-elenbaas_unfairness_2019}{}}%
+Elenbaas, Laura. 2019. {``Against Unfairness: {Young} Children's
+Judgments about Merit, Equity, and Equality.''} \emph{Journal of
+Experimental Child Psychology} 186 (October): 73--82.
+\url{https://doi.org/10.1016/j.jecp.2019.05.009}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-engelmann_children_2019}{}}%
+Engelmann, Jan M., and Michael Tomasello. 2019. {``Children's {Sense} of
+{Fairness} as {Equal Respect}.''} \emph{Trends in Cognitive Sciences} 23
+(6): 454--63. \url{https://doi.org/10.1016/j.tics.2019.03.001}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-erivwo_meritocracy_2021}{}}%
+Erivwo, Aroesiri, Elizabeth Varghese, Anthony Mathai, and Tasmia Afrin.
+2021. {``Meritocracy in the {Educational System},''} May.
+\url{https://doi.org/10.5281/ZENODO.4740695}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-evans_strong_2018}{}}%
+Evans, M. D. R., and Jonathan Kelley. 2018. {``Strong {Welfare States Do
+Not Intensify Public Support} for {Income Redistribution}, but {Even
+Reduce It} Among the {Prosperous}: {A Multilevel Analysis} of {Public
+Opinion} in 30 {Countries}.''} \emph{Societies} 8 (4): 105.
 \url{https://doi.org/10.3390/soc8040105}.
 
-\bibitem[\citeproctext]{ref-fernandez_positive_2013}
+\leavevmode\vadjust pre{\hypertarget{ref-fernandez_positive_2013}{}}%
 Fernandez, J. J., and A. M. Jaime-Castillo. 2013. {``Positive or
-Negative Policy Feedbacks? Explaining Popular Attitudes Towards
-Pragmatic Pension Policy Reforms.''} \emph{European Sociological Review}
-29 (4): 803--15. \url{https://doi.org/10.1093/esr/jcs059}.
-
-\bibitem[\citeproctext]{ref-frank_performance_2015}
+{Negative Policy Feedbacks}? {Explaining Popular Attitudes Towards
+Pragmatic Pension Policy Reforms}.''} \emph{European Sociological
+Review} 29 (4): 803--15. \url{https://doi.org/10.1093/esr/jcs059}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-frank_performance_2015}{}}%
 Frank, Douglas H., Klaus Wertenbroch, and William W. Maddux. 2015.
-{``Performance Pay or Redistribution? Cultural Differences in Just-World
-Beliefs and Preferences for Wage Inequality.''} \emph{Organizational
-Behavior and Human Decision Processes} 130 (September): 160--70.
-\url{https://doi.org/10.1016/j.obhdp.2015.04.001}.
-
-\bibitem[\citeproctext]{ref-garcia-sanchez_attitudes_2020}
+{``Performance Pay or Redistribution? {Cultural} Differences in
+Just-World Beliefs and Preferences for Wage Inequality.''}
+\emph{Organizational Behavior and Human Decision Processes} 130
+(September): 160--70. \url{https://doi.org/10.1016/j.obhdp.2015.04.001}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-garcia-sanchez_attitudes_2020}{}}%
 García-Sánchez, Efraín, Danny Osborne, Guillermo B. Willis, and Rosa
@@
 
-\bibitem[\citeproctext]{ref-goldthorpe_myth_2003}
+\leavevmode\vadjust pre{\hypertarget{ref-goldthorpe_myth_2003}{}}%
 Goldthorpe, John. 2003. {``The Myth of Education-Based Meritocracy.''}
@@
 
-\bibitem[\citeproctext]{ref-homan_being_2017}
+\leavevmode\vadjust pre{\hypertarget{ref-goudeau_hidden_2017}{}}%
+Goudeau, Sébastien, and Jean-Claude Croizet. 2017. {``Hidden
+{Advantages} and {Disadvantages} of {Social Class}: {How Classroom
+Settings Reproduce Social Inequality} by {Staging Unfair Comparison}.''}
+\emph{Psychological Science} 28 (2): 162--70.
+\url{https://doi.org/10.1177/0956797616676600}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-greitemeyer_subjective_2016}{}}%
+Greitemeyer, Tobias, and Christina Sagioglou. 2016. {``Subjective
+Socioeconomic Status Causes Aggression: {A} Test of the Theory of Social
+Deprivation.''} \emph{Journal of Personality and Social Psychology} 111
+(2): 178--94. \url{https://doi.org/10.1037/pspi0000058}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-harvey_breve_2015}{}}%
+Harvey, David. 2015. \emph{{Breve historia del neoliberalismo}}. Madrid
+(Espa{ñ}a): Ediciones Akal.
+
+\leavevmode\vadjust pre{\hypertarget{ref-homan_being_2017}{}}%
 Homan, Patricia, Lauren Valentino, and Emi Weed. 2017. {``Being and
-Becoming Poor: How Cultural Schemas Shape Beliefs about Poverty.''}
+{Becoming Poor}: {How Cultural Schemas Shape Beliefs About Poverty}.''}
 \emph{Social Forces} 95 (3): 1023--48.
@@
 
-\bibitem[\citeproctext]{ref-kluegel_beliefs_1987}
+\leavevmode\vadjust pre{\hypertarget{ref-huppert_development_2019}{}}%
+Huppert, Elizabeth, Jason M. Cowell, Yawei Cheng, Carlos
+Contreras-Ibáñez, Natalia Gomez-Sicard, Maria Luz Gonzalez-Gadea, David
+Huepe, et al. 2019. {``The Development of Children's Preferences for
+Equality and Equity Across 13 Individualistic and Collectivist
+Cultures.''} \emph{Developmental Science} 22 (2): e12729.
+\url{https://doi.org/10.1111/desc.12729}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-hvidberg_social_2023}{}}%
+Hvidberg, Kristoffer B, Claus T Kreiner, and Stefanie Stantcheva. 2023.
+{``Social {Positions} and {Fairness Views} on {Inequality}.''}
+\emph{Review of Economic Studies} 90 (6): 3083--3118.
+\url{https://doi.org/10.1093/restud/rdad019}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-iacoviello_collectivism_2019}{}}%
+Iacoviello, Vincenzo, and Fabio Lorenzi-Cioldi. 2019. {``Collectivism
+and {Individualism} in {Status Hierarchies}: {Socialization} and {Social
+Identity Explanations}.''} \emph{International Review of Social
+Psychology} 32 (1): 15. \url{https://doi.org/10.5334/irsp.285}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-jack_no_2016}{}}%
+Jack, Anthony Abraham. 2016. {``({No}) {Harm} in {Asking}: {Class},
+{Acquired Cultural Capital}, and {Academic Engagement} at an {Elite
+University}.''} \emph{Sociology of Education} 89 (1): 1--19.
+\url{https://doi.org/10.1177/0038040715614913}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-jasso_how_1999}{}}%
+Jasso, Guillermina. 1999. {``How {Much Injustice} Is {There} in the
+{World}? {Two New Justice Indexes}.''} \emph{American Sociological
+Review} 64 (1): 133--68.
+\url{https://doi.org/10.1177/000312249906400110}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-jensen_worlds_2008}{}}%
+Jensen, Carsten. 2008. {``Worlds of Welfare Services and Transfers.''}
+\emph{Journal of European Social Policy} 18 (2): 151--62.
+\url{https://doi.org/10.1177/0958928707087591}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-jonsson_institutional_2015}{}}%
+Jonsson, Anna-Carin, and Dennis Beach. 2015. {``Institutional
+Discrimination: {Stereotypes} and Social Reproduction of {`Class'} in
+the {Swedish} Upper-Secondary School.''} \emph{Social Psychology of
+Education} 18 (4): 703--17.
+\url{https://doi.org/10.1007/s11218-014-9279-1}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-jost_attitudinal_2000}{}}%
+Jost, John T., and Diana Burgess. 2000. {``Attitudinal {Ambivalence} and
+the {Conflict} Between {Group} and {System Justification Motives} in
+{Low Status Groups}.''} \emph{Personality and Social Psychology
+Bulletin} 26 (3): 293--305.
+\url{https://doi.org/10.1177/0146167200265003}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-khan_privilege_2011}{}}%
+Khan, Shamus. 2011. \emph{Privilege: {The Making} of an {Adolescent
+Elite} at {St}. {Paul}'s {School}}. Princeton, N.J.: Princeton
+University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kluegel_beliefs_1987}{}}%
 Kluegel, James R., and Eliot R. Smith. 1987. \emph{Beliefs about
-Inequality: Americans' Views of What Is and What Ought to Be}. New York:
-Routledge.
-
-\bibitem[\citeproctext]{ref-lampert_meritocratic_2013}
-Lampert, Khen. 2013. \emph{Meritocratic Education and Social
-Worthlessness}. London: Palgrave Macmillan UK.
+Inequality: {Americans}' Views of What Is and What Ought to Be}. London
+New York: Routledge, Taylor \& Francis Group.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kohn_social_1963}{}}%
+Kohn, Melvin L. 1963. {``Social {Class} and {Parent-Child
+Relationships}: {An Interpretation}.''} \emph{American Journal of
+Sociology} 68 (4): 471--80. \url{https://doi.org/10.1086/223403}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kohn_class_1969}{}}%
+Kohn, Melvin L., and Carmi Schooler. 1969. {``Class, {Occupation}, and
+{Orientation}.''} \emph{American Sociological Review} 34 (5): 659.
+\url{https://doi.org/10.2307/2092303}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kosse_prosociality_2020}{}}%
+Kosse, Fabian, and Michela M. Tincani. 2020. {``Prosociality Predicts
+Labor Market Success Around the World.''} \emph{Nature Communications}
+11 (1): 5298. \url{https://doi.org/10.1038/s41467-020-19007-1}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lampert_meritocratic_2013}{}}%
+Lampert, Khen. 2013. \emph{Meritocratic {Education} and {Social
+Worthlessness}}. London: Palgrave Macmillan UK.
 \url{https://doi.org/10.1057/9781137324894}.
 
-\bibitem[\citeproctext]{ref-lindh_public_2015}
-Lindh, Arvid. 2015. {``Public Opinion Against Markets? Attitudes Towards
-Market Distribution of Social Services -- a Comparison of 17
-Countries.''} \emph{Social Policy \& Administration} 49 (7): 887--910.
-\url{https://doi.org/10.1111/spol.12105}.
-
-\bibitem[\citeproctext]{ref-mijs_stratified_2016}
-Mijs, Jonathan. 2016. {``Stratified Failure: Educational Stratification
-and Students' Attributions of Their Mathematics Performance in 24
-Countries.''} \emph{Sociology of Education} 89 (2): 137--53.
-\url{https://doi.org/10.1177/0038040716636434}.
-
-\bibitem[\citeproctext]{ref-mijs_paradox_2021}
-Mijs, Jonathan J B. 2021. {``The Paradox of Inequality: Income
-Inequality and Belief in Meritocracy Go Hand in Hand.''}
-\emph{Socio-Economic Review} 19 (1): 7--35.
-\url{https://doi.org/10.1093/ser/mwy051}.
-
-\bibitem[\citeproctext]{ref-resh_sense_2010}
+\leavevmode\vadjust pre{\hypertarget{ref-lane_market_1986}{}}%
+Lane, Robert E. 1986. {``Market {Justice}, {Political Justice}.''}
+\emph{American Political Science Review} 80 (2): 383--402.
+\url{https://doi.org/10.2307/1958264}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lindh_public_2015}{}}%
+Lindh, Arvid. 2015. {``Public {Opinion} Against {Markets}? {Attitudes}
+Towards {Market Distribution} of {Social Services} -- {A Comparison} of
+17 {Countries}.''} \emph{Social Policy \& Administration} 49 (7):
+887--910. \url{https://doi.org/10.1111/spol.12105}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mcauliffe_developmental_2017}{}}%
+McAuliffe, Katherine, Peter R. Blake, Nikolaus Steinbeis, and Felix
+Warneken. 2017. {``The Developmental Foundations of Human Fairness.''}
+\emph{Nature Human Behaviour} 1 (2): 0042.
+\url{https://doi.org/10.1038/s41562-016-0042}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mijs_stratified_2016}{}}%
+Mijs, Jonathan. 2016. {``Stratified {Failure}: {Educational
+Stratification} and {Students}' {Attributions} of {Their Mathematics
+Performance} in 24 {Countries}.''} \emph{Sociology of Education} 89 (2):
+137--53. \url{https://doi.org/10.1177/0038040716636434}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mijs_paradox_2021}{}}%
+---------. 2021. {``The Paradox of Inequality: Income Inequality and
+Belief in Meritocracy Go Hand in Hand.''} \emph{Socio-Economic Review}
+19 (1): 7--35. \url{https://doi.org/10.1093/ser/mwy051}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mikus_children_2020}{}}%
+Mikus, Karoline, Nicole Tieben, and Pia S. Schober. 2020. {``Children's
+Conversion of Cultural Capital into Educational Success: The Symbolic
+and Skill-Generating Functions of Cultural Capital.''} \emph{British
+Journal of Sociology of Education} 41 (2): 197--217.
+\url{https://doi.org/10.1080/01425692.2019.1677454}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mishra_subjective_2015}{}}%
+Mishra, Sandeep, and R. Nicholas Carleton. 2015. {``Subjective Relative
+Deprivation Is Associated with Poorer Physical and Mental Health.''}
+\emph{Social Science \& Medicine} 147 (December): 144--49.
+\url{https://doi.org/10.1016/j.socscimed.2015.10.030}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-molyneux_neoliberal_2008}{}}%
+Molyneux, Maxine. 2008. {``The {`{Neoliberal Turn}'} and the {New Social
+Policy} in {Latin America}: {How Neoliberal}, {How New}?''}
+\emph{Development and Change} 39 (5): 775--97.
+\url{https://doi.org/10.1111/j.1467-7660.2008.00505.x}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-morris_representing_2022}{}}%
+Morris, Jennifer, John Reilly, Sergey Paltsev, Andrei Sokolov, and
+Kenneth Cox. 2022. {``Representing {Socio}-{Economic Uncertainty} in
+{Human System Models}.''} \emph{Earth's Future} 10 (4): e2021EF002239.
+\url{https://doi.org/10.1029/2021EF002239}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-pierson_increasing_2000}{}}%
+Pierson, Paul. 2000. {``Increasing {Returns}, {Path Dependence}, and the
+{Study} of {Politics}.''} \emph{American Political Science Review} 94
+(2): 251--67. \url{https://doi.org/10.2307/2586011}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-power_deprivationprotest_2018}{}}%
+Power, Séamus A. 2018. {``The {Deprivation-Protest Paradox}: {How} the
+{Perception} of {Unfair Economic Inequality Leads} to {Civic Unrest}.''}
+\emph{Current Anthropology} 59 (6): 765--89.
+\url{https://doi.org/10.1086/700679}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-power_relative_2020}{}}%
+Power, Séamus A, Thomas Madsen, and Thomas A Morton. 2020. {``Relative
+Deprivation and Revolt: Current and Future Directions.''} \emph{Current
+Opinion in Psychology} 35 (October): 119--24.
+\url{https://doi.org/10.1016/j.copsyc.2020.06.010}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-resh_sense_2010}{}}%
 Resh, Nura. 2010. {``Sense of Justice about Grades in School: Is It
@@
 
-\bibitem[\citeproctext]{ref-resh_sense_2017}
-Resh, Nura, and Clara Sabbagh. 2017. {``Sense of Justice in School and
-Civic Behavior.''} \emph{Social Psychology of Education} 20 (2):
-387--409. \url{https://doi.org/10.1007/s11218-017-9375-0}.
-
-\bibitem[\citeproctext]{ref-sandel_tyranny_2020}
-Sandel, Michael J. 2020. \emph{The Tyranny of Merit: What's Become of
+\leavevmode\vadjust pre{\hypertarget{ref-resh_sense_2014}{}}%
+Resh, Nura, and Clara Sabbagh. 2014. {``Sense of Justice in School and
+Civic Attitudes.''} \emph{Social Psychology of Education} 17 (1):
+51--72. \url{https://doi.org/10.1007/s11218-013-9240-8}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-resh_sense_2017}{}}%
+---------. 2017. {``Sense of Justice in School and Civic Behavior.''}
+\emph{Social Psychology of Education} 20 (2): 387--409.
+\url{https://doi.org/10.1007/s11218-017-9375-0}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-salamon_marketization_1993}{}}%
+Salamon, Lester M. 1993. {``The {Marketization} of {Welfare}: {Changing
+Nonprofit} and {For-Profit Roles} in the {American Welfare State}.''}
+\emph{Social Service Review} 67 (1): 16--39.
+\url{https://doi.org/10.1086/603963}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-salgado_inequality_2023}{}}%
+Salgado, Mauricio, and Javier Castillo. 2023. {``Inequality and
+{Stratification} in {Latin America}.''} In \emph{The {Oxford Handbook}
+of {Social Stratification}}, edited by Markus Gangl, Lucinda Platt,
+Javier G. Polavieja, and Herman G. van de Werfhorst, 0. 2023: Oxford
+University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sandel_tyranny_2020}{}}%
+Sandel, Michael J. 2020. \emph{The Tyranny of Merit: {What}'s Become of
 the Common Good?} First edition. New York: {Farrar, Straus and Giroux}.
 
-\bibitem[\citeproctext]{ref-vonhippel_test_2019}
-Von Hippel, Paul, and Caitlin Hamrock. 2019. {``Do Test Score Gaps Grow
-Before, During, or Between the School Years? Measurement Artifacts and
-What We Can Know in Spite of Them.''} \emph{Sociological Science} 6:
-43--80. \url{https://doi.org/10.15195/v6.a3}.
-
-\bibitem[\citeproctext]{ref-wegener_dominant_1995}
-Wegener, Bernd, and Stefan Liebig. 1995. {``Dominant Ideologies and the
-Variation of Distributive Justice Norms: A Comparison of East and West
-Germany, and the United States.''} In \emph{Social Justice and Political
-Change: Public Opinion in Capitalist and Post-Communist States}, edited
-by James R. Kluegel, David S. Mason, and Bernd Wegener, 239--59. New
-York: Walter de Gruyter.
-
-\bibitem[\citeproctext]{ref-wiederkehr_belief_2015}
+\leavevmode\vadjust pre{\hypertarget{ref-schunk_fairness_2023}{}}%
+Schunk, Daniel, and Isabell Zipperle. 2023. {``Fairness and Inequality
+Acceptance in Children and Adolescents: {A} Survey on Behaviors in
+Economic Experiments''} 37 (5): 1715--42.
+\url{https://doi.org/10.1111/joes.12553}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-shariff_income_2016}{}}%
+Shariff, Azim F., Dylan Wiwad, and Lara B. Aknin. 2016. {``Income
+{Mobility Breeds Tolerance} for {Income Inequality}: {Cross-National}
+and {Experimental Evidence}.''} \emph{Perspectives on Psychological
+Science} 11 (3): 373--80.
+\url{https://doi.org/10.1177/1745691616635596}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sigelman_development_1991}{}}%
+Sigelman, Carol K., and Kara A. Waitzman. 1991. {``The {Development} of
+{Distributive Justice Orientations}: {Contextual Influences} on
+{Children}'s {Resource Allocations}.''} \emph{Child Development} 62 (6):
+1367. \url{https://doi.org/10.2307/1130812}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sivesind_changing_2017}{}}%
+Sivesind, Karl Henrik. 2017. {``The {Changing Roles} of {For-Profit} and
+{Nonprofit Welfare Provision} in {Norway}, {Sweden}, and {Denmark}.''}
+In \emph{Promoting {Active Citizenship}}, edited by Karl Henrik Sivesind
+and Jo Saglie, 33--74. Cham: Springer International Publishing.
+\url{https://doi.org/10.1007/978-3-319-55381-8_2}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-smith_relative_2012}{}}%
+Smith, Heather J., Thomas F. Pettigrew, Gina M. Pippin, and Silvana
+Bialosiewicz. 2012. {``Relative {Deprivation}: {A Theoretical} and
+{Meta-Analytic Review}.''} \emph{Personality and Social Psychology
+Review} 16 (3): 203--32. \url{https://doi.org/10.1177/1088868311430825}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-soler-martinez_concerns_2023}{}}%
+Soler-Martínez, Francisco Miguel, Efraín García-Sánchez, and Guillermo
+B. Willis. 2023. {``Concerns {About Inequality} in {Health}, {Education}
+and {Income Jointly Predict Collective Actions}.''} \emph{Revista
+Latinoamericana de Psicolog{í}a} 55: 99.
+\url{https://doi.org/10.14349/rlp.2023.v55.12}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-stoy_worlds_2014}{}}%
+Stoy, Volquart. 2014. {``Worlds of {Welfare Services}: {From Discovery}
+to {Exploration}.''} \emph{Social Policy \& Administration} 48 (3):
+343--60. \url{https://doi.org/10.1111/spol.12006}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-streeck_citizens_2012}{}}%
+Streeck, Wolfgang. 2012. {``Citizens as {Customers}: {Considerations} on
+the {New Politics} of {Consumption}.''} \emph{New Left Review} 76.
+
+\leavevmode\vadjust pre{\hypertarget{ref-torche_intergenerational_2014}{}}%
+Torche, Florencia. 2014. {``Intergenerational {Mobility} and
+{Inequality}: {The Latin American Case}.''} \emph{Annual Review of
+Sociology} 40 (1): 619--42.
+\url{https://doi.org/10.1146/annurev-soc-071811-145521}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vonhippel_test_2019}{}}%
+Von Hippel, Paul, and Caitlin Hamrock. 2019. {``Do {Test Score Gaps
+Grow} Before, During, or Between the {School Years}? {Measurement
+Artifacts} and {What We Can Know} in {Spite} of {Them}.''}
+\emph{Sociological Science} 6: 43--80.
+\url{https://doi.org/10.15195/v6.a3}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-weaver_paths_2010}{}}%
+Weaver, Kent. 2010. {``Paths and {Forks} or {Chutes} and {Ladders}?:
+{Negative Feedbacks} and {Policy Regime Change}.''} \emph{Journal of
+Public Policy} 30 (2): 137--62.
+\url{https://doi.org/10.1017/S0143814X10000061}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wegener_dominant_1995}{}}%
+Wegener, Bernd, and Stefan Liebig. 1995. {``Dominant {Ideologies} and
+the {Variation} of {Distributive Justice Norms}: {A Comparison} of
+{East} and {West Germany}, and the {United States}.''} In \emph{Social
+{Justice} and {Political Change}: {Public Opinion} in {Capitalist} and
+{Post-Communist States}}, edited by James R. Kluegel, David S. Mason,
+and Bernd Wegener, 239--59. New York: Walter de Gruyter.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wiederkehr_belief_2015}{}}%
 Wiederkehr, Virginie, Virginie Bonnot, Silvia Krauth-Gruber, and Céline
@@
 Tool for Low Status Students.''} \emph{Frontiers in Psychology} 6.
-\url{https://www.frontiersin.org/articles/10.3389/fpsyg.2015.01053}.
-
-\bibitem[\citeproctext]{ref-young_rise_1958}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wynn_not_2018}{}}%
+Wynn, Karen, Paul Bloom, Ashley Jordan, Julia Marshall, and Mark
+Sheskin. 2018. {``Not {Noble Savages After All}: {Limits} to {Early
+Altruism}.''} \emph{Current Directions in Psychological Science} 27 (1):
+3--8. \url{https://doi.org/10.1177/0963721417734875}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-young_rise_1958}{}}%
 Young, Michael. 1958. \emph{The Rise of the Meritocracy}. New Brunswick,

### 2024-05-08 02-antecedentes.qmd
update plots and tables antecedentes y metodos

@@
 For instance, the 2019 International Social Survey Program (ISSP) provides insights into how adults perceive inequality in accessing welfare services.
-Figure 1 illustrates the variation in agreement levels regarding whether it is just for individuals with higher incomes to purchase better education or healthcare.
+@fig-issp illustrates the variation in agreement levels regarding whether it is just for individuals with higher incomes to purchase better education or healthcare.
 Notably, in 18 of the 29 countries analyzed, there is a greater justification for market inequality in healthcare access than in education.
@@
 
-```{r}
+```{r echo=FALSE}
+pacman::p_load(dplyr, ggplot2, haven)
+load("input/data/issp_rec.RData")
+```
+
+```{r echo=FALSE}
+#| label: fig-issp
 #| fig-cap: "Average of market justice preferences by country"
+
+base$country <- as.character(base$country)
+base$country <- substr(base$country, 4, nchar(base$country))
+base$country <- as_factor(base$country)
+
+base %>% ggplot(aes(mean, country)) +
+  theme_bw() +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  stat_summary(aes(xintercept = after_stat(x), y = 0), fun = mean, geom = "vline", orientation = "y", colour = "green") +
+  xlab("") +
+  ylab("Country")+
+  scale_x_continuous(limits = c(1, 5),
+                     breaks = seq(1, 5, by = 1),
+                     label = c("Very unjust,\n definitely wrong", 
+                               "Somewhat unjust,\n wrong", 
+                               "Neither just nor\n unjust, mixed feelings", 
+                               "Somewhat just,\n right", 
+                               "Very just,\n definitely right")) +
+  scale_color_discrete(name = NULL) +
+  scale_shape_discrete(name = NULL) +
+  theme(text = element_text(size = 9.35),
+#        legend.position = "bottom",
+        plot.margin = margin(b = 0))
 ```
@@
 
-```{r}
+```{r echo=FALSE}
 #| fig-cap: "Summary of hypotheses"
+#| label: fig-hypotheses
+
+knitr::include_graphics("input/img/hypotheses.png")
 ```

### 2024-05-08 03-methods.qmd
render paper
update plots and tables antecedentes y metodos

@@
 ---
-title: "03-methods"
 author: "Equipo Edumer"
@@
 
-Table 1 shows the items used, their response categories, and their frequencies.
+@tbl-desc-dependientes shows the items used, their response categories, and their frequencies.
 
-```{r}
+```{r echo=FALSE}
+pacman::p_load(summarytools, dplyr)
+load("input/data/dependientes.RData")
+
+```
+
+```{r, results='asis', echo=FALSE}
 #| tbl-cap: "Dependent variables"
+#| label: tbl-desc-dependientes
+
+df<-dfSummary(dependientes,
+               plain.ascii = FALSE,
+               style = "multiline",
+               tmp.img.dir = "/tmp",
+               graph.magnif = 0.75,
+               headings = F,  # encabezado
+               varnumbers = F, # num variable
+               labels.col = T, # etiquetas
+               na.col = F,    # missing
+               graph.col = F, # plot
+               valid.col = T, # n valido
+               col.widths = c(30,10,10,10))
+
+df$Variable <- NULL # delete variable column
+
+print(df)
 ```
@@
 Likewise, an index of access to technology includes the number of computers, tablets, and cell phones at home, as well as whether there is an Internet connection.
-Table 2 shows the items used, their response categories, and their frequency.
+@tbl-desc-independent shows the items used, their response categories, and their frequency.
 
-```{r}
+```{r echo=FALSE}
+load("input/data/independientes.RData")
+```
+
+```{r, results='asis', echo=FALSE}
 #| tbl-cap: "Independent variables"
+#| label: tbl-desc-independent
+
+df<-dfSummary(independientes,
+               plain.ascii = FALSE,
+               style = "multiline",
+               tmp.img.dir = "/tmp",
+               graph.magnif = 0.75,
+               headings = F,  # encabezado
+               varnumbers = F, # num variable
+               labels.col = T, # etiquetas
+               na.col = F,    # missing
+               graph.col = F, # plot
+               valid.col = T, # n valido
+               col.widths = c(30,10,10,10))
+
+df$Variable <- NULL # delete variable column
+
+print(df)
 ```
@@
 The school-level variables are the administrative dependency of the school, the socioeconomic classification made by the Ministry of Education, the level of performance in the SIMCE test of the school, and the proportion of parents with university or postgraduate degrees.
-Table 3 shows the items used, response categories, and frequency.
+@tbl-desc-school shows the items used, response categories, and frequency.
 
-```{r}
+```{r echo=FALSE}
+load("input/data/school.RData")
+```
+
+```{r, results='asis', echo=FALSE}
 #| tbl-cap: "School context variables"
+#| label: tbl-desc-school
+
+df<-dfSummary(school,
+               plain.ascii = FALSE,
+               style = "multiline",
+               tmp.img.dir = "/tmp",
+               graph.magnif = 0.75,
+               headings = F,  # encabezado
+               varnumbers = F, # num variable
+               labels.col = T, # etiquetas
+               na.col = F,    # missing
+               graph.col = F, # plot
+               valid.col = T, # n valido
+               col.widths = c(30,10,10,10))
+
+df$Variable <- NULL # delete variable column
+
+print(df)
 ```
@@
 
-The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open ScorrectlyOSF), the access to the document is available at this \[link\](https://doi.org/10.17605/OSF.IO/UFSDV).
+The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open Science (OSF), the access to the document is available at this [link](https://doi.org/10.17605/OSF.IO/UFSDV).
 The statistical analysis of this research was performed using the free software R version 4.1.3.

### 2024-05-08 04-analisis.qmd
render paper
update plots and tables antecedentes y metodos

@@
 ---
-title: "04-analisis"
 author: "Equipo Edumer"
@@
 
-```{r}
+```{r echo=FALSE}
+load("input/data/deservingness.RData")
+load("input/data/soc_effort.RData")
+load("input/data/soc_talent.RData")
+load("input/data/school_effort.RData")
+load("input/data/school_talent.RData")
+pacman::p_load(ggpubr)
+```
+
+```{r echo=FALSE}
 #| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
+#| fig-asp: 1.5
+
+plot_soc_effort <- soc_effort %>% ggplot(aes(esfuerzo_soc, mean)) +
+  theme_bw(base_size = 10) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  xlab("Social effort") +
+  ylab("Market justice preferences") +
+  scale_y_continuous(limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_discrete(name = NULL) +
+  scale_shape_discrete(name = NULL)
+
+plot_deservingness <- deservingness %>% ggplot(aes(merito_soc, mean)) +
+  theme_bw(base_size = 10) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  xlab("Deservingness") +
+  ylab("Market justice preferences") +
+  scale_y_continuous(limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_discrete(name = NULL) +
+  scale_shape_discrete(name = NULL)
+
+plot_soc_talent <- soc_talent %>% ggplot(aes(inteligencia_soc, mean)) +
+  theme_bw(base_size = 10) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  xlab("Social talent") +
+  ylab("Market justice preferences") +
+  scale_y_continuous(limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_discrete(name = NULL) +
+  scale_shape_discrete(name = NULL)
+
+plot_school_effort <- school_effort %>% ggplot(aes(esfuerzo_esc, mean)) +
+  theme_bw(base_size = 10) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  xlab("School effort") +
+  ylab("") +
+  scale_y_continuous(limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_discrete(name = NULL) +
+  scale_shape_discrete(name = NULL)
+
+plot_school_talent <- school_talent %>% ggplot(aes(inteligencia_esc, mean)) +
+  theme_bw(base_size = 10) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  xlab("School talent") +
+  ylab("") +
+  scale_y_continuous(limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_discrete(name = NULL) +
+  scale_shape_discrete(name = NULL)
+
+ggarrange(plot_soc_effort, plot_school_effort, plot_soc_talent, plot_school_talent, plot_deservingness, ncol = 2, nrow = 3, common.legend = TRUE)
 ```

### 2024-05-08 paper.tex
update plots and tables antecedentes y metodos

--- 2024-04-24 paper.tex
@@
 (\citeproc{ref-batruch_belief_2022}{Batruch et al. 2022};
-\citeproc{ref-wiederkehr_belief_2015}{Wiederkehr et al. 2015}).~
+\citeproc{ref-wiederkehr_belief_2015}{Wiederkehr et al. 2015}).
 
@@
 by a highly stratified educational system. In particular, we focus on
-market justice preferences (Lane, 1986), which refer to the preferences
-for distributing public goods (as health and education) based on
-criteria such as competition and payment capacity. Although from a
-rational point of view it could be expected an opposition to market
-justice by the underprivileged majority,~ we argue that in a social
-environment characterized by a promotion of meritocratic ideals - as the
-school system - would lead to the opposite: a larger market justice
-preferences.~
+market justice preferences (\citeproc{ref-lane_market_1986}{Lane 1986}),
+which refer to the preferences for distributing public goods (as health
+and education) based on criteria such as competition and payment
+capacity. Although from a rational point of view it could be expected an
+opposition to market justice by the underprivileged majority, we argue
+that in a social environment characterized by a promotion of
+meritocratic ideals - as the school system - would lead to the opposite:
+a larger market justice preferences.
 
@@
 emphasizing the so-called unfulfillable promise of meritocracy
-(\citeproc{ref-mijs_stratified_2016}{J. Mijs 2016}). Complementing this
+(\citeproc{ref-mijs_stratified_2016}{Mijs 2016}). Complementing this
 agenda, a second and emerging research area deals with subjective
-aspects of meritocracy, such as perceptions and beliefs.~
+aspects of meritocracy, such as perceptions and beliefs.
 
@@
 (\citeproc{ref-duru-bellat_who_2012}{Duru-Bellat and Tenret 2012};
-\citeproc{ref-castillo_meritocracia_2019}{Juan C. Castillo et al.
-2019}). This perception can vary greatly depending on individual
-experiences, social, economic, and cultural background. Some people may
-see meritocracy as a fair and just system that allows anyone to succeed
-based on their abilities and hard work. In contrast, others may view it
-as a myth or a cover for existing power dynamics and inequality, serving
-to maintain and even reinforce inequality
+\citeproc{ref-castillo_meritocracia_2019}{Castillo et al. 2019}). This
+perception can vary greatly depending on individual experiences, social,
+economic, and cultural background. Some people may see meritocracy as a
+fair and just system that allows anyone to succeed based on their
+abilities and hard work. In contrast, others may view it as a myth or a
+cover for existing power dynamics and inequality, serving to maintain
+and even reinforce inequality
 (\citeproc{ref-lampert_meritocratic_2013}{Lampert 2013};
-\citeproc{ref-mijs_paradox_2021}{J. J. B. Mijs 2021}). Based in this
-last perspective, we argue that individuals with a higher perception of
+\citeproc{ref-mijs_paradox_2021}{Mijs 2021}). Based in this last
+perspective, we argue that individuals with a higher perception of
 meritocracy will show a larger justification of social inequalities, as
@@
 willing to justify better services in these domains for those with
-higher incomes.~ Secondly, we focus on the student-age population as we
+higher incomes. Secondly, we focus on the student-age population as we
 point out that it is possible to track down the origin of meritocratic
@@
 of cultural beliefs by transmitting cultural norms, values, and
-expectations to young people.~
+expectations to young people.
 
@@
 significant attention has been paid to wage inequality, income
-distribution, and payment differentials in the literature {[}Juan Carlos
-Castillo (\citeproc{ref-castillo_legitimacy_2011}{2011}); {]} (Castillo,
-2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),~ there has
-been less examination of public beliefs about which life domains should
-be governed by market relations and even less about children's
-acceptance or rejection of these market principles. This oversight is
-notable given the extensive encroachment of market logic into public
-goods, welfare policy, and social services over the past five decades
-(Centeno \& Cohen, 2012; Harvey, 2007), affecting areas such as
+distribution, and payment differentials in the literature {[}Castillo
+(\citeproc{ref-castillo_legitimacy_2011}{2011}); Evans et al., 2010;
+Jasso (\citeproc{ref-jasso_how_1999}{1999}) ; Shariff, Wiwad, and Aknin
+(\citeproc{ref-shariff_income_2016}{2016}){]}, there has been less
+examination of public beliefs about which life domains should be
+governed by market relations and even less about children's acceptance
+or rejection of these market principles. This oversight is notable given
+the extensive encroachment of market logic into public goods, welfare
+policy, and social services over the past five decades
+(\citeproc{ref-centeno_arc_2012}{Centeno and Cohen 2012};
+\citeproc{ref-harvey_breve_2015}{Harvey 2015}), affecting areas such as
 pensions, health services, and education. The justification of social
@@
 justice, that is, affording legitimacy to the allocation of goods and
-services based on prices and individuals' ability to pay (Boltanski \&
-Chiapello- 2007; Lane 1986; Streeck 2012).
+services based on prices and individuals' ability to pay
+(\citeproc{ref-boltanski_new_2005}{Boltanski and Chiapello 2005};
+\citeproc{ref-lane_market_1986}{Lane 1986};
+\citeproc{ref-streeck_citizens_2012}{Streeck 2012}).
 
@@
 the market to stimulate wants without at the same time stimulating a
-sense of deserving more than one gets'' (1986, p.~384). Following the
-theory of relative deprivation --a social phenomenon arising when
-individuals cannot afford what most others in their environment can
-(Merton 1950)-- Lane notes that, in market settings, social comparisons
-are more likely to motivate increased effort rather than feelings of
-acute injustice because individuals attribute outcomes to their actions.
-Although empirical research has shown that, contrary to Lane's
-observation, relative deprivation, even in market settings, produces
-feelings of dissatisfaction, anger, and resentment that might motivate
-forms of collective action such as protests and revolt (Greitemeyer and
-Sagioglou 2016; Mishra and Carleton 2015; Smith et al.~2012; Power 2018;
-Power, Madsen, and Morton 2020), the conceptualization of market justice
-has been since then closely coupled to the merit principle for
-allocating outcomes: the claim that the unequal levels of well-being
-individuals enjoy ought to be, to some extent, a function of their
-talents and efforts, regardless of their needs or membership, the two
-latter being the realm of political justice and its closely coupled
-principles of need --allocating outcomes to those who require them
-most-- and equality --allocating the same outcome to everyone--, which
-have been at the center of welfarism (see Wilson 2003). The underlying
-notion of market justice also resonates in its application to welfare
-regimes, providing a framework for understanding the varied global
-approaches to managing social services.
+sense of deserving more than one gets'' {[}-Lane
+(\citeproc{ref-lane_market_1986}{1986}); p.~384{]}. Following the theory
+of relative deprivation --a social phenomenon arising when individuals
+cannot afford what most others in their environment can (Merton 1950)--
+Lane notes that, in market settings, social comparisons are more likely
+to motivate increased effort rather than feelings of acute injustice
+because individuals attribute outcomes to their actions. Although
+empirical research has shown that, contrary to Lane's observation,
+relative deprivation, even in market settings, produces feelings of
+dissatisfaction, anger, and resentment that might motivate forms of
+collective action such as protests and revolt
+(\citeproc{ref-greitemeyer_subjective_2016}{Greitemeyer and Sagioglou
+2016}; \citeproc{ref-mishra_subjective_2015}{Mishra and Carleton 2015};
+\citeproc{ref-smith_relative_2012}{Smith et al. 2012};
+\citeproc{ref-power_deprivationprotest_2018}{Séamus A. Power 2018};
+\citeproc{ref-power_relative_2020}{Séamus A. Power, Madsen, and Morton
+2020}), the conceptualization of market justice has been since then
+closely coupled to the merit principle for allocating outcomes: the
+claim that the unequal levels of well-being individuals enjoy ought to
+be, to some extent, a function of their talents and efforts, regardless
+of their needs or membership, the two latter being the realm of
+political justice and its closely coupled principles of need
+--allocating outcomes to those who require them most-- and equality
+--allocating the same outcome to everyone--, which have been at the
+center of welfarism (see Wilson 2003). The underlying notion of market
+justice also resonates in its application to welfare regimes, providing
+a framework for understanding the varied global approaches to managing
+social services.
 
@@
 nations, with substantial differences in funding and delivery methods
-(Jensen, 2008; Stoy, 2014). Nordic countries, for example, predominantly
-employ public agencies to produce and provide social services, funding
-these through collective taxation and offering them in kind to the
-majority of citizens. This system prioritizes political justice, placing
-it above market mechanisms in accessing services. In contrast, other
-countries rely more heavily on for-profit entities and private funding,
-where service distribution depends mainly on individual financial
-capacity of paying user fees, highlighting the influence of market
-justice in service allocation. The trend toward marketization of welfare
-services has been growing since the 1980s (Salamon, 1993), and this
+(\citeproc{ref-jensen_worlds_2008}{Jensen 2008};
+\citeproc{ref-stoy_worlds_2014}{Stoy 2014}). Nordic countries, for
+example, predominantly employ public agencies to produce and provide
+social services, funding these through collective taxation and offering
+them in kind to the majority of citizens. This system prioritizes
+political justice, placing it above market mechanisms in accessing
+services. In contrast, other countries rely more heavily on for-profit
+entities and private funding, where service distribution depends mainly
+on individual financial capacity of paying user fees, highlighting the
+influence of market justice in service allocation. The trend toward
+marketization of welfare services has been growing since the 1980s
+(\citeproc{ref-salamon_marketization_1993}{Salamon 1993}), and this
 shift is increasingly evident even in countries where market solutions
-have traditionally had a minor role in social policy (Sivesind, 2017).
+have traditionally had a minor role in social policy
+(\citeproc{ref-sivesind_changing_2017}{Sivesind 2017}).
 
@@
 (\citeproc{ref-fernandez_positive_2013}{Fernandez and Jaime-Castillo
-2013}) (Fernandez and Jaime-Castillo 2013; Pierson 2000; Weaver 2010),
-citizens' beliefs about market justice are likely also shaped by the
-institutional and social contexts they encounter. Indeed, the
-justification of inequality in access to essential services like
-education and health, based on one's ability to pay, shows significant
-variation across countries, as demonstrated by international surveys,
-although they do not usually include children in their samples. For
-instance, the 2019 International Social Survey Program (ISSP) provides
-insights into how adults perceive inequality in accessing welfare
-services. Figure 1 illustrates the variation in agreement levels
-regarding whether it is just for individuals with higher incomes to
-purchase better education or healthcare. Notably, in 18 of the 29
-countries analyzed, there is a greater justification for market
-inequality in healthcare access than in education. Nevertheless, the
-general sentiment typically ranges from `somewhat unjust' to 'neither
-just nor unjust, with mixed feelings.
+2013}; \citeproc{ref-pierson_increasing_2000}{Pierson 2000};
+\citeproc{ref-weaver_paths_2010}{Weaver 2010}), citizens' beliefs about
+market justice are likely also shaped by the institutional and social
+contexts they encounter. Indeed, the justification of inequality in
+access to essential services like education and health, based on one's
+ability to pay, shows significant variation across countries, as
+demonstrated by international surveys, although they do not usually
+include children in their samples. For instance, the 2019 International
+Social Survey Program (ISSP) provides insights into how adults perceive
+inequality in accessing welfare services. Figure~\ref{fig-issp}
+illustrates the variation in agreement levels regarding whether it is
+just for individuals with higher incomes to purchase better education or
+healthcare. Notably, in 18 of the 29 countries analyzed, there is a
+greater justification for market inequality in healthcare access than in
+education. Nevertheless, the general sentiment typically ranges from
+`somewhat unjust' to 'neither just nor unjust, with mixed feelings.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-issp-1.pdf}
+
+}
+
+\caption{\label{fig-issp}Average of market justice preferences by
+country}
+
+\end{figure}%
 
@@
 disapproval of market stratification of essential services. This finding
-is corroborated by Soler-Martínez and colleagues' (2023) research from
+is corroborated by Soler-Martínez, García-Sánchez, and Willis
+(\citeproc{ref-soler-martinez_concerns_2023}{2023}) research from
 Latinobarómetro 2020 across 18 Latin American countries, where concerns
@@
 typically driven by elite political decisions rather than grassroots
-demand.~
+demand.
 
@@
 responsible for their economic outcomes, a view that varies across the
-region
-(\citeproc{ref-buccaMeritBlameUnequal2016}{\textbf{buccaMeritBlameUnequal2016?}};
-\citeproc{ref-chongMysteryDiscriminationLatin2008}{\textbf{chongMysteryDiscriminationLatin2008?}};
-\citeproc{ref-torcheIntergenerationalMobilityInequality2014}{\textbf{torcheIntergenerationalMobilityInequality2014?}};
-\citeproc{ref-salgadoInequalityStratificationLatin2023}{\textbf{salgadoInequalityStratificationLatin2023?}}).
-The reliance on private welfare providers and widespread user fees
-(Molyneux, 2008) adds complexity to this context. Yet, research on
-children's justification of market-based inequalities in accessing
-welfare services remains limited, especially in Latin America,
-highlighting a significant gap in understanding how younger generations
-view market-based access to welfare and whether these views are
-associated with their meritocratic beliefs.~
+region (\citeproc{ref-bucca_merit_2016}{Bucca 2016};
+\citeproc{ref-chong_mystery_2008}{Chong and Ñopo 2008};
+\citeproc{ref-torche_intergenerational_2014}{Torche 2014};
+\citeproc{ref-salgado_inequality_2023}{Salgado and Castillo 2023}). The
+reliance on private welfare providers and widespread user fees
+(\citeproc{ref-molyneux_neoliberal_2008}{Molyneux 2008}) adds complexity
+to this context. Yet, research on children's justification of
+market-based inequalities in accessing welfare services remains limited,
+especially in Latin America, highlighting a significant gap in
+understanding how younger generations view market-based access to
+welfare and whether these views are associated with their meritocratic
+beliefs.
 
@@
 preferences (Castillo et al), whereby most of the link this subjective
-dimensions to individual socio-structural factors and~ context-level
+dimensions to individual socio-structural factors and context-level
 determinants. For instance, some studies have analyzed how those with
@@
 how greater economic inequality increases meritocratic beliefs
-(\citeproc{ref-mijs_paradox_2021a}{\textbf{mijs\_paradox\_2021a?}}), and
-how larger inequality affects meritocratic beliefs (Morris et al 2022).
-Based on these findings, a research agenda has been reinforced on the
+(\citeproc{ref-mijs_paradox_2021}{Mijs 2021}), and how larger inequality
+affects meritocratic beliefs
+(\citeproc{ref-morris_representing_2022}{Morris et al. 2022}). Based on
+these findings, a research agenda has been reinforced on the
 legitimizing role of meritocracy, in line with previous studies using
@@
 among those who endorsed meritocratic and equal opportunity beliefs. In
-the same line, experimental research by Durante \& Putterman (2009)
-subjects support less redistribution when the initial distribution is
-determined according to task performance.
+the same line, experimental research by Durante and Putterman
+(\citeproc{ref-durante_preferences_2009}{2009}) subjects support less
+redistribution when the initial distribution is determined according to
+task performance.
 
@@
 socialization institutions where achievement based on merit explains
-success (Erivwo et al 2021). Nevertheless, it is possible to find some
-initial works focused on the area of distributive justice at school that
-are closely related to meritocratic beliefs, as the ones by Resh and
-Sabbagh (quotes). Using justice in grade obtention as a measure of
-distributive justice and meritocracy (Sabbagh et al 2006), they find for
-instance that a larger sense of distributive justice about grades is
-associated to higher socio-economic status
-(\citeproc{ref-resh_sense_2010}{Resh 2010}), have a positive effect on
-liberal democratic orientation and on trust in people and in formal
-institutions (Resh \& Sabbagh 2013), and tend to refrain from violence
-and to engage to a greater extent in extra-curricular school activity
-and community volunteering (\citeproc{ref-resh_sense_2017}{Resh and
-Sabbagh 2017}).
+success (\citeproc{ref-erivwo_meritocracy_2021}{Erivwo et al. 2021}).
+Nevertheless, it is possible to find some initial works focused on the
+area of distributive justice at school that are closely related to
+meritocratic beliefs, as the ones by Resh and Sabbagh (quotes). Using
+justice in grade obtention as a measure of distributive justice and
+meritocracy (Sabbagh et al 2006), they find for instance that a larger
+sense of distributive justice about grades is associated to higher
+socio-economic status (\citeproc{ref-resh_sense_2010}{Resh 2010}), have
+a positive effect on liberal democratic orientation and on trust in
+people and in formal institutions (\citeproc{ref-resh_sense_2014}{Resh
+and Sabbagh 2014}), and tend to refrain from violence and to engage to a
+greater extent in extra-curricular school activity and community
+volunteering (\citeproc{ref-resh_sense_2017}{Resh and Sabbagh 2017}).
 
@@
 life. The differences in economic understanding across age groups are
-consistent with research on cognitive development (Choudhury et al.,
-2006). Adolescents with mature socio-cognitive abilities tend to express
-stronger preferences for fairness than infants and children (Wynn et
-al., 2017). As children grow older, they become more likely to behave
-fairly, with their early-emerging strict egalitarianism being replaced
-by an increasing endorsement of fairness principles and engagement in
-collaborative activities (Huppert et al., 2019; McAuliffe et al., 2017).
-In these activities, their fairness views consider individual
-contributions, merits, and circumstances (Almås et al., 2010; Huppert et
-al., 2019; Sigelman \& Waitzman, 1991). Engelmann and Tomasello (2019)
+consistent with research on cognitive development
+(\citeproc{ref-choudhury_social_2006}{Choudhury, Blakemore, and Charman
+2006}). Adolescents with mature socio-cognitive abilities tend to
+express stronger preferences for fairness than infants and children
+(\citeproc{ref-wynn_not_2018}{Wynn et al. 2018}). As children grow
+older, they become more likely to behave fairly, with their
+early-emerging strict egalitarianism being replaced by an increasing
+endorsement of fairness principles and engagement in collaborative
+activities (\citeproc{ref-huppert_development_2019}{Huppert et al.
+2019}; \citeproc{ref-mcauliffe_developmental_2017}{McAuliffe et al.
+2017}). In these activities, their fairness views consider individual
+contributions, merits, and circumstances
+(\citeproc{ref-almas_fairness_2010}{Almås et al. 2010};
+\citeproc{ref-huppert_development_2019}{Huppert et al. 2019};
+\citeproc{ref-sigelman_development_1991}{Sigelman and Waitzman 1991}).
+Engelmann and Tomasello (\citeproc{ref-engelmann_children_2019}{2019})
 claim that children's sense of fairness emerges at three years old, and
@@
 allocations are evaluated more negatively, but equitable and meritorious
-allocations are not evaluated more positively (Elenbaas, 2019).
+allocations are not evaluated more positively
+(\citeproc{ref-elenbaas_unfairness_2019}{Elenbaas 2019}).
 
@@
 behaviors by playing an essential role in the transmission of equity
-norms (Schunk \& Zipperle, 2023; Kosse et al., 2020). In fact, schools
-contribute to institutionalizing and reproducing inequality by promoting
-values, norms, practices, and languages familiar to higher-class
-families because the dominant group's culture shapes educational
-institutions (Bourdieu \& Passeron, 1990). Middle- and upper-class
-students are better equipped to face academic challenges and are more
-familiar with academic expectations (Mikus et al., 2019). Such
-familiarity represents cultural capital in educational contexts because
-higher-status students come to school ready to meet these expectations
-and reap the benefits (Jack, 2016; Khan, 2011). Conversely, lower-status
-children lacking cultural capital must catch up while experiencing
-inequitable comparisons (Goudeau \& Croizet, 2017). Additionally,
-academic achievement is treated as the outcome of dispositional factors
-(e.g., pupils' efforts and talents or lack of them) rather than the
-result of differential access to critical resources. Due to the
-meritocratic frame schools encourage, both low- and high-status
-individuals believe that students' success or failure is not due to
-their family background but rather to differences in efforts and talents
-(Darnon et al., 2018). In this sense, we believe that the perception of
-meritocracy can influence students' judgments about market justice
-preferences. Furthermore, we believe there is a difference between
-students' perceptions of meritocracy based on their own experience in
-school and what they perceive in society at large. Consistently, our
-first two hypotheses are:
+norms (\citeproc{ref-schunk_fairness_2023}{Schunk and Zipperle 2023};
+\citeproc{ref-kosse_prosociality_2020}{Kosse and Tincani 2020}). In
+fact, schools contribute to institutionalizing and reproducing
+inequality by promoting values, norms, practices, and languages familiar
+to higher-class families because the dominant group's culture shapes
+educational institutions
+(\citeproc{ref-bourdieu_reproduction_1990}{Bourdieu and Passeron 1990}).
+Middle- and upper-class students are better equipped to face academic
+challenges and are more familiar with academic expectations
+(\citeproc{ref-mikus_children_2020}{Mikus, Tieben, and Schober 2020}).
+Such familiarity represents cultural capital in educational contexts
+because higher-status students come to school ready to meet these
+expectations and reap the benefits (\citeproc{ref-jack_no_2016}{Jack
+2016}; \citeproc{ref-khan_privilege_2011}{Khan 2011}). Conversely,
+lower-status children lacking cultural capital must catch up while
+experiencing inequitable comparisons
+(\citeproc{ref-goudeau_hidden_2017}{Goudeau and Croizet 2017}).
+Additionally, academic achievement is treated as the outcome of
+dispositional factors (e.g., pupils' efforts and talents or lack of
+them) rather than the result of differential access to critical
+resources. Due to the meritocratic frame schools encourage, both low-
+and high-status individuals believe that students' success or failure is
+not due to their family background but rather to differences in efforts
+and talents (\citeproc{ref-darnon_where_2018}{Darnon et al. 2018}). In
+this sense, we believe that the perception of meritocracy can influence
+students' judgments about market justice preferences. Furthermore, we
+believe there is a difference between students' perceptions of
+meritocracy based on their own experience in school and what they
+perceive in society at large. Consistently, our first two hypotheses
+are:
 
@@
 children's and adolescent's market justice preferences. For example,
-Almås and colleagues (2017) found that adolescents from
-low-socioeconomic-status families are likelier to have an egalitarian
-fairness view and consider an equal distribution as fair in a situation
-with unequal merits. The authors speculate that differences in
-socialization practices across status groups might bring about, to a
-great extent, the fairness views of children and adolescents because
-social status seems to interact with these evaluations (e.g., Hvidberg
-et al., 2023).~
+Almås et al. (\citeproc{ref-almas_fairness_2017}{2017}) found that
+adolescents from low-socioeconomic-status families are likelier to have
+an egalitarian fairness view and consider an equal distribution as fair
+in a situation with unequal merits. The authors speculate that
+differences in socialization practices across status groups might bring
+about, to a great extent, the fairness views of children and adolescents
+because social status seems to interact with these evaluations (e.g.,
+\citeproc{ref-hvidberg_social_2023}{Hvidberg, Kreiner, and Stantcheva
+2023}).
 
@@
 curiosity, happiness, and consideration, while working-class parents
-promote deference, obedience, and conformity to authority (Kohn, 1963;
-Kohn \& Schooler, 1969). Although parents from all social backgrounds
-encourage individualism in their children, this shared norm translates
-into different forms in high and low social classes (1999). Acemoglu
-(2021) claimed that the values families impart to their children
-interact with social mobility. Because obedience is a valuable
-characteristic for employers, in low-wage and social mobility
-environments, low-income families impart values of obedience to their
-children to prevent disadvantaging them in labor markets. On the one
-hand, children from privileged families are socialized to adopt a clear
-conception of individualism that highlights their internal states,
-independence, and idiosyncrasies. In contrast, children from
-disadvantaged families are socialized to support a more balanced view of
-individualism that considers personal characteristics as resources to
-overcome collective impediments on the path to upward mobility
-(Iacoviello \& Lorenzi-Cioldi, 2019). In this way, we believe that there
-are differences in the socialization of values according to
-socioeconomic differences that could influence market justice
-preferences and, therefore, we propose the following hypothesis:
+promote deference, obedience, and conformity to authority
+(\citeproc{ref-kohn_social_1963}{Kohn 1963};
+\citeproc{ref-kohn_class_1969}{Kohn and Schooler 1969}). Although
+parents from all social backgrounds encourage individualism in their
+children, this shared norm translates into different forms in high and
+low social classes (1999). Acemoglu
+(\citeproc{ref-acemoglu_obedience_2021}{2021}) claimed that the values
+families impart to their children interact with social mobility. Because
+obedience is a valuable characteristic for employers, in low-wage and
+social mobility environments, low-income families impart values of
+obedience to their children to prevent disadvantaging them in labor
+markets. On the one hand, children from privileged families are
+socialized to adopt a clear conception of individualism that highlights
+their internal states, independence, and idiosyncrasies. In contrast,
+children from disadvantaged families are socialized to support a more
+balanced view of individualism that considers personal characteristics
+as resources to overcome collective impediments on the path to upward
+mobility (\citeproc{ref-iacoviello_collectivism_2019}{Iacoviello and
+Lorenzi-Cioldi 2019}). In this way, we believe that there are
+differences in the socialization of values according to socioeconomic
+differences that could influence market justice preferences and,
+therefore, we propose the following hypothesis:
 
@@
 significantly influences children's and adolescents' views on inequality
-and deservingness. For example, Jonsson and Beach's (2015) study
-revealed that higher-status adolescents in Sweden tend to perpetuate
-social class stereotypes while describing the vocational and academic
-tracks. Academic track students are depicted as wealthy, intelligent,
-ambitious, and diligent, while vocational track students are
-characterized as poor, unambitious, unintelligent, and lackadaisical.
-These stereotypes help individuals maintain a sense of superiority over
-others and legitimize the prevailing social hierarchies and economic
-disparities (Jost \& Burgess, 2000).
+and deservingness. For example, Jonsson and Beach
+(\citeproc{ref-jonsson_institutional_2015}{2015}) study revealed that
+higher-status adolescents in Sweden tend to perpetuate social class
+stereotypes while describing the vocational and academic tracks.
+Academic track students are depicted as wealthy, intelligent, ambitious,
+and diligent, while vocational track students are characterized as poor,
+unambitious, unintelligent, and lackadaisical. These stereotypes help
+individuals maintain a sense of superiority over others and legitimize
+the prevailing social hierarchies and economic disparities
+(\citeproc{ref-jost_attitudinal_2000}{Jost and Burgess 2000})
 
@@
 \subsubsection{Summary of hypotheses}\label{summary-of-hypotheses}
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{input/img/hypotheses.png}
+
+}
+
+\caption{\label{fig-hypotheses}Summary of hypotheses}
+
+\end{figure}%
 
@@
 
-Table 1 shows the items used, their response categories, and their
-frequencies.
+Table~\ref{tbl-desc-dependientes} shows the items used, their response
+categories, and their frequencies.
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.4286}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2551}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2143}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1020}}@{}}
+\caption{Dependent
+variables}\label{tbl-desc-dependientes}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Label
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Stats / Values
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Freqs (\% of Valid)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Valid
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Label
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Stats / Values
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Freqs (\% of Valid)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Valid
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+It is just that in Chile people with higher incomes can have better
+pensions than people with low incomes &
+\begin{minipage}[t]{\linewidth}\raggedright
+1. Strongly disagree\\
+2. Disagree\\
+3. Agree\\
+4. Strongly agree\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+1837 (30.6\%)\\
+1945 (32.4\%)\\
+1622 (27.0\%)\\
+608 (10.1\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+6012\\
+(95.9\%)\strut
+\end{minipage} \\
+It is just that in Chile people who can pay have a better education for
+their children & \begin{minipage}[t]{\linewidth}\raggedright
+1. Strongly disagree\\
+2. Disagree\\
+3. Agree\\
+4. Strongly agree\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+1766 (29.7\%)\\
+1732 (29.1\%)\\
+1704 (28.6\%)\\
+750 (12.6\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+5952\\
+(94.9\%)\strut
+\end{minipage} \\
+It is just that in Chile people with higher incomes can access better
+health services than people with low incomes &
+\begin{minipage}[t]{\linewidth}\raggedright
+1. Strongly disagree\\
+2. Disagree\\
+3. Agree\\
+4. Strongly agree\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+2254 (38.0\%)\\
+1685 (28.4\%)\\
+1401 (23.6\%)\\
+593 (10.0\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+5933\\
+(94.6\%)\strut
+\end{minipage} \\
+Inequality justification index &
+\begin{minipage}[t]{\linewidth}\raggedright
+Mean (sd) : 2.2 (0.9)\\
+min \textless{} med \textless{} max:\\
+1 \textless{} 2 \textless{} 4\\
+IQR (CV) : 1.7 (0.4)\strut
+\end{minipage} & 13 distinct values &
+\begin{minipage}[t]{\linewidth}\raggedright
+6077\\
+(96.9\%)\strut
+\end{minipage} \\
+\end{longtable}
 
@@
 number of computers, tablets, and cell phones at home, as well as
-whether there is an Internet connection. Table 2 shows the items used,
-their response categories, and their frequency.
+whether there is an Internet connection.
+Table~\ref{tbl-desc-independent} shows the items used, their response
+categories, and their frequency.
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3925}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3084}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1963}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1028}}@{}}
+\caption{Independent
+variables}\label{tbl-desc-independent}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Label
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Stats / Values
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Freqs (\% of Valid)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Valid
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Label
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Stats / Values
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Freqs (\% of Valid)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Valid
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Intelligence is important to get good grades &
+\begin{minipage}[t]{\linewidth}\raggedright
+1. Strongly disagree\\
+2. Disagree\\
+3. Agree\\
+4. Strongly agree\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+367 ( 6.1\%)\\
+920 (15.3\%)\\
+2970 (49.4\%)\\
+1760 (29.3\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+6017\\
+(95.9\%)\strut
+\end{minipage} \\
+Effort is important to get good grades &
+\begin{minipage}[t]{\linewidth}\raggedright
+1. Strongly disagree\\
+2. Disagree\\
+3. Agree\\
+4. Strongly agree\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+109 ( 1.8\%)\\
+88 ( 1.5\%)\\
+1427 (23.7\%)\\
+4406 (73.1\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+6030\\
+(96.1\%)\strut
+\end{minipage} \\
+In Chile, people are rewarded for their intelligence and skill &
+\begin{minipage}[t]{\linewidth}\raggedright
+1. Strongly disagree\\
+2. Disagree\\
+3. Agree\\
+4. Strongly agree\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+517 ( 9.0\%)\\
+1568 (27.3\%)\\
+2673 (46.6\%)\\
+983 (17.1\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+5741\\
+(91.5\%)\strut
+\end{minipage} \\
+In Chile, people are rewarded for their efforts &
+\begin{minipage}[t]{\linewidth}\raggedright
+1. Strongly disagree\\
+2. Disagree\\
+3. Agree\\
+4. Strongly agree\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+512 ( 8.7\%)\\
+1733 (29.4\%)\\
+2607 (44.2\%)\\
+1050 (17.8\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+5902\\
+(94.1\%)\strut
+\end{minipage} \\
+In Chile, people get what they deserve &
+\begin{minipage}[t]{\linewidth}\raggedright
+1. Strongly disagree\\
+2. Disagree\\
+3. Agree\\
+4. Strongly agree\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+604 (10.5\%)\\
+1911 (33.1\%)\\
+2388 (41.4\%)\\
+871 (15.1\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+5774\\
+(92.1\%)\strut
+\end{minipage} \\
+Parental educational level & \begin{minipage}[t]{\linewidth}\raggedright
+1. 8th grade or less\\
+2. Secondary Education\\
+3. Higher tec. education\\
+4. University or Postgraduat\\
+5. Missing\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+559 ( 8.9\%)\\
+1698 (27.1\%)\\
+960 (15.3\%)\\
+1080 (17.2\%)\\
+1975 (31.5\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+6272\\
+(100.0\%)\strut
+\end{minipage} \\
+Number of books at home & \begin{minipage}[t]{\linewidth}\raggedright
+1. Les than 25\\
+2. More than 25\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+3920 (63.2\%)\\
+2281 (36.8\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+6201\\
+(98.9\%)\strut
+\end{minipage} \\
+Technology access index & \begin{minipage}[t]{\linewidth}\raggedright
+Mean (sd) : 7.8 (2.5)\\
+min \textless{} med \textless{} max:\\
+0 \textless{} 8 \textless{} 12\\
+IQR (CV) : 3 (0.3)\strut
+\end{minipage} & 13 distinct values &
+\begin{minipage}[t]{\linewidth}\raggedright
+6272\\
+(100.0\%)\strut
+\end{minipage} \\
+\end{longtable}
 
@@
 Education, the level of performance in the SIMCE test of the school, and
-the proportion of parents with university or postgraduate degrees. Table
-3 shows the items used, response categories, and frequency.
+the proportion of parents with university or postgraduate degrees.
+Table~\ref{tbl-desc-school} shows the items used, response categories,
+and frequency.
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.4040}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2626}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2222}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1111}}@{}}
+\caption{School context variables}\label{tbl-desc-school}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Label
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Stats / Values
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Freqs (\% of Valid)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Valid
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Label
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Stats / Values
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Freqs (\% of Valid)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Valid
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Proportion of parents with university level by school &
+\begin{minipage}[t]{\linewidth}\raggedright
+Mean (sd) : 0.2 (0.2)\\
+min \textless{} med \textless{} max:\\
+0 \textless{} 0.1 \textless{} 0.9\\
+IQR (CV) : 0.2 (0.9)\strut
+\end{minipage} & 103 distinct values &
+\begin{minipage}[t]{\linewidth}\raggedright
+6272\\
+(100.0\%)\strut
+\end{minipage} \\
+SIMCE score by school & \begin{minipage}[t]{\linewidth}\raggedright
+1. Low\\
+2. Medium\\
+3. High\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+2091 (33.3\%)\\
+2091 (33.3\%)\\
+2090 (33.3\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+6272\\
+(100.0\%)\strut
+\end{minipage} \\
+Administrative dependency of school &
+\begin{minipage}[t]{\linewidth}\raggedright
+1. Public\\
+2. Subsidized private\\
+3. Private\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+2659 (42.4\%)\\
+3169 (50.5\%)\\
+444 ( 7.1\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+6272\\
+(100.0\%)\strut
+\end{minipage} \\
+Socioeconomic level of school &
+\begin{minipage}[t]{\linewidth}\raggedright
+1. Low\\
+2. Medium low\\
+3. Medium\\
+4. Medium high\\
+5. High\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+720 (11.5\%)\\
+2282 (36.4\%)\\
+1383 (22.1\%)\\
+1309 (20.9\%)\\
+578 ( 9.2\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+6272\\
+(100.0\%)\strut
+\end{minipage} \\
+\end{longtable}
 
@@
 The hypotheses of this research were pre-registered in the Open Science
-Framework platform of the Center for Open ScorrectlyOSF), the access to
+Framework platform of the Center for Open Science (OSF), the access to
 the document is available at this
-{[}link{]}(https://doi.org/10.17605/OSF.IO/UFSDV). The statistical
+\href{https://doi.org/10.17605/OSF.IO/UFSDV}{link}. The statistical
 analysis of this research was performed using the free software R
@@
 \subsubsection{Descriptive analysis}\label{descriptive-analysis}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/unnamed-chunk-11-1.pdf}
+
+}
+
+\caption{Market justice preferences in education, health and pensions by
+social and school meritocracy}
+
+\end{figure}%
 
@@
 students that attend schools with better achievement scores on average
-justify less inequality.~
+justify less inequality.
 
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-acemoglu_obedience_2021}
+Acemoglu, Daron. 2021. {``Obedience in the {Labor Market} and {Social
+Mobility}: {A Socio-Economic Approach}.''} w29125. Cambridge, MA:
+National Bureau of Economic Research.
+\url{https://doi.org/10.3386/w29125}.
+
+\bibitem[\citeproctext]{ref-almas_fairness_2017}
+Almås, Ingvild, Alexander W Cappelen, Kjell G Salvanes, Erik Ø Sørensen,
+and Bertil Tungodden. 2017. {``Fairness and Family Background.''}
+\emph{Politics, Philosophy \& Economics} 16 (2): 117--31.
+\url{https://doi.org/10.1177/1470594X15618966}.
+
+\bibitem[\citeproctext]{ref-almas_fairness_2010}
+Almås, Ingvild, Alexander W. Cappelen, Erik Ø. Sørensen, and Bertil
+Tungodden. 2010. {``Fairness and the {Development} of {Inequality
+Acceptance}.''} \emph{Science} 328 (5982): 1176--78.
+\url{https://doi.org/10.1126/science.1187300}.
+
 \bibitem[\citeproctext]{ref-almas_cutthroat_2020}
 Almås, Ingvild, Alexander W. Cappelen, and Bertil Tungodden. 2020.
-{``Cutthroat Capitalism Versus Cuddly Socialism: Are Americans More
-Meritocratic and Efficiency-Seeking Than Scandinavians?''} \emph{Journal
-of Political Economy} 128 (5): 1753--88.
+{``Cutthroat {Capitalism} Versus {Cuddly Socialism}: {Are Americans More
+Meritocratic} and {Efficiency-Seeking} Than {Scandinavians}?''}
+\emph{Journal of Political Economy} 128 (5): 1753--88.
 \url{https://doi.org/10.1086/705551}.
@@
 Batruch, Anatolia, Jolanda Jetten, Herman Van de Werfhorst, Céline
-Darnon, and Fabrizio Butera. 2022. {``Belief in School Meritocracy and
-the Legitimization of Social and Income Inequality.''} \emph{Social
-Psychological and Personality Science}, August, 194855062211110.
-\url{https://doi.org/10.1177/19485506221111017}.
+Darnon, and Fabrizio Butera. 2022. {``Belief in {School Meritocracy} and
+the {Legitimization} of {Social} and {Income Inequality}.''}
+\emph{Social Psychological and Personality Science}, August,
+194855062211110. \url{https://doi.org/10.1177/19485506221111017}.
+
+\bibitem[\citeproctext]{ref-boltanski_new_2005}
+Boltanski, Luc, and Eve Chiapello. 2005. \emph{The New Spirit of
+Capitalism}. London ; New York: Verso.
 
@@
 Bourdieu, Pierre, and Jean Claude Passeron. 1990. \emph{Reproduction in
-Education, Society and Culture}. Second Edition. Sage Publications Ltd.
+{Education}, {Society} and {Culture}}. Second Edition. Sage Publications
+Ltd.
+
+\bibitem[\citeproctext]{ref-bucca_merit_2016}
+Bucca, Mauricio. 2016. {``Merit and Blame in Unequal Societies:
+{Explaining Latin Americans}' Beliefs about Wealth and Poverty.''}
+\emph{Research in Social Stratification and Mobility} 44 (June):
+98--112. \url{https://doi.org/10.1016/j.rssm.2016.02.005}.
 
 \bibitem[\citeproctext]{ref-castillo_legitimacy_2011}
-Castillo, Juan Carlos. 2011. {``Legitimacy of Inequality in a Highly
-Unequal Context: Evidence from the Chilean Case.''} \emph{Social Justice
-Research} 24 (4): 314--40.
+Castillo, Juan Carlos. 2011. {``Legitimacy of {Inequality} in a {Highly
+Unequal Context}: {Evidence} from the {Chilean Case}.''} \emph{Social
+Justice Research} 24 (4): 314--40.
 \url{https://doi.org/10.1007/s11211-011-0144-5}.
@@
 \bibitem[\citeproctext]{ref-castillo_meritocracia_2019}
-Castillo, Juan C., Alex Torres, Jorge Atria, and Luis Maldonado. 2019.
-{``Meritocracia y Desigualdad Económica: Percepciones, Preferencias e
-Implicancias.''} \emph{Revista Internacional de Sociología} 77 (1): 117.
+Castillo, Juan Carlos, Alex Torres, Jorge Atria, and Luis Maldonado.
+2019. {``Meritocracia y Desigualdad Econ{ó}mica: {Percepciones},
+Preferencias e Implicancias.''} \emph{Revista Internacional de
+Sociolog{í}a} 77 (1): 117.
 \url{https://doi.org/10.3989/ris.2019.77.1.17.114}.
 
+\bibitem[\citeproctext]{ref-centeno_arc_2012}
+Centeno, Miguel A., and Joseph N. Cohen. 2012. {``The {Arc} of
+{Neoliberalism}.''} \emph{Annual Review of Sociology} 38 (1): 317--40.
+\url{https://doi.org/10.1146/annurev-soc-081309-150235}.
+
 \bibitem[\citeproctext]{ref-chafel_schooling_1997}
-Chafel, Judith A. 1997. {``Schooling, the Hidden Curriculum, and
-Children's Conceptions of Poverty.''} \emph{Social Policy Report} 11
-(1): 1--28. \url{https://doi.org/10.1002/j.2379-3988.1997.tb00004.x}.
+Chafel, Judith A. 1997. {``Schooling, the {Hidden Curriculum}, and
+{Children}'s {Conceptions} of {Poverty}.''} \emph{Social Policy Report}
+11 (1): 1--28. \url{https://doi.org/10.1002/j.2379-3988.1997.tb00004.x}.
+
+\bibitem[\citeproctext]{ref-chong_mystery_2008}
+Chong, Alberto, and Hugo Ñopo. 2008. {``The {Mystery} of
+{Discrimination} in {Latin America}.''} \emph{Econom{í}a} 8 (2):
+79--107. \url{https://doi.org/10.1353/eco.0.0005}.
+
+\bibitem[\citeproctext]{ref-choudhury_social_2006}
+Choudhury, Suparna, Sarah-Jayne Blakemore, and Tony Charman. 2006.
+{``Social Cognitive Development During Adolescence.''} \emph{Social
+Cognitive and Affective Neuroscience} 1 (3): 165--74.
+\url{https://doi.org/10.1093/scan/nsl024}.
+
+\bibitem[\citeproctext]{ref-darnon_where_2018}
+Darnon, Céline, Virginie Wiederkehr, Benoît Dompnier, and Delphine
+Martinot. 2018. {``{`{Where} There Is a Will, There Is a Way'}: {Belief}
+in School Meritocracy and the Social-Class Achievement Gap.''}
+\emph{British Journal of Social Psychology} 57 (1): 250--62.
+\url{https://doi.org/10.1111/bjso.12214}.
+
+\bibitem[\citeproctext]{ref-durante_preferences_2009}
+Durante, Ruben, and Louis G. Putterman. 2009. {``Preferences for
+{Redistribution} and {Perception} of {Fairness}: {An Experimental
+Study}.''} \emph{SSRN Electronic Journal}.
+\url{https://doi.org/10.2139/ssrn.1004573}.
 
 \bibitem[\citeproctext]{ref-duru-bellat_who_2012}
-Duru-Bellat, Marie, and Elise Tenret. 2012. {``Who's for Meritocracy?
-Individual and Contextual Variations in the Faith.''} \emph{Comparative
-Education Review} 56 (2): 223--47. \url{https://doi.org/10.1086/661290}.
+Duru-Bellat, Marie, and Elise Tenret. 2012. {``Who's for {Meritocracy}?
+{Individual} and {Contextual Variations} in the {Faith}.''}
+\emph{Comparative Education Review} 56 (2): 223--47.
+\url{https://doi.org/10.1086/661290}.
+
+\bibitem[\citeproctext]{ref-elenbaas_unfairness_2019}
+Elenbaas, Laura. 2019. {``Against Unfairness: {Young} Children's
+Judgments about Merit, Equity, and Equality.''} \emph{Journal of
+Experimental Child Psychology} 186 (October): 73--82.
+\url{https://doi.org/10.1016/j.jecp.2019.05.009}.
+
+\bibitem[\citeproctext]{ref-engelmann_children_2019}
+Engelmann, Jan M., and Michael Tomasello. 2019. {``Children's {Sense} of
+{Fairness} as {Equal Respect}.''} \emph{Trends in Cognitive Sciences} 23
+(6): 454--63. \url{https://doi.org/10.1016/j.tics.2019.03.001}.
+
+\bibitem[\citeproctext]{ref-erivwo_meritocracy_2021}
+Erivwo, Aroesiri, Elizabeth Varghese, Anthony Mathai, and Tasmia Afrin.
+2021. {``Meritocracy in the {Educational System},''} May.
+\url{https://doi.org/10.5281/ZENODO.4740695}.
 
 \bibitem[\citeproctext]{ref-evans_strong_2018}
-Evans, M. D. R., and Jonathan Kelley. 2018. {``Strong Welfare States Do
-Not Intensify Public Support for Income Redistribution, but Even Reduce
-It Among the Prosperous: A Multilevel Analysis of Public Opinion in 30
-Countries.''} \emph{Societies} 8 (4): 105.
+Evans, M. D. R., and Jonathan Kelley. 2018. {``Strong {Welfare States Do
+Not Intensify Public Support} for {Income Redistribution}, but {Even
+Reduce It} Among the {Prosperous}: {A Multilevel Analysis} of {Public
+Opinion} in 30 {Countries}.''} \emph{Societies} 8 (4): 105.
 \url{https://doi.org/10.3390/soc8040105}.
@@
 Fernandez, J. J., and A. M. Jaime-Castillo. 2013. {``Positive or
-Negative Policy Feedbacks? Explaining Popular Attitudes Towards
-Pragmatic Pension Policy Reforms.''} \emph{European Sociological Review}
-29 (4): 803--15. \url{https://doi.org/10.1093/esr/jcs059}.
+{Negative Policy Feedbacks}? {Explaining Popular Attitudes Towards
+Pragmatic Pension Policy Reforms}.''} \emph{European Sociological
+Review} 29 (4): 803--15. \url{https://doi.org/10.1093/esr/jcs059}.
 
@@
 Frank, Douglas H., Klaus Wertenbroch, and William W. Maddux. 2015.
-{``Performance Pay or Redistribution? Cultural Differences in Just-World
-Beliefs and Preferences for Wage Inequality.''} \emph{Organizational
-Behavior and Human Decision Processes} 130 (September): 160--70.
-\url{https://doi.org/10.1016/j.obhdp.2015.04.001}.
+{``Performance Pay or Redistribution? {Cultural} Differences in
+Just-World Beliefs and Preferences for Wage Inequality.''}
+\emph{Organizational Behavior and Human Decision Processes} 130
+(September): 160--70. \url{https://doi.org/10.1016/j.obhdp.2015.04.001}.
 
@@
 
+\bibitem[\citeproctext]{ref-goudeau_hidden_2017}
+Goudeau, Sébastien, and Jean-Claude Croizet. 2017. {``Hidden
+{Advantages} and {Disadvantages} of {Social Class}: {How Classroom
+Settings Reproduce Social Inequality} by {Staging Unfair Comparison}.''}
+\emph{Psychological Science} 28 (2): 162--70.
+\url{https://doi.org/10.1177/0956797616676600}.
+
+\bibitem[\citeproctext]{ref-greitemeyer_subjective_2016}
+Greitemeyer, Tobias, and Christina Sagioglou. 2016. {``Subjective
+Socioeconomic Status Causes Aggression: {A} Test of the Theory of Social
+Deprivation.''} \emph{Journal of Personality and Social Psychology} 111
+(2): 178--94. \url{https://doi.org/10.1037/pspi0000058}.
+
+\bibitem[\citeproctext]{ref-harvey_breve_2015}
+Harvey, David. 2015. \emph{{Breve historia del neoliberalismo}}. Madrid
+(Espa{ñ}a): Ediciones Akal.
+
 \bibitem[\citeproctext]{ref-homan_being_2017}
 Homan, Patricia, Lauren Valentino, and Emi Weed. 2017. {``Being and
-Becoming Poor: How Cultural Schemas Shape Beliefs about Poverty.''}
+{Becoming Poor}: {How Cultural Schemas Shape Beliefs About Poverty}.''}
 \emph{Social Forces} 95 (3): 1023--48.
@@
 
+\bibitem[\citeproctext]{ref-huppert_development_2019}
+Huppert, Elizabeth, Jason M. Cowell, Yawei Cheng, Carlos
+Contreras-Ibáñez, Natalia Gomez-Sicard, Maria Luz Gonzalez-Gadea, David
+Huepe, et al. 2019. {``The Development of Children's Preferences for
+Equality and Equity Across 13 Individualistic and Collectivist
+Cultures.''} \emph{Developmental Science} 22 (2): e12729.
+\url{https://doi.org/10.1111/desc.12729}.
+
+\bibitem[\citeproctext]{ref-hvidberg_social_2023}
+Hvidberg, Kristoffer B, Claus T Kreiner, and Stefanie Stantcheva. 2023.
+{``Social {Positions} and {Fairness Views} on {Inequality}.''}
+\emph{Review of Economic Studies} 90 (6): 3083--3118.
+\url{https://doi.org/10.1093/restud/rdad019}.
+
+\bibitem[\citeproctext]{ref-iacoviello_collectivism_2019}
+Iacoviello, Vincenzo, and Fabio Lorenzi-Cioldi. 2019. {``Collectivism
+and {Individualism} in {Status Hierarchies}: {Socialization} and {Social
+Identity Explanations}.''} \emph{International Review of Social
+Psychology} 32 (1): 15. \url{https://doi.org/10.5334/irsp.285}.
+
+\bibitem[\citeproctext]{ref-jack_no_2016}
+Jack, Anthony Abraham. 2016. {``({No}) {Harm} in {Asking}: {Class},
+{Acquired Cultural Capital}, and {Academic Engagement} at an {Elite
+University}.''} \emph{Sociology of Education} 89 (1): 1--19.
+\url{https://doi.org/10.1177/0038040715614913}.
+
+\bibitem[\citeproctext]{ref-jasso_how_1999}
+Jasso, Guillermina. 1999. {``How {Much Injustice} Is {There} in the
+{World}? {Two New Justice Indexes}.''} \emph{American Sociological
+Review} 64 (1): 133--68.
+\url{https://doi.org/10.1177/000312249906400110}.
+
+\bibitem[\citeproctext]{ref-jensen_worlds_2008}
+Jensen, Carsten. 2008. {``Worlds of Welfare Services and Transfers.''}
+\emph{Journal of European Social Policy} 18 (2): 151--62.
+\url{https://doi.org/10.1177/0958928707087591}.
+
+\bibitem[\citeproctext]{ref-jonsson_institutional_2015}
+Jonsson, Anna-Carin, and Dennis Beach. 2015. {``Institutional
+Discrimination: {Stereotypes} and Social Reproduction of {`Class'} in
+the {Swedish} Upper-Secondary School.''} \emph{Social Psychology of
+Education} 18 (4): 703--17.
+\url{https://doi.org/10.1007/s11218-014-9279-1}.
+
+\bibitem[\citeproctext]{ref-jost_attitudinal_2000}
+Jost, John T., and Diana Burgess. 2000. {``Attitudinal {Ambivalence} and
+the {Conflict} Between {Group} and {System Justification Motives} in
+{Low Status Groups}.''} \emph{Personality and Social Psychology
+Bulletin} 26 (3): 293--305.
+\url{https://doi.org/10.1177/0146167200265003}.
+
+\bibitem[\citeproctext]{ref-khan_privilege_2011}
+Khan, Shamus. 2011. \emph{Privilege: {The Making} of an {Adolescent
+Elite} at {St}. {Paul}'s {School}}. Princeton, N.J.: Princeton
+University Press.
+
 \bibitem[\citeproctext]{ref-kluegel_beliefs_1987}
 Kluegel, James R., and Eliot R. Smith. 1987. \emph{Beliefs about
-Inequality: Americans' Views of What Is and What Ought to Be}. New York:
-Routledge.
+Inequality: {Americans}' Views of What Is and What Ought to Be}. London
+New York: Routledge, Taylor \& Francis Group.
+
+\bibitem[\citeproctext]{ref-kohn_social_1963}
+Kohn, Melvin L. 1963. {``Social {Class} and {Parent-Child
+Relationships}: {An Interpretation}.''} \emph{American Journal of
+Sociology} 68 (4): 471--80. \url{https://doi.org/10.1086/223403}.
+
+\bibitem[\citeproctext]{ref-kohn_class_1969}
+Kohn, Melvin L., and Carmi Schooler. 1969. {``Class, {Occupation}, and
+{Orientation}.''} \emph{American Sociological Review} 34 (5): 659.
+\url{https://doi.org/10.2307/2092303}.
+
+\bibitem[\citeproctext]{ref-kosse_prosociality_2020}
+Kosse, Fabian, and Michela M. Tincani. 2020. {``Prosociality Predicts
+Labor Market Success Around the World.''} \emph{Nature Communications}
+11 (1): 5298. \url{https://doi.org/10.1038/s41467-020-19007-1}.
 
 \bibitem[\citeproctext]{ref-lampert_meritocratic_2013}
-Lampert, Khen. 2013. \emph{Meritocratic Education and Social
-Worthlessness}. London: Palgrave Macmillan UK.
+Lampert, Khen. 2013. \emph{Meritocratic {Education} and {Social
+Worthlessness}}. London: Palgrave Macmillan UK.
 \url{https://doi.org/10.1057/9781137324894}.
 
+\bibitem[\citeproctext]{ref-lane_market_1986}
+Lane, Robert E. 1986. {``Market {Justice}, {Political Justice}.''}
+\emph{American Political Science Review} 80 (2): 383--402.
+\url{https://doi.org/10.2307/1958264}.
+
 \bibitem[\citeproctext]{ref-lindh_public_2015}
-Lindh, Arvid. 2015. {``Public Opinion Against Markets? Attitudes Towards
-Market Distribution of Social Services -- a Comparison of 17
-Countries.''} \emph{Social Policy \& Administration} 49 (7): 887--910.
-\url{https://doi.org/10.1111/spol.12105}.
+Lindh, Arvid. 2015. {``Public {Opinion} Against {Markets}? {Attitudes}
+Towards {Market Distribution} of {Social Services} -- {A Comparison} of
+17 {Countries}.''} \emph{Social Policy \& Administration} 49 (7):
+887--910. \url{https://doi.org/10.1111/spol.12105}.
+
+\bibitem[\citeproctext]{ref-mcauliffe_developmental_2017}
+McAuliffe, Katherine, Peter R. Blake, Nikolaus Steinbeis, and Felix
+Warneken. 2017. {``The Developmental Foundations of Human Fairness.''}
+\emph{Nature Human Behaviour} 1 (2): 0042.
+\url{https://doi.org/10.1038/s41562-016-0042}.
 
 \bibitem[\citeproctext]{ref-mijs_stratified_2016}
-Mijs, Jonathan. 2016. {``Stratified Failure: Educational Stratification
-and Students' Attributions of Their Mathematics Performance in 24
-Countries.''} \emph{Sociology of Education} 89 (2): 137--53.
-\url{https://doi.org/10.1177/0038040716636434}.
+Mijs, Jonathan. 2016. {``Stratified {Failure}: {Educational
+Stratification} and {Students}' {Attributions} of {Their Mathematics
+Performance} in 24 {Countries}.''} \emph{Sociology of Education} 89 (2):
+137--53. \url{https://doi.org/10.1177/0038040716636434}.
 
 \bibitem[\citeproctext]{ref-mijs_paradox_2021}
-Mijs, Jonathan J B. 2021. {``The Paradox of Inequality: Income
-Inequality and Belief in Meritocracy Go Hand in Hand.''}
-\emph{Socio-Economic Review} 19 (1): 7--35.
-\url{https://doi.org/10.1093/ser/mwy051}.
+---------. 2021. {``The Paradox of Inequality: Income Inequality and
+Belief in Meritocracy Go Hand in Hand.''} \emph{Socio-Economic Review}
+19 (1): 7--35. \url{https://doi.org/10.1093/ser/mwy051}.
+
+\bibitem[\citeproctext]{ref-mikus_children_2020}
+Mikus, Karoline, Nicole Tieben, and Pia S. Schober. 2020. {``Children's
+Conversion of Cultural Capital into Educational Success: The Symbolic
+and Skill-Generating Functions of Cultural Capital.''} \emph{British
+Journal of Sociology of Education} 41 (2): 197--217.
+\url{https://doi.org/10.1080/01425692.2019.1677454}.
+
+\bibitem[\citeproctext]{ref-mishra_subjective_2015}
+Mishra, Sandeep, and R. Nicholas Carleton. 2015. {``Subjective Relative
+Deprivation Is Associated with Poorer Physical and Mental Health.''}
+\emph{Social Science \& Medicine} 147 (December): 144--49.
+\url{https://doi.org/10.1016/j.socscimed.2015.10.030}.
+
+\bibitem[\citeproctext]{ref-molyneux_neoliberal_2008}
+Molyneux, Maxine. 2008. {``The {`{Neoliberal Turn}'} and the {New Social
+Policy} in {Latin America}: {How Neoliberal}, {How New}?''}
+\emph{Development and Change} 39 (5): 775--97.
+\url{https://doi.org/10.1111/j.1467-7660.2008.00505.x}.
+
+\bibitem[\citeproctext]{ref-morris_representing_2022}
+Morris, Jennifer, John Reilly, Sergey Paltsev, Andrei Sokolov, and
+Kenneth Cox. 2022. {``Representing {Socio}-{Economic Uncertainty} in
+{Human System Models}.''} \emph{Earth's Future} 10 (4): e2021EF002239.
+\url{https://doi.org/10.1029/2021EF002239}.
+
+\bibitem[\citeproctext]{ref-pierson_increasing_2000}
+Pierson, Paul. 2000. {``Increasing {Returns}, {Path Dependence}, and the
+{Study} of {Politics}.''} \emph{American Political Science Review} 94
+(2): 251--67. \url{https://doi.org/10.2307/2586011}.
+
+\bibitem[\citeproctext]{ref-power_deprivationprotest_2018}
+Power, Séamus A. 2018. {``The {Deprivation-Protest Paradox}: {How} the
+{Perception} of {Unfair Economic Inequality Leads} to {Civic Unrest}.''}
+\emph{Current Anthropology} 59 (6): 765--89.
+\url{https://doi.org/10.1086/700679}.
+
+\bibitem[\citeproctext]{ref-power_relative_2020}
+Power, Séamus A, Thomas Madsen, and Thomas A Morton. 2020. {``Relative
+Deprivation and Revolt: Current and Future Directions.''} \emph{Current
+Opinion in Psychology} 35 (October): 119--24.
+\url{https://doi.org/10.1016/j.copsyc.2020.06.010}.
 
@@
 
+\bibitem[\citeproctext]{ref-resh_sense_2014}
+Resh, Nura, and Clara Sabbagh. 2014. {``Sense of Justice in School and
+Civic Attitudes.''} \emph{Social Psychology of Education} 17 (1):
+51--72. \url{https://doi.org/10.1007/s11218-013-9240-8}.
+
 \bibitem[\citeproctext]{ref-resh_sense_2017}
-Resh, Nura, and Clara Sabbagh. 2017. {``Sense of Justice in School and
-Civic Behavior.''} \emph{Social Psychology of Education} 20 (2):
-387--409. \url{https://doi.org/10.1007/s11218-017-9375-0}.
+---------. 2017. {``Sense of Justice in School and Civic Behavior.''}
+\emph{Social Psychology of Education} 20 (2): 387--409.
+\url{https://doi.org/10.1007/s11218-017-9375-0}.
+
+\bibitem[\citeproctext]{ref-salamon_marketization_1993}
+Salamon, Lester M. 1993. {``The {Marketization} of {Welfare}: {Changing
+Nonprofit} and {For-Profit Roles} in the {American Welfare State}.''}
+\emph{Social Service Review} 67 (1): 16--39.
+\url{https://doi.org/10.1086/603963}.
+
+\bibitem[\citeproctext]{ref-salgado_inequality_2023}
+Salgado, Mauricio, and Javier Castillo. 2023. {``Inequality and
+{Stratification} in {Latin America}.''} In \emph{The {Oxford Handbook}
+of {Social Stratification}}, edited by Markus Gangl, Lucinda Platt,
+Javier G. Polavieja, and Herman G. van de Werfhorst, 0. 2023: Oxford
+University Press.
 
 \bibitem[\citeproctext]{ref-sandel_tyranny_2020}
-Sandel, Michael J. 2020. \emph{The Tyranny of Merit: What's Become of
+Sandel, Michael J. 2020. \emph{The Tyranny of Merit: {What}'s Become of
 the Common Good?} First edition. New York: {Farrar, Straus and Giroux}.
 
+\bibitem[\citeproctext]{ref-schunk_fairness_2023}
+Schunk, Daniel, and Isabell Zipperle. 2023. {``Fairness and Inequality
+Acceptance in Children and Adolescents: {A} Survey on Behaviors in
+Economic Experiments''} 37 (5): 1715--42.
+\url{https://doi.org/10.1111/joes.12553}.
+
+\bibitem[\citeproctext]{ref-shariff_income_2016}
+Shariff, Azim F., Dylan Wiwad, and Lara B. Aknin. 2016. {``Income
+{Mobility Breeds Tolerance} for {Income Inequality}: {Cross-National}
+and {Experimental Evidence}.''} \emph{Perspectives on Psychological
+Science} 11 (3): 373--80.
+\url{https://doi.org/10.1177/1745691616635596}.
+
+\bibitem[\citeproctext]{ref-sigelman_development_1991}
+Sigelman, Carol K., and Kara A. Waitzman. 1991. {``The {Development} of
+{Distributive Justice Orientations}: {Contextual Influences} on
+{Children}'s {Resource Allocations}.''} \emph{Child Development} 62 (6):
+1367. \url{https://doi.org/10.2307/1130812}.
+
+\bibitem[\citeproctext]{ref-sivesind_changing_2017}
+Sivesind, Karl Henrik. 2017. {``The {Changing Roles} of {For-Profit} and
+{Nonprofit Welfare Provision} in {Norway}, {Sweden}, and {Denmark}.''}
+In \emph{Promoting {Active Citizenship}}, edited by Karl Henrik Sivesind
+and Jo Saglie, 33--74. Cham: Springer International Publishing.
+\url{https://doi.org/10.1007/978-3-319-55381-8_2}.
+
+\bibitem[\citeproctext]{ref-smith_relative_2012}
+Smith, Heather J., Thomas F. Pettigrew, Gina M. Pippin, and Silvana
+Bialosiewicz. 2012. {``Relative {Deprivation}: {A Theoretical} and
+{Meta-Analytic Review}.''} \emph{Personality and Social Psychology
+Review} 16 (3): 203--32. \url{https://doi.org/10.1177/1088868311430825}.
+
+\bibitem[\citeproctext]{ref-soler-martinez_concerns_2023}
+Soler-Martínez, Francisco Miguel, Efraín García-Sánchez, and Guillermo
+B. Willis. 2023. {``Concerns {About Inequality} in {Health}, {Education}
+and {Income Jointly Predict Collective Actions}.''} \emph{Revista
+Latinoamericana de Psicolog{í}a} 55: 99.
+\url{https://doi.org/10.14349/rlp.2023.v55.12}.
+
+\bibitem[\citeproctext]{ref-stoy_worlds_2014}
+Stoy, Volquart. 2014. {``Worlds of {Welfare Services}: {From Discovery}
+to {Exploration}.''} \emph{Social Policy \& Administration} 48 (3):
+343--60. \url{https://doi.org/10.1111/spol.12006}.
+
+\bibitem[\citeproctext]{ref-streeck_citizens_2012}
+Streeck, Wolfgang. 2012. {``Citizens as {Customers}: {Considerations} on
+the {New Politics} of {Consumption}.''} \emph{New Left Review} 76.
+
+\bibitem[\citeproctext]{ref-torche_intergenerational_2014}
+Torche, Florencia. 2014. {``Intergenerational {Mobility} and
+{Inequality}: {The Latin American Case}.''} \emph{Annual Review of
+Sociology} 40 (1): 619--42.
+\url{https://doi.org/10.1146/annurev-soc-071811-145521}.
+
 \bibitem[\citeproctext]{ref-vonhippel_test_2019}
-Von Hippel, Paul, and Caitlin Hamrock. 2019. {``Do Test Score Gaps Grow
-Before, During, or Between the School Years? Measurement Artifacts and
-What We Can Know in Spite of Them.''} \emph{Sociological Science} 6:
-43--80. \url{https://doi.org/10.15195/v6.a3}.
+Von Hippel, Paul, and Caitlin Hamrock. 2019. {``Do {Test Score Gaps
+Grow} Before, During, or Between the {School Years}? {Measurement
+Artifacts} and {What We Can Know} in {Spite} of {Them}.''}
+\emph{Sociological Science} 6: 43--80.
+\url{https://doi.org/10.15195/v6.a3}.
+
+\bibitem[\citeproctext]{ref-weaver_paths_2010}
+Weaver, Kent. 2010. {``Paths and {Forks} or {Chutes} and {Ladders}?:
+{Negative Feedbacks} and {Policy Regime Change}.''} \emph{Journal of
+Public Policy} 30 (2): 137--62.
+\url{https://doi.org/10.1017/S0143814X10000061}.
 
 \bibitem[\citeproctext]{ref-wegener_dominant_1995}
-Wegener, Bernd, and Stefan Liebig. 1995. {``Dominant Ideologies and the
-Variation of Distributive Justice Norms: A Comparison of East and West
-Germany, and the United States.''} In \emph{Social Justice and Political
-Change: Public Opinion in Capitalist and Post-Communist States}, edited
-by James R. Kluegel, David S. Mason, and Bernd Wegener, 239--59. New
-York: Walter de Gruyter.
+Wegener, Bernd, and Stefan Liebig. 1995. {``Dominant {Ideologies} and
+the {Variation} of {Distributive Justice Norms}: {A Comparison} of
+{East} and {West Germany}, and the {United States}.''} In \emph{Social
+{Justice} and {Political Change}: {Public Opinion} in {Capitalist} and
+{Post-Communist States}}, edited by James R. Kluegel, David S. Mason,
+and Bernd Wegener, 239--59. New York: Walter de Gruyter.
 
@@
 Tool for Low Status Students.''} \emph{Frontiers in Psychology} 6.
-\url{https://www.frontiersin.org/articles/10.3389/fpsyg.2015.01053}.
+
+\bibitem[\citeproctext]{ref-wynn_not_2018}
+Wynn, Karen, Paul Bloom, Ashley Jordan, Julia Marshall, and Mark
+Sheskin. 2018. {``Not {Noble Savages After All}: {Limits} to {Early
+Altruism}.''} \emph{Current Directions in Psychological Science} 27 (1):
+3--8. \url{https://doi.org/10.1177/0963721417734875}.
 

### 2024-05-09 02-antecedentes.qmd
update plots and tables analisis

@@
 #| fig-cap: "Average of market justice preferences by country"
+#| fig-asp: 0.65
 
@@
 
-```{r echo=FALSE}
+```{r echo=FALSE, out.width = "600px"}
 #| fig-cap: "Summary of hypotheses"
@@
 
-knitr::include_graphics("input/img/hypotheses.png")
+knitr::include_graphics("input/img/hypotheses.png",
+  auto_pdf = getOption("knitr.graphics.auto_pdf", TRUE))
 ```
-
-##

### 2024-05-09 04-analisis.qmd
update plots and tables analisis

--- 2024-04-23 04-analisis.qmd
@@
 
-```{r}
-#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
+```{r echo=FALSE}
+load("input/data/deservingness.RData")
+load("input/data/soc_effort.RData")
+load("input/data/soc_talent.RData")
+load("input/data/school_effort.RData")
+load("input/data/school_talent.RData")
+pacman::p_load(ggpubr)
 ```
 
-Figure 3 shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram en figure X).
+```{r echo=FALSE}
+#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
+#| fig-asp: 1.22
+#| label: fig-bivariate
+
+plot_soc_effort <- soc_effort %>% ggplot(aes(esfuerzo_soc, mean)) +
+  theme_bw(base_size = 10) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  xlab("Social effort") +
+  ylab("Market justice preferences") +
+  scale_y_continuous(limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_discrete(name = NULL) +
+  scale_shape_discrete(name = NULL)+
+  theme(legend.position = "none")
+
+plot_deservingness <- deservingness %>% ggplot(aes(merito_soc, mean)) +
+  theme_bw(base_size = 10) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  xlab("Deservingness") +
+  ylab("Market justice preferences") +
+  scale_y_continuous(limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_discrete(name = NULL) +
+  scale_shape_discrete(name = NULL)+
+  theme(legend.position = "none")
+
+plot_soc_talent <- soc_talent %>% ggplot(aes(inteligencia_soc, mean)) +
+  theme_bw(base_size = 10) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  xlab("Social talent") +
+  ylab("Market justice preferences") +
+  scale_y_continuous(limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_discrete(name = NULL) +
+  scale_shape_discrete(name = NULL)+
+  theme(legend.position = "none")
+
+plot_school_effort <- school_effort %>% ggplot(aes(esfuerzo_esc, mean)) +
+  theme_bw(base_size = 10) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  xlab("School effort") +
+  ylab("") +
+  scale_y_continuous(limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_discrete(name = NULL) +
+  scale_shape_discrete(name = NULL)+
+  theme(legend.position = "none")
+
+plot_school_talent <- school_talent %>% ggplot(aes(inteligencia_esc, mean)) +
+  theme_bw(base_size = 10) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  xlab("School talent") +
+  ylab("") +
+  scale_y_continuous(limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_discrete(name = NULL) +
+  scale_shape_discrete(name = NULL)
+
+leg <- get_legend(plot_school_talent)
+plot_school_talent <- plot_school_talent + theme(legend.position = "none")
+ggarrange(plot_soc_effort, plot_school_effort, plot_soc_talent, plot_school_talent, plot_deservingness, leg, ncol = 2, nrow = 3)
+```
+
+@fig-bivariate shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram in @fig-hypotheses).
 On the left, we observe the social meritocracy diagrams, while on the right the school meritocracy diagrams are shown.
@@
 Three Cumulative link mixed models were estimated for the ordinal dependent variables of justice in differential access to pensions, education, and health according to income.
-Figure 4 shows the estimation of this regression model containing all the variables used in the study for the three dependent variables separately.
+@fig-odds shows the estimation of this regression model containing all the variables used in the study for the three dependent variables separately.
 However, this figure shows only the effect of meritocracy variables on society and school as independent variables.
@@
 
-```{r}
-#| fig-cap: "Odds-ratios of justification of education, pensions and health by social and school meritocracy"
+```{r echo=FALSE}
+pacman::p_load(lme4, ordinal, sjPlot, texreg)
+load("input/data/data_rec.RData")
+data_pension <- data_rec %>% dplyr::select(-just_educ, -just_salud, -redistribucion)
+data_pension <- na.omit(data_pension)
+data_pension$just_pension_fact <- as_factor(data_pension$just_pension)
+reg5_pension <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_pension)
+data_salud <- data_rec %>% dplyr::select(-just_pension, -just_educ, -redistribucion)
+data_salud <- na.omit(data_salud)
+data_salud$just_salud_fact <- as_factor(data_salud$just_salud)
+reg5_salud <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_salud)
+data_educ <- data_rec %>% dplyr::select(-just_pension, -just_salud, -redistribucion)
+data_educ <- na.omit(data_educ)
+data_educ$just_educ_fact <- as_factor(data_educ$just_educ)
+reg5_educ <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_educ)
 ```
 
-Figure 4 shows that for the three dependent variables of justice in differential access to pensions, education, and health, the trend is similar.
+```{r echo=FALSE}
+#| fig-cap: "Odds-ratios of justification of education, pensions and health by social and school meritocracy"
+#| label: fig-odds
+#| fig-asp: 0.8
+
+plot_models(reg5_salud, reg5_educ, reg5_pension,  
+            rm.terms = c("1|2", "2|3", "3|4", "educacion_recEd Media", "educacion_recEd Tecnica", "educacion_recUniversidad o posgrado", "educacion_recNs/Nr", "libros_recMás de 25", "acc_tec", "internet_rec2", "internet_recNs/Nr", "mean_educ", "cod_depe2Part.
+subvencionado", "cod_depe2Part.
+privado", "cod_grupoMedio bajo", "cod_grupoMedio", "cod_grupoMedio alto", "cod_grupoAlto", "simceMedio", "simceAlto"), #quitar terminos
+            title = "", #quitar titulo
+            show.values = TRUE, #mostrar valor de efectos
+            dot.size = 1.3, #tamaño circulos
+            line.size = 1, #tamaño CI
+            value.size = 3.5, #tamaño valor efectoss
+            spacing = .7, #espacio entre efectos
+            grid = TRUE, # separa en un plot por modelo
+            vline.color = "red", # linea roja en punto neutro (1)
+            axis.labels = rev(c("School talent", 
+                            "School effort", 
+                            "Social talent",
+                            "Social effort", 
+                            "Deservingness"
+                            )), #con rev porque automatico los tira en otro orden
+            show.legend = FALSE,
+            m.labels = c("Health", "Education","Pension")) + # variables dependientes
+  theme_bw(base_size = 10)
+
+```
+
+@fig-odds shows that for the three dependent variables of justice in differential access to pensions, education, and health, the trend is similar.
 In the context of school meritocracy, the effects are mixed.
@@
 
-```{r}
-#| tbl-cap: "Individual effects of multilevel regression models"
+```{r echo=FALSE, include=FALSE}
+load("input/data/data_index.RData")
+reg0_bienestar <- lmer(bienestar ~ 1 + (1 | mrbd), data=data_bienestar)
+reg1_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_bienestar)
+reg2_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_bienestar)
+reg3_bienestar <- lmer(bienestar ~ 1 + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_bienestar)
+reg4_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_bienestar)
+
+test01<- anova(reg0_bienestar,reg1_bienestar,test = "Chisq")
+test02<- anova(reg0_bienestar,reg2_bienestar,test = "Chisq")
+test03<- anova(reg0_bienestar,reg3_bienestar,test = "Chisq")
+test04<- anova(reg0_bienestar,reg4_bienestar,test = "Chisq")
+
+test.pvalues1<- test01$`Pr(>Chisq)`[2]
+test.pvalues2<- test02$`Pr(>Chisq)`[2]
+test.pvalues3<- test03$`Pr(>Chisq)`[2]
+test.pvalues4<- test04$`Pr(>Chisq)`[2]
+
+deviance1 <- test01$deviance[2]
+deviance2 <- test02$deviance[2]
+deviance3 <- test03$deviance[2]
+deviance4 <- test04$deviance[2]
 ```
 
-Table 4 shows the results of the multilevel estimation for justice market preferences.
+```{r echo=FALSE, results='asis'}
+#| tbl-cap: "Individual effects of multilevel regression models"
+#| label: tbl-individual-reg
+
+texreg::texreg(list(reg1_bienestar, reg2_bienestar),
+        custom.model.names = c("Model 1",
+                               "Model 2"),
+        custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
+        omit.coef = "(internet_rec)",
+        caption = "",
+        custom.coef.names = c("Intercept",  
+                              "School talent",
+                              "School effort", 
+                              "Social talent",  
+                              "Social effort",
+                              "Deservingness",
+                              "Secondary", 
+                              "Higher tec.", 
+                              "University or posgraduate", 
+                              "Missing", 
+                              "More than 25 books (Ref. Less than 25)", 
+                              "Technology access"),
+        groups = list("Parental education (Ref.= 8th grade or less)"=7:10),
+        custom.gof.rows=list("Deviance" = c(deviance1,
+                                            deviance2),
+                             "Deviance Test (p)" = c(test.pvalues1,
+                                                       test.pvalues2))
+        )
+
+```
+
+@tbl-individual-reg shows the results of the multilevel estimation for justice market preferences.
 For this variable, the intraclass correlation obtained shows that the variation between schools corresponds to 4% of the variation of students’ preferences.
@@
 
-```{r}
+```{r echo=FALSE, results='asis'}
 #| tbl-cap: "Contextual effects of multilevel regression models"
+#| label: tbl-contextual-reg
+
+texreg::texreg(list(reg3_bienestar, reg4_bienestar),
+        custom.model.names = c("Model 1",
+                               "Model 2"),
+        custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
+        omit.coef = c("inteligencia_esc|esfuerzo_esc|inteligencia_soc|esfuerzo_soc|merito_soc|educacion_rec|libros_rec|acc_tec|internet_rec"),
+        caption = "",
+        custom.coef.names = c("Intercept", 
+                              "Prop. university level at school",
+                              "Subsidized private", 
+                              "Private", 
+                              "SES Medium low", 
+                              "SES Medium", 
+                              "SES Medium high", 
+                              "SES High", 
+                              "Simce Medium", 
+                              "Simce High"),
+        groups = list("Administrative dependence (Ref.= Public)"=3:4,
+                      "Socioeconomic level (Ref.= Low)"=5:8,
+                      "Achievement score (Ref.= Low)"=9:10),
+        custom.gof.rows=list("Deviance" = c(deviance3,
+                                            deviance4),
+                             "Deviance Test (p)" = c(test.pvalues3,
+                                                     test.pvalues4))
+        )
+
 ```
@@
 
-```{r}
+```{r echo=FALSE}
 #| tbl-cap: "Interactions effects"
+#| label: tbl-interact
+
+pacman::p_load(kableExtra)
+load("input/data/interact.RData")
+kbl(collapse_rows,
+    align = "lccccc", booktabs = T,
+    col.names =  c("Variable", 
+                   "School talent", 
+                   "School effort",
+                   "Social talent",
+                   "Social effort", 
+                   "Deservingness")) %>%
+  add_header_above(c(" " = 1,
+                     "Market Justice Preferences"=5), 
+                   bold = T)%>%  
+  kable_styling(latex_options = c("scale_down","HOLD_position"),
+                full_width = F) %>%
+  column_spec(1, bold = T, width = "4cm") %>%
+  column_spec(column = 2:6,width = "1.5cm") %>%
+  collapse_rows(columns = 1:2, valign = "middle") %>% 
+  footnote(general = "***p < 0.001, **p < 0.01, *p < 0.05",
+           footnote_as_chunk = T,general_title = "Note: ")
+
 ```
 
-The interaction terms in Table 6 suggest that students’ family background moderates the relationship between their meritocratic perceptions in Chile and their justification of inequality.
+The interaction terms in @tbl-interact suggest that students’ family background moderates the relationship between their meritocratic perceptions in Chile and their justification of inequality.
 The direction of these effects confirms our initial prediction.
@@
 That is, lower-status students (measured by parental education) justify more inequality when they adhere more strongly to deservingnessmeritocratic perceptions in Chile.
-We depict this moderating effect in Figure 1.
 This result confirms the enlightenment thesis (see above).
@@
 Thus, in line with our Hypothesis 5, high-status schools justify less inequality when, on average, their students have a greater perception of meritocracy in Chile, as measured by the three indicators we used (i.e., effort, deservingness, and talent).
-Finally, as we stated in Hypothesis 7, lowhigh-achieving schools justify more inequality when their students have, on average, stronger meritocratic beliefs in Chile.
-In Figure 25 we depict this moderating effect of school achievement for the relationship between deservingness and justification of inequality.
+Finally, as we stated in Hypothesis 7, low-achieving schools justify more inequality when their students have, on average, stronger meritocratic beliefs in Chile.
+In @fig-interaction we depict this moderating effect of school achievement for the relationship between deservingness and justification of inequality.
 We did not observe these moderating effects at the school level for students’ meritocratic beliefs in the school (i.e., the idea that effort or talent are important to get good grades).
 
-```{r}
+```{r echo=FALSE, include=FALSE}
+load("input/data/data_rec.RData")
+data_rec <- data_rec %>% rowwise() %>%  mutate(educacion_rec2 = case_when(educacion_rec=="Ed Basica"~"Low level",
+                                                 educacion_rec=="Ed Media"~"Low level",
+                                                 educacion_rec=="Ed Tecnica"~"Low level",
+                                                 educacion_rec=="Universidad o posgrado"~"University or Postgraduate",
+                                                 educacion_rec=="Ns/Nr"~"Missing"
+                                                 ))
+data_rec$educacion_rec2 <- factor(data_rec$educacion_rec2, levels = c("Low level", "University or Postgraduate", "Missing"))
+
+data_rec <- data_rec %>% rowwise() %>%  mutate(cod_grupo2  = case_when(cod_grupo=="Bajo"~"Bajo",
+                                                                       cod_grupo=="Medio bajo"~"Bajo",
+                                                                       cod_grupo=="Medio"~"Bajo",
+                                                                       cod_grupo=="Medio alto"~"Alto",
+                                                                       cod_grupo=="Alto"~"Alto"))
+data_rec$cod_grupo2 <- factor(data_rec$cod_grupo2, levels = c("Bajo", "Alto"))
+data_rec$simce <- factor(data_rec$simce, labels = c("Low", "Medium", "High"))
+
+data_bienestar <- na.omit(data_rec)
+#data_bienestar <- data_bienestar %>% mutate(pension_rec = 5-just_pension,
+#                                            educ_rec = 5-just_educ,
+#                                            salud_rec = 5-just_salud)
+data_bienestar <- data_bienestar %>% rowwise() %>% dplyr::mutate(bienestar = mean(c(just_educ, just_salud, just_pension), na.omit=TRUE))
+data_bienestar$bienestar <- round(data_bienestar$bienestar, 2)
+
+reg26.2_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + merito_soc*simce + internet_rec + (1 + simce | mrbd), data=data_bienestar)
+
+plot_interact<- plot_model(reg26.2_bienestar, type = "int", mdrt.values="minmax", title = "", colors = c("red", "white", "blue")) +
+  ylim(1,4)+
+  labs(x = "Deservingness", y="Market Justice Preferences", colour = "School Simce")+
+  theme_bw()
+```
+
+```{r echo=FALSE, warning=FALSE}
 #| fig-cap: "Interaction between deservingness and justification of inequality by school SIMCE achievement"
+#| label: fig-interaction
+
+plot_interact
+
 ```

### 2024-05-09 paper.tex
update plots and tables analisis

@@
 
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 \KOMAoption{captions}{tableheading}
@@
 
-\includegraphics{input/img/hypotheses.png}
+\includegraphics[width=6.25in,height=\textheight]{input/img/hypotheses.png}
 
@@
 \end{figure}%
-
-\subsection{}\label{section}
 
@@
 \end{minipage} \\
-Inequality justification index &
-\begin{minipage}[t]{\linewidth}\raggedright
+Market Justice Preferences & \begin{minipage}[t]{\linewidth}\raggedright
 Mean (sd) : 2.2 (0.9)\\
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics{paper_files/figure-pdf/unnamed-chunk-11-1.pdf}
+\begin{figure}
+
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-bivariate-1.pdf}
 
@@
 
-\caption{Market justice preferences in education, health and pensions by
-social and school meritocracy}
+\caption{\label{fig-bivariate}Market justice preferences in education,
+health and pensions by social and school meritocracy}
 
@@
 
-Figure 3 shows a series of graphs depicting the association between the
-variables of market justice preferences - in education, health, and
-pensions - and the variables of meritocratic perception at school
-(effort and talent) and in society (effort, talent, and deservingness)
-(see conceptual diagram en figure X). On the left, we observe the social
+Figure~\ref{fig-bivariate} shows a series of graphs depicting the
+association between the variables of market justice preferences - in
+education, health, and pensions - and the variables of meritocratic
+perception at school (effort and talent) and in society (effort, talent,
+and deservingness) (see conceptual diagram in
+Figure~\ref{fig-hypotheses}). On the left, we observe the social
 meritocracy diagrams, while on the right the school meritocracy diagrams
@@
 dependent variables of justice in differential access to pensions,
-education, and health according to income. Figure 4 shows the estimation
-of this regression model containing all the variables used in the study
-for the three dependent variables separately. However, this figure shows
-only the effect of meritocracy variables on society and school as
-independent variables. Complete models are available in the appendix.
-
-Figure 4 shows that for the three dependent variables of justice in
-differential access to pensions, education, and health, the trend is
-similar. In the context of school meritocracy, the effects are mixed. As
-school talent increases, the justification for differentiated access to
-pensions, education, and health increases; on the contrary, as the
-school effort variable increases, the justification for differentiated
-access to pensions, education, and health decreases, keeping the rest of
-the independent variables constant. Regarding the variables of social
-meritocracy, the three variables of talent, effort, and deservingness
-show that as these increase, the justification for differentiated access
-to pensions, education, and health also increases.
+education, and health according to income. Figure~\ref{fig-odds} shows
+the estimation of this regression model containing all the variables
+used in the study for the three dependent variables separately. However,
+this figure shows only the effect of meritocracy variables on society
+and school as independent variables. Complete models are available in
+the appendix.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-odds-1.pdf}
+
+}
+
+\caption{\label{fig-odds}Odds-ratios of justification of education,
+pensions and health by social and school meritocracy}
+
+\end{figure}%
+
+Figure~\ref{fig-odds} shows that for the three dependent variables of
+justice in differential access to pensions, education, and health, the
+trend is similar. In the context of school meritocracy, the effects are
+mixed. As school talent increases, the justification for differentiated
+access to pensions, education, and health increases; on the contrary, as
+the school effort variable increases, the justification for
+differentiated access to pensions, education, and health decreases,
+keeping the rest of the independent variables constant. Regarding the
+variables of social meritocracy, the three variables of talent, effort,
+and deservingness show that as these increase, the justification for
+differentiated access to pensions, education, and health also increases.
 
@@
 
-Table 4 shows the results of the multilevel estimation for justice
-market preferences. For this variable, the intraclass correlation
-obtained shows that the variation between schools corresponds to 4\% of
-the variation of students' preferences. This means that there is low
-variance between schools and therefore limits the possibilities of
-finding effects at the aggregate level.
+\begin{table}
+
+\caption{\label{tbl-individual-reg}Individual effects of multilevel
+regression models}
+
+\centering{
+
+\begin{center}
+\begin{tabular}{l c c}
+\hline
+ & Model 1 & Model 2 \\
+\hline
+Intercept                                    & $1.34^{***}$  & $1.37^{***}$  \\
+                                             & $(0.09)$      & $(0.10)$      \\
+School talent                                & $0.11^{***}$  & $0.10^{***}$  \\
+                                             & $(0.01)$      & $(0.01)$      \\
+School effort                                & $-0.12^{***}$ & $-0.12^{***}$ \\
+                                             & $(0.02)$      & $(0.02)$      \\
+Social talent                                & $0.07^{***}$  & $0.07^{***}$  \\
+                                             & $(0.02)$      & $(0.02)$      \\
+Social effort                                & $0.08^{***}$  & $0.08^{***}$  \\
+                                             & $(0.02)$      & $(0.02)$      \\
+Deservingness                                & $0.20^{***}$  & $0.20^{***}$  \\
+                                             & $(0.02)$      & $(0.02)$      \\
+Parental education (Ref.= 8th grade or less) &               &               \\
+                                             &               &               \\
+\quad Secondary                              &               & $-0.07$       \\
+                                             &               & $(0.05)$      \\
+\quad Higher tec.                            &               & $-0.10$       \\
+                                             &               & $(0.05)$      \\
+\quad University or posgraduate              &               & $-0.02$       \\
+                                             &               & $(0.05)$      \\
+\quad Missing                                &               & $-0.00$       \\
+                                             &               & $(0.05)$      \\
+More than 25 books (Ref. Less than 25)       &               & $-0.05$       \\
+                                             &               & $(0.03)$      \\
+Technology access                            &               & $0.01$        \\
+                                             &               & $(0.01)$      \\
+\hline
+Deviance                                     & $12368.85$    & $12355.60$    \\
+Deviance Test (p)                            & $0.00$        & $0.00$        \\
+AIC                                          & $12422.86$    & $12468.36$    \\
+BIC                                          & $12475.07$    & $12572.78$    \\
+Log Likelihood                               & $-6203.43$    & $-6218.18$    \\
+Num. obs.                                    & $5047$        & $5047$        \\
+Num. groups: mrbd                            & $231$         & $231$         \\
+Var: mrbd (Intercept)                        & $0.02$        & $0.02$        \\
+Var: Residual                                & $0.66$        & $0.66$        \\
+\hline
+\multicolumn{3}{l}{\scriptsize{*** p < 0.001; ** p < 0.01; * p < 0.05}}
+\end{tabular}
+\caption{}
+\label{table:coefficients}
+\end{center}
+
+}
+
+\end{table}%
+
+Table~\ref{tbl-individual-reg} shows the results of the multilevel
+estimation for justice market preferences. For this variable, the
+intraclass correlation obtained shows that the variation between schools
+corresponds to 4\% of the variation of students' preferences. This means
+that there is low variance between schools and therefore limits the
+possibilities of finding effects at the aggregate level.
 
@@
 
+\begin{table}
+
+\caption{\label{tbl-contextual-reg}Contextual effects of multilevel
+regression models}
+
+\centering{
+
+\begin{center}
+\begin{tabular}{l c c}
+\hline
+ & Model 1 & Model 2 \\
+\hline
+Intercept                                & $2.29^{***}$  & $1.42^{***}$  \\
+                                         & $(0.04)$      & $(0.11)$      \\
+Prop. university level at school         & $0.25$        & $0.18$        \\
+                                         & $(0.24)$      & $(0.22)$      \\
+Administrative dependence (Ref.= Public) &               &               \\
+                                         &               &               \\
+\quad Subsidized private                 & $0.01$        & $0.00$        \\
+                                         & $(0.04)$      & $(0.04)$      \\
+\quad Private                            & $-0.07$       & $-0.08$       \\
+                                         & $(0.14)$      & $(0.13)$      \\
+Socioeconomic level (Ref.= Low)          &               &               \\
+                                         &               &               \\
+\quad SES Medium low                     & $0.00$        & $0.02$        \\
+                                         & $(0.05)$      & $(0.05)$      \\
+\quad SES Medium                         & $-0.13$       & $-0.03$       \\
+                                         & $(0.07)$      & $(0.06)$      \\
+\quad SES Medium high                    & $-0.23^{*}$   & $-0.08$       \\
+                                         & $(0.09)$      & $(0.09)$      \\
+\quad SES High                           & $0.09$        & $0.25$        \\
+                                         & $(0.17)$      & $(0.15)$      \\
+Achievement score (Ref.= Low)            &               &               \\
+                                         &               &               \\
+\quad Simce Medium                       & $-0.09^{*}$   & $-0.10^{**}$  \\
+                                         & $(0.04)$      & $(0.04)$      \\
+\quad Simce High                         & $-0.27^{***}$ & $-0.24^{***}$ \\
+                                         & $(0.05)$      & $(0.04)$      \\
+\hline
+Deviance                                 & $12889.80$    & $12306.45$    \\
+Deviance Test (p)                        & $0.00$        & $0.00$        \\
+AIC                                      & $12954.69$    & $12473.31$    \\
+BIC                                      & $13033.01$    & $12636.48$    \\
+Log Likelihood                           & $-6465.35$    & $-6211.66$    \\
+Num. obs.                                & $5047$        & $5047$        \\
+Num. groups: mrbd                        & $231$         & $231$         \\
+Var: mrbd (Intercept)                    & $0.02$        & $0.01$        \\
+Var: Residual                            & $0.74$        & $0.66$        \\
+\hline
+\multicolumn{3}{l}{\scriptsize{*** p < 0.001; ** p < 0.01; * p < 0.05}}
+\end{tabular}
+\caption{}
+\label{table:coefficients}
+\end{center}
+
+}
+
+\end{table}%
+
 \subsubsection{Interactions effects}\label{interactions-effects}
 
-The interaction terms in Table 6 suggest that students' family
-background moderates the relationship between their meritocratic
+\begin{table}
+
+\caption{\label{tbl-interact}Interactions effects}
+
+\centering{
+
+[H]
+\centering
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
+\toprule
+\multicolumn{1}{c}{\textbf{ }} & \multicolumn{5}{c}{\textbf{Market Justice Preferences}} \\
+\cmidrule(l{3pt}r{3pt}){2-6}
+Variable & School talent & School effort & Social talent & Social effort & Deservingness\\
+\midrule
+\textbf{University or Postgraduate} & -0.034 & -0.012 & -0.073. & -0.079* & -0.118**\\
+\cmidrule{1-6}
+\textbf{More than 25 books} & -0.015 & -0.046 & -0.029 & -0.081** & -0.035\\
+\cmidrule{1-6}
+\textbf{School SES high} & 0.004 & 0.001 & -0.057. & -0.063* & -0.08**\\
+\cmidrule{1-6}
+\textbf{Prop. university level} & 0.008 & -0.021 & -0.123 & -0.171* & -0.162.\\
+\cmidrule{1-6}
+\textbf{Simce Medium} & 0.001 & 0.032 & -0.037 & -0.041 & -0.082*\\
+\cmidrule{1-6}
+\textbf{Simce High} & -0.016 & 0.008 & -0.043 & -0.063. & -0.087*\\
+\bottomrule
+\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: } ***p < 0.001, **p < 0.01, *p < 0.05}\\
+\end{tabular}}
+
+}
+
+\end{table}%
+
+The interaction terms in Table~\ref{tbl-interact} suggest that students'
+family background moderates the relationship between their meritocratic
 perceptions in Chile and their justification of inequality. The
@@
 inequality when they adhere more strongly to deservingnessmeritocratic
-perceptions in Chile. We depict this moderating effect in Figure 1. This
-result confirms the enlightenment thesis (see above). The same
-moderating effect is observed for students' family cultural capital: the
-relationship between effort (but not deservingness nor social talent)
-and justification of inequality becomes less positive for students with
-more than 25 books at home. Interestingly, family cultural capital also
-moderates the relationship between meritocratic beliefs at school (as
-measured by students' belief that effort is important to get good
-grades) and justification of inequality.
+perceptions in Chile. This result confirms the enlightenment thesis (see
+above). The same moderating effect is observed for students' family
+cultural capital: the relationship between effort (but not deservingness
+nor social talent) and justification of inequality becomes less positive
+for students with more than 25 books at home. Interestingly, family
+cultural capital also moderates the relationship between meritocratic
+beliefs at school (as measured by students' belief that effort is
+important to get good grades) and justification of inequality.
 
@@
 we used (i.e., effort, deservingness, and talent). Finally, as we stated
-in Hypothesis 7, lowhigh-achieving schools justify more inequality when
+in Hypothesis 7, low-achieving schools justify more inequality when
 their students have, on average, stronger meritocratic beliefs in Chile.
-In Figure 25 we depict this moderating effect of school achievement for
-the relationship between deservingness and justification of inequality.
-We did not observe these moderating effects at the school level for
-students' meritocratic beliefs in the school (i.e., the idea that effort
-or talent are important to get good grades).
+In Figure~\ref{fig-interaction} we depict this moderating effect of
+school achievement for the relationship between deservingness and
+justification of inequality. We did not observe these moderating effects
+at the school level for students' meritocratic beliefs in the school
+(i.e., the idea that effort or talent are important to get good grades).
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-interaction-1.pdf}
+
+}
+
+\caption{\label{fig-interaction}Interaction between deservingness and
+justification of inequality by school SIMCE achievement}
+
+\end{figure}%
 

### 2024-05-15 03-methods.qmd
update plots and tables analisis
Primer cambios
Cambios en method.qmd y analisis.qmd

@@
 
-## Methods
+## Data, variables and method
 
@@
 
-The main data source for the analysis is the First Study of Citizenship Education in Chile, carried out by the Education Quality Agency of the Ministry of Education.
-The application date was November 9, 2017.
-The target population of this study is Eighth-grade students from 242 schools.
-In the data, there are 8,589 students and 6,770 parents.
-This database was analyzed with the R package “ResponsePatterns” to detect possible repetitive and “careless” response patterns and thus contribute to a higher quality of research data (Gottfried, et al 2022).
-Responses from 171 students and 79 parents were removed, which, when merged, gave a total of 6,270 valid cases.
-
-The analysis of school variables includes data from the Ministry of Education's SIMCE 2017 database.
-This database contains information at the school level, such as the administrative dependency, its socioeconomic classification, and the achievement scores obtained in the mathematics and language census tests.
-It is available for free use on the MINEDUC \[web page\].
-
-After eliminating missing cases, the final sample used in the analysis was based on 5,047 students and parents of 231 schools for the dependent variable of access to social services.
+The main data source corresponds to the First Study of Civic Education in Chile, conducted by the Agency for Quality Education of the Ministry of Education in 2017.
+This database is based on the administration of a civic knowledge test to students and questionnaires to various members of the educational community.
+The target population consists of 8th-grade students in 242 schools across the country.
+Initially, the database comprises 8,589 student observations and 6,770 parent observations.
+However, considering the survey's unit of analysis and to ensure higher data quality, 171 student cases and 79 parent cases exhibiting repetitive and careless response patterns were removed (Gottfried et al., 2022).
+Additionally, data from the System for Measuring the Quality of Education (SIMCE) conducted by the Ministry of Education for the year 2017 are employed, providing school-level information such as administrative dependence, socioeconomic classification, and results obtained in standardized mathematics and language tests.
+After variable processing and removal of missing cases, the final database for this study includes a two-level stratified sample, composed of 5,047 students and parents (level 1) nested within 231 schools (level 2).
 
@@
 
-**Dependent variables**
+***Individual level***
 
-This study has three dependent variables related to the justification of social inequality in specific policy domains.
-The first asks whether access to social services should be conditional on income, i,.e., “It is just that in Chile people who can pay have a better education for their children”.
-Students rated their preferences using the following responses: “strongly disagree”, “Disagree”, “Agree”, and ”strongly agree”.
-An average index is built with these items (Cronbach's Alpha = 0,86).
-(Apéndice: items en español y su correspondiente traducción al inglés)
-
-@tbl-desc-dependientes shows the items used, their response categories, and their frequencies.
+**Justification of Inequality**: The dependent variable in this study is the justification of social inequality in specific policy domains, namely health, education, and pensions.
+This construct is measured through three variables addressing the degree of justification regarding whether access to social services in these domains should be income conditional.
+Justification of inequality in health is measured by the item "Is it fair in Chile that people with higher incomes can access better healthcare than those with lower incomes," while in education, it is measured using the statement "Is it fair in Chile that people who can afford it have better education for their children." In the case of pensions, it is measured through the item "Is it fair in Chile that people with higher incomes can have better pensions than those with lower incomes." In all cases, respondents declare their preferences on a Likert scale ranging from "completely disagree" (1) to "completely agree" (4).
+Additionally, we include a summarized indicator of preferences for market justice (REF), measured by an average index across these items (α = 0.86), with values ranging from 1 to 4, where higher values represent greater preferences for market justice (see @tbl-desc-dependientes).
 
 ```{r echo=FALSE}
-pacman::p_load(summarytools, dplyr)
+pacman::p_load(summarytools, dplyr, sjlabelled)
 load("input/data/dependientes.RData")
-
+dependientes$bienestar <- set_label(dependientes$bienestar, "Market Justice Preferences")
 ```
@@
 
-**Independent variables**
+**Perception of Meritocracy**: The main independent variable refers to the perception of meritocracy, operationalized through five items addressing the perception of rewards based on talent and intelligence at both the school and societal levels.
+At the school level, students respond to whether "Intelligence is important for getting good grades" and "Effort is important for getting good grades".
+At the societal level, students respond to the following questions: "In Chile, people are rewarded for their effort", "In Chile, people get what they deserve" and "In Chile, people are rewarded for their intelligence and skills".
+Each item was answered on a four-point Likert scale ranging from "completely disagree" (1) to "completely agree" (4).
 
-For the primary independent variable, the perception of meritocracy, five items address the perception of rewards according to effort and intelligence at the school and societal levels.
-At the school level, students answer whether “Intelligence is important to get good grades” and “Effort is important to get good grades”.
-At the societal level, students respond to the following questions: “In Chile, people are rewarded for their efforts”, “In Chile, people get what they deserve”, and “In Chile, people are rewarded for their intelligence and skills”.
-Each item was answered on a four-point scale ranging from “Strongly disagree” to “Strongly agree”.
-
-The rest of the independent variables are divided into individual and school levels.
-At the individual level, family socioeconomic status was measured by the parents' highest educational level and the number of books at home.
-Likewise, an index of access to technology includes the number of computers, tablets, and cell phones at home, as well as whether there is an Internet connection.
-@tbl-desc-independent shows the items used, their response categories, and their frequency.
+The analysis includes several individual control variables that have been identified as influential in recent research.
+These variables encompass family socioeconomic status, assessed by the highest level of parental education, the number of books in the household, and an index measuring access to technologies, all aimed at mitigating compositional effects (REFS).
+@tbl-desc-independent shows the independent variables used, their response categories and their frequencies.
 
@@
 
-The school-level variables are the administrative dependency of the school, the socioeconomic classification made by the Ministry of Education, the level of performance in the SIMCE test of the school, and the proportion of parents with university or postgraduate degrees.
-@tbl-desc-school shows the items used, response categories, and frequency.
+***Contextual level***
+
+This study examines four variables at the school level.
+Firstly, we assess the administrative dependence of schools, categorized as "Public," "Subsidized Private," or "Private." Secondly, we consider the socioeconomic classification of schools, determined by the Ministry of Education and measured on an ordinal scale with five categories ranging from "low" (1) to "high" (5).
+Thirdly, we evaluate the performance level of schools in SIMCE standardized tests, categorized as "low," "medium," or "high." Lastly, we analyze the proportion of parents with university or postgraduate education level at each school.
+The items used, response categories, and their frequencies are detailed in @tbl-desc-school.
 
@@
 
-The data has a hierarchical structure of students nested in schools, so the model estimation is performed in a multilevel (random effects) framework.
-This modeling approach lets us correctly estimate individual and contextual effects in a single model.
-We estimate cumulative link mixed models for the ordinal dependent variables, whereas we use linear mixed effects models for the average index of inequality justification.
+Given the hierarchical structure of the data, with students nested within schools, model estimation is performed within a multilevel framework.
+These models are appropriate for capturing both individual and contextual effects in a single analysis, allowing for the estimation of fixed effects between groups and random effects that vary from one group to another [@bell_fixed_2019; @hox_multilevel_2010].
+Cumulative link mixed models were employed for the ordinal dependent variables, while linear mixed-effects models were applied for the average market justice preference index.
 

### 2024-05-15 04-analisis.qmd
Primer cambios
Cambios en method.qmd y analisis.qmd

--- 2024-04-23 04-analisis.qmd
@@
 
-```{r}
-#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
+```{r packages, echo=FALSE}
+pacman::p_load(tidyverse,
+               ggpubr, 
+               lme4, 
+               ordinal,
+               sjPlot, 
+               texreg,
+               kableExtra)
+
+options(scipen=999)
+rm(list = ls())
+
 ```
 
-Figure 3 shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram en figure X).
+```{r data, echo=FALSE}
+
+load("input/data/data_index.RData")
+load("input/data/deservingness.RData")
+load("input/data/soc_effort.RData")
+load("input/data/soc_talent.RData")
+load("input/data/school_effort.RData")
+load("input/data/school_talent.RData")
+load("input/data/data_rec.RData")
+
+```
+
+```{r functions, echo=FALSE}
+
+miles <- function(x) {
+  format(round(as.numeric(x),0), big.mark = ".")
+}
+
+decimales <- function(x) {
+  format(round(as.numeric(x), 2), decimal.mark = ",")
+}
+
+# set theme
+
+#windowsFonts(`Times New Roman` = windowsFont("Times New Roman"))
+
+theme_nice <- function() {
+  theme_bw() +
+    theme(text = element_text(family = "serif"))
+
+}
+
+theme_set(theme_nice())
+
+```
+
+@fig-bivariate shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram in @fig-hypotheses).
 On the left, we observe the social meritocracy diagrams, while on the right the school meritocracy diagrams are shown.
@@
 
-### Cumulative link mixed models for the Justification of inequality in education, health and pensions
+```{r echo=FALSE}
+#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
+#| fig-asp: 1.22
+#| label: fig-bivariate
 
-Three Cumulative link mixed models were estimated for the ordinal dependent variables of justice in differential access to pensions, education, and health according to income.
-Figure 4 shows the estimation of this regression model containing all the variables used in the study for the three dependent variables separately.
-However, this figure shows only the effect of meritocracy variables on society and school as independent variables.
-Complete models are available in the appendix.
+plot_soc_effort <- soc_effort %>% 
+  mutate(justice = str_to_title(justice)) %>% 
+  ggplot(aes(x = esfuerzo_soc, y = mean)) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_grey() +
+  labs(x = "Social effort",
+       y = "Market justice preferences") +
+  theme(legend.position = "none",
+        text = element_text(size = 11))
 
-```{r}
-#| fig-cap: "Odds-ratios of justification of education, pensions and health by social and school meritocracy"
+plot_deservingness <- deservingness %>% 
+  mutate(justice = str_to_title(justice)) %>% 
+  ggplot(aes(x = merito_soc, y = mean)) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_grey() +
+  labs(x = "Deservingness",
+       y = "Market justice preferences") +
+  theme(legend.position = "none",
+        text = element_text(size = 11))
+
+plot_soc_talent <- soc_talent %>% 
+  mutate(justice = str_to_title(justice)) %>% 
+  ggplot(aes(x = inteligencia_soc, y = mean)) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_grey() +
+  labs(x = "Social talent",
+       y = "Market justice preferences") +
+  theme(legend.position = "none",
+        text = element_text(size = 11))
+
+plot_school_effort <- school_effort %>% 
+  mutate(justice = str_to_title(justice)) %>% 
+  ggplot(aes(x = esfuerzo_esc, y = mean)) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_grey() +
+  labs(x = "School effort",
+       y = "") +
+  theme(legend.position = "none",
+        text = element_text(size = 11))
+
+plot_school_talent <- school_talent %>% 
+  mutate(justice = str_to_title(justice)) %>% 
+  ggplot(aes(x = inteligencia_esc, y = mean)) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_grey() +
+  labs(x = "School talent",
+       y = "",
+       color = NULL,
+       shape = NULL) +
+  theme(text = element_text(size = 11))
+
+leg <- get_legend(plot_school_talent)
+
+plot_school_talent <- plot_school_talent + theme(legend.position = "none")
+
+ggarrange(plot_soc_effort, plot_school_effort, plot_soc_talent,
+          plot_school_talent, plot_deservingness, leg, ncol = 2, nrow = 3)
 ```
 
-Figure 4 shows that for the three dependent variables of justice in differential access to pensions, education, and health, the trend is similar.
-In the context of school meritocracy, the effects are mixed.
-As school talent increases, the justification for differentiated access to pensions, education, and health increases; on the contrary, as the school effort variable increases, the justification for differentiated access to pensions, education, and health decreases, keeping the rest of the independent variables constant.
-Regarding the variables of social meritocracy, the three variables of talent, effort, and deservingness show that as these increase, the justification for differentiated access to pensions, education, and health also increases.
+### Multivariate results
 
-### Multilevel regression models for market justice preferences
+The results of the cumulative link mixed models for the ordinal dependent variables of justice in differential access to education, health, and pensions by income are presented in @fig-odds.
+This figure illustrates the effect of meritocratic variables on society and school as independent variables; however, the models incorporate all variables utilized in the study for each of the three dependent variables separately (complete models are available in the [Appendix](#appendix)).
 
-```{r}
-#| tbl-cap: "Individual effects of multilevel regression models"
+```{r echo=FALSE}
+
+data_pension <- data_rec %>% dplyr::select(-just_educ, -just_salud, -redistribucion)
+data_pension <- na.omit(data_pension)
+data_pension$just_pension_fact <- as_factor(data_pension$just_pension)
+reg5_pension <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_pension)
+data_salud <- data_rec %>% dplyr::select(-just_pension, -just_educ, -redistribucion)
+data_salud <- na.omit(data_salud)
+data_salud$just_salud_fact <- as_factor(data_salud$just_salud)
+reg5_salud <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_salud)
+data_educ <- data_rec %>% dplyr::select(-just_pension, -just_salud, -redistribucion)
+data_educ <- na.omit(data_educ)
+data_educ$just_educ_fact <- as_factor(data_educ$just_educ)
+reg5_educ <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_educ)
 ```
 
-Table 4 shows the results of the multilevel estimation for justice market preferences.
-For this variable, the intraclass correlation obtained shows that the variation between schools corresponds to 4% of the variation of students’ preferences.
-This means that there is low variance between schools and therefore limits the possibilities of finding effects at the aggregate level.
+```{r echo=FALSE}
+#| fig-cap: "Odds-ratios of justification of pensions, education and health by social and school meritocracy"
+#| label: fig-odds
+#| fig-asp: 0.8
+
+plot_models(reg5_salud, reg5_educ, reg5_pension,  
+            rm.terms = c("1|2", "2|3", "3|4", "educacion_recEd Media", "educacion_recEd Tecnica", "educacion_recUniversidad o posgrado", "educacion_recNs/Nr", "libros_recMás de 25", "acc_tec", "internet_rec2", "internet_recNs/Nr", "mean_educ", "cod_depe2Part.
+subvencionado", "cod_depe2Part.
+privado", "cod_grupoMedio bajo", "cod_grupoMedio", "cod_grupoMedio alto", "cod_grupoAlto", "simceMedio", "simceAlto"), #quitar terminos
+            title = "", #quitar titulo
+            show.values = TRUE, #mostrar valor de efectos
+            dot.size = 1.3, #tamaño circulos
+            line.size = 1, #tamaño CI
+            value.size = 3.5, #tamaño valor efectoss
+            spacing = .7, #espacio entre efectos
+            grid = TRUE, # separa en un plot por modelo
+            vline.color = "firebrick", # linea roja en punto neutro (1)
+            axis.labels = rev(c("School talent", 
+                            "School effort", 
+                            "Social talent",
+                            "Social effort", 
+                            "Deservingness"
+                            )), #con rev porque automatico los tira en otro orden
+            show.legend = FALSE,
+            m.labels = c("Health", "Education","Pension"),
+            colors = gray(level = 0.2))  # variables dependientes
+
+```
+
+@fig-odds demonstrates a similar trend for the three dependent variables of justice in differential access to pensions, education, and health, with all effects being statistically significant.
+In the context of school meritocracy, the effects are mixed: as perceived school talent increases, the justification for differentiated access to these services increases; conversely, as perceived school effort increases, the justification decreases, holding all other variables constant.
+Regarding the variables of social meritocracy, the three variables of talent, effort, and deservingness show that as these increase, the justification for differentiated access to pensions, education, and health also increases.
+Overall, these results provide partial support for our first hypothesis (H1a), which posited that students who perceive higher levels of meritocracy in school would be more likely to justify differential access to pensions, health, and education.
+This is evidenced by the positive effect of perceiving talent as important for achieving good grades.
+However, the results strongly support our second hypothesis (H1b), as a higher perception of societal meritocracy among students tends to increase the justification for market-based justice in these social services.
+
+Contextual effects for ordinal models.
+
+```{r echo=FALSE, include=FALSE}
+
+reg0_bienestar <- lmer(bienestar ~ 1 + (1 | mrbd), data=data_bienestar)
+reg1_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_bienestar)
+reg2_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_bienestar)
+reg3_bienestar <- lmer(bienestar ~ 1 + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_bienestar)
+reg4_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_bienestar)
+
+test01<- anova(reg0_bienestar,reg1_bienestar,test = "Chisq")
+test02<- anova(reg0_bienestar,reg2_bienestar,test = "Chisq")
+test03<- anova(reg0_bienestar,reg3_bienestar,test = "Chisq")
+test04<- anova(reg0_bienestar,reg4_bienestar,test = "Chisq")
+
+test.pvalues1<- test01$`Pr(>Chisq)`[2]
+test.pvalues2<- test02$`Pr(>Chisq)`[2]
+test.pvalues3<- test03$`Pr(>Chisq)`[2]
+test.pvalues4<- test04$`Pr(>Chisq)`[2]
+
+deviance1 <- test01$deviance[2]
+deviance2 <- test02$deviance[2]
+deviance3 <- test03$deviance[2]
+deviance4 <- test04$deviance[2]
+```
+
+```{r echo=FALSE, results='asis'}
+#| tbl-cap: "Multilevel models for meritocracy perception and market justice preferences"
+#| label: tbl-individual-reg
+
+texreg::htmlreg(list(reg1_bienestar, reg2_bienestar),
+        custom.model.names = c("Model 1",
+                               "Model 2"),
+        custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
+        omit.coef = "(internet_rec)",
+        caption = "",
+        custom.coef.names = c("Intercept",  
+                              "School talent",
+                              "School effort", 
+                              "Social talent",  
+                              "Social effort",
+                              "Deservingness",
+                              "Secondary", 
+                              "Higher tec.", 
+                              "University or posgraduate", 
+                              "Missing", 
+                              "More than 25 books (Ref. Less than 25)", 
+                              "Technology access"),
+        groups = list("Parental education (Ref.= 8th grade or less)"=7:10),
+        custom.gof.rows=list("Deviance" = c(deviance1,
+                                            deviance2),
+                             "Deviance Test (p)" = c(test.pvalues1,
+                                                       test.pvalues2))
+        )
+
+```
+
+Regarding our measure of market justice preferences index, the results of the multilevel linear estimation are shown in @tbl-individual-reg.
+The intraclass correlation obtained for the null model (not shown) it's equal to 4% (ICC based on [@hox_multilevel_2010, p.15]), which indicates that only a small percentage of the total variance of the students' market justice preferences are associated with the characteristics of schools, and therefore limits the possibilities of finding effects at this second level of analysis.
 
@@
 
-```{r}
+```{r echo=FALSE, results='asis'}
 #| tbl-cap: "Contextual effects of multilevel regression models"
+#| label: tbl-contextual-reg
+
+texreg::htmlreg(list(reg3_bienestar, reg4_bienestar),
+        custom.model.names = c("Model 1",
+                               "Model 2"),
+        custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
+        omit.coef = c("inteligencia_esc|esfuerzo_esc|inteligencia_soc|esfuerzo_soc|merito_soc|educacion_rec|libros_rec|acc_tec|internet_rec"),
+        caption = "",
+        custom.coef.names = c("Intercept", 
+                              "Prop. university level at school",
+                              "Subsidized private", 
+                              "Private", 
+                              "SES Medium low", 
+                              "SES Medium", 
+                              "SES Medium high", 
+                              "SES High", 
+                              "Simce Medium", 
+                              "Simce High"),
+        groups = list("Administrative dependence (Ref.= Public)"=3:4,
+                      "Socioeconomic level (Ref.= Low)"=5:8,
+                      "Achievement score (Ref.= Low)"=9:10),
+        custom.gof.rows=list("Deviance" = c(deviance3,
+                                            deviance4),
+                             "Deviance Test (p)" = c(test.pvalues3,
+                                                     test.pvalues4))
+        )
+
 ```
 
-### Interactions effects
+```{r echo=FALSE}
+#| tbl-cap: "Interactions effects"
+#| label: tbl-interact
 
-```{r}
-#| tbl-cap: "Interactions effects"
+pacman::p_load(kableExtra)
+load("input/data/interact.RData")
+kbl(collapse_rows,
+    align = "lccccc", booktabs = T,
+    col.names =  c("Variable", 
+                   "School talent", 
+                   "School effort",
+                   "Social talent",
+                   "Social effort", 
+                   "Deservingness")) %>%
+  add_header_above(c(" " = 1,
+                     "Market Justice Preferences"=5), 
+                   bold = T)%>%  
+  kable_styling(latex_options = c("scale_down","HOLD_position"),
+                full_width = F) %>%
+  column_spec(1, bold = T, width = "4cm") %>%
+  column_spec(column = 2:6,width = "1.5cm") %>%
+  collapse_rows(columns = 1:2, valign = "middle") %>% 
+  footnote(general = "***p < 0.001, **p < 0.01, *p < 0.05",
+           footnote_as_chunk = T,general_title = "Note: ")
+
 ```
 
-The interaction terms in Table 6 suggest that students’ family background moderates the relationship between their meritocratic perceptions in Chile and their justification of inequality.
+The interaction terms in @tbl-interact suggest that students' family background moderates the relationship between their meritocratic perceptions in Chile and their justification of inequality.
 The direction of these effects confirms our initial prediction.
@@
 That is, lower-status students (measured by parental education) justify more inequality when they adhere more strongly to deservingnessmeritocratic perceptions in Chile.
-We depict this moderating effect in Figure 1.
 This result confirms the enlightenment thesis (see above).
-The same moderating effect is observed for students’ family cultural capital: the relationship between effort (but not deservingness nor social talent) and justification of inequality becomes less positive for students with more than 25 books at home.
-Interestingly, family cultural capital also moderates the relationship between meritocratic beliefs at school (as measured by students’ belief that effort is important to get good grades) and justification of inequality.
+The same moderating effect is observed for students' family cultural capital: the relationship between effort (but not deservingness nor social talent) and justification of inequality becomes less positive for students with more than 25 books at home.
+Interestingly, family cultural capital also moderates the relationship between meritocratic beliefs at school (as measured by students' belief that effort is important to get good grades) and justification of inequality.
 
-At the school level, we also observe moderating effects of school status and students’ meritocratic beliefs over their justification of inequality.
+At the school level, we also observe moderating effects of school status and students' meritocratic beliefs over their justification of inequality.
 Thus, in line with our Hypothesis 5, high-status schools justify less inequality when, on average, their students have a greater perception of meritocracy in Chile, as measured by the three indicators we used (i.e., effort, deservingness, and talent).
-Finally, as we stated in Hypothesis 7, lowhigh-achieving schools justify more inequality when their students have, on average, stronger meritocratic beliefs in Chile.
-In Figure 25 we depict this moderating effect of school achievement for the relationship between deservingness and justification of inequality.
-We did not observe these moderating effects at the school level for students’ meritocratic beliefs in the school (i.e., the idea that effort or talent are important to get good grades).
+Finally, as we stated in Hypothesis 7, low-achieving schools justify more inequality when their students have, on average, stronger meritocratic beliefs in Chile.
+In @fig-interaction we depict this moderating effect of school achievement for the relationship between deservingness and justification of inequality.
+We did not observe these moderating effects at the school level for students' meritocratic beliefs in the school (i.e., the idea that effort or talent are important to get good grades).
 
-```{r}
+```{r echo=FALSE, include=FALSE}
+load("input/data/data_rec.RData")
+data_rec <- data_rec %>% rowwise() %>%  mutate(educacion_rec2 = case_when(educacion_rec=="Ed Basica"~"Low level",
+                                                 educacion_rec=="Ed Media"~"Low level",
+                                                 educacion_rec=="Ed Tecnica"~"Low level",
+                                                 educacion_rec=="Universidad o posgrado"~"University or Postgraduate",
+                                                 educacion_rec=="Ns/Nr"~"Missing"
+                                                 ))
+data_rec$educacion_rec2 <- factor(data_rec$educacion_rec2, levels = c("Low level", "University or Postgraduate", "Missing"))
+
+data_rec <- data_rec %>% rowwise() %>%  mutate(cod_grupo2  = case_when(cod_grupo=="Bajo"~"Bajo",
+                                                                       cod_grupo=="Medio bajo"~"Bajo",
+                                                                       cod_grupo=="Medio"~"Bajo",
+                                                                       cod_grupo=="Medio alto"~"Alto",
+                                                                       cod_grupo=="Alto"~"Alto"))
+data_rec$cod_grupo2 <- factor(data_rec$cod_grupo2, levels = c("Bajo", "Alto"))
+data_rec$simce <- factor(data_rec$simce, labels = c("Low", "Medium", "High"))
+
+data_bienestar <- na.omit(data_rec)
+#data_bienestar <- data_bienestar %>% mutate(pension_rec = 5-just_pension,
+#                                            educ_rec = 5-just_educ,
+#                                            salud_rec = 5-just_salud)
+data_bienestar <- data_bienestar %>% rowwise() %>% dplyr::mutate(bienestar = mean(c(just_educ, just_salud, just_pension), na.omit=TRUE))
+data_bienestar$bienestar <- round(data_bienestar$bienestar, 2)
+
+reg26.2_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + merito_soc*simce + internet_rec + (1 + simce | mrbd), data=data_bienestar)
+
+plot_interact<- plot_model(reg26.2_bienestar, type = "int", mdrt.values="minmax", title = "", colors = c("red", "white", "blue")) +
+  ylim(1,4)+
+  labs(x = "Deservingness", y="Market Justice Preferences", colour = "School Simce")+
+  theme_bw()
+```
+
+```{r echo=FALSE, warning=FALSE}
 #| fig-cap: "Interaction between deservingness and justification of inequality by school SIMCE achievement"
+#| label: fig-interaction
+
+plot_interact
+
 ```

### 2024-05-15 paper.tex
Primer cambios
Cambios en method.qmd y analisis.qmd

@@
 \usepackage{amsmath,amssymb}
+\usepackage{lmodern}
 \usepackage{iftex}
@@
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-\usepackage{lmodern}
-\ifPDFTeX\else  
-    % xetex/luatex font selection
+  \setmainfont[]{Times New Roman}
 \fi
@@
 \makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
 \makeatother
 \makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
 \makeatother
@@
 
-\author{}
+\author{Equipo EDUMER}
 \date{}
@@
 \begin{document}
-
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
+
+\hypertarget{the-socialization-of-meritocracy-and-market-justice-preferences-at-school}{%
 \section{The socialization of meritocracy and market justice preferences
 at
-school}\label{the-socialization-of-meritocracy-and-market-justice-preferences-at-school}
-
-\subsection{Introduction}\label{introduction}
+school}\label{the-socialization-of-meritocracy-and-market-justice-preferences-at-school}}
+
+\hypertarget{introduction}{%
+\subsection{Introduction}\label{introduction}}
 
@@
 school level represents a threat to the promise of education and a
-meritocratic system (\citeproc{ref-bourdieu_reproduction_1990}{Bourdieu
-and Passeron 1990}). A large part of the research in this field at an
+meritocratic system
+(\protect\hyperlink{ref-bourdieu_reproduction_1990}{Bourdieu \&
+Passeron, 1990}). A large part of the research in this field at an
 international level has addressed the extent to which the social origin
 of students affects their academic results and their life opportunities
-(\citeproc{ref-vonhippel_test_2019}{Von Hippel and Hamrock 2019}),
-confirming that schools have severe difficulties in closing the gaps of
-origin. Besides this socioeconomic perspective on school opportunities,
-recent research has addressed to what extent inequalities in the school
-context are also influencing students' perceptions, beliefs, and
-attitudes: Are social inequalities perceived at the school context? Are
-they rejected by the students, particularly those who are worst-off in
-socioeconomic terms? Or, Is there evidence at the school level that
-social inequalities are tolerated and even justified?
-(\citeproc{ref-batruch_belief_2022}{Batruch et al. 2022};
-\citeproc{ref-wiederkehr_belief_2015}{Wiederkehr et al. 2015}).
+(\protect\hyperlink{ref-vonhippel_test_2019}{Von Hippel \& Hamrock,
+2019}), confirming that schools have severe difficulties in closing the
+gaps of origin. Besides this socioeconomic perspective on school
+opportunities, recent research has addressed to what extent inequalities
+in the school context are also influencing students' perceptions,
+beliefs, and attitudes: Are social inequalities perceived at the school
+context? Are they rejected by the students, particularly those who are
+worst-off in socioeconomic terms? Or, Is there evidence at the school
+level that social inequalities are tolerated and even justified?
+(\protect\hyperlink{ref-batruch_belief_2022}{Batruch et al., 2022};
+\protect\hyperlink{ref-wiederkehr_belief_2015}{Wiederkehr et al.,
+2015}).
 
@@
 by a highly stratified educational system. In particular, we focus on
-market justice preferences (\citeproc{ref-lane_market_1986}{Lane 1986}),
-which refer to the preferences for distributing public goods (as health
-and education) based on criteria such as competition and payment
-capacity. Although from a rational point of view it could be expected an
-opposition to market justice by the underprivileged majority, we argue
-that in a social environment characterized by a promotion of
-meritocratic ideals - as the school system - would lead to the opposite:
-a larger market justice preferences.
+market justice preferences
+(\protect\hyperlink{ref-lane_market_1986}{Lane, 1986}), which refer to
+the preferences for distributing public goods (as health and education)
+based on criteria such as competition and payment capacity. Although
+from a rational point of view it could be expected an opposition to
+market justice by the underprivileged majority, we argue that in a
+social environment characterized by a promotion of meritocratic ideals -
+as the school system - would lead to the opposite: a larger market
+justice preferences.
 
@@
 based on their abilities, knowledge, and achievements
-(\citeproc{ref-young_rise_1958}{Young 1958}). It is often seen as a way
-to create equal opportunities and fairness, as individuals can rise to
-positions of power and influence based on their own merit rather than
-their background or connections. However, some argue that meritocracy
-can actually lead to tolerating or even justifying social inequalities,
-as it can create a hierarchy where those who already have resources and
-advantages are more likely to succeed. In this regard, a great deal of
-academic research about meritocracy delves into the assessment of to
-what extent rewards and privileges in society are related to merit,
-emphasizing the so-called unfulfillable promise of meritocracy
-(\citeproc{ref-mijs_stratified_2016}{Mijs 2016}). Complementing this
-agenda, a second and emerging research area deals with subjective
-aspects of meritocracy, such as perceptions and beliefs.
+(\protect\hyperlink{ref-young_rise_1958}{Young, 1958}). It is often seen
+as a way to create equal opportunities and fairness, as individuals can
+rise to positions of power and influence based on their own merit rather
+than their background or connections. However, some argue that
+meritocracy can actually lead to tolerating or even justifying social
+inequalities, as it can create a hierarchy where those who already have
+resources and advantages are more likely to succeed. In this regard, a
+great deal of academic research about meritocracy delves into the
+assessment of to what extent rewards and privileges in society are
+related to merit, emphasizing the so-called unfulfillable promise of
+meritocracy (\protect\hyperlink{ref-mijs_stratified_2016}{Mijs, 2016}).
+Complementing this agenda, a second and emerging research area deals
+with subjective aspects of meritocracy, such as perceptions and beliefs.
 
@@
 understand the concept of meritocracy in their own society
-(\citeproc{ref-duru-bellat_who_2012}{Duru-Bellat and Tenret 2012};
-\citeproc{ref-castillo_meritocracia_2019}{Castillo et al. 2019}). This
-perception can vary greatly depending on individual experiences, social,
-economic, and cultural background. Some people may see meritocracy as a
-fair and just system that allows anyone to succeed based on their
-abilities and hard work. In contrast, others may view it as a myth or a
-cover for existing power dynamics and inequality, serving to maintain
-and even reinforce inequality
-(\citeproc{ref-lampert_meritocratic_2013}{Lampert 2013};
-\citeproc{ref-mijs_paradox_2021}{Mijs 2021}). Based in this last
-perspective, we argue that individuals with a higher perception of
+(\protect\hyperlink{ref-castillo_meritocracia_2019}{Castillo et al.,
+2019}; \protect\hyperlink{ref-duru-bellat_who_2012}{Duru-Bellat \&
+Tenret, 2012}). This perception can vary greatly depending on individual
+experiences, social, economic, and cultural background. Some people may
+see meritocracy as a fair and just system that allows anyone to succeed
+based on their abilities and hard work. In contrast, others may view it
+as a myth or a cover for existing power dynamics and inequality, serving
+to maintain and even reinforce inequality
+(\protect\hyperlink{ref-lampert_meritocratic_2013}{Lampert, 2013};
+\protect\hyperlink{ref-mijs_paradox_2021}{Mijs, 2021}). Based in this
+last perspective, we argue that individuals with a higher perception of
 meritocracy will show a larger justification of social inequalities, as
 individual achievement would be seen as rewarded and social policies as
-less necessary (\citeproc{ref-batruch_belief_2022}{Batruch et al.
-2022}).
+less necessary (\protect\hyperlink{ref-batruch_belief_2022}{Batruch et
+al., 2022}).
 
@@
 conditions of origin has been linked to the \emph{hidden curriculum}
-(\citeproc{ref-chafel_schooling_1997}{Chafel 1997}), whereby students
-learn about distributive norms in society and mechanisms of
+(\protect\hyperlink{ref-chafel_schooling_1997}{Chafel, 1997}), whereby
+students learn about distributive norms in society and mechanisms of
 justification of social differences. Based on recent studies that relate
 school meritocracy to the justification of economic inequalities in the
-adult population (\citeproc{ref-batruch_belief_2022}{Batruch et al.
-2022}; \citeproc{ref-wiederkehr_belief_2015}{Wiederkehr et al. 2015}),
-the central hypothesis guiding this research is that school-age students
-with a higher perception of meritocracy - both at school and at the
-societal level - will show a larger justification of social
-inequalities, as individual achievement would be seen as appropriately
-rewarded and social mechanisms for correcting inequalities as less
-necessary (\citeproc{ref-batruch_belief_2022}{Batruch et al. 2022}).
+adult population (\protect\hyperlink{ref-batruch_belief_2022}{Batruch et
+al., 2022}; \protect\hyperlink{ref-wiederkehr_belief_2015}{Wiederkehr et
+al., 2015}), the central hypothesis guiding this research is that
+school-age students with a higher perception of meritocracy - both at
+school and at the societal level - will show a larger justification of
+social inequalities, as individual achievement would be seen as
+appropriately rewarded and social mechanisms for correcting inequalities
+as less necessary (\protect\hyperlink{ref-batruch_belief_2022}{Batruch
+et al., 2022}).
 
@@
 perceptions of who deserves what and why
-(\citeproc{ref-kluegel_beliefs_1987}{Kluegel and Smith 1987}),
+(\protect\hyperlink{ref-kluegel_beliefs_1987}{Kluegel \& Smith, 1987}),
 highlights that people's explanations and justifications of social
 inequality are closely tied to their judgments of deservingness. The
-influence of ideologies (\citeproc{ref-wegener_dominant_1995}{Wegener
-and Liebig 1995}) and cultural schemas
-(\citeproc{ref-homan_being_2017}{Homan, Valentino, and Weed 2017}) is
-pivotal in shaping these explanations by offering symbolic
-representations that frame societal structures and expectations. While
-significant attention has been paid to wage inequality, income
-distribution, and payment differentials in the literature {[}Castillo
-(\citeproc{ref-castillo_legitimacy_2011}{2011}); Evans et al., 2010;
-Jasso (\citeproc{ref-jasso_how_1999}{1999}) ; Shariff, Wiwad, and Aknin
-(\citeproc{ref-shariff_income_2016}{2016}){]}, there has been less
-examination of public beliefs about which life domains should be
-governed by market relations and even less about children's acceptance
-or rejection of these market principles. This oversight is notable given
-the extensive encroachment of market logic into public goods, welfare
-policy, and social services over the past five decades
-(\citeproc{ref-centeno_arc_2012}{Centeno and Cohen 2012};
-\citeproc{ref-harvey_breve_2015}{Harvey 2015}), affecting areas such as
-pensions, health services, and education. The justification of social
-inequality based on marketype criteria has been conceptualized as the
-individuals' adherence to one specific justice evaluation, the market
-justice, that is, affording legitimacy to the allocation of goods and
-services based on prices and individuals' ability to pay
-(\citeproc{ref-boltanski_new_2005}{Boltanski and Chiapello 2005};
-\citeproc{ref-lane_market_1986}{Lane 1986};
-\citeproc{ref-streeck_citizens_2012}{Streeck 2012}).
+influence of ideologies
+(\protect\hyperlink{ref-wegener_dominant_1995}{Wegener \& Liebig, 1995})
+and cultural schemas (\protect\hyperlink{ref-homan_being_2017}{Homan et
+al., 2017}) is pivotal in shaping these explanations by offering
+symbolic representations that frame societal structures and
+expectations. While significant attention has been paid to wage
+inequality, income distribution, and payment differentials in the
+literature {[}Castillo
+(\protect\hyperlink{ref-castillo_legitimacy_2011}{2011}); Evans et al.,
+2010; Jasso (\protect\hyperlink{ref-jasso_how_1999}{1999}) ; Shariff et
+al. (\protect\hyperlink{ref-shariff_income_2016}{2016}){]}, there has
+been less examination of public beliefs about which life domains should
+be governed by market relations and even less about children's
+acceptance or rejection of these market principles. This oversight is
+notable given the extensive encroachment of market logic into public
+goods, welfare policy, and social services over the past five decades
+(\protect\hyperlink{ref-centeno_arc_2012}{Centeno \& Cohen, 2012};
+\protect\hyperlink{ref-harvey_breve_2015}{Harvey, 2015}), affecting
+areas such as pensions, health services, and education. The
+justification of social inequality based on marketype criteria has been
+conceptualized as the individuals' adherence to one specific justice
+evaluation, the market justice, that is, affording legitimacy to the
+allocation of goods and services based on prices and individuals'
+ability to pay (\protect\hyperlink{ref-boltanski_new_2005}{Boltanski \&
+Chiapello, 2005}; \protect\hyperlink{ref-lane_market_1986}{Lane, 1986};
+\protect\hyperlink{ref-streeck_citizens_2012}{Streeck, 2012}).
 
@@
 sense of deserving more than one gets'' {[}-Lane
-(\citeproc{ref-lane_market_1986}{1986}); p.~384{]}. Following the theory
-of relative deprivation --a social phenomenon arising when individuals
-cannot afford what most others in their environment can (Merton 1950)--
-Lane notes that, in market settings, social comparisons are more likely
-to motivate increased effort rather than feelings of acute injustice
-because individuals attribute outcomes to their actions. Although
-empirical research has shown that, contrary to Lane's observation,
-relative deprivation, even in market settings, produces feelings of
-dissatisfaction, anger, and resentment that might motivate forms of
-collective action such as protests and revolt
-(\citeproc{ref-greitemeyer_subjective_2016}{Greitemeyer and Sagioglou
-2016}; \citeproc{ref-mishra_subjective_2015}{Mishra and Carleton 2015};
-\citeproc{ref-smith_relative_2012}{Smith et al. 2012};
-\citeproc{ref-power_deprivationprotest_2018}{Séamus A. Power 2018};
-\citeproc{ref-power_relative_2020}{Séamus A. Power, Madsen, and Morton
-2020}), the conceptualization of market justice has been since then
+(\protect\hyperlink{ref-lane_market_1986}{1986}); p.~384{]}. Following
+the theory of relative deprivation --a social phenomenon arising when
+individuals cannot afford what most others in their environment can
+(Merton 1950)-- Lane notes that, in market settings, social comparisons
+are more likely to motivate increased effort rather than feelings of
+acute injustice because individuals attribute outcomes to their actions.
+Although empirical research has shown that, contrary to Lane's
+observation, relative deprivation, even in market settings, produces
+feelings of dissatisfaction, anger, and resentment that might motivate
+forms of collective action such as protests and revolt
+(\protect\hyperlink{ref-greitemeyer_subjective_2016}{Greitemeyer \&
+Sagioglou, 2016}; \protect\hyperlink{ref-mishra_subjective_2015}{Mishra
+\& Carleton, 2015};
+\protect\hyperlink{ref-power_deprivationprotest_2018}{Séamus A. Power,
+2018}; \protect\hyperlink{ref-power_relative_2020}{Séamus A. Power et
+al., 2020}; \protect\hyperlink{ref-smith_relative_2012}{Smith et al.,
+2012}), the conceptualization of market justice has been since then
 closely coupled to the merit principle for allocating outcomes: the
@@
 nations, with substantial differences in funding and delivery methods
-(\citeproc{ref-jensen_worlds_2008}{Jensen 2008};
-\citeproc{ref-stoy_worlds_2014}{Stoy 2014}). Nordic countries, for
-example, predominantly employ public agencies to produce and provide
+(\protect\hyperlink{ref-jensen_worlds_2008}{Jensen, 2008};
+\protect\hyperlink{ref-stoy_worlds_2014}{Stoy, 2014}). Nordic countries,
+for example, predominantly employ public agencies to produce and provide
 social services, funding these through collective taxation and offering
@@
 marketization of welfare services has been growing since the 1980s
-(\citeproc{ref-salamon_marketization_1993}{Salamon 1993}), and this
-shift is increasingly evident even in countries where market solutions
-have traditionally had a minor role in social policy
-(\citeproc{ref-sivesind_changing_2017}{Sivesind 2017}).
+(\protect\hyperlink{ref-salamon_marketization_1993}{Salamon, 1993}), and
+this shift is increasingly evident even in countries where market
+solutions have traditionally had a minor role in social policy
+(\protect\hyperlink{ref-sivesind_changing_2017}{Sivesind, 2017}).
 
@@
 previous policy trajectories
-(\citeproc{ref-fernandez_positive_2013}{Fernandez and Jaime-Castillo
-2013}; \citeproc{ref-pierson_increasing_2000}{Pierson 2000};
-\citeproc{ref-weaver_paths_2010}{Weaver 2010}), citizens' beliefs about
-market justice are likely also shaped by the institutional and social
-contexts they encounter. Indeed, the justification of inequality in
-access to essential services like education and health, based on one's
-ability to pay, shows significant variation across countries, as
-demonstrated by international surveys, although they do not usually
-include children in their samples. For instance, the 2019 International
-Social Survey Program (ISSP) provides insights into how adults perceive
-inequality in accessing welfare services. Figure~\ref{fig-issp}
-illustrates the variation in agreement levels regarding whether it is
-just for individuals with higher incomes to purchase better education or
-healthcare. Notably, in 18 of the 29 countries analyzed, there is a
-greater justification for market inequality in healthcare access than in
-education. Nevertheless, the general sentiment typically ranges from
-`somewhat unjust' to 'neither just nor unjust, with mixed feelings.
+(\protect\hyperlink{ref-fernandez_positive_2013}{Fernandez \&
+Jaime-Castillo, 2013};
+\protect\hyperlink{ref-pierson_increasing_2000}{Pierson, 2000};
+\protect\hyperlink{ref-weaver_paths_2010}{Weaver, 2010}), citizens'
+beliefs about market justice are likely also shaped by the institutional
+and social contexts they encounter. Indeed, the justification of
+inequality in access to essential services like education and health,
+based on one's ability to pay, shows significant variation across
+countries, as demonstrated by international surveys, although they do
+not usually include children in their samples. For instance, the 2019
+International Social Survey Program (ISSP) provides insights into how
+adults perceive inequality in accessing welfare services.
+Figure~\ref{fig-issp} illustrates the variation in agreement levels
+regarding whether it is just for individuals with higher incomes to
+purchase better education or healthcare. Notably, in 18 of the 29
+countries analyzed, there is a greater justification for market
+inequality in healthcare access than in education. Nevertheless, the
+general sentiment typically ranges from `somewhat unjust' to 'neither
+just nor unjust, with mixed feelings.
 
@@
 
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-issp-1.pdf}
+{\centering \includegraphics{paper_files/figure-pdf/fig-issp-1.pdf}
 
@@
 
-\end{figure}%
+\end{figure}
 
@@
 and health services, questions about public acceptance of these changes
-persist. Lindh (\citeproc{ref-lindh_public_2015}{2015}) analysis of ISSP
-2009 data from 17 OECD countries reveals a general lack of support for
-market-based distribution of social services, suggesting widespread
-disapproval of market stratification of essential services. This finding
-is corroborated by Soler-Martínez, García-Sánchez, and Willis
-(\citeproc{ref-soler-martinez_concerns_2023}{2023}) research from
-Latinobarómetro 2020 across 18 Latin American countries, where concerns
-about health and education access predominated over income inequality.
-These results indicate that reforms toward welfare marketization are
-typically driven by elite political decisions rather than grassroots
-demand.
+persist. Lindh (\protect\hyperlink{ref-lindh_public_2015}{2015})
+analysis of ISSP 2009 data from 17 OECD countries reveals a general lack
+of support for market-based distribution of social services, suggesting
+widespread disapproval of market stratification of essential services.
+This finding is corroborated by Soler-Martínez et al.
+(\protect\hyperlink{ref-soler-martinez_concerns_2023}{2023}) research
+from Latinobarómetro 2020 across 18 Latin American countries, where
+concerns about health and education access predominated over income
+inequality. These results indicate that reforms toward welfare
+marketization are typically driven by elite political decisions rather
+than grassroots demand.
 
@@
 responsible for their economic outcomes, a view that varies across the
-region (\citeproc{ref-bucca_merit_2016}{Bucca 2016};
-\citeproc{ref-chong_mystery_2008}{Chong and Ñopo 2008};
-\citeproc{ref-torche_intergenerational_2014}{Torche 2014};
-\citeproc{ref-salgado_inequality_2023}{Salgado and Castillo 2023}). The
-reliance on private welfare providers and widespread user fees
-(\citeproc{ref-molyneux_neoliberal_2008}{Molyneux 2008}) adds complexity
-to this context. Yet, research on children's justification of
-market-based inequalities in accessing welfare services remains limited,
-especially in Latin America, highlighting a significant gap in
-understanding how younger generations view market-based access to
+region (\protect\hyperlink{ref-bucca_merit_2016}{Bucca, 2016};
+\protect\hyperlink{ref-chong_mystery_2008}{Chong \& Ñopo, 2008};
+\protect\hyperlink{ref-salgado_inequality_2023}{Salgado \& Castillo,
+2023}; \protect\hyperlink{ref-torche_intergenerational_2014}{Torche,
+2014}). The reliance on private welfare providers and widespread user
+fees (\protect\hyperlink{ref-molyneux_neoliberal_2008}{Molyneux, 2008})
+adds complexity to this context. Yet, research on children's
+justification of market-based inequalities in accessing welfare services
+remains limited, especially in Latin America, highlighting a significant
+gap in understanding how younger generations view market-based access to
 welfare and whether these views are associated with their meritocratic
@@
 
-\subsection{Meritocracy}\label{meritocracy}
+\hypertarget{meritocracy}{%
+\subsection{Meritocracy}\label{meritocracy}}
 
@@
 signified a valued contribution to the common good.''
-(\citeproc{ref-sandel_tyranny_2020}{Sandel 2020}, pp.).
+(\protect\hyperlink{ref-sandel_tyranny_2020}{Sandel, 2020}, pp.).
 
@@
 mobility that threaten the meritocratic ideal
-(\citeproc{ref-goldthorpe_myth_2003}{Goldthorpe 2003}). More recently,
-sociology and social psychology research has attended to the subjective
-aspects vis-a-vis beliefs in meritocracy. The label of beliefs in this
-realm covers a series of areas, such as attitudes, perceptions, and
-preferences (Castillo et al), whereby most of the link this subjective
-dimensions to individual socio-structural factors and context-level
-determinants. For instance, some studies have analyzed how those with
-greater privileges believe more in meritocracy (Reynolds \& Chan 2014),
-how greater economic inequality increases meritocratic beliefs
-(\citeproc{ref-mijs_paradox_2021}{Mijs 2021}), and how larger inequality
-affects meritocratic beliefs
-(\citeproc{ref-morris_representing_2022}{Morris et al. 2022}). Based on
-these findings, a research agenda has been reinforced on the
+(\protect\hyperlink{ref-goldthorpe_myth_2003}{Goldthorpe, 2003}). More
+recently, sociology and social psychology research has attended to the
+subjective aspects vis-a-vis beliefs in meritocracy. The label of
+beliefs in this realm covers a series of areas, such as attitudes,
+perceptions, and preferences (Castillo et al), whereby most of the link
+this subjective dimensions to individual socio-structural factors and
+context-level determinants. For instance, some studies have analyzed how
+those with greater privileges believe more in meritocracy (Reynolds \&
+Chan 2014), how greater economic inequality increases meritocratic
+beliefs (\protect\hyperlink{ref-mijs_paradox_2021}{Mijs, 2021}), and how
+larger inequality affects meritocratic beliefs
+(\protect\hyperlink{ref-morris_representing_2022}{Morris et al., 2022}).
+Based on these findings, a research agenda has been reinforced on the
 legitimizing role of meritocracy, in line with previous studies using
@@
 negatively with support for redistributive compensation systems
-(\citeproc{ref-frank_performance_2015}{Frank, Wertenbroch, and Maddux
-2015}). Conversely, individuals tend to support redistribution when they
-believe that the disadvantaged lack the opportunities to succeed
-(\citeproc{ref-evans_strong_2018}{Evans and Kelley 2018}). Almås,
-Cappelen, and Tungodden (\citeproc{ref-almas_cutthroat_2020}{2020})
-found that in a relatively unequal society (the United States), the
-highly educated accept inequality significantly more than the less
-educated because they perceive inequality as justifiable owing to
-differences in productivity (i.e., merit), whereas in a relatively equal
-society (Norway), the less educated accept inequality more, but not
+(\protect\hyperlink{ref-frank_performance_2015}{Frank et al., 2015}).
+Conversely, individuals tend to support redistribution when they believe
+that the disadvantaged lack the opportunities to succeed
+(\protect\hyperlink{ref-evans_strong_2018}{Evans \& Kelley, 2018}).
+Almås et al. (\protect\hyperlink{ref-almas_cutthroat_2020}{2020}) found
+that in a relatively unequal society (the United States), the highly
+educated accept inequality significantly more than the less educated
+because they perceive inequality as justifiable owing to differences in
+productivity (i.e., merit), whereas in a relatively equal society
+(Norway), the less educated accept inequality more, but not
 significantly more than the highly educated because meritocratic values
-are less prevalent. Barr and Miller
-(\citeproc{ref-barr_effect_2020}{2020}) also addressed this triple
-interaction between the level of inequality in a society, the individual
-level of education, and the perceived origin of the disparity (either by
-luck or effort) to determine the extent to which inequality is accepted.
-They found that the interaction's mechanism varies depending on the
-compared societies. Finally, García-Sánchez et al.
-(\citeproc{ref-garcia-sanchez_attitudes_2020}{2020}), using data for 41
-countries from the International Social Survey Programme (ISSP), found
-that the perceived size of the income gap correlated positively with
-support for progressive taxation. Still, this association was weaker
-among those who endorsed meritocratic and equal opportunity beliefs. In
-the same line, experimental research by Durante and Putterman
-(\citeproc{ref-durante_preferences_2009}{2009}) subjects support less
-redistribution when the initial distribution is determined according to
-task performance.
+are less prevalent. Barr \& Miller
+(\protect\hyperlink{ref-barr_effect_2020}{2020}) also addressed this
+triple interaction between the level of inequality in a society, the
+individual level of education, and the perceived origin of the disparity
+(either by luck or effort) to determine the extent to which inequality
+is accepted. They found that the interaction's mechanism varies
+depending on the compared societies. Finally, García-Sánchez et al.
+(\protect\hyperlink{ref-garcia-sanchez_attitudes_2020}{2020}), using
+data for 41 countries from the International Social Survey Programme
+(ISSP), found that the perceived size of the income gap correlated
+positively with support for progressive taxation. Still, this
+association was weaker among those who endorsed meritocratic and equal
+opportunity beliefs. In the same line, experimental research by Durante
+\& Putterman (\protect\hyperlink{ref-durante_preferences_2009}{2009})
+subjects support less redistribution when the initial distribution is
+determined according to task performance.
 
@@
 socialization institutions where achievement based on merit explains
-success (\citeproc{ref-erivwo_meritocracy_2021}{Erivwo et al. 2021}).
-Nevertheless, it is possible to find some initial works focused on the
-area of distributive justice at school that are closely related to
-meritocratic beliefs, as the ones by Resh and Sabbagh (quotes). Using
+success (\protect\hyperlink{ref-erivwo_meritocracy_2021}{Erivwo et al.,
+2021}). Nevertheless, it is possible to find some initial works focused
+on the area of distributive justice at school that are closely related
+to meritocratic beliefs, as the ones by Resh and Sabbagh (quotes). Using
 justice in grade obtention as a measure of distributive justice and
@@
 sense of distributive justice about grades is associated to higher
-socio-economic status (\citeproc{ref-resh_sense_2010}{Resh 2010}), have
-a positive effect on liberal democratic orientation and on trust in
-people and in formal institutions (\citeproc{ref-resh_sense_2014}{Resh
-and Sabbagh 2014}), and tend to refrain from violence and to engage to a
-greater extent in extra-curricular school activity and community
-volunteering (\citeproc{ref-resh_sense_2017}{Resh and Sabbagh 2017}).
-
+socio-economic status (\protect\hyperlink{ref-resh_sense_2010}{Resh,
+2010}), have a positive effect on liberal democratic orientation and on
+trust in people and in formal institutions
+(\protect\hyperlink{ref-resh_sense_2014}{Resh \& Sabbagh, 2014}), and
+tend to refrain from violence and to engage to a greater extent in
+extra-curricular school activity and community volunteering
+(\protect\hyperlink{ref-resh_sense_2017}{Resh \& Sabbagh, 2017}).
+
+\hypertarget{childrens-judgments-of-inequality-schools-and-family-background}{%
 \subsection{Children's judgments of inequality, schools, and family
-background}\label{childrens-judgments-of-inequality-schools-and-family-background}
+background}\label{childrens-judgments-of-inequality-schools-and-family-background}}
 
@@
 consistent with research on cognitive development
-(\citeproc{ref-choudhury_social_2006}{Choudhury, Blakemore, and Charman
-2006}). Adolescents with mature socio-cognitive abilities tend to
-express stronger preferences for fairness than infants and children
-(\citeproc{ref-wynn_not_2018}{Wynn et al. 2018}). As children grow
-older, they become more likely to behave fairly, with their
+(\protect\hyperlink{ref-choudhury_social_2006}{Choudhury et al., 2006}).
+Adolescents with mature socio-cognitive abilities tend to express
+stronger preferences for fairness than infants and children
+(\protect\hyperlink{ref-wynn_not_2018}{Wynn et al., 2018}). As children
+grow older, they become more likely to behave fairly, with their
 early-emerging strict egalitarianism being replaced by an increasing
 endorsement of fairness principles and engagement in collaborative
-activities (\citeproc{ref-huppert_development_2019}{Huppert et al.
-2019}; \citeproc{ref-mcauliffe_developmental_2017}{McAuliffe et al.
+activities (\protect\hyperlink{ref-huppert_development_2019}{Huppert et
+al., 2019};
+\protect\hyperlink{ref-mcauliffe_developmental_2017}{McAuliffe et al.,
 2017}). In these activities, their fairness views consider individual
 contributions, merits, and circumstances
-(\citeproc{ref-almas_fairness_2010}{Almås et al. 2010};
-\citeproc{ref-huppert_development_2019}{Huppert et al. 2019};
-\citeproc{ref-sigelman_development_1991}{Sigelman and Waitzman 1991}).
-Engelmann and Tomasello (\citeproc{ref-engelmann_children_2019}{2019})
-claim that children's sense of fairness emerges at three years old, and
-we can observe it in collaborative activities, where they accept
-inequality if the procedure gives everyone an equal chance. Therefore,
-children at this age respond to unequal distributions based on
-interpersonal concerns, as they already demand equal respect. In any
-case, between 3 and 8 years of age, inequitable and anti-meritorious
-allocations are evaluated more negatively, but equitable and meritorious
-allocations are not evaluated more positively
-(\citeproc{ref-elenbaas_unfairness_2019}{Elenbaas 2019}).
+(\protect\hyperlink{ref-almas_fairness_2010}{Almås et al., 2010};
+\protect\hyperlink{ref-huppert_development_2019}{Huppert et al., 2019};
+\protect\hyperlink{ref-sigelman_development_1991}{Sigelman \& Waitzman,
+1991}). Engelmann \& Tomasello
+(\protect\hyperlink{ref-engelmann_children_2019}{2019}) claim that
+children's sense of fairness emerges at three years old, and we can
+observe it in collaborative activities, where they accept inequality if
+the procedure gives everyone an equal chance. Therefore, children at
+this age respond to unequal distributions based on interpersonal
+concerns, as they already demand equal respect. In any case, between 3
+and 8 years of age, inequitable and anti-meritorious allocations are
+evaluated more negatively, but equitable and meritorious allocations are
+not evaluated more positively
+(\protect\hyperlink{ref-elenbaas_unfairness_2019}{Elenbaas, 2019}).
 
@@
 behaviors by playing an essential role in the transmission of equity
-norms (\citeproc{ref-schunk_fairness_2023}{Schunk and Zipperle 2023};
-\citeproc{ref-kosse_prosociality_2020}{Kosse and Tincani 2020}). In
-fact, schools contribute to institutionalizing and reproducing
-inequality by promoting values, norms, practices, and languages familiar
-to higher-class families because the dominant group's culture shapes
-educational institutions
-(\citeproc{ref-bourdieu_reproduction_1990}{Bourdieu and Passeron 1990}).
-Middle- and upper-class students are better equipped to face academic
-challenges and are more familiar with academic expectations
-(\citeproc{ref-mikus_children_2020}{Mikus, Tieben, and Schober 2020}).
-Such familiarity represents cultural capital in educational contexts
-because higher-status students come to school ready to meet these
-expectations and reap the benefits (\citeproc{ref-jack_no_2016}{Jack
-2016}; \citeproc{ref-khan_privilege_2011}{Khan 2011}). Conversely,
+norms (\protect\hyperlink{ref-kosse_prosociality_2020}{Kosse \& Tincani,
+2020}; \protect\hyperlink{ref-schunk_fairness_2023}{Schunk \& Zipperle,
+2023}). In fact, schools contribute to institutionalizing and
+reproducing inequality by promoting values, norms, practices, and
+languages familiar to higher-class families because the dominant group's
+culture shapes educational institutions
+(\protect\hyperlink{ref-bourdieu_reproduction_1990}{Bourdieu \&
+Passeron, 1990}). Middle- and upper-class students are better equipped
+to face academic challenges and are more familiar with academic
+expectations (\protect\hyperlink{ref-mikus_children_2020}{Mikus et al.,
+2020}). Such familiarity represents cultural capital in educational
+contexts because higher-status students come to school ready to meet
+these expectations and reap the benefits
+(\protect\hyperlink{ref-jack_no_2016}{Jack, 2016};
+\protect\hyperlink{ref-khan_privilege_2011}{Khan, 2011}). Conversely,
 lower-status children lacking cultural capital must catch up while
 experiencing inequitable comparisons
-(\citeproc{ref-goudeau_hidden_2017}{Goudeau and Croizet 2017}).
+(\protect\hyperlink{ref-goudeau_hidden_2017}{Goudeau \& Croizet, 2017}).
 Additionally, academic achievement is treated as the outcome of
@@
 not due to their family background but rather to differences in efforts
-and talents (\citeproc{ref-darnon_where_2018}{Darnon et al. 2018}). In
-this sense, we believe that the perception of meritocracy can influence
-students' judgments about market justice preferences. Furthermore, we
-believe there is a difference between students' perceptions of
-meritocracy based on their own experience in school and what they
-perceive in society at large. Consistently, our first two hypotheses
-are:
+and talents (\protect\hyperlink{ref-darnon_where_2018}{Darnon et al.,
+2018}). In this sense, we believe that the perception of meritocracy can
+influence students' judgments about market justice preferences.
+Furthermore, we believe there is a difference between students'
+perceptions of meritocracy based on their own experience in school and
+what they perceive in society at large. Consistently, our first two
+hypotheses are:
 
@@
 children's and adolescent's market justice preferences. For example,
-Almås et al. (\citeproc{ref-almas_fairness_2017}{2017}) found that
-adolescents from low-socioeconomic-status families are likelier to have
-an egalitarian fairness view and consider an equal distribution as fair
-in a situation with unequal merits. The authors speculate that
+Almås et al. (\protect\hyperlink{ref-almas_fairness_2017}{2017}) found
+that adolescents from low-socioeconomic-status families are likelier to
+have an egalitarian fairness view and consider an equal distribution as
+fair in a situation with unequal merits. The authors speculate that
 differences in socialization practices across status groups might bring
@@
 because social status seems to interact with these evaluations (e.g.,
-\citeproc{ref-hvidberg_social_2023}{Hvidberg, Kreiner, and Stantcheva
-2023}).
+\protect\hyperlink{ref-hvidberg_social_2023}{Hvidberg et al., 2023}).
 
@@
 promote deference, obedience, and conformity to authority
-(\citeproc{ref-kohn_social_1963}{Kohn 1963};
-\citeproc{ref-kohn_class_1969}{Kohn and Schooler 1969}). Although
-parents from all social backgrounds encourage individualism in their
-children, this shared norm translates into different forms in high and
-low social classes (1999). Acemoglu
-(\citeproc{ref-acemoglu_obedience_2021}{2021}) claimed that the values
-families impart to their children interact with social mobility. Because
-obedience is a valuable characteristic for employers, in low-wage and
-social mobility environments, low-income families impart values of
-obedience to their children to prevent disadvantaging them in labor
-markets. On the one hand, children from privileged families are
+(\protect\hyperlink{ref-kohn_social_1963}{Kohn, 1963};
+\protect\hyperlink{ref-kohn_class_1969}{Kohn \& Schooler, 1969}).
+Although parents from all social backgrounds encourage individualism in
+their children, this shared norm translates into different forms in high
+and low social classes (1999). Acemoglu
+(\protect\hyperlink{ref-acemoglu_obedience_2021}{2021}) claimed that the
+values families impart to their children interact with social mobility.
+Because obedience is a valuable characteristic for employers, in
+low-wage and social mobility environments, low-income families impart
+values of obedience to their children to prevent disadvantaging them in
+labor markets. On the one hand, children from privileged families are
 socialized to adopt a clear conception of individualism that highlights
@@
 as resources to overcome collective impediments on the path to upward
-mobility (\citeproc{ref-iacoviello_collectivism_2019}{Iacoviello and
-Lorenzi-Cioldi 2019}). In this way, we believe that there are
+mobility
+(\protect\hyperlink{ref-iacoviello_collectivism_2019}{Iacoviello \&
+Lorenzi-Cioldi, 2019}). In this way, we believe that there are
 differences in the socialization of values according to socioeconomic
@@
 significantly influences children's and adolescents' views on inequality
-and deservingness. For example, Jonsson and Beach
-(\citeproc{ref-jonsson_institutional_2015}{2015}) study revealed that
-higher-status adolescents in Sweden tend to perpetuate social class
-stereotypes while describing the vocational and academic tracks.
-Academic track students are depicted as wealthy, intelligent, ambitious,
-and diligent, while vocational track students are characterized as poor,
-unambitious, unintelligent, and lackadaisical. These stereotypes help
-individuals maintain a sense of superiority over others and legitimize
-the prevailing social hierarchies and economic disparities
-(\citeproc{ref-jost_attitudinal_2000}{Jost and Burgess 2000})
+and deservingness. For example, Jonsson \& Beach
+(\protect\hyperlink{ref-jonsson_institutional_2015}{2015}) study
+revealed that higher-status adolescents in Sweden tend to perpetuate
+social class stereotypes while describing the vocational and academic
+tracks. Academic track students are depicted as wealthy, intelligent,
+ambitious, and diligent, while vocational track students are
+characterized as poor, unambitious, unintelligent, and lackadaisical.
+These stereotypes help individuals maintain a sense of superiority over
+others and legitimize the prevailing social hierarchies and economic
+disparities (\protect\hyperlink{ref-jost_attitudinal_2000}{Jost \&
+Burgess, 2000})
 
@@
 
-\subsubsection{Summary of hypotheses}\label{summary-of-hypotheses}
+\hypertarget{summary-of-hypotheses}{%
+\subsubsection{Summary of hypotheses}\label{summary-of-hypotheses}}
 
@@
 
-\centering{
-
-\includegraphics[width=6.25in,height=\textheight]{input/img/hypotheses.png}
+{\centering \includegraphics[width=6.25in,height=\textheight]{input/img/hypotheses.png}
 
@@
 
-\end{figure}%
-
-\subsection{Methods}\label{methods}
-
-\subsubsection{Data}\label{data}
-
-The main data source for the analysis is the First Study of Citizenship
-Education in Chile, carried out by the Education Quality Agency of the
-Ministry of Education. The application date was November 9, 2017. The
-target population of this study is Eighth-grade students from 242
-schools. In the data, there are 8,589 students and 6,770 parents. This
-database was analyzed with the R package ``ResponsePatterns'' to detect
-possible repetitive and ``careless'' response patterns and thus
-contribute to a higher quality of research data (Gottfried, et al 2022).
-Responses from 171 students and 79 parents were removed, which, when
-merged, gave a total of 6,270 valid cases.
-
-The analysis of school variables includes data from the Ministry of
-Education's SIMCE 2017 database. This database contains information at
-the school level, such as the administrative dependency, its
-socioeconomic classification, and the achievement scores obtained in the
-mathematics and language census tests. It is available for free use on
-the MINEDUC {[}web page{]}.
-
-After eliminating missing cases, the final sample used in the analysis
-was based on 5,047 students and parents of 231 schools for the dependent
-variable of access to social services.
-
-\subsubsection{Variables}\label{variables}
-
-\textbf{Dependent variables}
-
-This study has three dependent variables related to the justification of
-social inequality in specific policy domains. The first asks whether
-access to social services should be conditional on income, i,.e., ``It
-is just that in Chile people who can pay have a better education for
-their children''. Students rated their preferences using the following
-responses: ``strongly disagree'', ``Disagree'', ``Agree'', and
-''strongly agree''. An average index is built with these items
-(Cronbach's Alpha = 0,86). (Apéndice: items en español y su
-correspondiente traducción al inglés)
-
-Table~\ref{tbl-desc-dependientes} shows the items used, their response
-categories, and their frequencies.
-
+\end{figure}
+
+\hypertarget{methods}{%
+\subsection{Methods}\label{methods}}
+
+\hypertarget{data}{%
+\subsubsection{Data}\label{data}}
+
+The main data source corresponds to the First Study of Civic Education
+in Chile, conducted by the Agency for Quality Education of the Ministry
+of Education in 2017. This database is based on the administration of a
+civic knowledge test to students and questionnaires to various members
+of the educational community. The target population consists of
+8th-grade students in 242 schools across the country. Initially, the
+database comprises 8,589 student observations and 6,770 parent
+observations. However, considering the survey's unit of analysis and to
+ensure higher data quality, 171 student cases and 79 parent cases
+exhibiting repetitive and careless response patterns were removed
+(Gottfried et al., 2022). Additionally, data from the System for
+Measuring the Quality of Education (SIMCE) conducted by the Ministry of
+Education for the year 2017 are employed, providing school-level
+information such as administrative dependence, socioeconomic
+classification, and results obtained in standardized mathematics and
+language tests. After variable processing and removal of missing cases,
+the final database for this study includes a two-level stratified
+sample, composed of 5,047 students and parents (level 1) nested within
+231 schools (level 2).
+
+\hypertarget{variables}{%
+\subsubsection{Variables}\label{variables}}
+
+\textbf{\emph{Individual level}}
+
+\textbf{Justification of Inequality}: The dependent variable in this
+study is the justification of social inequality in specific policy
+domains, namely health, education, and pensions. This construct is
+measured through three variables addressing the degree of justification
+regarding whether access to social services in these domains should be
+income conditional. Justification of inequality in health is measured by
+the item ``Is it fair in Chile that people with higher incomes can
+access better healthcare than those with lower incomes,'' while in
+education, it is measured using the statement ``Is it fair in Chile that
+people who can afford it have better education for their children.'' In
+the case of pensions, it is measured through the item ``Is it fair in
+Chile that people with higher incomes can have better pensions than
+those with lower incomes.'' In all cases, respondents declare their
+preferences on a Likert scale ranging from ``completely disagree'' (1)
+to ``completely agree'' (4). Additionally, we include a summarized
+indicator of preferences for market justice (REF), measured by an
+average index across these items (α = 0.86), with values ranging from 1
+to 4, where higher values represent greater preferences for market
+justice (see Table~\ref{tbl-desc-dependientes}).
+
+\hypertarget{tbl-desc-dependientes}{}
 \begin{longtable}[]{@{}
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1020}}@{}}
-\caption{Dependent
-variables}\label{tbl-desc-dependientes}\tabularnewline
-\toprule\noalign{}
+\caption{\label{tbl-desc-dependientes}Dependent
+variables}\tabularnewline
+\toprule()
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule()
 \endfirsthead
-\toprule\noalign{}
+\toprule()
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule()
 \endhead
-\bottomrule\noalign{}
-\endlastfoot
 It is just that in Chile people with higher incomes can have better
@@
 \end{minipage} \\
+\bottomrule()
 \end{longtable}
 
-\textbf{Independent variables}
-
-For the primary independent variable, the perception of meritocracy,
-five items address the perception of rewards according to effort and
-intelligence at the school and societal levels. At the school level,
-students answer whether ``Intelligence is important to get good grades''
-and ``Effort is important to get good grades''. At the societal level,
-students respond to the following questions: ``In Chile, people are
-rewarded for their efforts'', ``In Chile, people get what they
-deserve'', and ``In Chile, people are rewarded for their intelligence
-and skills''. Each item was answered on a four-point scale ranging from
-``Strongly disagree'' to ``Strongly agree''.
-
-The rest of the independent variables are divided into individual and
-school levels. At the individual level, family socioeconomic status was
-measured by the parents' highest educational level and the number of
-books at home. Likewise, an index of access to technology includes the
-number of computers, tablets, and cell phones at home, as well as
-whether there is an Internet connection.
-Table~\ref{tbl-desc-independent} shows the items used, their response
-categories, and their frequency.
-
+\textbf{Perception of Meritocracy}: The main independent variable refers
+to the perception of meritocracy, operationalized through five items
+addressing the perception of rewards based on talent and intelligence at
+both the school and societal levels. At the school level, students
+respond to whether ``Intelligence is important for getting good grades''
+and ``Effort is important for getting good grades.'' At the societal
+level, students respond to the following questions: ``In Chile, people
+are rewarded for their effort,'' ``In Chile, people get what they
+deserve,'' and ``In Chile, people are rewarded for their intelligence
+and skills.'' Each item was answered on a four-point Likert scale
+ranging from ``completely disagree'' (1) to ``completely agree'' (4).
+
+The analysis includes several individual control variables that have
+been identified as influential in recent research. These variables
+encompass family socioeconomic status, assessed by the highest level of
+parental education, the number of books in the household, and an index
+measuring access to technologies, all aimed at mitigating compositional
+effects (REFS). Table~\ref{tbl-desc-independent} shows the independent
+variables used, their response categories and their frequencies.
+
+\hypertarget{tbl-desc-independent}{}
 \begin{longtable}[]{@{}
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1028}}@{}}
-\caption{Independent
-variables}\label{tbl-desc-independent}\tabularnewline
-\toprule\noalign{}
+\caption{\label{tbl-desc-independent}Independent
+variables}\tabularnewline
+\toprule()
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule()
 \endhead
-\bottomrule\noalign{}
-\endlastfoot
 Intelligence is important to get good grades &
@@
 \end{minipage} \\
+\bottomrule()
 \end{longtable}
 
-The school-level variables are the administrative dependency of the
-school, the socioeconomic classification made by the Ministry of
-Education, the level of performance in the SIMCE test of the school, and
-the proportion of parents with university or postgraduate degrees.
-Table~\ref{tbl-desc-school} shows the items used, response categories,
-and frequency.
-
+\textbf{\emph{Contextual level}}
+
+This study examines four variables at the school level. Firstly, we
+assess the administrative dependence of schools, categorized as
+``Public,'' ``Subsidized Private,'' or ``Private.'' Secondly, we
+consider the socioeconomic classification of schools, determined by the
+Ministry of Education and measured on an ordinal scale with five
+categories ranging from ``low'' (1) to ``high'' (5). Thirdly, we
+evaluate the performance level of schools in SIMCE standardized tests,
+categorized as ``low,'' ``medium,'' or ``high.'' Lastly, we analyze the
+proportion of parents with university or postgraduate education level at
+each school. The items used, response categories, and their frequencies
+are detailed in Table~\ref{tbl-desc-school}.
+
+\hypertarget{tbl-desc-school}{}
 \begin{longtable}[]{@{}
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1111}}@{}}
-\caption{School context variables}\label{tbl-desc-school}\tabularnewline
-\toprule\noalign{}
+\caption{\label{tbl-desc-school}School context variables}\tabularnewline
+\toprule()
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule()
 \endhead
-\bottomrule\noalign{}
-\endlastfoot
 Proportion of parents with university level by school &
@@
 \end{minipage} \\
+\bottomrule()
 \end{longtable}
 
-\subsubsection{Methods}\label{methods-1}
-
-The data has a hierarchical structure of students nested in schools, so
-the model estimation is performed in a multilevel (random effects)
-framework. This modeling approach lets us correctly estimate individual
-and contextual effects in a single model. We estimate cumulative link
-mixed models for the ordinal dependent variables, whereas we use linear
-mixed effects models for the average index of inequality justification.
+\hypertarget{methods-1}{%
+\subsubsection{Methods}\label{methods-1}}
+
+Given the hierarchical structure of the data, with students nested
+within schools, model estimation is performed within a multilevel
+framework. These models are appropriate for capturing both individual
+and contextual effects in a single analysis, allowing for the estimation
+of fixed effects between groups and random effects that vary from one
+group to another (\protect\hyperlink{ref-bell_fixed_2019}{Bell et al.,
+2019}; \protect\hyperlink{ref-hox_multilevel_2010}{Hox et al., 2010}).
+Cumulative link mixed models were employed for the ordinal dependent
+variables, while linear mixed-effects models were applied for the
+average market justice preference index.
 
@@
 
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-bivariate-1.pdf}
+{\centering \includegraphics{paper_files/figure-pdf/fig-bivariate-1.pdf}
 
@@
 
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-odds-1.pdf}
+{\centering \includegraphics{paper_files/figure-pdf/fig-odds-1.pdf}
 
@@
 
+\hypertarget{multilevel-regression-models-for-market-justice-preferences}{%
 \subsubsection{Multilevel regression models for market justice
-preferences}\label{multilevel-regression-models-for-market-justice-preferences}
-
+preferences}\label{multilevel-regression-models-for-market-justice-preferences}}
+
+\hypertarget{tbl-individual-reg}{}
 \begin{table}
-
-\caption{\label{tbl-individual-reg}Individual effects of multilevel
-regression models}
-
-\centering{
-
 \begin{center}
@@
 \end{tabular}
-\caption{}
+\caption{\label{tbl-individual-reg}Individual effects of multilevel regression models }
 \label{table:coefficients}
 \end{center}
-
-}
-
-\end{table}%
+\end{table}
 
@@
 
+\hypertarget{tbl-contextual-reg}{}
 \begin{table}
-
-\caption{\label{tbl-contextual-reg}Contextual effects of multilevel
-regression models}
-
-\centering{
-
 \begin{center}
@@
 \end{tabular}
-\caption{}
+\caption{\label{tbl-contextual-reg}Contextual effects of multilevel regression models }
 \label{table:coefficients}
 \end{center}
-
-}
-
-\end{table}%
-
-\subsubsection{Interactions effects}\label{interactions-effects}
-
-\begin{table}
-
-\caption{\label{tbl-interact}Interactions effects}
-
-\centering{
-
-[H]
+\end{table}
+
+\hypertarget{interactions-effects}{%
+\subsubsection{Interactions effects}\label{interactions-effects}}
+
+\hypertarget{tbl-interact}{}
+\begin{table}[H]
+\caption{\label{tbl-interact}Interactions effects }\tabularnewline
+
 \centering
-\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
+\resizebox{\linewidth}{!}{
 \begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
@@
 \end{tabular}}
-
-}
-
-\end{table}%
+\end{table}
 
@@
 
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-interaction-1.pdf}
+{\centering \includegraphics{paper_files/figure-pdf/fig-interaction-1.pdf}
 
@@
 
-\end{figure}%
-
-\subsection{Conclusion}\label{conclusion}
+\end{figure}
+
+\hypertarget{conclusion}{%
+\subsection{Conclusion}\label{conclusion}}
 
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5504}}@{}}
-\toprule\noalign{}
+\toprule()
 \endhead
-\bottomrule\noalign{}
-\endlastfoot
 \textbf{English translation} & \textbf{Original Spanish} \\
@@
 salud que las personas con ingresos más bajos \\
+\bottomrule()
 \end{longtable}
 
-\phantomsection\label{refs}
+\hypertarget{refs}{}
 \begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-acemoglu_obedience_2021}
-Acemoglu, Daron. 2021. {``Obedience in the {Labor Market} and {Social
-Mobility}: {A Socio-Economic Approach}.''} w29125. Cambridge, MA:
-National Bureau of Economic Research.
-\url{https://doi.org/10.3386/w29125}.
-
-\bibitem[\citeproctext]{ref-almas_fairness_2017}
-Almås, Ingvild, Alexander W Cappelen, Kjell G Salvanes, Erik Ø Sørensen,
-and Bertil Tungodden. 2017. {``Fairness and Family Background.''}
-\emph{Politics, Philosophy \& Economics} 16 (2): 117--31.
-\url{https://doi.org/10.1177/1470594X15618966}.
-
-\bibitem[\citeproctext]{ref-almas_fairness_2010}
-Almås, Ingvild, Alexander W. Cappelen, Erik Ø. Sørensen, and Bertil
-Tungodden. 2010. {``Fairness and the {Development} of {Inequality
-Acceptance}.''} \emph{Science} 328 (5982): 1176--78.
-\url{https://doi.org/10.1126/science.1187300}.
-
-\bibitem[\citeproctext]{ref-almas_cutthroat_2020}
-Almås, Ingvild, Alexander W. Cappelen, and Bertil Tungodden. 2020.
-{``Cutthroat {Capitalism} Versus {Cuddly Socialism}: {Are Americans More
-Meritocratic} and {Efficiency-Seeking} Than {Scandinavians}?''}
-\emph{Journal of Political Economy} 128 (5): 1753--88.
-\url{https://doi.org/10.1086/705551}.
-
-\bibitem[\citeproctext]{ref-barr_effect_2020}
-Barr, Abigail, and Luis Miller. 2020. {``The Effect of Education, Income
-Inequality and Merit on Inequality Acceptance.''} \emph{Journal of
-Economic Psychology} 80 (May): 102276.
-\url{https://doi.org/10.1016/j.joep.2020.102276}.
-
-\bibitem[\citeproctext]{ref-batruch_belief_2022}
-Batruch, Anatolia, Jolanda Jetten, Herman Van de Werfhorst, Céline
-Darnon, and Fabrizio Butera. 2022. {``Belief in {School Meritocracy} and
-the {Legitimization} of {Social} and {Income Inequality}.''}
-\emph{Social Psychological and Personality Science}, August,
-194855062211110. \url{https://doi.org/10.1177/19485506221111017}.
-
-\bibitem[\citeproctext]{ref-boltanski_new_2005}
-Boltanski, Luc, and Eve Chiapello. 2005. \emph{The New Spirit of
-Capitalism}. London ; New York: Verso.
-
-\bibitem[\citeproctext]{ref-bourdieu_reproduction_1990}
-Bourdieu, Pierre, and Jean Claude Passeron. 1990. \emph{Reproduction in
-{Education}, {Society} and {Culture}}. Second Edition. Sage Publications
-Ltd.
-
-\bibitem[\citeproctext]{ref-bucca_merit_2016}
-Bucca, Mauricio. 2016. {``Merit and Blame in Unequal Societies:
-{Explaining Latin Americans}' Beliefs about Wealth and Poverty.''}
-\emph{Research in Social Stratification and Mobility} 44 (June):
-98--112. \url{https://doi.org/10.1016/j.rssm.2016.02.005}.
-
-\bibitem[\citeproctext]{ref-castillo_legitimacy_2011}
-Castillo, Juan Carlos. 2011. {``Legitimacy of {Inequality} in a {Highly
-Unequal Context}: {Evidence} from the {Chilean Case}.''} \emph{Social
-Justice Research} 24 (4): 314--40.
-\url{https://doi.org/10.1007/s11211-011-0144-5}.
-
-\bibitem[\citeproctext]{ref-castillo_meritocracia_2019}
-Castillo, Juan Carlos, Alex Torres, Jorge Atria, and Luis Maldonado.
-2019. {``Meritocracia y Desigualdad Econ{ó}mica: {Percepciones},
-Preferencias e Implicancias.''} \emph{Revista Internacional de
-Sociolog{í}a} 77 (1): 117.
-\url{https://doi.org/10.3989/ris.2019.77.1.17.114}.
-
-\bibitem[\citeproctext]{ref-centeno_arc_2012}
-Centeno, Miguel A., and Joseph N. Cohen. 2012. {``The {Arc} of
-{Neoliberalism}.''} \emph{Annual Review of Sociology} 38 (1): 317--40.
-\url{https://doi.org/10.1146/annurev-soc-081309-150235}.
-
-\bibitem[\citeproctext]{ref-chafel_schooling_1997}
-Chafel, Judith A. 1997. {``Schooling, the {Hidden Curriculum}, and
-{Children}'s {Conceptions} of {Poverty}.''} \emph{Social Policy Report}
-11 (1): 1--28. \url{https://doi.org/10.1002/j.2379-3988.1997.tb00004.x}.
-
-\bibitem[\citeproctext]{ref-chong_mystery_2008}
-Chong, Alberto, and Hugo Ñopo. 2008. {``The {Mystery} of
-{Discrimination} in {Latin America}.''} \emph{Econom{í}a} 8 (2):
-79--107. \url{https://doi.org/10.1353/eco.0.0005}.
-
-\bibitem[\citeproctext]{ref-choudhury_social_2006}
-Choudhury, Suparna, Sarah-Jayne Blakemore, and Tony Charman. 2006.
-{``Social Cognitive Development During Adolescence.''} \emph{Social
-Cognitive and Affective Neuroscience} 1 (3): 165--74.
-\url{https://doi.org/10.1093/scan/nsl024}.
-
-\bibitem[\citeproctext]{ref-darnon_where_2018}
-Darnon, Céline, Virginie Wiederkehr, Benoît Dompnier, and Delphine
-Martinot. 2018. {``{`{Where} There Is a Will, There Is a Way'}: {Belief}
-in School Meritocracy and the Social-Class Achievement Gap.''}
-\emph{British Journal of Social Psychology} 57 (1): 250--62.
-\url{https://doi.org/10.1111/bjso.12214}.
-
-\bibitem[\citeproctext]{ref-durante_preferences_2009}
-Durante, Ruben, and Louis G. Putterman. 2009. {``Preferences for
+\leavevmode\vadjust pre{\hypertarget{ref-acemoglu_obedience_2021}{}}%
+Acemoglu, D. (2021). \emph{Obedience in the {Labor Market} and {Social
+Mobility}: {A Socio-Economic Approach}} (No. w29125) (p. w29125).
+Cambridge, MA: National Bureau of Economic Research.
+\url{https://doi.org/10.3386/w29125}
+
+\leavevmode\vadjust pre{\hypertarget{ref-almas_fairness_2017}{}}%
+Almås, I., Cappelen, A. W., Salvanes, K. G., Sørensen, E. Ø., \&
+Tungodden, B. (2017). Fairness and family background. \emph{Politics,
+Philosophy \& Economics}, \emph{16}(2), 117--131.
+\url{https://doi.org/10.1177/1470594X15618966}
+
+\leavevmode\vadjust pre{\hypertarget{ref-almas_fairness_2010}{}}%
+Almås, I., Cappelen, A. W., Sørensen, E. Ø., \& Tungodden, B. (2010).
+Fairness and the {Development} of {Inequality Acceptance}.
+\emph{Science}, \emph{328}(5982), 1176--1178.
+\url{https://doi.org/10.1126/science.1187300}
+
+\leavevmode\vadjust pre{\hypertarget{ref-almas_cutthroat_2020}{}}%
+Almås, I., Cappelen, A. W., \& Tungodden, B. (2020). Cutthroat
+{Capitalism} versus {Cuddly Socialism}: {Are Americans More
+Meritocratic} and {Efficiency-Seeking} than {Scandinavians}?
+\emph{Journal of Political Economy}, \emph{128}(5), 1753--1788.
+\url{https://doi.org/10.1086/705551}
+
+\leavevmode\vadjust pre{\hypertarget{ref-barr_effect_2020}{}}%
+Barr, A., \& Miller, L. (2020). The effect of education, income
+inequality and merit on inequality acceptance. \emph{Journal of Economic
+Psychology}, \emph{80}, 102276.
+\url{https://doi.org/10.1016/j.joep.2020.102276}
+
+\leavevmode\vadjust pre{\hypertarget{ref-batruch_belief_2022}{}}%
+Batruch, A., Jetten, J., Van de Werfhorst, H., Darnon, C., \& Butera, F.
+(2022). Belief in {School Meritocracy} and the {Legitimization} of
+{Social} and {Income Inequality}. \emph{Social Psychological and
+Personality Science}, 194855062211110.
+\url{https://doi.org/10.1177/19485506221111017}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bell_fixed_2019}{}}%
+Bell, A., Fairbrother, M., \& Jones, K. (2019). Fixed and random effects
+models: Making an informed choice. \emph{Quality \& Quantity},
+\emph{53}(2), 1051--1074.
+\url{https://doi.org/10.1007/s11135-018-0802-x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-boltanski_new_2005}{}}%
+Boltanski, L., \& Chiapello, E. (2005). \emph{The new spirit of
+capitalism}. London ; New York: Verso.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bourdieu_reproduction_1990}{}}%
+Bourdieu, P., \& Passeron, J. C. (1990). \emph{Reproduction in
+{Education}, {Society} and {Culture}} (Second Edition). Sage
+Publications Ltd.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bucca_merit_2016}{}}%
+Bucca, M. (2016). Merit and blame in unequal societies: {Explaining
+Latin Americans}' beliefs about wealth and poverty. \emph{Research in
+Social Stratification and Mobility}, \emph{44}, 98--112.
+\url{https://doi.org/10.1016/j.rssm.2016.02.005}
+
+\leavevmode\vadjust pre{\hypertarget{ref-castillo_legitimacy_2011}{}}%
+Castillo, J. C. (2011). Legitimacy of {Inequality} in a {Highly Unequal
+Context}: {Evidence} from the {Chilean Case}. \emph{Social Justice
+Research}, \emph{24}(4), 314--340.
+\url{https://doi.org/10.1007/s11211-011-0144-5}
+
+\leavevmode\vadjust pre{\hypertarget{ref-castillo_meritocracia_2019}{}}%
+Castillo, J. C., Torres, A., Atria, J., \& Maldonado, L. (2019).
+Meritocracia y desigualdad econ{ó}mica: {Percepciones}, preferencias e
+implicancias. \emph{Revista Internacional de Sociolog{í}a},
+\emph{77}(1), 117. \url{https://doi.org/10.3989/ris.2019.77.1.17.114}
+
+\leavevmode\vadjust pre{\hypertarget{ref-centeno_arc_2012}{}}%
+Centeno, M. A., \& Cohen, J. N. (2012). The {Arc} of {Neoliberalism}.
+\emph{Annual Review of Sociology}, \emph{38}(1), 317--340.
+\url{https://doi.org/10.1146/annurev-soc-081309-150235}
+
+\leavevmode\vadjust pre{\hypertarget{ref-chafel_schooling_1997}{}}%
+Chafel, J. A. (1997). Schooling, the {Hidden Curriculum}, and
+{Children}'s {Conceptions} of {Poverty}. \emph{Social Policy Report},
+\emph{11}(1), 1--28.
+\url{https://doi.org/10.1002/j.2379-3988.1997.tb00004.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-chong_mystery_2008}{}}%
+Chong, A., \& Ñopo, H. (2008). The {Mystery} of {Discrimination} in
+{Latin America}. \emph{Econom{í}a}, \emph{8}(2), 79--107.
+\url{https://doi.org/10.1353/eco.0.0005}
+
+\leavevmode\vadjust pre{\hypertarget{ref-choudhury_social_2006}{}}%
+Choudhury, S., Blakemore, S.-J., \& Charman, T. (2006). Social cognitive
+development during adolescence. \emph{Social Cognitive and Affective
+Neuroscience}, \emph{1}(3), 165--174.
+\url{https://doi.org/10.1093/scan/nsl024}
+
+\leavevmode\vadjust pre{\hypertarget{ref-darnon_where_2018}{}}%
+Darnon, C., Wiederkehr, V., Dompnier, B., \& Martinot, D. (2018).
+{``{Where} there is a will, there is a way''}: {Belief} in school
+meritocracy and the social-class achievement gap. \emph{British Journal
+of Social Psychology}, \emph{57}(1), 250--262.
+\url{https://doi.org/10.1111/bjso.12214}
+
+\leavevmode\vadjust pre{\hypertarget{ref-durante_preferences_2009}{}}%
+Durante, R., \& Putterman, L. G. (2009). Preferences for
 {Redistribution} and {Perception} of {Fairness}: {An Experimental
-Study}.''} \emph{SSRN Electronic Journal}.
-\url{https://doi.org/10.2139/ssrn.1004573}.
-
-\bibitem[\citeproctext]{ref-duru-bellat_who_2012}
-Duru-Bellat, Marie, and Elise Tenret. 2012. {``Who's for {Meritocracy}?
-{Individual} and {Contextual Variations} in the {Faith}.''}
-\emph{Comparative Education Review} 56 (2): 223--47.
-\url{https://doi.org/10.1086/661290}.
-
-\bibitem[\citeproctext]{ref-elenbaas_unfairness_2019}
-Elenbaas, Laura. 2019. {``Against Unfairness: {Young} Children's
-Judgments about Merit, Equity, and Equality.''} \emph{Journal of
-Experimental Child Psychology} 186 (October): 73--82.
-\url{https://doi.org/10.1016/j.jecp.2019.05.009}.
-
-\bibitem[\citeproctext]{ref-engelmann_children_2019}
-Engelmann, Jan M., and Michael Tomasello. 2019. {``Children's {Sense} of
-{Fairness} as {Equal Respect}.''} \emph{Trends in Cognitive Sciences} 23
-(6): 454--63. \url{https://doi.org/10.1016/j.tics.2019.03.001}.
-
-\bibitem[\citeproctext]{ref-erivwo_meritocracy_2021}
-Erivwo, Aroesiri, Elizabeth Varghese, Anthony Mathai, and Tasmia Afrin.
-2021. {``Meritocracy in the {Educational System},''} May.
-\url{https://doi.org/10.5281/ZENODO.4740695}.
-
-\bibitem[\citeproctext]{ref-evans_strong_2018}
-Evans, M. D. R., and Jonathan Kelley. 2018. {``Strong {Welfare States Do
-Not Intensify Public Support} for {Income Redistribution}, but {Even
-Reduce It} Among the {Prosperous}: {A Multilevel Analysis} of {Public
-Opinion} in 30 {Countries}.''} \emph{Societies} 8 (4): 105.
-\url{https://doi.org/10.3390/soc8040105}.
-
-\bibitem[\citeproctext]{ref-fernandez_positive_2013}
-Fernandez, J. J., and A. M. Jaime-Castillo. 2013. {``Positive or
-{Negative Policy Feedbacks}? {Explaining Popular Attitudes Towards
-Pragmatic Pension Policy Reforms}.''} \emph{European Sociological
-Review} 29 (4): 803--15. \url{https://doi.org/10.1093/esr/jcs059}.
-
-\bibitem[\citeproctext]{ref-frank_performance_2015}
-Frank, Douglas H., Klaus Wertenbroch, and William W. Maddux. 2015.
-{``Performance Pay or Redistribution? {Cultural} Differences in
-Just-World Beliefs and Preferences for Wage Inequality.''}
-\emph{Organizational Behavior and Human Decision Processes} 130
-(September): 160--70. \url{https://doi.org/10.1016/j.obhdp.2015.04.001}.
-
-\bibitem[\citeproctext]{ref-garcia-sanchez_attitudes_2020}
-García-Sánchez, Efraín, Danny Osborne, Guillermo B. Willis, and Rosa
-Rodríguez-Bailón. 2020. {``Attitudes Towards Redistribution and the
-Interplay Between Perceptions and Beliefs about Inequality.''}
-\emph{British Journal of Social Psychology} 59 (1): 111--36.
-\url{https://doi.org/10.1111/bjso.12326}.
-
-\bibitem[\citeproctext]{ref-goldthorpe_myth_2003}
-Goldthorpe, John. 2003. {``The Myth of Education-Based Meritocracy.''}
-\emph{New Economy} 10 (4): 234--39.
-\url{https://doi.org/10.1046/j.1468-0041.2003.00324.x}.
-
-\bibitem[\citeproctext]{ref-goudeau_hidden_2017}
-Goudeau, Sébastien, and Jean-Claude Croizet. 2017. {``Hidden
-{Advantages} and {Disadvantages} of {Social Class}: {How Classroom
-Settings Reproduce Social Inequality} by {Staging Unfair Comparison}.''}
-\emph{Psychological Science} 28 (2): 162--70.
-\url{https://doi.org/10.1177/0956797616676600}.
-
-\bibitem[\citeproctext]{ref-greitemeyer_subjective_2016}
-Greitemeyer, Tobias, and Christina Sagioglou. 2016. {``Subjective
-Socioeconomic Status Causes Aggression: {A} Test of the Theory of Social
-Deprivation.''} \emph{Journal of Personality and Social Psychology} 111
-(2): 178--94. \url{https://doi.org/10.1037/pspi0000058}.
-
-\bibitem[\citeproctext]{ref-harvey_breve_2015}
-Harvey, David. 2015. \emph{{Breve historia del neoliberalismo}}. Madrid
+Study}. \emph{SSRN Electronic Journal}.
+\url{https://doi.org/10.2139/ssrn.1004573}
+
+\leavevmode\vadjust pre{\hypertarget{ref-duru-bellat_who_2012}{}}%
+Duru-Bellat, M., \& Tenret, E. (2012). Who's for {Meritocracy}?
+{Individual} and {Contextual Variations} in the {Faith}.
+\emph{Comparative Education Review}, \emph{56}(2), 223--247.
+\url{https://doi.org/10.1086/661290}
+
+\leavevmode\vadjust pre{\hypertarget{ref-elenbaas_unfairness_2019}{}}%
+Elenbaas, L. (2019). Against unfairness: {Young} children's judgments
+about merit, equity, and equality. \emph{Journal of Experimental Child
+Psychology}, \emph{186}, 73--82.
+\url{https://doi.org/10.1016/j.jecp.2019.05.009}
+
+\leavevmode\vadjust pre{\hypertarget{ref-engelmann_children_2019}{}}%
+Engelmann, J. M., \& Tomasello, M. (2019). Children's {Sense} of
+{Fairness} as {Equal Respect}. \emph{Trends in Cognitive Sciences},
+\emph{23}(6), 454--463. \url{https://doi.org/10.1016/j.tics.2019.03.001}
+
+\leavevmode\vadjust pre{\hypertarget{ref-erivwo_meritocracy_2021}{}}%
+Erivwo, A., Varghese, E., Mathai, A., \& Afrin, T. (2021). Meritocracy
+in the {Educational System}.
+\url{https://doi.org/10.5281/ZENODO.4740695}
+
+\leavevmode\vadjust pre{\hypertarget{ref-evans_strong_2018}{}}%
+Evans, M. D. R., \& Kelley, J. (2018). Strong {Welfare States Do Not
+Intensify Public Support} for {Income Redistribution}, but {Even Reduce
+It} among the {Prosperous}: {A Multilevel Analysis} of {Public Opinion}
+in 30 {Countries}. \emph{Societies}, \emph{8}(4), 105.
+\url{https://doi.org/10.3390/soc8040105}
+
+\leavevmode\vadjust pre{\hypertarget{ref-fernandez_positive_2013}{}}%
+Fernandez, J. J., \& Jaime-Castillo, A. M. (2013). Positive or {Negative
+Policy Feedbacks}? {Explaining Popular Attitudes Towards Pragmatic
+Pension Policy Reforms}. \emph{European Sociological Review},
+\emph{29}(4), 803--815. \url{https://doi.org/10.1093/esr/jcs059}
+
+\leavevmode\vadjust pre{\hypertarget{ref-frank_performance_2015}{}}%
+Frank, D. H., Wertenbroch, K., \& Maddux, W. W. (2015). Performance pay
+or redistribution? {Cultural} differences in just-world beliefs and
+preferences for wage inequality. \emph{Organizational Behavior and Human
+Decision Processes}, \emph{130}, 160--170.
+\url{https://doi.org/10.1016/j.obhdp.2015.04.001}
+
+\leavevmode\vadjust pre{\hypertarget{ref-garcia-sanchez_attitudes_2020}{}}%
+García-Sánchez, E., Osborne, D., Willis, G. B., \& Rodríguez-Bailón, R.
+(2020). Attitudes towards redistribution and the interplay between
+perceptions and beliefs about inequality. \emph{British Journal of
+Social Psychology}, \emph{59}(1), 111--136.
+\url{https://doi.org/10.1111/bjso.12326}
+
+\leavevmode\vadjust pre{\hypertarget{ref-goldthorpe_myth_2003}{}}%
+Goldthorpe, J. (2003). The myth of education-based meritocracy.
+\emph{New Economy}, \emph{10}(4), 234--239.
+\url{https://doi.org/10.1046/j.1468-0041.2003.00324.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-goudeau_hidden_2017}{}}%
+Goudeau, S., \& Croizet, J.-C. (2017). Hidden {Advantages} and
+{Disadvantages} of {Social Class}: {How Classroom Settings Reproduce
+Social Inequality} by {Staging Unfair Comparison}. \emph{Psychological
+Science}, \emph{28}(2), 162--170.
+\url{https://doi.org/10.1177/0956797616676600}
+
+\leavevmode\vadjust pre{\hypertarget{ref-greitemeyer_subjective_2016}{}}%
+Greitemeyer, T., \& Sagioglou, C. (2016). Subjective socioeconomic
+status causes aggression: {A} test of the theory of social deprivation.
+\emph{Journal of Personality and Social Psychology}, \emph{111}(2),
+178--194. \url{https://doi.org/10.1037/pspi0000058}
+
+\leavevmode\vadjust pre{\hypertarget{ref-harvey_breve_2015}{}}%
+Harvey, D. (2015). \emph{{Breve historia del neoliberalismo}}. Madrid
 (Espa{ñ}a): Ediciones Akal.
 
-\bibitem[\citeproctext]{ref-homan_being_2017}
-Homan, Patricia, Lauren Valentino, and Emi Weed. 2017. {``Being and
-{Becoming Poor}: {How Cultural Schemas Shape Beliefs About Poverty}.''}
-\emph{Social Forces} 95 (3): 1023--48.
-\url{https://doi.org/10.1093/sf/sox007}.
-
-\bibitem[\citeproctext]{ref-huppert_development_2019}
-Huppert, Elizabeth, Jason M. Cowell, Yawei Cheng, Carlos
-Contreras-Ibáñez, Natalia Gomez-Sicard, Maria Luz Gonzalez-Gadea, David
-Huepe, et al. 2019. {``The Development of Children's Preferences for
-Equality and Equity Across 13 Individualistic and Collectivist
-Cultures.''} \emph{Developmental Science} 22 (2): e12729.
-\url{https://doi.org/10.1111/desc.12729}.
-
-\bibitem[\citeproctext]{ref-hvidberg_social_2023}
-Hvidberg, Kristoffer B, Claus T Kreiner, and Stefanie Stantcheva. 2023.
-{``Social {Positions} and {Fairness Views} on {Inequality}.''}
-\emph{Review of Economic Studies} 90 (6): 3083--3118.
-\url{https://doi.org/10.1093/restud/rdad019}.
-
-\bibitem[\citeproctext]{ref-iacoviello_collectivism_2019}
-Iacoviello, Vincenzo, and Fabio Lorenzi-Cioldi. 2019. {``Collectivism
-and {Individualism} in {Status Hierarchies}: {Socialization} and {Social
-Identity Explanations}.''} \emph{International Review of Social
-Psychology} 32 (1): 15. \url{https://doi.org/10.5334/irsp.285}.
-
-\bibitem[\citeproctext]{ref-jack_no_2016}
-Jack, Anthony Abraham. 2016. {``({No}) {Harm} in {Asking}: {Class},
-{Acquired Cultural Capital}, and {Academic Engagement} at an {Elite
-University}.''} \emph{Sociology of Education} 89 (1): 1--19.
-\url{https://doi.org/10.1177/0038040715614913}.
-
-\bibitem[\citeproctext]{ref-jasso_how_1999}
-Jasso, Guillermina. 1999. {``How {Much Injustice} Is {There} in the
-{World}? {Two New Justice Indexes}.''} \emph{American Sociological
-Review} 64 (1): 133--68.
-\url{https://doi.org/10.1177/000312249906400110}.
-
-\bibitem[\citeproctext]{ref-jensen_worlds_2008}
-Jensen, Carsten. 2008. {``Worlds of Welfare Services and Transfers.''}
-\emph{Journal of European Social Policy} 18 (2): 151--62.
-\url{https://doi.org/10.1177/0958928707087591}.
-
-\bibitem[\citeproctext]{ref-jonsson_institutional_2015}
-Jonsson, Anna-Carin, and Dennis Beach. 2015. {``Institutional
-Discrimination: {Stereotypes} and Social Reproduction of {`Class'} in
-the {Swedish} Upper-Secondary School.''} \emph{Social Psychology of
-Education} 18 (4): 703--17.
-\url{https://doi.org/10.1007/s11218-014-9279-1}.
-
-\bibitem[\citeproctext]{ref-jost_attitudinal_2000}
-Jost, John T., and Diana Burgess. 2000. {``Attitudinal {Ambivalence} and
-the {Conflict} Between {Group} and {System Justification Motives} in
-{Low Status Groups}.''} \emph{Personality and Social Psychology
-Bulletin} 26 (3): 293--305.
-\url{https://doi.org/10.1177/0146167200265003}.
-
-\bibitem[\citeproctext]{ref-khan_privilege_2011}
-Khan, Shamus. 2011. \emph{Privilege: {The Making} of an {Adolescent
-Elite} at {St}. {Paul}'s {School}}. Princeton, N.J.: Princeton
-University Press.
-
-\bibitem[\citeproctext]{ref-kluegel_beliefs_1987}
-Kluegel, James R., and Eliot R. Smith. 1987. \emph{Beliefs about
-Inequality: {Americans}' Views of What Is and What Ought to Be}. London
-New York: Routledge, Taylor \& Francis Group.
-
-\bibitem[\citeproctext]{ref-kohn_social_1963}
-Kohn, Melvin L. 1963. {``Social {Class} and {Parent-Child
-Relationships}: {An Interpretation}.''} \emph{American Journal of
-Sociology} 68 (4): 471--80. \url{https://doi.org/10.1086/223403}.
-
-\bibitem[\citeproctext]{ref-kohn_class_1969}
-Kohn, Melvin L., and Carmi Schooler. 1969. {``Class, {Occupation}, and
-{Orientation}.''} \emph{American Sociological Review} 34 (5): 659.
-\url{https://doi.org/10.2307/2092303}.
-
-\bibitem[\citeproctext]{ref-kosse_prosociality_2020}
-Kosse, Fabian, and Michela M. Tincani. 2020. {``Prosociality Predicts
-Labor Market Success Around the World.''} \emph{Nature Communications}
-11 (1): 5298. \url{https://doi.org/10.1038/s41467-020-19007-1}.
-
-\bibitem[\citeproctext]{ref-lampert_meritocratic_2013}
-Lampert, Khen. 2013. \emph{Meritocratic {Education} and {Social
+\leavevmode\vadjust pre{\hypertarget{ref-homan_being_2017}{}}%
+Homan, P., Valentino, L., \& Weed, E. (2017). Being and {Becoming Poor}:
+{How Cultural Schemas Shape Beliefs About Poverty}. \emph{Social
+Forces}, \emph{95}(3), 1023--1048.
+\url{https://doi.org/10.1093/sf/sox007}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hox_multilevel_2010}{}}%
+Hox, J., Moerbeek, M., \& Schoot, R. van de. (2010). \emph{Multilevel
+{Analysis}: {Techniques} and {Applications}, {Second Edition}} (2nd
+ed.). New York: Routledge. \url{https://doi.org/10.4324/9780203852279}
+
+\leavevmode\vadjust pre{\hypertarget{ref-huppert_development_2019}{}}%
+Huppert, E., Cowell, J. M., Cheng, Y., Contreras-Ibáñez, C.,
+Gomez-Sicard, N., Gonzalez-Gadea, M. L., \ldots{} Decety, J. (2019). The
+development of children's preferences for equality and equity across 13
+individualistic and collectivist cultures. \emph{Developmental Science},
+\emph{22}(2), e12729. \url{https://doi.org/10.1111/desc.12729}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hvidberg_social_2023}{}}%
+Hvidberg, K. B., Kreiner, C. T., \& Stantcheva, S. (2023). Social
+{Positions} and {Fairness Views} on {Inequality}. \emph{Review of
+Economic Studies}, \emph{90}(6), 3083--3118.
+\url{https://doi.org/10.1093/restud/rdad019}
+
+\leavevmode\vadjust pre{\hypertarget{ref-iacoviello_collectivism_2019}{}}%
+Iacoviello, V., \& Lorenzi-Cioldi, F. (2019). Collectivism and
+{Individualism} in {Status Hierarchies}: {Socialization} and {Social
+Identity Explanations}. \emph{International Review of Social
+Psychology}, \emph{32}(1), 15. \url{https://doi.org/10.5334/irsp.285}
+
+\leavevmode\vadjust pre{\hypertarget{ref-jack_no_2016}{}}%
+Jack, A. A. (2016). ({No}) {Harm} in {Asking}: {Class}, {Acquired
+Cultural Capital}, and {Academic Engagement} at an {Elite University}.
+\emph{Sociology of Education}, \emph{89}(1), 1--19.
+\url{https://doi.org/10.1177/0038040715614913}
+
+\leavevmode\vadjust pre{\hypertarget{ref-jasso_how_1999}{}}%
+Jasso, G. (1999). How {Much Injustice} is {There} in the {World}? {Two
+New Justice Indexes}. \emph{American Sociological Review}, \emph{64}(1),
+133--168. \url{https://doi.org/10.1177/000312249906400110}
+
+\leavevmode\vadjust pre{\hypertarget{ref-jensen_worlds_2008}{}}%
+Jensen, C. (2008). Worlds of welfare services and transfers.
+\emph{Journal of European Social Policy}, \emph{18}(2), 151--162.
+\url{https://doi.org/10.1177/0958928707087591}
+
+\leavevmode\vadjust pre{\hypertarget{ref-jonsson_institutional_2015}{}}%
+Jonsson, A.-C., \& Beach, D. (2015). Institutional discrimination:
+{Stereotypes} and social reproduction of {``class''} in the {Swedish}
+upper-secondary school. \emph{Social Psychology of Education},
+\emph{18}(4), 703--717. \url{https://doi.org/10.1007/s11218-014-9279-1}
+
+\leavevmode\vadjust pre{\hypertarget{ref-jost_attitudinal_2000}{}}%
+Jost, J. T., \& Burgess, D. (2000). Attitudinal {Ambivalence} and the
+{Conflict} between {Group} and {System Justification Motives} in {Low
+Status Groups}. \emph{Personality and Social Psychology Bulletin},
+\emph{26}(3), 293--305. \url{https://doi.org/10.1177/0146167200265003}
+
+\leavevmode\vadjust pre{\hypertarget{ref-khan_privilege_2011}{}}%
+Khan, S. (2011). \emph{Privilege: {The Making} of an {Adolescent Elite}
+at {St}. {Paul}'s {School}}. Princeton, N.J.: Princeton University
+Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kluegel_beliefs_1987}{}}%
+Kluegel, J. R., \& Smith, E. R. (1987). \emph{Beliefs about inequality:
+{Americans}' views of what is and what ought to be}. London New York:
+Routledge, Taylor \& Francis Group.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kohn_social_1963}{}}%
+Kohn, M. L. (1963). Social {Class} and {Parent-Child Relationships}: {An
+Interpretation}. \emph{American Journal of Sociology}, \emph{68}(4),
+471--480. \url{https://doi.org/10.1086/223403}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kohn_class_1969}{}}%
+Kohn, M. L., \& Schooler, C. (1969). Class, {Occupation}, and
+{Orientation}. \emph{American Sociological Review}, \emph{34}(5), 659.
+\url{https://doi.org/10.2307/2092303}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kosse_prosociality_2020}{}}%
+Kosse, F., \& Tincani, M. M. (2020). Prosociality predicts labor market
+success around the world. \emph{Nature Communications}, \emph{11}(1),
+5298. \url{https://doi.org/10.1038/s41467-020-19007-1}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lampert_meritocratic_2013}{}}%
+Lampert, K. (2013). \emph{Meritocratic {Education} and {Social
 Worthlessness}}. London: Palgrave Macmillan UK.
-\url{https://doi.org/10.1057/9781137324894}.
-
-\bibitem[\citeproctext]{ref-lane_market_1986}
-Lane, Robert E. 1986. {``Market {Justice}, {Political Justice}.''}
-\emph{American Political Science Review} 80 (2): 383--402.
-\url{https://doi.org/10.2307/1958264}.
-
-\bibitem[\citeproctext]{ref-lindh_public_2015}
-Lindh, Arvid. 2015. {``Public {Opinion} Against {Markets}? {Attitudes}
-Towards {Market Distribution} of {Social Services} -- {A Comparison} of
-17 {Countries}.''} \emph{Social Policy \& Administration} 49 (7):
-887--910. \url{https://doi.org/10.1111/spol.12105}.
-
-\bibitem[\citeproctext]{ref-mcauliffe_developmental_2017}
-McAuliffe, Katherine, Peter R. Blake, Nikolaus Steinbeis, and Felix
-Warneken. 2017. {``The Developmental Foundations of Human Fairness.''}
-\emph{Nature Human Behaviour} 1 (2): 0042.
-\url{https://doi.org/10.1038/s41562-016-0042}.
-
-\bibitem[\citeproctext]{ref-mijs_stratified_2016}
-Mijs, Jonathan. 2016. {``Stratified {Failure}: {Educational
-Stratification} and {Students}' {Attributions} of {Their Mathematics
-Performance} in 24 {Countries}.''} \emph{Sociology of Education} 89 (2):
-137--53. \url{https://doi.org/10.1177/0038040716636434}.
-
-\bibitem[\citeproctext]{ref-mijs_paradox_2021}
----------. 2021. {``The Paradox of Inequality: Income Inequality and
-Belief in Meritocracy Go Hand in Hand.''} \emph{Socio-Economic Review}
-19 (1): 7--35. \url{https://doi.org/10.1093/ser/mwy051}.
-
-\bibitem[\citeproctext]{ref-mikus_children_2020}
-Mikus, Karoline, Nicole Tieben, and Pia S. Schober. 2020. {``Children's
-Conversion of Cultural Capital into Educational Success: The Symbolic
-and Skill-Generating Functions of Cultural Capital.''} \emph{British
-Journal of Sociology of Education} 41 (2): 197--217.
-\url{https://doi.org/10.1080/01425692.2019.1677454}.
-
-\bibitem[\citeproctext]{ref-mishra_subjective_2015}
-Mishra, Sandeep, and R. Nicholas Carleton. 2015. {``Subjective Relative
-Deprivation Is Associated with Poorer Physical and Mental Health.''}
-\emph{Social Science \& Medicine} 147 (December): 144--49.
-\url{https://doi.org/10.1016/j.socscimed.2015.10.030}.
-
-\bibitem[\citeproctext]{ref-molyneux_neoliberal_2008}
-Molyneux, Maxine. 2008. {``The {`{Neoliberal Turn}'} and the {New Social
-Policy} in {Latin America}: {How Neoliberal}, {How New}?''}
-\emph{Development and Change} 39 (5): 775--97.
-\url{https://doi.org/10.1111/j.1467-7660.2008.00505.x}.
-
-\bibitem[\citeproctext]{ref-morris_representing_2022}
-Morris, Jennifer, John Reilly, Sergey Paltsev, Andrei Sokolov, and
-Kenneth Cox. 2022. {``Representing {Socio}-{Economic Uncertainty} in
-{Human System Models}.''} \emph{Earth's Future} 10 (4): e2021EF002239.
-\url{https://doi.org/10.1029/2021EF002239}.
-
-\bibitem[\citeproctext]{ref-pierson_increasing_2000}
-Pierson, Paul. 2000. {``Increasing {Returns}, {Path Dependence}, and the
-{Study} of {Politics}.''} \emph{American Political Science Review} 94
-(2): 251--67. \url{https://doi.org/10.2307/2586011}.
-
-\bibitem[\citeproctext]{ref-power_deprivationprotest_2018}
-Power, Séamus A. 2018. {``The {Deprivation-Protest Paradox}: {How} the
-{Perception} of {Unfair Economic Inequality Leads} to {Civic Unrest}.''}
-\emph{Current Anthropology} 59 (6): 765--89.
-\url{https://doi.org/10.1086/700679}.
-
-\bibitem[\citeproctext]{ref-power_relative_2020}
-Power, Séamus A, Thomas Madsen, and Thomas A Morton. 2020. {``Relative
-Deprivation and Revolt: Current and Future Directions.''} \emph{Current
-Opinion in Psychology} 35 (October): 119--24.
-\url{https://doi.org/10.1016/j.copsyc.2020.06.010}.
-
-\bibitem[\citeproctext]{ref-resh_sense_2010}
-Resh, Nura. 2010. {``Sense of Justice about Grades in School: Is It
-Stratified Like Academic Achievement?''} \emph{Social Psychology of
-Education} 13 (3): 313--29.
-\url{https://doi.org/10.1007/s11218-010-9117-z}.
-
-\bibitem[\citeproctext]{ref-resh_sense_2014}
-Resh, Nura, and Clara Sabbagh. 2014. {``Sense of Justice in School and
-Civic Attitudes.''} \emph{Social Psychology of Education} 17 (1):
-51--72. \url{https://doi.org/10.1007/s11218-013-9240-8}.
-
-\bibitem[\citeproctext]{ref-resh_sense_2017}
----------. 2017. {``Sense of Justice in School and Civic Behavior.''}
-\emph{Social Psychology of Education} 20 (2): 387--409.
-\url{https://doi.org/10.1007/s11218-017-9375-0}.
-
-\bibitem[\citeproctext]{ref-salamon_marketization_1993}
-Salamon, Lester M. 1993. {``The {Marketization} of {Welfare}: {Changing
-Nonprofit} and {For-Profit Roles} in the {American Welfare State}.''}
-\emph{Social Service Review} 67 (1): 16--39.
-\url{https://doi.org/10.1086/603963}.
-
-\bibitem[\citeproctext]{ref-salgado_inequality_2023}
-Salgado, Mauricio, and Javier Castillo. 2023. {``Inequality and
-{Stratification} in {Latin America}.''} In \emph{The {Oxford Handbook}
-of {Social Stratification}}, edited by Markus Gangl, Lucinda Platt,
-Javier G. Polavieja, and Herman G. van de Werfhorst, 0. 2023: Oxford
-University Press.
-
-\bibitem[\citeproctext]{ref-sandel_tyranny_2020}
-Sandel, Michael J. 2020. \emph{The Tyranny of Merit: {What}'s Become of
-the Common Good?} First edition. New York: {Farrar, Straus and Giroux}.
-
-\bibitem[\citeproctext]{ref-schunk_fairness_2023}
-Schunk, Daniel, and Isabell Zipperle. 2023. {``Fairness and Inequality
-Acceptance in Children and Adolescents: {A} Survey on Behaviors in
-Economic Experiments''} 37 (5): 1715--42.
-\url{https://doi.org/10.1111/joes.12553}.
-
-\bibitem[\citeproctext]{ref-shariff_income_2016}
-Shariff, Azim F., Dylan Wiwad, and Lara B. Aknin. 2016. {``Income
-{Mobility Breeds Tolerance} for {Income Inequality}: {Cross-National}
-and {Experimental Evidence}.''} \emph{Perspectives on Psychological
-Science} 11 (3): 373--80.
-\url{https://doi.org/10.1177/1745691616635596}.
-
-\bibitem[\citeproctext]{ref-sigelman_development_1991}
-Sigelman, Carol K., and Kara A. Waitzman. 1991. {``The {Development} of
+\url{https://doi.org/10.1057/9781137324894}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lane_market_1986}{}}%
+Lane, R. E. (1986). Market {Justice}, {Political Justice}.
+\emph{American Political Science Review}, \emph{80}(2), 383--402.
+\url{https://doi.org/10.2307/1958264}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lindh_public_2015}{}}%
+Lindh, A. (2015). Public {Opinion} against {Markets}? {Attitudes}
+towards {Market Distribution} of {Social Services} -- {A Comparison} of
+17 {Countries}. \emph{Social Policy \& Administration}, \emph{49}(7),
+887--910. \url{https://doi.org/10.1111/spol.12105}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mcauliffe_developmental_2017}{}}%
+McAuliffe, K., Blake, P. R., Steinbeis, N., \& Warneken, F. (2017). The
+developmental foundations of human fairness. \emph{Nature Human
+Behaviour}, \emph{1}(2), 0042.
+\url{https://doi.org/10.1038/s41562-016-0042}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mijs_stratified_2016}{}}%
+Mijs, J. (2016). Stratified {Failure}: {Educational Stratification} and
+{Students}' {Attributions} of {Their Mathematics Performance} in 24
+{Countries}. \emph{Sociology of Education}, \emph{89}(2), 137--153.
+\url{https://doi.org/10.1177/0038040716636434}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mijs_paradox_2021}{}}%
+Mijs, J. (2021). The paradox of inequality: Income inequality and belief
+in meritocracy go hand in hand. \emph{Socio-Economic Review},
+\emph{19}(1), 7--35. \url{https://doi.org/10.1093/ser/mwy051}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mikus_children_2020}{}}%
+Mikus, K., Tieben, N., \& Schober, P. S. (2020). Children's conversion
+of cultural capital into educational success: The symbolic and
+skill-generating functions of cultural capital. \emph{British Journal of
+Sociology of Education}, \emph{41}(2), 197--217.
+\url{https://doi.org/10.1080/01425692.2019.1677454}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mishra_subjective_2015}{}}%
+Mishra, S., \& Carleton, R. N. (2015). Subjective relative deprivation
+is associated with poorer physical and mental health. \emph{Social
+Science \& Medicine}, \emph{147}, 144--149.
+\url{https://doi.org/10.1016/j.socscimed.2015.10.030}
+
+\leavevmode\vadjust pre{\hypertarget{ref-molyneux_neoliberal_2008}{}}%
+Molyneux, M. (2008). The {``{Neoliberal Turn}''} and the {New Social
+Policy} in {Latin America}: {How Neoliberal}, {How New}?
+\emph{Development and Change}, \emph{39}(5), 775--797.
+\url{https://doi.org/10.1111/j.1467-7660.2008.00505.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-morris_representing_2022}{}}%
+Morris, J., Reilly, J., Paltsev, S., Sokolov, A., \& Cox, K. (2022).
+Representing {Socio}-{Economic Uncertainty} in {Human System Models}.
+\emph{Earth's Future}, \emph{10}(4), e2021EF002239.
+\url{https://doi.org/10.1029/2021EF002239}
+
+\leavevmode\vadjust pre{\hypertarget{ref-pierson_increasing_2000}{}}%
+Pierson, P. (2000). Increasing {Returns}, {Path Dependence}, and the
+{Study} of {Politics}. \emph{American Political Science Review},
+\emph{94}(2), 251--267. \url{https://doi.org/10.2307/2586011}
+
+\leavevmode\vadjust pre{\hypertarget{ref-power_deprivationprotest_2018}{}}%
+Power, Séamus A. (2018). The {Deprivation-Protest Paradox}: {How} the
+{Perception} of {Unfair Economic Inequality Leads} to {Civic Unrest}.
+\emph{Current Anthropology}, \emph{59}(6), 765--789.
+\url{https://doi.org/10.1086/700679}
+
+\leavevmode\vadjust pre{\hypertarget{ref-power_relative_2020}{}}%
+Power, Séamus A., Madsen, T., \& Morton, T. A. (2020). Relative
+deprivation and revolt: Current and future directions. \emph{Current
+Opinion in Psychology}, \emph{35}, 119--124.
+\url{https://doi.org/10.1016/j.copsyc.2020.06.010}
+
+\leavevmode\vadjust pre{\hypertarget{ref-resh_sense_2010}{}}%
+Resh, N. (2010). Sense of justice about grades in school: Is it
+stratified like academic achievement? \emph{Social Psychology of
+Education}, \emph{13}(3), 313--329.
+\url{https://doi.org/10.1007/s11218-010-9117-z}
+
+\leavevmode\vadjust pre{\hypertarget{ref-resh_sense_2014}{}}%
+Resh, N., \& Sabbagh, C. (2014). Sense of justice in school and civic
+attitudes. \emph{Social Psychology of Education}, \emph{17}(1), 51--72.
+\url{https://doi.org/10.1007/s11218-013-9240-8}
+
+\leavevmode\vadjust pre{\hypertarget{ref-resh_sense_2017}{}}%
+Resh, N., \& Sabbagh, C. (2017). Sense of justice in school and civic
+behavior. \emph{Social Psychology of Education}, \emph{20}(2), 387--409.
+\url{https://doi.org/10.1007/s11218-017-9375-0}
+
+\leavevmode\vadjust pre{\hypertarget{ref-salamon_marketization_1993}{}}%
+Salamon, L. M. (1993). The {Marketization} of {Welfare}: {Changing
+Nonprofit} and {For-Profit Roles} in the {American Welfare State}.
+\emph{Social Service Review}, \emph{67}(1), 16--39.
+\url{https://doi.org/10.1086/603963}
+
+\leavevmode\vadjust pre{\hypertarget{ref-salgado_inequality_2023}{}}%
+Salgado, M., \& Castillo, J. (2023). Inequality and {Stratification} in
+{Latin America}. In M. Gangl, L. Platt, J. G. Polavieja, \& H. G. van de
+Werfhorst (Eds.), \emph{The {Oxford Handbook} of {Social
+Stratification}} (p. 0). 2023: Oxford University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sandel_tyranny_2020}{}}%
+Sandel, M. J. (2020). \emph{The tyranny of merit: {What}'s become of the
+common good?} (First edition). New York: {Farrar, Straus and Giroux}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-schunk_fairness_2023}{}}%
+Schunk, D., \& Zipperle, I. (2023). Fairness and inequality acceptance
+in children and adolescents: {A} survey on behaviors in economic
+experiments, \emph{37}(5), 1715--1742.
+\url{https://doi.org/10.1111/joes.12553}
+
+\leavevmode\vadjust pre{\hypertarget{ref-shariff_income_2016}{}}%
+Shariff, A. F., Wiwad, D., \& Aknin, L. B. (2016). Income {Mobility
+Breeds Tolerance} for {Income Inequality}: {Cross-National} and
+{Experimental Evidence}. \emph{Perspectives on Psychological Science},
+\emph{11}(3), 373--380. \url{https://doi.org/10.1177/1745691616635596}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sigelman_development_1991}{}}%
+Sigelman, C. K., \& Waitzman, K. A. (1991). The {Development} of
 {Distributive Justice Orientations}: {Contextual Influences} on
-{Children}'s {Resource Allocations}.''} \emph{Child Development} 62 (6):
-1367. \url{https://doi.org/10.2307/1130812}.
-
-\bibitem[\citeproctext]{ref-sivesind_changing_2017}
-Sivesind, Karl Henrik. 2017. {``The {Changing Roles} of {For-Profit} and
-{Nonprofit Welfare Provision} in {Norway}, {Sweden}, and {Denmark}.''}
-In \emph{Promoting {Active Citizenship}}, edited by Karl Henrik Sivesind
-and Jo Saglie, 33--74. Cham: Springer International Publishing.
-\url{https://doi.org/10.1007/978-3-319-55381-8_2}.
-
-\bibitem[\citeproctext]{ref-smith_relative_2012}
-Smith, Heather J., Thomas F. Pettigrew, Gina M. Pippin, and Silvana
-Bialosiewicz. 2012. {``Relative {Deprivation}: {A Theoretical} and
-{Meta-Analytic Review}.''} \emph{Personality and Social Psychology
-Review} 16 (3): 203--32. \url{https://doi.org/10.1177/1088868311430825}.
-
-\bibitem[\citeproctext]{ref-soler-martinez_concerns_2023}
-Soler-Martínez, Francisco Miguel, Efraín García-Sánchez, and Guillermo
-B. Willis. 2023. {``Concerns {About Inequality} in {Health}, {Education}
-and {Income Jointly Predict Collective Actions}.''} \emph{Revista
-Latinoamericana de Psicolog{í}a} 55: 99.
-\url{https://doi.org/10.14349/rlp.2023.v55.12}.
-
-\bibitem[\citeproctext]{ref-stoy_worlds_2014}
-Stoy, Volquart. 2014. {``Worlds of {Welfare Services}: {From Discovery}
-to {Exploration}.''} \emph{Social Policy \& Administration} 48 (3):
-343--60. \url{https://doi.org/10.1111/spol.12006}.
-
-\bibitem[\citeproctext]{ref-streeck_citizens_2012}
-Streeck, Wolfgang. 2012. {``Citizens as {Customers}: {Considerations} on
-the {New Politics} of {Consumption}.''} \emph{New Left Review} 76.
-
-\bibitem[\citeproctext]{ref-torche_intergenerational_2014}
-Torche, Florencia. 2014. {``Intergenerational {Mobility} and
-{Inequality}: {The Latin American Case}.''} \emph{Annual Review of
-Sociology} 40 (1): 619--42.
-\url{https://doi.org/10.1146/annurev-soc-071811-145521}.
-
-\bibitem[\citeproctext]{ref-vonhippel_test_2019}
-Von Hippel, Paul, and Caitlin Hamrock. 2019. {``Do {Test Score Gaps
-Grow} Before, During, or Between the {School Years}? {Measurement
-Artifacts} and {What We Can Know} in {Spite} of {Them}.''}
-\emph{Sociological Science} 6: 43--80.
-\url{https://doi.org/10.15195/v6.a3}.
-
-\bibitem[\citeproctext]{ref-weaver_paths_2010}
-Weaver, Kent. 2010. {``Paths and {Forks} or {Chutes} and {Ladders}?:
-{Negative Feedbacks} and {Policy Regime Change}.''} \emph{Journal of
-Public Policy} 30 (2): 137--62.
-\url{https://doi.org/10.1017/S0143814X10000061}.
-
-\bibitem[\citeproctext]{ref-wegener_dominant_1995}
-Wegener, Bernd, and Stefan Liebig. 1995. {``Dominant {Ideologies} and
-the {Variation} of {Distributive Justice Norms}: {A Comparison} of
-{East} and {West Germany}, and the {United States}.''} In \emph{Social
-{Justice} and {Political Change}: {Public Opinion} in {Capitalist} and
-{Post-Communist States}}, edited by James R. Kluegel, David S. Mason,
-and Bernd Wegener, 239--59. New York: Walter de Gruyter.
-
-\bibitem[\citeproctext]{ref-wiederkehr_belief_2015}
-Wiederkehr, Virginie, Virginie Bonnot, Silvia Krauth-Gruber, and Céline
-Darnon. 2015. {``Belief in School Meritocracy as a System-Justifying
-Tool for Low Status Students.''} \emph{Frontiers in Psychology} 6.
-
-\bibitem[\citeproctext]{ref-wynn_not_2018}
-Wynn, Karen, Paul Bloom, Ashley Jordan, Julia Marshall, and Mark
-Sheskin. 2018. {``Not {Noble Savages After All}: {Limits} to {Early
-Altruism}.''} \emph{Current Directions in Psychological Science} 27 (1):
-3--8. \url{https://doi.org/10.1177/0963721417734875}.
-
-\bibitem[\citeproctext]{ref-young_rise_1958}
-Young, Michael. 1958. \emph{The Rise of the Meritocracy}. New Brunswick,
+{Children}'s {Resource Allocations}. \emph{Child Development},
+\emph{62}(6), 1367. \url{https://doi.org/10.2307/1130812}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sivesind_changing_2017}{}}%
+Sivesind, K. H. (2017). The {Changing Roles} of {For-Profit} and
+{Nonprofit Welfare Provision} in {Norway}, {Sweden}, and {Denmark}. In
+K. H. Sivesind \& J. Saglie (Eds.), \emph{Promoting {Active
+Citizenship}} (pp. 33--74). Cham: Springer International Publishing.
+\url{https://doi.org/10.1007/978-3-319-55381-8_2}
+
+\leavevmode\vadjust pre{\hypertarget{ref-smith_relative_2012}{}}%
+Smith, H. J., Pettigrew, T. F., Pippin, G. M., \& Bialosiewicz, S.
+(2012). Relative {Deprivation}: {A Theoretical} and {Meta-Analytic
+Review}. \emph{Personality and Social Psychology Review}, \emph{16}(3),
+203--232. \url{https://doi.org/10.1177/1088868311430825}
+
+\leavevmode\vadjust pre{\hypertarget{ref-soler-martinez_concerns_2023}{}}%
+Soler-Martínez, F. M., García-Sánchez, E., \& Willis, G. B. (2023).
+Concerns {About Inequality} in {Health}, {Education} and {Income Jointly
+Predict Collective Actions}. \emph{Revista Latinoamericana de
+Psicolog{í}a}, \emph{55}, 99.
+\url{https://doi.org/10.14349/rlp.2023.v55.12}
+
+\leavevmode\vadjust pre{\hypertarget{ref-stoy_worlds_2014}{}}%
+Stoy, V. (2014). Worlds of {Welfare Services}: {From Discovery} to
+{Exploration}. \emph{Social Policy \& Administration}, \emph{48}(3),
+343--360. \url{https://doi.org/10.1111/spol.12006}
+
+\leavevmode\vadjust pre{\hypertarget{ref-streeck_citizens_2012}{}}%
+Streeck, W. (2012). Citizens as {Customers}: {Considerations} on the
+{New Politics} of {Consumption}. \emph{New Left Review}, \emph{76}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-torche_intergenerational_2014}{}}%
+Torche, F. (2014). Intergenerational {Mobility} and {Inequality}: {The
+Latin American Case}. \emph{Annual Review of Sociology}, \emph{40}(1),
+619--642. \url{https://doi.org/10.1146/annurev-soc-071811-145521}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vonhippel_test_2019}{}}%
+Von Hippel, P., \& Hamrock, C. (2019). Do {Test Score Gaps Grow} before,
+during, or between the {School Years}? {Measurement Artifacts} and {What
+We Can Know} in {Spite} of {Them}. \emph{Sociological Science},
+\emph{6}, 43--80. \url{https://doi.org/10.15195/v6.a3}
+
+\leavevmode\vadjust pre{\hypertarget{ref-weaver_paths_2010}{}}%
+Weaver, K. (2010). Paths and {Forks} or {Chutes} and {Ladders}?:
+{Negative Feedbacks} and {Policy Regime Change}. \emph{Journal of Public
+Policy}, \emph{30}(2), 137--162.
+\url{https://doi.org/10.1017/S0143814X10000061}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wegener_dominant_1995}{}}%
+Wegener, B., \& Liebig, S. (1995). Dominant {Ideologies} and the
+{Variation} of {Distributive Justice Norms}: {A Comparison} of {East}
+and {West Germany}, and the {United States}. In J. R. Kluegel, D. S.
+Mason, \& B. Wegener (Eds.), \emph{Social {Justice} and {Political
+Change}: {Public Opinion} in {Capitalist} and {Post-Communist States}}
+(pp. 239--259). New York: Walter de Gruyter.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wiederkehr_belief_2015}{}}%
+Wiederkehr, V., Bonnot, V., Krauth-Gruber, S., \& Darnon, C. (2015).
+Belief in school meritocracy as a system-justifying tool for low status
+students. \emph{Frontiers in Psychology}, \emph{6}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wynn_not_2018}{}}%
+Wynn, K., Bloom, P., Jordan, A., Marshall, J., \& Sheskin, M. (2018).
+Not {Noble Savages After All}: {Limits} to {Early Altruism}.
+\emph{Current Directions in Psychological Science}, \emph{27}(1), 3--8.
+\url{https://doi.org/10.1177/0963721417734875}
+
+\leavevmode\vadjust pre{\hypertarget{ref-young_rise_1958}{}}%
+Young, M. (1958). \emph{The rise of the meritocracy}. New Brunswick,
 N.J., U.S.A: Transaction Publishers.

### 2024-05-15 paper.qmd
render paper

@@
 ---
-title: "market-justice-preferences"
-author: "Equipo EDUMER"
-format: pdf
+format:
+  html:
+    keep-tex: true
 editor: visual
 bibliography: "input/bib/just-ineq-merit.bib"
+csl: "input/bib/apa6.csl"
 link-citations: true
+mainfont: Times New Roman
 ---
@@
 
-##
+{{< include 01-introduction.qmd >}}
+
+{{< include 02-antecedentes.qmd >}}
+
+{{< include 03-methods.qmd >}}
+
+{{< include 04-analisis.qmd >}}
+
+{{< include 05-conclusiones.qmd >}}
+
+{{< include 06-anexos.qmd >}}

### 2024-05-16 01-introduction.qmd
update plots and tables antecedentes y metodos
intro update

--- 2024-04-24 01-introduction.qmd
@@
 author: "Equipo EDUMER"
-format: pdf
+format: html
 editor: visual
+bibliography: "input/bib/just-ineq-merit.bib"
+csl: "input/bib/apa6.csl"
 ---
@@
 Therefore, the consistent evidence of the high level of social reproduction at the school level represents a threat to the promise of education and a meritocratic system [@bourdieu_reproduction_1990].
-A large part of the research in this field at an international level has addressed the extent to which the social origin of students affects their academic results and their life opportunities [@vonhippel_test_2019], confirming that schools have severe difficulties in closing the gaps of origin.
-Besides this socioeconomic perspective on school opportunities, recent research has addressed to what extent inequalities in the school context are also influencing students' perceptions, beliefs, and attitudes: Are social inequalities perceived at the school context?
+A large part of the research in this field at an international level has addressed the extent to which the social origin of students affects their academic results and their life opportunities [@vonhippel_test_2019], confirming that schools have severe difficulties in closing the socio-economic and cultural gaps of origin.
+Besides this socioeconomic perspective on school opportunities, recent research has addressed to what extent inequalities in the school context are also influencing students' perceptions, beliefs, and attitudes: Are social inequalities even perceived at the school context?
 Are they rejected by the students, particularly those who are worst-off in socioeconomic terms?
@@
 [@batruch_belief_2022; @wiederkehr_belief_2015].
-
-In the present paper, we deal with the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system.
-In particular, we focus on market justice preferences (Lane, 1986), which refer to the preferences for distributing public goods (as health and education) based on criteria such as competition and payment capacity.
-Although from a rational point of view it could be expected an opposition to market justice by the underprivileged majority,  we argue that in a social environment characterized by a promotion of meritocratic ideals - as the school system - would lead to the opposite: a larger market justice preferences.
 
@@
 In this regard, a great deal of academic research about meritocracy delves into the assessment of to what extent rewards and privileges in society are related to merit, emphasizing the so-called unfulfillable promise of meritocracy [@mijs_stratified_2016].
-Complementing this agenda, a second and emerging research area deals with subjective aspects of meritocracy, such as perceptions and beliefs.
 
-\[definición de market justice, welfare marketization, (Lindh)\]
+In the present paper, we address the role of the perception of meritocracy on the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system.
+In particular, we focus in justifying inequalities in health, education and pensions, which are traditional social policy areas where acces to better services could be justified based on payment capacity, referred as to market justice [@lane_market_1986, @lindh_bringing_2023].
+Most of the research in this realm has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies [@batruch_belief_2022].
+Regarding schools, the way in which they deal with unequal conditions of origin has been linked to the *hidden curriculum* [@chafel_schooling_1997], whereby students learn about distributive norms in society and mechanisms of justification of social differences.
 
-\[contexto Chileno market justice, dar cifras/porcentajes de privatización de salud, educación y pensiones; asociar al tema de la calidad (lo público es peor)\]
+Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population [@batruch_belief_2022; @wiederkehr_belief_2015], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger justification of social inequalities, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary [@batruch_belief_2022].
+We focus on the student-age population as we point out that it is possible to track down the origin of meritocratic beliefs (and their consequences) to early socialization processes.
+To this regard, we take into account the family and the school as two main socialization agencies that play a significant role in the socialization of cultural beliefs by transmitting cultural norms, values, and expectations to young people.
 
-The perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society [@duru-bellat_who_2012; @castillo_meritocracia_2019].
-This perception can vary greatly depending on individual experiences, social, economic, and cultural background.
-Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work.
-In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2021].
-Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger justification of social inequalities, as individual achievement would be seen as rewarded and social policies as less necessary [@batruch_belief_2022].
-
-Most of the research that has related meritocratic beliefs to inequality justification so far has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies.
-Regarding schools, the way in which they deal with unequal conditions of origin has been linked to the *hidden curriculum* [@chafel_schooling_1997], whereby students learn about distributive norms in society and mechanisms of justification of social differences.
-Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population [@batruch_belief_2022; @wiederkehr_belief_2015], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger justification of social inequalities, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary [@batruch_belief_2022].
-
-\~
-
-\[introducir distinción sobre percepción de meritocracia a nivel escolar y a nivel de la sociedad\]
-
-\[resaltar 3 focos de innovación: market justice preferences at school level, relación con meritocracia, y esquema de meritocracia en la escuela y en la sociedad\]
-
-The present paper deals with the association between the perception of meritocracy and the justification of social inequalities, with two main focuses.
-Firstly, it assesses the justification of social inequalities in social policy domains, such as health, pensions, and education.
-We argue that individuals who perceive more meritocracy would be more willing to justify better services in these domains for those with higher incomes.  Secondly, we focus on the student-age population as we point out that it is possible to track down the origin of meritocratic beliefs (and their consequences) to early socialization processes.
-To this regard, we take into account the family and the school as two main socialization agencies that play a significant role in the socialization of cultural beliefs by transmitting cultural norms, values, and expectations to young people.
+++ párrafo contexto Chileno dar cifras/porcentajes de privatización de salud, educación y pensiones; asociar al tema de la calidad (lo público es peor); hay información sobre esto que se puede sacar del proyecto Fondecyt

### 2024-05-16 02-antecedentes.qmd
intro update

@@
 editor: visual
+bibliography: "input/bib/just-ineq-merit.bib"
+csl: "input/bib/apa6.csl"
 ---
@@
 
+{++_Esto viene de arriba, contextualizar_ The perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society [@duru-bellat_who_2012; @castillo_meritocracia_2019].
+This perception can vary greatly depending on individual experiences, social, economic, and cultural background.
+Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work.
+In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2021].
+Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger justification of social inequalities, as individual achievement would be seen as rewarded and social policies as less necessary ++}
+
 Research indicates various socialization practices at families and schools during childhood and adolescence that impact dispositional and behavioral tendencies concerning justifications of inequality in adult life.

### 2024-05-17 02-antecedentes.qmd
revisada introduccion y antecedentes, temas menores formato

--- 2024-05-09 02-antecedentes.qmd
@@
 author: "Equipo EDUMER"
-format: pdf
 editor: visual
+bibliography: "input/bib/just-ineq-merit.bib"
+csl: "input/bib/apa6.csl"
 ---
 
-## Justification of inequality
+## Justification of inequality and market justice
 
+The justification of social inequality based on marke-type criteria has been conceptualized as the individuals’ adherence to the deservingness of social goods and services (sucha as health, education and pensions) based on prices and individuals’ ability to pay [@boltanski_new_2005; @lane_market_1986; @streeck_citizens_2012].
 Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegel_beliefs_1987], highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness.
@@
 This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades [@centeno_arc_2012; @harvey_breve_2015], affecting areas such as pensions, health services, and education.
-The justification of social inequality based on marketype criteria has been conceptualized as the individuals’ adherence to one specific justice evaluation, the market justice, that is, affording legitimacy to the allocation of goods and services based on prices and individuals’ ability to pay [@boltanski_new_2005; @lane_market_1986; @streeck_citizens_2012].
 
 Robert E.
-Lane proposed the underpinnings of market justice, which he differentiated from political justice.
+Lane proposed the underpinnings of the concept of market justice, which he differentiated from political justice.
 For him, “it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets” [-@lane_market_1986; p.
 384].
-Following the theory of relative deprivation –a social phenomenon arising when individuals cannot afford what most others in their environment can (Merton 1950)– Lane notes that, in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions.
-Although empirical research has shown that, contrary to Lane’s observation, relative deprivation, even in market settings, produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action such as protests and revolt [@greitemeyer_subjective_2016; @mishra_subjective_2015; @smith_relative_2012; @power_deprivationprotest_2018; @power_relative_2020], the conceptualization of market justice has been since then closely coupled to the merit principle for allocating outcomes: the claim that the unequal levels of well-being individuals enjoy ought to be, to some extent, a function of their talents and efforts, regardless of their needs or membership, the two latter being the realm of political justice and its closely coupled principles of need –allocating outcomes to those who require them most– and equality –allocating the same outcome to everyone–, which have been at the center of welfarism (see Wilson 2003).
-The underlying notion of market justice also resonates in its application to welfare regimes, providing a framework for understanding the varied global approaches to managing social services.
+Contrary to the evidence that unequal distribution produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action [@greitemeyer_subjective_2016; @mishra_subjective_2015; @smith_relative_2012; @power_deprivationprotest_2018; @power_relative_2020], Lane points out that in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions.
+In this sense, unequal levels of well-being ought to be, to some extent, a function of their talents and efforts, instead of being based in distributive principles that characterize welfarism, such as need and equality (see Wilson 2003).
 
@@
 The trend toward marketization of welfare services has been growing since the 1980s [@salamon_marketization_1993], and this shift is increasingly evident even in countries where market solutions have traditionally had a minor role in social policy [@sivesind_changing_2017].
+The expansion of marketization has been related to larger justification of market mechanisms, whereby societies with larger private spending on services show larger market justice preferences [@lindh_public_2015].
 
-The question arises whether adults and children justify unequal access to welfare services based on market justice principles.
-Influenced by theories of policy feedback, which suggest that social welfare policies can reinforce (positive feedback) or undermine (negative feedback) previous policy trajectories [@fernandez_positive_2013; @pierson_increasing_2000; @weaver_paths_2010], citizens’ beliefs about market justice are likely also shaped by the institutional and social contexts they encounter.
-Indeed, the justification of inequality in access to essential services like education and health, based on one’s ability to pay, shows significant variation across countries, as demonstrated by international surveys, although they do not usually include children in their samples.
-For instance, the 2019 International Social Survey Program (ISSP) provides insights into how adults perceive inequality in accessing welfare services.
-@fig-issp illustrates the variation in agreement levels regarding whether it is just for individuals with higher incomes to purchase better education or healthcare.
-Notably, in 18 of the 29 countries analyzed, there is a greater justification for market inequality in healthcare access than in education.
-Nevertheless, the general sentiment typically ranges from 'somewhat unjust' to 'neither just nor unjust, with mixed feelings.
+Despite high-income inequality and limited social mobility in CHile and in Latin America in general, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@bucca_merit_2016; @chong_mystery_2008; @torche_intergenerational_2014; @salgado_inequality_2023].
+The reliance on private welfare providers and widespread user fees [@molyneux_neoliberal_2008] adds complexity to this context.
+Yet, research on children’s justification of market-based inequalities in accessing welfare services remains limited, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.
 
-```{r echo=FALSE}
-pacman::p_load(dplyr, ggplot2, haven)
-load("input/data/issp_rec.RData")
-```
+## Meritocratic perceptions and market justice
 
-```{r echo=FALSE}
-#| label: fig-issp
-#| fig-cap: "Average of market justice preferences by country"
-#| fig-asp: 0.65
+The original definition of merit is a combination of effort and talent [@young_rise_1958], and a meritocracy is a distributive system where merit is the main criteria for allocating valuable goods and rewards.
+From a sociological perspective, meritocracy has been used in research on social mobility to characterize societies with low mobility that threaten the meritocratic ideal [@goldthorpe_myth_2003].
+More recently, sociology and social psychology research has attended to subjective aspects related to the support for meritocratic principles in different societies, such as beliefs in meritocracy [@castillo_multidimensional_2023a; @mijs_unfulfillable_2016; @mijs_paradox_2019].
+The label of beliefs in this realm covers different concepts, but in this paper we refer to a conceptual model developed by @castillo_multidimensional_2023 which identify two main aspects: preferences and perceptions for meritocracy.
 
-base$country <- as.character(base$country)
-base$country <- substr(base$country, 4, nchar(base$country))
-base$country <- as_factor(base$country)
+While meritocratic preferences refer to a justification of distribution based on merit criteria (effort and talent), the perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society [@duru-bellat_who_2012; @castillo_meritocracia_2019].
+This perception can vary greatly depending on individual experiences, social, economic, and cultural background.
+Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work.
+In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2021].
+Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger market justice preferences, as individual achievement would be seen as rewarded and social policies as less necessary.
 
-base %>% ggplot(aes(mean, country)) +
-  theme_bw() +
-  geom_point(aes(color = justice, shape = justice), size = 2) +
-  stat_summary(aes(xintercept = after_stat(x), y = 0), fun = mean, geom = "vline", orientation = "y", colour = "green") +
-  xlab("") +
-  ylab("Country")+
-  scale_x_continuous(limits = c(1, 5),
-                     breaks = seq(1, 5, by = 1),
-                     label = c("Very unjust,\n definitely wrong", 
-                               "Somewhat unjust,\n wrong", 
-                               "Neither just nor\n unjust, mixed feelings", 
-                               "Somewhat just,\n right", 
-                               "Very just,\n definitely right")) +
-  scale_color_discrete(name = NULL) +
-  scale_shape_discrete(name = NULL) +
-  theme(text = element_text(size = 9.35),
-#        legend.position = "bottom",
-        plot.margin = margin(b = 0))
-```
+Some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2021], and how larger inequality affects meritocratic beliefs [@morris_representing_2022].
+A larger justification of meritocratic distribution have been related to less support for redistributive compensation systems [@frank_performance_2015].
+Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed [@evans_strong_2018].
+@almas_cutthroat_2020 found that in the US the highly educated accept inequality significantly more than the less educated because they perceive inequality as justifiable owing to differences in productivity (i.e., merit).
+@barr_effect_2020 found that in unequal societies, the highly educated accept inequality more than the low educated.
+Using data for 41 countries from the International Social Survey Programme (ISSP),@garcia-sanchez_attitudes_2020 found that the perceived size of the income gap correlated positively with support for progressive taxation, an association that was weaker among those who endorsed meritocratic and equal opportunity beliefs.
+In the same line, experimental research by @durante_preferences_2009 showed that subjects support less redistribution when the initial distribution is determined according to task performance.
 
-While contemporary policy developments have increasingly embraced the marketization of social welfare in areas such as education, pensions, and health services, questions about public acceptance of these changes persist.
-@lindh_public_2015 analysis of ISSP 2009 data from 17 OECD countries reveals a general lack of support for market-based distribution of social services, suggesting widespread disapproval of market stratification of essential services.
-This finding is corroborated by @soler-martinez_concerns_2023 research from Latinobarómetro 2020 across 18 Latin American countries, where concerns about health and education access predominated over income inequality.
-These results indicate that reforms toward welfare marketization are typically driven by elite political decisions rather than grassroots demand.
+## The socialization of market justice: families and schools
 
-Despite high-income inequality and limited social mobility in Latin America, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@bucca_merit_2016; @chong_mystery_2008; @torche_intergenerational_2014; @salgado_inequality_2023].
-The reliance on private welfare providers and widespread user fees [@molyneux_neoliberal_2008] adds complexity to this context.
-Yet, research on children’s justification of market-based inequalities in accessing welfare services remains limited, especially in Latin America, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.
-
-## Meritocracy
-
-The concept of meritocracy frequently appears nowadays when analyzing cultural determinants of social inequalities.
-In general, it is mentioned as a value associated with justice, as it would link efforts and talents with rewards in an equitable manner.
-This normative sense is quite far from its original formulation by @young_rise_1958 in the satirical novel "The Rise of Meritocracy", where it ironically represented a mechanism for reproducing the inequalities of origin.
-The meritocratic ideal had remained relatively unchallenged until a series of recent publications turned into its potential consequences for maintaining social inequality.
-Perhaps one of the most recent sources in this line is Michael Sandel's "The Tyranny of Merit", where he strongly questions the implications of carrying out the principle of merit in societies that do not guarantee equal opportunities and that generate a feeling of scarce recognition and appreciation of those who receive lesser rewards: "In society's eyes, and perhaps also their own, their work no longer signified a valued contribution to the common good." [@sandel_tyranny_2020, pp.].
-
-Empirical research on meritocracy has increased along with the philosophical-normative discussion on meritocracy in recent years.
-Particularly from a sociological perspective, meritocracy has been used in research on social mobility to characterize societies with low mobility that threaten the meritocratic ideal [@goldthorpe_myth_2003].
-More recently, sociology and social psychology research has attended to the subjective aspects vis-a-vis beliefs in meritocracy.
-The label of beliefs in this realm covers a series of areas, such as attitudes, perceptions, and preferences (Castillo et al), whereby most of the link this subjective dimensions to individual socio-structural factors and context-level determinants.
-For instance, some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2021], and how larger inequality affects meritocratic beliefs [@morris_representing_2022].
-Based on these findings, a research agenda has been reinforced on the legitimizing role of meritocracy, in line with previous studies using the concept of a just world (Lerner, Dalbert) and the theory of system justification (Jost & Major).
-
-How do meritocratic beliefs legitimize inequalities?
-Empirical studies have used experiments and surveys to address this question.
-For instance, the evidence suggests that just world beliefs correlate negatively with support for redistributive compensation systems [@frank_performance_2015].
-Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed [@evans_strong_2018].
-@almas_cutthroat_2020 found that in a relatively unequal society (the United States), the highly educated accept inequality significantly more than the less educated because they perceive inequality as justifiable owing to differences in productivity (i.e., merit), whereas in a relatively equal society (Norway), the less educated accept inequality more, but not significantly more than the highly educated because meritocratic values are less prevalent.
-@barr_effect_2020 also addressed this triple interaction between the level of inequality in a society, the individual level of education, and the perceived origin of the disparity (either by luck or effort) to determine the extent to which inequality is accepted.
-They found that the interaction's mechanism varies depending on the compared societies.
-Finally, @garcia-sanchez_attitudes_2020, using data for 41 countries from the International Social Survey Programme (ISSP), found that the perceived size of the income gap correlated positively with support for progressive taxation.
-Still, this association was weaker among those who endorsed meritocratic and equal opportunity beliefs.
-In the same line, experimental research by @durante_preferences_2009 subjects support less redistribution when the initial distribution is determined according to task performance.
-
-Research about meritocratic beliefs at school age is rather scarce, leaving a wide research gap as schools are one of the primary socialization institutions where achievement based on merit explains success [@erivwo_meritocracy_2021].
-Nevertheless, it is possible to find some initial works focused on the area of distributive justice at school that are closely related to meritocratic beliefs, as the ones by Resh and Sabbagh (quotes).
-Using justice in grade obtention as a measure of distributive justice and meritocracy (Sabbagh et al 2006), they find for instance that a larger sense of distributive justice about grades is associated to higher socio-economic status [@resh_sense_2010], have a positive effect on liberal democratic orientation and on trust in people and in formal institutions [@resh_sense_2014], and tend to refrain from violence and to engage to a greater extent in extra-curricular school activity and community volunteering [@resh_sense_2017].
-
-## Children's judgments of inequality, schools, and family background
-
-Research indicates various socialization practices at families and schools during childhood and adolescence that impact dispositional and behavioral tendencies concerning justifications of inequality in adult life.
-The differences in economic understanding across age groups are consistent with research on cognitive development [@choudhury_social_2006].
-Adolescents with mature socio-cognitive abilities tend to express stronger preferences for fairness than infants and children [@wynn_not_2018].
-As children grow older, they become more likely to behave fairly, with their early-emerging strict egalitarianism being replaced by an increasing endorsement of fairness principles and engagement in collaborative activities [@huppert_development_2019; @mcauliffe_developmental_2017].
-In these activities, their fairness views consider individual contributions, merits, and circumstances [@almas_fairness_2010; @huppert_development_2019; @sigelman_development_1991].
-@engelmann_children_2019 claim that children’s sense of fairness emerges at three years old, and we can observe it in collaborative activities, where they accept inequality if the procedure gives everyone an equal chance.
-Therefore, children at this age respond to unequal distributions based on interpersonal concerns, as they already demand equal respect.
-In any case, between 3 and 8 years of age, inequitable and anti-meritorious allocations are evaluated more negatively, but equitable and meritorious allocations are not evaluated more positively [@elenbaas_unfairness_2019].
-
-Some research shows that the social environment in which children develop, such as family and school, is associated with their prosocial behaviors by playing an essential role in the transmission of equity norms [@schunk_fairness_2023; @kosse_prosociality_2020].
-In fact, schools contribute to institutionalizing and reproducing inequality by promoting values, norms, practices, and languages familiar to higher-class families because the dominant group’s culture shapes educational institutions [@bourdieu_reproduction_1990].
+Schools contribute to institutionalizing and reproducing inequality by promoting values, norms, practices, and languages familiar to higher-class families because the dominant group’s culture shapes educational institutions [@bourdieu_reproduction_1990].
 Middle- and upper-class students are better equipped to face academic challenges and are more familiar with academic expectations [@mikus_children_2020].
@@
 Additionally, academic achievement is treated as the outcome of dispositional factors (e.g., pupils’ efforts and talents or lack of them) rather than the result of differential access to critical resources.
-Due to the meritocratic frame schools encourage, both low- and high-status individuals believe that students’ success or failure is not due to their family background but rather to differences in efforts and talents [@darnon_where_2018].
+Due to the meritocratic frame schools encourage, both low- and high-status individuals tend to believe that success or failure is not due to family background but rather to differences in efforts and talents [@darnon_where_2018].
 In this sense, we believe that the perception of meritocracy can influence students' judgments about market justice preferences.
-Furthermore, we believe there is a difference between students' perceptions of meritocracy based on their own experience in school and what they perceive in society at large.
-Consistently, our first two hypotheses are:
 
-H1a: Students who perceive that there is more meritocracy at school will show larger market justice preferences
+H1a: Students who perceive that there is more meritocracy in society will show larger market justice preferences
 
-H1b: Students who perceive that there is more meritocracy in society will show larger market justice preferences
+The perception of meritocracy has been mostly studied with general questions about rewards allocation based on effort and talent (ususally ingelligence).
+Nevertheless, when looking at school population it is possible to further consider the perception of meritocracy referred specifically to the school context.
+For instance, @resh_sense_2010 find that perception of justice in grades have a positive effect on liberal democratic orientation and on trust in people and in formal institutions [@resh_sense_2014].
+Attending to this evidence, we differentiate in this study between meritocratic perception in society at large, and meritocratic perception at school, proposing the next hypothesis:
 
-Family background and family socialization practices also contribute to children’s and adolescent’s market justice preferences.
-For example, @almas_fairness_2017 found that adolescents from low-socioeconomic-status families are likelier to have an egalitarian fairness view and consider an equal distribution as fair in a situation with unequal merits.
-The authors speculate that differences in socialization practices across status groups might bring about, to a great extent, the fairness views of children and adolescents because social status seems to interact with these evaluations [e.g., @hvidberg_social_2023].
+H1b: Students who perceive that there is more meritocracy at school will show larger market justice preferences
 
-The classic work of Kohn showed that middle-class parents value the expression of internal states and emotions, such as self-control, curiosity, happiness, and consideration, while working-class parents promote deference, obedience, and conformity to authority [@kohn_social_1963; @kohn_class_1969].
+Attending to the socialization within the family, the classic work of Kohn showed that middle-class parents value the expression of internal states and emotions, such as self-control, curiosity, happiness, and consideration, while working-class parents promote deference, obedience, and conformity to authority [@kohn_social_1963; @kohn_class_1969].
 Although parents from all social backgrounds encourage individualism in their children, this shared norm translates into different forms in high and low social classes (1999).
 @acemoglu_obedience_2021 claimed that the values families impart to their children interact with social mobility.
-Because obedience is a valuable characteristic for employers, in low-wage and social mobility environments, low-income families impart values of obedience to their children to prevent disadvantaging them in labor markets.
 On the one hand, children from privileged families are socialized to adopt a clear conception of individualism that highlights their internal states, independence, and idiosyncrasies.
 In contrast, children from disadvantaged families are socialized to support a more balanced view of individualism that considers personal characteristics as resources to overcome collective impediments on the path to upward mobility [@iacoviello_collectivism_2019].
-In this way, we believe that there are differences in the socialization of values according to socioeconomic differences that could influence market justice preferences and, therefore, we propose the following hypothesis:
+In the same line, @almas_fairness_2017 found that adolescents from low-socioeconomic-status families are likelier to have an egalitarian fairness view and consider an equal distribution as fair in a situation with unequal merits.
+Taking this into account, we believe that there will be differences in the socialization of values according to socioeconomic differences in families that could influence market justice preferences, which lead us to the following hypothesis:
 
@@
 Academic track students are depicted as wealthy, intelligent, ambitious, and diligent, while vocational track students are characterized as poor, unambitious, unintelligent, and lackadaisical.
-These stereotypes help individuals maintain a sense of superiority over others and legitimize the prevailing social hierarchies and economic disparities [@jost_attitudinal_2000]
+These stereotypes help individuals maintain a sense of superiority over others and legitimize the prevailing social hierarchies and economic disparities [@jost_attitudinal_2000].
+The link between school socioeconomic status and market preferences is certainly reinforced in societies with high school segregation, as is the case of Chile.
+Consistently, we expect that
 
@@
 
-H4: Students from schools with higher average levels of academic achievement will show larger market justice preferences.
+The last set of hypotheses deal with the role of meritocratic perception as a moderator of family and school status on market justice preferences.
+These hypotheses have more an exploratory character, as there are not clear references in the literature to this regard.
+The general argument deals with the role of socioeconomic status and meritocratic perceptions, which is contested in the literature.
+On the one hand, and based on a mere rational argument, it would be expected that those who succeed in terms of educational rewards and carreer paths - usuaylly those of higher status - perceive more meritocracy.
+Nevertheless, evidence regarding the legitimacy of meritocracy relies on consensual beliefs from the side of both winners and losers from the distributive system, pointing in the direction of fewer status differences when it comes to meritocratic beliefs [@wiederkehr_belief_2015].
+As this is the first study of this kind in the Chilean context, charaterized by extreme educational inequalities, we rely on a traditional and rational approach, proposing that:
 
-\[interacciones\]
+H4: Students from schools of higher social status will perceive more meritocracy.
 
-H5: The perception of meritocracy in school and society will moderate the effect of family social status on market justice preferences.
+H5: Students from families of higher social status will perceive more meritocracy.
 
-H6: The perception of meritocracy in school and society will moderate the effect of school status on market justice preferences.
+[aca falta hipothesis de school performance]
 
-H7: The perception of meritocracy in school and society will moderate the effect of school academic achievement on market justice preferences.
+The previous hypotheses serve as a sort of bridge for the final moderation hypotheses.
+As higher (family and school) status would be related to meritocratic perception (H5 and H6), and at the same time meritocratic perception would be related to larger market justice preferences (H1 and H2), we argue that meritocratic perceptions could play a moderating role: the association between high status and market justice would be stronger for those with a larger meritocratic perception:
 
-### Summary of hypotheses
+H6: The perception of meritocracy in school and society will moderate the effect of family social status on market justice preferences.
+
+H7: The perception of meritocracy in school and society will moderate the effect of school status on market justice preferences.
+
+### Summary of hypotheses @Kevin-carrasco, actualizar esquema
 

### 2024-05-17 paper.qmd
revisada introduccion y antecedentes, temas menores formato

@@
   html: 
+    toc: true
     keep-tex: true
 editor: visual
+smooth-scroll: true
+number-sections: true
 bibliography: "input/bib/just-ineq-merit.bib"
@@
 
-# The socialization of meritocracy and market justice preferences at school
+# The socialization of meritocracy and market justice preferences at school {.unnumbered}
+
+Abstract
 

### 2024-05-17 pruebas.qmd
revisada introduccion y antecedentes, temas menores formato

---
format:
  html:
    toc: true
    keep-tex: true
editor: visual
smooth-scroll: true
number-sections: true
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/apa6.csl"
link-citations: true
mainfont: Times New Roman
---

```{css, echo = FALSE}
.justify {
  text-align: justify !important
}
```

# The socialization of meritocracy and market justice preferences at school {.unnumbered}

Abstract

:::{.justify}
Robert E. Lane proposed the underpinnings of the concept of market justice, which he differentiated from political justice. For him, “it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets” [-Lane (1986); p. 384]. Contrary to the evidence that unequal distribution produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action (Greitemeyer & Sagioglou, 2016; Mishra & Carleton, 2015; Séamus A. Power, 2018; Séamus A. Power et al., 2020; Smith et al., 2012), Lane points out that in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions. In this sense, unequal levels of well-being ought to be, to some extent, a function of their talents and efforts, instead of being based in distributive principles that characterize welfarism, such as need and equality (see Wilson 2003).

:::

### 2024-05-23 03-methods.qmd
revisada introduccion y antecedentes, temas menores formato
corrección methods

@@
 author: "Equipo Edumer"
-format: pdf
 editor: visual
@@
 
-## Data, variables and method
+# Data, variables and method
 
-### Data
+## Data
 
-The main data source corresponds to the First Study of Civic Education in Chile, conducted by the Agency for Quality Education of the Ministry of Education in 2017.
-This database is based on the administration of a civic knowledge test to students and questionnaires to various members of the educational community.
-The target population consists of 8th-grade students in 242 schools across the country.
-Initially, the database comprises 8,589 student observations and 6,770 parent observations.
-However, considering the survey's unit of analysis and to ensure higher data quality, 171 student cases and 79 parent cases exhibiting repetitive and careless response patterns were removed (Gottfried et al., 2022).
-Additionally, data from the System for Measuring the Quality of Education (SIMCE) conducted by the Ministry of Education for the year 2017 are employed, providing school-level information such as administrative dependence, socioeconomic classification, and results obtained in standardized mathematics and language tests.
-After variable processing and removal of missing cases, the final database for this study includes a two-level stratified sample, composed of 5,047 students and parents (level 1) nested within 231 schools (level 2).
+## 🗒️ Answer
 
-### Variables
+The First Study of Civic Education in Chile, conducted by the Agency for Quality Education of the Ministry of Education in 2017, serves as the main data source.
+This database is composed by a civic knowledge test score and a series of items batteries measuring different aspects of citizenship.
+The target population includes 8th-grade students in 242 schools nationwide, as well as their parents and teachers.
+Initially, the database contains 8,589 student observations and 6,770 parent observations.
+To ensure higher data quality and considering the survey's unit of analysis, we removed 171 student cases and 79 parent cases that exhibited repetitive and careless response patterns (Gottfried et al., 2022).
+Additionally, we utilized data from the System for Measuring the Quality of Education (SIMCE) conducted by the Ministry of Education in 2017, which provides school-level information such as administrative dependence, socioeconomic classification, and results obtained in standardized mathematics and language tests.
+After processing the variables and removing missing cases, the final database for this study includes a two-level stratified sample composed of 5,047 students and parents (level 1) nested within 231 schools (level 2).
+
+## Variables
 
@@
 
-**Justification of Inequality**: The dependent variable in this study is the justification of social inequality in specific policy domains, namely health, education, and pensions.
-This construct is measured through three variables addressing the degree of justification regarding whether access to social services in these domains should be income conditional.
-Justification of inequality in health is measured by the item "Is it fair in Chile that people with higher incomes can access better healthcare than those with lower incomes," while in education, it is measured using the statement "Is it fair in Chile that people who can afford it have better education for their children." In the case of pensions, it is measured through the item "Is it fair in Chile that people with higher incomes can have better pensions than those with lower incomes." In all cases, respondents declare their preferences on a Likert scale ranging from "completely disagree" (1) to "completely agree" (4).
+**Justification of Inequality**: The dependent variables in this study are market justice preferences.
+This construct is measured through three variables addressing the degree of justification regarding whether access to social services in pensions, education and health should be income conditional.
+Justification of inequality in health is measured by the item "Is it fair in Chile that people with higher incomes can access better healthcare than those with lower incomes,", and the same is asked for education and pensions.
+In all cases, respondents declare their preferences on a Likert scale ranging from "completely disagree" (1) to "completely agree" (4).
 Additionally, we include a summarized indicator of preferences for market justice (REF), measured by an average index across these items (α = 0.86), with values ranging from 1 to 4, where higher values represent greater preferences for market justice (see @tbl-desc-dependientes).
+We analyzed these items independently as well as by constructing and average index called "market justice preferences".
 
@@
 The analysis includes several individual control variables that have been identified as influential in recent research.
-These variables encompass family socioeconomic status, assessed by the highest level of parental education, the number of books in the household, and an index measuring access to technologies, all aimed at mitigating compositional effects (REFS).
+These variables encompass family socioeconomic status - by the highest level of parental education -, the number of books in the household, and an index measuring access to technologies.
 @tbl-desc-independent shows the independent variables used, their response categories and their frequencies.
@@
 Firstly, we assess the administrative dependence of schools, categorized as "Public," "Subsidized Private," or "Private." Secondly, we consider the socioeconomic classification of schools, determined by the Ministry of Education and measured on an ordinal scale with five categories ranging from "low" (1) to "high" (5).
-Thirdly, we evaluate the performance level of schools in SIMCE standardized tests, categorized as "low," "medium," or "high." Lastly, we analyze the proportion of parents with university or postgraduate education level at each school.
-The items used, response categories, and their frequencies are detailed in @tbl-desc-school.
+Thirdly, we evaluate the performance level of schools in SIMCE standardized tests, categorized as "low," "medium," or "high".
+SIMCE (System of Measurement of Educational Quality) is a national test in areas such as language and mathematics which is applied yearly in different educational levels (REF).
+Lastly, we analyze the proportion of parents with university or postgraduate education level at each school.
+The variables, response categories, and their frequencies are detailed in @tbl-desc-school.
+{@kevin arreglar N L2 tabla}
 
@@
 
-### Methods
+## Methods
 

### 2024-05-27 03-methods.qmd
Actualizacion

@@
 
-# Data, variables and method
+## Data, variables and method
 
-## Data
-
-## 🗒️ Answer
+### Data
 
@@
 
-## Variables
+### Variables
 
@@
 **Justification of Inequality**: The dependent variables in this study are market justice preferences.
-This construct is measured through three variables addressing the degree of justification regarding whether access to social services in pensions, education and health should be income conditional.
-Justification of inequality in health is measured by the item "Is it fair in Chile that people with higher incomes can access better healthcare than those with lower incomes,", and the same is asked for education and pensions.
-In all cases, respondents declare their preferences on a Likert scale ranging from "completely disagree" (1) to "completely agree" (4).
-Additionally, we include a summarized indicator of preferences for market justice (REF), measured by an average index across these items (α = 0.86), with values ranging from 1 to 4, where higher values represent greater preferences for market justice (see @tbl-desc-dependientes).
-We analyzed these items independently as well as by constructing and average index called "market justice preferences".
+This construct is measured through three variables that address the degree of justification regarding whether access to social services in pensions, education, and health should be income conditional.
+Justification of inequality in health is measured by the item: "Is it fair in Chile that people with higher incomes can access better healthcare than those with lower incomes?" The same question is asked for education and pensions.
+In all cases, respondents indicate their preferences on a Likert scale ranging from "completely disagree" (1) to "completely agree" (4).
+Additionally, we include a summarized indicator of "market justice preferences", measured by an average index across these items (α = 0.86), with values ranging from 1 to 4, where higher values represent greater preferences for market justice (see @tbl-desc-dependientes).
+We analyzed these items independently as well as by the average index.
 
@@
 
-The analysis includes several individual control variables that have been identified as influential in recent research.
-These variables encompass family socioeconomic status - by the highest level of parental education -, the number of books in the household, and an index measuring access to technologies.
+**Family Socioeconomic Status**: The socioeconomic status of students' families is measured using two indicators.
+First, the highest educational level attained by the parents, with categories: "8th grade or less," "Secondary education," "Technical higher education," "University or postgraduate," and "No response." The inclusion of the "No response" category is due to its high frequency in the data; omitting it could obscure relevant associations.
+Second, the number of books in the household is used, categorized as "Less than 25" and "More than 25."
+
 @tbl-desc-independent shows the independent variables used, their response categories and their frequencies.
@@
 
-df<-dfSummary(independientes,
+df<-dfSummary(independientes[,-8],
                plain.ascii = FALSE,
@@
 
-This study examines four variables at the school level.
-Firstly, we assess the administrative dependence of schools, categorized as "Public," "Subsidized Private," or "Private." Secondly, we consider the socioeconomic classification of schools, determined by the Ministry of Education and measured on an ordinal scale with five categories ranging from "low" (1) to "high" (5).
-Thirdly, we evaluate the performance level of schools in SIMCE standardized tests, categorized as "low," "medium," or "high".
-SIMCE (System of Measurement of Educational Quality) is a national test in areas such as language and mathematics which is applied yearly in different educational levels (REF).
-Lastly, we analyze the proportion of parents with university or postgraduate education level at each school.
-The variables, response categories, and their frequencies are detailed in @tbl-desc-school.
-{@kevin arreglar N L2 tabla}
+This study focuses on two school-level characteristics: socioeconomic status and academic performance.
+Socioeconomic status is assessed using the Ministry of Education's classification, measured as an ordinal item with five categories ranging from "low" (1) to "high" (5).
+Academic performance is measured using the school's results in the SIMCE (System of Measurement of Educational Quality) standardized tests, which are administered yearly at different educational levels in areas such as language and mathematics.
+These results are categorized as "low," "medium," and "high." The items used, response categories, and their frequencies are detailed in @tbl-desc-school.
 
@@
 
-df<-dfSummary(school,
+df<-dfSummary(school[,c(2,4)],
                plain.ascii = FALSE,
@@
 
-## Methods
+***Controls***
+
+A series of individual control variables are included.
+At the individual level, an index of access to technology is constructed based on the number of computers, tablets, and cell phones in the household, as well as internet connectivity.
+At the contextual level, we use the administrative dependence of schools, classified as "Public," "Subsidized Private," or "Private," and the proportion of parents with university or postgraduate education at each school.
+
+### Methods
 

### 2024-05-27 04-analisis.qmd
revisada introduccion y antecedentes, temas menores formato
Actualizacion

@@
                texreg,
-               kableExtra)
+               kableExtra,
+               easystats)
 
@@
 rm(list = ls())
-
 ```
@@
 ```{r functions, echo=FALSE}
+
+odds_fun <- function(modelos) {
+   or <- texreg::extract(modelos)
+   or@coef <- exp(or@coef)
+   return(or)
+}
 
@@
 ```
+
+Likert plots
 
@@
 
-The results of the cumulative link mixed models for the ordinal dependent variables of justice in differential access to education, health, and pensions by income are presented in @fig-odds.
-This figure illustrates the effect of meritocratic variables on society and school as independent variables; however, the models incorporate all variables utilized in the study for each of the three dependent variables separately (complete models are available in the [Appendix](#appendix)).
+The results of the cumulative link mixed models for the ordinal dependent variables of justice in differential access to education, health, and pensions by income are presented in @tbl-ordinal-reg.
+This table shows the effects of individual and contextual-level independent variables related to our hypothesis; however, the models also incorporate the control variables utilized in the study for each of the three dependent variables separately (complete models are available in the [Appendix](#appendix)).
 
-```{r echo=FALSE}
+```{r echo=FALSE, include=FALSE}
 
-data_pension <- data_rec %>% dplyr::select(-just_educ, -just_salud, -redistribucion)
+data_pension <- data_rec %>% 
+    dplyr::select(-just_educ, -just_salud, -redistribucion)
 data_pension <- na.omit(data_pension)
 data_pension$just_pension_fact <- as_factor(data_pension$just_pension)
-reg5_pension <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_pension)
-data_salud <- data_rec %>% dplyr::select(-just_pension, -just_educ, -redistribucion)
+
+m0_p <- clmm(just_pension_fact ~ 1 + (1 | mrbd), data=data_pension)
+
+performance::icc(m0_p, by_group = T)
+
+m1_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_pension)
+
+m2_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
+    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_pension)
+
+m3_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
+    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
+    cod_grupo + simce + (1 | mrbd), data=data_pension)
+
+tbls <- list(m1_p, m2_p, m3_p)
+
+modelos_odd_pension <- map(tbls, odds_fun)
+
+data_salud <- data_rec %>% 
+    dplyr::select(-just_pension, -just_educ, -redistribucion)
 data_salud <- na.omit(data_salud)
 data_salud$just_salud_fact <- as_factor(data_salud$just_salud)
-reg5_salud <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_salud)
-data_educ <- data_rec %>% dplyr::select(-just_pension, -just_salud, -redistribucion)
+
+m0_h <- clmm(just_salud_fact ~ 1 + (1 | mrbd), data=data_salud)
+
+performance::icc(m0_h, by_group = T)
+
+m1_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_salud)
+
+m2_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
+    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_salud)
+
+m3_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
+    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
+    cod_grupo + simce + (1 | mrbd), data=data_salud)
+
+tbls <- list(m1_h, m2_h, m3_h)
+
+modelos_odd_health <- map(tbls, odds_fun)
+
+data_educ <- data_rec %>% 
+    dplyr::select(-just_pension, -just_salud, -redistribucion)
 data_educ <- na.omit(data_educ)
 data_educ$just_educ_fact <- as_factor(data_educ$just_educ)
-reg5_educ <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_educ)
-```
 
-```{r echo=FALSE}
-#| fig-cap: "Odds-ratios of justification of pensions, education and health by social and school meritocracy"
-#| label: fig-odds
-#| fig-asp: 0.8
+m0_e <- clmm(just_educ_fact ~ 1 + (1 | mrbd), data=data_educ)
 
-plot_models(reg5_salud, reg5_educ, reg5_pension,  
-            rm.terms = c("1|2", "2|3", "3|4", "educacion_recEd Media", "educacion_recEd Tecnica", "educacion_recUniversidad o posgrado", "educacion_recNs/Nr", "libros_recMás de 25", "acc_tec", "internet_rec2", "internet_recNs/Nr", "mean_educ", "cod_depe2Part.
-subvencionado", "cod_depe2Part.
-privado", "cod_grupoMedio bajo", "cod_grupoMedio", "cod_grupoMedio alto", "cod_grupoAlto", "simceMedio", "simceAlto"), #quitar terminos
-            title = "", #quitar titulo
-            show.values = TRUE, #mostrar valor de efectos
-            dot.size = 1.3, #tamaño circulos
-            line.size = 1, #tamaño CI
-            value.size = 3.5, #tamaño valor efectoss
-            spacing = .7, #espacio entre efectos
-            grid = TRUE, # separa en un plot por modelo
-            vline.color = "firebrick", # linea roja en punto neutro (1)
-            axis.labels = rev(c("School talent", 
-                            "School effort", 
-                            "Social talent",
-                            "Social effort", 
-                            "Deservingness"
-                            )), #con rev porque automatico los tira en otro orden
-            show.legend = FALSE,
-            m.labels = c("Health", "Education","Pension"),
-            colors = gray(level = 0.2))  # variables dependientes
+performance::icc(m0_e, by_group = T)
+
+m1_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_educ)
+
+m2_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
+    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_educ)
+
+m3_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
+    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
+    cod_grupo + simce + (1 | mrbd), data=data_educ)
+
+tbls <- list(m1_e, m2_e, m3_e)
+
+modelos_odd_educ <- map(tbls, odds_fun)
 
@@
 
-@fig-odds demonstrates a similar trend for the three dependent variables of justice in differential access to pensions, education, and health, with all effects being statistically significant.
+```{r echo=FALSE, results='asis'}
+#| tbl-cap: "Cumulative link multilevel models of differential acces justificacion of education, health and pensions"
+#| label: tbl-ordinal-reg
+
+ccoef <- list(
+  inteligencia_esc = "School talent",
+  esfuerzo_esc = "School effort",
+  inteligencia_soc = "Social talent",
+  esfuerzo_soc = "Social effort",
+  merito_soc = "Deservingness",
+  "educacion_recEd Media" = "Secondary",
+  "educacion_recEd Tecnica" = "Higher tec.",
+  "educacion_recUniversidad o posgrado" = "University or posgraduate",
+  "educacion_recNs/Nr" = "Missing",
+  mean_educ = "Prop. university level at school",
+  "cod_grupoMedio bajo" = "SES Medium low",
+  cod_grupoMedio = "SES Medium",
+  "cod_grupoMedio alto" = "SES Medium high",
+  cod_grupoAlto = "SES High",
+  simceMedio = "Simce Medium",
+  simceAlto = "Simce High"
+)
+
+texreg::htmlreg(l = list(modelos_odd_educ[[2]],
+                         modelos_odd_educ[[3]],
+                         modelos_odd_health[[2]],
+                         modelos_odd_health[[3]],
+                         modelos_odd_pension[[2]],
+                         modelos_odd_pension[[3]]),   
+                  stars = c(0.05, 0.01, 0.001),
+                  custom.note = "Note: Coefficients models in odds-ratio. Control variables are included. Standard errors in parentheses. %stars",
+                  custom.header = list("Education" = 1:2, "Health" = 3:4, "Pensions" = 5:6),
+                  custom.model.names = rep(c("Model 1", "Model 2"), 3),
+                  custom.coef.map = ccoef,
+                  leading.zero = T,
+                  scalebox = 0.60,
+                  groups = list("Parental education (Ref.= 8th grade or less)" = 6:9,
+                                "Socioeconomic level (Ref.= Low)" = 11:14,
+                                "Achievement score (Ref.= Low)" = 15:16),
+                 custom.gof.rows = list("Controls"=rep("Yes",6)))
+
+```
+
+@tbl-ordinal-reg demonstrates a similar trend for the three dependent variables of justice in differential access to pensions, education, and health, with all effects being statistically significant.
 In the context of school meritocracy, the effects are mixed: as perceived school talent increases, the justification for differentiated access to these services increases; conversely, as perceived school effort increases, the justification decreases, holding all other variables constant.
@@
 However, the results strongly support our second hypothesis (H1b), as a higher perception of societal meritocracy among students tends to increase the justification for market-based justice in these social services.
+CAMBIAR
 

### 2024-05-27 processing/analysis.r

@@ -1,0 +1,168 @@
+
+# 1. Packages -------------------------
+
+pacman::p_load(tidyverse,
+               ggpubr, 
+               lme4, 
+               ordinal,
+               sjPlot, 
+               texreg,
+               easystats,
+               performance)
+
+options(scipen=999)
+rm(list = ls())
+
+# 2. Data -------------------------
+
+load("input/data/data_index.RData")
+load("input/data/deservingness.RData")
+load("input/data/soc_effort.RData")
+load("input/data/soc_talent.RData")
+load("input/data/school_effort.RData")
+load("input/data/school_talent.RData")
+load("input/data/data_rec.RData")
+
+
+# 3. Analysis -------------------------
+
+odds_fun <- function(modelos) {
+   or <- texreg::extract(model = modelos, include.deviance = T)
+   or@coef <- exp(or@coef)
+   return(or)
+}
+
+
+# 3.1 Pensions ----
+
+data_pension <- data_rec %>% 
+    dplyr::select(-just_educ, -just_salud, -redistribucion)
+data_pension <- na.omit(data_pension)
+data_pension$just_pension_fact <- as_factor(data_pension$just_pension)
+
+# Null model
+m0_p <- clmm(just_pension_fact ~ 1 + (1 | mrbd), data=data_pension)
+
+performance::icc(m0_p, by_group = T) 
+## ICC School = 0.033
+
+# Model 1: Only meritocracy variables
+m1_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_pension) 
+
+# Model 2: Meritocracy variables + mean_educ + cod_depe2 + cod_grupo + simce
+m2_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + mean_educ + cod_depe2 + 
+    cod_grupo + simce + (1 | mrbd), data=data_pension) 
+
+# Model 3: Meritocracy variables + mean_educ + cod_depe2 + cod_grupo + simce + controls
[…]

### 2024-05-27 04-analisis.qmd
Update methods and analisis
update descriptive analisis market justice and meritocracy
update n lvl 2

--- 2024-05-15 04-analisis.qmd
@@
 
-## Analysis
+# Analysis
 
-### Descriptive analysis
+## Descriptive analysis
 
@@
                texreg,
-               kableExtra)
+               kableExtra,
+               sjPlot)
 
@@
 ```
+
+*Market justice preferences*
+
+```{r echo=FALSE}
+#| fig-cap: "Market justice preferences in education, health and pensions"
+#| fig-asp: 1.22
+#| label: fig-marketjustice
+
+data_rec %>% 
+  select(just_salud, just_educ, just_pension) %>% 
+  sjPlot::plot_stackfrq(expand.grid =TRUE,
+                      geom.colors = "PuBu",
+                      show.total = FALSE, 
+                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
+                      axis.labels = c("Health", "Education", "Pensions")) +
+  theme_light(base_size = 16) +
+  theme(legend.position="bottom", text = element_text(size = 14),
+        plot.caption = element_text(hjust = 1))
+```
+
+@fig-marketjustice shows the frequency distribution for the three market justice preference items (pensions, education, and health).
+The majority of respondents disagree (63.0% in total) with the idea that it is just for higher-income individuals to access better pensions.
+However, a notable portion (27.0%) agrees, and a smaller minority (10.1%) strongly agrees.
+Similar to pensions, a majority of respondents (58.8%) disagree with the idea that higher-income individuals should have access to better education.
+However, a considerable portion (28.6%) agrees, and a slightly larger minority (12.6%) strongly agrees compared to pensions.
+In the context of health, an even larger majority of respondents (66.4%) disagree with the notion that higher-income individuals should access better healthcare services.
+A smaller percentage (23.6%) agrees, and a minority (10.0%) strongly agrees.
+Overall, @fig-marketjustice shows a predominant sentiment against the justice of higher-income individuals accessing better services across all three areas.
+The strongest opposition is observed in the health sector.
+Despite the overall disagreement, a significant portion of respondents agree with the statement, particularly in education, where agreement and strong agreement are slightly higher compared to pensions and health.
+
+*School and social meritocracy*
+
+```{r echo=FALSE}
+#| fig-cap: "Social and school meritocracy"
+#| fig-asp: 1.22
+#| label: fig-meritocracy
+
+plot<-data_rec %>% 
+  select(merito_soc, esfuerzo_soc, inteligencia_soc, esfuerzo_esc, inteligencia_esc) %>% 
+  sjPlot::plot_stackfrq(expand.grid =TRUE,
+                      geom.colors = "PuBu",
+                      show.total = FALSE, 
+                      vjust=c(rep(c("center"),12), "bottom", "top", rep(c("center"), 6)),
+                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
+                      axis.labels = c("Deservingness", "Social effort", "Social talent", "School effort", "School talent")) +
+  theme_light(base_size = 16) +
+  theme(legend.position="bottom", text = element_text(size = 14),
+        plot.caption = element_text(hjust = 1))
+
+plot + 
+  annotate("text", x = 5.5, y = 0, label = "School meritocracy", size = 5, hjust = 0, vjust = 1) +
+  annotate("text", x = 3.5, y = 0, label = "Social meritocracy", size = 5, hjust = 0, vjust = 1)
+```
+
+@fig-meritocracy shows the frequency distribution for the five meritocracy items for the dimensions of school and social meritocracy.
+While a significant portion of respondents (29.2%) strongly agrees that talent is important to get good grades, a larger majority (78.6% combining agree and strongly agree) supports this notion.
+However, the strong agreement is not as pronounced as for school effort.
+An overwhelming majority (73.1%) agree that students effort is important to get good grades, which is a much higher percentage compared to the strong agreement for talent (29.2%).
+This indicates that people believe effort is more important than talent at school.
+There is a strong belief that both effort and talent are important, with a particularly high agreement for effort (73.1% strongly agree).
+
+In relation with social meritocracy, @fig-meritocracy shows that a majority of respondents (63.7% combining agree and strongly agree) agree that people are rewarded for their talent in society, with a notable portion (17.1%) strongly agreeing.
+A majority (62.0% combining agree and strongly agree) agree that people are rewarded for their effort in society.
+Similar to social talent, the perception is positive but less overwhelming than school meritocracy.
+The perception of whether people get what they deserve in society is more divided.
+While a slim majority (56.5% combining agree and strongly agree) agree, a significant minority (43.6%) disagrees, showing that this is the most contentious item among respondents.
+There is general agreement that people are rewarded for talent and effort in society, but the consensus is weaker compared to school meritocracy.
+Deservingness in society is the most contentious, with significant disagreement among respondents.
+
+*Market justice preferences in education, health and pensions by social and school meritocracy*
 
@@
 
-### Multivariate results
+## Models
 
@@
 
-```{r echo=FALSE}
+```{r echo=FALSE, warning=FALSE}
 
@@
 
-```{r echo=FALSE}
+```{r echo=FALSE, warning=FALSE}
 #| fig-cap: "Odds-ratios of justification of pensions, education and health by social and school meritocracy"
@@
             show.legend = FALSE,
-            m.labels = c("Health", "Education","Pension"),
+            m.labels = c("Health", "Education","Pensions"),
             colors = gray(level = 0.2))  # variables dependientes
@@
   collapse_rows(columns = 1:2, valign = "middle") %>% 
-  footnote(general = "***p < 0.001, **p < 0.01, *p < 0.05",
+  footnote(general = "*** p<0.001, ** p<0.01, * p<0.05",
            footnote_as_chunk = T,general_title = "Note: ")

### 2024-05-27 03-methods.qmd
Arreglos presentación modelos

@@
 
-## Data, variables and method
+# Data, variables and method
 
-### Data
+## Data
 
@@
 Initially, the database contains 8,589 student observations and 6,770 parent observations.
-To ensure higher data quality and considering the survey's unit of analysis, we removed 171 student cases and 79 parent cases that exhibited repetitive and careless response patterns (Gottfried et al., 2022).
+To ensure higher data quality and considering the survey's unit of analysis, we removed 171 student cases and 79 parent cases that exhibited repetitive and careless response patterns [@gottfried_autocorrelation_2022].
 Additionally, we utilized data from the System for Measuring the Quality of Education (SIMCE) conducted by the Ministry of Education in 2017, which provides school-level information such as administrative dependence, socioeconomic classification, and results obtained in standardized mathematics and language tests.
@@
 
-### Variables
+## Variables
 
@@
 
-**Justification of Inequality**: The dependent variables in this study are market justice preferences.
+**Market Justice Preferences**: The dependent variables in this study are market justice preferences.
 This construct is measured through three variables that address the degree of justification regarding whether access to social services in pensions, education, and health should be income conditional.
@@
 
-@tbl-desc-independent shows the independent variables used, their response categories and their frequencies.
+@tbl-desc-independent shows the individual level variables used, their response categories and their frequencies.
 
@@
 ```{r, results='asis', echo=FALSE}
-#| tbl-cap: "Independent variables"
+#| tbl-cap: "Individual level variables"
 #| label: tbl-desc-independent
@@
 Academic performance is measured using the school's results in the SIMCE (System of Measurement of Educational Quality) standardized tests, which are administered yearly at different educational levels in areas such as language and mathematics.
-These results are categorized as "low," "medium," and "high." The items used, response categories, and their frequencies are detailed in @tbl-desc-school.
+These results are categorized as "low," "medium," and "high." The contextual level items used, response categories, and their frequencies are detailed in @tbl-desc-school.
 
@@
 
-df<-dfSummary(school[,c(2,4)],
+df<-dfSummary(school,
                plain.ascii = FALSE,

### 2024-05-27 04-analisis.qmd
Arreglos presentación modelos

--- 2024-05-27 04-analisis.qmd
@@
 
-Likert plots
+@fig-marketjustice shows the frequency distribution for the three market justice preference items (pensions, education, and health).
+The majority of respondents disagree (63.0% in total) with the idea that it is just for higher-income individuals to access better pensions.
+However, a notable portion (27.0%) agrees, and a smaller minority (10.1%) strongly agrees.
+Similar to pensions, a majority of respondents (58.8%) disagree with the idea that higher-income individuals should have access to better education.
+However, a considerable portion (28.6%) agrees, and a slightly larger minority (12.6%) strongly agrees compared to pensions.
+In the context of health, an even larger majority of respondents (66.4%) disagree with the notion that higher-income individuals should access better healthcare services.
+A smaller percentage (23.6%) agrees, and a minority (10.0%) strongly agrees.
+Overall, @fig-marketjustice shows a predominant sentiment against the justice of higher-income individuals accessing better services across all three areas.
+The strongest opposition is observed in the health sector.
+Despite the overall disagreement, a significant portion of respondents agree with the statement, particularly in education, where agreement and strong agreement are slightly higher compared to pensions and health.
+
+```{r echo=FALSE}
+#| fig-cap: "Market justice preferences in education, health and pensions"
+#| fig-asp: 1.22
+#| label: fig-marketjustice
+
+data_rec %>% 
+  select(just_salud, just_pension, just_educ) %>% 
+  sjPlot::plot_stackfrq(expand.grid =TRUE,
+                      geom.colors = c("#7d7d7d","#ababab","#cecece","#ebebeb"),
+                      show.total = FALSE, 
+                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
+                      axis.labels = c("Health", "Pensions", "Education")) +
+ # theme_light(base_size = 16) +
+  theme(legend.position="bottom", text = element_text(size = 14),
+        plot.caption = element_text(hjust = 1))
+```
+
+@fig-meritocracy shows the frequency distribution for the five meritocracy items for the dimensions of school and social meritocracy.
+While a significant portion of respondents (29.2%) strongly agrees that talent is important to get good grades, a larger majority (78.6% combining agree and strongly agree) supports this notion.
+However, the strong agreement is not as pronounced as for school effort.
+An overwhelming majority (73.1%) agree that students effort is important to get good grades, which is a much higher percentage compared to the strong agreement for talent (29.2%).
+This indicates that people believe effort is more important than talent at school.
+There is a strong belief that both effort and talent are important, with a particularly high agreement for effort (73.1% strongly agree).
+
+```{r echo=FALSE}
+#| fig-cap: "Social and school meritocracy"
+#| fig-asp: 1.22
+#| label: fig-meritocracy
+
+plot<-data_rec %>% 
+  select(merito_soc, esfuerzo_soc, inteligencia_soc, esfuerzo_esc, inteligencia_esc) %>% 
+  sjPlot::plot_stackfrq(expand.grid =TRUE,
+                      geom.colors = c("#7d7d7d","#ababab","#cecece","#ebebeb"),
+                      show.total = FALSE, 
+                      vjust=c(rep(c("center"),12), "bottom", "top", rep(c("center"), 6)),
+                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
+                      axis.labels = c("Deservingness", "Social effort", "Social talent", "School effort", "School talent")) +
+  #theme_light(base_size = 16) +
+  theme(legend.position="bottom", text = element_text(size = 14),
+        plot.caption = element_text(hjust = 1))
+
+plot + 
+  annotate("text", x = 5.5, y = 0, label = "School meritocracy", size = 5, hjust = 0, vjust = 1) +
+  annotate("text", x = 3.5, y = 0, label = "Social meritocracy", size = 5, hjust = 0, vjust = 1)
+```
+
+In relation with social meritocracy, @fig-meritocracy shows that a majority of respondents (63.7% combining agree and strongly agree) agree that people are rewarded for their talent in society, with a notable portion (17.1%) strongly agreeing.
+A majority (62.0% combining agree and strongly agree) agree that people are rewarded for their effort in society.
+Similar to social talent, the perception is positive but less overwhelming than school meritocracy.
+The perception of whether people get what they deserve in society is more divided.
+While a slim majority (56.5% combining agree and strongly agree) agree, a significant minority (43.6%) disagrees, showing that this is the most contentious item among respondents.
+There is general agreement that people are rewarded for talent and effort in society, but the consensus is weaker compared to school meritocracy.
+Deservingness in society is the most contentious, with significant disagreement among respondents.
 
@@
 
-### Multivariate results
+## Multivariate results
 
-The results of the cumulative link mixed models for the ordinal dependent variables of justice in differential access to education, health, and pensions by income are presented in @tbl-ordinal-reg.
+The results of the cumulative link mixed models for the ordinal dependent variables of justice in differential access to health, pensions and education by income are presented in @tbl-ordinal-reg.
 This table shows the effects of individual and contextual-level independent variables related to our hypothesis; however, the models also incorporate the control variables utilized in the study for each of the three dependent variables separately (complete models are available in the [Appendix](#appendix)).
@@
   "educacion_recNs/Nr" = "Missing",
-  mean_educ = "Prop. university level at school",
   "cod_grupoMedio bajo" = "SES Medium low",
@@
 
-texreg::htmlreg(l = list(modelos_odd_educ[[2]],
-                         modelos_odd_educ[[3]],
-                         modelos_odd_health[[2]],
+texreg::htmlreg(l = list(modelos_odd_health[[2]],
                          modelos_odd_health[[3]],
                          modelos_odd_pension[[2]],
-                         modelos_odd_pension[[3]]),   
+                         modelos_odd_pension[[3]],
+                         modelos_odd_educ[[2]],
+                         modelos_odd_educ[[3]]),   
+                caption = NULL,
                 stars = c(0.05, 0.01, 0.001),
                 custom.note = "Note: Coefficients models in odds-ratio. Control variables are included. Standard errors in parentheses. %stars",
-                  custom.header = list("Education" = 1:2, "Health" = 3:4, "Pensions" = 5:6),
+                custom.header = list("Health" = 1:2, "Pensions" = 3:4, "Education" = 5:6),
                 custom.model.names = rep(c("Model 1", "Model 2"), 3),
@@
                 groups = list("Parental education (Ref.= 8th grade or less)" = 6:9,
-                                "Socioeconomic level (Ref.= Low)" = 11:14,
-                                "Achievement score (Ref.= Low)" = 15:16),
-                 custom.gof.rows = list("Controls"=rep("Yes",6)))
+                                "Socioeconomic level (Ref.= Low)" = 10:13,
+                                "Achievement score (Ref.= Low)" = 14:15),
+                custom.gof.rows = list("Controls"=rep("Yes",6)),
+                include.loglik = FALSE, include.aic = FALSE)
 
@@
 
-reg0_bienestar <- lmer(bienestar ~ 1 + (1 | mrbd), data=data_bienestar)
-reg1_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_bienestar)
-reg2_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_bienestar)
-reg3_bienestar <- lmer(bienestar ~ 1 + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_bienestar)
-reg4_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_bienestar)
+reg0_bienestar <- lmer(bienestar ~ 1 + (1 | mrbd), data = data_bienestar)
 
-test01<- anova(reg0_bienestar,reg1_bienestar,test = "Chisq")
-test02<- anova(reg0_bienestar,reg2_bienestar,test = "Chisq")
-test03<- anova(reg0_bienestar,reg3_bienestar,test = "Chisq")
-test04<- anova(reg0_bienestar,reg4_bienestar,test = "Chisq")
+performance::icc(reg0_bienestar, by_group = T)
 
-test.pvalues1<- test01$`Pr(>Chisq)`[2]
-test.pvalues2<- test02$`Pr(>Chisq)`[2]
-test.pvalues3<- test03$`Pr(>Chisq)`[2]
-test.pvalues4<- test04$`Pr(>Chisq)`[2]
+reg1_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
+                         inteligencia_soc + esfuerzo_soc + merito_soc + 
+                         acc_tec + internet_rec +
+                         (1 | mrbd), data=data_bienestar)
+
+reg2_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
+                         inteligencia_soc + esfuerzo_soc + merito_soc + 
+                         educacion_rec + libros_rec + acc_tec + internet_rec +
+                         (1 | mrbd), data=data_bienestar)
+
+reg3_bienestar <- lmer(bienestar ~ 1 + mean_educ + cod_depe2 + cod_grupo +
+                         simce + (1 | mrbd), data=data_bienestar)
+
+reg4_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + 
+                         inteligencia_soc + esfuerzo_esc + inteligencia_esc + 
+                         educacion_rec + libros_rec + acc_tec + internet_rec + 
+                         mean_educ + cod_depe2 + cod_grupo + 
+                         (1 | mrbd), data=data_bienestar)
+
+reg5_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + 
+                         inteligencia_soc + esfuerzo_esc + inteligencia_esc + 
+                         educacion_rec + libros_rec + acc_tec + internet_rec + 
+                         mean_educ + cod_depe2 + cod_grupo + simce + 
+                         (1 | mrbd), data=data_bienestar)
+
+test01 <- anova(reg0_bienestar, reg1_bienestar, test = "Chisq")
+test02 <- anova(reg0_bienestar, reg2_bienestar, test = "Chisq")
+test03 <- anova(reg0_bienestar, reg3_bienestar, test = "Chisq")
+test04 <- anova(reg0_bienestar, reg4_bienestar, test = "Chisq")
+test05 <- anova(reg0_bienestar, reg5_bienestar, test = "Chisq")
+
+test.pvalues1 <- test01$`Pr(>Chisq)`[2]
+test.pvalues2 <- test02$`Pr(>Chisq)`[2]
+test.pvalues3 <- test03$`Pr(>Chisq)`[2]
+test.pvalues4 <- test04$`Pr(>Chisq)`[2]
+test.pvalues5 <- test05$`Pr(>Chisq)`[2]
 
@@
 deviance4 <- test04$deviance[2]
+deviance5 <- test05$deviance[2]
 ```
@@
 
-texreg::htmlreg(list(reg1_bienestar, reg2_bienestar),
+ccoef <- list(
+  "(Intercept)" = "Intercept",
+  inteligencia_esc = "School talent",
+  esfuerzo_esc = "School effort",
+  inteligencia_soc = "Social talent",
+  esfuerzo_soc = "Social effort",
+  merito_soc = "Deservingness",
+  "educacion_recEd Media" = "Secondary",
+  "educacion_recEd Tecnica" = "Higher tec.",
+  "educacion_recUniversidad o posgrado" = "University or posgraduate",
+  "educacion_recNs/Nr" = "Missing",
+  "cod_grupoMedio bajo" = "SES Medium low",
+  cod_grupoMedio = "SES Medium",
+  "cod_grupoMedio alto" = "SES Medium high",
+  cod_grupoAlto = "SES High",
+  simceMedio = "Simce Medium",
+  simceAlto = "Simce High"
+)
+
+texreg::htmlreg(list(reg1_bienestar, reg2_bienestar, reg4_bienestar, reg5_bienestar),
                 custom.model.names = c("Model 1",
-                               "Model 2"),
-        custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
-        omit.coef = "(internet_rec)",
-        caption = "",
-        custom.coef.names = c("Intercept",  
-                              "School talent",
-                              "School effort", 
-                              "Social talent",  
-                              "Social effort",
-                              "Deservingness",
-                              "Secondary", 
-                              "Higher tec.", 
-                              "University or posgraduate", 
-                              "Missing", 
-                              "More than 25 books (Ref. Less than 25)", 
-                              "Technology access"),
-        groups = list("Parental education (Ref.= 8th grade or less)"=7:10),
-        custom.gof.rows=list("Deviance" = c(deviance1,
-                                            deviance2),
+                                       "Model 2",
+                                       "Model 3",
+                                       "Model 4"),
+                caption = NULL,
+                stars = c(0.05, 0.01, 0.001),
+                custom.note = "%stars",
+                custom.coef.map = ccoef,
+                leading.zero = T,
+                scalebox = 0.60,
+                groups = list("Parental education (Ref.= 8th grade or less)" = 7:10,
+                              "Socioeconomic level (Ref.= Low)" = 11:14,
+                              "Achievement score (Ref.= Low)" = 15:16),
+                custom.gof.rows = list("Controls"=rep("Yes",4),
+                                       "Deviance" = c(deviance1,
+                                                    deviance2,
+                                                    deviance4,
+                                                    deviance5),
                                        "Deviance Test (p)" = c(test.pvalues1,
-                                                       test.pvalues2))
-        )
+                                                             test.pvalues2,
+                                                             test.pvalues4,
+                                                             test.pvalues5)),
+                include.loglik = FALSE,include.aic = FALSE)
 
@@
 The total variance of model 5 according to Snijders and Bosker (2012) is 0.14.
-
-```{r echo=FALSE, results='asis'}
-#| tbl-cap: "Contextual effects of multilevel regression models"
-#| label: tbl-contextual-reg
-
-texreg::htmlreg(list(reg3_bienestar, reg4_bienestar),
-        custom.model.names = c("Model 1",
-                               "Model 2"),
-        custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
-        omit.coef = c("inteligencia_esc|esfuerzo_esc|inteligencia_soc|esfuerzo_soc|merito_soc|educacion_rec|libros_rec|acc_tec|internet_rec"),
-        caption = "",
-        custom.coef.names = c("Intercept", 
-                              "Prop. university level at school",
-                              "Subsidized private", 
-                              "Private", 
-                              "SES Medium low", 
-                              "SES Medium", 
-                              "SES Medium high", 
-                              "SES High", 
-                              "Simce Medium", 
-                              "Simce High"),
-        groups = list("Administrative dependence (Ref.= Public)"=3:4,
-                      "Socioeconomic level (Ref.= Low)"=5:8,
-                      "Achievement score (Ref.= Low)"=9:10),
-        custom.gof.rows=list("Deviance" = c(deviance3,
-                                            deviance4),
-                             "Deviance Test (p)" = c(test.pvalues3,
-                                                     test.pvalues4))
-        )
-
-```
 

### 2024-06-10 05-conclusiones.qmd
render paper
revisada introduccion y antecedentes, temas menores formato
discussion

@@
 ---
-title: "05-conclusiones"
 author: "Equipo EDUMER"
-format: pdf
 editor: visual
@@
 
-##
+# Discussion
+
+Our research aimed to examine the relationship between the perception of meritocracy and the justification of market justice preferences among eighth-grade students in Chile.
+Overall, our findings provided mixed evidence regarding our hypotheses.
+While we found support for some associations, others did not align with our initial expectations.
+Hypothesis H1a posited that students who perceive greater meritocracy in society would show stronger preferences for market justice.
+The results confirm this hypothesis, indicating a positive relationship between the perception that talent and effort are rewarded in society and the preferences for market justice in access to services such as health, education, and pensions.
+This finding aligns with previous research suggesting that individuals who believe in a meritocratic system are more likely to justify social inequalities [@batruch_belief_2022; @wiederkehr_belief_2015].
+Hypothesis H1b asserted that students who perceive greater meritocracy in school would show stronger preferences for market justice.
+The results show a mixed effect.
+The perception that talent is important for achieving good grades in school is associated with a greater preference for market justice in access to social services.
+Still, the perception that effort is important is associated with less preference for market justice.
+This highlights the influence of the school environment in shaping attitudes toward market justice, suggesting that meritocratic experiences in school can translate to broader societal views [@resh_sense_2010; @resh_sense_2014].
+
+Hypothesis H2 proposed that students from higher socioeconomic status families would show stronger preferences for market justice.
+However, the results indicate that this relationship is not as straightforward as expected.
+Although students from families with higher educational levels and more economic resources tend to prefer less market justice, this relationship is not statistically significant.
+This might be due to the influence of other contextual and cultural factors that modulate attitudes toward market justice [@kohn_social_1963; @kohn_class_1969; @almas_fairness_2017].
+
+Hypothesis H3 held that students from schools with higher socioeconomic status would show stronger preferences for market justice.
+The results did not confirm this hypothesis since although students from high-status schools tend to prefer more market justice, this result is not statistically significant.
+This finding, contrary to the existing literature [@jonsson_institutional_2015; @jost_attitudinal_2000], does not allow us to affirm the role of the school environment, particularly socioeconomic segregation, in influencing attitudes towards market justice and the acceptance of inequalities.
+
+Hypothesis Hx held that students from schools with higher school achievement would show stronger preferences for market justice.
+Contrary to that hypothesis, the results show that students from schools with greater achievement in the SIMCE test tend to prefer less market justice.
+
+Hypotheses H6 and H7 proposed that the perception of meritocracy would moderate the effect of socioeconomic status (both family and school) on market justice preferences.
+The results show that the perception of meritocracy, especially social meritocracy, amplifies the market justice preferences in schools with low levels of achievement and in families with parents with a lower educational level.
+This means that although those who perceive society as meritocratic have greater preferences towards market justice, this influence is less among those who have a higher socioeconomic status and attend schools with higher academic achievement.
+This finding highlights the complex interplay between socioeconomic status, school context, meritocratic beliefs, and preferences towards market justice.
+
+# Conclusions
+
+Future research could leverage the capabilities of longitudinal databases to explore contextual level.
+For example, statistical models could incorporate contextual information from administrative data sources, allowing for testing hypotheses that include contextual socioeconomic inequality and its dynamics over time.
+Additionally, this dataset could be used to compare inequality justification across different policy areas, such as health and pensions, and examine how meritocratic perceptions and beliefs affect these areas differently over time.
+Such studies would provide valuable insights for understanding societal and cultural changes and their impacts on solidarity and social cohesion.
 

### 2024-06-10 02-antecedentes.qmd
Actualizacion
Cambios

@@
 author: "Equipo EDUMER"
-editor: visual
 bibliography: "input/bib/just-ineq-merit.bib"
@@
 
+```{r echo=FALSE}
+pacman::p_load(tidyverse,
+               sjPlot,
+               sjlabelled,
+               sjmisc)
+
+options(scipen=999)
+rm(list = ls())
+
+load(file = "input/data/ELSOC_2016.RData")
+
+theme_nice <- function() {
+  theme_bw() +
+    theme(text = element_text(family = "serif"))
+
+}
+
+theme_set(theme_nice())
+```
+
 ## Justification of inequality and market justice
 
-The justification of social inequality based on marke-type criteria has been conceptualized as the individuals’ adherence to the deservingness of social goods and services (sucha as health, education and pensions) based on prices and individuals’ ability to pay [@boltanski_new_2005; @lane_market_1986; @streeck_citizens_2012].
-Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegel_beliefs_1987], highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness.
+The justification of social inequality based on marke-type criteria has been conceptualized as the individuals' adherence to the deservingness of social goods and services (sucha as health, education and pensions) based on prices and individuals' ability to pay [@boltanski_new_2005; @lane_market_1986; @streeck_citizens_2012].
+Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegel_beliefs_1987], highlights that people's explanations and justifications of social inequality are closely tied to their judgments of deservingness.
 The influence of ideologies [@wegener_dominant_1995] and cultural schemas [@homan_being_2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations.
-While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999 ; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles.
+While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999 ; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children's acceptance or rejection of these market principles.
 This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades [@centeno_arc_2012; @harvey_breve_2015], affecting areas such as pensions, health services, and education.
@@
 Lane proposed the underpinnings of the concept of market justice, which he differentiated from political justice.
-For him, “it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets” [-@lane_market_1986; p.
+For him, "it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets" [-@lane_market_1986; p.
 384].
@@
 
-Despite high-income inequality and limited social mobility in CHile and in Latin America in general, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@bucca_merit_2016; @chong_mystery_2008; @torche_intergenerational_2014; @salgado_inequality_2023].
+Despite high-income inequality and limited social mobility in Chile and in Latin America in general, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@bucca_merit_2016; @chong_mystery_2008; @torche_intergenerational_2014; @salgado_inequality_2023].
 The reliance on private welfare providers and widespread user fees [@molyneux_neoliberal_2008] adds complexity to this context.
-Yet, research on children’s justification of market-based inequalities in accessing welfare services remains limited, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.
+Yet, research on children's justification of market-based inequalities in accessing welfare services remains limited, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.
+
+```{r echo=FALSE, message=FALSE, error=FALSE}
+#| fig-cap: "Market justice preferences in health, pensions and education in the adult population"
+#| fig-asp: 0.6
+#| label: fig-elsoc
+
+elsoc_2016 <- elsoc_2016 %>% 
+  dplyr::select(Pensions = d02_01, 
+                Education = d02_02, 
+                Health = d02_03) %>% 
+  mutate_all(.funs = ~ set_na(., na = c(-666,-777,-888,-999), drop.levels = F) %>% 
+               sjmisc::rec(., rec = "rev", append = F) %>% 
+               factor(.,
+                      levels = c(1,2,3,4,5),
+                      labels = c("Strongly agree",
+                                 "Agree",
+                                 "Neither agree nor disagree",
+                                 "Desagree",
+                                 "Strongly desagree"))) %>% 
+  na.omit()
+
+elsoc_2016 %>% 
+  pivot_longer(cols = everything(),
+               names_to = "justice", 
+               values_to = "value") %>% 
+  mutate(justice = factor(justice, 
+                          levels = c("Health", "Pensions", "Education"))) %>% 
+  group_by(justice, value) %>%
+  summarise(total = n()) %>% 
+  mutate(prop = prop.table(total)*100) %>% 
+  ggplot(aes(x = value, y = prop, fill = value)) +
+  geom_col(position = "dodge", width = .9, alpha = 1, show.legend = T) +
+  scale_y_continuous(labels = function(prop){paste0(prop, "%")},
+                     limits = c(0,55),
+                     n.breaks = 7,
+                     expand = c(0,1)) +
+  facet_wrap(.~justice, scales = "fixed") +
+  geom_text(mapping = aes(label = scales::percent(prop, accuracy = 0.1, scale = 1, decimal.mark = ".")),
+            vjust = -0.5,
+            size = 3,
+            position = position_dodge(width = 0.8)) +
+  scale_fill_manual(values = c( "#F1EEF6", "#BDC9E1", "#b3b3b3ff", "#74A9CF", "#0570B0")) +
+  labs(x = NULL,
+       y = "Proportion (%)",
+       fill = NULL,
+       caption = paste0("Source: Authors calculation based on ELSOC 2016",
+                        " (N = ",dim(elsoc_2016)[1],")")) +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(size = 0))
+
+```
 
@@
 
-Schools contribute to institutionalizing and reproducing inequality by promoting values, norms, practices, and languages familiar to higher-class families because the dominant group’s culture shapes educational institutions [@bourdieu_reproduction_1990].
+Schools contribute to institutionalizing and reproducing inequality by promoting values, norms, practices, and languages familiar to higher-class families because the dominant group's culture shapes educational institutions [@bourdieu_reproduction_1990].
 Middle- and upper-class students are better equipped to face academic challenges and are more familiar with academic expectations [@mikus_children_2020].
@@
 Conversely, lower-status children lacking cultural capital must catch up while experiencing inequitable comparisons [@goudeau_hidden_2017].
-Additionally, academic achievement is treated as the outcome of dispositional factors (e.g., pupils’ efforts and talents or lack of them) rather than the result of differential access to critical resources.
+Additionally, academic achievement is treated as the outcome of dispositional factors (e.g., pupils' efforts and talents or lack of them) rather than the result of differential access to critical resources.
 Due to the meritocratic frame schools encourage, both low- and high-status individuals tend to believe that success or failure is not due to family background but rather to differences in efforts and talents [@darnon_where_2018].
@@
 
-Recent empirical research has demonstrated that the institutional design of schools, coupled with the meritocratic ideology it fosters, significantly influences children’s and adolescents’ views on inequality and deservingness.
+Recent empirical research has demonstrated that the institutional design of schools, coupled with the meritocratic ideology it fosters, significantly influences children's and adolescents' views on inequality and deservingness.
 For example, @jonsson_institutional_2015 study revealed that higher-status adolescents in Sweden tend to perpetuate social class stereotypes while describing the vocational and academic tracks.
@@
 
-[aca falta hipothesis de school performance]
+\[aca falta hipothesis de school performance\]
 
@@
 
-### Summary of hypotheses @Kevin-carrasco, actualizar esquema
+### Summary of hypotheses
 
@@
 
-knitr::include_graphics("input/img/hypotheses.png",
+knitr::include_graphics("input/img/hypothesis2.png",
   auto_pdf = getOption("knitr.graphics.auto_pdf", TRUE))

### 2024-06-10 04-analisis.qmd
Cambios

--- 2024-05-27 04-analisis.qmd
@@
 
-## Analysis
+# Results
 
-### Descriptive analysis
+## Descriptive
 
@@
 
-Likert plots
+@fig-marketjustice illustrates the distribution of responses to the three items on market justice preferences (healthcare, pensions, and education), indicating that most respondents tend to disagree or strongly disagree with the idea that it is just for people with higher incomes to have access to better services in these areas.
+The strongest opposition is observed in the healthcare sector, as the majority of respondents (66.4%) are against the idea that higher-income individuals should have access to better healthcare services.
+Similarly, regarding the concept of justice related to access to better pensions based on income, the level of disagreement reaches 63.0%.
+This decreases significantly in the field of education, as the percentage of respondents who disagree or strongly disagree with those with higher incomes having access to better education decreases to 58.8%.
+Only a small proportion agree that access to healthcare should be conditional on income (23.6%), and a minority strongly agree (10%), as is the case with pensions.
+However, this increases in the case of education, being the area where there is a significant proportion of respondents who agree or strongly agree with access based on market justice (41.2%).
+
+```{r echo=FALSE}
+#| fig-cap: "Market justice preferences in health, pensions and education"
+#| fig-asp: 0.6
+#| label: fig-marketjustice
+
+data_rec %>% 
+  select(just_salud, just_pension, just_educ) %>% 
+  sjPlot::plot_stackfrq(geom.size = 0.5,
+                        expand.grid =TRUE,
+                      geom.colors = "PuBu",
+                      show.total = FALSE, 
+                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
+                      axis.labels = c("Health", "Pensions", "Education")) +
+ # theme_light(base_size = 16) +
+  theme(legend.position="bottom", text = element_text(size = 14),
+        plot.caption = element_text(hjust = 1)) +
+       labs(caption = paste0("Source: Authors calculation based on First Study of Civic Education 2017",
+                             " (N = ",dim(data_rec)[1],")"))
+
+```
+
+Regarding the perception of meritocracy, @fig-meritocracy displays the frequency distribution of five items related to the dimensions of school and society.
+At the school level, there is a strong belief that both effort and talent are important for achieving good grades, with especially high agreement for effort.
+Specifically, while 78.6% agree or strongly agree on the importance of talent for obtaining good grades, this proportion rises to 96.8% for effort, indicating that respondents value effort more than talent in this context.
+
+```{r echo=FALSE}
+#| fig-cap: "Social and school meritocracy"
+#| fig-asp: 0.8
+#| label: fig-meritocracy
+
+plot<-data_rec %>% 
+  select(merito_soc, esfuerzo_soc, inteligencia_soc, esfuerzo_esc, inteligencia_esc) %>% 
+  sjPlot::plot_stackfrq(geom.size = 0.5,
+                        expand.grid =TRUE,
+                      geom.colors = "PuBu",
+                      show.total = FALSE, 
+                      vjust=c(rep(c("center"),12), "bottom", "top", rep(c("center"), 6)),
+                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
+                      axis.labels = c("Deservingness", "Social effort", "Social talent", "School effort", "School talent")) +
+  #theme_light(base_size = 16) +
+  theme(legend.position="bottom", text = element_text(size = 14),
+        plot.caption = element_text(hjust = 1))
+
+plot + 
+  annotate("text", x = 5.5, y = 0, label = "School meritocracy", size = 5, hjust = 0, vjust = 1) +
+  annotate("text", x = 3.5, y = 0, label = "Social meritocracy", size = 5, hjust = 0, vjust = 1) +
+       labs(caption = paste0("Source: Authors calculation based on First Study of Civic Education 2017",
+                             " (N = ",dim(data_rec)[1],")"))
+```
+
+At the societal level, @fig-meritocracy shows that a majority of respondents agree or strongly agree that individuals are rewarded for their talents (63.7%) and efforts (62%).
+Nevertheless, a notable proportion perceives that both effort and talent are not adequately rewarded in society, with 38.1% and 36.3% respectively.
+The perception of whether people get what they deserve in society is more divided; while a slim majority (56.5% combining agree and strongly agree) agrees, a significant minority (43.6%) disagrees, indicating that this is the most contentious item among respondents.
+There is general agreement that people are rewarded for talent and effort in society, but the consensus is weaker compared to school meritocracy.
 
@@
 plot_soc_effort <- soc_effort %>% 
-  mutate(justice = str_to_title(justice)) %>% 
+  mutate(justice = str_to_title(justice),
+         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
+  arrange(esfuerzo_soc, justice) %>% 
   ggplot(aes(x = esfuerzo_soc, y = mean)) +
@@
                                "Strongly\n agree")) +
-  scale_color_grey() +
+  scale_color_brewer(palette = "Paired") +
   labs(x = "Social effort",
@@
 plot_deservingness <- deservingness %>% 
-  mutate(justice = str_to_title(justice)) %>% 
+   mutate(justice = str_to_title(justice),
+         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
+  arrange(merito_soc, justice) %>% 
   ggplot(aes(x = merito_soc, y = mean)) +
@@
                                "Strongly\n agree")) +
-  scale_color_grey() +
+  scale_color_brewer(palette = "Paired") +
   labs(x = "Deservingness",
@@
 plot_soc_talent <- soc_talent %>% 
-  mutate(justice = str_to_title(justice)) %>% 
+   mutate(justice = str_to_title(justice),
+         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
+  arrange(inteligencia_soc, justice) %>%  
   ggplot(aes(x = inteligencia_soc, y = mean)) +
@@
                                "Strongly\n agree")) +
-  scale_color_grey() +
+  scale_color_brewer(palette = "Paired") +
   labs(x = "Social talent",
@@
 plot_school_effort <- school_effort %>% 
-  mutate(justice = str_to_title(justice)) %>% 
+  mutate(justice = str_to_title(justice),
+         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
+  arrange(esfuerzo_esc, justice) %>%  
   ggplot(aes(x = esfuerzo_esc, y = mean)) +
@@
                                "Strongly\n agree")) +
-  scale_color_grey() +
+  scale_color_brewer(palette = "Paired") +
   labs(x = "School effort",
@@
 plot_school_talent <- school_talent %>% 
-  mutate(justice = str_to_title(justice)) %>% 
+  mutate(justice = str_to_title(justice),
+         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
+  arrange(inteligencia_esc, justice) %>% 
   ggplot(aes(x = inteligencia_esc, y = mean)) +
@@
                                "Strongly\n agree")) +
-  scale_color_grey() +
+  scale_color_brewer(palette = "Paired") +
   labs(x = "School talent",
@@
 
-### Multivariate results
+## Multivariate
 
-The results of the cumulative link mixed models for the ordinal dependent variables of justice in differential access to education, health, and pensions by income are presented in @tbl-ordinal-reg.
-This table shows the effects of individual and contextual-level independent variables related to our hypothesis; however, the models also incorporate the control variables utilized in the study for each of the three dependent variables separately (complete models are available in the [Appendix](#appendix)).
+The results of the cumulative link mixed models for the ordinal dependent variables of justice in differential access to health, pensions and education by income are presented in @tbl-ordinal-reg[^1].
 
@@
 
-tbls <- list(m1_p, m2_p, m3_p)
+modelos_pension <- list(m1_p, m2_p, m3_p)
 
-modelos_odd_pension <- map(tbls, odds_fun)
+# modelos_odd_pension <- map(tbls, odds_fun)
 
@@
 
-tbls <- list(m1_h, m2_h, m3_h)
+modelos_health <- list(m1_h, m2_h, m3_h)
 
-modelos_odd_health <- map(tbls, odds_fun)
+# modelos_odd_health <- map(tbls, odds_fun)
 
@@
 
-tbls <- list(m1_e, m2_e, m3_e)
+modelos_educ <- list(m1_e, m2_e, m3_e)
 
-modelos_odd_educ <- map(tbls, odds_fun)
+#modelos_odd_educ <- map(tbls, odds_fun)
 
@@
   "educacion_recNs/Nr" = "Missing",
-  mean_educ = "Prop. university level at school",
+  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
   "cod_grupoMedio bajo" = "SES Medium low",
@@
 
-texreg::htmlreg(l = list(modelos_odd_educ[[2]],
-                         modelos_odd_educ[[3]],
-                         modelos_odd_health[[2]],
-                         modelos_odd_health[[3]],
-                         modelos_odd_pension[[2]],
-                         modelos_odd_pension[[3]]),   
+texreg::htmlreg(l = list(modelos_health[[2]],
+                         modelos_health[[3]],
+                         modelos_pension[[2]],
+                         modelos_pension[[3]],
+                         modelos_educ[[2]],
+                         modelos_educ[[3]]),   
+                caption = NULL,
                 stars = c(0.05, 0.01, 0.001),
-                  custom.note = "Note: Coefficients models in odds-ratio. Control variables are included. Standard errors in parentheses. %stars",
-                  custom.header = list("Education" = 1:2, "Health" = 3:4, "Pensions" = 5:6),
+                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars",
+                custom.header = list("Health" = 1:2, "Pensions" = 3:4, "Education" = 5:6),
                 custom.model.names = rep(c("Model 1", "Model 2"), 3),
@@
                                 "Achievement score (Ref.= Low)" = 15:16),
-                 custom.gof.rows = list("Controls"=rep("Yes",6)))
+                custom.gof.rows = list("Controls"=rep("Yes",6)),
+                include.loglik = FALSE, include.aic = FALSE)
 
@@
 
-texreg::htmlreg(list(reg1_bienestar, reg2_bienestar),
+ccoef <- list(
+  "(Intercept)" = "Intercept",
+  inteligencia_esc = "School talent",
+  esfuerzo_esc = "School effort",
+  inteligencia_soc = "Social talent",
+  esfuerzo_soc = "Social effort",
+  merito_soc = "Deservingness",
+  "educacion_recEd Media" = "Secondary",
+  "educacion_recEd Tecnica" = "Higher tec.",
+  "educacion_recUniversidad o posgrado" = "University or posgraduate",
+  "educacion_recNs/Nr" = "Missing",
+  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
+  "cod_grupoMedio bajo" = "SES Medium low",
+  cod_grupoMedio = "SES Medium",
+  "cod_grupoMedio alto" = "SES Medium high",
+  cod_grupoAlto = "SES High",
+  simceMedio = "Simce Medium",
+  simceAlto = "Simce High"
+)
+
+texreg::htmlreg(list(reg1_bienestar, reg2_bienestar, reg4_bienestar, reg5_bienestar),
                 custom.model.names = c("Model 1",
-                               "Model 2"),
-        custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
-        omit.coef = "(internet_rec)",
-        caption = "",
-        custom.coef.names = c("Intercept",  
-                              "School talent",
-                              "School effort", 
-                              "Social talent",  
-                              "Social effort",
-                              "Deservingness",
-                              "Secondary", 
-                              "Higher tec.", 
-                              "University or posgraduate", 
-                              "Missing", 
-                              "More than 25 books (Ref. Less than 25)", 
-                              "Technology access"),
-        groups = list("Parental education (Ref.= 8th grade or less)"=7:10),
-        custom.gof.rows=list("Deviance" = c(deviance1,
-                                            deviance2),
+                                       "Model 2",
+                                       "Model 3",
+                                       "Model 4"),
+                caption = NULL,
+                stars = c(0.05, 0.01, 0.001),
+                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars",
+                custom.coef.map = ccoef,
+                leading.zero = T,
+                scalebox = 0.60,
+                groups = list("Parental education (Ref.= 8th grade or less)" = 7:10,
+                              "Socioeconomic level (Ref.= Low)" = 12:15,
+                              "Achievement score (Ref.= Low)" = 16:17),
+                custom.gof.rows = list("Controls"=rep("Yes",4),
+                                       "Deviance" = c(deviance1,
+                                                    deviance2,
+                                                    deviance4,
+                                                    deviance5),
                                        "Deviance Test (p)" = c(test.pvalues1,
-                                                       test.pvalues2))
-        )
+                                                             test.pvalues2,
+                                                             test.pvalues4,
+                                                             test.pvalues5)),
+                include.loglik = FALSE,include.aic = FALSE)
 
@@
 ```
+
+[^1]: This table shows the effects of individual and contextual-level independent variables related to our hypothesis; however, the models also incorporate the control variables utilized in the study for each of the three dependent variables separately (complete models are available in the [Appendix](#appendix)).

### 2024-06-10 paper.qmd
Cambios

@@
     keep-tex: true
-editor: visual
 smooth-scroll: true

### 2024-06-11 05-conclusiones.qmd
update discusion

@@
 While we found support for some associations, others did not align with our initial expectations.
-Hypothesis H1a posited that students who perceive greater meritocracy in society would show stronger preferences for market justice.
-The results confirm this hypothesis, indicating a positive relationship between the perception that talent and effort are rewarded in society and the preferences for market justice in access to services such as health, education, and pensions.
+Hypothesis $H_{1a}$ posited that students who perceive greater meritocracy in society would show stronger preferences for market justice.
+The results support this hypothesis, indicating a positive relationship between the perception that talent and effort are rewarded in society and the preferences for market justice in access to services such as health, education, and pensions.
 This finding aligns with previous research suggesting that individuals who believe in a meritocratic system are more likely to justify social inequalities [@batruch_belief_2022; @wiederkehr_belief_2015].

### 2024-06-12 01-introduction.qmd
revisada introduccion y antecedentes, temas menores formato
Update intro and analysis

@@
 format: html
-editor: visual
 bibliography: "input/bib/just-ineq-merit.bib"
@@
 
-## Introduction
+# Introduction
 
@@
 In the present paper, we address the role of the perception of meritocracy on the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system.
-In particular, we focus in justifying inequalities in health, education and pensions, which are traditional social policy areas where acces to better services could be justified based on payment capacity, referred as to market justice [@lane_market_1986, @lindh_bringing_2023].
+In particular, we focus in justifying inequalities in health, education and pensions, which are traditional social policy areas where acces to better services could be justified based on payment capacity, referred as to market justice [@lane_market_1986, @lindh_bringing_2023].\
 Most of the research in this realm has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies [@batruch_belief_2022].
@@
 
-++ párrafo contexto Chileno dar cifras/porcentajes de privatización de salud, educación y pensiones; asociar al tema de la calidad (lo público es peor); hay información sobre esto que se puede sacar del proyecto Fondecyt
+The Chilean case is particularly intriguing for studying market justice preferences.
+This country is characterized by acute and persistent economic inequality, which stands out in Latin America and among OECD countries [@flores_top_2020].
+In Chile, the poorest 50% captures only 10% of the total income and has negative wealth, while the richest 1% receives almost 27% of the income and holds 49.6% of the wealth [@chancel_world_2022].
+Much of this inequality is linked to the deep neoliberal reforms that institutionalized privatization and commodification of various economic sectors, including public services, introduced during the dictatorship (1973-1989) and expanded in democracy through concessions, demand credits, and specific regulatory frameworks [@ffrench-davis_reformas_2018].
+This shift in economic policy has allowed the unprecedented emergence of markets in areas related to social reproduction, such as health, pensions, and education, with provision and access managed by private entities and segmented by individual payment capacity, heavily reliant on State subsidies [@boccardo_30_a].
+In health, although the majority of the population uses the public insurance system (78.9%), 15.3% are served by private insurers (CASEN, 2022).
+The pension system is based on individual capitalization, with mandatory contributions managed by private administrators investing in the financial market, currently involving 11 million contributors (Superintendencia de Pensiones, 2024).
+In education, 30.6% of school enrollment is in public schools, 54.0% in State-subsidized private schools, and 9.3% in fully private schools, generally attracting higher-income groups (Mineduc, 2023).

### 2024-06-12 02-antecedentes.qmd
Update intro and analysis

@@
 
-theme_nice <- function() {
-  theme_bw() +
-    theme(text = element_text(family = "serif"))
-
-}
-
-theme_set(theme_nice())
+theme_set(theme_bw())
 ```
@@
   mutate_all(.funs = ~ set_na(., na = c(-666,-777,-888,-999), drop.levels = F) %>% 
-               sjmisc::rec(., rec = "rev", append = F) %>% 
                factor(.,
                       levels = c(1,2,3,4,5),
-                      labels = c("Strongly agree",
+                      labels = c("Strongly desagree",
+                                 "Desagree",
+                                 "Neither agree nor disagree",
                                  "Agree",
-                                 "Neither agree nor disagree",
-                                 "Desagree",
-                                 "Strongly desagree"))) %>% 
+                                 "Strongly agree"))) %>% 
   na.omit()
@@
             position = position_dodge(width = 0.8)) +
-  scale_fill_manual(values = c( "#F1EEF6", "#BDC9E1", "#b3b3b3ff", "#74A9CF", "#0570B0")) +
+  scale_fill_manual(values = c("#F1EEF6", "#BDC9E1", "#b3b3b3ff", "#74A9CF", "#0570B0")) +
   labs(x = NULL,
@@
 From a sociological perspective, meritocracy has been used in research on social mobility to characterize societies with low mobility that threaten the meritocratic ideal [@goldthorpe_myth_2003].
-More recently, sociology and social psychology research has attended to subjective aspects related to the support for meritocratic principles in different societies, such as beliefs in meritocracy [@castillo_multidimensional_2023a; @mijs_unfulfillable_2016; @mijs_paradox_2019].
+More recently, sociology and social psychology research has attended to subjective aspects related to the support for meritocratic principles in different societies, such as beliefs in meritocracy [@castillo_multidimensional_2023; @mijs_unfulfillable_2016; @mijs_paradox_2019].
 The label of beliefs in this realm covers different concepts, but in this paper we refer to a conceptual model developed by @castillo_multidimensional_2023 which identify two main aspects: preferences and perceptions for meritocracy.

### 2024-06-12 04-analisis.qmd
Update intro and analysis

@@
 
-#windowsFonts(`Times New Roman` = windowsFont("Times New Roman"))
-
-theme_nice <- function() {
-  theme_bw() +
-    theme(text = element_text(family = "serif"))
-
-}
-
-theme_set(theme_nice())
+theme_set(theme_bw())
 
@@
 ggarrange(plot_soc_effort, plot_school_effort, plot_soc_talent,
-          plot_school_talent, plot_deservingness, leg, ncol = 2, nrow = 3)
+          plot_school_talent, plot_deservingness, leg, ncol = 2, nrow = 3) +
+  labs(caption = paste0("Source: Authors calculation based on First Study of Civic Education 2017",
+                             " (N = ",dim(data_rec)[1],")"))
 ```
@@
 
-The results of the cumulative link mixed models for the ordinal dependent variables of justice in differential access to health, pensions and education by income are presented in @tbl-ordinal-reg[^1].
+This section presents the results of the multilevel models.
+We begin by showing the results of the cumulative link mixed models in @tbl-ordinal-reg for the ordinal dependent variables of justice in health, pensions, and education separately, to determine if the effect of the predictors differs across these items.
+Then, we present the results of the linear mixed-effects models for the average index of preference for market justice in @tbl-lineal-reg[^1].
+For the health item, the intraclass correlation (ICC) obtained for the null model (not shown) indicates that 5% of the total variance of this item is associated with specific characteristics of the schools, while it is 3% for the pensions item and 4% for the education item (ICC based on [@hox_multilevel_2010, p.15]).
+For the average index of market justice preferences, the ICC reaches 4%, indicating that overall, only a small percentage of the total variance in students’ market justice preferences is associated with school characteristics, thus limiting the possibility of finding effects at this second level of analysis.
+
+[^1]: The tables shows the effects of individual and contextual-level independent variables related to our hypothesis; however, the models also incorporate the control variables utilized in the study for each of the three dependent variables separately and for the average index (complete models are available in the [Appendix](#appendix)).
 
@@
 ```{r echo=FALSE, results='asis'}
-#| tbl-cap: "Cumulative link multilevel models of differential acces justificacion of education, health and pensions"
+#| tbl-cap: "Cumulative link multilevel models of differential acces justificacion of health, pensions, and education"
 #| label: tbl-ordinal-reg
@@
 
-@tbl-ordinal-reg demonstrates a similar trend for the three dependent variables of justice in differential access to pensions, education, and health, with all effects being statistically significant.
+Model 1 in @tbl-ordinal-reg for the three dependent variables of justice in differential access to health, pensions, and education demonstrates a similar trend regarding the effect of the perception of meritocracy on these items, with all effects being statistically significant (*p* < 0.05).
+Regarding the variables of social meritocracy, the three variables—talent, effort, and deservingness—show that as these increase, the justification for differentiated access to pensions, education, and health also increases.
 In the context of school meritocracy, the effects are mixed: as perceived school talent increases, the justification for differentiated access to these services increases; conversely, as perceived school effort increases, the justification decreases, holding all other variables constant.
-Regarding the variables of social meritocracy, the three variables of talent, effort, and deservingness show that as these increase, the justification for differentiated access to pensions, education, and health also increases.
-Overall, these results provide partial support for our first hypothesis (H1a), which posited that students who perceive higher levels of meritocracy in school would be more likely to justify differential access to pensions, health, and education.
-This is evidenced by the positive effect of perceiving talent as important for achieving good grades.
-However, the results strongly support our second hypothesis (H1b), as a higher perception of societal meritocracy among students tends to increase the justification for market-based justice in these social services.
-CAMBIAR
 
-Contextual effects for ordinal models.
+Additionally, Model 1 shows the effect of family status, measured through the highest educational level of the parents and the number of books at home.
+For the three items—health, pensions, and education—a negative but statistically non-significant association with parents' educational level is observed.
+However, the results suggest a negative and statistically significant effect of the number of books at home for market justice in health and pensions, indicating that those with more than 25 books at home tend to justify less differentiated access in these areas compared to those with fewer than 25 books, controlling for other variables.
+This association is not observed for market justice in education.
+
+Model 2 in @tbl-ordinal-reg incorporates school-level variables for the three items of market justice in health, pensions, and education.
+Regarding the socioeconomic level of the schools, the results show that high socioeconomic status schools maintain greater preferences for market justice in education compared to low socioeconomic status schools, with a statistically significant association (*p* < 0.05).
+This effect is not observed for the health and pensions variables.
+Moreover, the academic performance of schools shows negative and statistically significant effects for the three items of market justice separately.
+The findings suggest that schools with higher academic performance tend to justify market allocation principles less for access to health, pensions, and education compared to low-performing schools.
+This difference is also observed in schools with medium performance, but only for the domains of health and pensions.
 
@@
 ```{r echo=FALSE, results='asis'}
-#| tbl-cap: "Multilevel models for meritocracy perception and market justice preferences"
-#| label: tbl-individual-reg
+#| tbl-cap: "Linear mixed-effects models for meritocracy perception and market justice preferences"
+#| label: tbl-lineal-reg
 
 ccoef <- list(
-  "(Intercept)" = "Intercept",
+  "(Intercept)" = "Constant",
   inteligencia_esc = "School talent",
@@
 ```
-
-[^1]: This table shows the effects of individual and contextual-level independent variables related to our hypothesis; however, the models also incorporate the control variables utilized in the study for each of the three dependent variables separately (complete models are available in the [Appendix](#appendix)).

### 2024-06-13 01-introduction.qmd
Update intro, antecedentes, metodos y analisis

@@
 Much of this inequality is linked to the deep neoliberal reforms that institutionalized privatization and commodification of various economic sectors, including public services, introduced during the dictatorship (1973-1989) and expanded in democracy through concessions, demand credits, and specific regulatory frameworks [@ffrench-davis_reformas_2018].
-This shift in economic policy has allowed the unprecedented emergence of markets in areas related to social reproduction, such as health, pensions, and education, with provision and access managed by private entities and segmented by individual payment capacity, heavily reliant on State subsidies [@boccardo_30_a].
+This shift in economic policy has allowed the unprecedented emergence of markets in areas related to social reproduction, such as health, pensions, and education, with provision and access managed by private entities and segmented by individual payment capacity, heavily reliant on State subsidies [@boccardo_30_2020].
 In health, although the majority of the population uses the public insurance system (78.9%), 15.3% are served by private insurers (CASEN, 2022).
@@
 In education, 30.6% of school enrollment is in public schools, 54.0% in State-subsidized private schools, and 9.3% in fully private schools, generally attracting higher-income groups (Mineduc, 2023).
+
+The provision and access to services in these areas under market principles correlate with the perceptions, preferences, and attitudes of Chilean society.
+Recent public opinion surveys conducted by the Center for Public Studies indicate that, generally, people prefer public insurance (64.1%) over private insurance (35.9%) if both cost the same.
+In pensions, Chileans show varied preferences regarding the public or private status of pension fund administrators; 52.3% favor both private and public institutions, 38.8% prefer only public institutions, and 8.6% prefer only private ones.
+This likely relates to the severe legitimacy crisis faced by private pension fund administrators, which has sparked social movements in recent years [@boccardo_30_2020].
+Interestingly, in education, there is a clear preference for private schools over public ones if both cost the same and are equally distant, with 63% preferring private and 37% public.

### 2024-06-13 03-methods.qmd
Cambios
update discusion
Update intro, antecedentes, metodos y analisis

@@
 
-df<-dfSummary(school,
+df<-dfSummary(school[,c(4,2)],
                plain.ascii = FALSE,
@@
 
-A series of individual control variables are included.
+A series of control variables are included.
 At the individual level, an index of access to technology is constructed based on the number of computers, tablets, and cell phones in the household, as well as internet connectivity.

### 2024-06-13 04-analisis.qmd
Update intro, antecedentes, metodos y analisis

@@
 However, this increases in the case of education, being the area where there is a significant proportion of respondents who agree or strongly agree with access based on market justice (41.2%).
+
+This contrasts with the pattern observed for these items in the Chilean adult population (see @fig-elsoc).
+While both age groups tend to disagree that higher-income individuals should access better health, pension, and education services, this tendency is more pronounced in adults than in students, who show greater agreement with market justice preferences in these areas.
+Additionally, the hierarchy of these domains differs between groups; adults justify market justice least in education, followed by health, and then pensions, whereas students justify market principles less in health, then pensions, and lastly in education.
 
@@
 For the health item, the intraclass correlation (ICC) obtained for the null model (not shown) indicates that 5% of the total variance of this item is associated with specific characteristics of the schools, while it is 3% for the pensions item and 4% for the education item (ICC based on [@hox_multilevel_2010, p.15]).
-For the average index of market justice preferences, the ICC reaches 4%, indicating that overall, only a small percentage of the total variance in students’ market justice preferences is associated with school characteristics, thus limiting the possibility of finding effects at this second level of analysis.
+For the average index of market justice preferences, the ICC reaches 4%, indicating that overall, only a small percentage of the total variance in students' market justice preferences is associated with school characteristics, thus limiting the possibility of finding effects at this second level of analysis.
 
@@
 
-Model 1 in @tbl-ordinal-reg for the three dependent variables of justice in differential access to health, pensions, and education demonstrates a similar trend regarding the effect of the perception of meritocracy on these items, with all effects being statistically significant (*p* < 0.05).
-Regarding the variables of social meritocracy, the three variables—talent, effort, and deservingness—show that as these increase, the justification for differentiated access to pensions, education, and health also increases.
+Model 1 in @tbl-ordinal-reg for the three dependent variables of justice in differential access to health, pensions, and education demonstrates a similar trend regarding the effect of the perception of meritocracy on these items, with all effects being statistically significant (*p* \< 0.05).
+Regarding the variables of social meritocracy, the three variables---talent, effort, and deservingness---show that as these increase, the justification for differentiated access to pensions, education, and health also increases.
 In the context of school meritocracy, the effects are mixed: as perceived school talent increases, the justification for differentiated access to these services increases; conversely, as perceived school effort increases, the justification decreases, holding all other variables constant.
 
-Additionally, Model 1 shows the effect of family status, measured through the highest educational level of the parents and the number of books at home.
-For the three items—health, pensions, and education—a negative but statistically non-significant association with parents' educational level is observed.
-However, the results suggest a negative and statistically significant effect of the number of books at home for market justice in health and pensions, indicating that those with more than 25 books at home tend to justify less differentiated access in these areas compared to those with fewer than 25 books, controlling for other variables.
-This association is not observed for market justice in education.
+Additionally, Model 1 examines family status through parents' educational level and the number of books at home.
+For health, pensions, and education, a negative but statistically non-significant association with parents' education is observed.
+However, having more than 25 books at home shows a negative and statistically significant effect on market justice in health and pensions, suggesting less justification for differentiated access compared to those with fewer books.
+This effect is not seen in education.
 
-Model 2 in @tbl-ordinal-reg incorporates school-level variables for the three items of market justice in health, pensions, and education.
-Regarding the socioeconomic level of the schools, the results show that high socioeconomic status schools maintain greater preferences for market justice in education compared to low socioeconomic status schools, with a statistically significant association (*p* < 0.05).
-This effect is not observed for the health and pensions variables.
-Moreover, the academic performance of schools shows negative and statistically significant effects for the three items of market justice separately.
-The findings suggest that schools with higher academic performance tend to justify market allocation principles less for access to health, pensions, and education compared to low-performing schools.
-This difference is also observed in schools with medium performance, but only for the domains of health and pensions.
+Model 2 in @tbl-ordinal-reg incorporates school-level variables.
+The results show that high socioeconomic level schools show greater preferences for market justice in education compared to low socioeconomic level schools, with statistically significant results (*p* \< 0.05).
+This effect is not seen for health and pensions.
+Moreover, the findings suggest that schools with higher academic performance shows negative and statistically significant effects across all three items, suggesting these schools justify market allocation principles less.
+This pattern is also observed in medium-performing schools, but only for health and pensions.
 
@@
 
-Regarding our measure of market justice preferences index, the results of the multilevel linear estimation are shown in @tbl-individual-reg.
-The intraclass correlation obtained for the null model (not shown) it's equal to 4% (ICC based on [@hox_multilevel_2010, p.15]), which indicates that only a small percentage of the total variance of the students' market justice preferences are associated with the characteristics of schools, and therefore limits the possibilities of finding effects at this second level of analysis.
+@tbl-lineal-reg presents the results of linear mixed-effects models for the market justice preference index.
+Model 1 shows that higher perceived societal meritocracy, through effort, talent, and deservingness, is positively and significantly associated with market justice preferences.
+School meritocracy perceptions yield mixed effects: perceived talent increases market justice preference, while perceived effort decreases it, both statistically significant (*p* \< 0.001).
+Model 2 includes family status, finding no statistically significant effects from parents' education or the number of books at home, both showing negative directions.
 
-Model 1 introduces social meritocratic variables: effort (whether efforts are rewarded), deservingness (people get what they deserve), and social talent (intelligence and skills are rewarded in society).
-In line with our hypotheses, the perception of a meritocratic society is positively related to the justification of inequality.
-Model 2 shows a mixed picture: while those perceiving that talent is rewarded also justify the inequality, the perception that effort is rewarded at school is negatively related to justification of inequality.
-Family background variables in Model 3 reveal that education and technology access are not related to justification of inequality, whereby we observe a negative impact of family cultural capital as measured by the number of books at home.
-While school socio-structural variables added in Model 4 show no significant effects, average achievement scores in the SIMCE test depict a negative relationship with the dependent variable, meaning that students that attend schools with better achievement scores on average justify less inequality.
-
-In relation to model fit, when comparing the deviance with a model without predictors (null model), all the models have a statistically significant difference, with model 5 having the lowest deviance.
-According to Raudenbush and Bryk's (2002) estimate of R2, the level 1 variance of model 5 is 0.11 and the level 2 variance is 0.72.
-The total variance of model 5 according to Snijders and Bosker (2012) is 0.14.
+Regarding school characteristics, Models 3 and 4 introduce socioeconomic level and academic performance.
+The results of Model 3 suggest no statistically significant relationship between school socioeconomic levels and market justice preferences.
+However, Model 4 indicates that schools with medium ($\beta$ = -0.10) and high ($\beta$ = -0.24) academic performance have lower market justice preferences compared to low-performing schools, with these effects being statistically significant when controlling for other variables.
 
@@
 load("input/data/interact.RData")
-kbl(collapse_rows,
+
+collapse_rows[c(-4),] %>% 
+  as_tibble() %>% 
+  kbl(.,
       align = "lccccc", booktabs = T,
@@
                      "Market Justice Preferences"=5), 
-                   bold = T)%>%  
+                   bold = T) %>% 
   kable_styling(latex_options = c("scale_down","HOLD_position"),
@@
   collapse_rows(columns = 1:2, valign = "middle") %>% 
-  footnote(general = "***p < 0.001, **p < 0.01, *p < 0.05",
-           footnote_as_chunk = T,general_title = "Note: ")
+  kableExtra::add_footnote(label = "Note: ***p < 0.001, **p < 0.01, *p < 0.05", 
+                           notation = "none")
 
@@
 
-The interaction terms in @tbl-interact suggest that students' family background moderates the relationship between their meritocratic perceptions in Chile and their justification of inequality.
-The direction of these effects confirms our initial prediction.
-Thus, in line with our Hypothesis 3, the relationship between social effort and justification of inequality becomes less positive for those students whose parents achieved university or postgraduate education.
-That is, lower-status students (measured by parental education) justify more inequality when they adhere more strongly to deservingnessmeritocratic perceptions in Chile.
-This result confirms the enlightenment thesis (see above).
-The same moderating effect is observed for students' family cultural capital: the relationship between effort (but not deservingness nor social talent) and justification of inequality becomes less positive for students with more than 25 books at home.
-Interestingly, family cultural capital also moderates the relationship between meritocratic beliefs at school (as measured by students' belief that effort is important to get good grades) and justification of inequality.
+@tbl-interact displays the interaction terms related to our hypothesis.
+The results suggest that students' family background moderates the relationship between their meritocratic perceptions and their justification of market allocation in social services.
+Specifically, the relationship between social effort and market justice preferences is less positive for students whose parents have university or postgraduate education.
+Lower-status students (measured by parental education) prefer more market justice when they strongly adhere to meritocratic perceptions of deservingness.
+A similar moderating effect is seen with family cultural capital: the relationship between effort and market justice preferences is less positive for students with more than 25 books at home.
 
-At the school level, we also observe moderating effects of school status and students' meritocratic beliefs over their justification of inequality.
-Thus, in line with our Hypothesis 5, high-status schools justify less inequality when, on average, their students have a greater perception of meritocracy in Chile, as measured by the three indicators we used (i.e., effort, deservingness, and talent).
-Finally, as we stated in Hypothesis 7, low-achieving schools justify more inequality when their students have, on average, stronger meritocratic beliefs in Chile.
-In @fig-interaction we depict this moderating effect of school achievement for the relationship between deservingness and justification of inequality.
-We did not observe these moderating effects at the school level for students' meritocratic beliefs in the school (i.e., the idea that effort or talent are important to get good grades).
+At the school level, we observe that school status and students' meritocratic beliefs moderate their preferences for market justice.
+High-status schools show less preference for market justice when their students, on average, have a greater perception of meritocracy, as measured by effort, deservingness, and talent.
+Conversely, low-achieving schools prefer more market justice when their students have stronger meritocratic beliefs.
+@fig-interaction illustrates this moderating effect of school achievement on the relationship between deservingness and market justice.
+No moderating effects were observed at the school level for students' meritocratic beliefs regarding effort or talent in achieving good grades.
 

### 2024-06-13 04-analisis.qmd

@@
 There is general agreement that people are rewarded for talent and effort in society, but the consensus is weaker compared to school meritocracy.
+
+*Market justice preferences*
+
+```{r echo=FALSE}
+#| fig-cap: "Market justice preferences in education, health and pensions"
+#| fig-asp: 1.22
+#| label: fig-marketjustice
+
+data_rec %>% 
+  select(just_salud, just_educ, just_pension) %>% 
+  sjPlot::plot_stackfrq(expand.grid =TRUE,
+                      geom.colors = "PuBu",
+                      show.total = FALSE, 
+                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
+                      axis.labels = c("Health", "Education", "Pensions")) +
+  theme_light(base_size = 16) +
+  theme(legend.position="bottom", text = element_text(size = 14),
+        plot.caption = element_text(hjust = 1))
+```
+
+@fig-marketjustice shows the frequency distribution for the three market justice preference items (pensions, education, and health).
+The majority of respondents disagree (63.0% in total) with the idea that it is just for higher-income individuals to access better pensions.
+However, a notable portion (27.0%) agrees, and a smaller minority (10.1%) strongly agrees.
+Similar to pensions, a majority of respondents (58.8%) disagree with the idea that higher-income individuals should have access to better education.
+However, a considerable portion (28.6%) agrees, and a slightly larger minority (12.6%) strongly agrees compared to pensions.
+In the context of health, an even larger majority of respondents (66.4%) disagree with the notion that higher-income individuals should access better healthcare services.
+A smaller percentage (23.6%) agrees, and a minority (10.0%) strongly agrees.
+Overall, @fig-marketjustice shows a predominant sentiment against the justice of higher-income individuals accessing better services across all three areas.
+The strongest opposition is observed in the health sector.
+Despite the overall disagreement, a significant portion of respondents agree with the statement, particularly in education, where agreement and strong agreement are slightly higher compared to pensions and health.
+
+*School and social meritocracy*
+
+```{r echo=FALSE}
+#| fig-cap: "Social and school meritocracy"
+#| fig-asp: 1.22
+#| label: fig-meritocracy
+
+plot<-data_rec %>% 
+  select(merito_soc, esfuerzo_soc, inteligencia_soc, esfuerzo_esc, inteligencia_esc) %>% 
+  sjPlot::plot_stackfrq(expand.grid =TRUE,
+                      geom.colors = "PuBu",
+                      show.total = FALSE, 
+                      vjust=c(rep(c("center"),12), "bottom", "top", rep(c("center"), 6)),
+                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
+                      axis.labels = c("Deservingness", "Social effort", "Social talent", "School effort", "School talent")) +
+  theme_light(base_size = 16) +
+  theme(legend.position="bottom", text = element_text(size = 14),
+        plot.caption = element_text(hjust = 1))
+
+plot + 
+  annotate("text", x = 5.5, y = 0, label = "School meritocracy", size = 5, hjust = 0, vjust = 1) +
+  annotate("text", x = 3.5, y = 0, label = "Social meritocracy", size = 5, hjust = 0, vjust = 1)
+```
+
+@fig-meritocracy shows the frequency distribution for the five meritocracy items for the dimensions of school and social meritocracy.
+While a significant portion of respondents (29.2%) strongly agrees that talent is important to get good grades, a larger majority (78.6% combining agree and strongly agree) supports this notion.
+However, the strong agreement is not as pronounced as for school effort.
+An overwhelming majority (73.1%) agree that students effort is important to get good grades, which is a much higher percentage compared to the strong agreement for talent (29.2%).
+This indicates that people believe effort is more important than talent at school.
+There is a strong belief that both effort and talent are important, with a particularly high agreement for effort (73.1% strongly agree).
+
+In relation with social meritocracy, @fig-meritocracy shows that a majority of respondents (63.7% combining agree and strongly agree) agree that people are rewarded for their talent in society, with a notable portion (17.1%) strongly agreeing.
+A majority (62.0% combining agree and strongly agree) agree that people are rewarded for their effort in society.
+Similar to social talent, the perception is positive but less overwhelming than school meritocracy.
+The perception of whether people get what they deserve in society is more divided.
+While a slim majority (56.5% combining agree and strongly agree) agree, a significant minority (43.6%) disagrees, showing that this is the most contentious item among respondents.
+There is general agreement that people are rewarded for talent and effort in society, but the consensus is weaker compared to school meritocracy.
+Deservingness in society is the most contentious, with significant disagreement among respondents.
+
+*Market justice preferences in education, health and pensions by social and school meritocracy*
 

### 2024-06-13 02-antecedentes.qmd

@@
 Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work.
-In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2021].
+In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2019].
 Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger market justice preferences, as individual achievement would be seen as rewarded and social policies as less necessary.
 
-Some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2021], and how larger inequality affects meritocratic beliefs [@morris_representing_2022].
+Some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2019], and how larger inequality affects meritocratic beliefs [@morris_representing_2022].
 A larger justification of meritocratic distribution have been related to less support for redistributive compensation systems [@frank_performance_2015].

### 2024-06-13 05-conclusiones.qmd
update discussion

@@
 
+The previous analyses were aimed at delving into the relationship between perceptions of meritocracy and preferences for market justice among students.
+Several models estimated how students' beliefs about the rewards for talent and effort in both society and school environments influence their attitudes towards access to social services like health, education, and pensions.
+The models also examined the impact of students' socioeconomic status and school achievement on these preferences, analyzing whether perceptions of meritocracy moderate these effects.
+The findings provide insights into the complex interplay between meritocratic beliefs, socioeconomic factors, and attitudes towards social justice.
+
+Hypothesis $H_{1a}$ posited that students who perceive greater meritocracy in society would show stronger preferences for market justice.
+The results support this hypothesis, indicating a positive relationship between the perception that talent and effort are rewarded in society and the preferences for market justice in access to services such as health, education, and pensions.
+This finding aligns with previous research suggesting that individuals who believe in a meritocratic system are more likely to justify social inequalities [@batruch_belief_2022; @wiederkehr_belief_2015].
+It is noteworthy that among the items measuring meritocratic perception, consistently the one about deservingness shows a larger effect when compared to the ones about talent and effort.
+Although the three items correlate, there is something distinctive about deservingness that probably encompasses other aspects beyond meritocracy, an aspect to which we will come back in the conclusions.
+Moving now to hypothesis $H_{1b}$, this is simply a variation of 1a, replacing the perception of meritocracy at society by the one at school.
+In this case we have only two items of meritocratic perception - unfortunately deservingness was not measured at school level - and the direction of the effects is not as straightforward as in the case of societal meritocracy in $H_{1a}$: positive for talent, but negative for effort.
+Still, one of the limitation for the comparison between perception of meritocracy and society is the difference of the items' phrasing.
+Whereas for society the item refers to the extent to to which efforts are rewarded, at school is about the importance of effort, and something that is important does not necessarily mean that it is enough to be rewarded.
+Furthermore, the school-effort item is the one with the lowest variation among all meritocratic items (more than 95% agree or strongly agree), and with the lowest effect size.
+
+Hypothesis $H_2$ proposed that students from higher socioeconomic status families would show stronger preferences for market justice.
+However, the results do not support this hypothesis since they indicate that this relationship is not as straightforward as expected.
+Although students from families with higher educational levels and more economic resources tend to prefer less market justice, this relationship is not statistically significant.
+This might be due to the influence of other contextual and cultural factors that modulate attitudes toward market justice [@kohn_social_1963; @kohn_class_1969; @almas_fairness_2017].
+Notably, socioeconomic status measured by parental educational level and the number of books at home do not explain variations in market justice preferences, suggesting that other variables, such as market justice preferences of parents, could play an essential role in the attitudinal socialization process.
+This aligns with the notion that beliefs and attitudes toward social justice are complex and influenced by multiple factors beyond just economic resources.
+
+Hypothesis $H_3$ held that students from schools with higher socioeconomic status would show stronger preferences for market justice.
+The results did not confirm this hypothesis since although students from high-status schools tend to prefer more market justice, this result is not statistically significant.
+This finding, contrary to the existing literature [@jonsson_institutional_2015; @jost_attitudinal_2000], does not allow us to affirm the hypothesis about the role of the school environment, particularly socioeconomic segregation, in influencing attitudes towards market justice and the acceptance of inequalities.
+One possible explanation is the role of school-level variables not fully captured in our models, such as the school's climate or peer influence.
+Additionally, the non-significant results may suggest that the effect of school socioeconomic status on market justice preferences is correlated to other factors, such as individual meritocratic beliefs.
+
+Hypothesis $H_x$ [Falta hipótesis school achievement] held that students from schools with higher school achievement would show stronger preferences for market justice.
+Contrary to that hypothesis, the results show that students from schools with greater achievement in the SIMCE test tend to prefer less market justice.
+This unexpected finding could be due to the nuanced relationship between academic success and social attitudes.
+Students in high-achieving schools may be more exposed to critical discussions about inequality and social justice, leading to a rejection of market-based justifications for access to social services.
+
+Hypotheses $H_6$ and $H_7$ proposed that the perception of meritocracy would moderate the effect of socioeconomic status (both family and school) on market justice preferences.
+Analyzing hypothesis $H_6$ on family socioeconomic status, the results show a negative interaction between meritocratic beliefs and socioeconomic status.
+However, only the social meritocracy variables have a significant effect.
+This means that the positive impact of social meritocracy variables on market justice preferences decreases in students with parents at a university or postgraduate level.
+On the other hand, the results of hypothesis $H_7$ follow a similar line: the perception of meritocracy, especially social meritocracy, moderates the market justice preferences and their association with socioeconomic level and school performance.
+The negative interaction in this context means that in schools with low levels of performance and low socioeconomic levels, the positive effect of the social meritocracy variables on market justice preferences is stronger, fulfilling what was stated in hypothesis $H_7$.
+These relations indicate that, although those who perceive society as meritocratic have greater preferences towards market justice, this influence is lower among those with a higher socioeconomic level and who attend schools with higher academic performance.
+This finding highlights the complex interaction between socioeconomic status, school context, meritocratic beliefs, and preferences toward market justice.
+The moderating role of meritocratic perceptions suggests that students' views on social justice are shaped by their socioeconomic background and beliefs about how society rewards effort and talent.
+This underlines the importance of addressing structural inequalities and individual beliefs in educational and social policies.
+
+# Conclusions
+
 Our research aimed to examine the relationship between the perception of meritocracy and the justification of market justice preferences among eighth-grade students in Chile.
@@
 While we found support for some associations, others did not align with our initial expectations.
-Hypothesis $H_{1a}$ posited that students who perceive greater meritocracy in society would show stronger preferences for market justice.
-The results support this hypothesis, indicating a positive relationship between the perception that talent and effort are rewarded in society and the preferences for market justice in access to services such as health, education, and pensions.
-This finding aligns with previous research suggesting that individuals who believe in a meritocratic system are more likely to justify social inequalities [@batruch_belief_2022; @wiederkehr_belief_2015].
-Hypothesis H1b asserted that students who perceive greater meritocracy in school would show stronger preferences for market justice.
-The results show a mixed effect.
-The perception that talent is important for achieving good grades in school is associated with a greater preference for market justice in access to social services.
-Still, the perception that effort is important is associated with less preference for market justice.
-This highlights the influence of the school environment in shaping attitudes toward market justice, suggesting that meritocratic experiences in school can translate to broader societal views [@resh_sense_2010; @resh_sense_2014].
-
-Hypothesis H2 proposed that students from higher socioeconomic status families would show stronger preferences for market justice.
-However, the results indicate that this relationship is not as straightforward as expected.
-Although students from families with higher educational levels and more economic resources tend to prefer less market justice, this relationship is not statistically significant.
-This might be due to the influence of other contextual and cultural factors that modulate attitudes toward market justice [@kohn_social_1963; @kohn_class_1969; @almas_fairness_2017].
-
-Hypothesis H3 held that students from schools with higher socioeconomic status would show stronger preferences for market justice.
-The results did not confirm this hypothesis since although students from high-status schools tend to prefer more market justice, this result is not statistically significant.
-This finding, contrary to the existing literature [@jonsson_institutional_2015; @jost_attitudinal_2000], does not allow us to affirm the role of the school environment, particularly socioeconomic segregation, in influencing attitudes towards market justice and the acceptance of inequalities.
-
-Hypothesis Hx held that students from schools with higher school achievement would show stronger preferences for market justice.
-Contrary to that hypothesis, the results show that students from schools with greater achievement in the SIMCE test tend to prefer less market justice.
-
-Hypotheses H6 and H7 proposed that the perception of meritocracy would moderate the effect of socioeconomic status (both family and school) on market justice preferences.
-The results show that the perception of meritocracy, especially social meritocracy, amplifies the market justice preferences in schools with low levels of achievement and in families with parents with a lower educational level.
-This means that although those who perceive society as meritocratic have greater preferences towards market justice, this influence is less among those who have a higher socioeconomic status and attend schools with higher academic achievement.
-This finding highlights the complex interplay between socioeconomic status, school context, meritocratic beliefs, and preferences towards market justice.
-
-# Conclusions
 

### 2024-06-14 05-conclusiones.qmd
correcciones

@@
 Still, one of the limitation for the comparison between perception of meritocracy and society is the difference of the items' phrasing.
-Whereas for society the item refers to the extent to to which efforts are rewarded, at school is about the importance of effort, and something that is important does not necessarily mean that it is enough to be rewarded.
+Whereas for society the item refers to the extent to to which efforts are rewarded, at school is about the importance of effort, whereby something that is important does not necessarily mean that it is enough to be rewarded.
 Furthermore, the school-effort item is the one with the lowest variation among all meritocratic items (more than 95% agree or strongly agree), and with the lowest effect size.
 
-Hypothesis $H_2$ proposed that students from higher socioeconomic status families would show stronger preferences for market justice.
-However, the results do not support this hypothesis since they indicate that this relationship is not as straightforward as expected.
+Hypothesis $H_2$ proposed that students from higher socioeconomic status families would show stronger preferences for market justice, which is not supported by the results.
 Although students from families with higher educational levels and more economic resources tend to prefer less market justice, this relationship is not statistically significant.
@@
 Notably, socioeconomic status measured by parental educational level and the number of books at home do not explain variations in market justice preferences, suggesting that other variables, such as market justice preferences of parents, could play an essential role in the attitudinal socialization process.
-This aligns with the notion that beliefs and attitudes toward social justice are complex and influenced by multiple factors beyond just economic resources.
 
@@
 
-Hypothesis $H_x$ [Falta hipótesis school achievement] held that students from schools with higher school achievement would show stronger preferences for market justice.
-Contrary to that hypothesis, the results show that students from schools with greater achievement in the SIMCE test tend to prefer less market justice.
-This unexpected finding could be due to the nuanced relationship between academic success and social attitudes.
-Students in high-achieving schools may be more exposed to critical discussions about inequality and social justice, leading to a rejection of market-based justifications for access to social services.
+Hypothesis $H_x$ [Falta hipótesis school achievement] held that students from schools with higher school achievement would show stronger preferences for market justice, following the argument that such school could foster meritocratic attitudes by showing a larger promotion of talent and effort in order to get better results.
+Nevertheless, the results show quite the opposite: schools with better academic results (as measured by a national standardized test) show in average less preferences for market justice.
+This finding is one of the most strongest and consistent throughout the models and it is relevante given the centrality of school performance in educational institutions.
+One posible explanation of this link is that students in high-achieving schools may be more exposed to discussions and to a deeper understanding about inequality and social justice, leading to a critical perspective towards the weight of market mechanism in the access to social services.
 
-Hypotheses $H_6$ and $H_7$ proposed that the perception of meritocracy would moderate the effect of socioeconomic status (both family and school) on market justice preferences.
-Analyzing hypothesis $H_6$ on family socioeconomic status, the results show a negative interaction between meritocratic beliefs and socioeconomic status.
-However, only the social meritocracy variables have a significant effect.
-This means that the positive impact of social meritocracy variables on market justice preferences decreases in students with parents at a university or postgraduate level.
-On the other hand, the results of hypothesis $H_7$ follow a similar line: the perception of meritocracy, especially social meritocracy, moderates the market justice preferences and their association with socioeconomic level and school performance.
-The negative interaction in this context means that in schools with low levels of performance and low socioeconomic levels, the positive effect of the social meritocracy variables on market justice preferences is stronger, fulfilling what was stated in hypothesis $H_7$.
-These relations indicate that, although those who perceive society as meritocratic have greater preferences towards market justice, this influence is lower among those with a higher socioeconomic level and who attend schools with higher academic performance.
+Hypotheses $H_6$ and $H_7$ proposed that family status would moderate the effect of meritocratic perceptions on market justice preferences.
+Hypothesis $H_6$ refers to family socioeconomic status, whereby the results show a negative interaction with two of the three perceptual variables at the society level.
+This means that the positive impact of the perception of social meritocracy on market justice preferences have a lower effect in students with parents with higher education and also in those with more books at home.
+Something similar occurs with hypothesis $H_7$, where now school level status moderates the effect of meritocracy variables, which in multilevel framework si referred as to cross-level interaction.
+The negative interaction in this context means that in schools of low socioeconomic levels, the positive effect of the social meritocracy variables on market justice preferences is stronger.
+These relations indicate that, although those who perceive society as meritocratic have greater preferences towards market justice, this influence is lower among those with a higher socioeconomic level.
 This finding highlights the complex interaction between socioeconomic status, school context, meritocratic beliefs, and preferences toward market justice.
@@
 
+- Performance: ver efecto SIMCE individual
+
 -   Efecto ilustrador de la educación? (esfuerzo -\> redistribución)

### 2024-06-14 01-introduction.qmd
revisión párrafos chile

@@
 In Chile, the poorest 50% captures only 10% of the total income and has negative wealth, while the richest 1% receives almost 27% of the income and holds 49.6% of the wealth [@chancel_world_2022].
-Much of this inequality is linked to the deep neoliberal reforms that institutionalized privatization and commodification of various economic sectors, including public services, introduced during the dictatorship (1973-1989) and expanded in democracy through concessions, demand credits, and specific regulatory frameworks [@ffrench-davis_reformas_2018].
+Much of this inequality has been attributed to the deep neoliberal reforms that institutionalized privatization and commodification of various economic sectors.
+Such reforms where introduced during the dictatorship (1973-1989) and expanded in democracy through concessions, demand credits, and specific regulatory frameworks [@ffrench-davis_reformas_2018].
 This shift in economic policy has allowed the unprecedented emergence of markets in areas related to social reproduction, such as health, pensions, and education, with provision and access managed by private entities and segmented by individual payment capacity, heavily reliant on State subsidies [@boccardo_30_2020].
-In health, although the majority of the population uses the public insurance system (78.9%), 15.3% are served by private insurers (CASEN, 2022).
-The pension system is based on individual capitalization, with mandatory contributions managed by private administrators investing in the financial market, currently involving 11 million contributors (Superintendencia de Pensiones, 2024).
-In education, 30.6% of school enrollment is in public schools, 54.0% in State-subsidized private schools, and 9.3% in fully private schools, generally attracting higher-income groups (Mineduc, 2023).
+In health, although the majority of the population uses the public insurance system (78.9%), the pension system is based on individual capitalization, with mandatory contributions managed by private administrators investing in the financial market, currently involving 11 million contributors (Superintendencia de Pensiones, 2024).
+In education, 30.6% of school enrollment is in public schools, 54.0% in state-subsidized (voucher) private schools, and 9.3% in fully private schools, generally attracting higher-income groups (Mineduc, 2023).
 
-The provision and access to services in these areas under market principles correlate with the perceptions, preferences, and attitudes of Chilean society.
-Recent public opinion surveys conducted by the Center for Public Studies indicate that, generally, people prefer public insurance (64.1%) over private insurance (35.9%) if both cost the same.
-In pensions, Chileans show varied preferences regarding the public or private status of pension fund administrators; 52.3% favor both private and public institutions, 38.8% prefer only public institutions, and 8.6% prefer only private ones.
+Some surveys in Chile have considered the perception and support for the market provision in public areas.
+Recent public opinion surveys conducted by the Center for Public Studies (CEP) indicate that, generally, people prefer public insurance (64.1%) over private insurance (35.9%) if both cost the same.
+In pensions, Chileans show varied preferences regarding the public or private status of pension fund administrators; 52.3% favor both private and public institutions, 38.8% prefer only public institutions, and only 8.6% prefer only private ones.
 This likely relates to the severe legitimacy crisis faced by private pension fund administrators, which has sparked social movements in recent years [@boccardo_30_2020].

### 2024-06-18 05-conclusiones.qmd
cambios NSE familiar y escolar

@@
 
-Hypothesis $H_2$ proposed that students from higher socioeconomic status families would show stronger preferences for market justice, which is not supported by the results.
+Hypothesis $H_2$ and hypothesis $H_3$ propose that socioeconomic status could influence students' preferences, but the results show different patterns.
+On the one hand, $H_2$ proposed that students from higher socioeconomic status families would show stronger preferences for market justice, which is not supported by the results.
 Although students from families with higher educational levels and more economic resources tend to prefer less market justice, this relationship is not statistically significant.
@@
 Notably, socioeconomic status measured by parental educational level and the number of books at home do not explain variations in market justice preferences, suggesting that other variables, such as market justice preferences of parents, could play an essential role in the attitudinal socialization process.
-
-Hypothesis $H_3$ held that students from schools with higher socioeconomic status would show stronger preferences for market justice.
-The results did not confirm this hypothesis since although students from high-status schools tend to prefer more market justice, this result is not statistically significant.
-This finding, contrary to the existing literature [@jonsson_institutional_2015; @jost_attitudinal_2000], does not allow us to affirm the hypothesis about the role of the school environment, particularly socioeconomic segregation, in influencing attitudes towards market justice and the acceptance of inequalities.
-One possible explanation is the role of school-level variables not fully captured in our models, such as the school's climate or peer influence.
-Additionally, the non-significant results may suggest that the effect of school socioeconomic status on market justice preferences is correlated to other factors, such as individual meritocratic beliefs.
+On the other hand, Hypothesis $H_3$ held that students from schools with higher socioeconomic status would show stronger preferences for market justice.
+The results confirm this hypothesis since students from high-status schools tend to prefer more market justice, compared to low-status schools.
+This finding, in line with the existing literature [@jonsson_institutional_2015; @jost_attitudinal_2000], does allow us to affirm the hypothesis about the role of the school environment, particularly socioeconomic segregation, in influencing attitudes towards market justice and the acceptance of inequalities.
+A possible explanation is the role of the context at the school level, which could have a more relevant role than the family socioeconomic status, since the highly stratified nature of the Chilean education system reinforces socioeconomic divisions and socializes students into different normative frameworks.
 
@@
 Nevertheless, the results show quite the opposite: schools with better academic results (as measured by a national standardized test) show in average less preferences for market justice.
-This finding is one of the most strongest and consistent throughout the models and it is relevante given the centrality of school performance in educational institutions.
-One posible explanation of this link is that students in high-achieving schools may be more exposed to discussions and to a deeper understanding about inequality and social justice, leading to a critical perspective towards the weight of market mechanism in the access to social services.
+This finding is one of the most strongest and consistent throughout the models and it is relevant given the centrality of school performance in educational institutions.
+One possible explanation of this link is that students in high-achieving schools may be more exposed to discussions and to a deeper understanding about inequality and social justice, leading to a critical perspective towards the weight of market mechanism in the access to social services.
 

### 2024-06-18 07-anexos.qmd
conclusiones

--- 2024-04-23 06-anexos.qmd
@@
 ---
-title: "06-anexos"
 author: "Equipo EDUMER"
@@
 ---
+
+## Appendix {#appendix}
 
@@
 |                                                                                                                    |                                                                                                                                              |
-|--------------------------------------------------------------------------------------------------------------------|----------------------------------------------------------------------------------------------------------------------------------------------|
+|--------------------------------|---------------------------------------|
 | **English translation**                                                                                            | **Original Spanish**                                                                                                                         |

### 2024-06-18 02-antecedentes.qmd

@@
 These stereotypes help individuals maintain a sense of superiority over others and legitimize the prevailing social hierarchies and economic disparities [@jost_attitudinal_2000].
-The link between school socioeconomic status and market preferences is certainly reinforced in societies with high school segregation, as is the case of Chile.
-Consistently, we expect that
+In this line, and besides status, we also expect that schools that exhibit an average better performance scores in standardized large scale assessments would also show larger market justice preferences.
+The correspondent hypotheses are:
 
@@
 
-The last set of hypotheses deal with the role of meritocratic perception as a moderator of family and school status on market justice preferences.
-These hypotheses have more an exploratory character, as there are not clear references in the literature to this regard.
-The general argument deals with the role of socioeconomic status and meritocratic perceptions, which is contested in the literature.
-On the one hand, and based on a mere rational argument, it would be expected that those who succeed in terms of educational rewards and carreer paths - usuaylly those of higher status - perceive more meritocracy.
-Nevertheless, evidence regarding the legitimacy of meritocracy relies on consensual beliefs from the side of both winners and losers from the distributive system, pointing in the direction of fewer status differences when it comes to meritocratic beliefs [@wiederkehr_belief_2015].
-As this is the first study of this kind in the Chilean context, charaterized by extreme educational inequalities, we rely on a traditional and rational approach, proposing that:
+H4: Students from schools with an average better performance in standardized achivement tests will show larger market justice preferences.
 
-H4: Students from schools of higher social status will perceive more meritocracy.
+The last set of hypotheses deal with the moderation effects.
+We propose that the link between meritocracy and market justice could be stronger for those of high status and attendig to school with larger achievement scores.
+In this regard, the effecs of status and merit would not be independent, as it would be expected that those who succeed in terms of educational rewards and carreer paths - usuaylly those of higher status - perceive more meritocracy, which would reflect on a larger market justice prefences.
+In similar terms, schools of better performance would promote meritocratic perception, sthrengthening market justice prefences:
 
-H5: Students from families of higher social status will perceive more meritocracy.
+H5: The family social status will moderate the effect of meritocratic perception on market justice preferences.
 
-\[aca falta hipothesis de school performance\]
-
-The previous hypotheses serve as a sort of bridge for the final moderation hypotheses.
-As higher (family and school) status would be related to meritocratic perception (H5 and H6), and at the same time meritocratic perception would be related to larger market justice preferences (H1 and H2), we argue that meritocratic perceptions could play a moderating role: the association between high status and market justice would be stronger for those with a larger meritocratic perception:
-
-H6: The perception of meritocracy in school and society will moderate the effect of family social status on market justice preferences.
-
-H7: The perception of meritocracy in school and society will moderate the effect of school status on market justice preferences.
+H5: The school performance in standardized achievement tests will moderate the effect of meritocratic perceptions on market justice preferences.
 

### 2024-06-18 04-analisis.qmd

--- 2024-06-13 04-analisis.qmd
@@
 ```{r echo=FALSE, include=FALSE}
+sjmisc::frq(data_rec$educacion_rec)
+data_rec <- data_rec %>% mutate(universitario = case_when(educacion_rec=="Ed Basica"~"No",
+                                                          educacion_rec=="Ed Media"~"No",
+                                                          educacion_rec=="Ed Tecnica"~"No",
+                                                          educacion_rec=="Universidad o posgrado"~"Si",
+                                                          educacion_rec=="Ns/Nr"~"Ns/Nr"))
+data_rec$universitario <- factor(data_rec$universitario, levels = c("No", "Si", "Ns/Nr"))
+sjmisc::frq(data_rec$universitario)
+
+data_rec <- data_rec %>% mutate(school_nse = case_when(cod_grupo=="Bajo"~"Bajo",
+                                                          cod_grupo=="Medio bajo"~"Medio",
+                                                          cod_grupo=="Medio"~"Medio",
+                                                          cod_grupo=="Medio alto"~"Medio",
+                                                          cod_grupo=="Alto"~"Alto"))
+sjmisc::frq(data_rec$school_nse)
+data_rec$school_nse <- factor(data_rec$school_nse, levels = c("Bajo", "Medio", "Alto"))
 
@@
 m2_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
     libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_pension)
@@
 m3_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
     libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
-    cod_grupo + simce + (1 | mrbd), data=data_pension)
+    school_nse + simce + (1 | mrbd), data=data_pension)
 
@@
 m2_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
     libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_salud)
@@
 m3_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
     libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
-    cod_grupo + simce + (1 | mrbd), data=data_salud)
+    school_nse + simce + (1 | mrbd), data=data_salud)
 
@@
 m2_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
     libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_educ)
@@
 m3_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
     libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
-    cod_grupo + simce + (1 | mrbd), data=data_educ)
+    school_nse + simce + (1 | mrbd), data=data_educ)
 
@@
   merito_soc = "Deservingness",
-  "educacion_recEd Media" = "Secondary",
-  "educacion_recEd Tecnica" = "Higher tec.",
-  "educacion_recUniversidad o posgrado" = "University or posgraduate",
-  "educacion_recNs/Nr" = "Missing",
+  "universitarioSi" = "University or posgraduate",
+  "universitarioNs/Nr" = "Missing",
   "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
-  "cod_grupoMedio bajo" = "SES Medium low",
-  cod_grupoMedio = "SES Medium",
-  "cod_grupoMedio alto" = "SES Medium high",
-  cod_grupoAlto = "SES High",
+  "school_nseMedio" = "SES Medium",
+  school_nseAlto = "SES High",
   simceMedio = "Simce Medium",
@@
                 scalebox = 0.60,
-                groups = list("Parental education (Ref.= 8th grade or less)" = 6:9,
-                                "Socioeconomic level (Ref.= Low)" = 11:14,
-                                "Achievement score (Ref.= Low)" = 15:16),
+                groups = list("Parental education (Ref.= Non-university)" = 6:7,
+                                "Socioeconomic level (Ref.= Low)" = 9:10,
+                                "Achievement score (Ref.= Low)" = 11:12),
                 custom.gof.rows = list("Controls"=rep("Yes",6)),
@@
 Model 2 in @tbl-ordinal-reg incorporates school-level variables.
-The results show that high socioeconomic level schools show greater preferences for market justice in education compared to low socioeconomic level schools, with statistically significant results (*p* \< 0.05).
-This effect is not seen for health and pensions.
+The results show that high socioeconomic level schools show greater preferences for market justice in health, pensions, and education compared to low socioeconomic level schools, with statistically significant results (*p* \< 0.05).
 Moreover, the findings suggest that schools with higher academic performance shows negative and statistically significant effects across all three items, suggesting these schools justify market allocation principles less.
-This pattern is also observed in medium-performing schools, but only for health and pensions.
 
 ```{r echo=FALSE, include=FALSE}
+data_bienestar <- data_bienestar %>% mutate(universitario = case_when(educacion_rec=="Ed Basica"~"No",
+                                                          educacion_rec=="Ed Media"~"No",
+                                                          educacion_rec=="Ed Tecnica"~"No",
+                                                          educacion_rec=="Universidad o posgrado"~"Si",
+                                                          educacion_rec=="Ns/Nr"~"Ns/Nr"))
+data_bienestar$universitario <- factor(data_bienestar$universitario, levels = c("No", "Si", "Ns/Nr"))
+
+sjmisc::frq(data_bienestar$cod_grupo)
+data_bienestar <- data_bienestar %>% mutate(school_nse = case_when(cod_grupo=="Bajo"~"Bajo",
+                                                          cod_grupo=="Medio bajo"~"Medio",
+                                                          cod_grupo=="Medio"~"Medio",
+                                                          cod_grupo=="Medio alto"~"Medio",
+                                                          cod_grupo=="Alto"~"Alto"))
+sjmisc::frq(data_bienestar$school_nse)
+data_bienestar$school_nse <- factor(data_bienestar$school_nse, levels = c("Bajo", "Medio", "Alto"))
 
@@
                          inteligencia_soc + esfuerzo_soc + merito_soc + 
-                         educacion_rec + libros_rec + acc_tec + internet_rec +
+                         universitario + libros_rec + acc_tec + internet_rec +
                          (1 | mrbd), data=data_bienestar)
 
-reg3_bienestar <- lmer(bienestar ~ 1 + mean_educ + cod_depe2 + cod_grupo +
+reg3_bienestar <- lmer(bienestar ~ 1 + mean_educ + cod_depe2 + school_nse +
                          simce + (1 | mrbd), data=data_bienestar)
@@
                          inteligencia_soc + esfuerzo_esc + inteligencia_esc + 
-                         educacion_rec + libros_rec + acc_tec + internet_rec + 
-                         mean_educ + cod_depe2 + cod_grupo + 
+                         universitario + libros_rec + acc_tec + internet_rec + 
+                         mean_educ + cod_depe2 + school_nse + 
                          (1 | mrbd), data=data_bienestar)
@@
                          inteligencia_soc + esfuerzo_esc + inteligencia_esc + 
-                         educacion_rec + libros_rec + acc_tec + internet_rec + 
-                         mean_educ + cod_depe2 + cod_grupo + simce + 
+                         universitario + libros_rec + acc_tec + internet_rec + 
+                         mean_educ + cod_depe2 + school_nse + simce + 
                          (1 | mrbd), data=data_bienestar)
+
+knitreg(list(reg1_bienestar, reg2_bienestar, reg3_bienestar, reg4_bienestar, reg5_bienestar))
 
@@
                 scalebox = 0.60,
-                groups = list("Parental education (Ref.= 8th grade or less)" = 7:10,
-                              "Socioeconomic level (Ref.= Low)" = 12:15,
-                              "Achievement score (Ref.= Low)" = 16:17),
+                groups = list("Parental education (Ref.= Non-university)" = 7:8,
+                              "Socioeconomic level (Ref.= Low)" = 10:11,
+                              "Achievement score (Ref.= Low)" = 12:13),
                 custom.gof.rows = list("Controls"=rep("Yes",4),
@@
 School meritocracy perceptions yield mixed effects: perceived talent increases market justice preference, while perceived effort decreases it, both statistically significant (*p* \< 0.001).
-Model 2 includes family status, finding no statistically significant effects from parents' education or the number of books at home, both showing negative directions.
+Model 2 includes family status, finding no statistically significant effects from parents' education or the number of books at home, showing positive direction for university or posgraduate, and negative direction for books at home.
 
 Regarding school characteristics, Models 3 and 4 introduce socioeconomic level and academic performance.
-The results of Model 3 suggest no statistically significant relationship between school socioeconomic levels and market justice preferences.
-However, Model 4 indicates that schools with medium ($\beta$ = -0.10) and high ($\beta$ = -0.24) academic performance have lower market justice preferences compared to low-performing schools, with these effects being statistically significant when controlling for other variables.
+The results of Model 3 suggest statistically significant relationship between school socioeconomic levels and market justice preferences, specifically those school with high socieconomic level ($\beta$ = 0.37) compared to low socioeconomic level.
+However, Model 4 indicates that schools with medium ($\beta$ = -0.11) and high ($\beta$ = -0.26) academic performance have lower market justice preferences compared to low-performing schools, with these effects being statistically significant when controlling for other variables.
 

### 2024-06-18 05-discussion.qmd

--- 2024-06-18 05-conclusiones.qmd
@@
 This underlines the importance of addressing structural inequalities and individual beliefs in educational and social policies.
-
-# Conclusions
-
-Our research aimed to examine the relationship between the perception of meritocracy and the justification of market justice preferences among eighth-grade students in Chile.
-Overall, our findings provided mixed evidence regarding our hypotheses.
-While we found support for some associations, others did not align with our initial expectations.
-
-Future research could leverage the capabilities of longitudinal databases to explore contextual level.
-For example, statistical models could incorporate contextual information from administrative data sources, allowing for testing hypotheses that include contextual socioeconomic inequality and its dynamics over time.
-Additionally, this dataset could be used to compare inequality justification across different policy areas, such as health and pensions, and examine how meritocratic perceptions and beliefs affect these areas differently over time.
-Such studies would provide valuable insights for understanding societal and cultural changes and their impacts on solidarity and social cohesion.
-
-- Performance: ver efecto SIMCE individual
-
--   Efecto ilustrador de la educación? (esfuerzo -\> redistribución)

### 2024-06-19 01-introduction.qmd
review JC

@@
 It is often seen as a way to create equal opportunities and fairness, as individuals can rise to positions of power and influence based on their own merit rather than their background or connections.
-However, some argue that meritocracy can actually lead to tolerating or even justifying social inequalities, as it can create a hierarchy where those who already have resources and advantages are more likely to succeed.
+However, some argue that meritocracy can actually lead to tolerating or even justifying social inequalities, as it can create a hierarchy where those who already have resources and advantages are more likely to succeed [@sandel_tyranny_2020;@mcnamee_meritocracy_2004].
 In this regard, a great deal of academic research about meritocracy delves into the assessment of to what extent rewards and privileges in society are related to merit, emphasizing the so-called unfulfillable promise of meritocracy [@mijs_stratified_2016].
@@
 In education, 30.6% of school enrollment is in public schools, 54.0% in state-subsidized (voucher) private schools, and 9.3% in fully private schools, generally attracting higher-income groups (Mineduc, 2023).
-
-Some surveys in Chile have considered the perception and support for the market provision in public areas.
-Recent public opinion surveys conducted by the Center for Public Studies (CEP) indicate that, generally, people prefer public insurance (64.1%) over private insurance (35.9%) if both cost the same.
-In pensions, Chileans show varied preferences regarding the public or private status of pension fund administrators; 52.3% favor both private and public institutions, 38.8% prefer only public institutions, and only 8.6% prefer only private ones.
-This likely relates to the severe legitimacy crisis faced by private pension fund administrators, which has sparked social movements in recent years [@boccardo_30_2020].
-Interestingly, in education, there is a clear preference for private schools over public ones if both cost the same and are equally distant, with 63% preferring private and 37% public.

### 2024-06-19 02-antecedentes.qmd
review JC

@@
 
-The justification of social inequality based on marke-type criteria has been conceptualized as the individuals' adherence to the deservingness of social goods and services (sucha as health, education and pensions) based on prices and individuals' ability to pay [@boltanski_new_2005; @lane_market_1986; @streeck_citizens_2012].
+The justification of social inequality based on market-type criteria has been conceptualized as the individuals' adherence to the deservingness of social goods and services (such as health, education, and pensions) based on prices and individuals' ability to pay [@boltanski_new_2005; @lane_market_1986; @streeck_citizens_2012].
 Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegel_beliefs_1987], highlights that people's explanations and justifications of social inequality are closely tied to their judgments of deservingness.
@@
 
+There are substantial differences in funding and delivery methods in the management of social services across nations [@jensen_worlds_2008; @stoy_worlds_2014].
+Nordic countries, for example, predominantly employ public agencies to produce and provide social services, funding these through collective taxation and offering them in kind to the majority of citizens.
+This system prioritizes social justice, placing it above market mechanisms in accessing services.
+In contrast, other countries rely more heavily on for-profit entities and private funding, where service distribution depends mainly on individual financial capacity to pay user fees, highlighting the influence of market justice in service allocation.
+The trend toward marketization of welfare services has been growing since the 1980s [@salamon_marketization_1993], and this shift is increasingly evident even in countries where market solutions have traditionally had a minor role in social policy [@sivesind_changing_2017].
+The expansion of marketization has been related to a larger justification of market mechanisms, whereby societies with larger private spending on services show larger market justice preferences [@lindh_public_2015].
+
 Robert E.
@@
 384].
-Contrary to the evidence that unequal distribution produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action [@greitemeyer_subjective_2016; @mishra_subjective_2015; @smith_relative_2012; @power_deprivationprotest_2018; @power_relative_2020], Lane points out that in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions.
-In this sense, unequal levels of well-being ought to be, to some extent, a function of their talents and efforts, instead of being based in distributive principles that characterize welfarism, such as need and equality (see Wilson 2003).
+Contrary to the evidence that unequal distribution produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action [@greitemeyer_subjective_2016; @mishra_subjective_2015; @smith_relative_2012; @power_deprivationprotest_2018; @power_relative_2020], Lane pointed out that in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions.
+In this sense, unequal levels of well-being would be, to some extent, a function of their talents and efforts, instead of being based on distributive principles that characterize welfare states, such as need and equality (see Wilson 2003).
 
-The management of social services manifests in varying approaches across nations, with substantial differences in funding and delivery methods [@jensen_worlds_2008; @stoy_worlds_2014].
-Nordic countries, for example, predominantly employ public agencies to produce and provide social services, funding these through collective taxation and offering them in kind to the majority of citizens.
-This system prioritizes political justice, placing it above market mechanisms in accessing services.
-In contrast, other countries rely more heavily on for-profit entities and private funding, where service distribution depends mainly on individual financial capacity of paying user fees, highlighting the influence of market justice in service allocation.
-The trend toward marketization of welfare services has been growing since the 1980s [@salamon_marketization_1993], and this shift is increasingly evident even in countries where market solutions have traditionally had a minor role in social policy [@sivesind_changing_2017].
-The expansion of marketization has been related to larger justification of market mechanisms, whereby societies with larger private spending on services show larger market justice preferences [@lindh_public_2015].
-
-Despite high-income inequality and limited social mobility in Chile and in Latin America in general, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@bucca_merit_2016; @chong_mystery_2008; @torche_intergenerational_2014; @salgado_inequality_2023].
-The reliance on private welfare providers and widespread user fees [@molyneux_neoliberal_2008] adds complexity to this context.
-Yet, research on children's justification of market-based inequalities in accessing welfare services remains limited, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.
-
-```{r echo=FALSE, message=FALSE, error=FALSE}
-#| fig-cap: "Market justice preferences in health, pensions and education in the adult population"
-#| fig-asp: 0.6
-#| label: fig-elsoc
-
-elsoc_2016 <- elsoc_2016 %>% 
-  dplyr::select(Pensions = d02_01, 
-                Education = d02_02, 
-                Health = d02_03) %>% 
-  mutate_all(.funs = ~ set_na(., na = c(-666,-777,-888,-999), drop.levels = F) %>% 
-               factor(.,
-                      levels = c(1,2,3,4,5),
-                      labels = c("Strongly desagree",
-                                 "Desagree",
-                                 "Neither agree nor disagree",
-                                 "Agree",
-                                 "Strongly agree"))) %>% 
-  na.omit()
-
-elsoc_2016 %>% 
-  pivot_longer(cols = everything(),
-               names_to = "justice", 
-               values_to = "value") %>% 
-  mutate(justice = factor(justice, 
-                          levels = c("Health", "Pensions", "Education"))) %>% 
-  group_by(justice, value) %>%
-  summarise(total = n()) %>% 
-  mutate(prop = prop.table(total)*100) %>% 
-  ggplot(aes(x = value, y = prop, fill = value)) +
-  geom_col(position = "dodge", width = .9, alpha = 1, show.legend = T) +
-  scale_y_continuous(labels = function(prop){paste0(prop, "%")},
-                     limits = c(0,55),
-                     n.breaks = 7,
-                     expand = c(0,1)) +
-  facet_wrap(.~justice, scales = "fixed") +
-  geom_text(mapping = aes(label = scales::percent(prop, accuracy = 0.1, scale = 1, decimal.mark = ".")),
-            vjust = -0.5,
-            size = 3,
-            position = position_dodge(width = 0.8)) +
-  scale_fill_manual(values = c("#F1EEF6", "#BDC9E1", "#b3b3b3ff", "#74A9CF", "#0570B0")) +
-  labs(x = NULL,
-       y = "Proportion (%)",
-       fill = NULL,
-       caption = paste0("Source: Authors calculation based on ELSOC 2016",
-                        " (N = ",dim(elsoc_2016)[1],")")) +
-  theme(legend.position = "bottom",
-        axis.text.x = element_text(size = 0))
-
-```
+Despite high-income inequality and limited social mobility in Chile, and in Latin America in general, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@bucca_merit_2016; @chong_mystery_2008; @torche_intergenerational_2014; @salgado_inequality_2023].
+The reliance on private welfare providers and widespread user fees [@molyneux_neoliberal_2008] adds complexity to this context, as reflected in surveys conducted by the Center for Public Studies (CEP).
+According to this data, 35.9% prefer private health insurance, and 63% would prefer private education.
+Yet, research on children's justification in this area remains limited, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.
 
@@
 
-The original definition of merit is a combination of effort and talent [@young_rise_1958], and a meritocracy is a distributive system where merit is the main criteria for allocating valuable goods and rewards.
+The original definition of merit is a combination of effort and talent [@young_rise_1958], and a meritocracy is a distributive system where merit is the main criterion for allocating valuable goods and rewards.
 From a sociological perspective, meritocracy has been used in research on social mobility to characterize societies with low mobility that threaten the meritocratic ideal [@goldthorpe_myth_2003].
 More recently, sociology and social psychology research has attended to subjective aspects related to the support for meritocratic principles in different societies, such as beliefs in meritocracy [@castillo_multidimensional_2023; @mijs_unfulfillable_2016; @mijs_paradox_2019].
-The label of beliefs in this realm covers different concepts, but in this paper we refer to a conceptual model developed by @castillo_multidimensional_2023 which identify two main aspects: preferences and perceptions for meritocracy.
 
-While meritocratic preferences refer to a justification of distribution based on merit criteria (effort and talent), the perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society [@duru-bellat_who_2012; @castillo_meritocracia_2019].
-This perception can vary greatly depending on individual experiences, social, economic, and cultural background.
+Meritocratic beliefs can cover two types of subjective processes: preferences and perceptions [@castillo_multidimensional_2023].
+While meritocratic preferences refer to a justification of distribution based on merit criteria (effort and talent), the perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their society [@duru-bellat_who_2012; @castillo_meritocracia_2019].
+The perception can vary greatly depending on individual experiences, as well as social, economic, and cultural background.
 Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work.
 In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2019].
-Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger market justice preferences, as individual achievement would be seen as rewarded and social policies as less necessary.
+Some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2019], and how larger inequality decreases it [@morris_representing_2022].
 
-Some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2019], and how larger inequality affects meritocratic beliefs [@morris_representing_2022].
-A larger justification of meritocratic distribution have been related to less support for redistributive compensation systems [@frank_performance_2015].
-Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed [@evans_strong_2018].
+A larger justification of meritocratic distribution has been related to less support for redistributive compensation systems [@frank_performance_2015], as individual achievement would be seen as rewarded and social policies as less necessary.
 @almas_cutthroat_2020 found that in the US the highly educated accept inequality significantly more than the less educated because they perceive inequality as justifiable owing to differences in productivity (i.e., merit).
 @barr_effect_2020 found that in unequal societies, the highly educated accept inequality more than the low educated.
-Using data for 41 countries from the International Social Survey Programme (ISSP),@garcia-sanchez_attitudes_2020 found that the perceived size of the income gap correlated positively with support for progressive taxation, an association that was weaker among those who endorsed meritocratic and equal opportunity beliefs.
-In the same line, experimental research by @durante_preferences_2009 showed that subjects support less redistribution when the initial distribution is determined according to task performance.
-
-## The socialization of market justice: families and schools
+Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed [@evans_strong_2018].
 
@@
 Additionally, academic achievement is treated as the outcome of dispositional factors (e.g., pupils' efforts and talents or lack of them) rather than the result of differential access to critical resources.
-Due to the meritocratic frame schools encourage, both low- and high-status individuals tend to believe that success or failure is not due to family background but rather to differences in efforts and talents [@darnon_where_2018].
-In this sense, we believe that the perception of meritocracy can influence students' judgments about market justice preferences.
+Due to the meritocratic frame schools encourage, both low- and high-status individuals tend to believe that success or failure is not due to the family background but rather to differences in efforts and talents [@darnon_where_2018].
+In this sense, we believe that the perception of meritocracy can influence students' judgments about market justice preferences, leading to our first hypothesis:
 
-H1a: Students who perceive that there is more meritocracy in society will show larger market justice preferences
+$H_{1a}$: Students who perceive that there is more meritocracy in society will show larger market justice preferences
 
-The perception of meritocracy has been mostly studied with general questions about rewards allocation based on effort and talent (ususally ingelligence).
-Nevertheless, when looking at school population it is possible to further consider the perception of meritocracy referred specifically to the school context.
-For instance, @resh_sense_2010 find that perception of justice in grades have a positive effect on liberal democratic orientation and on trust in people and in formal institutions [@resh_sense_2014].
+The perception of meritocracy has been mostly studied with general questions about reward allocation based on effort and talent (usually intelligence).
+Nevertheless, when looking at the school population it is possible to further consider the perception of meritocracy referred specifically to the school context.
+For instance, @resh_sense_2010 find that perception of justice in grades has a positive effect on liberal democratic orientation, and trust in people and formal institutions [@resh_sense_2014].
 Attending to this evidence, we differentiate in this study between meritocratic perception in society at large, and meritocratic perception at school, proposing the next hypothesis:
 
-H1b: Students who perceive that there is more meritocracy at school will show larger market justice preferences
+$H_{1b}$: Students who perceive that there is more meritocracy at school will show larger market justice preferences
+
+## The socialization of market justice: families and schools
 
@@
 In the same line, @almas_fairness_2017 found that adolescents from low-socioeconomic-status families are likelier to have an egalitarian fairness view and consider an equal distribution as fair in a situation with unequal merits.
-Taking this into account, we believe that there will be differences in the socialization of values according to socioeconomic differences in families that could influence market justice preferences, which lead us to the following hypothesis:
+Taking this into account, we believe that there will be differences in the socialization of values according to socioeconomic differences in families that could influence market justice preferences, which leads us to the following hypothesis:
 
-H2: Students from families of higher social status will show larger market justice preferences.
+$H_2$: Students from families of higher social status will show larger market justice preferences.
 
@@
 These stereotypes help individuals maintain a sense of superiority over others and legitimize the prevailing social hierarchies and economic disparities [@jost_attitudinal_2000].
-In this line, and besides status, we also expect that schools that exhibit an average better performance scores in standardized large scale assessments would also show larger market justice preferences.
+In this line, and besides status, we also expect that schools that exhibit average better performance scores in standardized large-scale assessments would also show larger market justice preferences.
 The correspondent hypotheses are:
 
-H3: Students from schools of higher social status will show larger market justice preferences.
+$H_3$: Students from schools of higher social status will show larger market justice preferences.
 
-H4: Students from schools with an average better performance in standardized achivement tests will show larger market justice preferences.
+$H_4$: Students from schools with an average better performance in standardized achievement tests will show larger market justice preferences.
 
-The last set of hypotheses deal with the moderation effects.
-We propose that the link between meritocracy and market justice could be stronger for those of high status and attendig to school with larger achievement scores.
-In this regard, the effecs of status and merit would not be independent, as it would be expected that those who succeed in terms of educational rewards and carreer paths - usuaylly those of higher status - perceive more meritocracy, which would reflect on a larger market justice prefences.
-In similar terms, schools of better performance would promote meritocratic perception, sthrengthening market justice prefences:
+The last set of hypotheses deals with the moderation effects.
+We propose that the link between meritocracy and market justice could be stronger for those of high status and attending schools with larger achievement scores.
+In this regard, the effects of status and merit would not be independent, as it would be expected that those who succeed in terms of educational rewards and career paths - usually those of higher status - perceive more meritocracy, which would reflect on larger market justice preferences.
+In similar terms, schools of better performance would promote meritocratic perception, strengthening market justice preferences:
 
-H5: The family social status will moderate the effect of meritocratic perception on market justice preferences.
+$H_5$: The family's social status will moderate the effect of meritocratic perception on market justice preferences.
 
-H5: The school performance in standardized achievement tests will moderate the effect of meritocratic perceptions on market justice preferences.
+$H_6$: The school's performance in standardized achievement tests will moderate the effect of meritocratic perceptions on market justice preferences.
 
-### Summary of hypotheses
+Figure @fig-hypotheses depicts a summary of the research hypothesis described before.
+Market justice preferences are our main concept under study (vis-a-vis dependent variable), which refers to the justification of better access to health, education, and pensions based on payment capacities (as we detail further in the methods section).
+The explanatory concepts (independent variables) are at two levels: student (individual) variables and school (context) variables.
+The moderation hypotheses are represented through arrows pointing to the arrow between meritocratic perception and market justice preferences, which will be estimated through interaction effects.
 

### 2024-06-21 01-introduction.qmd
add processing docs and fixed references

@@
 It is often seen as a way to create equal opportunities and fairness, as individuals can rise to positions of power and influence based on their own merit rather than their background or connections.
-However, some argue that meritocracy can actually lead to tolerating or even justifying social inequalities, as it can create a hierarchy where those who already have resources and advantages are more likely to succeed [@sandel_tyranny_2020;@mcnamee_meritocracy_2004].
+However, some argue that meritocracy can actually lead to tolerating or even justifying social inequalities, as it can create a hierarchy where those who already have resources and advantages are more likely to succeed [@sandel_tyranny_2020; @mcnamee_meritocracy_2004].
 In this regard, a great deal of academic research about meritocracy delves into the assessment of to what extent rewards and privileges in society are related to merit, emphasizing the so-called unfulfillable promise of meritocracy [@mijs_stratified_2016].
@@
 In the present paper, we address the role of the perception of meritocracy on the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system.
-In particular, we focus in justifying inequalities in health, education and pensions, which are traditional social policy areas where acces to better services could be justified based on payment capacity, referred as to market justice [@lane_market_1986, @lindh_bringing_2023].\
+In particular, we focus in justifying inequalities in health, education and pensions, which are traditional social policy areas where access to better services could be justified based on payment capacity, referred as to market justice [@lane_market_1986; @lindh_bringing_2023].
 Most of the research in this realm has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies [@batruch_belief_2022].

### 2024-06-21 02-antecedentes.qmd
add processing docs and fixed references

@@
 The influence of ideologies [@wegener_dominant_1995] and cultural schemas [@homan_being_2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations.
-While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999 ; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children's acceptance or rejection of these market principles.
+While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children's acceptance or rejection of these market principles.
 This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades [@centeno_arc_2012; @harvey_breve_2015], affecting areas such as pensions, health services, and education.
@@
 Lane proposed the underpinnings of the concept of market justice, which he differentiated from political justice.
-For him, "it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets" [-@lane_market_1986; p.
+For him, "it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets" [-@lane_market_1986 p.
 384].
@@
 Schools contribute to institutionalizing and reproducing inequality by promoting values, norms, practices, and languages familiar to higher-class families because the dominant group's culture shapes educational institutions [@bourdieu_reproduction_1990].
-Middle- and upper-class students are better equipped to face academic challenges and are more familiar with academic expectations [@mikus_children_2020].
+Middle- and upper-class students are better equipped to face academic challenges and are more familiar with academic expectations [@mikus_children_2019].
 Such familiarity represents cultural capital in educational contexts because higher-status students come to school ready to meet these expectations and reap the benefits [@jack_no_2016; @khan_privilege_2011].
@@
 Attending to the socialization within the family, the classic work of Kohn showed that middle-class parents value the expression of internal states and emotions, such as self-control, curiosity, happiness, and consideration, while working-class parents promote deference, obedience, and conformity to authority [@kohn_social_1963; @kohn_class_1969].
-Although parents from all social backgrounds encourage individualism in their children, this shared norm translates into different forms in high and low social classes (1999).
+Although parents from all social backgrounds encourage individualism in their children, this shared norm translates into different forms in high and low social classes [-@kohn_class_1969].
 @acemoglu_obedience_2021 claimed that the values families impart to their children interact with social mobility.

### 2024-06-21 04-analisis.qmd
render paper
add processing docs and fixed references

@@
 author: "Equipo Edumer"
-editor: visual
+bibliography: "input/bib/just-ineq-merit.bib"
+csl: "input/bib/apa6.csl"
 ---
@@
 
-```{r echo=FALSE, include=FALSE}
+```{r echo=FALSE, include=FALSE, eval=FALSE}
 sjmisc::frq(data_rec$educacion_rec)
@@
 
-```{r echo=FALSE, results='asis'}
+```{r echo=FALSE, results='asis', eval=FALSE}
 #| tbl-cap: "Cumulative link multilevel models of differential acces justificacion of health, pensions, and education"
@@
                 custom.gof.rows = list("Controls"=rep("Yes",6)),
-                include.loglik = FALSE, include.aic = FALSE)
+                include.loglik = FALSE, include.aic = FALSE,
+                file = "output/tables/tab-ordinalreg.html")
 
+webshot::webshot(url = "output/tables/tab-ordinalreg.html", file = "output/tables/tab-ordinalreg.png")
 ```
+
+![](output/tables/tab-ordinalreg.png)
 
@@
                          (1 | mrbd), data=data_bienestar)
-
-knitreg(list(reg1_bienestar, reg2_bienestar, reg3_bienestar, reg4_bienestar, reg5_bienestar))
 

### 2024-06-21 05-discussion.qmd
add processing docs and fixed references

@@
 author: "Equipo EDUMER"
-editor: visual
+bibliography: "input/bib/just-ineq-merit.bib"
+csl: "input/bib/apa6.csl"
 ---
@@
 
-Hypothesis $H_x$ [Falta hipótesis school achievement] held that students from schools with higher school achievement would show stronger preferences for market justice, following the argument that such school could foster meritocratic attitudes by showing a larger promotion of talent and effort in order to get better results.
+Hypothesis $H_x$ \[Falta hipótesis school achievement\] held that students from schools with higher school achievement would show stronger preferences for market justice, following the argument that such school could foster meritocratic attitudes by showing a larger promotion of talent and effort in order to get better results.
 Nevertheless, the results show quite the opposite: schools with better academic results (as measured by a national standardized test) show in average less preferences for market justice.

### 2024-06-21 06-conclusion.qmd
add processing docs and fixed references

---
author: "Equipo EDUMER"
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/apa6.csl"
---

# Conclusion

Our research aimed to examine the relationship between the perception of meritocracy and the justification of market justice preferences among eighth-grade students in Chile. From our knowledge this is the first study addressing market justice preferences at school level, finding that students exhibit large preferences for market justice (about one third agree or strongly agree with them), a share that is quite larger when compared with evidence in adult population. Perception of meritocracy is also high, particularly when it comes to the reward of effort, with an striking eighty six percent of agreement. Regarding these descriptive indicators, a a first message coming from this study is that there is something particular in school-age when it comes to perceptions and preferences about inequality. This can relate to development aspects (Kim & Gewirtz, 2019, Rizzo & Killen, 2020) but it could be something specific to the Chilean case, which requires further comparative as well as longitudinal research in the area. In any case, teaching about the origins and characteristics of social inequality is not part of the chilean school curriculum, calling for a more dedicated consideration of these topics in areas such as citizensip education as well as in history and social science education.

The core of this paper was the association between meritocratic perceptions and preferences for market justice. In general, we find that those who perceive that effort and talent/intelligence are rewarded, are more willing to agree with that richer individuals can have better health, education, and pensions. This is not the first time that meritocratic perceptions (and also beliefs) have been related with the legitimation of social inequalities [@mijs_paradox_2019; @darnon_where_2018], but the consideration of school level population as well as market justice preferences allows adding evidence regarding the role of meritocracy regarding inequality beliefs: legitimation of market criteria in social services seems to begin at school age, and what students think of meritocracy is linked to it. Although there are still several open questions, as for instance differences between meritocratic perceptions in school and in the society at large, this raises up the relevance of what is done (or not) in schools in order to challenge the meritocratic ideals.

The consideration of the school context opened some caveats in this research. We found contrary evidence for the association between status (as measured by average education), school performance, and market justice prefecences: high status and high achievement at school level is asociated with less market justice preferences. This is opposed to the rational-interest argument that suggests that those doing better - both economically and academically- would demand less redistribution (Gonthier,2016; Runst, 2018; Bullock, 2021) and therefore are expected to challenge market mechanisms. A possible explanation is that education contributes to form a crirical view of markets in redistribution, nevertheless further reserach in this area is needed in order to assert the specific mechanisms that might be playing a role. All in all, from this study we know that school contexts are not trivial to this regard, calling for larger attention to school age research on the formation of inequality preferences and beliefs.

### 2024-06-21 processing/03-justpension.Rmd
add processing docs and fixed references

---
title: "01-justpension"
author: "Equipo EDUMER"
date: '2022-06-27'
output: 
  html_document: 
    toc: yes
    code_folding: hide
    toc_float: 
      collapsed: true
      smooth_scroll: false
      number_sections: true
---

```{r}
pacman::p_load(haven, dplyr, summarytools, sjmisc, car, sjlabelled, sjPlot, lme4, stargazer, corrplot, ltm, texreg, ordinal, MASS, webshot2)
load("input/data/proc/data_rec.RData")
```

# Modelos de regresión logística ordenada (ordinal logit) (CLMM del paquete ordinal - Cumulative link mixed models)

## Es justo que las personas de ingresos altos puedan optar a mejores pensiones

```{r}
data_pension <- data_rec %>% dplyr::select(-just_educ, -just_salud, -redistribucion)
data_pension <- na.omit(data_pension)
```

## Intra class correlation

```{r}
data_pension$just_pension_fact <- as_factor(data_pension$just_pension)
reg0_pension <- clmm(just_pension_fact ~ 1 + (1 | mrbd), data=data_pension)
performance::icc(reg0_pension)
```

```{r }
reg1_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + (1 | mrbd), data=data_pension)
reg2_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + (1 | mrbd), data=data_pension)
reg3_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_pension)
reg4_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + (1 | mrbd), data=data_pension)
reg5_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_pension)
```

```{r results='asis'}
htmlreg(list(reg0_pension, reg1_pension, reg2_pension, reg3_pension, reg4_pension, reg5_pension), 
        custom.model.names = c("Model 0", 
                               "Model 1",
                               "Model 2",
                               "Model 3",
                               "Model 4",
                               "Model 5"), 
        custom.note = "$***p <$ 0.001, $**p <$ 0.01, $*p <$ 0.05", 
        omit.coef = "(internet_rec)",
        custom.coef.names = c("Strongly Disagree|Disagree",
                              "Disagree|Agree",
                              "Agree|Strongly Agree",
                              "Social effort",
                              "Deservingness", 
                              "Social talent", 
                              "School effort", 
                              "School talent", 
                              "Secondary <br> <i>(Ref. 8th grade or less)</i>", 
                              "Higher tec.", 
                              "University or posgraduate", 
                              "Missing", 
                              "More than 25 books <br> <i>(Ref. Less than 25)</i>", 
                              "Technology access",
                              "Prop. university level at school",
                              "Subsidized private <br> <i>(Ref. Public)</i>", 
                              "Private", 
                              "SES Medium low <br> <i>(Ref. Low)</i>", 
                              "SES Medium", 
                              "SES Medium high", 
                              "SES High", 
                              "Simce Medium <i>(Ref. Low)</i>", 
                              "Simce High"),
        caption.above = TRUE,
        file = "output/tables/reg_justpension.html")

webshot(url ="output/tables/reg_justpension.html" ,file ="output/tables/reg_justpension.png")
```

### Interacciones

*meritocracia x ses familiar (educacion)*

```{r}
reg10_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*esfuerzo_soc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_pension)
reg11_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*merito_soc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_pension)
reg12_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*inteligencia_soc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_pension)
reg13_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*esfuerzo_esc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_pension)
reg14_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*inteligencia_esc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_pension)
```

```{r results='asis'}
htmlreg(list(reg9_pension, reg10_pension, reg11_pension, reg12_pension, reg13_pension, reg14_pension))
```

*meritocracia x ses familiar (libros)*

```{r}
reg15_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*esfuerzo_soc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_pension)
reg16_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*merito_soc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_pension)
reg17_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*inteligencia_soc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_pension)
reg18_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*esfuerzo_esc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_pension)
reg19_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*inteligencia_esc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_pension)
```

```{r results='asis'}
htmlreg(list(reg9_pension, reg15_pension, reg16_pension, reg17_pension, reg18_pension, reg19_pension))
```

*meritocracia x ses escuela*

```{r}
reg20_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*esfuerzo_soc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_pension)
reg21_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*merito_soc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_pension)
reg22_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*inteligencia_soc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_pension)
reg23_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*esfuerzo_esc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_pension)
reg24_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*inteligencia_esc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_pension)
```

```{r results='asis'}
htmlreg(list(reg9_pension, reg20_pension, reg21_pension, reg22_pension, reg23_pension, reg24_pension))
```

*meritocracia x simce*

```{r}
reg25_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*esfuerzo_soc + internet_rec + (1 + simce | mrbd), data=data_pension)
reg26_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*merito_soc + internet_rec + (1 + simce | mrbd), data=data_pension)
reg27_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*inteligencia_soc + internet_rec + (1 + simce | mrbd), data=data_pension)
reg28_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*esfuerzo_esc + internet_rec + (1 + simce | mrbd), data=data_pension)
reg29_pension <- clmm(just_pension_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*inteligencia_esc + internet_rec + (1 + simce | mrbd), data=data_pension)
```

```{r results='asis'}
htmlreg(list(reg9_pension, reg25_pension, reg26_pension, reg27_pension, reg28_pension, reg29_pension))
```

### 2024-06-21 processing/04-justsalud.Rmd
add processing docs and fixed references

---
title: "03-justsalud"
author: "Equipo EDUMER"
date: '2022-06-27'
output: 
  html_document: 
    toc: yes
    code_folding: hide
    toc_float: 
      collapsed: true
      smooth_scroll: false
      number_sections: true
---

```{r}
pacman::p_load(haven, dplyr, summarytools, sjmisc, car, sjlabelled, sjPlot, lme4, stargazer, corrplot, ltm, texreg, ordinal, MASS, webshot2)
load("input/data/proc/data_rec.RData")
```

# Modelos de regresión logística ordenada (ordinal logit) (CLMM del paquete ordinal - Cumulative link mixed models)

## Es justo que las personas de ingresos altos puedan optar a mejor salud

```{r}
data_salud <- data_rec %>% dplyr::select(-just_pension, -just_educ, -redistribucion)
data_salud <- na.omit(data_salud)
```

## Intra class correlation

```{r}
data_salud$just_salud_fact <- as_factor(data_salud$just_salud)
reg0_salud <- clmm(just_salud_fact ~ 1 + (1 | mrbd), data=data_salud)
performance::icc(reg0_salud)
```

```{r}
reg1_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + (1 | mrbd), data=data_salud)
reg2_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + (1 | mrbd), data=data_salud)
reg3_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_salud)
reg4_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + (1 | mrbd), data=data_salud)
reg5_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_salud)
```

```{r results='asis'}
htmlreg(list(reg0_salud, reg1_salud, reg2_salud, reg3_salud, reg4_salud, reg5_salud), 
        custom.model.names = c("Model 0", 
                               "Model 1",
                               "Model 2",
                               "Model 3",
                               "Model 4",
                               "Model 5"), 
        custom.note = "$***p <$ 0.001, $**p <$ 0.01, $*p <$ 0.05", 
        omit.coef = "(internet_rec)",
        custom.coef.names = c("Strongly Disagree|Disagree",
                              "Disagree|Agree",
                              "Agree|Strongly Agree",
                              "Social effort",
                              "Deservingness", 
                              "Social talent", 
                              "School effort", 
                              "School talent", 
                              "Secondary <br> <i>(Ref. 8th grade or less)</i>", 
                              "Higher tec.", 
                              "University or posgraduate", 
                              "Missing", 
                              "More than 25 books <br> <i>(Ref. Less than 25)</i>", 
                              "Technology access",
                              "Prop. university level at school",
                              "Subsidized private <br> <i>(Ref. Public)</i>", 
                              "Private", 
                              "SES Medium low <br> <i>(Ref. Low)</i>", 
                              "SES Medium", 
                              "SES Medium high", 
                              "SES High", 
                              "Simce Medium <i>(Ref. Low)</i>", 
                              "Simce High"),
        caption.above = TRUE,
        file = "output/tables/reg_justsalud.html")

webshot(url ="output/tables/reg_justsalud.html" ,file ="output/tables/reg_justsalud.png")
```

### Interacciones

*meritocracia x ses familiar (educacion)*

```{r}
reg10_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*esfuerzo_soc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_salud)
reg11_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*merito_soc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_salud)
reg12_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*inteligencia_soc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_salud)
reg13_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*esfuerzo_esc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_salud)
reg14_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*inteligencia_esc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_salud)
```

```{r results='asis'}
htmlreg(list(reg9_salud, reg10_salud, reg11_salud, reg12_salud, reg13_salud, reg14_salud))
```

*meritocracia x ses familiar (libros)*

```{r}
reg15_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*esfuerzo_soc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_salud)
reg16_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*merito_soc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_salud)
reg17_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*inteligencia_soc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_salud)
reg18_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*esfuerzo_esc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_salud)
reg19_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*inteligencia_esc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_salud)
```

```{r results='asis'}
htmlreg(list(reg9_salud, reg15_salud, reg16_salud, reg17_salud, reg18_salud, reg19_salud))
```

*meritocracia x ses escuela*

```{r}
reg20_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*esfuerzo_soc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_salud)
reg21_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*merito_soc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_salud)
reg22_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*inteligencia_soc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_salud)
reg23_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*esfuerzo_esc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_salud)
reg24_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*inteligencia_esc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_salud)
```

```{r results='asis'}
htmlreg(list(reg9_salud, reg20_salud, reg21_salud, reg22_salud, reg23_salud, reg24_salud))
```

*meritocracia x simce*

```{r}
reg25_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*esfuerzo_soc + internet_rec + (1 + simce | mrbd), data=data_salud)
reg26_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*merito_soc + internet_rec + (1 + simce | mrbd), data=data_salud)
reg27_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*inteligencia_soc + internet_rec + (1 + simce | mrbd), data=data_salud)
reg28_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*esfuerzo_esc + internet_rec + (1 + simce | mrbd), data=data_salud)
reg29_salud <- clmm(just_salud_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*inteligencia_esc + internet_rec + (1 + simce | mrbd), data=data_salud)
```

```{r results='asis'}
htmlreg(list(reg9_salud, reg25_salud, reg26_salud, reg27_salud, reg28_salud, reg29_salud))
```

### 2024-06-21 processing/05-justeduc.Rmd
add processing docs and fixed references

---
title: "02-justeduc"
author: "Equipo EDUMER"
date: '2022-06-27'
output: 
  html_document: 
    toc: yes
    code_folding: hide
    toc_float: 
      collapsed: true
      smooth_scroll: false
      number_sections: true
---

```{r}
pacman::p_load(haven, dplyr, summarytools, sjmisc, car, sjlabelled, sjPlot, lme4, stargazer, corrplot, ltm, texreg, ordinal, MASS, webshot2)

load("input/data/proc/data_rec.RData")
```

# Modelos de regresión logística ordenada (ordinal logit) (CLMM del paquete ordinal - Cumulative link mixed models)

## Es justo que las personas de ingresos altos puedan optar a mejor educacion

```{r}
data_educ <- data_rec %>% dplyr::select(-just_pension, -just_salud, -redistribucion)
data_educ <- na.omit(data_educ)
```

## Intra class correlation

```{r}
data_educ$just_educ_fact <- as_factor(data_educ$just_educ)
reg0_educ <- clmm(just_educ_fact ~ 1 + (1 | mrbd), data = data_educ)
performance::icc(reg0_educ)
```

```{r}
reg1_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + (1 | mrbd), data=data_educ)
reg2_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + (1 | mrbd), data=data_educ)
reg3_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_educ)
reg4_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + (1 | mrbd), data=data_educ)
reg5_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_educ)
```
```{r results='asis'}
htmlreg(list(reg0_educ, reg1_educ, reg2_educ, reg3_educ, reg4_educ, reg5_educ), 
        custom.model.names = c("Model 0", 
                               "Model 1",
                               "Model 2",
                               "Model 3",
                               "Model 4",
                               "Model 5"), 
        custom.note = "$***p <$ 0.001, $**p <$ 0.01, $*p <$ 0.05", 
        omit.coef = "(internet_rec)",
        custom.coef.names = c("Strongly Disagree|Disagree",
                              "Disagree|Agree",
                              "Agree|Strongly Agree",
                               "Social effort",
                              "Deservingness", 
                              "Social talent", 
                              "School effort", 
                              "School talent", 
                              "Secondary <br> <i>(Ref. 8th grade or less)</i>", 
                              "Higher tec.", 
                              "University or posgraduate", 
                              "Missing", 
                              "More than 25 books <br> <i>(Ref. Less than 25)</i>", 
                              "Technology access",
                              "Prop. university level at school",
                              "Subsidized private <br> <i>(Ref. Public)</i>", 
                              "Private", 
                              "SES Medium low <br> <i>(Ref. Low)</i>", 
                              "SES Medium", 
                              "SES Medium high", 
                              "SES High", 
                              "Simce Medium <i>(Ref. Low)</i>", 
                              "Simce High"),
        caption.above = TRUE,
        file = "output/tables/reg_justeduc.html")

webshot(url ="output/tables/reg_justeduc.html" ,file ="output/tables/reg_justeduc.png")
```

### Interacciones

*meritocracia x ses familiar (educacion)*

```{r}
reg10_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*esfuerzo_soc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_educ)
reg11_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*merito_soc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_educ)
reg12_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*inteligencia_soc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_educ)
reg13_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*esfuerzo_esc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_educ)
reg14_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec*inteligencia_esc + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + educacion_rec | mrbd), data=data_educ)
```

```{r results='asis'}
htmlreg(list(reg9_educ, reg10_educ, reg11_educ, reg12_educ, reg13_educ, reg14_educ))
```

*meritocracia x ses familiar (libros)*

```{r}
reg15_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*esfuerzo_soc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_educ)
reg16_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*merito_soc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_educ)
reg17_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*inteligencia_soc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_educ)
reg18_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*esfuerzo_esc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_educ)
reg19_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec*inteligencia_esc + acc_tec + cod_depe2 + cod_grupo + simce + internet_rec + (1 + libros_rec | mrbd), data=data_educ)
```

```{r results='asis'}
htmlreg(list(reg9_educ, reg15_educ, reg16_educ, reg17_educ, reg18_educ, reg19_educ))
```

*meritocracia x ses escuela*

```{r}
reg20_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*esfuerzo_soc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_educ)
reg21_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*merito_soc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_educ)
reg22_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*inteligencia_soc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_educ)
reg23_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*esfuerzo_esc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_educ)
reg24_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo*inteligencia_esc + simce + internet_rec + (1 + cod_grupo | mrbd), data=data_educ)
```

```{r results='asis'}
htmlreg(list(reg9_educ, reg20_educ, reg21_educ, reg22_educ, reg23_educ, reg24_educ))
```

*meritocracia x simce*

```{r}
reg25_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*esfuerzo_soc + internet_rec + (1 + simce | mrbd), data=data_educ)
reg26_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*merito_soc + internet_rec + (1 + simce | mrbd), data=data_educ)
reg27_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*inteligencia_soc + internet_rec + (1 + simce | mrbd), data=data_educ)
reg28_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*esfuerzo_esc + internet_rec + (1 + simce | mrbd), data=data_educ)
reg29_educ <- clmm(just_educ_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + cod_depe2 + cod_grupo + simce*inteligencia_esc + internet_rec + (1 + simce | mrbd), data=data_educ)
```

```{r results='asis'}
htmlreg(list(reg9_educ, reg25_educ, reg26_educ, reg27_educ, reg28_educ, reg29_educ))
```

### 2024-06-21 processing/06-redis.Rmd
add processing docs and fixed references

---
title: "04-redis"
author: "Equipo EDUMER"
date: '2022-06-27'
output: 
  html_document: 
    toc: yes
    code_folding: hide
    toc_float: 
      collapsed: true
      smooth_scroll: false
      number_sections: true
---

```{r}
pacman::p_load(haven, dplyr, summarytools, sjmisc, car, sjlabelled, sjPlot, lme4, stargazer, corrplot, ltm, texreg, ordinal, MASS, webshot)
load("input/data/proc/alumnos_patterns.RData")
load("input/data/proc/apoderados_patterns.RData")
```

```{r}
data_est <- alumnos_patterns %>% dplyr::select(idalumno,
                                               sexo,
                                               codigoCurso,
                                               mrbd,
                                               cod_depe2, #dependencia administrativa
                                               cod_grupo, #categorizacion socioeconomica
                                               simce_lect=prom_lect8b_rbd, #Promedio lectura escuela
                                               simce_mate=prom_mate8b_rbd, # Promedio matematica escuela
                                               inteligencia_esc=est_p6_1, # inteligencia escolar
                                               esfuerzo_esc=est_p6_2, # esfuerzo escolar
                                               esfuerzo_soc=est_p38_2, #esfuerzo social
                                               merito_soc=est_p38_3, #merito social
                                               inteligencia_soc=est_p38_4, #inteligencia social
                                               just_pension=est_p39_1, #justicia pension
                                               just_educ=est_p39_2, # justicia educacion
                                               just_salud=est_p39_3, # justicia salud
                                               redistribucion=est_p39_4, # redistribucion
                                               libros=est_p3, # cantidad de libros
                                               pc=est_p4_1, # tiene pc
                                               tablet=est_p4_2, # tiene tablet
                                               celular=est_p4_3 # tiene celular
                                               ) %>% as.data.frame(.)

data_apod <- apoderados_patterns %>% dplyr::select(idalumno,
                                               educacion=educ_max, # nivel educacional mas alto
                                               internet=apod_p9 # conexion a internet en la casa
                                               ) %>% as.data.frame(.)

data_apod <- data_apod %>%  set_na(internet, na = c(0,99), drop.levels = FALSE, as.tag = FALSE) # recode missings de doble marca (99) y vacío (0) a NA
```

## merge bases
```{r}
data <- left_join(data_est, data_apod, by="idalumno")
rm(apoderados_patterns, alumnos_patterns, data_est, data_apod)
```

```{r}
#Educacion
data$educacion_rec <- ifelse(is.na(data$educacion), "Ns/Nr", data$educacion)
data <- data %>% rowwise() %>%  mutate(educacion_rec = case_when(educacion_rec==1~"Ed Basica",
                                                 educacion_rec==2~"Ed Basica",
                                                 educacion_rec==3~"Ed Basica",
                                                 educacion_rec==4~"Ed Media",
                                                 educacion_rec==5~"Ed Media",
                                                 educacion_rec==6~"Ed Tecnica",
                                                 educacion_rec==7~"Ed Tecnica",
                                                 educacion_rec==8~"Universidad o posgrado",
                                                 educacion_rec==9~"Universidad o posgrado",
                                                 educacion_rec==10~"Universidad o posgrado",
                                                 educacion_rec=="Ns/Nr"~"Ns/Nr"
                                                 ))
data$educacion_rec <- factor(data$educacion_rec, levels = c("Ed Basica", "Ed Media", "Ed Tecnica", "Universidad o posgrado", "Ns/Nr"))

## Proporcion de Universidad o posgrado por escuela
data <- data %>% group_by(mrbd) %>%
  mutate(mean_educ = mean(educacion_rec=="Universidad o posgrado", na.rm=T))

# Internet
data$internet <- factor(data$internet, labels = c("Si", "No"))
data$internet_rec <- ifelse(is.na(data$internet), "Ns/Nr", data$internet)

# libros
data$libros_rec <- as.numeric(data$libros)
data <- data %>% rowwise() %>% mutate(libros_rec = case_when(libros_rec==1 ~ "Menos de 25",
                                                             libros_rec==2 ~ "Menos de 25",
                                                             libros_rec==3 ~ "Más de 25",
                                                             libros_rec==4 ~ "Más de 25",
                                                             libros_rec==5 ~ "Más de 25"))
data$libros_rec <- factor(data$libros_rec, levels = c("Menos de 25", "Más de 25"))
```

```{r}
data <- data %>% rowwise() %>% mutate(prom_simce = mean(simce_lect,simce_mate))
summary(data$prom_simce)
data$simce <- ntile(data$prom_simce,3)
table(data$simce)
data$simce <- factor(data$simce, labels = c("Bajo", "Medio", "Alto"))
```

```{r}
# Acceso a tecnologia
data$pc <- as.numeric(data$pc)
data$tablet <- as.numeric(data$tablet)
data$celular <- as.numeric(data$celular)

corrplot.mixed(cor(dplyr::select(data, pc, tablet, celular),
                   method = "spearman",
                   use = "complete.obs"))
```

```{r}
ltm::cronbach.alpha(data  %>%
  dplyr::select(pc, tablet, celular), na.rm=TRUE)
```

```{r}
data <- data %>% rowwise() %>% dplyr::mutate(acc_tec = sum(pc, tablet, celular, na.rm = T))
summary(data$acc_tec)
```

```{r}
# Variables meritocracia
data$inteligencia_esc <- as.numeric(data$inteligencia_esc)
data$esfuerzo_esc <- as.numeric(data$esfuerzo_esc)
data$esfuerzo_soc <- as.numeric(data$esfuerzo_soc)
data$merito_soc <- as.numeric(data$merito_soc)
data$inteligencia_soc <- as.numeric(data$inteligencia_soc)
data$just_pension <- as.numeric(data$just_pension)
data$just_salud <- as.numeric(data$just_salud)
data$just_educ <- as.numeric(data$just_educ)
data$redistribucion <- as.numeric(data$redistribucion)
```

```{r}
data_rec <- data %>% dplyr::select(-educacion, -internet, -simce_lect, -simce_mate, -libros, -pc, -tablet, -celular)
```

# Modelos de regresión logística ordenada (ordinal logit) (CLMM del paquete ordinal - Cumulative link mixed models)

## El gobierno deberia hacer algo para disminuir las brechas de ingresos

```{r}
data_redis <- data_rec %>% dplyr::select(-just_pension, -just_educ, -just_salud)
data_redis <- na.omit(data_redis)
data_redis$redistribucion_fact <- as_factor(data_redis$redistribucion)
```

## Intraclass correlation

```{r}
reg0_redis <- clmm(redistribucion_fact ~ 1 + (1 | mrbd), data=data_redis)
performance::icc(reg0_redis)
```

## Regresiones

```{r}
reg1_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + (1 | mrbd), data=data_redis)
reg2_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + (1 | mrbd), data=data_redis)
reg3_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_redis)
reg4_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + (1 | mrbd), data=data_redis)
reg5_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_redis)
```

```{r results='asis'}
htmlreg(list(reg0_redis, reg1_redis, reg2_redis, reg3_redis, reg4_redis, reg5_redis), 
        custom.model.names = c("Nulo",
                               "Modelo 1",
                               "Modelo 2",
                               "Modelo 3",
                               "Modelo 4",
                               "Modelo 5"), 
        custom.note = "$***p <$ 0.001, $**p <$ 0.01, $*p <$ 0.05", 
        omit.coef = "(internet_rec)",
        custom.coef.names = c("Muy desacuerdo|Desacuerdo",
                              "Desacuerdo|De acuerdo",
                              "De acuerdo|Muy de acuerdo",
                              "Esfuerzo social", 
                              "Merito social", 
                              "Inteligencia social", 
                              "Esfuerzo escolar", 
                              "Inteligencia escolar", 
                              "Ed media <br> <i>(Ref. Ed Basica)</i>", 
                              "Ed tecnica", "Universidad o posgrado", 
                              "Ns/Nr", 
                              "Más de 25 libros <br> <i>(Ref. Menos de 25)</i>", 
                              "Acceso a tecnología",
                              "Prop. universitarios escuela",
                              "Part. subv <br> <i>(Ref. Municipal)</i>", 
                              "Part. priv", 
                              "SES Medio bajo <br> <i>(Ref. Bajo)</i>", 
                              "SES Medio", 
                              "SES Medio alto", 
                              "SES Alto", 
                              "Simce_Medio <i>(Ref. Bajo)</i>", 
                              "Simce_alto"))
```

```{r results='asis'}
htmlreg(list(reg0_redis, reg1_redis, reg2_redis, reg3_redis, reg4_redis, reg5_redis), 
        custom.model.names = c("Model 0", 
                               "Model 1",
                               "Model 2",
                               "Model 3",
                               "Model 4",
                               "Model 5"), 
        custom.note = "$***p <$ 0.001, $**p <$ 0.01, $*p <$ 0.05", 
        omit.coef = "(internet_rec)",
        custom.coef.names = c("Strongly Disagree|Disagree",
                              "Disagree|Agree",
                              "Agree|Strongly Agree",
                               "Social effort",
                              "Deservingness", 
                              "Social talent", 
                              "School effort", 
                              "School talent", 
                              "Secondary <br> <i>(Ref. 8th grade or less)</i>", 
                              "Higher tec.", 
                              "University or posgraduate", 
                              "Missing", 
                              "More than 25 books <br> <i>(Ref. Less than 25)</i>", 
                              "Technology access",
                              "Prop. university level at school",
                              "Subsidized private <br> <i>(Ref. Public)</i>", 
                              "Private", 
                              "SES Medium low <br> <i>(Ref. Low)</i>", 
                              "SES Medium", 
                              "SES Medium high", 
                              "SES High", 
                              "Simce Medium <i>(Ref. Low)</i>", 
                              "Simce High"),
        caption = "Redistribution",
        caption.above = TRUE,
        file = "output/tables/reg_redis.html")
webshot(url ="output/tables/reg_redis.html" ,file ="output/tables/reg_redis.png")
webshot(url ="output/tables/reg_redis.html" ,file ="output/tables/reg_redis.pdf")
```

```{r}
labels.1 <- c("1|2", "2|3", "3|4",
              "Esfuerzo social",
                              "Merito social", 
                              "Inteligencia social", 
                              "Esfuerzo escolar", 
                              "Inteligencia escolar", 
                              "Ed media", 
                              "Ed tecnica", "Universidad o posgrado", 
                              "Ns/Nr", 
                              "Más de 25 libros", 
                              "Acceso a tecnología",
                              "Prop. universitarios escuela",
                              "Part. subv", 
                              "Part. priv", 
                              "SES Medio bajo", 
                              "SES Medio", 
                              "SES Medio alto", 
                              "SES Alto", 
                              "Simce_Medio", 
                              "Simce_alto",
                              "Internet",
                              "NS/NR"
                              )
labels2.1 <- rev(labels.1)
plot2 <- plot_model(reg10.1_redis, title = "", show.values = TRUE, value.offset = .5, value.size = 4, axis.labels = labels2.1)

ggsave(plot2, file="output/graphs/plot2.png")
```

```{r}
easystats::model_dashboard(reg5_redis, output_file = "dashboard_redistribution.html", output_dir = "output/tables")
```

### 2024-06-21 processing/08-coef_plots.Rmd
add processing docs and fixed references

---
title: "coef plot"
author: "equipo EDUMER"
date: "2022-12-13"
output: html_document
---

```{r}
pacman::p_load(dplyr, sjPlot, lme4, stargazer, corrplot, texreg, ordinal, MASS, webshot2, ggplot2, sjlabelled, tidyverse, ggpubr)

load("input/data/proc/data_rec.RData")
```

# Gráfico bivariado
```{r}
soc_effort = data_rec %>% filter(!is.na(esfuerzo_soc)) %>% 
  group_by(esfuerzo_soc) %>%
  summarise(education = mean(just_educ, na.rm = T),
            pensions = mean(just_pension, na.rm = T),
            health = mean(just_salud, na.rm = T))

soc_effort <- soc_effort %>% pivot_longer(c("health", "education", "pensions"), 
                              names_to="justice", 
                              values_to="mean")

deservingness = data_rec %>% filter(!is.na(merito_soc)) %>% 
  group_by(merito_soc) %>%
  summarise(education = mean(just_educ, na.rm = T),
            pensions = mean(just_pension, na.rm = T),
            health = mean(just_salud, na.rm = T))

deservingness <- deservingness %>% pivot_longer(c("health", "education", "pensions"), 
                              names_to="justice", 
                              values_to="mean")

soc_talent = data_rec %>% filter(!is.na(inteligencia_soc)) %>% 
  group_by(inteligencia_soc) %>%
  summarise(education = mean(just_educ, na.rm = T),
            pensions = mean(just_pension, na.rm = T),
            health = mean(just_salud, na.rm = T))

soc_talent <- soc_talent %>% pivot_longer(c("health", "education", "pensions"), 
                              names_to="justice", 
                              values_to="mean")

school_effort = data_rec %>% filter(!is.na(esfuerzo_esc)) %>% 
  group_by(esfuerzo_esc) %>%
  summarise(education = mean(just_educ, na.rm = T),
            pensions = mean(just_pension, na.rm = T),
            health = mean(just_salud, na.rm = T))

school_effort <- school_effort %>% pivot_longer(c("health", "education", "pensions"), 
                              names_to="justice", 
                              values_to="mean")

school_talent = data_rec %>% filter(!is.na(inteligencia_esc)) %>% 
  group_by(inteligencia_esc) %>%
  summarise(education = mean(just_educ, na.rm = T),
            pensions = mean(just_pension, na.rm = T),
            health = mean(just_salud, na.rm = T))

school_talent <- school_talent %>% pivot_longer(c("health", "education", "pensions"), 
                              names_to="justice", 
                              values_to="mean")
```

```{r}
plot_soc_effort <- soc_effort %>% ggplot(aes(esfuerzo_soc, mean)) +
  theme_bw(base_size = 15) +
  geom_point(aes(color = justice, shape = justice), size = 4) +
  xlab("Social effort") +
  ylab("Inequality Justification") +
  scale_y_continuous(limits = c(1.5, 3)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly 
disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly 
agree")) +
  labs(color = "Justification of inequality", shape = "Justification of inequality")

plot_deservingness <- deservingness %>% ggplot(aes(merito_soc, mean)) +
  theme_bw(base_size = 15) +
  geom_point(aes(color = justice, shape = justice), size = 4) +
  xlab("Deservingness") +
  ylab("Inequality Justification") +
  scale_y_continuous(limits = c(1.5, 3)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly 
disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly 
agree")) +
  labs(color = "Justification of inequality", shape = "Justification of inequality")

plot_soc_talent <- soc_talent %>% ggplot(aes(inteligencia_soc, mean)) +
  theme_bw(base_size = 15) +
  geom_point(aes(color = justice, shape = justice), size = 4) +
  xlab("Social talent") +
  ylab("Inequality Justification") +
  scale_y_continuous(limits = c(1.5, 3)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly 
disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly 
agree")) +
  labs(color = "Justification of inequality", shape = "Justification of inequality")

plot_school_effort <- school_effort %>% ggplot(aes(esfuerzo_esc, mean)) +
  theme_bw(base_size = 15) +
  geom_point(aes(color = justice, shape = justice), size = 4) +
  xlab("School effort") +
  ylab("") +
  scale_y_continuous(limits = c(1.5, 3)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly 
disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly 
agree")) +
  labs(color = "Justification of inequality", shape = "Justification of inequality")

plot_school_talent <- school_talent %>% ggplot(aes(inteligencia_esc, mean)) +
  theme_bw(base_size = 15) +
  geom_point(aes(color = justice, shape = justice), size = 4) +
  xlab("School talent") +
  ylab("") +
  scale_y_continuous(limits = c(1.5, 3)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly 
disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly 
agree")) +
  labs(color = "Justification of inequality", shape = "Justification of inequality")

plot_merit_justice <- ggarrange(plot_soc_effort, plot_school_effort, plot_soc_talent, plot_school_talent, plot_deservingness, common.legend = TRUE, ncol = 2, nrow = 3, legend = "bottom")
plot_merit_justice

ggsave(plot_merit_justice, file="output/graphs/plot_merit_justice.png", height = 10, width = 10)

plot_merit_justice_school <- ggarrange(plot_school_effort, plot_school_talent, common.legend = TRUE, legend = "bottom")
plot_merit_justice_school
ggsave(plot_merit_justice_school, file="output/graphs/plot_merit_justice_school.png", height = 7, width = 10)

plot_merit_justice_soc <- ggarrange(plot_soc_effort, plot_soc_talent, plot_deservingness, common.legend = TRUE, ncol = 3, nrow = 1, legend = "bottom")
plot_merit_justice_soc
ggsave(plot_merit_justice_soc, file="output/graphs/plot_merit_justice_soc.png", height = 7, width = 10)
```

## Correlaciones

```{r}
merit_cor<-
  data_rec %>%
  dplyr::select(just_pension, just_educ, just_salud, esfuerzo_soc, inteligencia_soc, merito_soc, esfuerzo_esc, inteligencia_esc) %>%
  lavaan::lavCor(., ordered=names(.))
diag(merit_cor) = NA
rownames(merit_cor) <- c("A. Pension",
                         "B. Education",
                         "C. Health",
                         "D. Social effort",
                         "E. Social talent",
                         "F. Deservingness",
                         "G. School effort",
                         "H. School talent")
colnames(merit_cor) <-c("(A)", "(B)","(C)", "(D)", "(E)", "(F)", "(G)", "(H)")

png("output/graphs/merit_cor.png",width=600,height=300, pointsize = 13)
corrplot::corrplot(merit_cor,
  method = "color",
  addCoef.col = "#000390",
  type = "upper",
  tl.col = "black",
  col=colorRampPalette(c("white","#0068DC"))(12),
  bg = "white",
  na.label = "-")

dev.off()
```

## Es justo que las personas de ingresos altos puedan optar a mejores pensiones

```{r}
data_pension <- data_rec %>% dplyr::select(-just_educ, -just_salud, -redistribucion)
data_pension <- na.omit(data_pension)
data_pension$just_pension_fact <- as_factor(data_pension$just_pension)
reg2_pension <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_pension)
reg5_pension <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_pension)
```

## Es justo que las personas de ingresos altos puedan optar a mejor salud

```{r}
data_salud <- data_rec %>% dplyr::select(-just_pension, -just_educ, -redistribucion)
data_salud <- na.omit(data_salud)
data_salud$just_salud_fact <- as_factor(data_salud$just_salud)
reg2_salud <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_salud)
reg5_salud <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_salud)
```

## Es justo que las personas de ingresos altos puedan optar a mejor educacion

```{r}
data_educ <- data_rec %>% dplyr::select(-just_pension, -just_salud, -redistribucion)
data_educ <- na.omit(data_educ)
data_educ$just_educ_fact <- as_factor(data_educ$just_educ)
reg2_educ <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_educ)
reg5_educ <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_educ)
```

# coef plot con modelo 2 (solo variables meritocracia)

```{r eval=FALSE}
plots <- plot_models(reg2_salud, reg2_pension, reg2_educ, 
            rm.terms = c("1|2", "2|3", "3|4"), #quitar terminos
            title = "", #quitar titulo
            show.values = TRUE, #mostrar valor de efectos
            dot.size = 3, #tamaño circulos
            value.size = 3.5, #tamaño valor efectoss
            spacing = .7, #espacio entre efectos
            axis.labels = rev(c("School talent", 
                            "School effort", 
                            "Social talent",
                            "Social effort", 
                            "Deservingness")), #con rev porque automatico los tira en otro orden
            m.labels = c("Health", "Pension", "Education")) # variables dependientes

ggsave(plots, file="output/graphs/plots.png")
```

```{r eval=FALSE}
data_bienestar <- na.omit(data_rec)
data_bienestar <- data_bienestar %>% rowwise() %>% dplyr::mutate(bienestar = mean(c(just_pension, just_educ, just_salud), na.omit=TRUE))
data_bienestar$bienestar <- round(data_bienestar$bienestar, 2)

reg5_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_bienestar)
```

# coef plot modelo 5 (todas las variables)

```{r}
plots2<-plot_models(reg5_salud, reg5_educ, reg5_pension,  
            rm.terms = c("1|2", "2|3", "3|4", "educacion_recEd Media", "educacion_recEd Tecnica", "educacion_recUniversidad o posgrado", "educacion_recNs/Nr", "libros_recMás de 25", "acc_tec", "internet_rec2", "internet_recNs/Nr", "mean_educ", "cod_depe2Part. subvencionado", "cod_depe2Part. privado", "cod_grupoMedio bajo", "cod_grupoMedio", "cod_grupoMedio alto", "cod_grupoAlto", "simceMedio", "simceAlto"), #quitar terminos
            title = "", #quitar titulo
            show.values = TRUE, #mostrar valor de efectos
            dot.size = 1.3, #tamaño circulos
            line.size = 1, #tamaño CI
            value.size = 3.5, #tamaño valor efectoss
            spacing = .7, #espacio entre efectos
            grid = TRUE, # separa en un plot por modelo
            vline.color = "red", # linea roja en punto neutro (1)
            axis.labels = rev(c("School talent", 
                            "School effort", 
                            "Social talent",
                            "Social effort", 
                            "Deservingness"
                            )), #con rev porque automatico los tira en otro orden
            show.legend = FALSE,
            m.labels = c("Health", "Education","Pension")) + # variables dependientes
  theme_bw(base_size = 16)

plots2

ggsave(plots2, file="output/graphs/plots2.png", width = 10, height = 7)
```

#### Mismo plot pero en español para presentación
```{r}
plots2_esp<-plot_models(reg5_salud, reg5_pension, reg5_educ, 
            rm.terms = c("1|2", "2|3", "3|4", "educacion_recEd Media", "educacion_recEd Tecnica", "educacion_recUniversidad o posgrado", "educacion_recNs/Nr", "libros_recMás de 25", "acc_tec", "internet_rec2", "internet_recNs/Nr", "mean_educ", "cod_depe2Part. subvencionado", "cod_depe2Part. privado", "cod_grupoMedio bajo", "cod_grupoMedio", "cod_grupoMedio alto", "cod_grupoAlto", "simceMedio", "simceAlto"), #quitar terminos
            title = "", #quitar titulo
            show.values = TRUE, #mostrar valor de efectos
            dot.size = 1.3, #tamaño circulos
            line.size = 1, #tamaño CI
            value.size = 3.5, #tamaño valor efectoss
            spacing = .7, #espacio entre efectos
            grid = TRUE, # separa en un plot por modelo
            vline.color = "red", # linea roja en punto neutro (1)
            axis.labels = rev(c("Esfuerzo social", 
                            "Merecimiento", 
                            "Talento social", 
                            "Esfuerzo escolar", 
                            "Talento escolar")), #con rev porque automatico los tira en otro orden
            show.legend = FALSE,
            m.labels = c("Salud", "Pensiones", "Educación")) + # variables dependientes
  theme_bw(base_size = 16)

plots2_esp

ggsave(plots2_esp, file="output/graphs/plots2_esp.png", width = 10, height = 7)
```

### 2024-06-21 processing/issp2019.Rmd
add processing docs and fixed references

---
title: "ISSP2019"
author: "Equipo EDUMER"
date: "2022-11-28"
output: html_document
---

```{r}
pacman::p_load(dplyr, ggplot2, sjPlot, haven, sjlabelled, tidyverse)

data <- read_sav("input/data/original/ISSP2019.sav")
```

```{r}
table(data$v30)
summary(data$v30)
data$v30 <- factor(data$v30, 
                   levels = c(5,4,3,2,1),
                   labels = c("Very unjust, definitely wrong", 
                              "Somewhat unjust, wrong", 
                              "Neither just nor unjust, mixed feelings", 
                              "Somewhat just, right", 
                              "Very just, definitely right"
                              ))

data$v30 <- set_label(data$v30, label = "Just or unjust that people with higher incomes can buy better health care than people with lower incomes?")

table(data$v31)
summary(data$v31)
data$v31 <- factor(data$v31, 
                   levels = c(5,4,3,2,1),
                   labels = c("Very unjust, definitely wrong", 
                              "Somewhat unjust, wrong", 
                              "Neither just nor unjust, mixed feelings", 
                              "Somewhat just, right", 
                              "Very just, definitely right"
                              ))

data$v31 <- set_label(data$v31, label = "Just or unjust that people with higher incomes can buy better education for children than people with lower incomes?")
```

```{r}
issp2019 <- sjPlot::plot_stackfrq(dplyr::select(data, v30, v31), geom.size = 0.4) +
  theme_bw() +
  theme(legend.position="bottom")
issp2019

ggsave(issp2019, file = "output/graphs/issp2019.png",device = "png",width = 25,height = 13,dpi = "retina",units = "cm")
```

# Cleveland plot

```{r}
data$v30 <- as.numeric(data$v30)
data$v31 <- as.numeric(data$v31)

data$country <- factor(data$country, labels = get_labels(data$country))

health <- data %>% group_by(country) %>% 
  summarise(Health = mean(v30, na.rm=TRUE)) %>% 
  arrange(Health) %>% 
  mutate(country = factor(country, levels = .$country))

education <- data %>% group_by(country) %>%
  summarise(Education = mean(v31, na.rm=TRUE)) %>% 
  arrange(Education) %>% 
  mutate(country = factor(country, levels = .$country))

base <- left_join(health, education, by="country")

base <- base[order(base$Health), ]

base <- base %>% pivot_longer(c("Health", "Education"), 
                              names_to="justice", 
                              values_to="mean")
```

```{r}
cleveland_issp <- base %>% ggplot(aes(mean, country)) +
  theme_bw() +
  geom_point(aes(color = justice, shape = justice), size = 3) +
  stat_summary(aes(xintercept = after_stat(x), y = 0), fun = mean, geom = "vline", orientation = "y", colour = "green") +
  xlab("average") +
  scale_x_continuous(limits = c(1, 5),
                     breaks = seq(1, 5, by = 1),
                     label = c("Very unjust, definitely wrong", 
                               "Somewhat unjust, wrong", 
                               "Neither just nor unjust, 
  mixed feelings", 
                               "Somewhat just, right", 
                               "Very just, definitely right"))
cleveland_issp

ggsave(cleveland_issp, file = "output/graphs/cleveland_issp.png",device = "png",width = 25,height = 13,dpi = "retina",units = "cm")
```

# Solo Chile

```{r}
data_chile <- data %>% filter(country==152)
data_chile$v30 <- as.numeric(data_chile$v30)
data_chile$v31 <- as.numeric(data_chile$v31)
issp2019_chile <- sjPlot::plot_stackfrq(select(data_chile, v30, v31), geom.size = 0.4) +
  theme_bw() +
  theme(legend.position="bottom")
issp2019_chile

ggsave(issp2019_chile, file = "output/graphs/issp2019_chile.png",device = "png",width = 25,height = 13,dpi = "retina",units = "cm")
```

### 2024-06-24 01-introduction.qmd
Revision Andreas y Render completo
Revision Andreas:
- Falta discussion de H7
- Referencias que faltan agregarlas a Zotero (Kevin)

--- 2024-06-19 01-introduction.qmd
@@
 Since its origins, educational institutions have been related to the idea of social mobility and access to better opportunities.
-Therefore, the consistent evidence of the high level of social reproduction at the school level represents a threat to the promise of education and a meritocratic system  [@bourdieu_reproduction_1990].
+Despite this, the consistent evidence of the high level of social reproduction at the school level represents a threat to the promise of education and a meritocratic system  [@bourdieu_reproduction_1990].
 A large part of the research in this field at an international  level has addressed the extent to which the social origin of students affects their academic results and their life opportunities [@vonhippel_test_2019], confirming that schools have severe difficulties in closing the socio-economic and cultural gaps of origin.
@@
 In the present paper, we address the role of the perception of meritocracy on the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system.
-In particular, we focus in justifying inequalities in health, education and pensions, which are traditional social policy areas where acces to better services could be justified based on payment capacity, referred as to market justice [@lane_market_1986, @lindh_bringing_2023].\
+In particular, we focus in justifying inequalities in health, education and pensions, which are traditional social policy areas where acces to better services could be justified based on payment capacity, referred as to market justice [@lane_market_1986, @lindh_bringing_2023].
 Most of the research in this realm has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies [@batruch_belief_2022].
@@
 
-Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population [@batruch_belief_2022; @wiederkehr_belief_2015], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger justification of social inequalities, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary [@batruch_belief_2022].
+Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population [@batruch_belief_2022; @wiederkehr_belief_2015], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger market justice preferences, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary [@batruch_belief_2022].
 We focus on the student-age population as we point out that it is possible to track down the origin of meritocratic beliefs (and their consequences) to early socialization processes.
@@
 This shift in economic policy has allowed the unprecedented emergence of markets in areas related to social reproduction, such as health, pensions, and education, with provision and access managed by private entities and segmented by individual payment capacity, heavily reliant on State subsidies [@boccardo_30_2020].
-In health, although the majority of the population uses the public insurance system (78.9%), the pension system is based on individual capitalization, with mandatory contributions managed by private administrators investing in the financial market, currently involving 11 million contributors (Superintendencia de Pensiones, 2024).
-In education, 30.6% of school enrollment is in public schools, 54.0% in state-subsidized (voucher) private schools, and 9.3% in fully private schools, generally attracting higher-income groups (Mineduc, 2023).
+In health, although the majority of the population uses the public insurance system (78.9%), 15.3% are served by private insurers (Observatorio Social, 2023).
+The pension system is based on individual capitalization, with mandatory contributions managed by private administrators investing in the financial market, currently involving 11 million contributors (Superintendencia de Pensiones, 2024).
+In education, 30.6% of school enrollment is in public schools, 54.0% in state-subsidized (voucher) private schools, and 9.3% in fully private schools, generally attracting higher-income groups (Ministerio de Educación, 2023).

### 2024-06-24 04-analisis.qmd
Revision Andreas y Render completo
Revision Andreas:
- Falta discussion de H7
- Referencias que faltan agregarlas a Zotero (Kevin)

--- 2024-06-18 04-analisis.qmd
@@
                kableExtra,
-               easystats)
+               easystats,
+               ggeffects)
 
@@
 
-This contrasts with the pattern observed for these items in the Chilean adult population (see @fig-elsoc).
-While both age groups tend to disagree that higher-income individuals should access better health, pension, and education services, this tendency is more pronounced in adults than in students, who show greater agreement with market justice preferences in these areas.
-Additionally, the hierarchy of these domains differs between groups; adults justify market justice least in education, followed by health, and then pensions, whereas students justify market principles less in health, then pensions, and lastly in education.
-
 ```{r echo=FALSE}
@@
         plot.caption = element_text(hjust = 1)) +
-       labs(caption = paste0("Source: Authors calculation based on First Study of Civic Education 2017",
-                             " (N = ",dim(data_rec)[1],")"))
+       labs(caption = "Source: own elaboration based on First Study of Civic Education 2017")
 
@@
   annotate("text", x = 3.5, y = 0, label = "Social meritocracy", size = 5, hjust = 0, vjust = 1) +
-       labs(caption = paste0("Source: Authors calculation based on First Study of Civic Education 2017",
-                             " (N = ",dim(data_rec)[1],")"))
+       labs(caption = "Source: own elaboration based on First Study of Civic Education 2017")
 ```
@@
           plot_school_talent, plot_deservingness, leg, ncol = 2, nrow = 3) +
-  labs(caption = paste0("Source: Authors calculation based on First Study of Civic Education 2017",
-                             " (N = ",dim(data_rec)[1],")"))
+  labs(caption = "Source: own elaboration based on First Study of Civic Education 2017")
 ```
@@
                 stars = c(0.05, 0.01, 0.001),
-                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars",
+                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses.
+Control variables are included.
+%stars.
+Source: own elaboration based on First Study of Civic Education 2017.",
                 custom.header = list("Health" = 1:2, "Pensions" = 3:4, "Education" = 5:6),
@@
 
-knitreg(list(reg1_bienestar, reg2_bienestar, reg3_bienestar, reg4_bienestar, reg5_bienestar))
-
 test01 <- anova(reg0_bienestar, reg1_bienestar, test = "Chisq")
@@
                 stars = c(0.05, 0.01, 0.001),
-                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars",
+                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses.
+Control variables are included.
+%stars.
+Source: own elaboration based on First Study of Civic Education 2017.",
                 custom.coef.map = ccoef,
@@
   collapse_rows(columns = 1:2, valign = "middle") %>% 
-  kableExtra::add_footnote(label = "Note: ***p < 0.001, **p < 0.01, *p < 0.05", 
+  kableExtra::add_footnote(label = "Note: ***p < 0.001, **p < 0.01, *p < 0.05. Source: own elaboration based on First Study of Civic Education 2017.", 
                            notation = "none")
@@
 
-plot_interact<- plot_model(reg26.2_bienestar, type = "int", mdrt.values="minmax", title = "", colors = c("red", "white", "blue")) +
-  ylim(1,4)+
-  labs(x = "Deservingness", y="Market Justice Preferences", colour = "School Simce")+
-  theme_bw()
+plot_interact <- ggpredict(model = reg26.2_bienestar, terms = c("merito_soc", "simce")) %>% 
+  filter(group != "Medium") %>% 
+  plot() +
+  ylim(1,4) +
+  labs(title = NULL,
+       x = "Deservingness", 
+       y ="Market Justice Preferences", 
+       colour = "School Simce",
+       caption = "Source: own elaboration based on First Study of Civic Education 2017")
 ```
@@
 
-plot_interact
+plot_interact +
+  theme_bw()
 

### 2024-06-24 05-discussion.qmd
Revision Andreas y Render completo
Revision Andreas:
- Falta discussion de H7
- Referencias que faltan agregarlas a Zotero (Kevin)

--- 2024-06-18 05-discussion.qmd
@@
 
-The previous analyses were aimed at delving into the relationship between perceptions of meritocracy and preferences for market justice among students.
-Several models estimated how students' beliefs about the rewards for talent and effort in both society and school environments influence their attitudes towards access to social services like health, education, and pensions.
-The models also examined the impact of students' socioeconomic status and school achievement on these preferences, analyzing whether perceptions of meritocracy moderate these effects.
-The findings provide insights into the complex interplay between meritocratic beliefs, socioeconomic factors, and attitudes towards social justice.
+The previous analyses aimed to explore the relationship between perceptions of meritocracy and preferences for market justice among students in Chilean society.
+Several models estimated how students' beliefs about the rewards for talent and effort in both societal and school contexts influence their attitudes toward access to social services like health, education, and pensions based on pay capacity.
+Additionally, the models examined the impact of students' socioeconomic status and school achievement on these preferences, assessing whether perceptions of meritocracy moderate these effects.
+The findings offer insights into the complex interplay between meritocratic beliefs, socioeconomic factors, and attitudes toward market justice criteria.
 
@@
 
-Hypothesis $H_x$ [Falta hipótesis school achievement] held that students from schools with higher school achievement would show stronger preferences for market justice, following the argument that such school could foster meritocratic attitudes by showing a larger promotion of talent and effort in order to get better results.
+Hypothesis $H_4$ held that students from schools with higher school achievement would show stronger preferences for market justice, following the argument that such school could foster meritocratic attitudes by showing a larger promotion of talent and effort in order to get better results.
 Nevertheless, the results show quite the opposite: schools with better academic results (as measured by a national standardized test) show in average less preferences for market justice.
@@
 
-Hypotheses $H_6$ and $H_7$ proposed that family status would moderate the effect of meritocratic perceptions on market justice preferences.
-Hypothesis $H_6$ refers to family socioeconomic status, whereby the results show a negative interaction with two of the three perceptual variables at the society level.
+Hypotheses $H_5$ and $H_6$ proposed that socioeconomic status would moderate the effect of meritocratic perceptions on market justice preferences.
+Hypothesis $H_5$ refers to family socioeconomic status, whereby the results show a negative interaction with two of the three perceptual variables at the society level.
 This means that the positive impact of the perception of social meritocracy on market justice preferences have a lower effect in students with parents with higher education and also in those with more books at home.
-Something similar occurs with hypothesis $H_7$, where now school level status moderates the effect of meritocracy variables, which in multilevel framework si referred as to cross-level interaction.
+Something similar occurs with hypothesis $H_6$, where now school level status moderates the effect of meritocracy variables, which in multilevel framework si referred as to cross-level interaction.
 The negative interaction in this context means that in schools of low socioeconomic levels, the positive effect of the social meritocracy variables on market justice preferences is stronger.
@@
 This underlines the importance of addressing structural inequalities and individual beliefs in educational and social policies.
+
+$H_7$ Falta solo esto

### 2024-06-24 06-conclusion.qmd
Revision Andreas y Render completo
Revision Andreas:
- Falta discussion de H7
- Referencias que faltan agregarlas a Zotero (Kevin)

@@
 In general, we find that those who perceive that effort and talent/intelligence are rewarded, are more willing to agree with that richer individuals can have better health, education, and pensions.
-This is not the first time that meritocratic perceptions (and also beliefs) have been related with the legitimation of social inequalities [@mijs_paradox_2019; @darnon_where_2018], but the consideration of school level population as well as market justice preferences allows adding evidence regarding the role of meritocracy regarding inequality beliefs: legitimation of market criteria in social services seems to begin at school age, and what students think of meritocracy is linked to it.
+This is not the first time that meritocratic perceptions (and also beliefs) have been related with the legitimation of social inequalities (Mijs, 2019; Darnon 2017), but the consideration of school level population as well as market justice preferences allows adding evidence regarding the role of meritocracy regarding inequality beliefs: legitimation of market criteria in social services seems to begin at school age, and what students think of meritocracy is linked to it.
 Although there are still several open questions, as for instance differences between meritocratic perceptions in school and in the society at large, this raises up the relevance of what is done (or not) in schools in order to challenge the meritocratic ideals.
@@
 The consideration of the school context opened some caveats in this research.
-We found contrary evidence for the association between status (as measured by average education), school performance, and market justice prefecences: high status and high achievement at school level is asociated with less market justice preferences.
+We found contrary evidence for the association between status, school performance, and market justice prefecences: high status and high achievement at school level is asociated with less market justice preferences.
 This is opposed to the rational-interest argument that suggests that those doing better - both economically and academically- would demand less redistribution (Gonthier,2016; Runst, 2018; Bullock, 2021) and therefore are expected to challenge market mechanisms.

### 2024-06-25 05-discussion.qmd
Update varios y render

@@
 
-$H_7$ Falta solo esto
+Hypothesis $H7$ posited that the academic performance of schools could moderate the effect of meritocratic perceptions on preferences for market justice.
+This hypothesis is based on the argument that the level of academic performance in schools entails a differentiated promotion of meritocratic perceptions to achieve better results, which would, in turn, reinforce or diminish preferences for market justice.
+The results of the interactions between levels suggest that, indeed, in higher-performing schools, the effect of meritocratic perceptions on the justification of service provision based on market principles is weaker compared to lower-performing schools.
+Similarly to the effect of academic performance in $H4$, a possible explanation for this moderation is that students in high-performing schools may be more exposed to knowledge about the effects of inequality and social justice.
+This exposure translates into an enlightening effect of education on inequality, which can mitigate the impact of meritocracy on market justice preferences.

### 2024-06-28 05-discussion.qmd
revision JC

@@
 
-Hypothesis $H_4$ held that students from schools with higher school achievement would show stronger preferences for market justice, following the argument that such school could foster meritocratic attitudes by showing a larger promotion of talent and effort in order to get better results.
+Hypothesis $H_4$ held that students from schools with higher school achievement would show stronger preferences for market justice, following the argument that such schools could foster meritocratic attitudes by showing a larger promotion of talent and effort in order to get better results.
 Nevertheless, the results show quite the opposite: schools with better academic results (as measured by a national standardized test) show in average less preferences for market justice.
 This finding is one of the most strongest and consistent throughout the models and it is relevant given the centrality of school performance in educational institutions.
-One possible explanation of this link is that students in high-achieving schools may be more exposed to discussions and to a deeper understanding about inequality and social justice, leading to a critical perspective towards the weight of market mechanism in the access to social services.
+One possible explanation of this link is that students in high-achieving schools may be more exposed to discussions and a deeper understanding of inequality and social justice, leading to a critical perspective toward the weight of market mechanisms in the access to social services.
 
 Hypotheses $H_5$ and $H_6$ proposed that socioeconomic status would moderate the effect of meritocratic perceptions on market justice preferences.
-Hypothesis $H_5$ refers to family socioeconomic status, whereby the results show a negative interaction with two of the three perceptual variables at the society level.
-This means that the positive impact of the perception of social meritocracy on market justice preferences have a lower effect in students with parents with higher education and also in those with more books at home.
-Something similar occurs with hypothesis $H_6$, where now school level status moderates the effect of meritocracy variables, which in multilevel framework si referred as to cross-level interaction.
+Hypothesis $H_5$ refers to family socioeconomic status, whereby the results show a negative interaction with two of the three perceptual variables at the societal level.
+This means that the positive impact of the perception of social meritocracy on market justice preferences had a lower effect on students with parents with higher education and also on those with more books at home.
+Something similar occurs with hypothesis $H_6$, where now school-level status moderates the effect of meritocracy variables, which in multilevel framework is referred as to cross-level interaction.
 The negative interaction in this context means that in schools of low socioeconomic levels, the positive effect of the social meritocracy variables on market justice preferences is stronger.

### 2024-06-28 02-antecedentes.qmd

--- 2024-06-19 02-antecedentes.qmd
@@
 The influence of ideologies [@wegener_dominant_1995] and cultural schemas [@homan_being_2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations.
-While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999 ; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children's acceptance or rejection of these market principles.
+While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999 ; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations [@lindh_bringing_2023] and even less about children's acceptance or rejection of these market principles.
 This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades [@centeno_arc_2012; @harvey_breve_2015], affecting areas such as pensions, health services, and education.
@@
 Lane proposed the underpinnings of the concept of market justice, which he differentiated from political justice.
-For him, "it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets" [-@lane_market_1986; p.
+For him, "it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets" [-@lane_market_1986, p.
 384].
-Contrary to the evidence that unequal distribution produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action [@greitemeyer_subjective_2016; @mishra_subjective_2015; @smith_relative_2012; @power_deprivationprotest_2018; @power_relative_2020], Lane pointed out that in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions.
+Contrary to the evidence that unequal distribution produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action [@greitemeyer_subjective_2016; @mishra_subjective_2015; @smith_relative_2012; @power_deprivationprotest_2018], Lane pointed out that in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions.
 In this sense, unequal levels of well-being would be, to some extent, a function of their talents and efforts, instead of being based on distributive principles that characterize welfare states, such as need and equality (see Wilson 2003).
@@
 According to this data, 35.9% prefer private health insurance, and 63% would prefer private education.
-Yet, research on children's justification in this area remains limited, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.
+Yet, research on children's justification in this area remains limited, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs (Centro de Estudios Públicos, 2024).
 
@@
 
-$H_6$: The school's performance in standardized achievement tests will moderate the effect of meritocratic perceptions on market justice preferences.
+$H_6$: The school's status will moderate the effect of meritocratic perceptions on market justice preferences.
+
+$H_7$: The school's performance in standardized achievement tests will moderate the effect of meritocratic perceptions on market justice preferences.
 

### 2024-06-28 04-analisis.qmd

@@
   kable_styling(latex_options = c("scale_down","HOLD_position"),
-                full_width = F) %>% 
+                full_width = T) %>% 
   column_spec(1, bold = T, width = "4cm") %>%

### 2024-06-28 06-conclusion.qmd

@@
 The consideration of the school context opened some caveats in this research.
-We found contrary evidence for the association between status, school performance, and market justice prefecences: high status and high achievement at school level is asociated with less market justice preferences.
+We found contrary evidence for the association between status, school performance, and market justice preferences: high status and high achievement at the school level are associated with less market justice preferences.
 This is opposed to the rational-interest argument that suggests that those doing better - both economically and academically- would demand less redistribution (Gonthier,2016; Runst, 2018; Bullock, 2021) and therefore are expected to challenge market mechanisms.
-A possible explanation is that education contributes to form a crirical view of markets in redistribution, nevertheless further reserach in this area is needed in order to assert the specific mechanisms that might be playing a role.
-All in all, from this study we know that school contexts are not trivial to this regard, calling for larger attention to school age research on the formation of inequality preferences and beliefs.
+A possible explanation is that education contributes to forming a critical view of markets in redistribution, nevertheless, further research in this area is needed to assert the specific mechanisms that might be playing a role.
+All in all, from this study we know that school contexts are not trivial in this regard, calling for larger attention to school-age research on the formation of inequality preferences and beliefs.

### 2024-06-28 06-conclusion.qmd

@@
 In general, we find that those who perceive that effort and talent/intelligence are rewarded, are more willing to agree with that richer individuals can have better health, education, and pensions.
-This is not the first time that meritocratic perceptions (and also beliefs) have been related with the legitimation of social inequalities (Mijs, 2019; Darnon 2017), but the consideration of school level population as well as market justice preferences allows adding evidence regarding the role of meritocracy regarding inequality beliefs: legitimation of market criteria in social services seems to begin at school age, and what students think of meritocracy is linked to it.
+This is not the first time that meritocratic perceptions (and also beliefs) have been related with the legitimation of social inequalities [@mijs_paradox_2019; @darnon_where_2018], but the consideration of school level population as well as market justice preferences allows adding evidence regarding the role of meritocracy regarding inequality beliefs: legitimation of market criteria in social services seems to begin at school age, and what students think of meritocracy is linked to it.
 Although there are still several open questions, as for instance differences between meritocratic perceptions in school and in the society at large, this raises up the relevance of what is done (or not) in schools in order to challenge the meritocratic ideals.

### 2024-06-28 03-methods.qmd
add processing docs and fixed references

@@
 The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open Science (OSF), the access to the document is available at this [link](https://doi.org/10.17605/OSF.IO/UFSDV).
-The statistical analysis of this research was performed using the free software R version 4.1.3.
+The statistical analysis for this research was conducted using the *lme4* package in R version 4.1.3.

### 2024-06-28 processing/0-patterns.Rmd

---
title: "0-patterns"
author: "Equipo EDUMER"
date: "2022-08-07"
output: html_document
---

# Identificar patrones de respuesta en R

base de datos: primer estudio de formacion ciudadana en chile 2017 (base alumnos)

```{r}
pacman::p_load(dplyr, responsePatterns, haven, sjmisc)

load("input/data/original/base_estudiantes.RData")
alumnos <- base_estudiantes %>% select(-codigoCurso, -serie, -mrbd, -numerolista, -sexo, -puntaje, -ponderador, -mrun, -agno, -cod_com_alu, -nom_com_alu, -prom_gral, -rbdi, -cod_depe1, -cod_depe2, -cod_grupo, -prom_lect8b_rbd, -prom_mate8b_rbd) #eliminé variable id curso y "serie". 
#La idea es solo mantener una variable id y el resto de variables que posean un rango 
#similar de valores de respuesta posibles
```

* El paquete estima las autocorrelaciones de todas las variables, para cada caso.

"Para cada encuestado, se calculan los coeficientes de autocorrelación para todos los valores razonables de lag (k) y se mantiene el coeficiente de autocorrelación absoluto más alto. El valor resultante debe indicar el nivel máximo absoluto de asociación en las respuestas del encuestado en función del orden de respuesta, y su respectivo k debe indicar la longitud del patrón de respuesta que se repite" (Gottfried, Jezek & Kralova, 2021)

```{r}
# calcula las auto correlaciones
rp <- responsePatterns::rp.acors(alumnos, max.lag = 10, id.var = "idalumno") 
```

* El paquete permite ver un histograma de la distribución de las autocorrelaciones más altas de cada uno de los casos según su patrón (k)

```{r}
rp.hist(rp)
```

* Para una mejor visualización de los resultados se eliminan NA 0 y 99 y se estima de nuevo

```{r}
alumnos[2:220] <- alumnos[2:220] %>%  set_na(., na = c(0, 99), drop.levels = FALSE, as.tag = FALSE)
rp2 <- responsePatterns::rp.acors(alumnos, max.lag = 10, id.var = "idalumno") 
```

* El reporte del paquete entrega los índices importantes (la autocorrelacion mas alta y el patron al que pertenece). Esta funcion no entrega los valores de autocorrelacion para cada k, solo el mas alto

```{r}
indices <- rp.indices(rp, round = 2, include.coefs = F)

head(indices, 10)
```

* Como entrega una distribucion en percentiles, es posible enfocarnos solo en los más altos

```{r}
remove <- dplyr::filter(indices, percentile==0 | percentile>=99)
dim(remove) # hay 430 casos con una autocorrelacion alta

head(remove, 10)
```

* Finalmente, l-s autores recomiendan revisar cada caso por separado (en lo posible) antes de determinar qué casos excluir. Esto se puede hacer con la siguiente función

```{r}
rp.plot(rp2, rowname = 31)
```

```{r}
base <- indices %>% select(percentile)
load("input/data/original/estudiantes_labelled.RData")
base_patt <- cbind(base_estudiantes, base)

alumnos_patterns <- dplyr::filter(base_patt, percentile!=0 & percentile<99) %>% select(-percentile)
save(alumnos_patterns, file = "input/data/proc/alumnos_patterns.RData")
```

# Apoderados

se repiten los mismos pasos, pero para la base de apoderados

```{r}
load("input/data/original/base_apoderados.RData")
apoderados <- base_apoderados %>% select(-codigoCurso, -serie, -apod_p1_1, -apod_p1_2)
```

```{r}
# calcula las auto correlaciones
rp3 <- responsePatterns::rp.acors(apoderados, max.lag = 10, id.var = "idalumno")
```

```{r}
rp.hist(rp3)
```

```{r}
apoderados[2:137] <- apoderados[2:137] %>%  set_na(., na = c(0, 99), drop.levels = FALSE, as.tag = FALSE)
rp4 <- responsePatterns::rp.acors(apoderados, max.lag = 10, id.var = "idalumno") 
```

```{r}
indices2 <- rp.indices(rp3, round = 2, include.coefs = F)

head(indices2, 10)
```

```{r}
remove2 <- dplyr::filter(indices2, percentile==0 | percentile>=99)
dim(remove2) # hay 430 casos con una autocorrelacion alta

head(remove2, 10)
```

```{r}
rp.plot(rp4, rowname = 147)
```

```{r}
base2 <- indices2 %>% select(percentile)
load("input/data/original/apoderados_labelled.RData")
base_patt2 <- cbind(base_apoderados, base2)

apoderados_patterns <- dplyr::filter(base_patt2, percentile!=0 & percentile<99) %>% select(-percentile)
save(apoderados_patterns, file = "input/data/proc/apoderados_patterns.RData")
```

### 2024-06-28 processing/01-descriptivos-meritocracia.Rmd

---
title: "descriptivos-meritocracia"
author: "Equipo EDUMER"
date: "29-11-2021"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    code_folding: hide
    toc_float:
      collapsed: true
      smooth_scroll: false
      number_sections: true
---

```{r}
pacman::p_load(haven, dplyr, summarytools, sjmisc, car, sjlabelled, webshot, sjPlot, ggplot2, corrplot, ltm)
load("input/data/proc/alumnos_patterns.RData")
load("input/data/proc/apoderados_patterns.RData")
```

```{r}
data_est <- alumnos_patterns %>% dplyr::select(idalumno,
                                               sexo,
                                               codigoCurso,
                                               mrbd,
                                               cod_depe2, #dependencia administrativa
                                               cod_grupo, #categorizacion socioeconomica
                                               simce_lect=prom_lect8b_rbd, #Promedio lectura escuela
                                               simce_mate=prom_mate8b_rbd, # Promedio matematica escuela
                                               inteligencia_esc=est_p6_1, # inteligencia escolar
                                               esfuerzo_esc=est_p6_2, # esfuerzo escolar
                                               esfuerzo_soc=est_p38_2, #esfuerzo social
                                               merito_soc=est_p38_3, #merito social
                                               inteligencia_soc=est_p38_4, #inteligencia social
                                               just_pension=est_p39_1, #justicia pension
                                               just_educ=est_p39_2, # justicia educacion
                                               just_salud=est_p39_3, # justicia salud
                                               redistribucion=est_p39_4, # redistribucion
                                               libros=est_p3, # cantidad de libros
                                               pc=est_p4_1, # tiene pc
                                               tablet=est_p4_2, # tiene tablet
                                               celular=est_p4_3 # tiene celular
                                               ) %>% as.data.frame(.)

data_apod <- apoderados_patterns %>% dplyr::select(idalumno,
                                               educacion=educ_max, # nivel educacional mas alto
                                               internet=apod_p9 # conexion a internet en la casa
                                               ) %>% as.data.frame(.)
```

## merge bases
```{r}
data <- merge(data_est, data_apod, by="idalumno")
rm(apoderados_patterns, alumnos_patterns, data_est, data_apod)
```

```{r}
# Educacion. 10 categorías, se recodifican en 4 y una de no sabe y no responde
data$educacion_rec <- ifelse(is.na(data$educacion), "Ns/Nr", data$educacion)
data <- data %>% rowwise() %>%  mutate(educacion_rec = case_when(educacion_rec==1~"8th grade or less",
                                                 educacion_rec==2~"8th grade or less",
                                                 educacion_rec==3~"8th grade or less",
                                                 educacion_rec==4~"Secondary Education",
                                                 educacion_rec==5~"Secondary Education",
                                                 educacion_rec==6~"Higher tec. education",
                                                 educacion_rec==7~"Higher tec. education",
                                                 educacion_rec==8~"University or Postgraduate",
                                                 educacion_rec==9~"University or Postgraduate",
                                                 educacion_rec==10~"University or Postgraduate",
                                                 educacion_rec=="Ns/Nr"~"Missing"
                                                 ))
data$educacion_rec <- factor(data$educacion_rec, levels = c("8th grade or less", "Secondary Education", "Higher tec. education", "University or Postgraduate", "Missing"))
data$educacion_rec <- set_label(data$educacion_rec,label = "Parental educational level")

## Proporcion de Universidad o posgrado por escuela
data <- data %>% group_by(mrbd) %>%
  mutate(mean_educ = mean(educacion_rec=="University or Postgraduate", na.rm=T))
data$mean_educ <- set_label(data$mean_educ,label = "Proportion of parents with university level by school")

# Internet
data <- data %>% set_na(internet, na = c(0,99), drop.levels = FALSE, as.tag = FALSE) # recode missings de doble marca (99) y vacío (0) a NA

data$internet_rec <- ifelse(is.na(data$internet), "Missing", data$internet)
data$internet_rec <- factor(data$internet_rec, levels = c("1", "2"), labels = c("Yes", "No"))
# libros
data$libros_rec <- as.numeric(data$libros)
data <- data %>% rowwise() %>% mutate(libros_rec = case_when(libros_rec==1 ~ "Les than 25",
                                                             libros_rec==2 ~ "Les than 25",
                                                             libros_rec==3 ~ "More than 25",
                                                             libros_rec==4 ~ "More than 25",
                                                             libros_rec==5 ~ "More than 25"))
data$libros_rec <- factor(data$libros_rec, levels = c("Les than 25", "More than 25"))
data$libros_rec <- set_label(data$libros_rec,label = "Number of books at home")
## Promedio simce
data <- data %>% rowwise() %>% mutate(prom_simce = mean(simce_lect,simce_mate))
summary(data$prom_simce)
data$simce <- ntile(data$prom_simce,3)
table(data$simce)
data$simce <- factor(data$simce, labels = c("Low", "Medium", "High"))
data$simce <- set_label(data$simce,label = "SIMCE score by school")
```

```{r}
# Acceso a tecnologia
data$pc <- as.numeric(data$pc)
data$tablet <- as.numeric(data$tablet)
data$celular <- as.numeric(data$celular)

corrplot.mixed(cor(dplyr::select(data, pc, tablet, celular),
                   method = "spearman",
                   use = "complete.obs"))
```

```{r}
ltm::cronbach.alpha(data  %>%
  dplyr::select(pc, tablet, celular), na.rm=TRUE)
```

```{r}
data <- data %>% rowwise() %>% dplyr::mutate(acc_tec = sum(pc, tablet, celular, na.rm = T))
summary(data$acc_tec)
data$acc_tec <- set_label(data$acc_tec,label = "Technology access index")
```

# Dependientes
```{r}
dependientes <- data %>% dplyr::select(just_pension, just_educ, just_salud)

dependientes$pension_rec <- as.numeric(dependientes$just_pension)
dependientes$educ_rec <- as.numeric(dependientes$just_salud)
dependientes$salud_rec <- as.numeric(dependientes$just_educ)
#dependientes <- dependientes %>% dplyr::mutate(pension_rec = 5-pension_rec,
#                                            educ_rec = 5-educ_rec,
#                                            salud_rec = 5-salud_rec)

ltm::cronbach.alpha(dependientes  %>%
  dplyr::select(educ_rec, salud_rec, pension_rec), na.rm=TRUE)

dependientes <- dependientes %>% rowwise() %>% dplyr::mutate(bienestar = mean(c(educ_rec, salud_rec, pension_rec), na.rm=TRUE))
dependientes$bienestar <- round(dependientes$bienestar, 2)
dependientes <- dependientes %>% dplyr::select(-pension_rec, -educ_rec, -salud_rec)

dependientes$just_pension <- factor(dependientes$just_pension, levels = c("Muy desacuerdo", "Desacuerdo", "De acuerdo", "Muy de acuerdo"), labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"))
dependientes$just_pension <- set_label(dependientes$just_pension, label = "It is just that in Chile people with higher incomes can have better pensions than people with low incomes")

dependientes$just_salud <- factor(dependientes$just_salud, levels = c("Muy desacuerdo", "Desacuerdo", "De acuerdo", "Muy de acuerdo"), labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"))
dependientes$just_salud <- set_label(dependientes$just_salud,label = "It is just that in Chile people with higher incomes can access better health services than people with low incomes")

dependientes$just_educ <- factor(dependientes$just_educ, levels = c("Muy desacuerdo", "Desacuerdo", "De acuerdo", "Muy de acuerdo"), labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"))
dependientes$just_educ <- set_label(dependientes$just_educ,label = "It is just that in Chile people who can pay have a better education for their children")

dependientes$bienestar <- set_label(dependientes$bienestar,label = "Market Justice Preferences")

#dependientes$redistribucion <- factor(dependientes$redistribucion, levels = c("Muy desacuerdo", "Desacuerdo", "De acuerdo", "Muy de acuerdo"), labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"))
#dependientes$redistribucion <- set_label(dependientes$redistribucion,label = "The Chilean government should do something to reduce the income gap between the rich and the poor.")
```

```{r, results='asis'}
df<-dfSummary(dependientes,
               plain.ascii = TRUE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(30,10,10,10))

df$Variable <- NULL # delete variable column
df
view(df, footnote=NA, file = "output/tables/desc01.html")
webshot::webshot(url ="output/tables/desc01.html" ,file ="output/tables/desc01.png")
```

# independientes
```{r}
independientes <- data %>% dplyr::select(inteligencia_esc, esfuerzo_esc, inteligencia_soc, esfuerzo_soc, merito_soc,  educacion_rec, libros_rec, acc_tec)

independientes$inteligencia_esc <- factor(independientes$inteligencia_esc, levels = c("Muy desacuerdo", "Desacuerdo", "De acuerdo", "Muy de acuerdo"), labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"))
independientes$inteligencia_esc <- set_label(independientes$inteligencia_esc, label = "Intelligence is important to get good grades")

independientes$esfuerzo_esc <- factor(independientes$esfuerzo_esc, levels = c("Muy desacuerdo", "Desacuerdo", "De acuerdo", "Muy de acuerdo"), labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"))
independientes$esfuerzo_esc <- set_label(independientes$esfuerzo_esc, label = "Effort is important to get good grades")

independientes$esfuerzo_soc <- factor(independientes$esfuerzo_soc, levels = c("Muy desacuerdo", "Desacuerdo", "De acuerdo", "Muy de acuerdo"), labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"))
independientes$esfuerzo_soc <- set_label(independientes$esfuerzo_soc, label = "In Chile, people are rewarded for their efforts")

independientes$merito_soc <- factor(independientes$merito_soc, levels = c("Muy desacuerdo", "Desacuerdo", "De acuerdo", "Muy de acuerdo"), labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"))
independientes$merito_soc <- set_label(independientes$merito_soc, label = "In Chile, people get what they deserve")

independientes$inteligencia_soc <- factor(independientes$inteligencia_soc, levels = c("Muy desacuerdo", "Desacuerdo", "De acuerdo", "Muy de acuerdo"), labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"))
independientes$inteligencia_soc <- set_label(independientes$inteligencia_soc, label = "In Chile, people are rewarded for their intelligence and skill")

df2<- dfSummary(independientes,
               plain.ascii = FALSE,
               style = "grid",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(20,30,10,10,10))

df2$Variable <- c("School talent", "School effort", "Social talent", "Social effort", "Deservingness", "Education", "Books", "Technology")
df2
view(df2, footnote=NA, file = "output/tables/desc02.html")
webshot(url ="output/tables/desc02.html" ,file ="output/tables/desc02.png")
```

# School

```{r}
school <- data %>%
  group_by(mrbd) %>%
  summarise(
    mean_educ = mean(mean_educ, na.rm = TRUE),
    simce = first(simce),
    cod_depe2 = first(cod_depe2),
    cod_grupo = first(cod_grupo)
  )

school$cod_depe2  <- factor(school$cod_depe2, levels = c("Municipal", "Part. subvencionado", "Part. privado"), labels = c("Public", "Subsidized private", "Private"))
school$cod_depe2 <- set_label(school$cod_depe2, label = "Administrative dependency of school")

school$cod_grupo  <- factor(school$cod_grupo, levels = c("Bajo", "Medio bajo", "Medio", "Medio alto", "Alto"), labels = c("Low", "Medium low", "Medium", "Medium high", "High"))
school$cod_grupo <- set_label(school$cod_grupo, label = "Socioeconomic level of school")

school$mean_educ <- set_label(school$mean_educ, label = "Proportion of parents with university level by school")

school$simce <- set_label(school$simce, label = "SIMCE score achievement by school")

school <- school %>% dplyr::select(-mrbd)

save(school, file="input/data/school.RData")

df3<- dfSummary(school,
               plain.ascii = FALSE,
               style = "grid",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(10,30,10,10,10))
df3$Variable <- c("Prop. university level
at school",
                  "Achievement", "Dependency", "Socioeconomic level")
view(df3, footnote=NA, file = "output/tables/desc03.html")
webshot(url ="output/tables/desc03.html" ,file ="output/tables/desc03.png")
```

# graficos descriptivos

## dependientes
```{r}
dependientes <- sjPlot::plot_stackfrq(dplyr::select(data, just_educ, just_pension, just_salud), geom.size = 0.4) +
  theme_bw() +
  theme(legend.position="bottom")
dependientes
ggsave(dependientes, file = "output/graphs/dependientes.png",device = "png",width = 25,height = 13,dpi = "retina",units = "cm")
```

## independientes

```{r}
independientes <- sjPlot::plot_stackfrq(dplyr::select(data, inteligencia_esc, esfuerzo_esc, inteligencia_soc, merito_soc, esfuerzo_soc)) +
  theme_bw() +
  theme(legend.position="bottom")

ggsave(independientes, file = "output/graphs/independientes.png",device = "png",width = 25,height = 13,dpi = "retina",units = "cm")
```

### 2024-06-28 processing/02-analisis-multinivel.Rmd

---
title: "analisis-multinivel"
author: "Equipo EDUMER"
date: '2022-06-13'
output: 
  html_document: 
    toc: yes
    code_folding: hide
    toc_float: 
      collapsed: true
      smooth_scroll: false
      number_sections: true
---

```{r}
pacman::p_load(haven, dplyr, summarytools, sjmisc, car, sjlabelled, sjPlot, lme4, stargazer, corrplot, ltm, texreg, ordinal, MASS, webshot, ggplot2, r2mlm)
load("input/data/proc/alumnos_patterns.RData")
load("input/data/proc/apoderados_patterns.RData")
```

```{r}
# Seleccionar variables de interes
data_est <- alumnos_patterns %>% dplyr::select(idalumno,
                                               sexo,
                                               codigoCurso,
                                               mrbd,
                                               cod_depe2, #dependencia administrativa
                                               cod_grupo, #categorizacion socioeconomica
                                               simce_lect=prom_lect8b_rbd, #Promedio lectura escuela
                                               simce_mate=prom_mate8b_rbd, # Promedio matematica escuela
                                               inteligencia_esc=est_p6_1, # inteligencia escolar
                                               esfuerzo_esc=est_p6_2, # esfuerzo escolar
                                               esfuerzo_soc=est_p38_2, #esfuerzo social
                                               merito_soc=est_p38_3, #merito social
                                               inteligencia_soc=est_p38_4, #inteligencia social
                                               just_pension=est_p39_1, #justicia pension
                                               just_educ=est_p39_2, # justicia educacion
                                               just_salud=est_p39_3, # justicia salud
                                               redistribucion=est_p39_4, # redistribucion
                                               libros=est_p3, # cantidad de libros
                                               pc=est_p4_1, # tiene pc
                                               tablet=est_p4_2, # tiene tablet
                                               celular=est_p4_3 # tiene celular
                                               ) %>% as.data.frame(.)

data_apod <- apoderados_patterns %>% dplyr::select(idalumno,
                                               educacion=educ_max, # nivel educacional mas alto
                                               internet=apod_p9 # conexion a internet en la casa
                                               ) %>% as.data.frame(.)

data_apod <- data_apod %>%  set_na(internet, na = c(0,99), drop.levels = FALSE, as.tag = FALSE) # recode missings de doble marca (99) y vacío (0) a NA
```

## merge bases
```{r}
data <- merge(data_est, data_apod, by="idalumno")
rm(apoderados_patterns, alumnos_patterns, data_est, data_apod)
```

```{r}
# Educacion. 10 categorías, se recodifican en 4 y una de no sabe y no responde
data$educacion_rec <- ifelse(is.na(data$educacion), "Ns/Nr", data$educacion)
data <- data %>% rowwise() %>%  mutate(educacion_rec = case_when(educacion_rec==1~"Ed Basica",
                                                 educacion_rec==2~"Ed Basica",
                                                 educacion_rec==3~"Ed Basica",
                                                 educacion_rec==4~"Ed Media",
                                                 educacion_rec==5~"Ed Media",
                                                 educacion_rec==6~"Ed Tecnica",
                                                 educacion_rec==7~"Ed Tecnica",
                                                 educacion_rec==8~"Universidad o posgrado",
                                                 educacion_rec==9~"Universidad o posgrado",
                                                 educacion_rec==10~"Universidad o posgrado",
                                                 educacion_rec=="Ns/Nr"~"Ns/Nr"
                                                 ))
data$educacion_rec <- factor(data$educacion_rec, levels = c("Ed Basica", "Ed Media", "Ed Tecnica", "Universidad o posgrado", "Ns/Nr"))

## Proporcion de Universidad o posgrado por escuela
data <- data %>% group_by(mrbd) %>%
  mutate(mean_educ = mean(educacion_rec=="Universidad o posgrado", na.rm=T))

# Internet
data$internet <- factor(data$internet, labels = c("Si", "No"))
data$internet_rec <- ifelse(is.na(data$internet), "Ns/Nr", data$internet)

# libros
data$libros_rec <- as.numeric(data$libros)
data <- data %>% rowwise() %>% mutate(libros_rec = case_when(libros_rec==1 ~ "Menos de 25",
                                                             libros_rec==2 ~ "Menos de 25",
                                                             libros_rec==3 ~ "Más de 25",
                                                             libros_rec==4 ~ "Más de 25",
                                                             libros_rec==5 ~ "Más de 25"))
data$libros_rec <- factor(data$libros_rec, levels = c("Menos de 25", "Más de 25"))

## Promedio simce
data <- data %>% rowwise() %>% mutate(prom_simce = mean(simce_lect,simce_mate))
summary(data$prom_simce)
data$simce <- ntile(data$prom_simce,3)
table(data$simce)
data$simce <- factor(data$simce, labels = c("Bajo", "Medio", "Alto"))
```

```{r}
# Acceso a tecnologia
data$pc <- as.numeric(data$pc)
data$tablet <- as.numeric(data$tablet)
data$celular <- as.numeric(data$celular)

corrplot.mixed(cor(dplyr::select(data, pc, tablet, celular),
                   method = "spearman",
                   use = "complete.obs"))
```

```{r}
ltm::cronbach.alpha(data  %>%
  dplyr::select(pc, tablet, celular), na.rm=TRUE)
```

```{r}
data <- data %>% rowwise() %>% dplyr::mutate(acc_tec = sum(pc, tablet, celular, na.rm = T))
summary(data$acc_tec)
```

```{r}
# Variables meritocracia
data$inteligencia_esc <- as.numeric(data$inteligencia_esc)
data$esfuerzo_esc <- as.numeric(data$esfuerzo_esc)
data$esfuerzo_soc <- as.numeric(data$esfuerzo_soc)
data$merito_soc <- as.numeric(data$merito_soc)
data$inteligencia_soc <- as.numeric(data$inteligencia_soc)
data$just_pension <- as.numeric(data$just_pension)
data$just_salud <- as.numeric(data$just_salud)
data$just_educ <- as.numeric(data$just_educ)
data$redistribucion <- as.numeric(data$redistribucion)
```

```{r}
data_rec <- data %>% dplyr::select(-educacion, -internet, -simce_lect, -simce_mate, -libros, -pc, -tablet, -celular)

save(data_rec, file = "input/data/proc/data_rec.RData")
```

# Modelos de regresión multinivel (lmer)

## indice variables acceso a bienestar social

```{r}
data_bienestar <- na.omit(data_rec)

table(data_bienestar$just_pension)

corrplot.mixed(cor(dplyr::select(data_bienestar, just_pension, just_educ, just_salud, redistribucion),
                   method = "spearman",
                   use = "complete.obs"))
```

```{r}
ltm::cronbach.alpha(data_bienestar  %>%
  dplyr::select(just_educ, just_salud, just_pension), na.rm=TRUE)
```

```{r}
data_bienestar <- data_bienestar %>% rowwise() %>% dplyr::mutate(bienestar = mean(c(just_pension, just_educ, just_salud), na.omit=TRUE))
data_bienestar$bienestar <- round(data_bienestar$bienestar, 2)
summary(data_bienestar$bienestar)
```

## Intra class correlation

```{r}
reg0_bienestar <- lmer(bienestar ~ 1 + (1 | mrbd), data=data_bienestar)
performance::icc(reg0_bienestar)
```

## Regresiones multinivel

```{r}
reg1_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_bienestar)
reg2_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_bienestar)

reg3_bienestar <- lmer(bienestar ~ 1 + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_bienestar)
reg4_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_bienestar)
```

### Test deviance

```{r}
test01<- anova(reg0_bienestar,reg1_bienestar,test = "Chisq")
test02<- anova(reg0_bienestar,reg2_bienestar,test = "Chisq")
test03<- anova(reg0_bienestar,reg3_bienestar,test = "Chisq")
test04<- anova(reg0_bienestar,reg4_bienestar,test = "Chisq")

test.pvalues1<- test01$`Pr(>Chisq)`[2]
test.pvalues2<- test02$`Pr(>Chisq)`[2]
test.pvalues3<- test03$`Pr(>Chisq)`[2]
test.pvalues4<- test04$`Pr(>Chisq)`[2]

deviance1 <- test01$deviance[2]
deviance2 <- test02$deviance[2]
deviance3 <- test03$deviance[2]
deviance4 <- test04$deviance[2]
```

### Tabla de regresión

```{r results='asis'}
htmlreg(list(reg1_bienestar, reg2_bienestar, reg3_bienestar, reg4_bienestar),
        custom.model.names = c("Modelo 1",
                               "Modelo 2",
                               "Modelo 3",
                               "Modelo 4",
                               "Modelo 5"), 
        custom.note = "$***p <$ 0.001, $**p <$ 0.01, $*p <$ 0.05", 
        omit.coef = "(internet_rec)",
        custom.coef.names = c("Intercepto", 
                              "Esfuerzo social", 
                              "Merito social", 
                              "Inteligencia social", 
                              "Esfuerzo escolar", 
                              "Inteligencia escolar", 
                              "Ed media <br> <i>(Ref. Ed Basica)</i>", 
                              "Ed tecnica", "Universidad o posgrado", 
                              "Ns/Nr", 
                              "Más de 25 libros <br> <i>(Ref. Menos de 25)</i>", 
                              "Acceso a tecnología",
                              "Prop. universitarios escuela",
                              "Part. subv <br> <i>(Ref. Municipal)</i>", 
                              "Part. priv", 
                              "SES Medio bajo <br> <i>(Ref. Bajo)</i>", 
                              "SES Medio", 
                              "SES Medio alto", 
                              "SES Alto", 
                              "Simce_Medio <i>(Ref. Bajo)</i>", 
                              "Simce_alto")
                             )
```

### Individual effects

```{r results='asis'}
## ojo aquí, creo que para los p*** sin exportar en html se puede usar "$^{***}$ p < 0.001; $^{**}$ p < 0.01; $^{*}$ p < 0.05

htmlreg(list(reg1_bienestar, reg2_bienestar),
        custom.model.names = c("Model 1",
                               "Model 2"),
        doctype = FALSE,
        custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
        omit.coef = "(internet_rec)",
        custom.coef.names = c("Intercept",  
                              "School talent",
                              "School effort", 
                              "Social talent",  
                              "Social effort",
                              "Deservingness",
                              "Secondary <br> <i>(Ref. 8th grade or less)</i>", 
                              "Higher tec.", 
                              "University or posgraduate", 
                              "Missing", 
                              "More than 25 books <br> <i>(Ref. Less than 25)</i>", 
                              "Technology access"),
        caption = "Justification of Inequality",
        caption.above = TRUE,
        custom.gof.rows=list("Deviance" = c(deviance1,
                                            deviance2),
                             "Deviance Test (p)" = c(test.pvalues1,
                                                       test.pvalues2)),
        file = "output/tables/reg_acceso_individual.html")

webshot2::webshot(url ="output/tables/reg_acceso_individual.html" ,file ="output/tables/reg_acceso_individual.png")
```

### Contextual effects

```{r results='asis'}
## ojo aquí, creo que para los p*** sin exportar en html se puede usar "$^{***}$ p < 0.001; $^{**}$ p < 0.01; $^{*}$ p < 0.05

htmlreg(list(reg3_bienestar, reg4_bienestar),
        custom.model.names = c("Model 1",
                               "Model 2"),
        doctype = FALSE,
        custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
        omit.coef = c("inteligencia_esc|esfuerzo_esc|inteligencia_soc|esfuerzo_soc|merito_soc|educacion_rec|libros_rec|acc_tec|internet_rec"),
        custom.coef.names = c("Intercept", 
                              "Prop. university level at school",
                              "Subsidized private <br> <i>(Ref. Public)</i>", 
                              "Private", 
                              "SES Medium low <br> <i>(Ref. Low)</i>", 
                              "SES Medium", 
                              "SES Medium high", 
                              "SES High", 
                              "Simce Medium <i>(Ref. Low)</i>", 
                              "Simce High"),
        caption = "Justification of Inequality",
        caption.above = TRUE,
        custom.gof.rows=list("Deviance" = c(deviance3,
                                            deviance4),
                             "Deviance Test (p)" = c(test.pvalues3,
                                                     test.pvalues4)),
        file = "output/tables/reg_acceso_contextual.html")

webshot2::webshot(url ="output/tables/reg_acceso_contextual.html" ,file ="output/tables/reg_acceso_contextual.png")
```

```{r}
labels <- c("Esfuerzo social", 
                              "Merito social", 
                              "Inteligencia social", 
                              "Esfuerzo escolar", 
                              "Inteligencia escolar", 
                              "Ed media", 
                              "Ed tecnica", "Universidad o posgrado", 
                              "Ns/Nr", 
                              "Más de 25 libros", 
                              "Acceso a tecnología",
                              "Internet",
                              "NS/NR",
                              "Prop. universitarios escuela",
                              "Part. subv", 
                              "Part. priv", 
                              "SES Medio bajo", 
                              "SES Medio", 
                              "SES Medio alto", 
                              "SES Alto", 
                              "Simce_Medio", 
                              "Simce_alto"
                              )
labels2 <- rev(labels)
plot1 <- plot_model(reg5_bienestar, 
                    title = "",
                    axis.labels = labels2,
                    show.values = TRUE, 
                    value.offset = .4)
plot1
ggsave(plot1, file="output/graphs/plot1.png")
```

```{r}
easystats::model_dashboard(reg5_bienestar, output_file = "dashboard_bienestar.html", output_dir = "output/tables")
```

## R2

```{r}
mitml::multilevelR2(reg5_bienestar)

```

### 2024-06-28 processing/07-interacciones.Rmd

---
title: "05-interacciones"
author: "Equipo EDUMER"
date: "2022-08-24"
output: html_document
---

## indice variables acceso a bienestar social

```{r}
pacman::p_load(haven, dplyr, summarytools, sjmisc, car, sjlabelled, sjPlot, lme4, stargazer, corrplot, ltm, texreg, ordinal, MASS, webshot, gtools, kableExtra, ggplot2)
load("input/data/proc/data_rec.RData")

data_rec <- data_rec %>% rowwise() %>%  mutate(educacion_rec2 = case_when(educacion_rec=="Ed Basica"~"Low level",
                                                 educacion_rec=="Ed Media"~"Low level",
                                                 educacion_rec=="Ed Tecnica"~"Low level",
                                                 educacion_rec=="Universidad o posgrado"~"University or Postgraduate",
                                                 educacion_rec=="Ns/Nr"~"Missing"
                                                 ))
data_rec$educacion_rec2 <- factor(data_rec$educacion_rec2, levels = c("Low level", "University or Postgraduate", "Missing"))

data_rec <- data_rec %>% rowwise() %>%  mutate(cod_grupo2  = case_when(cod_grupo=="Bajo"~"Bajo",
                                                                       cod_grupo=="Medio bajo"~"Bajo",
                                                                       cod_grupo=="Medio"~"Bajo",
                                                                       cod_grupo=="Medio alto"~"Alto",
                                                                       cod_grupo=="Alto"~"Alto"))
data_rec$cod_grupo2 <- factor(data_rec$cod_grupo2, levels = c("Bajo", "Alto"))
data_rec$simce <- factor(data_rec$simce, labels = c("Low", "Medium", "High"))

data_bienestar <- na.omit(data_rec)
#data_bienestar <- data_bienestar %>% mutate(pension_rec = 5-just_pension,
#                                            educ_rec = 5-just_educ,
#                                            salud_rec = 5-just_salud)
data_bienestar <- data_bienestar %>% rowwise() %>% dplyr::mutate(bienestar = mean(c(just_educ, just_salud, just_pension), na.omit=TRUE))
data_bienestar$bienestar <- round(data_bienestar$bienestar, 2)

reg10.1_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 | mrbd), data=data_bienestar)
```

*meritocracia x ses familiar (educacion)*

```{r}
reg10_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2*esfuerzo_soc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)

reg11_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2*merito_soc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)

reg12_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2*inteligencia_soc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)

reg13_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2*esfuerzo_esc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)

reg14_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2*inteligencia_esc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)
```

```{r results='asis'}
htmlreg(list(reg10.1_bienestar, reg10_bienestar, reg11_bienestar, reg12_bienestar, reg13_bienestar, reg14_bienestar))
```

```{r}
plot_model(reg11_bienestar, type = "int", mdrt.values="meansd")
```

```{r}
plot_model(reg12_bienestar, type = "int", mdrt.values="meansd")
```

*meritocracia x ses familiar (libros)*

```{r}
reg15_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec*esfuerzo_soc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
reg16_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec*merito_soc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
reg17_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec*inteligencia_soc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
reg18_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2 + libros_rec*esfuerzo_esc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
reg19_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2 + libros_rec*inteligencia_esc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
```

```{r results='asis'}
htmlreg(list(reg10.1_bienestar, reg15_bienestar, reg16_bienestar, reg17_bienestar, reg18_bienestar, reg19_bienestar))
```

```{r}
plot_model(reg15_bienestar, type = "int", mdrt.values="meansd")
```

```{r}
plot_model(reg17_bienestar, type = "int", mdrt.values="meansd")
```

```{r}
plot_model(reg18_bienestar, type = "int", mdrt.values="meansd")
```

*meritocracia x ses escuela*

```{r}
reg20_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*esfuerzo_soc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
reg21_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*merito_soc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
reg22_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*inteligencia_soc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
reg23_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*esfuerzo_esc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
reg24_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*inteligencia_esc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
```

```{r results='asis'}
htmlreg(list(reg10.1_bienestar, reg20_bienestar, reg21_bienestar, reg22_bienestar, reg23_bienestar, reg24_bienestar))
```

```{r}
plot_model(reg21_bienestar, type = "int", mdrt.values="meansd")
```

```{r}
plot_model(reg22_bienestar, type = "int", mdrt.values="meansd")
```

*meritocracia x prop. educacion universitaria por escuela*

```{r}
reg20.1_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*esfuerzo_soc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
reg21.1_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*merito_soc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
reg22.1_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*inteligencia_soc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
reg23.1_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*esfuerzo_esc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
reg24.1_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*inteligencia_esc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
```

```{r results='asis'}
htmlreg(list(reg10.1_bienestar, reg20.1_bienestar, reg21.1_bienestar, reg22.1_bienestar, reg23.1_bienestar, reg24.1_bienestar))
```

```{r}
plot_model(reg20.1_bienestar, type = "int", mdrt.values="meansd")
```

```{r}
plot_model(reg21.1_bienestar, type = "int", mdrt.values="meansd")
```

```{r}
plot_model(reg22.1_bienestar, type = "int", mdrt.values="meansd")
```

*meritocracia x simce*

```{r}
reg25_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*esfuerzo_soc + internet_rec + (1 + simce | mrbd), data=data_bienestar)

reg26_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*merito_soc + internet_rec + (1 + simce | mrbd), data=data_bienestar)

reg27_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*inteligencia_soc + internet_rec + (1 + simce | mrbd), data=data_bienestar)

reg28_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*esfuerzo_esc + internet_rec + (1 + simce | mrbd), data=data_bienestar)

reg29_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*inteligencia_esc + internet_rec + (1 + simce | mrbd), data=data_bienestar)
```

```{r results='asis'}
htmlreg(list(reg10.1_bienestar, reg25_bienestar, reg26_bienestar, reg27_bienestar, reg28_bienestar, reg29_bienestar))
```

```{r}
plot_model(reg25_bienestar, type = "int", mdrt.values="meansd")
```

```{r}
plot_model(reg26_bienestar, type = "int", mdrt.values="meansd")
```

```{r}
plot_model(reg27_bienestar, type = "int", mdrt.values="meansd")
```

# Modelos de regresión logística ordenada (ordinal logit) (CLMM del paquete ordinal - Cumulative link mixed models)

## El gobierno deberia hacer algo para disminuir las brechas de ingresos

```{r}
data_redis <- data_rec %>% dplyr::select(-just_pension, -just_educ, -just_salud)
data_redis <- na.omit(data_redis)
data_redis$redistribucion_fact <- as_factor(data_redis$redistribucion)

reg10.1_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 | mrbd), data=data_redis)
```

### Interacciones

*meritocracia x ses familiar (educacion)*

```{r}
reg10_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2*esfuerzo_soc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_redis)
reg11_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2*merito_soc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_redis)
reg12_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2*inteligencia_soc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_redis)
reg13_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2*esfuerzo_esc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_redis)
reg14_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2*inteligencia_esc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_redis)
```

```{r results='asis'}
htmlreg(list(reg10.1_redis, reg10_redis, reg11_redis, reg12_redis, reg13_redis, reg14_redis))
```

*meritocracia x ses familiar (libros)*

```{r}
reg15_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec*esfuerzo_soc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_redis)
reg16_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec*merito_soc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_redis)
reg17_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec*inteligencia_soc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_redis)
reg18_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec*esfuerzo_esc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_redis)
reg19_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec*inteligencia_esc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_redis)
```

```{r results='asis'}
htmlreg(list(reg10.1_redis, reg15_redis, reg16_redis, reg17_redis, reg18_redis, reg19_redis))
```

*meritocracia x ses escuela*

```{r}
reg20_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + cod_grupo2*esfuerzo_soc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_redis)

reg21_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + cod_grupo2*merito_soc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_redis)

reg22_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + cod_grupo2*inteligencia_soc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_redis)

reg23_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + cod_grupo2*esfuerzo_esc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_redis)

reg24_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + cod_grupo2*inteligencia_esc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_redis)
```

```{r results='asis'}
htmlreg(list(reg20_redis, reg21_redis, reg22_redis, reg23_redis, reg24_redis))
```

*meritocracia x prop. educacion universitaria por escuela*

```{r}
reg20.1_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + mean_educ*esfuerzo_soc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_redis)

reg21.1_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + mean_educ*merito_soc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_redis)

reg22.1_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + mean_educ*inteligencia_soc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_redis)

reg23.1_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + mean_educ*esfuerzo_esc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_redis)

reg24.1_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + mean_educ*inteligencia_esc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_redis)
```

```{r results='asis'}
htmlreg(list(reg20.1_redis, reg21.1_redis, reg22.1_redis, reg23.1_redis, reg24.1_redis))
```

*meritocracia x simce*

```{r}
reg25_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*esfuerzo_soc + internet_rec + (1 + simce | mrbd), data=data_redis)
reg26_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*merito_soc + internet_rec + (1 + simce | mrbd), data=data_redis)
reg27_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*inteligencia_soc + internet_rec + (1 + simce | mrbd), data=data_redis)
reg28_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*esfuerzo_esc + internet_rec + (1 + simce | mrbd), data=data_redis)
reg29_redis <- clmm(redistribucion_fact ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*inteligencia_esc + internet_rec + (1 + simce | mrbd), data=data_redis)
```

```{r results='asis'}
htmlreg(list(reg25_redis, reg26_redis, reg27_redis, reg28_redis, reg29_redis))
```

# Comparacion interacciones significativas

```{r}
sjPlot::tab_model(list(reg10.1_bienestar, reg11_bienestar, reg12_bienestar, reg15_bienestar, reg18_bienestar, reg21_bienestar, reg21.1_bienestar, reg25_bienestar, reg26_bienestar, reg27_bienestar), show.ci=FALSE)

sjPlot::tab_model(list(reg10.1_bienestar, reg11_bienestar, reg12_bienestar, reg15_bienestar, reg18_bienestar, reg21_bienestar, reg21.1_bienestar, reg25_bienestar, reg26_bienestar, reg27_bienestar), # los modelos estimados
  show.ci=FALSE, # no mostrar intervalo de confianza (por defecto lo hace)
  p.style = "stars", # asteriscos de significación estadística
  dv.labels = c("Model 1",
                "Model 2",
                "Model 3",
                "Model 4",
                "Model 5",
                "Model 6",
                "Model 7",
                "Model 8",
                "Model 9",
                "Model 10"), # etiquetas de modelos o variables dep.
  pred.labels = c("Intercept",
                              "Social effort", 
                              "Social merit", 
                              "Social intelligence", 
                              "School effort", 
                              "School intelligence", 
                              "Secondary <br> <i>(Ref. 8th grade or less)</i>", 
                              "Higher tec.", 
                              "University", 
                              "Missing", 
                              "More than 25 books <br> <i>(Ref. Less than 25)</i>", 
                              "Technology access",
                              "Prop. university level at school",
                              "Subsidized private <br> <i>(Ref. Public)</i>", 
                              "Private", 
                              "SES Medium low <br> <i>(Ref. Low)</i>", 
                              "SES Medium", 
                              "SES Medium high", 
                              "SES High", 
                              "Simce Medium <i>(Ref. Low)</i>", 
                              "Simce High",
                  "Internet <br> <i>(Ref. No)</i>",
                  "Missing",
                              "Secondary*socialmerit",
                              "Tec.*socialmerit",
                              "university*socialmerit",
                              "missing*socialmerit",
                              "Secondary*socialintelligence",
                              "Tec.*socialintelligence",
                              "university*socialintelligence",
                              "missing*socialintelligence",
                              "More than 25 books*socialeffort",
                              "More than 25 books*schooleffort",
                              "SES Medium low*socialmerit", 
                              "SES Medium*socialmerit", 
                              "SES Medium high*socialmerit", 
                              "SES High*socialmerit",
                              "prop. university*socialmerit",
                              "Simce medium*socialeffort",
                              "Simce High*socialeffort",
                              "Simce medium*socialmerit",
                              "Simce high*socialmerit",
                              "Simce medium*socialintelligence",
                              "Simce high*socialintelligence"),
  string.pred = "Predictors", string.est = "β") # nombre predictores y símbolo beta en tabla
```

```{r}
htmlreg(list(reg10.1_bienestar, reg11_bienestar, reg12_bienestar, reg15_bienestar, reg18_bienestar, reg21_bienestar, reg21.1_bienestar, reg25_bienestar, reg26_bienestar, reg27_bienestar),
        custom.model.names = c("Model 1",
                               "Model 2",
                               "Model 3",
                               "Model 4",
                               "Model 5",
                               "Model 6",
                               "Model 7",
                               "Model 8",
                               "Model 9",
                               "Model 10"), 
        custom.note = "$***p <$ 0.001, $**p <$ 0.01, $*p <$ 0.05", 
        omit.coef = "(internet_rec)",
        custom.coef.names = c("Intercept",
                              "Social effort", 
                              "Social merit", 
                              "Social intelligence", 
                              "School effort", 
                              "School intelligence", 
                              "Secondary <br> <i>(Ref. 8th grade or less)</i>", 
                              "Higher tec.", 
                              "University", 
                              "Missing", 
                              "More than 25 books <br> <i>(Ref. Less than 25)</i>", 
                              "Technology access",
                              "Prop. university level at school",
                              "Subsidized private <br> <i>(Ref. Public)</i>", 
                              "Private", 
                              "SES Medium low <br> <i>(Ref. Low)</i>", 
                              "SES Medium", 
                              "SES Medium high", 
                              "SES High", 
                              "Simce Medium <i>(Ref. Low)</i>", 
                              "Simce High",
                              "Secondary*socialmerit",
                              "Tec.*socialmerit",
                              "university*socialmerit",
                              "missing*socialmerit",
                              "Secondary*socialintelligence",
                              "Tec.*socialintelligence",
                              "university*socialintelligence",
                              "missing*socialintelligence",
                              "More than 25 books*socialeffort",
                              "More than 25 books*schooleffort",
                              "SES Medium low*socialmerit", 
                              "SES Medium*socialmerit", 
                              "SES Medium high*socialmerit", 
                              "SES High*socialmerit",
                              "prop. university*socialmerit",
                              "Simce medium*socialeffort",
                              "Simce High*socialeffort",
                              "Simce medium*socialmerit",
                              "Simce high*socialmerit",
                              "Simce medium*socialintelligence",
                              "Simce high*socialintelligence"
                              ),
        file = "output/tables/interaction1.html")
```

```{r}
screenreg(list(reg25_redis, reg26_redis))
```

# Tabla comparativa

Se extraen coeficientes y pvalues de las interacciones

```{r}
# educacion
interact1 <- texreg::extract(reg10_bienestar)
interact1_1 <- cbind(interact1@coef[19], interact1@pvalues[19])

interact2 <- texreg::extract(reg11_bienestar)
interact2_1 <- cbind(interact2@coef[19], interact2@pvalues[19])

interact3 <- texreg::extract(reg12_bienestar)
interact3_1 <- cbind(interact3@coef[19], interact3@pvalues[19])

interact4 <- texreg::extract(reg13_bienestar)
interact4_1 <- cbind(interact4@coef[19], interact4@pvalues[19])

interact5 <- texreg::extract(reg14_bienestar)
interact5_1 <- cbind(interact5@coef[19], interact5@pvalues[19])

#libros
interact6 <- texreg::extract(reg15_bienestar)
interact6_1 <- cbind(interact6@coef[19], interact6@pvalues[19])

interact7 <- texreg::extract(reg16_bienestar)
interact7_1 <- cbind(interact7@coef[19], interact7@pvalues[19])

interact8 <- texreg::extract(reg17_bienestar)
interact8_1 <- cbind(interact8@coef[19], interact8@pvalues[19])

interact9 <- texreg::extract(reg18_bienestar)
interact9_1 <- cbind(interact9@coef[19], interact9@pvalues[19])

interact10 <- texreg::extract(reg19_bienestar)
interact10_1 <- cbind(interact10@coef[19], interact10@pvalues[19])

# Ses school
interact11 <- texreg::extract(reg20_bienestar)
interact11_1 <- cbind(interact11@coef[19], interact11@pvalues[19])

interact12 <- texreg::extract(reg21_bienestar)
interact12_1 <- cbind(interact12@coef[19], interact12@pvalues[19])

interact13 <- texreg::extract(reg22_bienestar)
interact13_1 <- cbind(interact13@coef[19], interact13@pvalues[19])

interact14 <- texreg::extract(reg23_bienestar)
interact14_1 <- cbind(interact14@coef[19], interact14@pvalues[19])

interact15 <- texreg::extract(reg24_bienestar)
interact15_1 <- cbind(interact15@coef[19], interact15@pvalues[19])

# prop educacion school
interact16 <- texreg::extract(reg20.1_bienestar)
interact16_1 <- cbind(interact16@coef[19], interact16@pvalues[19])

interact17 <- texreg::extract(reg21.1_bienestar)
interact17_1 <- cbind(interact17@coef[19], interact17@pvalues[19])

interact18 <- texreg::extract(reg22.1_bienestar)
interact18_1 <- cbind(interact18@coef[19], interact18@pvalues[19])

interact19 <- texreg::extract(reg23.1_bienestar)
interact19_1 <- cbind(interact19@coef[19], interact19@pvalues[19])

interact20 <- texreg::extract(reg24.1_bienestar)
interact20_1 <- cbind(interact20@coef[19], interact20@pvalues[19])

# Simce
interact21 <- texreg::extract(reg25_bienestar)
interact21_1 <- cbind(interact21@coef[19:20], interact21@pvalues[19:20])

interact22 <- texreg::extract(reg26_bienestar)
interact22_1 <- cbind(interact22@coef[19:20], interact22@pvalues[19:20])

interact23 <- texreg::extract(reg27_bienestar)
interact23_1 <- cbind(interact23@coef[19:20], interact23@pvalues[19:20])

interact24 <- texreg::extract(reg28_bienestar)
interact24_1 <- cbind(interact24@coef[19:20], interact24@pvalues[19:20])

interact25 <- texreg::extract(reg29_bienestar)
interact25_1 <- cbind(interact25@coef[19:20], interact25@pvalues[19:20])
```

```{r}
soc_effort = 
  c(paste0(round(interact1_1[,1],3), stars.pval(interact1_1[,2])),
    paste0(round(interact6_1[,1],3), stars.pval(interact6_1[,2])),
    paste0(round(interact11_1[,1],3), stars.pval(interact11_1[,2])),
    paste0(round(interact16_1[,1],3), stars.pval(interact16_1[,2])),
    paste0(round(interact21_1[,1],3), stars.pval(interact21_1[,2])))

soc_merit =
  c(paste0(round(interact2_1[,1],3), stars.pval(interact2_1[,2])),
    paste0(round(interact7_1[,1],3), stars.pval(interact7_1[,2])),
    paste0(round(interact12_1[,1],3), stars.pval(interact12_1[,2])),
    paste0(round(interact17_1[,1],3), stars.pval(interact17_1[,2])),
    paste0(round(interact22_1[,1],3), stars.pval(interact22_1[,2])))

soc_intel =
  c(paste0(round(interact3_1[,1],3), stars.pval(interact3_1[,2])),
    paste0(round(interact8_1[,1],3), stars.pval(interact8_1[,2])),
    paste0(round(interact13_1[,1],3), stars.pval(interact13_1[,2])),
    paste0(round(interact18_1[,1],3), stars.pval(interact18_1[,2])),
    paste0(round(interact23_1[,1],3), stars.pval(interact23_1[,2])))

school_effort =
  c(paste0(round(interact4_1[,1],3), stars.pval(interact4_1[,2])),
    paste0(round(interact9_1[,1],3), stars.pval(interact9_1[,2])),
    paste0(round(interact14_1[,1],3), stars.pval(interact14_1[,2])),
    paste0(round(interact19_1[,1],3), stars.pval(interact19_1[,2])),
    paste0(round(interact24_1[,1],3), stars.pval(interact24_1[,2])))

school_intel =
  c(paste0(round(interact5_1[,1],3), stars.pval(interact5_1[,2])),
    paste0(round(interact10_1[,1],3), stars.pval(interact10_1[,2])),
    paste0(round(interact15_1[,1],3), stars.pval(interact15_1[,2])),
    paste0(round(interact20_1[,1],3), stars.pval(interact20_1[,2])),
    paste0(round(interact25_1[,1],3), stars.pval(interact25_1[,2])))

collapse_rows <-
  data.frame(vars = c("University or Postgraduate", 
                      "More than 25 books", 
                      "School SES high", 
                      "Prop. university level",
                      "Simce Medium", 
                      "Simce High"),
             school_intel,
             school_effort,
             soc_intel,  
             soc_effort, 
             soc_merit 
             ) #espacio privado
```

### Estas interacciones de redistribucion no se incluirán en la tabla final
```{r, eval=FALSE}
# educacion
interact1.1 <- texreg::extract(reg10_redis)
interact1.1_1 <- cbind(interact1.1@coef[18:19], interact1.1@pvalues[18:19])

interact2.1 <- texreg::extract(reg11_redis)
interact2.1_1 <- cbind(interact2.1@coef[18:19], interact2.1@pvalues[18:19])

interact3.1 <- texreg::extract(reg12_redis)
interact3.1_1 <- cbind(interact3.1@coef[18:19], interact3.1@pvalues[18:19])

interact4.1 <- texreg::extract(reg13_redis)
interact4.1_1 <- cbind(interact4.1@coef[18:19], interact4.1@pvalues[18:19])

interact5.1 <- texreg::extract(reg14_redis)
interact5.1_1 <- cbind(interact5.1@coef[18:19], interact5.1@pvalues[18:19])

#libros
interact6.1 <- texreg::extract(reg15_redis)
interact6.1_1 <- cbind(interact6.1@coef[18], interact6.1@pvalues[18])

interact7.1 <- texreg::extract(reg16_redis)
interact7.1_1 <- cbind(interact7.1@coef[18], interact7.1@pvalues[18])

interact8.1 <- texreg::extract(reg17_redis)
interact8.1_1 <- cbind(interact8.1@coef[18], interact8.1@pvalues[18])

interact9.1 <- texreg::extract(reg18_redis)
interact9.1_1 <- cbind(interact9.1@coef[18], interact9.1@pvalues[18])

interact10.1 <- texreg::extract(reg19_redis)
interact10.1_1 <- cbind(interact10.1@coef[18], interact10.1@pvalues[18])

# Ses school
interact11.1 <- texreg::extract(reg20_redis)
interact11.1_1 <- cbind(interact11.1@coef[18], interact11.1@pvalues[18])

interact12.1 <- texreg::extract(reg21_redis)
interact12.1_1 <- cbind(interact12.1@coef[18], interact12.1@pvalues[18])

interact13.1 <- texreg::extract(reg22_redis)
interact13.1_1 <- cbind(interact13.1@coef[18], interact13.1@pvalues[18])

interact14.1 <- texreg::extract(reg23_redis)
interact14.1_1 <- cbind(interact14.1@coef[18], interact14.1@pvalues[18])

interact15.1 <- texreg::extract(reg24_redis)
interact15.1_1 <- cbind(interact15.1@coef[18], interact15.1@pvalues[18])

# prop educacion school
interact16.1 <- texreg::extract(reg20.1_redis)
interact16.1_1 <- cbind(interact16.1@coef[18], interact16.1@pvalues[18])

interact17.1 <- texreg::extract(reg21.1_redis)
interact17.1_1 <- cbind(interact17.1@coef[18], interact17.1@pvalues[18])

interact18.1 <- texreg::extract(reg22.1_redis)
interact18.1_1 <- cbind(interact18.1@coef[18], interact18.1@pvalues[18])

interact19.1 <- texreg::extract(reg23.1_redis)
interact19.1_1 <- cbind(interact19.1@coef[18], interact19.1@pvalues[18])

interact20.1 <- texreg::extract(reg24.1_redis)
interact20.1_1 <- cbind(interact20.1@coef[18], interact20.1@pvalues[18])

# Simce
interact21.1 <- texreg::extract(reg25_redis)
interact21.1_1 <- cbind(interact21.1@coef[18:19], interact21.1@pvalues[18:19])

interact22.1 <- texreg::extract(reg26_redis)
interact22.1_1 <- cbind(interact22.1@coef[18:19], interact22.1@pvalues[18:19])

interact23.1 <- texreg::extract(reg27_redis)
interact23.1_1 <- cbind(interact23.1@coef[18:19], interact23.1@pvalues[18:19])

interact24.1 <- texreg::extract(reg28_redis)
interact24.1_1 <- cbind(interact24.1@coef[18:19], interact24.1@pvalues[18:19])

interact25.1 <- texreg::extract(reg29_redis)
interact25.1_1 <- cbind(interact25.1@coef[18:19], interact25.1@pvalues[18:19])
```

```{r, eval=FALSE}
soc_effort.1 = 
  c(paste0(round(interact1.1_1[,1],3), stars.pval(interact1.1_1[,2])),
    paste0(round(interact6.1_1[,1],3), stars.pval(interact6.1_1[,2])),
    paste0(round(interact11.1_1[,1],3), stars.pval(interact11.1_1[,2])),
    paste0(round(interact16.1_1[,1],3), stars.pval(interact16.1_1[,2])),
    paste0(round(interact21.1_1[,1],3), stars.pval(interact21.1_1[,2])))

soc_merit.1 =
  c(paste0(round(interact2.1_1[,1],3), stars.pval(interact2.1_1[,2])),
    paste0(round(interact7.1_1[,1],3), stars.pval(interact7.1_1[,2])),
    paste0(round(interact12.1_1[,1],3), stars.pval(interact12.1_1[,2])),
    paste0(round(interact17.1_1[,1],3), stars.pval(interact17.1_1[,2])),
    paste0(round(interact22.1_1[,1],3), stars.pval(interact22.1_1[,2])))

soc_intel.1 =
  c(paste0(round(interact3.1_1[,1],3), stars.pval(interact3.1_1[,2])),
    paste0(round(interact8.1_1[,1],3), stars.pval(interact8.1_1[,2])),
    paste0(round(interact13.1_1[,1],3), stars.pval(interact13.1_1[,2])),
    paste0(round(interact18.1_1[,1],3), stars.pval(interact18.1_1[,2])),
    paste0(round(interact23.1_1[,1],3), stars.pval(interact23.1_1[,2])))

school_effort.1 =
  c(paste0(round(interact4.1_1[,1],3), stars.pval(interact4.1_1[,2])),
    paste0(round(interact9.1_1[,1],3), stars.pval(interact9.1_1[,2])),
    paste0(round(interact14.1_1[,1],3), stars.pval(interact14.1_1[,2])),
    paste0(round(interact19.1_1[,1],3), stars.pval(interact19.1_1[,2])),
    paste0(round(interact24.1_1[,1],3), stars.pval(interact24.1_1[,2])))

school_intel.1 =
  c(paste0(round(interact5.1_1[,1],3), stars.pval(interact5.1_1[,2])),
    paste0(round(interact10.1_1[,1],3), stars.pval(interact10.1_1[,2])),
    paste0(round(interact15.1_1[,1],3), stars.pval(interact15.1_1[,2])),
    paste0(round(interact20.1_1[,1],3), stars.pval(interact20.1_1[,2])),
    paste0(round(interact25.1_1[,1],3), stars.pval(interact25.1_1[,2])))

collapse_rows.1 <-
  data.frame(vars = c("University or Postgraduate", 
                      "Ed. Missing", 
                      "More than 25 books",
                      "School SES high", 
                      "Prop. university level",
                      "Simce Medium", 
                      "Simce High"),
             soc_effort.1, 
             soc_merit.1, 
             soc_intel.1, 
             school_effort.1, 
             school_intel.1) #espacio privado
```

```{r}
collapse_cols <- left_join(collapse_rows, collapse_rows.1, by="vars")
save(collapse_cols, file="input/data/proc/collapse_cols2.RData")
```

, 
    caption = "Interactions",
    align = "lcccccccccc", booktabs = T,
    col.names =  c("Variable",
                   "Social effort", 
                   "Social merit",
                   "Social intelligence", 
                   "School effort", 
                   "School intelligence",
                   "Social effort", 
                   "Social merit",
                   "Social intelligence", 
                   "School effort", 
                   "School intelligence"

```{r}
kbl(collapse_rows,
    align = "lccccc", booktabs = T,
    col.names =  c("Variable", 
                   "School talent", 
                   "School effort",
                   "Social talent",
                   "Social effort", 
                   "Deservingness")) %>%
  add_header_above(c(" " = 1,
                     "Justification of Inequality"=5), 
                   bold = T)%>%  
  kable_styling(latex_options = c("scale_down","HOLD_position"),
                full_width = F) %>%
  column_spec(1, bold = T, width = "4cm") %>%
  column_spec(column = 2:6,width = "1.5cm") %>%
  collapse_rows(columns = 1:2, valign = "middle") %>% 
  footnote(general = "***p < 0.001, **p < 0.01, *p < 0.05",
           footnote_as_chunk = T,general_title = "Note: ") %>% 
  save_kable("output/tables/interactions3.png")
```

# Grafico de interacciones

```{r}
int1 <- plot_model(reg11_bienestar, type = "int", mdrt.values="meansd", title = "") +
  theme_bw() +
  labs(x = "Parental educational level", y="Justification of Inequality", colour = "Deservingness")

ggsave(int1, file="output/graphs/int1.png")
```

```{r}
int2 <- plot_model(reg26_bienestar, type = "int", mdrt.values="meansd", title = "") +
  theme_bw() +
  labs(x = "School Simce", y="Justification of Inequality", colour = "Deservingness")

ggsave(int2, file="output/graphs/int2.png")
```

#### Gráfico sin categoría missing (con código sale sin intervalos de confianza; se deja Missing en blanco para ocultar)
```{r}
reg11.2_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + merito_soc*educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)
#no se xq no funciona el de abajo
int1.2 <- sjPlot::plot_model(reg11.2_bienestar, type = "int", title = "", colors = c("red", "blue", "white")) +
  theme_bw() +
  labs(x ="Deservingness" , y="Justification of Inequality", colour = "Parental educational level")
int1.2
ggsave(int1.2, file="output/graphs/int1.2.png")
```

```{r}
reg26.2_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + merito_soc*simce + internet_rec + (1 + simce | mrbd), data=data_bienestar)
```
```{r}
int2.2 <- plot_model(reg26.2_bienestar, type = "int", mdrt.values="minmax", title = "", colors = c("red", "white", "blue")) +
  theme_bw() +
  ylim(1,4)+
  labs(x = "Deservingness", y="Justification of Inequality", colour = "School Simce")
int2.2
ggsave(int2.2, file="output/graphs/int2.2.png")
```

* Lo mismo en español
```{r}
int2.2_esp <- plot_model(reg26.2_bienestar, type = "int", mdrt.values="minmax", title = "", colors = c("red", "white", "blue")) +
  theme_bw() +
  ylim(1,4)+
  labs(x = "Merecimiento", y="Justificación de la desigualdad", colour = "Simce escolar")

int2.2_esp
ggsave(int2.2_esp, file="output/graphs/int2.2_esp.png")
```

```{r}
int2.3 <- plot_model(reg26.2_bienestar, type = "int", mdrt.values="minmax", title = "", colors = c("red", "white", "blue")) +
  theme_bw() +
  labs(x = "Mérito en la sociedad", y="Índice de acceso a servicios sociales", colour = "Logro Simce")
int2.3
ggsave(int2.3, file="output/graphs/int2.3.png")
```

### 2024-06-28 02-antecedentes.qmd

@@
 According to this data, 35.9% prefer private health insurance, and 63% would prefer private education.
-Yet, research on children's justification in this area remains limited, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs (Centro de Estudios Públicos, 2024).
+Yet, research on children's justification in this area remains limited, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs [@centrodeestudiospublicos_estudio_2024].
 
@@
 
-Figure @fig-hypotheses depicts a summary of the research hypothesis described before.
+@fig-hypotheses depicts a summary of the research hypothesis described before.
 Market justice preferences are our main concept under study (vis-a-vis dependent variable), which refers to the justification of better access to health, education, and pensions based on payment capacities (as we detail further in the methods section).

### 2024-07-03 01-introduction.qmd

@@
 In the present paper, we address the role of the perception of meritocracy on the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system.
-In particular, we focus in justifying inequalities in health, education and pensions, which are traditional social policy areas where acces to better services could be justified based on payment capacity, referred as to market justice [@lane_market_1986, @lindh_bringing_2023].
+In particular, we focus in justifying inequalities in health, education and pensions, which are traditional social policy areas where acces to better services could be justified based on payment capacity, referred as to market justice [@lane_market_1986; @lindh_bringing_2023].
 Most of the research in this realm has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies [@batruch_belief_2022].
@@
 This shift in economic policy has allowed the unprecedented emergence of markets in areas related to social reproduction, such as health, pensions, and education, with provision and access managed by private entities and segmented by individual payment capacity, heavily reliant on State subsidies [@boccardo_30_2020].
-In health, although the majority of the population uses the public insurance system (78.9%), 15.3% are served by private insurers (Observatorio Social, 2023).
-The pension system is based on individual capitalization, with mandatory contributions managed by private administrators investing in the financial market, currently involving 11 million contributors (Superintendencia de Pensiones, 2024).
-In education, 30.6% of school enrollment is in public schools, 54.0% in state-subsidized (voucher) private schools, and 9.3% in fully private schools, generally attracting higher-income groups (Ministerio de Educación, 2023).
+In health, although the majority of the population uses the public insurance system (78.9%), 15.3% are served by private insurers [@observatoriosocial_estadisticas_2024].
+The pension system is based on individual capitalization, with mandatory contributions managed by private administrators investing in the financial market, currently involving 11 million contributors [@superintendenciadepensiones_estadisticas_2024].
+In education, 30.6% of school enrollment is in public schools, 54.0% in state-subsidized (voucher) private schools, and 9.3% in fully private schools, generally attracting higher-income groups [@ministeriodeeducacion_resumen_2023].

### 2024-07-03 04-analisis.qmd
correccion analisis jc

@@
 ```
+---
+author: "Equipo Edumer"
+bibliography: "input/bib/just-ineq-merit.bib"
+csl: "input/bib/apa6.csl"
+---
+
+# Results
+
+## Descriptive
+
+```{r packages, echo=FALSE}
+pacman::p_load(tidyverse,
+               ggpubr, 
+               lme4,
+               ordinal,
+               sjPlot, 
+               texreg,
+               kableExtra,
+               easystats,
+               ggeffects)
+
+options(scipen=999)
+rm(list = ls())
+```
+
+```{r data, echo=FALSE}
+
+load("input/data/data_index.RData")
+load("input/data/deservingness.RData")
+load("input/data/soc_effort.RData")
+load("input/data/soc_talent.RData")
+load("input/data/school_effort.RData")
+load("input/data/school_talent.RData")
+load("input/data/data_rec.RData")
+
+```
+
+```{r functions, echo=FALSE}
+
+odds_fun <- function(modelos) {
+   or <- texreg::extract(modelos)
+   or@coef <- exp(or@coef)
+   return(or)
+}
+
+miles <- function(x) {
+  format(round(as.numeric(x),0), big.mark = ".")
+}
+
+decimales <- function(x) {
+  format(round(as.numeric(x), 2), decimal.mark = ",")
+}
+
+# set theme
+
+theme_set(theme_bw())
+
+```
+
+@fig-marketjustice illustrates the distribution of responses to the three items on market justice preferences (healthcare, pensions, and education), indicating that most respondents tend to disagree or strongly disagree with the idea that it is just for people with higher incomes to have access to better services in these areas.
+The strongest opposition is observed in the healthcare sector, as the majority of respondents (66.4%) are against the idea that higher-income individuals should have access to better healthcare services.
+Similarly, regarding the concept of justice related to access to better pensions based on income, the level of disagreement reaches 63.0%.
+This decreases significantly in the field of education, as the percentage of respondents who disagree or strongly disagree with those with higher incomes having access to better education decreases to 58.8%.
+Only a small proportion agree that access to healthcare should be conditional on income (23.6%), and a minority strongly agree (10%), as is the case with pensions.
+However, this increases in the case of education, being the area where there is a significant proportion of respondents who agree or strongly agree with access based on market justice (41.2%).
+
+```{r echo=FALSE}
+#| fig-cap: "Market justice preferences in health, pensions and education"
+#| fig-asp: 0.6
+#| label: fig-marketjustice
+
+data_rec %>% 
+  select(just_salud, just_pension, just_educ) %>% 
+  sjPlot::plot_stackfrq(geom.size = 0.5,
+                        expand.grid =TRUE,
+                      geom.colors = "PuBu",
+                      show.total = FALSE, 
+                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
+                      axis.labels = c("Health", "Pensions", "Education")) +
+ # theme_light(base_size = 16) +
+  theme(legend.position="bottom", text = element_text(size = 14),
+        plot.caption = element_text(hjust = 1)) +
+       labs(caption = "Source: own elaboration based on First Study of Civic Education 2017")
+
+```
+
+Regarding the perception of meritocracy, @fig-meritocracy displays the frequency distribution of five items related to the dimensions of school and society.
+At the school level, there is a strong belief that both effort and talent are important for achieving good grades, with especially high agreement for effort.
+Specifically, while 78.6% agree or strongly agree on the importance of talent for obtaining good grades, this proportion rises to 96.8% for effort, indicating that respondents value effort more than talent in this context.
+
+```{r echo=FALSE}
+#| fig-cap: "Social and school meritocracy"
+#| fig-asp: 0.8
+#| label: fig-meritocracy
+
+plot<-data_rec %>% 
+  select(merito_soc, esfuerzo_soc, inteligencia_soc, esfuerzo_esc, inteligencia_esc) %>% 
+  sjPlot::plot_stackfrq(geom.size = 0.5,
+                        expand.grid =TRUE,
+                      geom.colors = "PuBu",
+                      show.total = FALSE, 
+                      vjust=c(rep(c("center"),12), "bottom", "top", rep(c("center"), 6)),
+                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
+                      axis.labels = c("Deservingness", "Social effort", "Social talent", "School effort", "School talent")) +
+  #theme_light(base_size = 16) +
+  theme(legend.position="bottom", text = element_text(size = 14),
+        plot.caption = element_text(hjust = 1))
+
+plot + 
+  annotate("text", x = 5.5, y = 0, label = "School meritocracy", size = 5, hjust = 0, vjust = 1) +
+  annotate("text", x = 3.5, y = 0, label = "Social meritocracy", size = 5, hjust = 0, vjust = 1) +
+       labs(caption = "Source: own elaboration based on First Study of Civic Education 2017")
+```
+
+At the societal level, @fig-meritocracy shows that a majority of respondents agree or strongly agree that individuals are rewarded for their talents (63.7%) and efforts (62%).
+Nevertheless, a notable proportion perceives that both effort and talent are not adequately rewarded in society, with 38.1% and 36.3% respectively.
+The perception of whether people get what they deserve in society is more divided; while a slim majority (56.5% combining agree and strongly agree) agrees, a significant minority (43.6%) disagrees, indicating that this is the most contentious item among respondents.
+There is general agreement that people are rewarded for talent and effort in society, but the consensus is weaker compared to school meritocracy.
+
+@fig-bivariate shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram in @fig-hypotheses).
+In general we observe a positive association between meritocracy and market justice preferences for the societal questions, but this is not as straightforward for the school items, particularly for the one referred to effort.
+
+```{r echo=FALSE}
+#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
+#| fig-asp: 1.22
+#| label: fig-bivariate
+
+plot_soc_effort <- soc_effort %>% 
+  mutate(justice = str_to_title(justice),
+         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
+  arrange(esfuerzo_soc, justice) %>% 
+  ggplot(aes(x = esfuerzo_soc, y = mean)) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_brewer(palette = "Paired") +
+  labs(x = "Social effort",
+       y = "Market justice preferences") +
+  theme(legend.position = "none",
+        text = element_text(size = 11))
+
+plot_deservingness <- deservingness %>% 
+   mutate(justice = str_to_title(justice),
+         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
+  arrange(merito_soc, justice) %>% 
+  ggplot(aes(x = merito_soc, y = mean)) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_brewer(palette = "Paired") +
+  labs(x = "Deservingness",
+       y = "Market justice preferences") +
+  theme(legend.position = "none",
+        text = element_text(size = 11))
+
+plot_soc_talent <- soc_talent %>% 
+   mutate(justice = str_to_title(justice),
+         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
+  arrange(inteligencia_soc, justice) %>%  
+  ggplot(aes(x = inteligencia_soc, y = mean)) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_brewer(palette = "Paired") +
+  labs(x = "Social talent",
+       y = "Market justice preferences") +
+  theme(legend.position = "none",
+        text = element_text(size = 11))
+
+plot_school_effort <- school_effort %>% 
+  mutate(justice = str_to_title(justice),
+         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
+  arrange(esfuerzo_esc, justice) %>%  
+  ggplot(aes(x = esfuerzo_esc, y = mean)) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_brewer(palette = "Paired") +
+  labs(x = "School effort",
+       y = "") +
+  theme(legend.position = "none",
+        text = element_text(size = 11))
+
+plot_school_talent <- school_talent %>% 
+  mutate(justice = str_to_title(justice),
+         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
+  arrange(inteligencia_esc, justice) %>% 
+  ggplot(aes(x = inteligencia_esc, y = mean)) +
+  geom_point(aes(color = justice, shape = justice), size = 2) +
+  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
+  scale_x_continuous(limits = c(1, 4),
+                     breaks = seq(1, 4, by = 1),
+                     label = c("Strongly\n disagree", 
+                               "Disagree", 
+                               "Agree", 
+                               "Strongly\n agree")) +
+  scale_color_brewer(palette = "Paired") +
+  labs(x = "School talent",
+       y = "",
+       color = NULL,
+       shape = NULL) +
+  theme(text = element_text(size = 11))
+
+leg <- get_legend(plot_school_talent)
+
+plot_school_talent <- plot_school_talent + theme(legend.position = "none")
+
+ggarrange(plot_soc_effort, plot_school_effort, plot_soc_talent,
+          plot_school_talent, plot_deservingness, leg, ncol = 2, nrow = 3) +
+  labs(caption = "Source: own elaboration based on First Study of Civic Education 2017")
+```
+
+## Multivariate
+
+This section presents the results of the multilevel models.
+We begin by showing the results of the cumulative link mixed models in @tbl-ordinal-reg for the ordinal dependent variables of justice in health, pensions, and education separately in order to compare the effect of the predictors.
+Then, we present the results of the linear mixed-effects models for the average index of preference for market justice in @tbl-lineal-reg[^1].
+
+For the health item, the intraclass correlation (ICC) obtained for the null model (not shown) indicates that 5% of the total variance of this item is associated with specific characteristics of the schools, while it is 3% for the pensions item and 4% for the education item (ICC based on [@hox_multilevel_2010, p.15]).
+For the average index of market justice preferences, the ICC reaches 4%, indicating that overall, only a small percentage of the total variance in students' market justice preferences is associated with school characteristics, thus limiting the possibility of finding effects at this second level of analysis.
+
+[^1]: The table shows the effects of individual and contextual-level independent variables related to our hypotheses; however, the models also incorporate the control variables for each of the three dependent variables separately and for the average index (complete models are available in the [Appendix](#appendix)).
+
+```{r echo=FALSE, include=FALSE, eval=FALSE}
+sjmisc::frq(data_rec$educacion_rec)
+data_rec <- data_rec %>% mutate(universitario = case_when(educacion_rec=="Ed Basica"~"No",
+                                                          educacion_rec=="Ed Media"~"No",
+                                                          educacion_rec=="Ed Tecnica"~"No",
+                                                          educacion_rec=="Universidad o posgrado"~"Si",
+                                                          educacion_rec=="Ns/Nr"~"Ns/Nr"))
+data_rec$universitario <- factor(data_rec$universitario, levels = c("No", "Si", "Ns/Nr"))
+sjmisc::frq(data_rec$universitario)
+
+data_rec <- data_rec %>% mutate(school_nse = case_when(cod_grupo=="Bajo"~"Bajo",
+                                                          cod_grupo=="Medio bajo"~"Medio",
+                                                          cod_grupo=="Medio"~"Medio",
+                                                          cod_grupo=="Medio alto"~"Medio",
+                                                          cod_grupo=="Alto"~"Alto"))
+sjmisc::frq(data_rec$school_nse)
+data_rec$school_nse <- factor(data_rec$school_nse, levels = c("Bajo", "Medio", "Alto"))
+
+data_pension <- data_rec %>% 
+    dplyr::select(-just_educ, -just_salud, -redistribucion)
+data_pension <- na.omit(data_pension)
+data_pension$just_pension_fact <- as_factor(data_pension$just_pension)
+
+m0_p <- clmm(just_pension_fact ~ 1 + (1 | mrbd), data=data_pension)
+
+performance::icc(m0_p, by_group = T)
+
+m1_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_pension)
+
+m2_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
+    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_pension)
+
+m3_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
+    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
+    school_nse + simce + (1 | mrbd), data=data_pension)
+
+modelos_pension <- list(m1_p, m2_p, m3_p)
+
+# modelos_odd_pension <- map(tbls, odds_fun)
+
+data_salud <- data_rec %>% 
+    dplyr::select(-just_pension, -just_educ, -redistribucion)
+data_salud <- na.omit(data_salud)
+data_salud$just_salud_fact <- as_factor(data_salud$just_salud)
+
+m0_h <- clmm(just_salud_fact ~ 1 + (1 | mrbd), data=data_salud)
+
+performance::icc(m0_h, by_group = T)
+
+m1_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_salud)
+
+m2_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
+    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_salud)
+
+m3_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
+    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
+    school_nse + simce + (1 | mrbd), data=data_salud)
+
+modelos_health <- list(m1_h, m2_h, m3_h)
+
+# modelos_odd_health <- map(tbls, odds_fun)
+
+data_educ <- data_rec %>% 
+    dplyr::select(-just_pension, -just_salud, -redistribucion)
+data_educ <- na.omit(data_educ)
+data_educ$just_educ_fact <- as_factor(data_educ$just_educ)
+
+m0_e <- clmm(just_educ_fact ~ 1 + (1 | mrbd), data=data_educ)
+
+performance::icc(m0_e, by_group = T)
+
+m1_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_educ)
+
+m2_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
+    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_educ)
+
+m3_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
+    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
+    school_nse + simce + (1 | mrbd), data=data_educ)
+
+modelos_educ <- list(m1_e, m2_e, m3_e)
+
+#modelos_odd_educ <- map(tbls, odds_fun)
+
+```
+
+```{r echo=FALSE, results='asis', eval=FALSE}
+#| tbl-cap: "Cumulative link multilevel models of differential acces justificacion of health, pensions, and education"
+#| label: tbl-ordinal-reg
+
+ccoef <- list(
+  inteligencia_esc = "School talent",
+  esfuerzo_esc = "School effort",
+  inteligencia_soc = "Social talent",
+  esfuerzo_soc = "Social effort",
+  merito_soc = "Deservingness",
+  "universitarioSi" = "University or posgraduate",
+  "universitarioNs/Nr" = "Missing",
+  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
+  "school_nseMedio" = "SES Medium",
+  school_nseAlto = "SES High",
+  simceMedio = "Simce Medium",
+  simceAlto = "Simce High"
+)
+
+texreg::htmlreg(l = list(modelos_health[[2]],
+                         modelos_health[[3]],
+                         modelos_pension[[2]],
+                         modelos_pension[[3]],
+                         modelos_educ[[2]],
+                         modelos_educ[[3]]),   
+                caption = NULL,
+                stars = c(0.05, 0.01, 0.001),
+                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses.
+Control variables are included.
+%stars.
+Source: own elaboration based on First Study of Civic Education 2017.",
+                custom.header = list("Health" = 1:2, "Pensions" = 3:4, "Education" = 5:6),
+                custom.model.names = rep(c("Model 1", "Model 2"), 3),
+                custom.coef.map = ccoef,
+                leading.zero = T,
+                scalebox = 0.60,
+                groups = list("Parental education (Ref.= Non-university)" = 6:7,
+                                "Socioeconomic level (Ref.= Low)" = 9:10,
+                                "Achievement score (Ref.= Low)" = 11:12),
+                custom.gof.rows = list("Controls"=rep("Yes",6)),
+                include.loglik = FALSE, include.aic = FALSE,
+                file = "output/tables/tab-ordinalreg.html")
+
+webshot::webshot(url = "output/tables/tab-ordinalreg.html", file = "output/tables/tab-ordinalreg.png")
+```
+
+![](output/tables/tab-ordinalreg.png)
+
+Model 1 in @tbl-ordinal-reg for the three dependent variables of justice in differential access to health, pensions, and education demonstrates a similar trend regarding the effect of the perception of meritocracy on these items, with all effects being statistically significant (*p* \< 0.05).
+Regarding the variables of social meritocracy, the three variables---talent, effort, and deservingness---show that as these increase, the justification for differentiated access to pensions, education, and health also increases.
+In the context of school meritocracy, the effects are mixed: as perceived school talent increases, the justification for differentiated access to these services increases; conversely, as perceived school effort increases, the justification decreases, holding all other variables constant.
+
+Additionally, Model 1 examines family status through parents' educational level and the number of books at home.
+For health, pensions, and education, a negative but statistically non-significant association with parents' education is observed.
+However, having more than 25 books at home shows a negative and statistically significant effect on market justice in health and pensions, suggesting less justification for differentiated access compared to those with fewer books.
+
+Model 2 in @tbl-ordinal-reg incorporates school-level variables.
+The results show that high socioeconomic level schools show greater preferences for market justice in health, pensions, and education compared to low socioeconomic level schools, with statistically significant results (*p* \< 0.05).
+Moreover, and contrary to our hypothesis, schools with higher academic performance show negative and statistically significant effects across the three items.
+
+```{r echo=FALSE, include=FALSE}
+data_bienestar <- data_bienestar %>% mutate(universitario = case_when(educacion_rec=="Ed Basica"~"No",
+                                                          educacion_rec=="Ed Media"~"No",
+                                                          educacion_rec=="Ed Tecnica"~"No",
+                                                          educacion_rec=="Universidad o posgrado"~"Si",
+                                                          educacion_rec=="Ns/Nr"~"Ns/Nr"))
+data_bienestar$universitario <- factor(data_bienestar$universitario, levels = c("No", "Si", "Ns/Nr"))
+
+sjmisc::frq(data_bienestar$cod_grupo)
+data_bienestar <- data_bienestar %>% mutate(school_nse = case_when(cod_grupo=="Bajo"~"Bajo",
+                                                          cod_grupo=="Medio bajo"~"Medio",
+                                                          cod_grupo=="Medio"~"Medio",
+                                                          cod_grupo=="Medio alto"~"Medio",
+                                                          cod_grupo=="Alto"~"Alto"))
+sjmisc::frq(data_bienestar$school_nse)
+data_bienestar$school_nse <- factor(data_bienestar$school_nse, levels = c("Bajo", "Medio", "Alto"))
+
+reg0_bienestar <- lmer(bienestar ~ 1 + (1 | mrbd), data = data_bienestar)
+
+performance::icc(reg0_bienestar, by_group = T)
+
+reg1_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
+                         inteligencia_soc + esfuerzo_soc + merito_soc + 
+                         acc_tec + internet_rec +
+                         (1 | mrbd), data=data_bienestar)
+
+reg2_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
+                         inteligencia_soc + esfuerzo_soc + merito_soc + 
+                         universitario + libros_rec + acc_tec + internet_rec +
+                         (1 | mrbd), data=data_bienestar)
+
+reg3_bienestar <- lmer(bienestar ~ 1 + mean_educ + cod_depe2 + school_nse +
+                         simce + (1 | mrbd), data=data_bienestar)
+
+reg4_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + 
+                         inteligencia_soc + esfuerzo_esc + inteligencia_esc + 
+                         universitario + libros_rec + acc_tec + internet_rec + 
+                         mean_educ + cod_depe2 + school_nse + 
+                         (1 | mrbd), data=data_bienestar)
+
+reg5_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + 
+                         inteligencia_soc + esfuerzo_esc + inteligencia_esc + 
+                         universitario + libros_rec + acc_tec + internet_rec + 
+                         mean_educ + cod_depe2 + school_nse + simce + 
+                         (1 | mrbd), data=data_bienestar)
+
+test01 <- anova(reg0_bienestar, reg1_bienestar, test = "Chisq")
+test02 <- anova(reg0_bienestar, reg2_bienestar, test = "Chisq")
+test03 <- anova(reg0_bienestar, reg3_bienestar, test = "Chisq")
+test04 <- anova(reg0_bienestar, reg4_bienestar, test = "Chisq")
+test05 <- anova(reg0_bienestar, reg5_bienestar, test = "Chisq")
+
+test.pvalues1 <- test01$`Pr(>Chisq)`[2]
+test.pvalues2 <- test02$`Pr(>Chisq)`[2]
+test.pvalues3 <- test03$`Pr(>Chisq)`[2]
+test.pvalues4 <- test04$`Pr(>Chisq)`[2]
+test.pvalues5 <- test05$`Pr(>Chisq)`[2]
+
+deviance1 <- test01$deviance[2]
+deviance2 <- test02$deviance[2]
+deviance3 <- test03$deviance[2]
+deviance4 <- test04$deviance[2]
+deviance5 <- test05$deviance[2]
+```
+
+```{r echo=FALSE, results='asis'}
+#| tbl-cap: "Linear mixed-effects models for meritocracy perception and market justice preferences"
+#| label: tbl-lineal-reg
+
+ccoef <- list(
+  "(Intercept)" = "Constant",
+  inteligencia_esc = "School talent",
+  esfuerzo_esc = "School effort",
+  inteligencia_soc = "Social talent",
+  esfuerzo_soc = "Social effort",
+  merito_soc = "Deservingness",
+  "universitarioSi" = "University or posgraduate",
+  "universitarioNs/Nr" = "Missing",
+  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
+  "school_nseMedio" = "SES Medium",
+  school_nseAlto = "SES High",
+  simceMedio = "Simce Medium",
+  simceAlto = "Simce High"
+)
+
+texreg::htmlreg(list(reg1_bienestar, reg2_bienestar, reg4_bienestar, reg5_bienestar),
+                custom.model.names = c("Model 1",
+                                       "Model 2",
+                                       "Model 3",
+                                       "Model 4"),
+                caption = NULL,
+                stars = c(0.05, 0.01, 0.001),
+                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses.
+Control variables are included.
+%stars.
+Source: own elaboration based on First Study of Civic Education 2017.",
+                custom.coef.map = ccoef,
+                leading.zero = T,
+                scalebox = 0.60,
+                groups = list("Parental education (Ref.= Non-university)" = 7:8,
+                              "Socioeconomic level (Ref.= Low)" = 10:11,
+                              "Achievement score (Ref.= Low)" = 12:13),
+                custom.gof.rows = list("Controls"=rep("Yes",4),
+                                       "Deviance" = c(deviance1,
+                                                    deviance2,
+                                                    deviance4,
+                                                    deviance5),
+                                       "Deviance Test (p)" = c(test.pvalues1,
+                                                             test.pvalues2,
+                                                             test.pvalues4,
+                                                             test.pvalues5)),
+                include.loglik = FALSE,include.aic = FALSE)
+
+```
+
+@tbl-lineal-reg presents the results of linear mixed-effects models for the market justice preference index.
+Model 1 shows that higher perceived societal meritocracy, through effort, talent, and deservingness, is positively and significantly associated with market justice preferences.
+School meritocracy perceptions yield mixed effects: perceived talent increases market justice preference, while perceived effort decreases it, both statistically significant (*p* \< 0.001).
+Model 2 includes family status, finding no statistically significant effects from parents' education or the number of books at home.
+Regarding school characteristics, Models 3 and 4 introduce socioeconomic level and academic performance.
+The results of Model 3 suggest a statistically significant relationship between school socioeconomic levels and market justice preferences, specifically those schools with high socioeconomic levels ($\beta$ = 0.37) compared to low socioeconomic levels.
+However, Model 4 indicates that schools with medium ($\beta$ = -0.11) and high ($\beta$ = -0.26) academic performance present lower market justice preferences compared to low-performing schools, with these effects being statistically significant when controlling for other variables.
+
+```{r echo=FALSE}
+#| tbl-cap: "Interactions effects"
+#| label: tbl-interact
+
+pacman::p_load(kableExtra)
+load("input/data/interact.RData")
+
+collapse_rows[c(-4),] %>% 
+  as_tibble() %>% 
+  kbl(.,
+      align = "lccccc", booktabs = T,
+      col.names =  c("Variable", 
+                     "School talent", 
+                     "School effort",
+                     "Social talent",
+                     "Social effort", 
+                     "Deservingness")) %>% 
+  add_header_above(c(" " = 1,
+                     "Market Justice Preferences"=5), 
+                   bold = T) %>% 
+  kable_styling(latex_options = c("scale_down","HOLD_position"),
+                full_width = T) %>% 
+  column_spec(1, bold = T, width = "4cm") %>%
+  column_spec(column = 2:6,width = "1.5cm") %>%
+  collapse_rows(columns = 1:2, valign = "middle") %>% 
+  kableExtra::add_footnote(label = "Note: ***p < 0.001, **p < 0.01, *p < 0.05. Source: own elaboration based on First Study of Civic Education 2017.", 
+                           notation = "none")
+
+```
+
+@tbl-interact displays the interaction terms related to our hypotheses.
+The results suggest that students' family background moderates the relationship between their meritocratic perceptions and their justification of market allocation in social services.
+Specifically, the relationship between social effort and market justice preferences is less positive for students whose parents have university or postgraduate education.
+Lower-status students (measured by parental education) prefer more market justice when they strongly adhere to meritocratic perceptions of deservingness.
+A similar moderating effect is seen with family cultural capital: the relationship between effort and market justice preferences is less positive for students with more than 25 books at home.
+
+At the school level, we observe that school status and students' meritocratic beliefs moderate their preferences for market justice.
+High-status schools show less preference for market justice when their students, on average, have a greater perception of meritocracy, as measured by effort, deservingness, and talent.
+Conversely, low-achieving schools prefer more market justice when their students have stronger meritocratic beliefs.
+@fig-interaction illustrates this moderating effect of school achievement on the relationship between deservingness and market justice.
+No moderating effects were observed at the school level for students' meritocratic beliefs regarding effort or talent in achieving good grades.
+
+```{r echo=FALSE, include=FALSE}
+load("input/data/data_rec.RData")
+data_rec <- data_rec %>% rowwise() %>%  mutate(educacion_rec2 = case_when(educacion_rec=="Ed Basica"~"Low level",
+                                                 educacion_rec=="Ed Media"~"Low level",
+                                                 educacion_rec=="Ed Tecnica"~"Low level",
+                                                 educacion_rec=="Universidad o posgrado"~"University or Postgraduate",
+                                                 educacion_rec=="Ns/Nr"~"Missing"
+                                                 ))
+data_rec$educacion_rec2 <- factor(data_rec$educacion_rec2, levels = c("Low level", "University or Postgraduate", "Missing"))
+
+data_rec <- data_rec %>% rowwise() %>%  mutate(cod_grupo2  = case_when(cod_grupo=="Bajo"~"Bajo",
+                                                                       cod_grupo=="Medio bajo"~"Bajo",
+                                                                       cod_grupo=="Medio"~"Bajo",
+                                                                       cod_grupo=="Medio alto"~"Alto",
+                                                                       cod_grupo=="Alto"~"Alto"))
+data_rec$cod_grupo2 <- factor(data_rec$cod_grupo2, levels = c("Bajo", "Alto"))
+data_rec$simce <- factor(data_rec$simce, labels = c("Low", "Medium", "High"))
+
+data_bienestar <- na.omit(data_rec)
+#data_bienestar <- data_bienestar %>% mutate(pension_rec = 5-just_pension,
+#                                            educ_rec = 5-just_educ,
+#                                            salud_rec = 5-just_salud)
+data_bienestar <- data_bienestar %>% rowwise() %>% dplyr::mutate(bienestar = mean(c(just_educ, just_salud, just_pension), na.omit=TRUE))
+data_bienestar$bienestar <- round(data_bienestar$bienestar, 2)
+
+reg26.2_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + merito_soc*simce + internet_rec + (1 + simce | mrbd), data=data_bienestar)
+
+plot_interact <- ggpredict(model = reg26.2_bienestar, terms = c("merito_soc", "simce")) %>% 
+  filter(group != "Medium") %>% 
+  plot() +
+  ylim(1,4) +
+  labs(title = NULL,
+       x = "Deservingness", 
+       y ="Market Justice Preferences", 
+       colour = "School Simce",
+       caption = "Source: own elaboration based on First Study of Civic Education 2017")
+```
+
+```{r echo=FALSE, warning=FALSE}
+#| fig-cap: "Interaction between deservingness and justification of inequality by school SIMCE achievement"
+#| label: fig-interaction
+
+plot_interact +
+  theme_bw()
+
+```

### 2024-07-04 02-antecedentes.qmd

@@
 The influence of ideologies [@wegener_dominant_1995] and cultural schemas [@homan_being_2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations.
-While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999 ; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations [@lindh_bringing_2023] and even less about children's acceptance or rejection of these market principles.
+While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; @kelley_legitimate_2021; @jasso_how_1999 ; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations [@lindh_bringing_2023] and even less about children's acceptance or rejection of these market principles.
 This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades [@centeno_arc_2012; @harvey_breve_2015], affecting areas such as pensions, health services, and education.
@@
 In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2019].
-Some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2019], and how larger inequality decreases it [@morris_representing_2022].
+Some studies have analyzed how those with greater privileges believe more in meritocracy [@reynolds_perceptions_2014], how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2019], and how larger inequality decreases it [@morris_representing_2022].
 

### 2024-07-04 06-conclusion.qmd

@@
 Regarding these descriptive indicators, a a first message coming from this study is that there is something particular in school-age when it comes to perceptions and preferences about inequality.
-This can relate to development aspects (Kim & Gewirtz, 2019, Rizzo & Killen, 2020) but it could be something specific to the Chilean case, which requires further comparative as well as longitudinal research in the area.
-In any case, teaching about the origins and characteristics of social inequality is not part of the chilean school curriculum, calling for a more dedicated consideration of these topics in areas such as citizensip education as well as in history and social science education.
+This can relate to development aspects [@kim_socioeconomic_2020; @rizzo_children_2020] but it could be something specific to the Chilean case, which requires further comparative as well as longitudinal research in the area.
+In any case, teaching about the origins and characteristics of social inequality is not part of the chilean school curriculum, calling for a more dedicated consideration of these topics in areas such as citizenship education as well as in history and social science education.
 
@@
 We found contrary evidence for the association between status, school performance, and market justice preferences: high status and high achievement at the school level are associated with less market justice preferences.
-This is opposed to the rational-interest argument that suggests that those doing better - both economically and academically- would demand less redistribution (Gonthier,2016; Runst, 2018; Bullock, 2021) and therefore are expected to challenge market mechanisms.
+This is opposed to the rational-interest argument that suggests that those doing better - both economically and academically- would demand less redistribution [@gonthier_parallel_2017; @runst_does_2018; @bullock_education_2021] and therefore are expected to challenge market mechanisms.
 A possible explanation is that education contributes to forming a critical view of markets in redistribution, nevertheless, further research in this area is needed to assert the specific mechanisms that might be playing a role.

### 2024-07-04 02-antecedentes.qmd
Update edit submit

--- 2024-06-28 02-antecedentes.qmd
@@
 The influence of ideologies [@wegener_dominant_1995] and cultural schemas [@homan_being_2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations.
-While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999 ; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations [@lindh_bringing_2023] and even less about children's acceptance or rejection of these market principles.
+While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations [@lindh_bringing_2023] and even less about children's acceptance or rejection of these market principles.
 This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades [@centeno_arc_2012; @harvey_breve_2015], affecting areas such as pensions, health services, and education.

### 2024-07-04 04-analisis.qmd
Update edit submit

@@
 load("input/data/data_rec.RData")
+load("input/data/mlm_models.RData")
+load("input/data/interact.RData")
 
@@
 #| fig-cap: "Market justice preferences in health, pensions and education"
-#| fig-asp: 0.6
+#| fig-asp: 0.4
 #| label: fig-marketjustice
@@
   select(just_salud, just_pension, just_educ) %>% 
-  sjPlot::plot_stackfrq(geom.size = 0.5,
+  sjPlot::plot_stackfrq(geom.size = 0.6,
                         expand.grid =TRUE,
@@
 #| fig-cap: "Social and school meritocracy"
-#| fig-asp: 0.8
+#| fig-asp: 0.6
 #| label: fig-meritocracy
+plot_a <- data_rec %>% 
+  select(inteligencia_esc, esfuerzo_esc) %>% 
+  sjPlot::plot_stackfrq(expand.grid =TRUE,
+                        geom.colors = "PuBu",
+                        show.total = FALSE, 
+                        vjust = c(rep(c("center"),4), "bottom", "top", rep(c("center"), 2)),
+                        legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
+                        axis.labels = c("School talent", "School effort")) +
+  theme(legend.position = "bottom", text = element_text(size = 14),
+        plot.caption = element_text(hjust = 1))
 
-plot<-data_rec %>% 
-  select(merito_soc, esfuerzo_soc, inteligencia_soc, esfuerzo_esc, inteligencia_esc) %>% 
-  sjPlot::plot_stackfrq(geom.size = 0.5,
+plot_b <- data_rec %>% 
+  select(merito_soc, inteligencia_soc, esfuerzo_soc) %>% 
+  sjPlot::plot_stackfrq(geom.size = 0.6,
                         expand.grid =TRUE,
@@
                         show.total = FALSE, 
-                      vjust=c(rep(c("center"),12), "bottom", "top", rep(c("center"), 6)),
+                        vjust = "center",
                         legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
-                      axis.labels = c("Deservingness", "Social effort", "Social talent", "School effort", "School talent")) +
-  #theme_light(base_size = 16) +
-  theme(legend.position="bottom", text = element_text(size = 14),
+                        axis.labels = c("Deservingness", "Social talent", "Social effort")) +
+  theme(legend.position = "bottom", text = element_text(size = 14),
         plot.caption = element_text(hjust = 1))
 
-plot + 
-  annotate("text", x = 5.5, y = 0, label = "School meritocracy", size = 5, hjust = 0, vjust = 1) +
-  annotate("text", x = 3.5, y = 0, label = "Social meritocracy", size = 5, hjust = 0, vjust = 1) +
-       labs(caption = "Source: own elaboration based on First Study of Civic Education 2017")
-```
-
-At the societal level, @fig-meritocracy shows that a majority of respondents agree or strongly agree that individuals are rewarded for their talents (63.7%) and efforts (62%).
-Nevertheless, a notable proportion perceives that both effort and talent are not adequately rewarded in society, with 38.1% and 36.3% respectively.
-The perception of whether people get what they deserve in society is more divided; while a slim majority (56.5% combining agree and strongly agree) agrees, a significant minority (43.6%) disagrees, indicating that this is the most contentious item among respondents.
-There is general agreement that people are rewarded for talent and effort in society, but the consensus is weaker compared to school meritocracy.
-
-@fig-bivariate shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram in @fig-hypotheses).
-On the left, we observe the social meritocracy diagrams, while on the right the school meritocracy diagrams are shown.
-For the three variables of perception of meritocracy in society the relationship is clear, since the average of market justice preferences increases the more there is agreement that people are rewarded for their effort, merit, and talent.
-This relationship needs to be clarified in the case of the variables of perception of meritocracy at school.
-To the extent that there is more agreement that the perception of talent is essential for obtaining good grades, the average of market justice preferences increases, but this relationship is not as clear as with the variables of meritocracy in society.
-In addition, the graph does not show a clear trend in the relationship between the perception that effort is essential to obtain good grades and market justice preferences.
-
-```{r echo=FALSE}
-#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
-#| fig-asp: 1.22
-#| label: fig-bivariate
-
-plot_soc_effort <- soc_effort %>% 
-  mutate(justice = str_to_title(justice),
-         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
-  arrange(esfuerzo_soc, justice) %>% 
-  ggplot(aes(x = esfuerzo_soc, y = mean)) +
-  geom_point(aes(color = justice, shape = justice), size = 2) +
-  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
-  scale_x_continuous(limits = c(1, 4),
-                     breaks = seq(1, 4, by = 1),
-                     label = c("Strongly\n disagree", 
-                               "Disagree", 
-                               "Agree", 
-                               "Strongly\n agree")) +
-  scale_color_brewer(palette = "Paired") +
-  labs(x = "Social effort",
-       y = "Market justice preferences") +
-  theme(legend.position = "none",
-        text = element_text(size = 11))
-
-plot_deservingness <- deservingness %>% 
-   mutate(justice = str_to_title(justice),
-         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
-  arrange(merito_soc, justice) %>% 
-  ggplot(aes(x = merito_soc, y = mean)) +
-  geom_point(aes(color = justice, shape = justice), size = 2) +
-  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
-  scale_x_continuous(limits = c(1, 4),
-                     breaks = seq(1, 4, by = 1),
-                     label = c("Strongly\n disagree", 
-                               "Disagree", 
-                               "Agree", 
-                               "Strongly\n agree")) +
-  scale_color_brewer(palette = "Paired") +
-  labs(x = "Deservingness",
-       y = "Market justice preferences") +
-  theme(legend.position = "none",
-        text = element_text(size = 11))
-
-plot_soc_talent <- soc_talent %>% 
-   mutate(justice = str_to_title(justice),
-         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
-  arrange(inteligencia_soc, justice) %>%  
-  ggplot(aes(x = inteligencia_soc, y = mean)) +
-  geom_point(aes(color = justice, shape = justice), size = 2) +
-  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
-  scale_x_continuous(limits = c(1, 4),
-                     breaks = seq(1, 4, by = 1),
-                     label = c("Strongly\n disagree", 
-                               "Disagree", 
-                               "Agree", 
-                               "Strongly\n agree")) +
-  scale_color_brewer(palette = "Paired") +
-  labs(x = "Social talent",
-       y = "Market justice preferences") +
-  theme(legend.position = "none",
-        text = element_text(size = 11))
-
-plot_school_effort <- school_effort %>% 
-  mutate(justice = str_to_title(justice),
-         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
-  arrange(esfuerzo_esc, justice) %>%  
-  ggplot(aes(x = esfuerzo_esc, y = mean)) +
-  geom_point(aes(color = justice, shape = justice), size = 2) +
-  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
-  scale_x_continuous(limits = c(1, 4),
-                     breaks = seq(1, 4, by = 1),
-                     label = c("Strongly\n disagree", 
-                               "Disagree", 
-                               "Agree", 
-                               "Strongly\n agree")) +
-  scale_color_brewer(palette = "Paired") +
-  labs(x = "School effort",
-       y = "") +
-  theme(legend.position = "none",
-        text = element_text(size = 11))
-
-plot_school_talent <- school_talent %>% 
-  mutate(justice = str_to_title(justice),
-         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
-  arrange(inteligencia_esc, justice) %>% 
-  ggplot(aes(x = inteligencia_esc, y = mean)) +
-  geom_point(aes(color = justice, shape = justice), size = 2) +
-  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
-  scale_x_continuous(limits = c(1, 4),
-                     breaks = seq(1, 4, by = 1),
-                     label = c("Strongly\n disagree", 
-                               "Disagree", 
-                               "Agree", 
-                               "Strongly\n agree")) +
-  scale_color_brewer(palette = "Paired") +
-  labs(x = "School talent",
-       y = "",
-       color = NULL,
-       shape = NULL) +
-  theme(text = element_text(size = 11))
-
-leg <- get_legend(plot_school_talent)
-
-plot_school_talent <- plot_school_talent + theme(legend.position = "none")
-
-ggarrange(plot_soc_effort, plot_school_effort, plot_soc_talent,
-          plot_school_talent, plot_deservingness, leg, ncol = 2, nrow = 3) +
-  labs(caption = "Source: own elaboration based on First Study of Civic Education 2017")
-```
-
-## Multivariate
-
-This section presents the results of the multilevel models.
-We begin by showing the results of the cumulative link mixed models in @tbl-ordinal-reg for the ordinal dependent variables of justice in health, pensions, and education separately, to determine if the effect of the predictors differs across these items.
-Then, we present the results of the linear mixed-effects models for the average index of preference for market justice in @tbl-lineal-reg[^1].
-For the health item, the intraclass correlation (ICC) obtained for the null model (not shown) indicates that 5% of the total variance of this item is associated with specific characteristics of the schools, while it is 3% for the pensions item and 4% for the education item (ICC based on [@hox_multilevel_2010, p.15]).
-For the average index of market justice preferences, the ICC reaches 4%, indicating that overall, only a small percentage of the total variance in students' market justice preferences is associated with school characteristics, thus limiting the possibility of finding effects at this second level of analysis.
-
-[^1]: The tables shows the effects of individual and contextual-level independent variables related to our hypothesis; however, the models also incorporate the control variables utilized in the study for each of the three dependent variables separately and for the average index (complete models are available in the [Appendix](#appendix)).
-
-```{r echo=FALSE, include=FALSE, eval=FALSE}
-sjmisc::frq(data_rec$educacion_rec)
-data_rec <- data_rec %>% mutate(universitario = case_when(educacion_rec=="Ed Basica"~"No",
-                                                          educacion_rec=="Ed Media"~"No",
-                                                          educacion_rec=="Ed Tecnica"~"No",
-                                                          educacion_rec=="Universidad o posgrado"~"Si",
-                                                          educacion_rec=="Ns/Nr"~"Ns/Nr"))
-data_rec$universitario <- factor(data_rec$universitario, levels = c("No", "Si", "Ns/Nr"))
-sjmisc::frq(data_rec$universitario)
-
-data_rec <- data_rec %>% mutate(school_nse = case_when(cod_grupo=="Bajo"~"Bajo",
-                                                          cod_grupo=="Medio bajo"~"Medio",
-                                                          cod_grupo=="Medio"~"Medio",
-                                                          cod_grupo=="Medio alto"~"Medio",
-                                                          cod_grupo=="Alto"~"Alto"))
-sjmisc::frq(data_rec$school_nse)
-data_rec$school_nse <- factor(data_rec$school_nse, levels = c("Bajo", "Medio", "Alto"))
-
-data_pension <- data_rec %>% 
-    dplyr::select(-just_educ, -just_salud, -redistribucion)
-data_pension <- na.omit(data_pension)
-data_pension$just_pension_fact <- as_factor(data_pension$just_pension)
-
-m0_p <- clmm(just_pension_fact ~ 1 + (1 | mrbd), data=data_pension)
-
-performance::icc(m0_p, by_group = T)
-
-m1_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_pension)
-
-m2_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
-    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_pension)
-
-m3_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
-    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
-    school_nse + simce + (1 | mrbd), data=data_pension)
-
-modelos_pension <- list(m1_p, m2_p, m3_p)
-
-# modelos_odd_pension <- map(tbls, odds_fun)
-
-data_salud <- data_rec %>% 
-    dplyr::select(-just_pension, -just_educ, -redistribucion)
-data_salud <- na.omit(data_salud)
-data_salud$just_salud_fact <- as_factor(data_salud$just_salud)
-
-m0_h <- clmm(just_salud_fact ~ 1 + (1 | mrbd), data=data_salud)
-
-performance::icc(m0_h, by_group = T)
-
-m1_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_salud)
-
-m2_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
-    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_salud)
-
-m3_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
-    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
-    school_nse + simce + (1 | mrbd), data=data_salud)
-
-modelos_health <- list(m1_h, m2_h, m3_h)
-
-# modelos_odd_health <- map(tbls, odds_fun)
-
-data_educ <- data_rec %>% 
-    dplyr::select(-just_pension, -just_salud, -redistribucion)
-data_educ <- na.omit(data_educ)
-data_educ$just_educ_fact <- as_factor(data_educ$just_educ)
-
-m0_e <- clmm(just_educ_fact ~ 1 + (1 | mrbd), data=data_educ)
-
-performance::icc(m0_e, by_group = T)
-
-m1_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_educ)
-
-m2_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
-    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_educ)
-
-m3_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
-    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
-    school_nse + simce + (1 | mrbd), data=data_educ)
-
-modelos_educ <- list(m1_e, m2_e, m3_e)
-
-#modelos_odd_educ <- map(tbls, odds_fun)
-
-```
-
-```{r echo=FALSE, results='asis', eval=FALSE}
-#| tbl-cap: "Cumulative link multilevel models of differential acces justificacion of health, pensions, and education"
-#| label: tbl-ordinal-reg
-
-ccoef <- list(
-  inteligencia_esc = "School talent",
-  esfuerzo_esc = "School effort",
-  inteligencia_soc = "Social talent",
-  esfuerzo_soc = "Social effort",
-  merito_soc = "Deservingness",
-  "universitarioSi" = "University or posgraduate",
-  "universitarioNs/Nr" = "Missing",
-  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
-  "school_nseMedio" = "SES Medium",
-  school_nseAlto = "SES High",
-  simceMedio = "Simce Medium",
-  simceAlto = "Simce High"
-)
-
-texreg::htmlreg(l = list(modelos_health[[2]],
-                         modelos_health[[3]],
-                         modelos_pension[[2]],
-                         modelos_pension[[3]],
-                         modelos_educ[[2]],
-                         modelos_educ[[3]]),   
-                caption = NULL,
-                stars = c(0.05, 0.01, 0.001),
-                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses.
-Control variables are included.
-%stars.
-Source: own elaboration based on First Study of Civic Education 2017.",
-                custom.header = list("Health" = 1:2, "Pensions" = 3:4, "Education" = 5:6),
-                custom.model.names = rep(c("Model 1", "Model 2"), 3),
-                custom.coef.map = ccoef,
-                leading.zero = T,
-                scalebox = 0.60,
-                groups = list("Parental education (Ref.= Non-university)" = 6:7,
-                                "Socioeconomic level (Ref.= Low)" = 9:10,
-                                "Achievement score (Ref.= Low)" = 11:12),
-                custom.gof.rows = list("Controls"=rep("Yes",6)),
-                include.loglik = FALSE, include.aic = FALSE,
-                file = "output/tables/tab-ordinalreg.html")
-
-webshot::webshot(url = "output/tables/tab-ordinalreg.html", file = "output/tables/tab-ordinalreg.png")
-```
-
-![](output/tables/tab-ordinalreg.png)
-
-Model 1 in @tbl-ordinal-reg for the three dependent variables of justice in differential access to health, pensions, and education demonstrates a similar trend regarding the effect of the perception of meritocracy on these items, with all effects being statistically significant (*p* \< 0.05).
-Regarding the variables of social meritocracy, the three variables---talent, effort, and deservingness---show that as these increase, the justification for differentiated access to pensions, education, and health also increases.
-In the context of school meritocracy, the effects are mixed: as perceived school talent increases, the justification for differentiated access to these services increases; conversely, as perceived school effort increases, the justification decreases, holding all other variables constant.
-
-Additionally, Model 1 examines family status through parents' educational level and the number of books at home.
-For health, pensions, and education, a negative but statistically non-significant association with parents' education is observed.
-However, having more than 25 books at home shows a negative and statistically significant effect on market justice in health and pensions, suggesting less justification for differentiated access compared to those with fewer books.
-This effect is not seen in education.
-
-Model 2 in @tbl-ordinal-reg incorporates school-level variables.
-The results show that high socioeconomic level schools show greater preferences for market justice in health, pensions, and education compared to low socioeconomic level schools, with statistically significant results (*p* \< 0.05).
-Moreover, the findings suggest that schools with higher academic performance shows negative and statistically significant effects across all three items, suggesting these schools justify market allocation principles less.
-
-```{r echo=FALSE, include=FALSE}
-data_bienestar <- data_bienestar %>% mutate(universitario = case_when(educacion_rec=="Ed Basica"~"No",
-                                                          educacion_rec=="Ed Media"~"No",
-                                                          educacion_rec=="Ed Tecnica"~"No",
-                                                          educacion_rec=="Universidad o posgrado"~"Si",
-                                                          educacion_rec=="Ns/Nr"~"Ns/Nr"))
-data_bienestar$universitario <- factor(data_bienestar$universitario, levels = c("No", "Si", "Ns/Nr"))
-
-sjmisc::frq(data_bienestar$cod_grupo)
-data_bienestar <- data_bienestar %>% mutate(school_nse = case_when(cod_grupo=="Bajo"~"Bajo",
-                                                          cod_grupo=="Medio bajo"~"Medio",
-                                                          cod_grupo=="Medio"~"Medio",
-                                                          cod_grupo=="Medio alto"~"Medio",
-                                                          cod_grupo=="Alto"~"Alto"))
-sjmisc::frq(data_bienestar$school_nse)
-data_bienestar$school_nse <- factor(data_bienestar$school_nse, levels = c("Bajo", "Medio", "Alto"))
-
-reg0_bienestar <- lmer(bienestar ~ 1 + (1 | mrbd), data = data_bienestar)
-
-performance::icc(reg0_bienestar, by_group = T)
-
-reg1_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
-                         inteligencia_soc + esfuerzo_soc + merito_soc + 
-                         acc_tec + internet_rec +
-                         (1 | mrbd), data=data_bienestar)
-
-reg2_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
-                         inteligencia_soc + esfuerzo_soc + merito_soc + 
-                         universitario + libros_rec + acc_tec + internet_rec +
-                         (1 | mrbd), data=data_bienestar)
-
-reg3_bienestar <- lmer(bienestar ~ 1 + mean_educ + cod_depe2 + school_nse +
-                         simce + (1 | mrbd), data=data_bienestar)
-
-reg4_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + 
-                         inteligencia_soc + esfuerzo_esc + inteligencia_esc + 
-                         universitario + libros_rec + acc_tec + internet_rec + 
-                         mean_educ + cod_depe2 + school_nse + 
-                         (1 | mrbd), data=data_bienestar)
-
-reg5_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + 
-                         inteligencia_soc + esfuerzo_esc + inteligencia_esc + 
-                         universitario + libros_rec + acc_tec + internet_rec + 
-                         mean_educ + cod_depe2 + school_nse + simce + 
-                         (1 | mrbd), data=data_bienestar)
-
-test01 <- anova(reg0_bienestar, reg1_bienestar, test = "Chisq")
-test02 <- anova(reg0_bienestar, reg2_bienestar, test = "Chisq")
-test03 <- anova(reg0_bienestar, reg3_bienestar, test = "Chisq")
-test04 <- anova(reg0_bienestar, reg4_bienestar, test = "Chisq")
-test05 <- anova(reg0_bienestar, reg5_bienestar, test = "Chisq")
-
-test.pvalues1 <- test01$`Pr(>Chisq)`[2]
-test.pvalues2 <- test02$`Pr(>Chisq)`[2]
-test.pvalues3 <- test03$`Pr(>Chisq)`[2]
-test.pvalues4 <- test04$`Pr(>Chisq)`[2]
-test.pvalues5 <- test05$`Pr(>Chisq)`[2]
-
-deviance1 <- test01$deviance[2]
-deviance2 <- test02$deviance[2]
-deviance3 <- test03$deviance[2]
-deviance4 <- test04$deviance[2]
-deviance5 <- test05$deviance[2]
-```
-
-```{r echo=FALSE, results='asis'}
-#| tbl-cap: "Linear mixed-effects models for meritocracy perception and market justice preferences"
-#| label: tbl-lineal-reg
-
-ccoef <- list(
-  "(Intercept)" = "Constant",
-  inteligencia_esc = "School talent",
-  esfuerzo_esc = "School effort",
-  inteligencia_soc = "Social talent",
-  esfuerzo_soc = "Social effort",
-  merito_soc = "Deservingness",
-  "universitarioSi" = "University or posgraduate",
-  "universitarioNs/Nr" = "Missing",
-  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
-  "school_nseMedio" = "SES Medium",
-  school_nseAlto = "SES High",
-  simceMedio = "Simce Medium",
-  simceAlto = "Simce High"
-)
-
-texreg::htmlreg(list(reg1_bienestar, reg2_bienestar, reg4_bienestar, reg5_bienestar),
-                custom.model.names = c("Model 1",
-                                       "Model 2",
-                                       "Model 3",
-                                       "Model 4"),
-                caption = NULL,
-                stars = c(0.05, 0.01, 0.001),
-                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses.
-Control variables are included.
-%stars.
-Source: own elaboration based on First Study of Civic Education 2017.",
-                custom.coef.map = ccoef,
-                leading.zero = T,
-                scalebox = 0.60,
-                groups = list("Parental education (Ref.= Non-university)" = 7:8,
-                              "Socioeconomic level (Ref.= Low)" = 10:11,
-                              "Achievement score (Ref.= Low)" = 12:13),
-                custom.gof.rows = list("Controls"=rep("Yes",4),
-                                       "Deviance" = c(deviance1,
-                                                    deviance2,
-                                                    deviance4,
-                                                    deviance5),
-                                       "Deviance Test (p)" = c(test.pvalues1,
-                                                             test.pvalues2,
-                                                             test.pvalues4,
-                                                             test.pvalues5)),
-                include.loglik = FALSE,include.aic = FALSE)
-
-```
-
-@tbl-lineal-reg presents the results of linear mixed-effects models for the market justice preference index.
-Model 1 shows that higher perceived societal meritocracy, through effort, talent, and deservingness, is positively and significantly associated with market justice preferences.
-School meritocracy perceptions yield mixed effects: perceived talent increases market justice preference, while perceived effort decreases it, both statistically significant (*p* \< 0.001).
-Model 2 includes family status, finding no statistically significant effects from parents' education or the number of books at home, showing positive direction for university or posgraduate, and negative direction for books at home.
-
-Regarding school characteristics, Models 3 and 4 introduce socioeconomic level and academic performance.
-The results of Model 3 suggest statistically significant relationship between school socioeconomic levels and market justice preferences, specifically those school with high socieconomic level ($\beta$ = 0.37) compared to low socioeconomic level.
-However, Model 4 indicates that schools with medium ($\beta$ = -0.11) and high ($\beta$ = -0.26) academic performance have lower market justice preferences compared to low-performing schools, with these effects being statistically significant when controlling for other variables.
-
-```{r echo=FALSE}
-#| tbl-cap: "Interactions effects"
-#| label: tbl-interact
-
-pacman::p_load(kableExtra)
-load("input/data/interact.RData")
-
-collapse_rows[c(-4),] %>% 
-  as_tibble() %>% 
-  kbl(.,
-      align = "lccccc", booktabs = T,
-      col.names =  c("Variable", 
-                     "School talent", 
-                     "School effort",
-                     "Social talent",
-                     "Social effort", 
-                     "Deservingness")) %>% 
-  add_header_above(c(" " = 1,
-                     "Market Justice Preferences"=5), 
-                   bold = T) %>% 
-  kable_styling(latex_options = c("scale_down","HOLD_position"),
-                full_width = T) %>% 
-  column_spec(1, bold = T, width = "4cm") %>%
-  column_spec(column = 2:6,width = "1.5cm") %>%
-  collapse_rows(columns = 1:2, valign = "middle") %>% 
-  kableExtra::add_footnote(label = "Note: ***p < 0.001, **p < 0.01, *p < 0.05. Source: own elaboration based on First Study of Civic Education 2017.", 
-                           notation = "none")
-
-```
-
-@tbl-interact displays the interaction terms related to our hypothesis.
-The results suggest that students' family background moderates the relationship between their meritocratic perceptions and their justification of market allocation in social services.
-Specifically, the relationship between social effort and market justice preferences is less positive for students whose parents have university or postgraduate education.
-Lower-status students (measured by parental education) prefer more market justice when they strongly adhere to meritocratic perceptions of deservingness.
-A similar moderating effect is seen with family cultural capital: the relationship between effort and market justice preferences is less positive for students with more than 25 books at home.
-
-At the school level, we observe that school status and students' meritocratic beliefs moderate their preferences for market justice.
-High-status schools show less preference for market justice when their students, on average, have a greater perception of meritocracy, as measured by effort, deservingness, and talent.
-Conversely, low-achieving schools prefer more market justice when their students have stronger meritocratic beliefs.
-@fig-interaction illustrates this moderating effect of school achievement on the relationship between deservingness and market justice.
-No moderating effects were observed at the school level for students' meritocratic beliefs regarding effort or talent in achieving good grades.
-
-```{r echo=FALSE, include=FALSE}
-load("input/data/data_rec.RData")
-data_rec <- data_rec %>% rowwise() %>%  mutate(educacion_rec2 = case_when(educacion_rec=="Ed Basica"~"Low level",
-                                                 educacion_rec=="Ed Media"~"Low level",
-                                                 educacion_rec=="Ed Tecnica"~"Low level",
-                                                 educacion_rec=="Universidad o posgrado"~"University or Postgraduate",
-                                                 educacion_rec=="Ns/Nr"~"Missing"
-                                                 ))
-data_rec$educacion_rec2 <- factor(data_rec$educacion_rec2, levels = c("Low level", "University or Postgraduate", "Missing"))
-
-data_rec <- data_rec %>% rowwise() %>%  mutate(cod_grupo2  = case_when(cod_grupo=="Bajo"~"Bajo",
-                                                                       cod_grupo=="Medio bajo"~"Bajo",
-                                                                       cod_grupo=="Medio"~"Bajo",
-                                                                       cod_grupo=="Medio alto"~"Alto",
-                                                                       cod_grupo=="Alto"~"Alto"))
-data_rec$cod_grupo2 <- factor(data_rec$cod_grupo2, levels = c("Bajo", "Alto"))
-data_rec$simce <- factor(data_rec$simce, labels = c("Low", "Medium", "High"))
-
-data_bienestar <- na.omit(data_rec)
-#data_bienestar <- data_bienestar %>% mutate(pension_rec = 5-just_pension,
-#                                            educ_rec = 5-just_educ,
-#                                            salud_rec = 5-just_salud)
-data_bienestar <- data_bienestar %>% rowwise() %>% dplyr::mutate(bienestar = mean(c(just_educ, just_salud, just_pension), na.omit=TRUE))
-data_bienestar$bienestar <- round(data_bienestar$bienestar, 2)
-
-reg26.2_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + merito_soc*simce + internet_rec + (1 + simce | mrbd), data=data_bienestar)
-
-plot_interact <- ggpredict(model = reg26.2_bienestar, terms = c("merito_soc", "simce")) %>% 
-  filter(group != "Medium") %>% 
-  plot() +
-  ylim(1,4) +
-  labs(title = NULL,
-       x = "Deservingness", 
-       y ="Market Justice Preferences", 
-       colour = "School Simce",
-       caption = "Source: own elaboration based on First Study of Civic Education 2017")
-```
-
-```{r echo=FALSE, warning=FALSE}
-#| fig-cap: "Interaction between deservingness and justification of inequality by school SIMCE achievement"
-#| label: fig-interaction
-
-plot_interact +
-  theme_bw()
-
-```
----
-author: "Equipo Edumer"
-bibliography: "input/bib/just-ineq-merit.bib"
-csl: "input/bib/apa6.csl"
----
-
-# Results
-
-## Descriptive
-
-```{r packages, echo=FALSE}
-pacman::p_load(tidyverse,
-               ggpubr, 
-               lme4,
-               ordinal,
-               sjPlot, 
-               texreg,
-               kableExtra,
-               easystats,
-               ggeffects)
-
-options(scipen=999)
-rm(list = ls())
-```
-
-```{r data, echo=FALSE}
-
-load("input/data/data_index.RData")
-load("input/data/deservingness.RData")
-load("input/data/soc_effort.RData")
-load("input/data/soc_talent.RData")
-load("input/data/school_effort.RData")
-load("input/data/school_talent.RData")
-load("input/data/data_rec.RData")
-
-```
-
-```{r functions, echo=FALSE}
-
-odds_fun <- function(modelos) {
-   or <- texreg::extract(modelos)
-   or@coef <- exp(or@coef)
-   return(or)
-}
-
-miles <- function(x) {
-  format(round(as.numeric(x),0), big.mark = ".")
-}
-
-decimales <- function(x) {
-  format(round(as.numeric(x), 2), decimal.mark = ",")
-}
-
-# set theme
-
-theme_set(theme_bw())
-
-```
-
-@fig-marketjustice illustrates the distribution of responses to the three items on market justice preferences (healthcare, pensions, and education), indicating that most respondents tend to disagree or strongly disagree with the idea that it is just for people with higher incomes to have access to better services in these areas.
-The strongest opposition is observed in the healthcare sector, as the majority of respondents (66.4%) are against the idea that higher-income individuals should have access to better healthcare services.
-Similarly, regarding the concept of justice related to access to better pensions based on income, the level of disagreement reaches 63.0%.
-This decreases significantly in the field of education, as the percentage of respondents who disagree or strongly disagree with those with higher incomes having access to better education decreases to 58.8%.
-Only a small proportion agree that access to healthcare should be conditional on income (23.6%), and a minority strongly agree (10%), as is the case with pensions.
-However, this increases in the case of education, being the area where there is a significant proportion of respondents who agree or strongly agree with access based on market justice (41.2%).
-
-```{r echo=FALSE}
-#| fig-cap: "Market justice preferences in health, pensions and education"
-#| fig-asp: 0.6
-#| label: fig-marketjustice
-
-data_rec %>% 
-  select(just_salud, just_pension, just_educ) %>% 
-  sjPlot::plot_stackfrq(geom.size = 0.5,
-                        expand.grid =TRUE,
-                      geom.colors = "PuBu",
-                      show.total = FALSE, 
-                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
-                      axis.labels = c("Health", "Pensions", "Education")) +
- # theme_light(base_size = 16) +
-  theme(legend.position="bottom", text = element_text(size = 14),
-        plot.caption = element_text(hjust = 1)) +
+ggarrange(plot_a, plot_b, 
+          common.legend = T, 
+          ncol = 1,
+          legend = "bottom", 
+          labels = c("School", "Society")
+          ) +
+  theme(plot.caption = element_text(hjust = 1, size = 11)) +
   labs(caption = "Source: own elaboration based on First Study of Civic Education 2017")
 
-```
-
-Regarding the perception of meritocracy, @fig-meritocracy displays the frequency distribution of five items related to the dimensions of school and society.
-At the school level, there is a strong belief that both effort and talent are important for achieving good grades, with especially high agreement for effort.
-Specifically, while 78.6% agree or strongly agree on the importance of talent for obtaining good grades, this proportion rises to 96.8% for effort, indicating that respondents value effort more than talent in this context.
-
-```{r echo=FALSE}
-#| fig-cap: "Social and school meritocracy"
-#| fig-asp: 0.8
-#| label: fig-meritocracy
-
-plot<-data_rec %>% 
-  select(merito_soc, esfuerzo_soc, inteligencia_soc, esfuerzo_esc, inteligencia_esc) %>% 
-  sjPlot::plot_stackfrq(geom.size = 0.5,
-                        expand.grid =TRUE,
-                      geom.colors = "PuBu",
-                      show.total = FALSE, 
-                      vjust=c(rep(c("center"),12), "bottom", "top", rep(c("center"), 6)),
-                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
-                      axis.labels = c("Deservingness", "Social effort", "Social talent", "School effort", "School talent")) +
-  #theme_light(base_size = 16) +
-  theme(legend.position="bottom", text = element_text(size = 14),
-        plot.caption = element_text(hjust = 1))
-
-plot + 
-  annotate("text", x = 5.5, y = 0, label = "School meritocracy", size = 5, hjust = 0, vjust = 1) +
-  annotate("text", x = 3.5, y = 0, label = "Social meritocracy", size = 5, hjust = 0, vjust = 1) +
-       labs(caption = "Source: own elaboration based on First Study of Civic Education 2017")
 ```
@@
           plot_school_talent, plot_deservingness, leg, ncol = 2, nrow = 3) +
+ theme(plot.caption = element_text(hjust = 1, size = 11)) +
   labs(caption = "Source: own elaboration based on First Study of Civic Education 2017")
@@
 
-For the health item, the intraclass correlation (ICC) obtained for the null model (not shown) indicates that 5% of the total variance of this item is associated with specific characteristics of the schools, while it is 3% for the pensions item and 4% for the education item (ICC based on [@hox_multilevel_2010, p.15]).
+For the health item, the intraclass correlation (ICC) obtained for the null model (not shown) indicates that 5% of the total variance of this item is associated with specific characteristics of the schools, while it is 3% for the pensions item and 4% for the education item [ICC based on @hox_multilevel_2010, p.15].
 For the average index of market justice preferences, the ICC reaches 4%, indicating that overall, only a small percentage of the total variance in students' market justice preferences is associated with school characteristics, thus limiting the possibility of finding effects at this second level of analysis.
 
-[^1]: The table shows the effects of individual and contextual-level independent variables related to our hypotheses; however, the models also incorporate the control variables for each of the three dependent variables separately and for the average index (complete models are available in the [Appendix](#appendix)).
+[^1]: The table shows the effects of individual and contextual-level independent variables related to our hypotheses; however, the models also incorporate the control variables for each of the three dependent variables separately and for the average index (complete models are available in the [Supplementary Material](#appendix)).
 
-```{r echo=FALSE, include=FALSE, eval=FALSE}
-sjmisc::frq(data_rec$educacion_rec)
-data_rec <- data_rec %>% mutate(universitario = case_when(educacion_rec=="Ed Basica"~"No",
-                                                          educacion_rec=="Ed Media"~"No",
-                                                          educacion_rec=="Ed Tecnica"~"No",
-                                                          educacion_rec=="Universidad o posgrado"~"Si",
-                                                          educacion_rec=="Ns/Nr"~"Ns/Nr"))
-data_rec$universitario <- factor(data_rec$universitario, levels = c("No", "Si", "Ns/Nr"))
-sjmisc::frq(data_rec$universitario)
-
-data_rec <- data_rec %>% mutate(school_nse = case_when(cod_grupo=="Bajo"~"Bajo",
-                                                          cod_grupo=="Medio bajo"~"Medio",
-                                                          cod_grupo=="Medio"~"Medio",
-                                                          cod_grupo=="Medio alto"~"Medio",
-                                                          cod_grupo=="Alto"~"Alto"))
-sjmisc::frq(data_rec$school_nse)
-data_rec$school_nse <- factor(data_rec$school_nse, levels = c("Bajo", "Medio", "Alto"))
-
-data_pension <- data_rec %>% 
-    dplyr::select(-just_educ, -just_salud, -redistribucion)
-data_pension <- na.omit(data_pension)
-data_pension$just_pension_fact <- as_factor(data_pension$just_pension)
-
-m0_p <- clmm(just_pension_fact ~ 1 + (1 | mrbd), data=data_pension)
-
-performance::icc(m0_p, by_group = T)
-
-m1_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_pension)
-
-m2_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
-    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_pension)
-
-m3_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
-    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
-    school_nse + simce + (1 | mrbd), data=data_pension)
-
-modelos_pension <- list(m1_p, m2_p, m3_p)
-
-# modelos_odd_pension <- map(tbls, odds_fun)
-
-data_salud <- data_rec %>% 
-    dplyr::select(-just_pension, -just_educ, -redistribucion)
-data_salud <- na.omit(data_salud)
-data_salud$just_salud_fact <- as_factor(data_salud$just_salud)
-
-m0_h <- clmm(just_salud_fact ~ 1 + (1 | mrbd), data=data_salud)
-
-performance::icc(m0_h, by_group = T)
-
-m1_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_salud)
-
-m2_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
-    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_salud)
-
-m3_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
-    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
-    school_nse + simce + (1 | mrbd), data=data_salud)
+```{r echo=FALSE, results='asis'}
+#| tbl-cap: "Cumulative link multilevel models of differential acces justificacion of health, pensions, and education"
+#| label: tbl-ordinal-reg
 
 modelos_health <- list(m1_h, m2_h, m3_h)
-
-# modelos_odd_health <- map(tbls, odds_fun)
-
-data_educ <- data_rec %>% 
-    dplyr::select(-just_pension, -just_salud, -redistribucion)
-data_educ <- na.omit(data_educ)
-data_educ$just_educ_fact <- as_factor(data_educ$just_educ)
-
-m0_e <- clmm(just_educ_fact ~ 1 + (1 | mrbd), data=data_educ)
-
-performance::icc(m0_e, by_group = T)
-
-m1_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_educ)
-
-m2_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
-    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_educ)
-
-m3_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
-    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
-    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
-    school_nse + simce + (1 | mrbd), data=data_educ)
-
+modelos_pension <- list(m1_p, m2_p, m3_p)
 modelos_educ <- list(m1_e, m2_e, m3_e)
-
-#modelos_odd_educ <- map(tbls, odds_fun)
-
-```
-
-```{r echo=FALSE, results='asis', eval=FALSE}
-#| tbl-cap: "Cumulative link multilevel models of differential acces justificacion of health, pensions, and education"
-#| label: tbl-ordinal-reg
 
@@
 
-texreg::htmlreg(l = list(modelos_health[[2]],
-                         modelos_health[[3]],
-                         modelos_pension[[2]],
-                         modelos_pension[[3]],
-                         modelos_educ[[2]],
-                         modelos_educ[[3]]),   
+texreg::texreg(
+  l = list(modelos_health[[2]], modelos_health[[3]], modelos_pension[[2]], modelos_pension[[3]], modelos_educ[[2]], modelos_educ[[3]]),
   caption = NULL,
   stars = c(0.05, 0.01, 0.001),
-                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses.
+  custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses.
 Control variables are included.
-%stars.
-Source: own elaboration based on First Study of Civic Education 2017.",
+%stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",threeparttable = T,
   custom.header = list("Health" = 1:2, "Pensions" = 3:4, "Education" = 5:6),
@@
   custom.coef.map = ccoef,
-                leading.zero = T,
-                scalebox = 0.60,
-                groups = list("Parental education (Ref.= Non-university)" = 6:7,
-                                "Socioeconomic level (Ref.= Low)" = 9:10,
-                                "Achievement score (Ref.= Low)" = 11:12),
-                custom.gof.rows = list("Controls"=rep("Yes",6)),
-                include.loglik = FALSE, include.aic = FALSE,
-                file = "output/tables/tab-ordinalreg.html")
+  leading.zero = TRUE,
+  use.packages = FALSE,
+  booktabs = TRUE,
+  scalebox = 0.8,
+  groups = list("Parental education (Ref.= Non-university)" = 6:7, "Socioeconomic level (Ref.= Low)" = 9:10, "Achievement score (Ref.= Low)" = 11:12),
+  custom.gof.rows = list("Controls" = rep("Yes", 6)),
+  include.loglik = FALSE,
+  include.aic = FALSE,
+)
 
-webshot::webshot(url = "output/tables/tab-ordinalreg.html", file = "output/tables/tab-ordinalreg.png")
 ```
-
-![](output/tables/tab-ordinalreg.png)
 
@@
 Moreover, and contrary to our hypothesis, schools with higher academic performance show negative and statistically significant effects across the three items.
-
-```{r echo=FALSE, include=FALSE}
-data_bienestar <- data_bienestar %>% mutate(universitario = case_when(educacion_rec=="Ed Basica"~"No",
-                                                          educacion_rec=="Ed Media"~"No",
-                                                          educacion_rec=="Ed Tecnica"~"No",
-                                                          educacion_rec=="Universidad o posgrado"~"Si",
-                                                          educacion_rec=="Ns/Nr"~"Ns/Nr"))
-data_bienestar$universitario <- factor(data_bienestar$universitario, levels = c("No", "Si", "Ns/Nr"))
-
-sjmisc::frq(data_bienestar$cod_grupo)
-data_bienestar <- data_bienestar %>% mutate(school_nse = case_when(cod_grupo=="Bajo"~"Bajo",
-                                                          cod_grupo=="Medio bajo"~"Medio",
-                                                          cod_grupo=="Medio"~"Medio",
-                                                          cod_grupo=="Medio alto"~"Medio",
-                                                          cod_grupo=="Alto"~"Alto"))
-sjmisc::frq(data_bienestar$school_nse)
-data_bienestar$school_nse <- factor(data_bienestar$school_nse, levels = c("Bajo", "Medio", "Alto"))
-
-reg0_bienestar <- lmer(bienestar ~ 1 + (1 | mrbd), data = data_bienestar)
-
-performance::icc(reg0_bienestar, by_group = T)
-
-reg1_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
-                         inteligencia_soc + esfuerzo_soc + merito_soc + 
-                         acc_tec + internet_rec +
-                         (1 | mrbd), data=data_bienestar)
-
-reg2_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
-                         inteligencia_soc + esfuerzo_soc + merito_soc + 
-                         universitario + libros_rec + acc_tec + internet_rec +
-                         (1 | mrbd), data=data_bienestar)
-
-reg3_bienestar <- lmer(bienestar ~ 1 + mean_educ + cod_depe2 + school_nse +
-                         simce + (1 | mrbd), data=data_bienestar)
-
-reg4_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + 
-                         inteligencia_soc + esfuerzo_esc + inteligencia_esc + 
-                         universitario + libros_rec + acc_tec + internet_rec + 
-                         mean_educ + cod_depe2 + school_nse + 
-                         (1 | mrbd), data=data_bienestar)
-
-reg5_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + 
-                         inteligencia_soc + esfuerzo_esc + inteligencia_esc + 
-                         universitario + libros_rec + acc_tec + internet_rec + 
-                         mean_educ + cod_depe2 + school_nse + simce + 
-                         (1 | mrbd), data=data_bienestar)
-
-test01 <- anova(reg0_bienestar, reg1_bienestar, test = "Chisq")
-test02 <- anova(reg0_bienestar, reg2_bienestar, test = "Chisq")
-test03 <- anova(reg0_bienestar, reg3_bienestar, test = "Chisq")
-test04 <- anova(reg0_bienestar, reg4_bienestar, test = "Chisq")
-test05 <- anova(reg0_bienestar, reg5_bienestar, test = "Chisq")
-
-test.pvalues1 <- test01$`Pr(>Chisq)`[2]
-test.pvalues2 <- test02$`Pr(>Chisq)`[2]
-test.pvalues3 <- test03$`Pr(>Chisq)`[2]
-test.pvalues4 <- test04$`Pr(>Chisq)`[2]
-test.pvalues5 <- test05$`Pr(>Chisq)`[2]
-
-deviance1 <- test01$deviance[2]
-deviance2 <- test02$deviance[2]
-deviance3 <- test03$deviance[2]
-deviance4 <- test04$deviance[2]
-deviance5 <- test05$deviance[2]
-```
 
@@
 
-texreg::htmlreg(list(reg1_bienestar, reg2_bienestar, reg4_bienestar, reg5_bienestar),
+texreg::texreg(list(reg1_bienestar, reg2_bienestar, reg4_bienestar, reg5_bienestar),
                 custom.model.names = c("Model 1",
@@
                 stars = c(0.05, 0.01, 0.001),
-                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses.
+                custom.coef.map = ccoef,
+               custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses.
 Control variables are included.
-%stars.
-Source: own elaboration based on First Study of Civic Education 2017.",
-                custom.coef.map = ccoef,
+%stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",threeparttable = T,
                 leading.zero = T,
-                scalebox = 0.60,
+                use.packages = F,
+                booktabs = TRUE,
+                scalebox = 0.85,
                 groups = list("Parental education (Ref.= Non-university)" = 7:8,
@@
 #| label: tbl-interact
-
-pacman::p_load(kableExtra)
-load("input/data/interact.RData")
 
@@
 
-```{r echo=FALSE, include=FALSE}
-load("input/data/data_rec.RData")
-data_rec <- data_rec %>% rowwise() %>%  mutate(educacion_rec2 = case_when(educacion_rec=="Ed Basica"~"Low level",
-                                                 educacion_rec=="Ed Media"~"Low level",
-                                                 educacion_rec=="Ed Tecnica"~"Low level",
-                                                 educacion_rec=="Universidad o posgrado"~"University or Postgraduate",
-                                                 educacion_rec=="Ns/Nr"~"Missing"
-                                                 ))
-data_rec$educacion_rec2 <- factor(data_rec$educacion_rec2, levels = c("Low level", "University or Postgraduate", "Missing"))
+```{r echo=FALSE, warning=FALSE, message=FALSE}
+#| fig-cap: "Interaction between deservingness and justification of inequality by school SIMCE achievement"
+#| label: fig-interaction
 
-data_rec <- data_rec %>% rowwise() %>%  mutate(cod_grupo2  = case_when(cod_grupo=="Bajo"~"Bajo",
-                                                                       cod_grupo=="Medio bajo"~"Bajo",
-                                                                       cod_grupo=="Medio"~"Bajo",
-                                                                       cod_grupo=="Medio alto"~"Alto",
-                                                                       cod_grupo=="Alto"~"Alto"))
-data_rec$cod_grupo2 <- factor(data_rec$cod_grupo2, levels = c("Bajo", "Alto"))
-data_rec$simce <- factor(data_rec$simce, labels = c("Low", "Medium", "High"))
-
-data_bienestar <- na.omit(data_rec)
-#data_bienestar <- data_bienestar %>% mutate(pension_rec = 5-just_pension,
-#                                            educ_rec = 5-just_educ,
-#                                            salud_rec = 5-just_salud)
-data_bienestar <- data_bienestar %>% rowwise() %>% dplyr::mutate(bienestar = mean(c(just_educ, just_salud, just_pension), na.omit=TRUE))
-data_bienestar$bienestar <- round(data_bienestar$bienestar, 2)
-
-reg26.2_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + merito_soc*simce + internet_rec + (1 + simce | mrbd), data=data_bienestar)
-
-plot_interact <- ggpredict(model = reg26.2_bienestar, terms = c("merito_soc", "simce")) %>% 
+ggpredict(model = reg26.2_bienestar, terms = c("merito_soc", "simce")) %>% 
   filter(group != "Medium") %>% 
@@
        colour = "School Simce",
-       caption = "Source: own elaboration based on First Study of Civic Education 2017")
-```
-
-```{r echo=FALSE, warning=FALSE}
-#| fig-cap: "Interaction between deservingness and justification of inequality by school SIMCE achievement"
-#| label: fig-interaction
-
-plot_interact +
+       caption = "Source: own elaboration based on First Study of Civic Education 2017") +
   theme_bw()

### 2024-07-04 07-supplementary-material.qmd
Update edit submit

--- 2024-06-18 07-anexos.qmd
@@
 author: "Equipo EDUMER"
-format: pdf
-editor: visual
+bibliography: "input/bib/just-ineq-merit.bib"
+csl: "input/bib/apa6.csl"
 ---
 
-## Appendix {#appendix}
+## Supplementary Material {#appendix}
 
@@
 |                                                                                                                    |                                                                                                                                              |
-|--------------------------------|---------------------------------------|
+|---------------------------------|---------------------------------------|
 | **English translation**                                                                                            | **Original Spanish**                                                                                                                         |

### 2024-07-04 back_matter.qmd
Update edit submit

---
title: "The Socialization of Meritocracy and Market Justice Preferences at School"
subtitle: "Back Matter"
format:
  hikmah-pdf: 
    indent: false
  hikmah-manuscript-pdf: 
    indent: false
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/apa6.csl"
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
papersize: letter
colorlinks: true
fig-height: 4
fig-width: 7.5
header-includes:
  - \usepackage{threeparttable}
---

{{< include 07-supplementary-material.qmd >}}

## Acknowledgments

The authors acknowledge the support of ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) grant ANID/FONDAP 1513009.

## Authors' Contributions

The following statements should be used "Conceptualization, X.X. and Y.Y.; Methodology, X.X.; Software, X.X.; Validation, X.X., Y.Y. and Z.Z.; Formal Analysis, X.X.; Investigation, X.X.; Resources, X.X.; Data Curation, X.X.; Writing – Original Draft Preparation, X.X.; Writing – Review & Editing, X.X.; Visualization, X.X.; Supervision, X.X.; Project Administration, X.X.; Funding Acquisition, Y.Y.”

## Conflict of Interest

The authors report no competing interests to declare.

## Funding

This research received no external funding.

## Data Avaibility Statement

The original data presented in the study are openly available in <https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/>

Access the data, code, and analysis presented in the study are openly available in <https://github.com/educacion-meritocracia/market-justice-preferences-school>.

## References

### 2024-07-04 cover_letter.qmd
Update edit submit

---
title: "The Socialization of Meritocracy and Market Justice Preferences at School"
subtitle: "Cover Letter"
format:
  hikmah-pdf: 
    indent: false
  hikmah-manuscript-pdf: 
    indent: false
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/apa6.csl"
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
papersize: letter
colorlinks: true
fig-height: 4
fig-width: 7.5
header-includes:
  - \usepackage{threeparttable}
---

This article evaluates

## Authors Identification

Juan Carlos Castillo, Department of Sociology, Universidad de Chile, Santiago de Chile[^*].  
ORCID: 0000-0003-1265-7854

Mauricio Salgado, School of Sociology, Universidad Andrés Bello.  
ORCID: 0000-0002-8108-6154

Kevin Carrasco, Center for Social Conflict and Cohesion Studies.  
ORCID: 0000-0002-9348-8361

Andreas Laffert, Institute of Sociology, Pontificia Universidad Católica de Chile.  
ORCID: 0000-0002-9008-2454

## Exclusive Submission Statement

The authors confirm that this article has been submitted exclusively to this journal.

[^*]: Corresponding author (juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) - Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile). The authors acknowledge the support of ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) grant ANID/FONDAP 1513009.

### 2024-07-04 paper.qmd
add quarto now extension
Update edit submit

--- 2024-05-15 paper.qmd
@@
 ---
+title: "The socialization of meritocracy and market justice preferences at school"
 format:
-  html:
-    keep-tex: true
-editor: visual
+  hikmah-pdf: 
+    indent: false
+  hikmah-manuscript-pdf: 
+    indent: false
+author:
+  - name: Juan Carlos Castillo
+    affiliations: Department of Sociology, Universidad de Chile
+    email: juancastillov@uchile.cl
+    attributes:
+      corresponding: true
+  - name: Mauricio Salgado
+    affiliations: School of Sociology, Universidad Andrés Bello
+  - name: Kevin Carrasco
+    affiliations: Center for Social Conflict and Cohesion Studies
+  - name: Andreas Laffert
+    affiliations: Institute of Sociology, Universidad Católica de Chile
+abstract: |
+  This article evaluates novel approaches to do some really important things.
+keywords: 
+  - open-source 
+  - scientific publishing
+  - reproducible research
 bibliography: "input/bib/just-ineq-merit.bib"
@@
 link-citations: true
+linestretch: 1.15       
 mainfont: Times New Roman
+fontsize: 12pt          
+geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
+papersize: letter
+colorlinks: true
+fig-height: 4
+fig-width: 7.5
+header-includes:
+  - \usepackage{threeparttable}
 ---
-
-# The socialization of meritocracy and market justice preferences at school
 
@@
 
-{{< include 05-conclusiones.qmd >}}
+{{< include 05-discussion.qmd >}}
 
-{{< include 06-anexos.qmd >}}
+{{< include 06-conclusion.qmd >}}

### 2024-07-04 processing/02-analisis-actualizado.Rmd
Update edit submit

---
title: "Análisis actualizados"
author: "Equipo EDUMER"
date: "`r Sys.setlocale('LC_TIME', 'es_ES.UTF-8'); format(Sys.Date(), '%A %d, %B %Y')`"
output: 
  html_document: 
    toc: yes
    code_folding: hide
    toc_float: 
      collapsed: true
      smooth_scroll: false
      number_sections: true
---

```{r packages, echo=FALSE}
pacman::p_load(tidyverse,
               ggpubr, 
               lme4,
               ordinal,
               sjPlot, 
               texreg,
               kableExtra,
               easystats,
               ggeffects,
               webshot)

options(scipen=999)
rm(list = ls())
```

```{r data, echo=FALSE}

load("../input/data/data_index.RData")
load("../input/data/data_rec.RData")
webshot::install_phantomjs(force = T)
```

## Cumulative link mixed models

```{r echo=FALSE, include=FALSE, eval=FALSE}
sjmisc::frq(data_rec$educacion_rec)
data_rec <- data_rec %>% mutate(universitario = case_when(educacion_rec=="Ed Basica"~"No",
                                                          educacion_rec=="Ed Media"~"No",
                                                          educacion_rec=="Ed Tecnica"~"No",
                                                          educacion_rec=="Universidad o posgrado"~"Si",
                                                          educacion_rec=="Ns/Nr"~"Ns/Nr"))
data_rec$universitario <- factor(data_rec$universitario, levels = c("No", "Si", "Ns/Nr"))
sjmisc::frq(data_rec$universitario)

data_rec <- data_rec %>% mutate(school_nse = case_when(cod_grupo=="Bajo"~"Bajo",
                                                          cod_grupo=="Medio bajo"~"Medio",
                                                          cod_grupo=="Medio"~"Medio",
                                                          cod_grupo=="Medio alto"~"Medio",
                                                          cod_grupo=="Alto"~"Alto"))
sjmisc::frq(data_rec$school_nse)
data_rec$school_nse <- factor(data_rec$school_nse, levels = c("Bajo", "Medio", "Alto"))

data_pension <- data_rec %>% 
    dplyr::select(-just_educ, -just_salud, -redistribucion)
data_pension <- na.omit(data_pension)
data_pension$just_pension_fact <- as_factor(data_pension$just_pension)

m0_p <- clmm(just_pension_fact ~ 1 + (1 | mrbd), data=data_pension)

performance::icc(m0_p, by_group = T)

m1_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_pension)

m2_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_pension)

m3_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
    school_nse + simce + (1 | mrbd), data=data_pension)

modelos_pension <- list(m1_p, m2_p, m3_p)

# modelos_odd_pension <- map(tbls, odds_fun)

data_salud <- data_rec %>% 
    dplyr::select(-just_pension, -just_educ, -redistribucion)
data_salud <- na.omit(data_salud)
data_salud$just_salud_fact <- as_factor(data_salud$just_salud)

m0_h <- clmm(just_salud_fact ~ 1 + (1 | mrbd), data=data_salud)

performance::icc(m0_h, by_group = T)

m1_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_salud)

m2_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_salud)

m3_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
    school_nse + simce + (1 | mrbd), data=data_salud)

modelos_health <- list(m1_h, m2_h, m3_h)

# modelos_odd_health <- map(tbls, odds_fun)

data_educ <- data_rec %>% 
    dplyr::select(-just_pension, -just_salud, -redistribucion)
data_educ <- na.omit(data_educ)
data_educ$just_educ_fact <- as_factor(data_educ$just_educ)

m0_e <- clmm(just_educ_fact ~ 1 + (1 | mrbd), data=data_educ)

performance::icc(m0_e, by_group = T)

m1_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_educ)

m2_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_educ)

m3_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
    school_nse + simce + (1 | mrbd), data=data_educ)

modelos_educ <- list(m1_e, m2_e, m3_e)

#modelos_odd_educ <- map(tbls, odds_fun)

```

```{r echo=FALSE, results='asis', eval=FALSE}
#| tbl-cap: "Cumulative link multilevel models of differential acces justificacion of health, pensions, and education"
#| label: tbl-ordinal-reg

ccoef <- list(
  inteligencia_esc = "School talent",
  esfuerzo_esc = "School effort",
  inteligencia_soc = "Social talent",
  esfuerzo_soc = "Social effort",
  merito_soc = "Deservingness",
  "universitarioSi" = "University or posgraduate",
  "universitarioNs/Nr" = "Missing",
  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
  "school_nseMedio" = "SES Medium",
  school_nseAlto = "SES High",
  simceMedio = "Simce Medium",
  simceAlto = "Simce High"
)

texreg::htmlreg(l = list(modelos_health[[2]],
                         modelos_health[[3]],
                         modelos_pension[[2]],
                         modelos_pension[[3]],
                         modelos_educ[[2]],
                         modelos_educ[[3]]),   
                caption = NULL,
                stars = c(0.05, 0.01, 0.001),
                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars. Source: own elaboration based on First Study of Civic Education 2017.",
                custom.header = list("Health" = 1:2, "Pensions" = 3:4, "Education" = 5:6),
                custom.model.names = rep(c("Model 1", "Model 2"), 3),
                custom.coef.map = ccoef,
                leading.zero = T,
                scalebox = 0.60,
                groups = list("Parental education (Ref.= Non-university)" = 6:7,
                                "Socioeconomic level (Ref.= Low)" = 9:10,
                                "Achievement score (Ref.= Low)" = 11:12),
                custom.gof.rows = list("Controls"=rep("Yes",6)),
                include.loglik = FALSE, include.aic = FALSE,
                file = "../output/tables/tab-ordinalreg.html")

webshot::webshot(url = "../output/tables/tab-ordinalreg.html", file = "../output/tables/tab-ordinalreg.png")
```

## Mixed effects models

```{r echo=FALSE, include=FALSE}
data_bienestar <- data_bienestar %>% mutate(universitario = case_when(educacion_rec=="Ed Basica"~"No",
                                                          educacion_rec=="Ed Media"~"No",
                                                          educacion_rec=="Ed Tecnica"~"No",
                                                          educacion_rec=="Universidad o posgrado"~"Si",
                                                          educacion_rec=="Ns/Nr"~"Ns/Nr"))
data_bienestar$universitario <- factor(data_bienestar$universitario, levels = c("No", "Si", "Ns/Nr"))

sjmisc::frq(data_bienestar$cod_grupo)
data_bienestar <- data_bienestar %>% mutate(school_nse = case_when(cod_grupo=="Bajo"~"Bajo",
                                                          cod_grupo=="Medio bajo"~"Medio",
                                                          cod_grupo=="Medio"~"Medio",
                                                          cod_grupo=="Medio alto"~"Medio",
                                                          cod_grupo=="Alto"~"Alto"))
sjmisc::frq(data_bienestar$school_nse)
data_bienestar$school_nse <- factor(data_bienestar$school_nse, levels = c("Bajo", "Medio", "Alto"))

reg0_bienestar <- lmer(bienestar ~ 1 + (1 | mrbd), data = data_bienestar)

performance::icc(reg0_bienestar, by_group = T)

reg1_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
                         inteligencia_soc + esfuerzo_soc + merito_soc + 
                         acc_tec + internet_rec +
                         (1 | mrbd), data=data_bienestar)

reg2_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
                         inteligencia_soc + esfuerzo_soc + merito_soc + 
                         universitario + libros_rec + acc_tec + internet_rec +
                         (1 | mrbd), data=data_bienestar)

reg3_bienestar <- lmer(bienestar ~ 1 + mean_educ + cod_depe2 + school_nse +
                         simce + (1 | mrbd), data=data_bienestar)

reg4_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + 
                         inteligencia_soc + esfuerzo_esc + inteligencia_esc + 
                         universitario + libros_rec + acc_tec + internet_rec + 
                         mean_educ + cod_depe2 + school_nse + 
                         (1 | mrbd), data=data_bienestar)

reg5_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + 
                         inteligencia_soc + esfuerzo_esc + inteligencia_esc + 
                         universitario + libros_rec + acc_tec + internet_rec + 
                         mean_educ + cod_depe2 + school_nse + simce + 
                         (1 | mrbd), data=data_bienestar)

test01 <- anova(reg0_bienestar, reg1_bienestar, test = "Chisq")
test02 <- anova(reg0_bienestar, reg2_bienestar, test = "Chisq")
test03 <- anova(reg0_bienestar, reg3_bienestar, test = "Chisq")
test04 <- anova(reg0_bienestar, reg4_bienestar, test = "Chisq")
test05 <- anova(reg0_bienestar, reg5_bienestar, test = "Chisq")

test.pvalues1 <- test01$`Pr(>Chisq)`[2]
test.pvalues2 <- test02$`Pr(>Chisq)`[2]
test.pvalues3 <- test03$`Pr(>Chisq)`[2]
test.pvalues4 <- test04$`Pr(>Chisq)`[2]
test.pvalues5 <- test05$`Pr(>Chisq)`[2]

deviance1 <- test01$deviance[2]
deviance2 <- test02$deviance[2]
deviance3 <- test03$deviance[2]
deviance4 <- test04$deviance[2]
deviance5 <- test05$deviance[2]
```

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Linear mixed-effects models for meritocracy perception and market justice preferences"
#| label: tbl-lineal-reg

ccoef <- list(
  "(Intercept)" = "Constant",
  inteligencia_esc = "School talent",
  esfuerzo_esc = "School effort",
  inteligencia_soc = "Social talent",
  esfuerzo_soc = "Social effort",
  merito_soc = "Deservingness",
  "universitarioSi" = "University or posgraduate",
  "universitarioNs/Nr" = "Missing",
  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
  "school_nseMedio" = "SES Medium",
  school_nseAlto = "SES High",
  simceMedio = "Simce Medium",
  simceAlto = "Simce High"
)

texreg::htmlreg(list(reg1_bienestar, reg2_bienestar, reg4_bienestar, reg5_bienestar),
                custom.model.names = c("Model 1",
                                       "Model 2",
                                       "Model 3",
                                       "Model 4"),
                caption = NULL,
                stars = c(0.05, 0.01, 0.001),
                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars. Source: own elaboration based on First Study of Civic Education 2017.",
                custom.coef.map = ccoef,
                leading.zero = T,
                scalebox = 0.60,
                groups = list("Parental education (Ref.= Non-university)" = 7:8,
                              "Socioeconomic level (Ref.= Low)" = 10:11,
                              "Achievement score (Ref.= Low)" = 12:13),
                custom.gof.rows = list("Controls"=rep("Yes",4),
                                       "Deviance" = c(deviance1,
                                                    deviance2,
                                                    deviance4,
                                                    deviance5),
                                       "Deviance Test (p)" = c(test.pvalues1,
                                                             test.pvalues2,
                                                             test.pvalues4,
                                                             test.pvalues5)),
                include.loglik = FALSE,include.aic = FALSE,
                file = "../output/tables/tab-linealreg.html")

webshot::webshot(url = "../output/tables/tab-linealreg.html", file = "../output/tables/tab-linealreg.png")

```

```{r echo=FALSE, include=FALSE}
load("../input/data/data_rec.RData")
data_rec <- data_rec %>% rowwise() %>%  mutate(educacion_rec2 = case_when(educacion_rec=="Ed Basica"~"Low level",
                                                 educacion_rec=="Ed Media"~"Low level",
                                                 educacion_rec=="Ed Tecnica"~"Low level",
                                                 educacion_rec=="Universidad o posgrado"~"University or Postgraduate",
                                                 educacion_rec=="Ns/Nr"~"Missing"
                                                 ))
data_rec$educacion_rec2 <- factor(data_rec$educacion_rec2, levels = c("Low level", "University or Postgraduate", "Missing"))

data_rec <- data_rec %>% rowwise() %>%  mutate(cod_grupo2  = case_when(cod_grupo=="Bajo"~"Bajo",
                                                                       cod_grupo=="Medio bajo"~"Bajo",
                                                                       cod_grupo=="Medio"~"Bajo",
                                                                       cod_grupo=="Medio alto"~"Alto",
                                                                       cod_grupo=="Alto"~"Alto"))
data_rec$cod_grupo2 <- factor(data_rec$cod_grupo2, levels = c("Bajo", "Alto"))
data_rec$simce <- factor(data_rec$simce, labels = c("Low", "Medium", "High"))

data_bienestar <- na.omit(data_rec)
#data_bienestar <- data_bienestar %>% mutate(pension_rec = 5-just_pension,
#                                            educ_rec = 5-just_educ,
#                                            salud_rec = 5-just_salud)
data_bienestar <- data_bienestar %>% rowwise() %>% dplyr::mutate(bienestar = mean(c(just_educ, just_salud, just_pension), na.omit=TRUE))
data_bienestar$bienestar <- round(data_bienestar$bienestar, 2)

reg26.2_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + merito_soc*simce + internet_rec + (1 + simce | mrbd), data=data_bienestar)

```

```{r echo=FALSE}
save(m0_h, m1_h, m2_h, m3_h, 
     m0_p, m1_p, m2_p, m3_p,
     m0_e, m1_e, m2_e, m3_e,
     reg0_bienestar, reg1_bienestar, reg2_bienestar, reg4_bienestar, reg5_bienestar,
     deviance1, deviance2, deviance3, deviance4, deviance5,
     test.pvalues1, test.pvalues2, test.pvalues3, test.pvalues4, test.pvalues5,
     reg26.2_bienestar, file = "../input/data/mlm_models.RData"
     )
```

### 2024-07-08 04-analisis.qmd
Ultimos cambios: paper, front matter, cover letter, back matter

@@
 bibliography: "input/bib/just-ineq-merit.bib"
-csl: "input/bib/apa6.csl"
+csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
 ---
@@
   scalebox = 0.8,
-  groups = list("Parental education (Ref.= Non-university)" = 6:7, "Socioeconomic level (Ref.= Low)" = 9:10, "Achievement score (Ref.= Low)" = 11:12),
-  custom.gof.rows = list("Controls" = rep("Yes", 6)),
   include.loglik = FALSE,
   include.aic = FALSE,
+  groups = list("Parental education (Ref.= Non-university)" = 6:7, "Socioeconomic level (Ref.= Low)" = 9:10, "Achievement score (Ref.= Low)" = 11:12),
+  custom.gof.names = c("BIC", 
+                       "Num.obs",
+                       "Num. groups: Schools",
+                       "Var: Schools (Intercept)"),
+  custom.gof.rows = list("Controls"=rep("Yes", 6))
 )
@@
 
-texreg::texreg(list(reg1_bienestar, reg2_bienestar, reg4_bienestar, reg5_bienestar),
+texreg::texreg(list(reg1_bienestar, reg2_bienestar, 
+                       reg4_bienestar, reg5_bienestar),
                custom.model.names = c("Model 1",
@@
                scalebox = 0.85,
+               include.loglik = FALSE,
+               include.aic = FALSE,
                groups = list("Parental education (Ref.= Non-university)" = 7:8,
@@
                              "Achievement score (Ref.= Low)" = 12:13),
+               custom.gof.names = c("BIC", 
+                                    "Num.obs",
+                                    "Num. groups: Schools",
+                                    "Var: Schools (Intercept)",
+                                    "Var: Residual"),
                custom.gof.rows = list("Controls"=rep("Yes",4),
@@
                                                               test.pvalues4,
-                                                             test.pvalues5)),
-                include.loglik = FALSE,include.aic = FALSE)
+                                                              test.pvalues5))
+               )
 

### 2024-07-08 07-supplementary-material.qmd
Ultimos cambios: paper, front matter, cover letter, back matter

--- 2024-06-18 07-anexos.qmd
@@
 author: "Equipo EDUMER"
-format: pdf
-editor: visual
+bibliography: "input/bib/just-ineq-merit.bib"
+csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
 ---
 
-## Appendix {#appendix}
+# Supplementary Material {#appendix}
 
-```{r}
-#| tbl-cap: "Translation of inequality justification items"
+```{r load, include=FALSE}
+pacman::p_load(tidyverse,
+               ggpubr, 
+               lme4,
+               ordinal,
+               sjPlot, 
+               texreg,
+               kableExtra,
+               easystats,
+               ggeffects,
+               knitr,
+               webshot2)
+
+options(scipen=999)
+rm(list = ls())
+
+load("input/data/data_index.RData")
+load("input/data/data_rec.RData")
+load("input/data/mlm_models.RData")
 ```
 
-|                                                                                                                    |                                                                                                                                              |
-|--------------------------------|---------------------------------------|
-| **English translation**                                                                                            | **Original Spanish**                                                                                                                         |
-| It is just that in Chile people who cna pay have a better education for their children                             | Es justo que en Chile las personas que puedan pagar tengan una mejor educación para sus hijos.
-|
-| It is just that in Chile people with higher incomes can have better pensions than people with low incomes          | Es justo que en Chile las personas con mayores ingresos puedan tener mejores pensiones que las personas de ingresos más bajos                |
-| It is just that in Chile people with higher incomes can access better health services than people with low incomes | Es justo que en Chile las personas con mayores ingresos puedan acceder a una mejor atención de salud que las personas con ingresos más bajos |
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = F,
+                      warning = F,
+                      error = F, 
+                      message = F) 
+```
 
-```{r}
-#| fig-cap: "Access to better health services"
+## Translated items for market justice preferences
+
+```{r echo=FALSE}
+#| tbl-cap: "Translation of items for health, pensions, and education"
+#| label: tbl-translation
+
+data <- data.frame(
+  English = c("It is just that in Chile people with higher incomes can access better health services than people with low incomes",
+              "It is just that in Chile people with higher incomes can have better pensions than people with low incomes",
+              "It is just that in Chile people who can pay have a better education for their children"),
+  Spanish = c("Es justo que en Chile las personas con mayores ingresos puedan acceder a una mejor atención de salud que las personas con ingresos más bajos",
+              "Es justo que en Chile las personas con mayores ingresos puedan tener mejores pensiones que las personas de ingresos más bajos",
+              "Es justo que en Chile las personas que puedan pagar tengan una mejor educación para sus hijos")
+  )
+
+kableExtra::kable(data, 
+      col.names = c("English Translation", "Original Spanish"), 
+      booktabs = T, escape = F, align = 'l') %>%
+  kableExtra::kable_styling(latex_options = "HOLD_position", bootstrap_options = "striped", full_width = T)  %>%
+  row_spec(0, bold = T)
+
+```
+
+## Complete multilevel models
+
+```{r echo=FALSE}
+
+data_rec <- data_rec %>% mutate(universitario = case_when(educacion_rec=="Ed Basica"~"No",
+                                                          educacion_rec=="Ed Media"~"No",
+                                                          educacion_rec=="Ed Tecnica"~"No",
+                                                          educacion_rec=="Universidad o posgrado"~"Si",
+                                                          educacion_rec=="Ns/Nr"~"Ns/Nr"))
+data_rec$universitario <- factor(data_rec$universitario, levels = c("No", "Si", "Ns/Nr"))
+
+data_rec <- data_rec %>% mutate(school_nse = case_when(cod_grupo=="Bajo"~"Bajo",
+                                                          cod_grupo=="Medio bajo"~"Medio",
+                                                          cod_grupo=="Medio"~"Medio",
+                                                          cod_grupo=="Medio alto"~"Medio",
+                                                          cod_grupo=="Alto"~"Alto"))
+data_rec$school_nse <- factor(data_rec$school_nse, levels = c("Bajo", "Medio", "Alto"))
+
+data_salud <- data_rec %>% 
+    dplyr::select(-just_pension, -just_educ, -redistribucion)
+data_salud <- na.omit(data_salud)
+data_salud$just_salud_fact <- as_factor(data_salud$just_salud)
+
+m0_h <- clmm(just_salud_fact ~ 1 + (1 | mrbd), data=data_salud)
+
+m1_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_salud)
+
+m2_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
+    libros_rec + (1 | mrbd), data=data_salud)
+
+m3_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
+    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_salud)
+
+m4_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
+    libros_rec + acc_tec + internet_rec + school_nse + simce + (1 | mrbd), data=data_salud)
+
+m5_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
+    libros_rec + acc_tec + internet_rec + school_nse + simce +
+    mean_educ + cod_depe2 +(1 | mrbd), data=data_salud)
+
+modelos_health <- list(m0_h, m1_h, m2_h, m3_h, m4_h, m5_h)
+
+```
+
+```{r echo=FALSE, results='asis'}
+#| tbl-cap: "Cumulative Link Multilevel Models for Justifying Differential Access to Health Care"
+#| label: tbl-ordinal-health
+
+ccoef <- list(
+  "1|2" = "Strongly Disagree|Disagree",
+  "2|3" = "Disagree|Agree",
+  "3|4" = "Agree|Strongly Agree",
+  inteligencia_esc = "School talent",
+  esfuerzo_esc = "School effort",
+  inteligencia_soc = "Social talent",
+  esfuerzo_soc = "Social effort",
+  merito_soc = "Deservingness",
+  universitarioSi = "University or posgraduate",
+  "universitarioNs/Nr" = "Missing",
+  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
+  acc_tec = "Technology acces",
+  internet_rec2 = "No internet connectivity (Ref.= Internet connectivity)",
+  school_nseMedio = "SES Medium",
+  school_nseAlto = "SES High",
+  simceMedio = "Simce Medium",
+  simceAlto = "Simce High",
+  mean_educ = "Prop. university level at school",
+  "cod_depe2Part. subvencionado" = "Subsidized Private",
+  "cod_depe2Part. privado" = "Private")
+
+texreg::texreg(
+  l = modelos_health,
+  caption = NULL,
+  stars = c(0.05, 0.01, 0.001),
+  custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. %stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
+  threeparttable = T,
+  custom.model.names = paste0(rep("Model ",6), seq(0,5)),
+  custom.coef.map = ccoef,
+  leading.zero = TRUE,
+  use.packages = FALSE,
+  booktabs = TRUE,
+  scalebox = 0.8,
+  groups = list("Parental education (Ref.= Non-university)" = 9:10, "Socioeconomic level (Ref.= Low)" = 14:15, "Achievement score (Ref.= Low)" = 16:17, "School dependence (Ref.= Public)" = 19:20),
+  custom.gof.names = c("Log Likelihood",
+                       "AIC", 
+                       "BIC", 
+                       "Num.obs",
+                       "Num. groups: Schools",
+                       "Var: Schools (Intercept)"))
+
+```
+
+```{r echo=FALSE}
+data_pension <- data_rec %>% 
+    dplyr::select(-just_educ, -just_salud, -redistribucion)
+data_pension <- na.omit(data_pension)
+data_pension$just_pension_fact <- as_factor(data_pension$just_pension)
+
+m0_p <- clmm(just_pension_fact ~ 1 + (1 | mrbd), data=data_pension)
+
+m1_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc +
+             esfuerzo_esc + inteligencia_soc + esfuerzo_soc +
+             merito_soc + (1 | mrbd), data=data_pension)
+
+m2_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc +
+             esfuerzo_esc + inteligencia_soc + esfuerzo_soc +
+             merito_soc + universitario + libros_rec +
+             (1 | mrbd), data=data_pension)
+
+m3_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc +
+             esfuerzo_esc + inteligencia_soc + esfuerzo_soc +
+             merito_soc + universitario + libros_rec + acc_tec +
+             internet_rec + (1 | mrbd), data=data_pension)
+
+m4_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc +
+             esfuerzo_esc + inteligencia_soc + esfuerzo_soc +
+             merito_soc + universitario + libros_rec + acc_tec + 
+             internet_rec + school_nse + simce + (1 | mrbd),
+             data=data_pension)
+
+m5_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc +
+             esfuerzo_esc + inteligencia_soc + esfuerzo_soc +
+             merito_soc + universitario + libros_rec + acc_tec + 
+             internet_rec + school_nse + simce + mean_educ + 
+             cod_depe2 + (1 | mrbd), data=data_pension)
+
+modelos_pension <- list(m0_p, m1_p, m2_p, m3_p, m4_p, m5_p)
+
+```
+
+```{r echo=FALSE, results='asis'}
+#| tbl-cap: "Cumulative Link Multilevel Models for Justifying Differential Access to Pensions"
+#| label: tbl-ordinal-pension
+
+texreg::texreg(
+  l = modelos_pension,
+  caption = NULL,
+  stars = c(0.05, 0.01, 0.001),
+  custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. %stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
+  threeparttable = T,
+  custom.model.names = paste0(rep("Model ",6), seq(0,5)),
+  custom.coef.map = ccoef,
+  leading.zero = TRUE,
+  use.packages = FALSE,
+  booktabs = TRUE,
+  scalebox = 0.7,
+  groups = list("Parental education (Ref.= Non-university)" = 9:10, "Socioeconomic level (Ref.= Low)" = 14:15, "Achievement score (Ref.= Low)" = 16:17, "School dependence (Ref.= Public)" = 19:20),
+  custom.gof.names = c("Log Likelihood",
+                       "AIC", 
+                       "BIC", 
+                       "Num.obs",
+                       "Num. groups: Schools",
+                       "Var: Schools (Intercept)"))
+
+```
+
+```{r echo=FALSE}
+data_educ <- data_rec %>% 
+    dplyr::select(-just_pension, -just_salud, -redistribucion)
+data_educ <- na.omit(data_educ)
+data_educ$just_educ_fact <- as_factor(data_educ$just_educ)
+
+m0_e <- clmm(just_educ_fact ~ 1 + (1 | mrbd), data=data_educ)
+
+m1_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_educ)
+
+m2_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
+    libros_rec + (1 | mrbd), data=data_educ)
+
+m3_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
+    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_educ)
+
+m4_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
+    libros_rec + acc_tec + internet_rec + school_nse + simce + (1 | mrbd), data=data_educ)
+
+m5_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
+    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
+    libros_rec + acc_tec + internet_rec + school_nse + simce +
+    mean_educ + cod_depe2 + (1 | mrbd), data=data_educ)
+
+modelos_educ <- list(m0_e, m1_e, m2_e, m3_e, m4_e, m5_e)
+
+```
+
+```{r echo=FALSE, results='asis'}
+#| tbl-cap: "Cumulative Link Multilevel Models for Justifying Differential Access to Education"
+#| label: tbl-ordinal-educ
+
+texreg::texreg(
+  l = modelos_educ,
+  caption = NULL,
+  stars = c(0.05, 0.01, 0.001),
+  custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. %stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
+  threeparttable = T,
+  custom.model.names = paste0(rep("Model ",6), seq(0,5)),
+  custom.coef.map = ccoef,
+  leading.zero = TRUE,
+  use.packages = FALSE,
+  booktabs = TRUE,
+  scalebox = 0.7,
+  groups = list("Parental education (Ref.= Non-university)" = 9:10, "Socioeconomic level (Ref.= Low)" = 14:15, "Achievement score (Ref.= Low)" = 16:17, "School dependence (Ref.= Public)" = 19:20),
+  custom.gof.names = c("Log Likelihood",
+                       "AIC", 
+                       "BIC", 
+                       "Num.obs",
+                       "Num. groups: Schools",
+                       "Var: Schools (Intercept)"))
+
+```
+
+```{r echo=FALSE, include=FALSE}
+
+data_bienestar <- data_bienestar %>% mutate(universitario = case_when(educacion_rec=="Ed Basica"~"No",
+                                                          educacion_rec=="Ed Media"~"No",
+                                                          educacion_rec=="Ed Tecnica"~"No",
+                                                          educacion_rec=="Universidad o posgrado"~"Si",
+                                                          educacion_rec=="Ns/Nr"~"Ns/Nr"))
+data_bienestar$universitario <- factor(data_bienestar$universitario, levels = c("No", "Si", "Ns/Nr"))
+
+data_bienestar <- data_bienestar %>% mutate(school_nse = case_when(cod_grupo=="Bajo"~"Bajo",
+                                                          cod_grupo=="Medio bajo"~"Medio",
+                                                          cod_grupo=="Medio"~"Medio",
+                                                          cod_grupo=="Medio alto"~"Medio",
+                                                          cod_grupo=="Alto"~"Alto"))
+data_bienestar$school_nse <- factor(data_bienestar$school_nse, levels = c("Bajo", "Medio", "Alto"))
+
+reg0_bienestar <- lmer(bienestar ~ 1 + (1 | mrbd), data = data_bienestar)
+
+reg1_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
+                         inteligencia_soc + esfuerzo_soc + merito_soc + 
+                         (1 | mrbd), data=data_bienestar)
+
+reg2_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
+                         inteligencia_soc + esfuerzo_soc + merito_soc + 
+                         universitario + libros_rec + (1 | mrbd), data=data_bienestar)
+
+reg3_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
+                         inteligencia_soc + esfuerzo_soc + merito_soc + 
+                         universitario + libros_rec + acc_tec + 
+                         internet_rec +  (1 | mrbd), data=data_bienestar)
+
+reg4_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
+                         inteligencia_soc + esfuerzo_soc + merito_soc + 
+                         universitario + libros_rec + acc_tec + 
+                         internet_rec +  school_nse + simce + 
+                         (1 | mrbd), data=data_bienestar)
+
+reg5_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
+                         inteligencia_soc + esfuerzo_soc + merito_soc + 
+                         universitario + libros_rec + acc_tec + 
+                         internet_rec +  school_nse + simce + 
+                         mean_educ + cod_depe2 + 
+                         (1 | mrbd), data=data_bienestar)
+```
+
+```{r echo=FALSE, results='asis'}
+#| tbl-cap: "Linear mixed-effects models for meritocracy perception and market justice preferences"
+#| label: tbl-lineal-reg
+ccoef <- list(
+ "(Intercept)" = "Constant",
+  inteligencia_esc = "School talent",
+  esfuerzo_esc = "School effort",
+  inteligencia_soc = "Social talent",
+  esfuerzo_soc = "Social effort",
+  merito_soc = "Deservingness",
+  universitarioSi = "University or posgraduate",
+  "universitarioNs/Nr" = "Missing",
+  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
+  acc_tec = "Technology acces",
+  internet_rec2 = "No internet connectivity (Ref.= Internet connectivity)",
+  school_nseMedio = "SES Medium",
+  school_nseAlto = "SES High",
+  simceMedio = "Simce Medium",
+  simceAlto = "Simce High",
+  mean_educ = "Prop. university level at school",
+  "cod_depe2Part. subvencionado" = "Subsidized Private",
+  "cod_depe2Part. privado" = "Private")
+
+texreg::texreg(
+  l = list(reg0_bienestar, reg1_bienestar, reg2_bienestar, reg3_bienestar, reg4_bienestar, reg5_bienestar),
+  caption = NULL,
+  stars = c(0.05, 0.01, 0.001),
+  custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. %stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
+  threeparttable = T,
+  custom.model.names = paste0(rep("Model ",6), seq(0,5)),
+  custom.coef.map = ccoef,
+  leading.zero = TRUE,
+  use.packages = FALSE,
+  booktabs = TRUE,
+  scalebox = 0.7,
+  groups = list("Parental education (Ref.= Non-university)" = 7:8, "Socioeconomic level (Ref.= Low)" = 12:13, "Achievement score (Ref.= Low)" = 14:15, "School dependence (Ref.= Public)" = 16:18),
+   custom.gof.names = c("AIC", 
+                               "BIC", 
+                               "Log Likelihood", 
+                               "Num.obs",
+                               "Num. groups: Schools",
+                               "Var: Schools (Intercept)",
+                               "Var: Residual"))
+
 ```
@@
 ```{r}
-#| fig-cap: "Access to better education"
+
+load("input/data/data_rec.RData")
+
+data_rec <- data_rec %>% rowwise() %>%  mutate(educacion_rec2 = case_when(educacion_rec=="Ed Basica"~"Low level",
+                                                 educacion_rec=="Ed Media"~"Low level",
+                                                 educacion_rec=="Ed Tecnica"~"Low level",
+                                                 educacion_rec=="Universidad o posgrado"~"University or Postgraduate",
+                                                 educacion_rec=="Ns/Nr"~"Missing"
+                                                 ))
+data_rec$educacion_rec2 <- factor(data_rec$educacion_rec2, levels = c("Low level", "University or Postgraduate", "Missing"))
+
+data_rec <- data_rec %>% rowwise() %>%  mutate(cod_grupo2  = case_when(cod_grupo=="Bajo"~"Bajo",
+                                                                       cod_grupo=="Medio bajo"~"Bajo",
+                                                                       cod_grupo=="Medio"~"Bajo",
+                                                                       cod_grupo=="Medio alto"~"Alto",
+                                                                       cod_grupo=="Alto"~"Alto"))
+data_rec$cod_grupo2 <- factor(data_rec$cod_grupo2, levels = c("Bajo", "Alto"))
+data_rec$simce <- factor(data_rec$simce, labels = c("Low", "Medium", "High"))
+
+data_bienestar <- na.omit(data_rec)
+#data_bienestar <- data_bienestar %>% mutate(pension_rec = 5-just_pension,
+#                                            educ_rec = 5-just_educ,
+#                                            salud_rec = 5-just_salud)
+data_bienestar <- data_bienestar %>% rowwise() %>% dplyr::mutate(bienestar = mean(c(just_educ, just_salud, just_pension), na.omit=TRUE))
+data_bienestar$bienestar <- round(data_bienestar$bienestar, 2)
+
 ```
@@
 ```{r}
-#| fig-cap: "Access to better pensions"
+
+# meritocracia x ses familiar (educacion)
+reg10_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2*esfuerzo_soc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)
+
+reg11_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2*merito_soc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)
+
+reg12_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2*inteligencia_soc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)
+
+reg13_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2*esfuerzo_esc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)
+
+reg14_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2*inteligencia_esc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)
 ```
+
+```{r echo=FALSE, results='asis'}
+#| tbl-cap: "Linear mixed-effects models for meritocracy perception, parental education and market justice preferences"
+#| label: tbl-interac-educparent
+
+ccoef <- list(
+  "(Intercept)" = "Constant",
+  inteligencia_esc = "School talent",
+  esfuerzo_esc = "School effort",
+  inteligencia_soc = "Social talent",
+  esfuerzo_soc = "Social effort",
+  merito_soc = "Deservingness",
+  "educacion_rec2University or Postgraduate" = "University or posgraduate",
+  "educacion_rec2Missing" = "Missing",
+  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
+  cod_grupo2Alto = "SES High (Ref.= Low)",
+  simceMedium = "Simce Medium",
+  simceHigh = "Simce High",
+  "educacion_rec2University or Postgraduate:esfuerzo_soc" = "University or posgraduate * Social effort",
+  "educacion_rec2Missing:esfuerzo_soc" = "Missing * Social effort",
+  "educacion_rec2University or Postgraduate:merito_soc" = "University or posgraduate * Deservingness",
+  "educacion_rec2Missing:merito_soc" = "Missing * Deservingness",
+  "educacion_rec2University or Postgraduate:inteligencia_soc" = "University or posgraduate * Social talent",
+  "educacion_rec2Missing:inteligencia_soc" = "Missing * Social talent",
+  "educacion_rec2University or Postgraduate:esfuerzo_esc" = "University or posgraduate * School effort",
+  "educacion_rec2Missing:esfuerzo_esc" = "Missing * School effort",
+  "educacion_rec2University or Postgraduate:inteligencia_esc" = "University or posgraduate * School talent",
+  "educacion_rec2Missing:inteligencia_esc" = "Missing * School talent"
+  )
+
+# meritocracia x ses familiar (educacion)
+
+texreg::texreg(l = list(reg10_bienestar, reg11_bienestar,
+                   reg12_bienestar, reg13_bienestar,
+                   reg14_bienestar),
+          caption = NULL,
+          stars = c(0.05, 0.01, 0.001),
+          custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses.
+Control variables are included.
+%stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
+          threeparttable = T,
+          custom.model.names = paste0(rep("Model ",5), seq(1,5)),
+          custom.coef.map = ccoef,
+          leading.zero = TRUE,
+          use.packages = FALSE,
+          booktabs = TRUE,
+          scalebox = 0.7,
+          groups = list("Parental education (Ref.= Non-university)" = 7:8,
+                        "Achievement score (Ref.= Low)" = 11:12,
+                        "Parental education x Meritocracy" = 13:22),
+          custom.gof.names = c("AIC", 
+                               "BIC", 
+                               "Log Likelihood", 
+                               "Num.obs",
+                               "Num. groups: Schools",
+                               "Var: Schools (Intercept)",
+                               "Var: Schools University or Postgraduate",
+                               "Var: Schools Missing",
+                               "Cov: Schools (Intercept) University or Postgraduate",
+                               "Cov: Schools (Intercept) Missing",
+                               "Cov: Schools University or Postgraduate, Missing",
+                               "Var: Residual"),
+          custom.gof.rows = list("Controls"=rep("Yes",5)))
+
+```
+
+```{r}
+
+# meritocracia x ses familiar (libros)
+
+reg15_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec*esfuerzo_soc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
+reg16_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec*merito_soc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
+reg17_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec*inteligencia_soc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
+reg18_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2 + libros_rec*esfuerzo_esc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
+reg19_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2 + libros_rec*inteligencia_esc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
+```
+
+```{r echo=FALSE, results='asis'}
+#| tbl-cap: "Linear mixed-effects models for meritocracy perception, books at home and market justice preferences"
+#| label: tbl-interac-books
+
+ccoef <- list(
+  "(Intercept)" = "Constant",
+  inteligencia_esc = "School talent",
+  esfuerzo_esc = "School effort",
+  inteligencia_soc = "Social talent",
+  esfuerzo_soc = "Social effort",
+  merito_soc = "Deservingness",
+  "educacion_rec2University or Postgraduate" = "University or posgraduate",
+  "educacion_rec2Missing" = "Missing",
+  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
+  cod_grupo2Alto = "SES High (Ref.= Low)",
+  simceMedium = "Simce Medium",
+  simceHigh = "Simce High",
+  "libros_recMás de 25:esfuerzo_soc" = "More than 25 books * Social effort",
+  "libros_recMás de 25:merito_soc" = "More than 25 books * Deservingness",
+  "libros_recMás de 25:inteligencia_soc" = "More than 25 books * Social talent",
+  "libros_recMás de 25:esfuerzo_esc" = "More than 25 books * School effort",
+  "libros_recMás de 25:inteligencia_esc" = "More than 25 books * School talent"
+  )
+
+# meritocracia x ses familiar (libros)
+
+texreg::texreg(l = list(reg15_bienestar, reg16_bienestar,
+                   reg17_bienestar, reg18_bienestar,
+                   reg19_bienestar),
+          caption = NULL,
+          stars = c(0.05, 0.01, 0.001),
+          custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses.
+Control variables are included.
+%stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
+          threeparttable = T,
+          custom.model.names = paste0(rep("Model ",5), seq(1,5)),
+          custom.coef.map = ccoef,
+          leading.zero = TRUE,
+          use.packages = FALSE,
+          booktabs = TRUE,
+          scalebox = 0.7,
+          groups = list("Parental education (Ref.= Non-university)" = 7:8,
+                        "Achievement score (Ref.= Low)" = 11:12,
+                        "Books at home x Meritocracy" = 13:17),
+          custom.gof.names = c("AIC", 
+                               "BIC", 
+                               "Log Likelihood", 
+                               "Num.obs",
+                               "Num. groups: Schools",
+                               "Var: Schools (Intercept)",
+                               "Var: Schools More than 25 books",
+                               "Cov: Schools (Intercept) More than 25 books",
+                               "Var: Residual"),
+          custom.gof.rows = list("Controls"=rep("Yes",5)))
+
+```
+
+```{r}
+
+# meritocracia x ses escuela
+
+reg20_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*esfuerzo_soc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
+reg21_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*merito_soc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
+reg22_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*inteligencia_soc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
+reg23_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*esfuerzo_esc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
+reg24_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*inteligencia_esc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
+```
+
+```{r echo=FALSE, results='asis'}
+#| tbl-cap: "Linear mixed-effects models for meritocracy perception, school socioeconomic status and market justice preferences"
+#| label: tbl-interac-schoolses
+
+ccoef <- list(
+  "(Intercept)" = "Constant",
+  inteligencia_esc = "School talent",
+  esfuerzo_esc = "School effort",
+  inteligencia_soc = "Social talent",
+  esfuerzo_soc = "Social effort",
+  merito_soc = "Deservingness",
+  "educacion_rec2University or Postgraduate" = "University or posgraduate",
+  "educacion_rec2Missing" = "Missing",
+  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
+  cod_grupo2Alto = "SES High (Ref.= Low)",
+  simceMedium = "Simce Medium",
+  simceHigh = "Simce High",
+  "cod_grupo2Alto:esfuerzo_soc" = "SES High * Social effort",
+  "cod_grupo2Alto:merito_soc" = "SES High * Deservingness",
+  "cod_grupo2Alto:inteligencia_soc" = "SES High * Social talent",
+  "cod_grupo2Alto:esfuerzo_esc" = "SES High * School effort",
+  "cod_grupo2Alto:inteligencia_esc" = "SES High * School talent"
+  )
+
+# meritocracia x ses escuela
+
+texreg::texreg(l = list(reg20_bienestar, reg21_bienestar,
+                   reg22_bienestar, reg23_bienestar,
+                   reg24_bienestar),
+          caption = NULL,
+          stars = c(0.05, 0.01, 0.001),
+          custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses.
+Control variables are included.
+%stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
+          threeparttable = T,
+          custom.model.names = paste0(rep("Model ",5), seq(1,5)),
+          custom.coef.map = ccoef,
+          leading.zero = TRUE,
+          use.packages = FALSE,
+          booktabs = TRUE,
+          scalebox = 0.7,
+          groups = list("Parental education (Ref.= Non-university)" = 7:8,
+                        "Achievement score (Ref.= Low)" = 11:12,
+                        "School SES x Meritocracy" = 13:17),
+          custom.gof.names = c("AIC", 
+                               "BIC", 
+                               "Log Likelihood", 
+                               "Num.obs",
+                               "Num. groups: Schools",
+                               "Var: Schools (Intercept)",
+                               "Var: Schools SES High",
+                               "Cov: Schools (Intercept) SES High",
+                               "Var: Residual"),
+          custom.gof.rows = list("Controls"=rep("Yes",5)))
+
+```
+
+```{r}
+
+#meritocracia x prop. educacion universitaria por escuela
+
+reg20.1_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*esfuerzo_soc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
+reg21.1_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*merito_soc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
+reg22.1_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*inteligencia_soc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
+reg23.1_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*esfuerzo_esc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
+reg24.1_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*inteligencia_esc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
+```
+
+```{r}
+# meritocracia x simce
+
+reg25_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*esfuerzo_soc + internet_rec + (1 + simce | mrbd), data=data_bienestar)
+
+reg26_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*merito_soc + internet_rec + (1 + simce | mrbd), data=data_bienestar)
+
+reg27_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*inteligencia_soc + internet_rec + (1 + simce | mrbd), data=data_bienestar)
+
+reg28_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*esfuerzo_esc + internet_rec + (1 + simce | mrbd), data=data_bienestar)
+
+reg29_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*inteligencia_esc + internet_rec + (1 + simce | mrbd), data=data_bienestar)
+```
+
+```{r echo=FALSE, results='asis'}
+#| tbl-cap: "Linear mixed-effects models for meritocracy perception, school perfomance status and market justice preferences"
+#| label: tbl-interac-schoolperfo
+
+ccoef <- list(
+  "(Intercept)" = "Constant",
+  inteligencia_esc = "School talent",
+  esfuerzo_esc = "School effort",
+  inteligencia_soc = "Social talent",
+  esfuerzo_soc = "Social effort",
+  merito_soc = "Deservingness",
+  "educacion_rec2University or Postgraduate" = "University or posgraduate",
+  "educacion_rec2Missing" = "Missing",
+  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
+  cod_grupo2Alto = "SES High (Ref.= Low)",
+  simceMedium = "Simce Medium",
+  simceHigh = "Simce High",
+  "simceMedium:esfuerzo_soc" = "Simce Medium * Social effort",
+  "simceHigh:esfuerzo_soc" = "Simce High * Social effort",
+  "simceMedium:merito_soc" = "Simce Medium * Deservingness",
+  "simceHigh:merito_soc" = "Simce High * Deservingness",
+  "simceMedium:inteligencia_soc" = "Simce Medium * Social talent",
+  "simceHigh:inteligencia_soc" = "Simce High * Social talent",
+  "simceMedium:esfuerzo_esc" = "Simce Medium * School effort",
+  "simceHigh:esfuerzo_esc" = "Simce High * School effort",
+  "simceMedium:inteligencia_esc" = "Simce Medium * School talent",
+  "simceHigh:inteligencia_esc" = "Simce High * School talent"
+  )
+
+# meritocracia x perfomance escuela
+
+texreg::texreg(l = list(reg25_bienestar, reg26_bienestar,
+                   reg27_bienestar, reg28_bienestar,
+                   reg29_bienestar),
+          caption = NULL,
+          stars = c(0.05, 0.01, 0.001),
+          custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses.
+Control variables are included.
+%stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
+          threeparttable = T,
+          custom.model.names = paste0(rep("Model ",5), seq(1,5)),
+          custom.coef.map = ccoef,
+          leading.zero = TRUE,
+          use.packages = FALSE,
+          booktabs = TRUE,
+          scalebox = 0.7,
+          groups = list("Parental education (Ref.= Non-university)" = 7:8,
+                        "Achievement score (Ref.= Low)" = 11:12,
+                        "School perfomance (SIMCE) x Meritocracy" = 13:22),
+          custom.gof.names = c("AIC", 
+                               "BIC", 
+                               "Log Likelihood", 
+                               "Num.obs",
+                               "Num. groups: Schools",
+                               "Var: Schools (Intercept)",
+                               "Var: Schools Simce Medium",
+                               "Var: Schools Simce High",
+                               "Cov: Schools (Intercept) Simce Medium",
+                               "Cov: Schools (Intercept) Simce High",
+                               "Cov: Schools Simce Medium, Simce High",
+                               "Var: Residual"),
+          custom.gof.rows = list("Controls"=rep("Yes",5)))
+
+```

### 2024-07-08 cover_letter.qmd
Ultimos cambios: paper, front matter, cover letter, back matter

@@
 bibliography: "input/bib/just-ineq-merit.bib"
-csl: "input/bib/apa6.csl"
+csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
 link-citations: true

### 2024-07-08 paper.qmd
Ultimos cambios: paper, front matter, cover letter, back matter

--- 2024-06-10 paper.qmd
@@
 ---
+title: "The socialization of meritocracy and market justice preferences at school"
 format:
-  html: 
-    toc: true
-    keep-tex: true
-smooth-scroll: true
-number-sections: true
+  hikmah-pdf: 
+    indent: false
+  hikmah-manuscript-pdf: 
+    indent: false
+author:
+  - name: Juan Carlos Castillo
+    email: juancastillov@uchile.cl
+    orcid: 0000-0003-1265-7854
+    affiliations: 
+        name: Department of Sociology, Universidad de Chile
+        country: Chile
+    attributes:
+      corresponding: true
+  - name: Mauricio Salgado
+    email: msalgado@unab.cl
+    orcid: 0000-0002-8108-6154
+    affiliations: 
+        name: Centro de Estudios Públicos, Chile
+        country: Chile
+  - name: Kevin Carrasco
+    email: kevin.carrasco@ug.uchile.cl
+    orcid: 0000-0002-9348-8361
+    affiliations: 
+        name: Center for Social Conflict and Cohesion Studies
+        country: Chile
+  - name: Andreas Laffert
+    email: alaffertt@estudiante.uc.cl
+    orcid: 0000-0002-9008-2454
+    affiliations: 
+        name: Institute of Sociology, Universidad Católica de Chile
+        country: Chile
+abstract: |
+  This article evaluates novel approaches to do some really important things.
+keywords: 
+  - open-source 
+  - scientific publishing
+  - reproducible research
 bibliography: "input/bib/just-ineq-merit.bib"
-csl: "input/bib/apa6.csl"
+csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
 link-citations: true
+linestretch: 1.15       
 mainfont: Times New Roman
+fontsize: 12pt          
+geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
+papersize: letter
+colorlinks: true
+fig-height: 4
+fig-width: 7.5
+header-includes:
+  - \usepackage{threeparttable}
 ---
-
-# The socialization of meritocracy and market justice preferences at school {.unnumbered}
-
-Abstract
 

### 2024-07-19 02-antecedentes.qmd
Ultimos cambios: paper, front matter, cover letter, back matter
Versión final paper.pdf

@@
 The reliance on private welfare providers and widespread user fees [@molyneux_neoliberal_2008] adds complexity to this context, as reflected in surveys conducted by the Center for Public Studies (CEP).
-According to this data, 35.9% prefer private health insurance, and 63% would prefer private education.
-Yet, research on children's justification in this area remains limited, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs [@centrodeestudiospublicos_estudio_2024].
+According to this data, 35.9% prefer private health insurance, and 63% would prefer private education [@centrodeestudiospublicos_estudio_2024].
+Yet, research on children's justification in this area remains limited, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.
 

### 2024-07-19 paper.qmd
Update con enlaces
Versión final paper.pdf

--- 2024-06-10 paper.qmd
@@
 ---
+title: "The socialization of meritocracy and market justice preferences at school"
 format:
-  html: 
-    toc: true
-    keep-tex: true
-smooth-scroll: true
-number-sections: true
+  hikmah-pdf: 
+    indent: false
+  hikmah-manuscript-pdf: 
+    indent: false
+author:
+  - name: Juan Carlos Castillo
+    email: juancastillov@uchile.cl
+    orcid: 0000-0003-1265-7854
+    affiliations: 
+        name: Department of Sociology, Universidad de Chile
+        country: Chile
+    attributes:
+      corresponding: true
+  - name: Mauricio Salgado
+    email: msalgado@unab.cl
+    orcid: 0000-0002-8108-6154
+    affiliations: 
+        name: Centro de Estudios Públicos, Chile
+        country: Chile
+  - name: Kevin Carrasco
+    email: kevin.carrasco@ug.uchile.cl
+    orcid: 0000-0002-9348-8361
+    affiliations: 
+        name: Center for Social Conflict and Cohesion Studies
+        country: Chile
+  - name: Andreas Laffert
+    email: alaffertt@estudiante.uc.cl
+    orcid: 0000-0002-9008-2454
+    affiliations: 
+        name: Institute of Sociology, Universidad Católica de Chile
+        country: Chile
+abstract: |
+  Previous research has shown that schools justify student performance differences based on meritocratic ideals.
+However, the link between school meritocracy and students’ justification of inequalities has received little attention.
+This paper examines how school and family socialization influence students’ beliefs about meritocracy and market-based justifications for access to health, pensions, and education in Chile.
+Using data from the First Study of Civic Education in Chile, we employ a multilevel modeling strategy to analyze 5,047 students and parents nested within 231 schools.
+Our findings reveal that a significant proportion of Chilean students agree with market justice principles, at a higher rate than adults.
+Most students endorse meritocratic views, particularly the reward of effort.
+Multilevel estimations show strong associations between meritocratic beliefs and market justice preferences: students who believe in the reward of effort and talent are more likely to justify inequalities based on financial capacity.
+At the school level, market justice preferences are higher in high-status schools but lower in schools with better achievement.
+Additionally, the conditional influence of meritocracy beliefs weakens in schools with higher socioeconomic status and performance.
+These results suggest that meritocratic beliefs and market justice preferences begin early and are influenced by the school environment.
+keywords: 
+  - market justice 
+  - meritocracy
+  - socialization
+  - family
+  - schools
+  - Chile
 bibliography: "input/bib/just-ineq-merit.bib"
-csl: "input/bib/apa6.csl"
+csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
 link-citations: true
+linestretch: 1.15       
 mainfont: Times New Roman
+fontsize: 12pt          
+geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
+papersize: letter
+colorlinks: true
+fig-height: 4
+fig-width: 7.5
+header-includes:
+  - \usepackage{threeparttable}
 ---
 
-# The socialization of meritocracy and market justice preferences at school {.unnumbered}
-
-Abstract
+This document was last modified at {{< modified >}}
+and it was last rendered at {{< now >}}.
 

### 2024-07-22 01-introduction.qmd
Update edit submit
Ultimos cambios: paper, front matter, cover letter, back matter

@@
 author: "Equipo EDUMER"
-format: html
 bibliography: "input/bib/just-ineq-merit.bib"
-csl: "input/bib/apa6.csl"
+csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
 ---
@@
 In the present paper, we address the role of the perception of meritocracy on the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system.
-In particular, we focus in justifying inequalities in health, education and pensions, which are traditional social policy areas where acces to better services could be justified based on payment capacity, referred as to market justice [@lane_market_1986; @lindh_bringing_2023].
+In particular, we focus on justifying inequalities in health, education and pensions, which are traditional social policy areas where access to better services could be justified based on payment capacity, referred to as market justice [@lane_market_1986; @lindh_bringing_2023].
 Most of the research in this realm has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies [@batruch_belief_2022].

### 2024-07-22 07-supplementary-material.qmd

@@
 
-# Supplementary Material {#appendix}
+# Supplementary Material
 
@@
 
-## Translated items for market justice preferences
+# Translated items for market justice preferences
 
@@
       booktabs = T, escape = F, align = 'l') %>%
-  kableExtra::kable_styling(latex_options = "HOLD_position", bootstrap_options = "striped", full_width = T)  %>%
+  kableExtra::kable_styling(latex_options = "hold_position", bootstrap_options = "striped", full_width = T)  %>%
   row_spec(0, bold = T)
@@
 
-## Complete multilevel models
+# Complete multilevel models
 
@@
   booktabs = TRUE,
-  scalebox = 0.8,
+  scalebox = 0.7,
   groups = list("Parental education (Ref.= Non-university)" = 9:10, "Socioeconomic level (Ref.= Low)" = 14:15, "Achievement score (Ref.= Low)" = 16:17, "School dependence (Ref.= Public)" = 19:20),

### 2024-07-22 back_matter.qmd

@@
 
-{{< include 07-supplementary-material.qmd >}}
-
-## Acknowledgments
+# Acknowledgments
 
@@
 
-## Authors' Contributions
+# Authors' Contributions
 
-The following statements should be used "Conceptualization, X.X.
-and Y.Y.; Methodology, X.X.; Software, X.X.; Validation, X.X., Y.Y.
-and Z.Z.; Formal Analysis, X.X.; Investigation, X.X.; Resources, X.X.; Data Curation, X.X.; Writing – Original Draft Preparation, X.X.; Writing – Review & Editing, X.X.; Visualization, X.X.; Supervision, X.X.; Project Administration, X.X.; Funding Acquisition, Y.Y.”
+Conceptualization, Juan Carlos Castillo & Mauricio Salgado; Methodology, Juan Carlos Castillo, Kevin Carrasco & Andreas Laffert; Software, Kevin Carrasco & Andreas Laffert; Formal Analysis, Kevin Carrasco & Andreas Laffert; Writing – Original Draft Preparation, Juan Carlos Castillo, Mauricio Salgado, Kevin Carrasco & Andreas Laffert; Visualization, Kevin Carrasco & Andreas Laffert; Supervision, Juan Carlos Castillo.
 
-## Conflict of Interest
+# Conflict of Interest
 
@@
 
-## Funding
+# Funding
 
@@
 
-## Data Avaibility Statement
+# Data Avaibility Statement
 
@@
 
-## References
+{{< pagebreak >}}
+
+{{< include 07-supplementary-material.qmd >}}

### 2024-07-22 02-antecedentes.qmd
Ajustes nuevos
Ajustes finales de:
- Referencias
- Cover Letter
- Funding
- Paper blinded

@@
 The influence of ideologies [@wegener_dominant_1995] and cultural schemas [@homan_being_2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations.
-While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations [@lindh_bringing_2023] and even less about children's acceptance or rejection of these market principles.
+While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; @evans_family_2010; @jasso_how_1999; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations [@lindh_bringing_2023] and even less about children's acceptance or rejection of these market principles.
 This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades [@centeno_arc_2012; @harvey_breve_2015], affecting areas such as pensions, health services, and education.
@@
 Contrary to the evidence that unequal distribution produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action [@greitemeyer_subjective_2016; @mishra_subjective_2015; @smith_relative_2012; @power_deprivationprotest_2018], Lane pointed out that in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions.
-In this sense, unequal levels of well-being would be, to some extent, a function of their talents and efforts, instead of being based on distributive principles that characterize welfare states, such as need and equality (see Wilson 2003).
+In this sense, unequal levels of well-being would be, to some extent, a function of their talents and efforts, instead of being based on distributive principles that characterize welfare states, such as need and equality (see [@wilson_role_2003]).
 

### 2024-07-22 cover_letter.qmd
Ajustes nuevos
Ajustes finales de:
- Referencias
- Cover Letter
- Funding
- Paper blinded

@@
 
-This article evaluates
+Previous research has shown that schools justify student performance differences based on meritocratic ideals.
+However, the link between school meritocracy and students’ justification of inequalities has received little attention.
+This paper examines how school and family socialization influence students’ beliefs about meritocracy and market-based justifications for access to health, pensions, and education in Chile.
+Using data from the First Study of Civic Education in Chile, we employ a multilevel modeling strategy to analyze 5,047 students and parents nested within 231 schools.
+Our findings reveal that a significant proportion of Chilean students agree with market justice principles, at a higher rate than adults.
+Most students endorse meritocratic views, particularly the reward of effort.
+Multilevel estimations show strong associations between meritocratic beliefs and market justice preferences: students who believe in the reward of effort and talent are more likely to justify inequalities based on financial capacity.
+At the school level, market justice preferences are higher in high-status schools but lower in schools with better achievement.
+Additionally, the conditional influence of meritocracy beliefs weakens in schools with higher socioeconomic status and performance.
+These results suggest that meritocratic beliefs and market justice preferences begin early and are influenced by the school environment.
 
@@
 [^*]: Corresponding author (juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) - Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile).
-The authors acknowledge the support of ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) grant ANID/FONDAP 1513009.

### 2024-07-22 paper-blinded.qmd
Ajustes nuevos
Ajustes finales de:
- Referencias
- Cover Letter
- Funding
- Paper blinded

--- 2024-07-19 paper.qmd
@@
     indent: false
-author:
-  - name: Juan Carlos Castillo
-    email: juancastillov@uchile.cl
-    orcid: 0000-0003-1265-7854
-    affiliations: 
-        name: Department of Sociology, Universidad de Chile
-        country: Chile
-    attributes:
-      corresponding: true
-  - name: Mauricio Salgado
-    email: msalgado@unab.cl
-    orcid: 0000-0002-8108-6154
-    affiliations: 
-        name: Centro de Estudios Públicos, Chile
-        country: Chile
-  - name: Kevin Carrasco
-    email: kevin.carrasco@ug.uchile.cl
-    orcid: 0000-0002-9348-8361
-    affiliations: 
-        name: Center for Social Conflict and Cohesion Studies
-        country: Chile
-  - name: Andreas Laffert
-    email: alaffertt@estudiante.uc.cl
-    orcid: 0000-0002-9008-2454
-    affiliations: 
-        name: Institute of Sociology, Universidad Católica de Chile
-        country: Chile
 abstract: |
@@
 
-This document was last modified at {{< modified >}}
-and it was last rendered at {{< now >}}.
-
 {{< include 01-introduction.qmd >}}

### 2024-07-23 05-discussion.qmd
Ultimos cambios: paper, front matter, cover letter, back matter
template mdpi
agregue un nuevo qmd para el template mdpi

@@
----
-author: "Equipo EDUMER"
-editor: visual
----
-
 # Discussion
@@
 Still, one of the limitation for the comparison between perception of meritocracy and society is the difference of the items' phrasing.
-Whereas for society the item refers to the extent to to which efforts are rewarded, at school is about the importance of effort, whereby something that is important does not necessarily mean that it is enough to be rewarded.
+Whereas for society the item refers to the extent to which efforts are rewarded, at school is about the importance of effort, whereby something that is important does not necessarily mean that it is enough to be rewarded.
 Furthermore, the school-effort item is the one with the lowest variation among all meritocratic items (more than 95% agree or strongly agree), and with the lowest effect size.

### 2024-07-23 06-conclusion.qmd
template mdpi
agregue un nuevo qmd para el template mdpi

--- 2024-06-28 06-conclusion.qmd
@@
----
-author: "Equipo EDUMER"
-bibliography: "input/bib/just-ineq-merit.bib"
-csl: "input/bib/apa6.csl"
----
-
 # Conclusion
@@
 Regarding these descriptive indicators, a a first message coming from this study is that there is something particular in school-age when it comes to perceptions and preferences about inequality.
-This can relate to development aspects (Kim & Gewirtz, 2019, Rizzo & Killen, 2020) but it could be something specific to the Chilean case, which requires further comparative as well as longitudinal research in the area.
+This can relate to development aspects [@kim_socioeconomic_2020, @rizzo_children_2020] but it could be something specific to the Chilean case, which requires further comparative as well as longitudinal research in the area.
 In any case, teaching about the origins and characteristics of social inequality is not part of the chilean school curriculum, calling for a more dedicated consideration of these topics in areas such as citizensip education as well as in history and social science education.
@@
 We found contrary evidence for the association between status, school performance, and market justice preferences: high status and high achievement at the school level are associated with less market justice preferences.
-This is opposed to the rational-interest argument that suggests that those doing better - both economically and academically- would demand less redistribution (Gonthier,2016; Runst, 2018; Bullock, 2021) and therefore are expected to challenge market mechanisms.
+This is opposed to the rational-interest argument that suggests that those doing better - both economically and academically- would demand less redistribution [@gonthier_parallel_2017, @runst_does_2018, @bullock_education_2021] and therefore are expected to challenge market mechanisms.
 A possible explanation is that education contributes to forming a critical view of markets in redistribution, nevertheless, further research in this area is needed to assert the specific mechanisms that might be playing a role.

### 2024-07-24 MDPI-submit/template-mdpi.tex
Archivos submit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
    behavsci,
    article,
    submit,
moreauthors
]{mdpi}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% TODO: Add custom LaTeX header directives here
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{plainnat}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The socialization of meritocracy and market justice preferences at school},
  pdfauthor={Juan Carlos Castillo; Mauricio Salgado; Kevin Carrasco; Andreas Laffert},
  pdfkeywords={market
justice, meritocracy, socialization, family, schools, Chile},
  pdfcreator={LaTeX via pandoc},
  unicode=true,
	bookmarksopen={true},
	pdffitwindow=true, 
	colorlinks=true, 
	linkcolor=bluecite, 
	citecolor=bluecite, 
	urlcolor=bluecite, 
	hyperfootnotes=true, 
	pdfstartview={FitH},
	pdfpagemode= UseNone}

\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{\the\year{}}
\copyrightyear{\the\year{}}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{ } 
\daterevised{ } % Comment out if no revised date
\dateaccepted{ } 
\datepublished{ } 
%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%\pdfoutput=1 % Uncommented for upload to arXiv.org
%\CorrStatement{yes}  % For updates

\Title{ The socialization of meritocracy and market justice preferences
at school }

% MDPI internal command: Title for citation in the left column
\TitleCitation{ The socialization of meritocracy and market justice
preferences at school }

% Author Orchid ID: enter ID or remove command
%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
\newcommand{\orcidauthorA}{0000-0003-1265-7854} 
\newcommand{\orcidauthorB}{0000-0002-8108-6154} 
\newcommand{\orcidauthorC}{0000-0002-9348-8361} 
\newcommand{\orcidauthorD}{0000-0002-9008-2454}

% Authors, for the paper (add full first names)
%\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{2,\ddagger}$ and Firstname Lastname $^{2,}$*}

\Author{Juan Carlos Castillo$^{1}$\orcidA{}, 
Mauricio Salgado$^{2}$\orcidB{}, 
Kevin Carrasco$^{3}$\orcidC{}~and~Andreas Laffert$^{4}$\orcidD{}
}
%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{
  Juan Carlos Castillo,
  Mauricio Salgado,
  Kevin Carrasco,
  Andreas Laffert}

% MDPI internal command: Authors, for citation in the left column
%\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
%\usepackage{xstring}
\newcommand\getfirst[1]{\getfirstaux#1\relax} \def\getfirstaux#1#2\relax{#1}
\newcommand{\authorformatted}{Castillo, \getfirst{Juan Carlos}.; Salgado, \getfirst{Mauricio}.; Carrasco, \getfirst{Kevin}.; Laffert, \getfirst{Andreas}.}
\AuthorCitation{\authorformatted}

% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address {%
$^{1}$ \quad Department of Sociology, Universidad de Chile\\
$^{2}$ \quad CEP, Centro de Estudios Públicos\\
$^{3}$ \quad COES, Center for Social Conflict and Cohesion Studies\\
$^{4}$ \quad Institute of Sociology, Pontificia Universidad Católica de Chile\\
}

%\address{
%$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
%$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
%% see: https://quarto.org/docs/journals/authors.html
\corres{Correspondence: juancastillov@uchile.cl}

% Current address and/or shared authorship
%\firstnote{Current address: Affiliation.}  % Current address should not be the same as any items in the Affiliation section.
%\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{ Previous research has shown that schools justify student
performance differences based on meritocratic ideals. However, the link
between school meritocracy and students' justification of inequalities
has received little attention. This paper examines how school and family
socialization influence students' beliefs about meritocracy and
market-based justifications for access to health, pensions, and
education in Chile. Using data from the First Study of Civic Education
in Chile, we employ a multilevel modeling strategy to analyze 5,047
students and parents nested within 231 schools. Our findings reveal that
a significant proportion of Chilean students agree with market justice
principles, at a higher rate than adults. Most students endorse
meritocratic views, particularly the reward of effort. Multilevel
estimations show strong associations between meritocratic beliefs and
market justice preferences: students who believe in the reward of effort
and talent are more likely to justify inequalities based on financial
capacity. At the school level, market justice preferences are higher in
high-status schools but lower in schools with better achievement.
Additionally, the conditional influence of meritocracy beliefs weakens
in schools with higher socioeconomic status and performance. These
results suggest that meritocratic beliefs and market justice preferences
begin early and are influenced by the school environment. }

% Keywords
\keyword{market
justice; meritocracy; socialization; family; schools; Chile}

\begin{document}

\section{Introduction}\label{introduction}

Since its origins, educational institutions have been related to the
idea of social mobility and access to better opportunities. Despite
this, the consistent evidence of the high level of social reproduction
at the school level represents a threat to the promise of education and
a meritocratic system \citep{bourdieu_reproduction_1990}. A large part
of the research in this field at an international level has addressed
the extent to which the social origin of students affects their academic
results and their life opportunities \citep{vonhippel_test_2019},
confirming that schools have severe difficulties in closing the
socio-economic and cultural gaps of origin. Besides this socioeconomic
perspective on school opportunities, recent research has addressed to
what extent inequalities in the school context are also influencing
students' perceptions, beliefs, and attitudes: Are social inequalities
even perceived at the school context? Are they rejected by the students,
particularly those who are worst-off in socioeconomic terms? Or, Is
there evidence at the school level that social inequalities are
tolerated and even justified?
\citep{batruch_belief_2022, wiederkehr_belief_2015}.

Given that the school environment has an important focus on performance,
achievement and acknowledgment, meritocracy has been one of the
principal concepts used for understanding and even for justifying
performance differences among students. Meritocracy is a distributive
system based on the belief that people should be rewarded and promoted
based on their abilities, knowledge, and achievements
\citep{young_rise_1958}. It is often seen as a way to create equal
opportunities and fairness, as individuals can rise to positions of
power and influence based on their own merit rather than their
background or connections. However, some argue that meritocracy can
actually lead to tolerating or even justifying social inequalities, as
it can create a hierarchy where those who already have resources and
advantages are more likely to succeed
\citep{sandel_tyranny_2020, mcnamee_meritocracy_2004}. In this regard, a
great deal of academic research about meritocracy delves into the
assessment of to what extent rewards and privileges in society are
related to merit, emphasizing the so-called unfulfillable promise of
meritocracy \citep{mijs_stratified_2016}.

In the present paper, we address the role of the perception of
meritocracy on the justification of social inequalities by eighth-grade
students in Chile, a country characterized by a highly stratified
educational system. In particular, we focus on justifying inequalities
in health, education and pensions, which are traditional social policy
areas where access to better services could be justified based on
payment capacity, referred to as market justice
\citep{lane_market_1986, lindh_bringing_2023}. Most of the research in
this realm has only considered adults, leaving aside the study of how
beliefs in this field develop at student age as well as the impact of
the school context and the family as the main socialization agencies
\citep{batruch_belief_2022}. Regarding schools, the way in which they
deal with unequal conditions of origin has been linked to the
\emph{hidden curriculum} \citep{chafel_schooling_1997}, whereby students
learn about distributive norms in society and mechanisms of
justification of social differences.

Based on recent studies that relate school meritocracy to the
justification of economic inequalities in the adult population
\citep{batruch_belief_2022, wiederkehr_belief_2015}, the central
hypothesis guiding this research is that school-age students with a
higher perception of meritocracy - both at school and at the societal
level - will show a larger market justice preferences, as individual
achievement would be seen as appropriately rewarded and social
mechanisms for correcting inequalities as less necessary
\citep{batruch_belief_2022}. We focus on the student-age population as
we point out that it is possible to track down the origin of
meritocratic beliefs (and their consequences) to early socialization
processes. To this regard, we take into account the family and the
school as two main socialization agencies that play a significant role
in the socialization of cultural beliefs by transmitting cultural norms,
values, and expectations to young people.

The Chilean case is particularly intriguing for studying market justice
preferences. This country is characterized by acute and persistent
economic inequality, which stands out in Latin America and among OECD
countries \citep{flores_top_2020}. In Chile, the poorest 50\% captures
only 10\% of the total income and has negative wealth, while the richest
1\% receives almost 27\% of the income and holds 49.6\% of the wealth
\citep{chancel_world_2022}. Much of this inequality has been attributed
to the deep neoliberal reforms that institutionalized privatization and
commodification of various economic sectors. Such reforms where
introduced during the dictatorship (1973-1989) and expanded in democracy
through concessions, demand credits, and specific regulatory frameworks
\citep{ffrench-davis_reformas_2018}. This shift in economic policy has
allowed the unprecedented emergence of markets in areas related to
social reproduction, such as health, pensions, and education, with
provision and access managed by private entities and segmented by
individual payment capacity, heavily reliant on State subsidies
\citep{boccardo_30_2020}. In health, although the majority of the
population uses the public insurance system (78.9\%), 15.3\% are served
by private insurers \citep{observatoriosocial_estadisticas_2024}. The
pension system is based on individual capitalization, with mandatory
contributions managed by private administrators investing in the
financial market, currently involving 11 million contributors
\citep{superintendenciadepensiones_estadisticas_2024}. In education,
30.6\% of school enrollment is in public schools, 54.0\% in
state-subsidized (voucher) private schools, and 9.3\% in fully private
schools, generally attracting higher-income groups
\citep{ministeriodeeducacion_resumen_2023}.

\subsection{Justification of inequality and market
justice}\label{justification-of-inequality-and-market-justice}

The justification of social inequality based on market-type criteria has
been conceptualized as the individuals' adherence to the deservingness
of social goods and services (such as health, education, and pensions)
based on prices and individuals' ability to pay
\citep{boltanski_new_2005, lane_market_1986, streeck_citizens_2012}.
Research on social stratification beliefs, which explore individual
perceptions of who deserves what and why \citep{kluegel_beliefs_1987},
highlights that people's explanations and justifications of social
inequality are closely tied to their judgments of deservingness. The
influence of ideologies \citep{wegener_dominant_1995} and cultural
schemas \citep{homan_being_2017} is pivotal in shaping these
explanations by offering symbolic representations that frame societal
structures and expectations. While significant attention has been paid
to wage inequality, income distribution, and payment differentials in
the literature
\citep{castillo_legitimacy_2011, evans_family_2010, jasso_how_1999, shariff_income_2016},
there has been less examination of public beliefs about which life
domains should be governed by market relations
\citep{lindh_bringing_2023} and even less about children's acceptance or
rejection of these market principles. This oversight is notable given
the extensive encroachment of market logic into public goods, welfare
policy, and social services over the past five decades
\citep{centeno_arc_2012, harvey_breve_2015}, affecting areas such as
pensions, health services, and education.

There are substantial differences in funding and delivery methods in the
management of social services across nations
\citep{jensen_worlds_2008, stoy_worlds_2014}. Nordic countries, for
example, predominantly employ public agencies to produce and provide
social services, funding these through collective taxation and offering
them in kind to the majority of citizens. This system prioritizes social
justice, placing it above market mechanisms in accessing services. In
contrast, other countries rely more heavily on for-profit entities and
private funding, where service distribution depends mainly on individual
financial capacity to pay user fees, highlighting the influence of
market justice in service allocation. The trend toward marketization of
welfare services has been growing since the 1980s
\citep{salamon_marketization_1993}, and this shift is increasingly
evident even in countries where market solutions have traditionally had
a minor role in social policy \citep{sivesind_changing_2017}. The
expansion of marketization has been related to a larger justification of
market mechanisms, whereby societies with larger private spending on
services show larger market justice preferences
\citep{lindh_public_2015}.

Robert E. Lane proposed the underpinnings of the concept of market
justice, which he differentiated from political justice. For him, ``it
is the genius of the market to stimulate wants without at the same time
stimulating a sense of deserving more than one gets''
\citeyearpar[p.~384]{lane_market_1986}. Contrary to the evidence that
unequal distribution produces feelings of dissatisfaction, anger, and
resentment that might motivate forms of collective action
\citep{greitemeyer_subjective_2016, mishra_subjective_2015, smith_relative_2012, power_deprivationprotest_2018},
Lane pointed out that in market settings, social comparisons are more
likely to motivate increased effort rather than feelings of acute
injustice because individuals attribute outcomes to their actions. In
this sense, unequal levels of well-being would be, to some extent, a
function of their talents and efforts, instead of being based on
distributive principles that characterize welfare states, such as need
and equality (see \citep{wilson_role_2003}).

Despite high-income inequality and limited social mobility in Chile, and
in Latin America in general, there is a prevalent belief that
individuals are solely responsible for their economic outcomes, a view
that varies across the region
\citep{bucca_merit_2016, chong_mystery_2008, torche_intergenerational_2014, salgado_inequality_2023}.
The reliance on private welfare providers and widespread user fees
\citep{molyneux_neoliberal_2008} adds complexity to this context, as
reflected in surveys conducted by the Center for Public Studies (CEP).
According to this data, 35.9\% prefer private health insurance, and 63\%
would prefer private education
\citep{centrodeestudiospublicos_estudio_2024}. Yet, research on
children's justification in this area remains limited, highlighting a
significant gap in understanding how younger generations view
market-based access to welfare and whether these views are associated
with their meritocratic beliefs.

\subsection{Meritocratic perceptions and market
justice}\label{meritocratic-perceptions-and-market-justice}

The original definition of merit is a combination of effort and talent
\citep{young_rise_1958}, and a meritocracy is a distributive system
where merit is the main criterion for allocating valuable goods and
rewards. From a sociological perspective, meritocracy has been used in
research on social mobility to characterize societies with low mobility
that threaten the meritocratic ideal \citep{goldthorpe_myth_2003}. More
recently, sociology and social psychology research has attended to
subjective aspects related to the support for meritocratic principles in
different societies, such as beliefs in meritocracy
\citep{castillo_multidimensional_2023, mijs_unfulfillable_2016, mijs_paradox_2019}.

Meritocratic beliefs can cover two types of subjective processes:
preferences and perceptions \citep{castillo_multidimensional_2023}.
While meritocratic preferences refer to a justification of distribution
based on merit criteria (effort and talent), the perception of
meritocracy refers to how individuals view and understand the concept of
meritocracy in their society
\citep{duru-bellat_who_2012, castillo_meritocracia_2019}. The perception
can vary greatly depending on individual experiences, as well as social,
economic, and cultural background. Some people may see meritocracy as a
fair and just system that allows anyone to succeed based on their
abilities and hard work. In contrast, others may view it as a myth or a
cover for existing power dynamics and inequality, serving to maintain
and even reinforce inequality
\citep{lampert_meritocratic_2013, mijs_paradox_2019}. Some studies have
analyzed how those with greater privileges believe more in meritocracy
\citep{reynolds_perceptions_2014}, how greater economic inequality
increases meritocratic beliefs \citep{mijs_paradox_2019}, and how larger
inequality decreases it \citep{morris_representing_2022}.

A larger justification of meritocratic distribution has been related to
less support for redistributive compensation systems
\citep{frank_performance_2015}, as individual achievement would be seen
as rewarded and social policies as less necessary.
\citet{almas_cutthroat_2020} found that in the US the highly educated
accept inequality significantly more than the less educated because they
perceive inequality as justifiable owing to differences in productivity
(i.e., merit). \citet{barr_effect_2020} found that in unequal societies,
the highly educated accept inequality more than the low educated.
Conversely, individuals tend to support redistribution when they believe
that the disadvantaged lack the opportunities to succeed
\citep{evans_strong_2018}.

Schools contribute to institutionalizing and reproducing inequality by
promoting values, norms, practices, and languages familiar to
higher-class families because the dominant group's culture shapes
educational institutions \citep{bourdieu_reproduction_1990}. Middle- and
upper-class students are better equipped to face academic challenges and
are more familiar with academic expectations
\citep{mikus_children_2019}. Such familiarity represents cultural
capital in educational contexts because higher-status students come to
school ready to meet these expectations and reap the benefits
\citep{jack_no_2016, khan_privilege_2011}. Conversely, lower-status
children lacking cultural capital must catch up while experiencing
inequitable comparisons \citep{goudeau_hidden_2017}. Additionally,
academic achievement is treated as the outcome of dispositional factors
(e.g., pupils' efforts and talents or lack of them) rather than the
result of differential access to critical resources. Due to the
meritocratic frame schools encourage, both low- and high-status
individuals tend to believe that success or failure is not due to the
family background but rather to differences in efforts and talents
\citep{darnon_where_2018}. In this sense, we believe that the perception
of meritocracy can influence students' judgments about market justice
preferences, leading to our first hypothesis:

\(H_{1a}\): Students who perceive that there is more meritocracy in
society will show larger market justice preferences

The perception of meritocracy has been mostly studied with general
questions about reward allocation based on effort and talent (usually
intelligence). Nevertheless, when looking at the school population it is
possible to further consider the perception of meritocracy referred
specifically to the school context. For instance,
\citet{resh_sense_2010} find that perception of justice in grades has a
positive effect on liberal democratic orientation, and trust in people
and formal institutions \citep{resh_sense_2014}. Attending to this
evidence, we differentiate in this study between meritocratic perception
in society at large, and meritocratic perception at school, proposing
the next hypothesis:

\(H_{1b}\): Students who perceive that there is more meritocracy at
school will show larger market justice preferences

\subsection{The socialization of market justice: families and
schools}\label{the-socialization-of-market-justice-families-and-schools}

Attending to the socialization within the family, the classic work of
Kohn showed that middle-class parents value the expression of internal
states and emotions, such as self-control, curiosity, happiness, and
consideration, while working-class parents promote deference, obedience,
and conformity to authority \citep{kohn_social_1963, kohn_class_1969}.
Although parents from all social backgrounds encourage individualism in
their children, this shared norm translates into different forms in high
and low social classes \citeyearpar{kohn_class_1969}.
\citet{acemoglu_obedience_2021} claimed that the values families impart
to their children interact with social mobility. On the one hand,
children from privileged families are socialized to adopt a clear
conception of individualism that highlights their internal states,
independence, and idiosyncrasies. In contrast, children from
disadvantaged families are socialized to support a more balanced view of
individualism that considers personal characteristics as resources to
overcome collective impediments on the path to upward mobility
\citep{iacoviello_collectivism_2019}. In the same line,
\citet{almas_fairness_2017} found that adolescents from
low-socioeconomic-status families are likelier to have an egalitarian
fairness view and consider an equal distribution as fair in a situation
with unequal merits. Taking this into account, we believe that there
will be differences in the socialization of values according to
socioeconomic differences in families that could influence market
justice preferences, which leads us to the following hypothesis:

\(H_2\): Students from families of higher social status will show larger
market justice preferences.

Recent empirical research has demonstrated that the institutional design
of schools, coupled with the meritocratic ideology it fosters,
significantly influences children's and adolescents' views on inequality
and deservingness. For example, \citet{jonsson_institutional_2015} study
revealed that higher-status adolescents in Sweden tend to perpetuate
social class stereotypes while describing the vocational and academic
tracks. Academic track students are depicted as wealthy, intelligent,
ambitious, and diligent, while vocational track students are
characterized as poor, unambitious, unintelligent, and lackadaisical.
These stereotypes help individuals maintain a sense of superiority over
others and legitimize the prevailing social hierarchies and economic
disparities \citep{jost_attitudinal_2000}. In this line, and besides
status, we also expect that schools that exhibit average better
performance scores in standardized large-scale assessments would also
show larger market justice preferences. The correspondent hypotheses
are:

\(H_3\): Students from schools of higher social status will show larger
market justice preferences.

\(H_4\): Students from schools with an average better performance in
standardized achievement tests will show larger market justice
preferences.

The last set of hypotheses deals with the moderation effects. We propose
that the link between meritocracy and market justice could be stronger
for those of high status and attending schools with larger achievement
scores. In this regard, the effects of status and merit would not be
independent, as it would be expected that those who succeed in terms of
educational rewards and career paths - usually those of higher status -
perceive more meritocracy, which would reflect on larger market justice
preferences. In similar terms, schools of better performance would
promote meritocratic perception, strengthening market justice
preferences:

\(H_5\): The family's social status will moderate the effect of
meritocratic perception on market justice preferences.

\(H_6\): The school's status will moderate the effect of meritocratic
perceptions on market justice preferences.

\(H_7\): The school's performance in standardized achievement tests will
moderate the effect of meritocratic perceptions on market justice
preferences.

Figure~\ref{fig-hypotheses} depicts a summary of the research hypothesis
described before. Market justice preferences are our main concept under
study (vis-a-vis dependent variable), which refers to the justification
of better access to health, education, and pensions based on payment
capacities (as we detail further in the methods section). The
explanatory concepts (independent variables) are at two levels: student
(individual) variables and school (context) variables. The moderation
hypotheses are represented through arrows pointing to the arrow between
meritocratic perception and market justice preferences, which will be
estimated through interaction effects.

\begin{figure}

\centering{

\includegraphics[width=6.25in,height=\textheight]{input/img/hypothesis2.png}

}

\caption{\label{fig-hypotheses}Summary of hypotheses}

\end{figure}%

\section{Data, variables and method}\label{data-variables-and-method}

\subsection{Data}\label{data}

The First Study of Civic Education in Chile, conducted by the Agency for
Quality Education of the Ministry of Education in 2017, serves as the
main data source. This database is composed by a civic knowledge test
score and a series of items batteries measuring different aspects of
citizenship. The target population includes 8th-grade students in 242
schools nationwide, as well as their parents and teachers. Initially,
the database contains 8,589 student observations and 6,770 parent
observations. To ensure higher data quality and considering the survey's
unit of analysis, we removed 171 student cases and 79 parent cases that
exhibited repetitive and careless response patterns
\citep{gottfried_autocorrelation_2022}. Additionally, we utilized data
from the System for Measuring the Quality of Education (SIMCE) conducted
by the Ministry of Education in 2017, which provides school-level
information such as administrative dependence, socioeconomic
classification, and results obtained in standardized mathematics and
language tests. After processing the variables and removing missing
cases, the final database for this study includes a two-level stratified
sample composed of 5,047 students and parents (level 1) nested within
231 schools (level 2).

\subsection{Variables}\label{variables}

\textbf{\emph{Individual level}}

\textbf{Market Justice Preferences}: The dependent variables in this
study are market justice preferences. This construct is measured through
three variables that address the degree of justification regarding
whether access to social services in pensions, education, and health
should be income conditional. Justification of inequality in health is
measured by the item: ``Is it fair in Chile that people with higher
incomes can access better healthcare than those with lower incomes?''
The same question is asked for education and pensions. In all cases,
respondents indicate their preferences on a Likert scale ranging from
``completely disagree'' (1) to ``completely agree'' (4). Additionally,
we include a summarized indicator of ``market justice preferences'',
measured by an average index across these items (α = 0.86), with values
ranging from 1 to 4, where higher values represent greater preferences
for market justice (see Table~\ref{tbl-desc-dependientes}). We analyzed
these items independently as well as by the average index.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.4286}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2551}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2143}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1020}}@{}}
\caption{Dependent
variables}\label{tbl-desc-dependientes}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
It is just that in Chile people with higher incomes can have better
pensions than people with low incomes &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly disagree\\
2. Disagree\\
3. Agree\\
4. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1837 (30.6\%)\\
1945 (32.4\%)\\
1622 (27.0\%)\\
608 (10.1\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
6012\\
(95.9\%)\strut
\end{minipage} \\
It is just that in Chile people who can pay have a better education for
their children & \begin{minipage}[t]{\linewidth}\raggedright
1. Strongly disagree\\
2. Disagree\\
3. Agree\\
4. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1766 (29.7\%)\\
1732 (29.1\%)\\
1704 (28.6\%)\\
750 (12.6\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
5952\\
(94.9\%)\strut
\end{minipage} \\
It is just that in Chile people with higher incomes can access better
health services than people with low incomes &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly disagree\\
2. Disagree\\
3. Agree\\
4. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
2254 (38.0\%)\\
1685 (28.4\%)\\
1401 (23.6\%)\\
593 (10.0\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
5933\\
(94.6\%)\strut
\end{minipage} \\
Market Justice Preferences & \begin{minipage}[t]{\linewidth}\raggedright
Mean (sd) : 2.2 (0.9)\\
min \textless{} med \textless{} max:\\
1 \textless{} 2 \textless{} 4\\
IQR (CV) : 1.7 (0.4)\strut
\end{minipage} & 13 distinct values &
\begin{minipage}[t]{\linewidth}\raggedright
6077\\
(96.9\%)\strut
\end{minipage} \\
\end{longtable}

\textbf{Perception of Meritocracy}: The main independent variable refers
to the perception of meritocracy, operationalized through five items
addressing the perception of rewards based on talent and intelligence at
both the school and societal levels. At the school level, students
respond to whether ``Intelligence is important for getting good grades''
and ``Effort is important for getting good grades''. At the societal
level, students respond to the following questions: ``In Chile, people
are rewarded for their effort'', ``In Chile, people get what they
deserve'' and ``In Chile, people are rewarded for their intelligence and
skills''. Each item was answered on a four-point Likert scale ranging
from ``completely disagree'' (1) to ``completely agree'' (4).

\textbf{Family Socioeconomic Status}: The socioeconomic status of
students' families is measured using two indicators. First, the highest
educational level attained by the parents, with categories: ``8th grade
or less,'' ``Secondary education,'' ``Technical higher education,''
``University or postgraduate,'' and ``No response.'' The inclusion of
the ``No response'' category is due to its high frequency in the data;
omitting it could obscure relevant associations. Second, the number of
books in the household is used, categorized as ``Less than 25'' and
``More than 25.''

Table~\ref{tbl-desc-independent} shows the individual level variables
used, their response categories and their frequencies.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3925}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3084}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1963}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1028}}@{}}
\caption{Individual level
variables}\label{tbl-desc-independent}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Intelligence is important to get good grades &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly disagree\\
2. Disagree\\
3. Agree\\
4. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
367 ( 6.1\%)\\
920 (15.3\%)\\
2970 (49.4\%)\\
1760 (29.3\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
6017\\
(95.9\%)\strut
\end{minipage} \\
Effort is important to get good grades &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly disagree\\
2. Disagree\\
3. Agree\\
4. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
109 ( 1.8\%)\\
88 ( 1.5\%)\\
1427 (23.7\%)\\
4406 (73.1\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
6030\\
(96.1\%)\strut
\end{minipage} \\
In Chile, people are rewarded for their intelligence and skill &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly disagree\\
2. Disagree\\
3. Agree\\
4. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
517 ( 9.0\%)\\
1568 (27.3\%)\\
2673 (46.6\%)\\
983 (17.1\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
5741\\
(91.5\%)\strut
\end{minipage} \\
In Chile, people are rewarded for their efforts &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly disagree\\
2. Disagree\\
3. Agree\\
4. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
512 ( 8.7\%)\\
1733 (29.4\%)\\
2607 (44.2\%)\\
1050 (17.8\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
5902\\
(94.1\%)\strut
\end{minipage} \\
In Chile, people get what they deserve &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly disagree\\
2. Disagree\\
3. Agree\\
4. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
604 (10.5\%)\\
1911 (33.1\%)\\
2388 (41.4\%)\\
871 (15.1\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
5774\\
(92.1\%)\strut
\end{minipage} \\
Parental educational level & \begin{minipage}[t]{\linewidth}\raggedright
1. 8th grade or less\\
2. Secondary Education\\
3. Higher tec. education\\
4. University or Postgraduat\\
5. Missing\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
559 ( 8.9\%)\\
1698 (27.1\%)\\
960 (15.3\%)\\
1080 (17.2\%)\\
1975 (31.5\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
6272\\
(100.0\%)\strut
\end{minipage} \\
Number of books at home & \begin{minipage}[t]{\linewidth}\raggedright
1. Les than 25\\
2. More than 25\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3920 (63.2\%)\\
2281 (36.8\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
6201\\
(98.9\%)\strut
\end{minipage} \\
\end{longtable}

\textbf{\emph{Contextual level}}

This study focuses on two school-level characteristics: socioeconomic
status and academic performance. Socioeconomic status is assessed using
the Ministry of Education's classification, measured as an ordinal item
with five categories ranging from ``low'' (1) to ``high'' (5). Academic
performance is measured using the school's results in the SIMCE (System
of Measurement of Educational Quality) standardized tests, which are
administered yearly at different educational levels in areas such as
language and mathematics. These results are categorized as ``low,''
``medium,'' and ``high.'' The contextual level items used, response
categories, and their frequencies are detailed in
Table~\ref{tbl-desc-school}.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.4138}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2184}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2414}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1264}}@{}}
\caption{School context variables}\label{tbl-desc-school}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Socioeconomic level of school &
\begin{minipage}[t]{\linewidth}\raggedright
1. Low\\
2. Medium low\\
3. Medium\\
4. Medium high\\
5. High\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
49 (21.1\%)\\
92 (39.7\%)\\
43 (18.5\%)\\
29 (12.5\%)\\
19 ( 8.2\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
232\\
(100.0\%)\strut
\end{minipage} \\
SIMCE score achievement by school &
\begin{minipage}[t]{\linewidth}\raggedright
1. Low\\
2. Medium\\
3. High\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
108 (46.6\%)\\
69 (29.7\%)\\
55 (23.7\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
232\\
(100.0\%)\strut
\end{minipage} \\
\end{longtable}

\textbf{\emph{Controls}}

A series of control variables are included. At the individual level, an
index of access to technology is constructed based on the number of
computers, tablets, and cell phones in the household, as well as
internet connectivity. At the contextual level, we use the
administrative dependence of schools, classified as ``Public,''
``Subsidized Private,'' or ``Private,'' and the proportion of parents
with university or postgraduate education at each school.

\subsection{Methods}\label{methods}

Given the hierarchical structure of the data, with students nested
within schools, model estimation is performed within a multilevel
framework. These models are appropriate for capturing both individual
and contextual effects in a single analysis, allowing for the estimation
of fixed effects between groups and random effects that vary from one
group to another \citep{bell_fixed_2019, hox_multilevel_2010}.
Cumulative link mixed models were employed for the ordinal dependent
variables, while linear mixed-effects models were applied for the
average market justice preference index.

The hypotheses of this research were pre-registered in the Open Science
Framework platform of the Center for Open Science (OSF), the access to
the document is available at this
\href{https://doi.org/10.17605/OSF.IO/UFSDV}{link}. The statistical
analysis for this research was conducted using the \emph{lme4} package
in R version 4.1.3.

\section{Results}\label{results}

\subsection{Descriptive}\label{descriptive}

Figure~\ref{fig-marketjustice} illustrates the distribution of responses
to the three items on market justice preferences (healthcare, pensions,
and education), indicating that most respondents tend to disagree or
strongly disagree with the idea that it is just for people with higher
incomes to have access to better services in these areas. The strongest
opposition is observed in the healthcare sector, as the majority of
respondents (66.4\%) are against the idea that higher-income individuals
should have access to better healthcare services. Similarly, regarding
the concept of justice related to access to better pensions based on
income, the level of disagreement reaches 63.0\%. This decreases
significantly in the field of education, as the percentage of
respondents who disagree or strongly disagree with those with higher
incomes having access to better education decreases to 58.8\%. Only a
small proportion agree that access to healthcare should be conditional
on income (23.6\%), and a minority strongly agree (10\%), as is the case
with pensions. However, this increases in the case of education, being
the area where there is a significant proportion of respondents who
agree or strongly agree with access based on market justice (41.2\%).

\begin{figure}

\centering{

\includegraphics{template-mdpi_files/figure-pdf/fig-marketjustice-1.pdf}

}

\caption{\label{fig-marketjustice}Market justice preferences in health,
pensions and education}

\end{figure}%

Regarding the perception of meritocracy, Figure~\ref{fig-meritocracy}
displays the frequency distribution of five items related to the
dimensions of school and society. At the school level, there is a strong
belief that both effort and talent are important for achieving good
grades, with especially high agreement for effort. Specifically, while
78.6\% agree or strongly agree on the importance of talent for obtaining
good grades, this proportion rises to 96.8\% for effort, indicating that
respondents value effort more than talent in this context.

\begin{figure}

\centering{

\includegraphics{template-mdpi_files/figure-pdf/fig-meritocracy-1.pdf}

}

\caption{\label{fig-meritocracy}Social and school meritocracy}

\end{figure}%

At the societal level, Figure~\ref{fig-meritocracy} shows that a
majority of respondents agree or strongly agree that individuals are
rewarded for their talents (63.7\%) and efforts (62\%). Nevertheless, a
notable proportion perceives that both effort and talent are not
adequately rewarded in society, with 38.1\% and 36.3\% respectively. The
perception of whether people get what they deserve in society is more
divided; while a slim majority (56.5\% combining agree and strongly
agree) agrees, a significant minority (43.6\%) disagrees, indicating
that this is the most contentious item among respondents. There is
general agreement that people are rewarded for talent and effort in
society, but the consensus is weaker compared to school meritocracy.

Figure~\ref{fig-bivariate} shows a series of graphs depicting the
association between the variables of market justice preferences - in
education, health, and pensions - and the variables of meritocratic
perception at school (effort and talent) and in society (effort, talent,
and deservingness) (see conceptual diagram in
Figure~\ref{fig-hypotheses}). In general we observe a positive
association between meritocracy and market justice preferences for the
societal questions, but this is not as straightforward for the school
items, particularly for the one referred to effort.

\begin{figure}

\centering{

\includegraphics{template-mdpi_files/figure-pdf/fig-bivariate-1.pdf}

}

\caption{\label{fig-bivariate}Market justice preferences in education,
health and pensions by social and school meritocracy}

\end{figure}%

\subsection{Multivariate}\label{multivariate}

This section presents the results of the multilevel models. We begin by
showing the results of the cumulative link mixed models in
Table~\ref{tbl-ordinal-reg} for the ordinal dependent variables of
justice in health, pensions, and education separately in order to
compare the effect of the predictors. Then, we present the results of
the linear mixed-effects models for the average index of preference for
market justice in Table~\ref{tbl-lineal-reg}\footnote{The table shows
  the effects of individual and contextual-level independent variables
  related to our hypotheses; however, the models also incorporate the
  control variables for each of the three dependent variables separately
  and for the average index (complete models are available in the
  \hyperref[appendix]{Supplementary Material}).}.

For the health item, the intraclass correlation (ICC) obtained for the
null model (not shown) indicates that 5\% of the total variance of this
item is associated with specific characteristics of the schools, while
it is 3\% for the pensions item and 4\% for the education item
\citep[ICC based on][p.15]{hox_multilevel_2010}. For the average index
of market justice preferences, the ICC reaches 4\%, indicating that
overall, only a small percentage of the total variance in students'
market justice preferences is associated with school characteristics,
thus limiting the possibility of finding effects at this second level of
analysis.

\begin{table}

\caption{\label{tbl-ordinal-reg}Cumulative link multilevel models of
differential acces justificacion of health, pensions, and education}

\centering{

\begin{center}
\scalebox{0.8}{
\begin{threeparttable}
\begin{tabular}{l c c c c c c}
\toprule
 & \multicolumn{2}{c}{Health} & \multicolumn{2}{c}{Pensions} & \multicolumn{2}{c}{Education} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
 & Model 1 & Model 2 & Model 1 & Model 2 & Model 1 & Model 2 \\
\midrule
School talent                             & $0.23^{***}$  & $0.22^{***}$  & $0.17^{***}$  & $0.16^{***}$  & $0.25^{***}$  & $0.25^{***}$  \\
                                          & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
School effort                             & $-0.26^{***}$ & $-0.26^{***}$ & $-0.26^{***}$ & $-0.25^{***}$ & $-0.23^{***}$ & $-0.23^{***}$ \\
                                          & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
Social talent                             & $0.16^{***}$  & $0.16^{***}$  & $0.12^{**}$   & $0.13^{**}$   & $0.14^{**}$   & $0.15^{***}$  \\
                                          & $(0.05)$      & $(0.05)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
Social effort                             & $0.11^{*}$    & $0.11^{*}$    & $0.27^{***}$  & $0.27^{***}$  & $0.15^{**}$   & $0.15^{**}$   \\
                                          & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      \\
Deservingness                             & $0.52^{***}$  & $0.51^{***}$  & $0.38^{***}$  & $0.37^{***}$  & $0.44^{***}$  & $0.43^{***}$  \\
                                          & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      \\
Parental education (Ref.= Non-university) &               &               &               &               &               &               \\
                                          &               &               &               &               &               &               \\
\quad University or posgraduate           & $0.05$        & $0.08$        & $0.14$        & $0.12$        & $0.09$        & $0.07$        \\
                                          & $(0.08)$      & $(0.08)$      & $(0.08)$      & $(0.08)$      & $(0.08)$      & $(0.08)$      \\
\quad Missing                             & $0.10$        & $0.11$        & $0.12^{*}$    & $0.11$        & $0.13^{*}$    & $0.12^{*}$    \\
                                          & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      \\
More than 25 books (Ref.= Less than 25)   & $-0.16^{**}$  & $-0.13^{*}$   & $-0.13^{*}$   & $-0.13^{*}$   & $-0.07$       & $-0.07$       \\
                                          & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      \\
Socioeconomic level (Ref.= Low)           &               &               &               &               &               &               \\
                                          &               &               &               &               &               &               \\
\quad SES Medium                          &               & $0.08$        &               & $0.02$        &               & $0.08$        \\
                                          &               & $(0.10)$      &               & $(0.10)$      &               & $(0.10)$      \\
\quad SES High                            &               & $0.71^{*}$    &               & $0.72^{**}$   &               & $0.93^{***}$  \\
                                          &               & $(0.28)$      &               & $(0.26)$      &               & $(0.28)$      \\
Achievement score (Ref.= Low)             &               &               &               &               &               &               \\
                                          &               &               &               &               &               &               \\
\quad Simce Medium                        &               & $-0.29^{***}$ &               & $-0.19^{*}$   &               & $-0.16^{*}$   \\
                                          &               & $(0.08)$      &               & $(0.08)$      &               & $(0.08)$      \\
\quad Simce High                          &               & $-0.56^{***}$ &               & $-0.53^{***}$ &               & $-0.46^{***}$ \\
                                          &               & $(0.10)$      &               & $(0.09)$      &               & $(0.10)$      \\
\midrule
Controls                                  & Yes           & Yes           & Yes           & Yes           & Yes           & Yes           \\
BIC                                       & $12747.58$    & $12761.47$    & $13103.81$    & $13112.87$    & $13329.33$    & $13352.18$    \\
Num.obs                                   & $5144$        & $5144$        & $5177$        & $5177$        & $5159$        & $5159$        \\
Num. groups: Schools                      & $231$         & $231$         & $232$         & $232$         & $231$         & $231$         \\
Var: Schools (Intercept)                  & $0.11$        & $0.05$        & $0.09$        & $0.04$        & $0.10$        & $0.05$        \\
\bottomrule
\end{tabular}
\begin{tablenotes}[flushleft]
\scriptsize{\item Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. $^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$ \\ \item Source: own elaboration based on First Study of Civic Education 2017.}
\end{tablenotes}
\end{threeparttable}
}
\caption{}
\label{table:coefficients}
\end{center}

}

\end{table}%

Model 1 in Table~\ref{tbl-ordinal-reg} for the three dependent variables
of justice in differential access to health, pensions, and education
demonstrates a similar trend regarding the effect of the perception of
meritocracy on these items, with all effects being statistically
significant (\emph{p} \textless{} 0.05). Regarding the variables of
social meritocracy, the three variables---talent, effort, and
deservingness---show that as these increase, the justification for
differentiated access to pensions, education, and health also increases.
In the context of school meritocracy, the effects are mixed: as
perceived school talent increases, the justification for differentiated
access to these services increases; conversely, as perceived school
effort increases, the justification decreases, holding all other
variables constant.

Additionally, Model 1 examines family status through parents'
educational level and the number of books at home. For health, pensions,
and education, a negative but statistically non-significant association
with parents' education is observed. However, having more than 25 books
at home shows a negative and statistically significant effect on market
justice in health and pensions, suggesting less justification for
differentiated access compared to those with fewer books.

Model 2 in Table~\ref{tbl-ordinal-reg} incorporates school-level
variables. The results show that high socioeconomic level schools show
greater preferences for market justice in health, pensions, and
education compared to low socioeconomic level schools, with
statistically significant results (\emph{p} \textless{} 0.05). Moreover,
and contrary to our hypothesis, schools with higher academic performance
show negative and statistically significant effects across the three
items.

\begin{table}

\caption{\label{tbl-lineal-reg}Linear mixed-effects models for
meritocracy perception and market justice preferences}

\centering{

\begin{center}
\scalebox{0.85}{
\begin{threeparttable}
\begin{tabular}{l c c c c}
\toprule
 & Model 1 & Model 2 & Model 3 & Model 4 \\
\midrule
Constant                                  & $1.31^{***}$  & $1.30^{***}$  & $1.36^{***}$  & $1.39^{***}$  \\
                                          & $(0.09)$      & $(0.10)$      & $(0.10)$      & $(0.10)$      \\
School talent                             & $0.11^{***}$  & $0.11^{***}$  & $0.10^{***}$  & $0.10^{***}$  \\
                                          & $(0.01)$      & $(0.01)$      & $(0.01)$      & $(0.01)$      \\
School effort                             & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ \\
                                          & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
Social talent                             & $0.07^{***}$  & $0.07^{***}$  & $0.07^{***}$  & $0.07^{***}$  \\
                                          & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
Social effort                             & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  \\
                                          & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
Deservingness                             & $0.20^{***}$  & $0.20^{***}$  & $0.20^{***}$  & $0.20^{***}$  \\
                                          & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
Parental education (Ref.= Non-university) &               &               &               &               \\
                                          &               &               &               &               \\
\quad University or posgraduate           &               & $0.05$        & $0.05$        & $0.05$        \\
                                          &               & $(0.03)$      & $(0.04)$      & $(0.04)$      \\
\quad Missing                             &               & $0.06^{*}$    & $0.06^{*}$    & $0.06^{*}$    \\
                                          &               & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
More than 25 books (Ref.= Less than 25)   &               & $-0.05$       & $-0.05$       & $-0.04$       \\
                                          &               & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
Socioeconomic level (Ref.= Low)           &               &               &               &               \\
                                          &               &               &               &               \\
\quad SES Medium                          &               &               & $-0.01$       & $0.01$        \\
                                          &               &               & $(0.05)$      & $(0.05)$      \\
\quad SES High                            &               &               & $0.31^{*}$    & $0.37^{**}$   \\
                                          &               &               & $(0.14)$      & $(0.13)$      \\
Achievement score (Ref.= Low)             &               &               &               &               \\
                                          &               &               &               &               \\
\quad Simce Medium                        &               &               &               & $-0.11^{**}$  \\
                                          &               &               &               & $(0.04)$      \\
\quad Simce High                          &               &               &               & $-0.26^{***}$ \\
                                          &               &               &               & $(0.04)$      \\
\midrule
Controls                                  & Yes           & Yes           & Yes           & Yes           \\
Deviance                                  & $12368.01$    & $12359.40$    & $12344.10$    & $12310.90$    \\
Deviance Test (p)                         & $0.00$        & $0.00$        & $0.00$        & $0.00$        \\
BIC                                       & $12517.46$    & $12550.41$    & $12595.11$    & $12589.48$    \\
Num.obs                                   & $5047$        & $5047$        & $5047$        & $5047$        \\
Num. groups: Schools                      & $231$         & $231$         & $231$         & $231$         \\
Var: Schools (Intercept)                  & $0.02$        & $0.02$        & $0.02$        & $0.01$        \\
Var: Residual                             & $0.66$        & $0.66$        & $0.67$        & $0.66$        \\
\bottomrule
\end{tabular}
\begin{tablenotes}[flushleft]
\scriptsize{\item Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. $^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$ \\ \item Source: own elaboration based on First Study of Civic Education 2017.}
\end{tablenotes}
\end{threeparttable}
}
\caption{}
\label{table:coefficients}
\end{center}

}

\end{table}%

Table~\ref{tbl-lineal-reg} presents the results of linear mixed-effects
models for the market justice preference index. Model 1 shows that
higher perceived societal meritocracy, through effort, talent, and
deservingness, is positively and significantly associated with market
justice preferences. School meritocracy perceptions yield mixed effects:
perceived talent increases market justice preference, while perceived
effort decreases it, both statistically significant (\emph{p}
\textless{} 0.001). Model 2 includes family status, finding no
statistically significant effects from parents' education or the number
of books at home. Regarding school characteristics, Models 3 and 4
introduce socioeconomic level and academic performance. The results of
Model 3 suggest a statistically significant relationship between school
socioeconomic levels and market justice preferences, specifically those
schools with high socioeconomic levels (\(\beta\) = 0.37) compared to
low socioeconomic levels. However, Model 4 indicates that schools with
medium (\(\beta\) = -0.11) and high (\(\beta\) = -0.26) academic
performance present lower market justice preferences compared to
low-performing schools, with these effects being statistically
significant when controlling for other variables.

\begin{table}

\caption{\label{tbl-interact}Interactions effects}

\centering{

\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabu} to \linewidth {>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{5}{c}{\textbf{Market Justice Preferences}} \\
\cmidrule(l{3pt}r{3pt}){2-6}
Variable & School talent & School effort & Social talent & Social effort & Deservingness\\
\midrule
\textbf{University or Postgraduate} & -0.034 & -0.012 & -0.073. & -0.079* & -0.118**\\
\cmidrule{1-6}
\textbf{More than 25 books} & -0.015 & -0.046 & -0.029 & -0.081** & -0.035\\
\cmidrule{1-6}
\textbf{School SES high} & 0.004 & 0.001 & -0.057. & -0.063* & -0.08**\\
\cmidrule{1-6}
\textbf{Simce Medium} & 0.001 & 0.032 & -0.037 & -0.041 & -0.082*\\
\cmidrule{1-6}
\textbf{Simce High} & -0.016 & 0.008 & -0.043 & -0.063. & -0.087*\\
\bottomrule
\multicolumn{6}{l}{\textsuperscript{} Note: ***p < 0.001, **p < 0.01, *p < 0.05. Source: own elaboration based on First Study of Civic}\\
\multicolumn{6}{l}{Education 2017.}\\
\end{tabu}}

}

\end{table}%

Table~\ref{tbl-interact} displays the interaction terms related to our
hypotheses. The results suggest that students' family background
moderates the relationship between their meritocratic perceptions and
their justification of market allocation in social services.
Specifically, the relationship between social effort and market justice
preferences is less positive for students whose parents have university
or postgraduate education. Lower-status students (measured by parental
education) prefer more market justice when they strongly adhere to
meritocratic perceptions of deservingness. A similar moderating effect
is seen with family cultural capital: the relationship between effort
and market justice preferences is less positive for students with more
than 25 books at home.

At the school level, we observe that school status and students'
meritocratic beliefs moderate their preferences for market justice.
High-status schools show less preference for market justice when their
students, on average, have a greater perception of meritocracy, as
measured by effort, deservingness, and talent. Conversely, low-achieving
schools prefer more market justice when their students have stronger
meritocratic beliefs. Figure~\ref{fig-interaction} illustrates this
moderating effect of school achievement on the relationship between
deservingness and market justice. No moderating effects were observed at
the school level for students' meritocratic beliefs regarding effort or
talent in achieving good grades.

\begin{figure}

\centering{

\includegraphics{template-mdpi_files/figure-pdf/fig-interaction-1.pdf}

}

\caption{\label{fig-interaction}Interaction between deservingness and
justification of inequality by school SIMCE achievement}

\end{figure}%

\section{Discussion}\label{discussion}

The previous analyses aimed to explore the relationship between
perceptions of meritocracy and preferences for market justice among
students in Chilean society. Several models estimated how students'
beliefs about the rewards for talent and effort in both societal and
school contexts influence their attitudes toward access to social
services like health, education, and pensions based on pay capacity.
Additionally, the models examined the impact of students' socioeconomic
status and school achievement on these preferences, assessing whether
perceptions of meritocracy moderate these effects. The findings offer
insights into the complex interplay between meritocratic beliefs,
socioeconomic factors, and attitudes toward market justice criteria.

Hypothesis \(H_{1a}\) posited that students who perceive greater
meritocracy in society would show stronger preferences for market
justice. The results support this hypothesis, indicating a positive
relationship between the perception that talent and effort are rewarded
in society and the preferences for market justice in access to services
such as health, education, and pensions. This finding aligns with
previous research suggesting that individuals who believe in a
meritocratic system are more likely to justify social inequalities
\citep{batruch_belief_2022, wiederkehr_belief_2015}. It is noteworthy
that among the items measuring meritocratic perception, consistently the
one about deservingness shows a larger effect when compared to the ones
about talent and effort. Although the three items correlate, there is
something distinctive about deservingness that probably encompasses
other aspects beyond meritocracy, an aspect to which we will come back
in the conclusions. Moving now to hypothesis \(H_{1b}\), this is simply
a variation of 1a, replacing the perception of meritocracy at society by
the one at school. In this case we have only two items of meritocratic
perception - unfortunately deservingness was not measured at school
level - and the direction of the effects is not as straightforward as in
the case of societal meritocracy in \(H_{1a}\): positive for talent, but
negative for effort. Still, one of the limitation for the comparison
between perception of meritocracy and society is the difference of the
items' phrasing. Whereas for society the item refers to the extent to
which efforts are rewarded, at school is about the importance of effort,
whereby something that is important does not necessarily mean that it is
enough to be rewarded. Furthermore, the school-effort item is the one
with the lowest variation among all meritocratic items (more than 95\%
agree or strongly agree), and with the lowest effect size.

Hypothesis \(H_2\) and hypothesis \(H_3\) propose that socioeconomic
status could influence students' preferences, but the results show
different patterns. On the one hand, \(H_2\) proposed that students from
higher socioeconomic status families would show stronger preferences for
market justice, which is not supported by the results. Although students
from families with higher educational levels and more economic resources
tend to prefer less market justice, this relationship is not
statistically significant. This might be due to the influence of other
contextual and cultural factors that modulate attitudes toward market
justice \citep{kohn_social_1963, kohn_class_1969, almas_fairness_2017}.
Notably, socioeconomic status measured by parental educational level and
the number of books at home do not explain variations in market justice
preferences, suggesting that other variables, such as market justice
preferences of parents, could play an essential role in the attitudinal
socialization process. On the other hand, Hypothesis \(H_3\) held that
students from schools with higher socioeconomic status would show
stronger preferences for market justice. The results confirm this
hypothesis since students from high-status schools tend to prefer more
market justice, compared to low-status schools. This finding, in line
with the existing literature
\citep{jonsson_institutional_2015, jost_attitudinal_2000}, does allow us
to affirm the hypothesis about the role of the school environment,
particularly socioeconomic segregation, in influencing attitudes towards
market justice and the acceptance of inequalities. A possible
explanation is the role of the context at the school level, which could
have a more relevant role than the family socioeconomic status, since
the highly stratified nature of the Chilean education system reinforces
socioeconomic divisions and socializes students into different normative
frameworks.

Hypothesis \(H_4\) held that students from schools with higher school
achievement would show stronger preferences for market justice,
following the argument that such schools could foster meritocratic
attitudes by showing a larger promotion of talent and effort in order to
get better results. Nevertheless, the results show quite the opposite:
schools with better academic results (as measured by a national
standardized test) show in average less preferences for market justice.
This finding is one of the most strongest and consistent throughout the
models and it is relevant given the centrality of school performance in
educational institutions. One possible explanation of this link is that
students in high-achieving schools may be more exposed to discussions
and a deeper understanding of inequality and social justice, leading to
a critical perspective toward the weight of market mechanisms in the
access to social services.

Hypotheses \(H_5\) and \(H_6\) proposed that socioeconomic status would
moderate the effect of meritocratic perceptions on market justice
preferences. Hypothesis \(H_5\) refers to family socioeconomic status,
whereby the results show a negative interaction with two of the three
perceptual variables at the societal level. This means that the positive
impact of the perception of social meritocracy on market justice
preferences had a lower effect on students with parents with higher
education and also on those with more books at home. Something similar
occurs with hypothesis \(H_6\), where now school-level status moderates
the effect of meritocracy variables, which in multilevel framework is
referred as to cross-level interaction. The negative interaction in this
context means that in schools of low socioeconomic levels, the positive
effect of the social meritocracy variables on market justice preferences
is stronger. These relations indicate that, although those who perceive
society as meritocratic have greater preferences towards market justice,
this influence is lower among those with a higher socioeconomic level.
This finding highlights the complex interaction between socioeconomic
status, school context, meritocratic beliefs, and preferences toward
market justice. The moderating role of meritocratic perceptions suggests
that students' views on social justice are shaped by their socioeconomic
background and beliefs about how society rewards effort and talent. This
underlines the importance of addressing structural inequalities and
individual beliefs in educational and social policies.

Hypothesis \(H7\) posited that the academic performance of schools could
moderate the effect of meritocratic perceptions on preferences for
market justice. This hypothesis is based on the argument that the level
of academic performance in schools entails a differentiated promotion of
meritocratic perceptions to achieve better results, which would, in
turn, reinforce or diminish preferences for market justice. The results
of the interactions between levels suggest that, indeed, in
higher-performing schools, the effect of meritocratic perceptions on the
justification of service provision based on market principles is weaker
compared to lower-performing schools. Similarly to the effect of
academic performance in \(H4\), a possible explanation for this
moderation is that students in high-performing schools may be more
exposed to knowledge about the effects of inequality and social justice.
This exposure translates into an enlightening effect of education on
inequality, which can mitigate the impact of meritocracy on market
justice preferences.

\section{Conclusion}\label{conclusion}

Our research aimed to examine the relationship between the perception of
meritocracy and the justification of market justice preferences among
eighth-grade students in Chile. From our knowledge this is the first
study addressing market justice preferences at school level, finding
that students exhibit large preferences for market justice (about one
third agree or strongly agree with them), a share that is quite larger
when compared with evidence in adult population. Perception of
meritocracy is also high, particularly when it comes to the reward of
effort, with an striking eighty six percent of agreement. Regarding
these descriptive indicators, a a first message coming from this study
is that there is something particular in school-age when it comes to
perceptions and preferences about inequality. This can relate to
development aspects
\citep[\citet{rizzo_children_2020}]{kim_socioeconomic_2020} but it could
be something specific to the Chilean case, which requires further
comparative as well as longitudinal research in the area. In any case,
teaching about the origins and characteristics of social inequality is
not part of the chilean school curriculum, calling for a more dedicated
consideration of these topics in areas such as citizensip education as
well as in history and social science education.

The core of this paper was the association between meritocratic
perceptions and preferences for market justice. In general, we find that
those who perceive that effort and talent/intelligence are rewarded, are
more willing to agree with that richer individuals can have better
health, education, and pensions. This is not the first time that
meritocratic perceptions (and also beliefs) have been related with the
legitimation of social inequalities
\citep{mijs_paradox_2019, darnon_where_2018}, but the consideration of
school level population as well as market justice preferences allows
adding evidence regarding the role of meritocracy regarding inequality
beliefs: legitimation of market criteria in social services seems to
begin at school age, and what students think of meritocracy is linked to
it. Although there are still several open questions, as for instance
differences between meritocratic perceptions in school and in the
society at large, this raises up the relevance of what is done (or not)
in schools in order to challenge the meritocratic ideals.

The consideration of the school context opened some caveats in this
research. We found contrary evidence for the association between status,
school performance, and market justice preferences: high status and high
achievement at the school level are associated with less market justice
preferences. This is opposed to the rational-interest argument that
suggests that those doing better - both economically and academically-
would demand less redistribution \citep[\citet{runst_does_2018},
\citet{bullock_education_2021}]{gonthier_parallel_2017} and therefore
are expected to challenge market mechanisms. A possible explanation is
that education contributes to forming a critical view of markets in
redistribution, nevertheless, further research in this area is needed to
assert the specific mechanisms that might be playing a role. All in all,
from this study we know that school contexts are not trivial in this
regard, calling for larger attention to school-age research on the
formation of inequality preferences and beliefs.

%\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work~reported.}

\authorcontributions{Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: \getfirst{Juan Carlos}.\getfirst{Castillo}.:   Conceptualization, Supervision, Methodology, Writing
- review and
editing; \getfirst{Mauricio}.\getfirst{Salgado}.:   Conceptualization, Methodology, Writing
- review and
editing; \getfirst{Kevin}.\getfirst{Carrasco}.:   Methodology, formal
analysis, software, visualization, Writing - review and
editing; \getfirst{Andreas}.\getfirst{Laffert}.:   Methodology, formal
analysis, software, visualization, Writing - review and editing}

%\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''. Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}

\funding{The authors acknowledge the support of ANID/FONDECYT grant
1210847 and the Center for Conflict and Social Cohesion Studies (COES)
grant ANID/FONDAP 1513009.}

%\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study. You might choose to exclude this statement if the study did not require ethical approval. Please note that the Editorial Office might ask you for further information. Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans. OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals. OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.}

%\informedconsent{Any research article describing a study involving humans should contain this statement. Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans. You might also choose to exclude this statement if the study did not involve humans. Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves). Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}

%\dataavailability{We encourage all authors of articles published in MDPI journals to share their research data. In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study. Where no new data were created, or where data is unavailable due to privacy or ethical restrictions, a statement is still required. Suggested Data Availability Statements are available in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}.}

\dataavailability{The original data presented in the study are openly
available in
\url{https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/}}

% Only for journal Nursing Reports
%\publicinvolvement{Please describe how the public (patients, consumers, carers) were involved in the research. Consider reporting against the GRIPP2 (Guidance for Reporting Involvement of Patients and the Public) checklist. If the public were not involved in any aspect of the research add: ``No public involvement in any aspect of this research''.}

% Only for journal Nursing Reports
%\guidelinesstandards{Please add a statement indicating which reporting guideline was used when drafting the report. For example, ``This manuscript was drafted against the XXX (the full name of reporting guidelines and citation) for XXX (type of research) research''. A complete list of reporting guidelines can be accessed via the equator network: \url{https://www.equator-network.org/}.}

%\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections. This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}

%\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflicts of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript; or in the decision to publish the results must be declared in this section. If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the results''.}

\conflictsofinterest{The authors declare no conflict of interest.}

\begin{adjustwidth}{-\extralength}{0cm}
%\printendnotes[custom] % Un-comment to print a list of endnotes

\reftitle{References}

  \bibliography{input/bib/just-ineq-merit.bib}

\PublishersNote{}
\end{adjustwidth}\end{document}

### 2024-07-24 back_matter.qmd
Ajustes nuevos
Ajustes finales de:
- Referencias
- Cover Letter
- Funding
- Paper blinded
Archivos submit

--- 2024-07-04 back_matter.qmd
@@
 title: "The Socialization of Meritocracy and Market Justice Preferences at School"
-subtitle: "Back Matter"
+subtitle: "Supplementary Material"
 format:
@@
     indent: false
+    keep-tex: true
   hikmah-manuscript-pdf: 
     indent: false
+    keep-tex: true
 bibliography: "input/bib/just-ineq-merit.bib"
-csl: "input/bib/apa6.csl"
+csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
 link-citations: true
@@
 {{< include 07-supplementary-material.qmd >}}
-
-## Acknowledgments
-
-The authors acknowledge the support of ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) grant ANID/FONDAP 1513009.
-
-## Authors' Contributions
-
-The following statements should be used "Conceptualization, X.X.
-and Y.Y.; Methodology, X.X.; Software, X.X.; Validation, X.X., Y.Y.
-and Z.Z.; Formal Analysis, X.X.; Investigation, X.X.; Resources, X.X.; Data Curation, X.X.; Writing – Original Draft Preparation, X.X.; Writing – Review & Editing, X.X.; Visualization, X.X.; Supervision, X.X.; Project Administration, X.X.; Funding Acquisition, Y.Y.”
-
-## Conflict of Interest
-
-The authors report no competing interests to declare.
-
-## Funding
-
-This research received no external funding.
-
-## Data Avaibility Statement
-
-The original data presented in the study are openly available in <https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/>
-
-Access the data, code, and analysis presented in the study are openly available in <https://github.com/educacion-meritocracia/market-justice-preferences-school>.
-
-## References

### 2024-07-24 cover_letter.qmd
Archivos submit

@@
     indent: false
+    keep-tex: true
   hikmah-manuscript-pdf: 
     indent: false
+    keep-tex: true
 bibliography: "input/bib/just-ineq-merit.bib"

### 2024-07-25 02-antecedentes.qmd
template mdpi
agregue un nuevo qmd para el template mdpi
Ref arreglada

@@
----
-author: "Equipo EDUMER"
-bibliography: "input/bib/just-ineq-merit.bib"
-csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
----
-
 ```{r echo=FALSE}
@@
 Attending to the socialization within the family, the classic work of Kohn showed that middle-class parents value the expression of internal states and emotions, such as self-control, curiosity, happiness, and consideration, while working-class parents promote deference, obedience, and conformity to authority [@kohn_social_1963; @kohn_class_1969].
-Although parents from all social backgrounds encourage individualism in their children, this shared norm translates into different forms in high and low social classes [-@kohn_class_1969].
+Although parents from all social backgrounds encourage individualism in their children, this shared norm translates into different forms in high and low social classes [@kohn_class_1969].
 @acemoglu_obedience_2021 claimed that the values families impart to their children interact with social mobility.

### 2024-07-25 MDPI-submit/template-mdpi.tex
Ref arreglada

@@
 their children, this shared norm translates into different forms in high
-and low social classes \citeyearpar{kohn_class_1969}.
+and low social classes \citep{kohn_class_1969}.
 \citet{acemoglu_obedience_2021} claimed that the values families impart

### 2024-08-05 MDPI-submit/template-mdpi.tex
Actualización paper anonimizado en .tex y orden de repo

@@
   pdftitle={The socialization of meritocracy and market justice preferences at school},
-  pdfauthor={Juan Carlos Castillo; Mauricio Salgado; Kevin Carrasco; Andreas Laffert},
+  % pdfauthor={Juan Carlos Castillo; Mauricio Salgado; Kevin Carrasco; Andreas Laffert},
   pdfkeywords={market
@@
 
-\Author{Juan Carlos Castillo$^{1}$\orcidA{}, 
-Mauricio Salgado$^{2}$\orcidB{}, 
-Kevin Carrasco$^{3}$\orcidC{}~and~Andreas Laffert$^{4}$\orcidD{}
-}
+%\Author{Juan Carlos Castillo$^{1}$\orcidA{}, 
+%Mauricio Salgado$^{2}$\orcidB{}, 
+%Kevin Carrasco$^{3}$\orcidC{}~and~Andreas Laffert$^{4}$\orcidD{}
+%}
 %\longauthorlist{yes}
@@
 % MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{
-  Juan Carlos Castillo,
-  Mauricio Salgado,
-  Kevin Carrasco,
-  Andreas Laffert}
+%\AuthorNames{
+%  Juan Carlos Castillo,
+%  Mauricio Salgado,
+%  Kevin Carrasco,
+%  Andreas Laffert}
 
@@
 %\usepackage{xstring}
-\newcommand\getfirst[1]{\getfirstaux#1\relax} \def\getfirstaux#1#2\relax{#1}
-\newcommand{\authorformatted}{Castillo, \getfirst{Juan Carlos}.; Salgado, \getfirst{Mauricio}.; Carrasco, \getfirst{Kevin}.; Laffert, \getfirst{Andreas}.}
-\AuthorCitation{\authorformatted}
+%\newcommand\getfirst[1]{\getfirstaux#1\relax} \def\getfirstaux#1#2\relax{#1}
+%\newcommand{\authorformatted}{Castillo, \getfirst{Juan Carlos}.; Salgado, \getfirst{Mauricio}.; Carrasco, \getfirst{Kevin}.; Laffert, \getfirst{Andreas}.}
+%\AuthorCitation{\authorformatted}
 
@@
 % Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
-\address {%
-$^{1}$ \quad Department of Sociology, Universidad de Chile\\
-$^{2}$ \quad CEP, Centro de Estudios Públicos\\
-$^{3}$ \quad COES, Center for Social Conflict and Cohesion Studies\\
-$^{4}$ \quad Institute of Sociology, Pontificia Universidad Católica de Chile\\
-}
+%\address {%
+%$^{1}$ \quad Department of Sociology, Universidad de Chile\\
+%$^{2}$ \quad CEP, Centro de Estudios Públicos\\
+%$^{3}$ \quad COES, Center for Social Conflict and Cohesion Studies\\
+%$^{4}$ \quad Institute of Sociology, Pontificia Universidad Católica de Chile\\
+%}
 
@@
 %% see: https://quarto.org/docs/journals/authors.html
-\corres{Correspondence: juancastillov@uchile.cl}
+%\corres{Correspondence: juancastillov@uchile.cl}
 
@@
 stimulating a sense of deserving more than one gets''
-\citeyearpar[p.~384]{lane_market_1986}. Contrary to the evidence that
+\citep[p.384]{lane_market_1986}. Contrary to the evidence that
 unequal distribution produces feelings of dissatisfaction, anger, and
@@
 development aspects
-\citep[\citet{rizzo_children_2020}]{kim_socioeconomic_2020} but it could
+\citep{rizzo_children_2020, kim_socioeconomic_2020} but it could
 be something specific to the Chilean case, which requires further
@@
 suggests that those doing better - both economically and academically-
-would demand less redistribution \citep[\citet{runst_does_2018},
-\citet{bullock_education_2021}]{gonthier_parallel_2017} and therefore
+would demand less redistribution \citep{runst_does_2018,bullock_education_2021,gonthier_parallel_2017} and therefore
 are expected to challenge market mechanisms. A possible explanation is
@@
 available in
-\url{https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/}}
+\url{https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/}
+
+Access to the data, code, and analysis presented in the study are openly available in \url{https://github.com/educacion-meritocracia/market-justice-preferences-school}
+
+}
 

### 2024-08-05 papelera/paper.qmd
Actualización paper anonimizado en .tex y orden de repo

--- 2024-07-22 paper-blinded.qmd
@@
 format:
-  hikmah-pdf: 
-    indent: false
-  hikmah-manuscript-pdf: 
-    indent: false
+  mdpi-pdf:
+    keep-tex: true
+mdpi_journal: behavsci
+article_type: article
+article_status: submit
+author:
+  - name: Juan Carlos Castillo
+    email: juancastillov@uchile.cl
+    affiliations: 
+        name: Department of Sociology, Universidad de Chile
+        country: Chile
+    attributes:
+      corresponding: true
+  - name: Mauricio Salgado
+    email: msalgado@unab.cl
+    affiliations: 
+        name: Centro de Estudios Públicos, Chile
+        country: Chile
+  - name: Kevin Carrasco
+    email: kevin.carrasco@ug.uchile.cl
+    affiliations: 
+        name: Center for Social Conflict and Cohesion Studies
+        country: Chile
+  - name: Andreas Laffert
+    email: alaffertt@estudiante.uc.cl
+    affiliations: 
+        name: Institute of Sociology, Universidad Católica de Chile
+        country: Chile
 abstract: |
@@
 csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
-link-citations: true
-linestretch: 1.15       
-mainfont: Times New Roman
-fontsize: 12pt          
-geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
-papersize: letter
-colorlinks: true
-fig-height: 4
-fig-width: 7.5
 header-includes:
   - \usepackage{threeparttable}
+  - \usepackage{graphicx}
+  - \includegraphics{pics/logo-mdpi.pdf}
 ---

### 2024-08-05 papelera/paper.tex
Actualización paper anonimizado en .tex y orden de repo

--- 2024-07-24 MDPI-submit/template-mdpi.tex
@@
 \usepackage{xcolor}
+\usepackage{threeparttable}
+\usepackage{graphicx}
+\includegraphics{pics/logo-mdpi.pdf}
 \makeatletter
@@
   pdftitle={The socialization of meritocracy and market justice preferences at school},
-  pdfauthor={Juan Carlos Castillo; Mauricio Salgado; Kevin Carrasco; Andreas Laffert},
+  pdfauthor={Equipo EDUMER},
   pdfkeywords={market
@@
 %\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
-\newcommand{\orcidauthorA}{0000-0003-1265-7854} 
-\newcommand{\orcidauthorB}{0000-0002-8108-6154} 
-\newcommand{\orcidauthorC}{0000-0002-9348-8361} 
-\newcommand{\orcidauthorD}{0000-0002-9008-2454}
 
@@
 
-\Author{Juan Carlos Castillo$^{1}$\orcidA{}, 
-Mauricio Salgado$^{2}$\orcidB{}, 
-Kevin Carrasco$^{3}$\orcidC{}~and~Andreas Laffert$^{4}$\orcidD{}
+\Author{Juan Carlos Castillo$^{aff-1}$, 
+Mauricio Salgado$^{aff-2}$, 
+Kevin Carrasco$^{aff-3}$~and~Andreas Laffert$^{aff-4}$
 }
@@
 \address {%
-$^{1}$ \quad Department of Sociology, Universidad de Chile\\
-$^{2}$ \quad CEP, Centro de Estudios Públicos\\
-$^{3}$ \quad COES, Center for Social Conflict and Cohesion Studies\\
-$^{4}$ \quad Institute of Sociology, Pontificia Universidad Católica de Chile\\
+$^{aff-1}$ \quad , Department of Sociology, Universidad de Chile\\
+$^{aff-2}$ \quad , Centro de Estudios Públicos, Chile\\
+$^{aff-3}$ \quad , Center for Social Conflict and Cohesion Studies\\
+$^{aff-4}$ \quad , Institute of Sociology, Universidad Católica de
+Chile\\
 }
@@
 
-\includegraphics{template-mdpi_files/figure-pdf/fig-marketjustice-1.pdf}
+\includegraphics{paper_files/figure-pdf/fig-marketjustice-1.pdf}
 
@@
 
-\includegraphics{template-mdpi_files/figure-pdf/fig-meritocracy-1.pdf}
+\includegraphics{paper_files/figure-pdf/fig-meritocracy-1.pdf}
 
@@
 
-\includegraphics{template-mdpi_files/figure-pdf/fig-bivariate-1.pdf}
+\includegraphics{paper_files/figure-pdf/fig-bivariate-1.pdf}
 
@@
 
-\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
+\resizebox{\linewidth}{!}{
 \begin{tabu} to \linewidth {>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
@@
 
-\includegraphics{template-mdpi_files/figure-pdf/fig-interaction-1.pdf}
+\includegraphics{paper_files/figure-pdf/fig-interaction-1.pdf}
 
@@
 
-\authorcontributions{Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: \getfirst{Juan Carlos}.\getfirst{Castillo}.:   Conceptualization, Supervision, Methodology, Writing
-- review and
-editing; \getfirst{Mauricio}.\getfirst{Salgado}.:   Conceptualization, Methodology, Writing
-- review and
-editing; \getfirst{Kevin}.\getfirst{Carrasco}.:   Methodology, formal
-analysis, software, visualization, Writing - review and
-editing; \getfirst{Andreas}.\getfirst{Laffert}.:   Methodology, formal
-analysis, software, visualization, Writing - review and editing}
+}
 
@@
 Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
-
-\funding{The authors acknowledge the support of ANID/FONDECYT grant
-1210847 and the Center for Conflict and Social Cohesion Studies (COES)
-grant ANID/FONDAP 1513009.}
 
@@
 
-\dataavailability{The original data presented in the study are openly
-available in
-\url{https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/}}
-
 % Only for journal Nursing Reports
@@
 
-\conflictsofinterest{The authors declare no conflict of interest.}
-
 \begin{adjustwidth}{-\extralength}{0cm}

### 2024-08-05 paper.qmd
Actualización paper anonimizado en .tex y orden de repo

@@
 format:
-  hikmah-pdf: 
-    indent: false
-  hikmah-manuscript-pdf: 
-    indent: false
+  mdpi-pdf:
+    keep-tex: true
+mdpi_journal: behavsci
+article_type: article
+article_status: submit
 author:
   - name: Juan Carlos Castillo
+    corresponding: true
+    orcid: 0000-0003-1265-7854
     email: juancastillov@uchile.cl
-    orcid: 0000-0003-1265-7854
+    role:
+      - Conceptualization
+      - Supervision
+      - Methodology
+      - Writing - review and editing
     affiliations:
-        name: Department of Sociology, Universidad de Chile
+      - name: "Universidad de Chile"
+        department: Department of Sociology
+        id: 1
         country: Chile
-    attributes:
-      corresponding: true
   - name: Mauricio Salgado
-    email: msalgado@unab.cl
     orcid: 0000-0002-8108-6154
+    role:
+      - Conceptualization
+      - Methodology
+      - Writing - review and editing    
     affiliations:
-        name: Centro de Estudios Públicos, Chile
+      - name: Centro de Estudios Públicos
+        id: 2
+        department: CEP
         country: Chile
   - name: Kevin Carrasco
-    email: kevin.carrasco@ug.uchile.cl
     orcid: 0000-0002-9348-8361
+    role:
+      - Methodology
+      - formal analysis
+      - software
+      - visualization
+      - Writing - review and editing
     affiliations:
-        name: Center for Social Conflict and Cohesion Studies
+      - name: Center for Social Conflict and Cohesion Studies
+        id: 3
+        department: COES
         country: Chile
   - name: Andreas Laffert
-    email: alaffertt@estudiante.uc.cl
     orcid: 0000-0002-9008-2454
+    role:
+      - Methodology
+      - formal analysis
+      - software
+      - visualization
+      - Writing - review and editing  
     affiliations:
-        name: Institute of Sociology, Universidad Católica de Chile
+      - name: Universidad Católica de Chile
+        id: 4
+        department: Institute of Sociology
         country: Chile
+fundingstatement: | 
+  The authors acknowledge the support of ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) grant ANID/FONDAP 1513009.
+dataavailability: |
+  The original data presented in the study are openly available in <https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/>
+conflictsofinterest: |
+  The authors declare no conflict of interest.
 abstract: |
@@
 These results suggest that meritocratic beliefs and market justice preferences begin early and are influenced by the school environment.
-keywords: 
-  - market justice 
-  - meritocracy
-  - socialization
-  - family
-  - schools
-  - Chile
+keywords: [market justice, Meritocracy, socialization, family, schools, Chile]
 bibliography: "input/bib/just-ineq-merit.bib"
-csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
-link-citations: true
-linestretch: 1.15       
-mainfont: Times New Roman
-fontsize: 12pt          
-geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
-papersize: letter
-colorlinks: true
-fig-height: 4
-fig-width: 7.5
-header-includes:
-  - \usepackage{threeparttable}
 ---
-
-This document was last modified at {{< modified >}}
-and it was last rendered at {{< now >}}.
 

### 2024-08-05 paper.tex
Actualización paper anonimizado en .tex y orden de repo

--- 2024-05-08 paper.tex
@@
 \documentclass[
-  letterpaper,
-  DIV=11,
-  numbers=noendperiod]{scrartcl}
+    behavsci,
+    article,
+    submit,
+moreauthors
+]{mdpi}
 
@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 
@@
 \makeatother
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
-  \fi
-  % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
-\KOMAoption{captions}{tableheading}
+
+% TODO: Add custom LaTeX header directives here
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 \makeatletter
@@
 \fi
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
 \usepackage{bookmark}
@@
 \hypersetup{
-  pdfauthor={Equipo EDUMER},
+  pdftitle={The socialization of meritocracy and market justice preferences at school},
+  pdfauthor={Juan Carlos Castillo; Mauricio Salgado; Kevin Carrasco; Andreas Laffert},
+  pdfkeywords={market
+justice, Meritocracy, socialization, family, schools, Chile},
+  pdfcreator={LaTeX via pandoc},
+  unicode=true,
+	bookmarksopen={true},
+	pdffitwindow=true, 
 	colorlinks=true, 
-  linkcolor={blue},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={Blue},
-  pdfcreator={LaTeX via pandoc}}
-
-\author{}
-\date{}
+	linkcolor=bluecite, 
+	citecolor=bluecite, 
+	urlcolor=bluecite, 
+	hyperfootnotes=true, 
+	pdfstartview={FitH},
+	pdfpagemode= UseNone}
+
+\firstpage{1} 
+\makeatletter 
+\setcounter{page}{\@firstpage} 
+\makeatother
+\pubvolume{1}
+\issuenum{1}
+\articlenumber{0}
+\pubyear{\the\year{}}
+\copyrightyear{\the\year{}}
+%\externaleditor{Academic Editor: Firstname Lastname}
+\datereceived{ } 
+\daterevised{ } % Comment out if no revised date
+\dateaccepted{ } 
+\datepublished{ } 
+%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
+%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
+\hreflink{https://doi.org/} % If needed use \linebreak
+%\doinum{}
+%\pdfoutput=1 % Uncommented for upload to arXiv.org
+%\CorrStatement{yes}  % For updates
+
+\Title{ The socialization of meritocracy and market justice preferences
+at school }
+
+% MDPI internal command: Title for citation in the left column
+\TitleCitation{ The socialization of meritocracy and market justice
+preferences at school }
+
+% Author Orchid ID: enter ID or remove command
+%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
+%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorA}{0000-0003-1265-7854} 
+\newcommand{\orcidauthorB}{0000-0002-8108-6154} 
+\newcommand{\orcidauthorC}{0000-0002-9348-8361} 
+\newcommand{\orcidauthorD}{0000-0002-9008-2454}
+
+% Authors, for the paper (add full first names)
+%\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{2,\ddagger}$ and Firstname Lastname $^{2,}$*}
+
+\Author{Juan Carlos Castillo$^{1}$\orcidA{}, 
+Mauricio Salgado$^{2}$\orcidB{}, 
+Kevin Carrasco$^{3}$\orcidC{}~and~Andreas Laffert$^{4}$\orcidD{}
+}
+%\longauthorlist{yes}
+
+% MDPI internal command: Authors, for metadata in PDF
+\AuthorNames{
+  Juan Carlos Castillo,
+  Mauricio Salgado,
+  Kevin Carrasco,
+  Andreas Laffert}
+
+% MDPI internal command: Authors, for citation in the left column
+%\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
+%\usepackage{xstring}
+\newcommand\getfirst[1]{\getfirstaux#1\relax} \def\getfirstaux#1#2\relax{#1}
+\newcommand{\authorformatted}{Castillo, \getfirst{Juan Carlos}.; Salgado, \getfirst{Mauricio}.; Carrasco, \getfirst{Kevin}.; Laffert, \getfirst{Andreas}.}
+\AuthorCitation{\authorformatted}
+
+% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+\address {%
+$^{1}$ \quad Department of Sociology, Universidad de Chile\\
+$^{2}$ \quad CEP, Centro de Estudios Públicos\\
+$^{3}$ \quad COES, Center for Social Conflict and Cohesion Studies\\
+$^{4}$ \quad Institute of Sociology, Universidad Católica de Chile\\
+}
+
+%\address{
+%$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
+%$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
+
+% Contact information of the corresponding author
+%% see: https://quarto.org/docs/journals/authors.html
+\corres{Correspondence: juancastillov@uchile.cl}
+
+% Current address and/or shared authorship
+%\firstnote{Current address: Affiliation.}  % Current address should not be the same as any items in the Affiliation section.
+%\secondnote{These authors contributed equally to this work.}
+% The commands \thirdnote{} till \eighthnote{} are available for further notes
+
+%\simplesumm{} % Simple summary
+
+%\conference{} % An extended version of a conference paper
+
+% Abstract (Do not insert blank lines, i.e. \\) 
+\abstract{ Previous research has shown that schools justify student
+performance differences based on meritocratic ideals. However, the link
+between school meritocracy and students' justification of inequalities
+has received little attention. This paper examines how school and family
+socialization influence students' beliefs about meritocracy and
+market-based justifications for access to health, pensions, and
+education in Chile. Using data from the First Study of Civic Education
+in Chile, we employ a multilevel modeling strategy to analyze 5,047
+students and parents nested within 231 schools. Our findings reveal that
+a significant proportion of Chilean students agree with market justice
+principles, at a higher rate than adults. Most students endorse
+meritocratic views, particularly the reward of effort. Multilevel
+estimations show strong associations between meritocratic beliefs and
+market justice preferences: students who believe in the reward of effort
+and talent are more likely to justify inequalities based on financial
+capacity. At the school level, market justice preferences are higher in
+high-status schools but lower in schools with better achievement.
+Additionally, the conditional influence of meritocracy beliefs weakens
+in schools with higher socioeconomic status and performance. These
+results suggest that meritocratic beliefs and market justice preferences
+begin early and are influenced by the school environment. }
+
+% Keywords
+\keyword{market
+justice; Meritocracy; socialization; family; schools; Chile}
 
@@
 
-\section{The socialization of meritocracy and market justice preferences
-at
-school}\label{the-socialization-of-meritocracy-and-market-justice-preferences-at-school}
-
-\subsection{Introduction}\label{introduction}
+\section{Introduction}\label{introduction}
 
 Since its origins, educational institutions have been related to the
-idea of social mobility and access to better opportunities. Therefore,
-the consistent evidence of the high level of social reproduction at the
-school level represents a threat to the promise of education and a
-meritocratic system (\citeproc{ref-bourdieu_reproduction_1990}{Bourdieu
-and Passeron 1990}). A large part of the research in this field at an
-international level has addressed the extent to which the social origin
-of students affects their academic results and their life opportunities
-(\citeproc{ref-vonhippel_test_2019}{Von Hippel and Hamrock 2019}),
-confirming that schools have severe difficulties in closing the gaps of
-origin. Besides this socioeconomic perspective on school opportunities,
-recent research has addressed to what extent inequalities in the school
-context are also influencing students' perceptions, beliefs, and
-attitudes: Are social inequalities perceived at the school context? Are
-they rejected by the students, particularly those who are worst-off in
-socioeconomic terms? Or, Is there evidence at the school level that
-social inequalities are tolerated and even justified?
-(\citeproc{ref-batruch_belief_2022}{Batruch et al. 2022};
-\citeproc{ref-wiederkehr_belief_2015}{Wiederkehr et al. 2015}).
-
-In the present paper, we deal with the justification of social
-inequalities by eighth-grade students in Chile, a country characterized
-by a highly stratified educational system. In particular, we focus on
-market justice preferences (\citeproc{ref-lane_market_1986}{Lane 1986}),
-which refer to the preferences for distributing public goods (as health
-and education) based on criteria such as competition and payment
-capacity. Although from a rational point of view it could be expected an
-opposition to market justice by the underprivileged majority, we argue
-that in a social environment characterized by a promotion of
-meritocratic ideals - as the school system - would lead to the opposite:
-a larger market justice preferences.
+idea of social mobility and access to better opportunities. Despite
+this, the consistent evidence of the high level of social reproduction
+at the school level represents a threat to the promise of education and
+a meritocratic system \citep{bourdieu_reproduction_1990}. A large part
+of the research in this field at an international level has addressed
+the extent to which the social origin of students affects their academic
+results and their life opportunities \citep{vonhippel_test_2019},
+confirming that schools have severe difficulties in closing the
+socio-economic and cultural gaps of origin. Besides this socioeconomic
+perspective on school opportunities, recent research has addressed to
+what extent inequalities in the school context are also influencing
+students' perceptions, beliefs, and attitudes: Are social inequalities
+even perceived at the school context? Are they rejected by the students,
+particularly those who are worst-off in socioeconomic terms? Or, Is
+there evidence at the school level that social inequalities are
+tolerated and even justified?
+\citep{batruch_belief_2022, wiederkehr_belief_2015}.
 
@@
 based on their abilities, knowledge, and achievements
-(\citeproc{ref-young_rise_1958}{Young 1958}). It is often seen as a way
-to create equal opportunities and fairness, as individuals can rise to
-positions of power and influence based on their own merit rather than
-their background or connections. However, some argue that meritocracy
-can actually lead to tolerating or even justifying social inequalities,
-as it can create a hierarchy where those who already have resources and
-advantages are more likely to succeed. In this regard, a great deal of
-academic research about meritocracy delves into the assessment of to
-what extent rewards and privileges in society are related to merit,
-emphasizing the so-called unfulfillable promise of meritocracy
-(\citeproc{ref-mijs_stratified_2016}{Mijs 2016}). Complementing this
-agenda, a second and emerging research area deals with subjective
-aspects of meritocracy, such as perceptions and beliefs.
-
-{[}definición de market justice, welfare marketization, (Lindh){]}
-
-{[}contexto Chileno market justice, dar cifras/porcentajes de
-privatización de salud, educación y pensiones; asociar al tema de la
-calidad (lo público es peor){]}
-
-The perception of meritocracy refers to how individuals view and
-understand the concept of meritocracy in their own society
-(\citeproc{ref-duru-bellat_who_2012}{Duru-Bellat and Tenret 2012};
-\citeproc{ref-castillo_meritocracia_2019}{Castillo et al. 2019}). This
-perception can vary greatly depending on individual experiences, social,
+\citep{young_rise_1958}. It is often seen as a way to create equal
+opportunities and fairness, as individuals can rise to positions of
+power and influence based on their own merit rather than their
+background or connections. However, some argue that meritocracy can
+actually lead to tolerating or even justifying social inequalities, as
+it can create a hierarchy where those who already have resources and
+advantages are more likely to succeed
+\citep{sandel_tyranny_2020, mcnamee_meritocracy_2004}. In this regard, a
+great deal of academic research about meritocracy delves into the
+assessment of to what extent rewards and privileges in society are
+related to merit, emphasizing the so-called unfulfillable promise of
+meritocracy \citep{mijs_stratified_2016}.
+
+In the present paper, we address the role of the perception of
+meritocracy on the justification of social inequalities by eighth-grade
+students in Chile, a country characterized by a highly stratified
+educational system. In particular, we focus on justifying inequalities
+in health, education and pensions, which are traditional social policy
+areas where access to better services could be justified based on
+payment capacity, referred to as market justice
+\citep{lane_market_1986, lindh_bringing_2023}. Most of the research in
+this realm has only considered adults, leaving aside the study of how
+beliefs in this field develop at student age as well as the impact of
+the school context and the family as the main socialization agencies
+\citep{batruch_belief_2022}. Regarding schools, the way in which they
+deal with unequal conditions of origin has been linked to the
+\emph{hidden curriculum} \citep{chafel_schooling_1997}, whereby students
+learn about distributive norms in society and mechanisms of
+justification of social differences.
+
+Based on recent studies that relate school meritocracy to the
+justification of economic inequalities in the adult population
+\citep{batruch_belief_2022, wiederkehr_belief_2015}, the central
+hypothesis guiding this research is that school-age students with a
+higher perception of meritocracy - both at school and at the societal
+level - will show a larger market justice preferences, as individual
+achievement would be seen as appropriately rewarded and social
+mechanisms for correcting inequalities as less necessary
+\citep{batruch_belief_2022}. We focus on the student-age population as
+we point out that it is possible to track down the origin of
+meritocratic beliefs (and their consequences) to early socialization
+processes. To this regard, we take into account the family and the
+school as two main socialization agencies that play a significant role
+in the socialization of cultural beliefs by transmitting cultural norms,
+values, and expectations to young people.
+
+The Chilean case is particularly intriguing for studying market justice
+preferences. This country is characterized by acute and persistent
+economic inequality, which stands out in Latin America and among OECD
+countries \citep{flores_top_2020}. In Chile, the poorest 50\% captures
+only 10\% of the total income and has negative wealth, while the richest
+1\% receives almost 27\% of the income and holds 49.6\% of the wealth
+\citep{chancel_world_2022}. Much of this inequality has been attributed
+to the deep neoliberal reforms that institutionalized privatization and
+commodification of various economic sectors. Such reforms where
+introduced during the dictatorship (1973-1989) and expanded in democracy
+through concessions, demand credits, and specific regulatory frameworks
+\citep{ffrench-davis_reformas_2018}. This shift in economic policy has
+allowed the unprecedented emergence of markets in areas related to
+social reproduction, such as health, pensions, and education, with
+provision and access managed by private entities and segmented by
+individual payment capacity, heavily reliant on State subsidies
+\citep{boccardo_30_2020}. In health, although the majority of the
+population uses the public insurance system (78.9\%), 15.3\% are served
+by private insurers \citep{observatoriosocial_estadisticas_2024}. The
+pension system is based on individual capitalization, with mandatory
+contributions managed by private administrators investing in the
+financial market, currently involving 11 million contributors
+\citep{superintendenciadepensiones_estadisticas_2024}. In education,
+30.6\% of school enrollment is in public schools, 54.0\% in
+state-subsidized (voucher) private schools, and 9.3\% in fully private
+schools, generally attracting higher-income groups
+\citep{ministeriodeeducacion_resumen_2023}.
+
+\subsection{Justification of inequality and market
+justice}\label{justification-of-inequality-and-market-justice}
+
+The justification of social inequality based on market-type criteria has
+been conceptualized as the individuals' adherence to the deservingness
+of social goods and services (such as health, education, and pensions)
+based on prices and individuals' ability to pay
+\citep{boltanski_new_2005, lane_market_1986, streeck_citizens_2012}.
+Research on social stratification beliefs, which explore individual
+perceptions of who deserves what and why \citep{kluegel_beliefs_1987},
+highlights that people's explanations and justifications of social
+inequality are closely tied to their judgments of deservingness. The
+influence of ideologies \citep{wegener_dominant_1995} and cultural
+schemas \citep{homan_being_2017} is pivotal in shaping these
+explanations by offering symbolic representations that frame societal
+structures and expectations. While significant attention has been paid
+to wage inequality, income distribution, and payment differentials in
+the literature
+\citep{castillo_legitimacy_2011, evans_family_2010, jasso_how_1999, shariff_income_2016},
+there has been less examination of public beliefs about which life
+domains should be governed by market relations
+\citep{lindh_bringing_2023} and even less about children's acceptance or
+rejection of these market principles. This oversight is notable given
+the extensive encroachment of market logic into public goods, welfare
+policy, and social services over the past five decades
+\citep{centeno_arc_2012, harvey_breve_2015}, affecting areas such as
+pensions, health services, and education.
+
+There are substantial differences in funding and delivery methods in the
+management of social services across nations
+\citep{jensen_worlds_2008, stoy_worlds_2014}. Nordic countries, for
+example, predominantly employ public agencies to produce and provide
+social services, funding these through collective taxation and offering
+them in kind to the majority of citizens. This system prioritizes social
+justice, placing it above market mechanisms in accessing services. In
+contrast, other countries rely more heavily on for-profit entities and
+private funding, where service distribution depends mainly on individual
+financial capacity to pay user fees, highlighting the influence of
+market justice in service allocation. The trend toward marketization of
+welfare services has been growing since the 1980s
+\citep{salamon_marketization_1993}, and this shift is increasingly
+evident even in countries where market solutions have traditionally had
+a minor role in social policy \citep{sivesind_changing_2017}. The
+expansion of marketization has been related to a larger justification of
+market mechanisms, whereby societies with larger private spending on
+services show larger market justice preferences
+\citep{lindh_public_2015}.
+
+Robert E. Lane proposed the underpinnings of the concept of market
+justice, which he differentiated from political justice. For him, ``it
+is the genius of the market to stimulate wants without at the same time
+stimulating a sense of deserving more than one gets''
+\citeyearpar[p.~384]{lane_market_1986}. Contrary to the evidence that
+unequal distribution produces feelings of dissatisfaction, anger, and
+resentment that might motivate forms of collective action
+\citep{greitemeyer_subjective_2016, mishra_subjective_2015, smith_relative_2012, power_deprivationprotest_2018},
+Lane pointed out that in market settings, social comparisons are more
+likely to motivate increased effort rather than feelings of acute
+injustice because individuals attribute outcomes to their actions. In
+this sense, unequal levels of well-being would be, to some extent, a
+function of their talents and efforts, instead of being based on
+distributive principles that characterize welfare states, such as need
+and equality (see \citep{wilson_role_2003}).
+
+Despite high-income inequality and limited social mobility in Chile, and
+in Latin America in general, there is a prevalent belief that
+individuals are solely responsible for their economic outcomes, a view
+that varies across the region
+\citep{bucca_merit_2016, chong_mystery_2008, torche_intergenerational_2014, salgado_inequality_2023}.
+The reliance on private welfare providers and widespread user fees
+\citep{molyneux_neoliberal_2008} adds complexity to this context, as
+reflected in surveys conducted by the Center for Public Studies (CEP).
+According to this data, 35.9\% prefer private health insurance, and 63\%
+would prefer private education
+\citep{centrodeestudiospublicos_estudio_2024}. Yet, research on
+children's justification in this area remains limited, highlighting a
+significant gap in understanding how younger generations view
+market-based access to welfare and whether these views are associated
+with their meritocratic beliefs.
+
+\subsection{Meritocratic perceptions and market
+justice}\label{meritocratic-perceptions-and-market-justice}
+
+The original definition of merit is a combination of effort and talent
+\citep{young_rise_1958}, and a meritocracy is a distributive system
+where merit is the main criterion for allocating valuable goods and
+rewards. From a sociological perspective, meritocracy has been used in
+research on social mobility to characterize societies with low mobility
+that threaten the meritocratic ideal \citep{goldthorpe_myth_2003}. More
+recently, sociology and social psychology research has attended to
+subjective aspects related to the support for meritocratic principles in
+different societies, such as beliefs in meritocracy
+\citep{castillo_multidimensional_2023, mijs_unfulfillable_2016, mijs_paradox_2019}.
+
+Meritocratic beliefs can cover two types of subjective processes:
+preferences and perceptions \citep{castillo_multidimensional_2023}.
+While meritocratic preferences refer to a justification of distribution
+based on merit criteria (effort and talent), the perception of
+meritocracy refers to how individuals view and understand the concept of
+meritocracy in their society
+\citep{duru-bellat_who_2012, castillo_meritocracia_2019}. The perception
+can vary greatly depending on individual experiences, as well as social,
 economic, and cultural background. Some people may see meritocracy as a
@@
 and even reinforce inequality
-(\citeproc{ref-lampert_meritocratic_2013}{Lampert 2013};
-\citeproc{ref-mijs_paradox_2021}{Mijs 2021}). Based in this last
-perspective, we argue that individuals with a higher perception of
-meritocracy will show a larger justification of social inequalities, as
-individual achievement would be seen as rewarded and social policies as
-less necessary (\citeproc{ref-batruch_belief_2022}{Batruch et al.
-2022}).
-
-Most of the research that has related meritocratic beliefs to inequality
-justification so far has only considered adults, leaving aside the study
-of how beliefs in this field develop at student age as well as the
-impact of the school context and the family as the main socialization
-agencies. Regarding schools, the way in which they deal with unequal
-conditions of origin has been linked to the \emph{hidden curriculum}
-(\citeproc{ref-chafel_schooling_1997}{Chafel 1997}), whereby students
-learn about distributive norms in society and mechanisms of
-justification of social differences. Based on recent studies that relate
-school meritocracy to the justification of economic inequalities in the
-adult population (\citeproc{ref-batruch_belief_2022}{Batruch et al.
-2022}; \citeproc{ref-wiederkehr_belief_2015}{Wiederkehr et al. 2015}),
-the central hypothesis guiding this research is that school-age students
-with a higher perception of meritocracy - both at school and at the
-societal level - will show a larger justification of social
-inequalities, as individual achievement would be seen as appropriately
-rewarded and social mechanisms for correcting inequalities as less
-necessary (\citeproc{ref-batruch_belief_2022}{Batruch et al. 2022}).
-
-\textasciitilde{}
-
-{[}introducir distinción sobre percepción de meritocracia a nivel
-escolar y a nivel de la sociedad{]}
-
-{[}resaltar 3 focos de innovación: market justice preferences at school
-level, relación con meritocracia, y esquema de meritocracia en la
-escuela y en la sociedad{]}
-
-The present paper deals with the association between the perception of
-meritocracy and the justification of social inequalities, with two main
-focuses. Firstly, it assesses the justification of social inequalities
-in social policy domains, such as health, pensions, and education. We
-argue that individuals who perceive more meritocracy would be more
-willing to justify better services in these domains for those with
-higher incomes. Secondly, we focus on the student-age population as we
-point out that it is possible to track down the origin of meritocratic
-beliefs (and their consequences) to early socialization processes. To
-this regard, we take into account the family and the school as two main
-socialization agencies that play a significant role in the socialization
-of cultural beliefs by transmitting cultural norms, values, and
-expectations to young people.
-
-\subsection{Justification of
-inequality}\label{justification-of-inequality}
-
-Research on social stratification beliefs, which explore individual
-perceptions of who deserves what and why
-(\citeproc{ref-kluegel_beliefs_1987}{Kluegel and Smith 1987}),
-highlights that people's explanations and justifications of social
-inequality are closely tied to their judgments of deservingness. The
-influence of ideologies (\citeproc{ref-wegener_dominant_1995}{Wegener
-and Liebig 1995}) and cultural schemas
-(\citeproc{ref-homan_being_2017}{Homan, Valentino, and Weed 2017}) is
-pivotal in shaping these explanations by offering symbolic
-representations that frame societal structures and expectations. While
-significant attention has been paid to wage inequality, income
-distribution, and payment differentials in the literature {[}Castillo
-(\citeproc{ref-castillo_legitimacy_2011}{2011}); Evans et al., 2010;
-Jasso (\citeproc{ref-jasso_how_1999}{1999}) ; Shariff, Wiwad, and Aknin
-(\citeproc{ref-shariff_income_2016}{2016}){]}, there has been less
-examination of public beliefs about which life domains should be
-governed by market relations and even less about children's acceptance
-or rejection of these market principles. This oversight is notable given
-the extensive encroachment of market logic into public goods, welfare
-policy, and social services over the past five decades
-(\citeproc{ref-centeno_arc_2012}{Centeno and Cohen 2012};
-\citeproc{ref-harvey_breve_2015}{Harvey 2015}), affecting areas such as
-pensions, health services, and education. The justification of social
-inequality based on marketype criteria has been conceptualized as the
-individuals' adherence to one specific justice evaluation, the market
-justice, that is, affording legitimacy to the allocation of goods and
-services based on prices and individuals' ability to pay
-(\citeproc{ref-boltanski_new_2005}{Boltanski and Chiapello 2005};
-\citeproc{ref-lane_market_1986}{Lane 1986};
-\citeproc{ref-streeck_citizens_2012}{Streeck 2012}).
-
-Robert E. Lane proposed the underpinnings of market justice, which he
-differentiated from political justice. For him, ``it is the genius of
-the market to stimulate wants without at the same time stimulating a
-sense of deserving more than one gets'' {[}-Lane
-(\citeproc{ref-lane_market_1986}{1986}); p.~384{]}. Following the theory
-of relative deprivation --a social phenomenon arising when individuals
-cannot afford what most others in their environment can (Merton 1950)--
-Lane notes that, in market settings, social comparisons are more likely
-to motivate increased effort rather than feelings of acute injustice
-because individuals attribute outcomes to their actions. Although
-empirical research has shown that, contrary to Lane's observation,
-relative deprivation, even in market settings, produces feelings of
-dissatisfaction, anger, and resentment that might motivate forms of
-collective action such as protests and revolt
-(\citeproc{ref-greitemeyer_subjective_2016}{Greitemeyer and Sagioglou
-2016}; \citeproc{ref-mishra_subjective_2015}{Mishra and Carleton 2015};
-\citeproc{ref-smith_relative_2012}{Smith et al. 2012};
-\citeproc{ref-power_deprivationprotest_2018}{Séamus A. Power 2018};
-\citeproc{ref-power_relative_2020}{Séamus A. Power, Madsen, and Morton
-2020}), the conceptualization of market justice has been since then
-closely coupled to the merit principle for allocating outcomes: the
-claim that the unequal levels of well-being individuals enjoy ought to
-be, to some extent, a function of their talents and efforts, regardless
-of their needs or membership, the two latter being the realm of
-political justice and its closely coupled principles of need
---allocating outcomes to those who require them most-- and equality
---allocating the same outcome to everyone--, which have been at the
-center of welfarism (see Wilson 2003). The underlying notion of market
-justice also resonates in its application to welfare regimes, providing
-a framework for understanding the varied global approaches to managing
-social services.
-
-The management of social services manifests in varying approaches across
-nations, with substantial differences in funding and delivery methods
-(\citeproc{ref-jensen_worlds_2008}{Jensen 2008};
-\citeproc{ref-stoy_worlds_2014}{Stoy 2014}). Nordic countries, for
-example, predominantly employ public agencies to produce and provide
-social services, funding these through collective taxation and offering
-them in kind to the majority of citizens. This system prioritizes
-political justice, placing it above market mechanisms in accessing
-services. In contrast, other countries rely more heavily on for-profit
-entities and private funding, where service distribution depends mainly
-on individual financial capacity of paying user fees, highlighting the
-influence of market justice in service allocation. The trend toward
-marketization of welfare services has been growing since the 1980s
-(\citeproc{ref-salamon_marketization_1993}{Salamon 1993}), and this
-shift is increasingly evident even in countries where market solutions
-have traditionally had a minor role in social policy
-(\citeproc{ref-sivesind_changing_2017}{Sivesind 2017}).
-
-The question arises whether adults and children justify unequal access
-to welfare services based on market justice principles. Influenced by
-theories of policy feedback, which suggest that social welfare policies
-can reinforce (positive feedback) or undermine (negative feedback)
-previous policy trajectories
-(\citeproc{ref-fernandez_positive_2013}{Fernandez and Jaime-Castillo
-2013}; \citeproc{ref-pierson_increasing_2000}{Pierson 2000};
-\citeproc{ref-weaver_paths_2010}{Weaver 2010}), citizens' beliefs about
-market justice are likely also shaped by the institutional and social
-contexts they encounter. Indeed, the justification of inequality in
-access to essential services like education and health, based on one's
-ability to pay, shows significant variation across countries, as
-demonstrated by international surveys, although they do not usually
-include children in their samples. For instance, the 2019 International
-Social Survey Program (ISSP) provides insights into how adults perceive
-inequality in accessing welfare services. Figure~\ref{fig-issp}
-illustrates the variation in agreement levels regarding whether it is
-just for individuals with higher incomes to purchase better education or
-healthcare. Notably, in 18 of the 29 countries analyzed, there is a
-greater justification for market inequality in healthcare access than in
-education. Nevertheless, the general sentiment typically ranges from
-`somewhat unjust' to 'neither just nor unjust, with mixed feelings.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-issp-1.pdf}
-
-}
-
-\caption{\label{fig-issp}Average of market justice preferences by
-country}
-
-\end{figure}%
-
-While contemporary policy developments have increasingly embraced the
-marketization of social welfare in areas such as education, pensions,
-and health services, questions about public acceptance of these changes
-persist. Lindh (\citeproc{ref-lindh_public_2015}{2015}) analysis of ISSP
-2009 data from 17 OECD countries reveals a general lack of support for
-market-based distribution of social services, suggesting widespread
-disapproval of market stratification of essential services. This finding
-is corroborated by Soler-Martínez, García-Sánchez, and Willis
-(\citeproc{ref-soler-martinez_concerns_2023}{2023}) research from
-Latinobarómetro 2020 across 18 Latin American countries, where concerns
-about health and education access predominated over income inequality.
-These results indicate that reforms toward welfare marketization are
-typically driven by elite political decisions rather than grassroots
-demand.
-
-Despite high-income inequality and limited social mobility in Latin
-America, there is a prevalent belief that individuals are solely
-responsible for their economic outcomes, a view that varies across the
-region (\citeproc{ref-bucca_merit_2016}{Bucca 2016};
-\citeproc{ref-chong_mystery_2008}{Chong and Ñopo 2008};
-\citeproc{ref-torche_intergenerational_2014}{Torche 2014};
-\citeproc{ref-salgado_inequality_2023}{Salgado and Castillo 2023}). The
-reliance on private welfare providers and widespread user fees
-(\citeproc{ref-molyneux_neoliberal_2008}{Molyneux 2008}) adds complexity
-to this context. Yet, research on children's justification of
-market-based inequalities in accessing welfare services remains limited,
-especially in Latin America, highlighting a significant gap in
-understanding how younger generations view market-based access to
-welfare and whether these views are associated with their meritocratic
-beliefs.
-
-\subsection{Meritocracy}\label{meritocracy}
-
-The concept of meritocracy frequently appears nowadays when analyzing
-cultural determinants of social inequalities. In general, it is
-mentioned as a value associated with justice, as it would link efforts
-and talents with rewards in an equitable manner. This normative sense is
-quite far from its original formulation by Young
-(\citeproc{ref-young_rise_1958}{1958}) in the satirical novel ``The Rise
-of Meritocracy'', where it ironically represented a mechanism for
-reproducing the inequalities of origin. The meritocratic ideal had
-remained relatively unchallenged until a series of recent publications
-turned into its potential consequences for maintaining social
-inequality. Perhaps one of the most recent sources in this line is
-Michael Sandel's ``The Tyranny of Merit'', where he strongly questions
-the implications of carrying out the principle of merit in societies
-that do not guarantee equal opportunities and that generate a feeling of
-scarce recognition and appreciation of those who receive lesser rewards:
-``In society's eyes, and perhaps also their own, their work no longer
-signified a valued contribution to the common good.''
-(\citeproc{ref-sandel_tyranny_2020}{Sandel 2020}, pp.).
-
-Empirical research on meritocracy has increased along with the
-philosophical-normative discussion on meritocracy in recent years.
-Particularly from a sociological perspective, meritocracy has been used
-in research on social mobility to characterize societies with low
-mobility that threaten the meritocratic ideal
-(\citeproc{ref-goldthorpe_myth_2003}{Goldthorpe 2003}). More recently,
-sociology and social psychology research has attended to the subjective
-aspects vis-a-vis beliefs in meritocracy. The label of beliefs in this
-realm covers a series of areas, such as attitudes, perceptions, and
-preferences (Castillo et al), whereby most of the link this subjective
-dimensions to individual socio-structural factors and context-level
-determinants. For instance, some studies have analyzed how those with
-greater privileges believe more in meritocracy (Reynolds \& Chan 2014),
-how greater economic inequality increases meritocratic beliefs
-(\citeproc{ref-mijs_paradox_2021}{Mijs 2021}), and how larger inequality
-affects meritocratic beliefs
-(\citeproc{ref-morris_representing_2022}{Morris et al. 2022}). Based on
-these findings, a research agenda has been reinforced on the
-legitimizing role of meritocracy, in line with previous studies using
-the concept of a just world (Lerner, Dalbert) and the theory of system
-justification (Jost \& Major).
-
-How do meritocratic beliefs legitimize inequalities? Empirical studies
-have used experiments and surveys to address this question. For
-instance, the evidence suggests that just world beliefs correlate
-negatively with support for redistributive compensation systems
-(\citeproc{ref-frank_performance_2015}{Frank, Wertenbroch, and Maddux
-2015}). Conversely, individuals tend to support redistribution when they
-believe that the disadvantaged lack the opportunities to succeed
-(\citeproc{ref-evans_strong_2018}{Evans and Kelley 2018}). Almås,
-Cappelen, and Tungodden (\citeproc{ref-almas_cutthroat_2020}{2020})
-found that in a relatively unequal society (the United States), the
-highly educated accept inequality significantly more than the less
-educated because they perceive inequality as justifiable owing to
-differences in productivity (i.e., merit), whereas in a relatively equal
-society (Norway), the less educated accept inequality more, but not
-significantly more than the highly educated because meritocratic values
-are less prevalent. Barr and Miller
-(\citeproc{ref-barr_effect_2020}{2020}) also addressed this triple
-interaction between the level of inequality in a society, the individual
-level of education, and the perceived origin of the disparity (either by
-luck or effort) to determine the extent to which inequality is accepted.
-They found that the interaction's mechanism varies depending on the
-compared societies. Finally, García-Sánchez et al.
-(\citeproc{ref-garcia-sanchez_attitudes_2020}{2020}), using data for 41
-countries from the International Social Survey Programme (ISSP), found
-that the perceived size of the income gap correlated positively with
-support for progressive taxation. Still, this association was weaker
-among those who endorsed meritocratic and equal opportunity beliefs. In
-the same line, experimental research by Durante and Putterman
-(\citeproc{ref-durante_preferences_2009}{2009}) subjects support less
-redistribution when the initial distribution is determined according to
-task performance.
-
-Research about meritocratic beliefs at school age is rather scarce,
-leaving a wide research gap as schools are one of the primary
-socialization institutions where achievement based on merit explains
-success (\citeproc{ref-erivwo_meritocracy_2021}{Erivwo et al. 2021}).
-Nevertheless, it is possible to find some initial works focused on the
-area of distributive justice at school that are closely related to
-meritocratic beliefs, as the ones by Resh and Sabbagh (quotes). Using
-justice in grade obtention as a measure of distributive justice and
-meritocracy (Sabbagh et al 2006), they find for instance that a larger
-sense of distributive justice about grades is associated to higher
-socio-economic status (\citeproc{ref-resh_sense_2010}{Resh 2010}), have
-a positive effect on liberal democratic orientation and on trust in
-people and in formal institutions (\citeproc{ref-resh_sense_2014}{Resh
-and Sabbagh 2014}), and tend to refrain from violence and to engage to a
-greater extent in extra-curricular school activity and community
-volunteering (\citeproc{ref-resh_sense_2017}{Resh and Sabbagh 2017}).
-
-\subsection{Children's judgments of inequality, schools, and family
-background}\label{childrens-judgments-of-inequality-schools-and-family-background}
-
-Research indicates various socialization practices at families and
-schools during childhood and adolescence that impact dispositional and
-behavioral tendencies concerning justifications of inequality in adult
-life. The differences in economic understanding across age groups are
-consistent with research on cognitive development
-(\citeproc{ref-choudhury_social_2006}{Choudhury, Blakemore, and Charman
-2006}). Adolescents with mature socio-cognitive abilities tend to
-express stronger preferences for fairness than infants and children
-(\citeproc{ref-wynn_not_2018}{Wynn et al. 2018}). As children grow
-older, they become more likely to behave fairly, with their
-early-emerging strict egalitarianism being replaced by an increasing
-endorsement of fairness principles and engagement in collaborative
-activities (\citeproc{ref-huppert_development_2019}{Huppert et al.
-2019}; \citeproc{ref-mcauliffe_developmental_2017}{McAuliffe et al.
-2017}). In these activities, their fairness views consider individual
-contributions, merits, and circumstances
-(\citeproc{ref-almas_fairness_2010}{Almås et al. 2010};
-\citeproc{ref-huppert_development_2019}{Huppert et al. 2019};
-\citeproc{ref-sigelman_development_1991}{Sigelman and Waitzman 1991}).
-Engelmann and Tomasello (\citeproc{ref-engelmann_children_2019}{2019})
-claim that children's sense of fairness emerges at three years old, and
-we can observe it in collaborative activities, where they accept
-inequality if the procedure gives everyone an equal chance. Therefore,
-children at this age respond to unequal distributions based on
-interpersonal concerns, as they already demand equal respect. In any
-case, between 3 and 8 years of age, inequitable and anti-meritorious
-allocations are evaluated more negatively, but equitable and meritorious
-allocations are not evaluated more positively
-(\citeproc{ref-elenbaas_unfairness_2019}{Elenbaas 2019}).
-
-Some research shows that the social environment in which children
-develop, such as family and school, is associated with their prosocial
-behaviors by playing an essential role in the transmission of equity
-norms (\citeproc{ref-schunk_fairness_2023}{Schunk and Zipperle 2023};
-\citeproc{ref-kosse_prosociality_2020}{Kosse and Tincani 2020}). In
-fact, schools contribute to institutionalizing and reproducing
-inequality by promoting values, norms, practices, and languages familiar
-to higher-class families because the dominant group's culture shapes
-educational institutions
-(\citeproc{ref-bourdieu_reproduction_1990}{Bourdieu and Passeron 1990}).
-Middle- and upper-class students are better equipped to face academic
-challenges and are more familiar with academic expectations
-(\citeproc{ref-mikus_children_2020}{Mikus, Tieben, and Schober 2020}).
-Such familiarity represents cultural capital in educational contexts
-because higher-status students come to school ready to meet these
-expectations and reap the benefits (\citeproc{ref-jack_no_2016}{Jack
-2016}; \citeproc{ref-khan_privilege_2011}{Khan 2011}). Conversely,
-lower-status children lacking cultural capital must catch up while
-experiencing inequitable comparisons
-(\citeproc{ref-goudeau_hidden_2017}{Goudeau and Croizet 2017}).
-Additionally, academic achievement is treated as the outcome of
-dispositional factors (e.g., pupils' efforts and talents or lack of
-them) rather than the result of differential access to critical
-resources. Due to the meritocratic frame schools encourage, both low-
-and high-status individuals believe that students' success or failure is
-not due to their family background but rather to differences in efforts
-and talents (\citeproc{ref-darnon_where_2018}{Darnon et al. 2018}). In
-this sense, we believe that the perception of meritocracy can influence
-students' judgments about market justice preferences. Furthermore, we
-believe there is a difference between students' perceptions of
-meritocracy based on their own experience in school and what they
-perceive in society at large. Consistently, our first two hypotheses
-are:
-
-H1a: Students who perceive that there is more meritocracy at school will
-show larger market justice preferences
-
-H1b: Students who perceive that there is more meritocracy in society
-will show larger market justice preferences
-
-Family background and family socialization practices also contribute to
-children's and adolescent's market justice preferences. For example,
-Almås et al. (\citeproc{ref-almas_fairness_2017}{2017}) found that
-adolescents from low-socioeconomic-status families are likelier to have
-an egalitarian fairness view and consider an equal distribution as fair
-in a situation with unequal merits. The authors speculate that
-differences in socialization practices across status groups might bring
-about, to a great extent, the fairness views of children and adolescents
-because social status seems to interact with these evaluations (e.g.,
-\citeproc{ref-hvidberg_social_2023}{Hvidberg, Kreiner, and Stantcheva
-2023}).
-
-The classic work of Kohn showed that middle-class parents value the
-expression of internal states and emotions, such as self-control,
-curiosity, happiness, and consideration, while working-class parents
-promote deference, obedience, and conformity to authority
-(\citeproc{ref-kohn_social_1963}{Kohn 1963};
-\citeproc{ref-kohn_class_1969}{Kohn and Schooler 1969}). Although
-parents from all social backgrounds encourage individualism in their
-children, this shared norm translates into different forms in high and
-low social classes (1999). Acemoglu
-(\citeproc{ref-acemoglu_obedience_2021}{2021}) claimed that the values
-families impart to their children interact with social mobility. Because
-obedience is a valuable characteristic for employers, in low-wage and
-social mobility environments, low-income families impart values of
-obedience to their children to prevent disadvantaging them in labor
-markets. On the one hand, children from privileged families are
-socialized to adopt a clear conception of individualism that highlights
-their internal states, independence, and idiosyncrasies. In contrast,
-children from disadvantaged families are socialized to support a more
-balanced view of individualism that considers personal characteristics
-as resources to overcome collective impediments on the path to upward
-mobility (\citeproc{ref-iacoviello_collectivism_2019}{Iacoviello and
-Lorenzi-Cioldi 2019}). In this way, we believe that there are
-differences in the socialization of values according to socioeconomic
-differences that could influence market justice preferences and,
-therefore, we propose the following hypothesis:
-
-H2: Students from families of higher social status will show larger
+\citep{lampert_meritocratic_2013, mijs_paradox_2019}. Some studies have
+analyzed how those with greater privileges believe more in meritocracy
+\citep{reynolds_perceptions_2014}, how greater economic inequality
+increases meritocratic beliefs \citep{mijs_paradox_2019}, and how larger
+inequality decreases it \citep{morris_representing_2022}.
+
+A larger justification of meritocratic distribution has been related to
+less support for redistributive compensation systems
+\citep{frank_performance_2015}, as individual achievement would be seen
+as rewarded and social policies as less necessary.
+\citet{almas_cutthroat_2020} found that in the US the highly educated
+accept inequality significantly more than the less educated because they
+perceive inequality as justifiable owing to differences in productivity
+(i.e., merit). \citet{barr_effect_2020} found that in unequal societies,
+the highly educated accept inequality more than the low educated.
+Conversely, individuals tend to support redistribution when they believe
+that the disadvantaged lack the opportunities to succeed
+\citep{evans_strong_2018}.
+
+Schools contribute to institutionalizing and reproducing inequality by
+promoting values, norms, practices, and languages familiar to
+higher-class families because the dominant group's culture shapes
+educational institutions \citep{bourdieu_reproduction_1990}. Middle- and
+upper-class students are better equipped to face academic challenges and
+are more familiar with academic expectations
+\citep{mikus_children_2019}. Such familiarity represents cultural
+capital in educational contexts because higher-status students come to
+school ready to meet these expectations and reap the benefits
+\citep{jack_no_2016, khan_privilege_2011}. Conversely, lower-status
+children lacking cultural capital must catch up while experiencing
+inequitable comparisons \citep{goudeau_hidden_2017}. Additionally,
+academic achievement is treated as the outcome of dispositional factors
+(e.g., pupils' efforts and talents or lack of them) rather than the
+result of differential access to critical resources. Due to the
+meritocratic frame schools encourage, both low- and high-status
+individuals tend to believe that success or failure is not due to the
+family background but rather to differences in efforts and talents
+\citep{darnon_where_2018}. In this sense, we believe that the perception
+of meritocracy can influence students' judgments about market justice
+preferences, leading to our first hypothesis:
+
+\(H_{1a}\): Students who perceive that there is more meritocracy in
+society will show larger market justice preferences
+
+The perception of meritocracy has been mostly studied with general
+questions about reward allocation based on effort and talent (usually
+intelligence). Nevertheless, when looking at the school population it is
+possible to further consider the perception of meritocracy referred
+specifically to the school context. For instance,
+\citet{resh_sense_2010} find that perception of justice in grades has a
+positive effect on liberal democratic orientation, and trust in people
+and formal institutions \citep{resh_sense_2014}. Attending to this
+evidence, we differentiate in this study between meritocratic perception
+in society at large, and meritocratic perception at school, proposing
+the next hypothesis:
+
+\(H_{1b}\): Students who perceive that there is more meritocracy at
+school will show larger market justice preferences
+
+\subsection{The socialization of market justice: families and
+schools}\label{the-socialization-of-market-justice-families-and-schools}
+
+Attending to the socialization within the family, the classic work of
+Kohn showed that middle-class parents value the expression of internal
+states and emotions, such as self-control, curiosity, happiness, and
+consideration, while working-class parents promote deference, obedience,
+and conformity to authority \citep{kohn_social_1963, kohn_class_1969}.
+Although parents from all social backgrounds encourage individualism in
+their children, this shared norm translates into different forms in high
+and low social classes \citeyearpar{kohn_class_1969}.
+\citet{acemoglu_obedience_2021} claimed that the values families impart
+to their children interact with social mobility. On the one hand,
+children from privileged families are socialized to adopt a clear
+conception of individualism that highlights their internal states,
+independence, and idiosyncrasies. In contrast, children from
+disadvantaged families are socialized to support a more balanced view of
+individualism that considers personal characteristics as resources to
+overcome collective impediments on the path to upward mobility
+\citep{iacoviello_collectivism_2019}. In the same line,
+\citet{almas_fairness_2017} found that adolescents from
+low-socioeconomic-status families are likelier to have an egalitarian
+fairness view and consider an equal distribution as fair in a situation
+with unequal merits. Taking this into account, we believe that there
+will be differences in the socialization of values according to
+socioeconomic differences in families that could influence market
+justice preferences, which leads us to the following hypothesis:
+
+\(H_2\): Students from families of higher social status will show larger
 market justice preferences.
@@
 significantly influences children's and adolescents' views on inequality
-and deservingness. For example, Jonsson and Beach
-(\citeproc{ref-jonsson_institutional_2015}{2015}) study revealed that
-higher-status adolescents in Sweden tend to perpetuate social class
-stereotypes while describing the vocational and academic tracks.
-Academic track students are depicted as wealthy, intelligent, ambitious,
-and diligent, while vocational track students are characterized as poor,
-unambitious, unintelligent, and lackadaisical. These stereotypes help
-individuals maintain a sense of superiority over others and legitimize
-the prevailing social hierarchies and economic disparities
-(\citeproc{ref-jost_attitudinal_2000}{Jost and Burgess 2000})
-
-H3: Students from schools of higher social status will show larger
+and deservingness. For example, \citet{jonsson_institutional_2015} study
+revealed that higher-status adolescents in Sweden tend to perpetuate
+social class stereotypes while describing the vocational and academic
+tracks. Academic track students are depicted as wealthy, intelligent,
+ambitious, and diligent, while vocational track students are
+characterized as poor, unambitious, unintelligent, and lackadaisical.
+These stereotypes help individuals maintain a sense of superiority over
+others and legitimize the prevailing social hierarchies and economic
+disparities \citep{jost_attitudinal_2000}. In this line, and besides
+status, we also expect that schools that exhibit average better
+performance scores in standardized large-scale assessments would also
+show larger market justice preferences. The correspondent hypotheses
+are:
+
+\(H_3\): Students from schools of higher social status will show larger
 market justice preferences.
 
-H4: Students from schools with higher average levels of academic
-achievement will show larger market justice preferences.
-
-{[}interacciones{]}
-
-H5: The perception of meritocracy in school and society will moderate
-the effect of family social status on market justice preferences.
-
-H6: The perception of meritocracy in school and society will moderate
-the effect of school status on market justice preferences.
-
-H7: The perception of meritocracy in school and society will moderate
-the effect of school academic achievement on market justice preferences.
-
-\subsubsection{Summary of hypotheses}\label{summary-of-hypotheses}
+\(H_4\): Students from schools with an average better performance in
+standardized achievement tests will show larger market justice
+preferences.
+
+The last set of hypotheses deals with the moderation effects. We propose
+that the link between meritocracy and market justice could be stronger
+for those of high status and attending schools with larger achievement
+scores. In this regard, the effects of status and merit would not be
+independent, as it would be expected that those who succeed in terms of
+educational rewards and career paths - usually those of higher status -
+perceive more meritocracy, which would reflect on larger market justice
+preferences. In similar terms, schools of better performance would
+promote meritocratic perception, strengthening market justice
+preferences:
+
+\(H_5\): The family's social status will moderate the effect of
+meritocratic perception on market justice preferences.
+
+\(H_6\): The school's status will moderate the effect of meritocratic
+perceptions on market justice preferences.
+
+\(H_7\): The school's performance in standardized achievement tests will
+moderate the effect of meritocratic perceptions on market justice
+preferences.
+
+Figure~\ref{fig-hypotheses} depicts a summary of the research hypothesis
+described before. Market justice preferences are our main concept under
+study (vis-a-vis dependent variable), which refers to the justification
+of better access to health, education, and pensions based on payment
+capacities (as we detail further in the methods section). The
+explanatory concepts (independent variables) are at two levels: student
+(individual) variables and school (context) variables. The moderation
+hypotheses are represented through arrows pointing to the arrow between
+meritocratic perception and market justice preferences, which will be
+estimated through interaction effects.
 
@@
 
-\includegraphics{input/img/hypotheses.png}
+\includegraphics[width=6.25in,height=\textheight]{input/img/hypothesis2.png}
 
@@
 
-\subsection{}\label{section}
-
-\subsection{Methods}\label{methods}
-
-\subsubsection{Data}\label{data}
-
-The main data source for the analysis is the First Study of Citizenship
-Education in Chile, carried out by the Education Quality Agency of the
-Ministry of Education. The application date was November 9, 2017. The
-target population of this study is Eighth-grade students from 242
-schools. In the data, there are 8,589 students and 6,770 parents. This
-database was analyzed with the R package ``ResponsePatterns'' to detect
-possible repetitive and ``careless'' response patterns and thus
-contribute to a higher quality of research data (Gottfried, et al 2022).
-Responses from 171 students and 79 parents were removed, which, when
-merged, gave a total of 6,270 valid cases.
-
-The analysis of school variables includes data from the Ministry of
-Education's SIMCE 2017 database. This database contains information at
-the school level, such as the administrative dependency, its
-socioeconomic classification, and the achievement scores obtained in the
-mathematics and language census tests. It is available for free use on
-the MINEDUC {[}web page{]}.
-
-After eliminating missing cases, the final sample used in the analysis
-was based on 5,047 students and parents of 231 schools for the dependent
-variable of access to social services.
-
-\subsubsection{Variables}\label{variables}
-
-\textbf{Dependent variables}
-
-This study has three dependent variables related to the justification of
-social inequality in specific policy domains. The first asks whether
-access to social services should be conditional on income, i,.e., ``It
-is just that in Chile people who can pay have a better education for
-their children''. Students rated their preferences using the following
-responses: ``strongly disagree'', ``Disagree'', ``Agree'', and
-''strongly agree''. An average index is built with these items
-(Cronbach's Alpha = 0,86). (Apéndice: items en español y su
-correspondiente traducción al inglés)
-
-Table~\ref{tbl-desc-dependientes} shows the items used, their response
-categories, and their frequencies.
+\section{Data, variables and method}\label{data-variables-and-method}
+
+\subsection{Data}\label{data}
+
+The First Study of Civic Education in Chile, conducted by the Agency for
+Quality Education of the Ministry of Education in 2017, serves as the
+main data source. This database is composed by a civic knowledge test
+score and a series of items batteries measuring different aspects of
+citizenship. The target population includes 8th-grade students in 242
+schools nationwide, as well as their parents and teachers. Initially,
+the database contains 8,589 student observations and 6,770 parent
+observations. To ensure higher data quality and considering the survey's
+unit of analysis, we removed 171 student cases and 79 parent cases that
+exhibited repetitive and careless response patterns
+\citep{gottfried_autocorrelation_2022}. Additionally, we utilized data
+from the System for Measuring the Quality of Education (SIMCE) conducted
+by the Ministry of Education in 2017, which provides school-level
+information such as administrative dependence, socioeconomic
+classification, and results obtained in standardized mathematics and
+language tests. After processing the variables and removing missing
+cases, the final database for this study includes a two-level stratified
+sample composed of 5,047 students and parents (level 1) nested within
+231 schools (level 2).
+
+\subsection{Variables}\label{variables}
+
+\textbf{\emph{Individual level}}
+
+\textbf{Market Justice Preferences}: The dependent variables in this
+study are market justice preferences. This construct is measured through
+three variables that address the degree of justification regarding
+whether access to social services in pensions, education, and health
+should be income conditional. Justification of inequality in health is
+measured by the item: ``Is it fair in Chile that people with higher
+incomes can access better healthcare than those with lower incomes?''
+The same question is asked for education and pensions. In all cases,
+respondents indicate their preferences on a Likert scale ranging from
+``completely disagree'' (1) to ``completely agree'' (4). Additionally,
+we include a summarized indicator of ``market justice preferences'',
+measured by an average index across these items (α = 0.86), with values
+ranging from 1 to 4, where higher values represent greater preferences
+for market justice (see Table~\ref{tbl-desc-dependientes}). We analyzed
+these items independently as well as by the average index.
 
@@
 
-\textbf{Independent variables}
-
-For the primary independent variable, the perception of meritocracy,
-five items address the perception of rewards according to effort and
-intelligence at the school and societal levels. At the school level,
-students answer whether ``Intelligence is important to get good grades''
-and ``Effort is important to get good grades''. At the societal level,
-students respond to the following questions: ``In Chile, people are
-rewarded for their efforts'', ``In Chile, people get what they
-deserve'', and ``In Chile, people are rewarded for their intelligence
-and skills''. Each item was answered on a four-point scale ranging from
-``Strongly disagree'' to ``Strongly agree''.
-
-The rest of the independent variables are divided into individual and
-school levels. At the individual level, family socioeconomic status was
-measured by the parents' highest educational level and the number of
-books at home. Likewise, an index of access to technology includes the
-number of computers, tablets, and cell phones at home, as well as
-whether there is an Internet connection.
-Table~\ref{tbl-desc-independent} shows the items used, their response
-categories, and their frequency.
+\textbf{Perception of Meritocracy}: The main independent variable refers
+to the perception of meritocracy, operationalized through five items
+addressing the perception of rewards based on talent and intelligence at
+both the school and societal levels. At the school level, students
+respond to whether ``Intelligence is important for getting good grades''
+and ``Effort is important for getting good grades''. At the societal
+level, students respond to the following questions: ``In Chile, people
+are rewarded for their effort'', ``In Chile, people get what they
+deserve'' and ``In Chile, people are rewarded for their intelligence and
+skills''. Each item was answered on a four-point Likert scale ranging
+from ``completely disagree'' (1) to ``completely agree'' (4).
+
+\textbf{Family Socioeconomic Status}: The socioeconomic status of
+students' families is measured using two indicators. First, the highest
+educational level attained by the parents, with categories: ``8th grade
+or less,'' ``Secondary education,'' ``Technical higher education,''
+``University or postgraduate,'' and ``No response.'' The inclusion of
+the ``No response'' category is due to its high frequency in the data;
+omitting it could obscure relevant associations. Second, the number of
+books in the household is used, categorized as ``Less than 25'' and
+``More than 25.''
+
+Table~\ref{tbl-desc-independent} shows the individual level variables
+used, their response categories and their frequencies.
 
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1028}}@{}}
-\caption{Independent
+\caption{Individual level
 variables}\label{tbl-desc-independent}\tabularnewline
@@
 \end{minipage} \\
-Technology access index & \begin{minipage}[t]{\linewidth}\raggedright
-Mean (sd) : 7.8 (2.5)\\
-min \textless{} med \textless{} max:\\
-0 \textless{} 8 \textless{} 12\\
-IQR (CV) : 3 (0.3)\strut
-\end{minipage} & 13 distinct values &
-\begin{minipage}[t]{\linewidth}\raggedright
-6272\\
-(100.0\%)\strut
-\end{minipage} \\
 \end{longtable}
 
-The school-level variables are the administrative dependency of the
-school, the socioeconomic classification made by the Ministry of
-Education, the level of performance in the SIMCE test of the school, and
-the proportion of parents with university or postgraduate degrees.
-Table~\ref{tbl-desc-school} shows the items used, response categories,
-and frequency.
+\textbf{\emph{Contextual level}}
+
+This study focuses on two school-level characteristics: socioeconomic
+status and academic performance. Socioeconomic status is assessed using
+the Ministry of Education's classification, measured as an ordinal item
+with five categories ranging from ``low'' (1) to ``high'' (5). Academic
+performance is measured using the school's results in the SIMCE (System
+of Measurement of Educational Quality) standardized tests, which are
+administered yearly at different educational levels in areas such as
+language and mathematics. These results are categorized as ``low,''
+``medium,'' and ``high.'' The contextual level items used, response
+categories, and their frequencies are detailed in
+Table~\ref{tbl-desc-school}.
 
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.4040}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2626}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2222}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1111}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.4138}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2184}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2414}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1264}}@{}}
 \caption{School context variables}\label{tbl-desc-school}\tabularnewline
@@
 \endlastfoot
-Proportion of parents with university level by school &
-\begin{minipage}[t]{\linewidth}\raggedright
-Mean (sd) : 0.2 (0.2)\\
-min \textless{} med \textless{} max:\\
-0 \textless{} 0.1 \textless{} 0.9\\
-IQR (CV) : 0.2 (0.9)\strut
-\end{minipage} & 103 distinct values &
-\begin{minipage}[t]{\linewidth}\raggedright
-6272\\
-(100.0\%)\strut
-\end{minipage} \\
-SIMCE score by school & \begin{minipage}[t]{\linewidth}\raggedright
-1. Low\\
-2. Medium\\
-3. High\strut
-\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
-2091 (33.3\%)\\
-2091 (33.3\%)\\
-2090 (33.3\%)\strut
-\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
-6272\\
-(100.0\%)\strut
-\end{minipage} \\
-Administrative dependency of school &
-\begin{minipage}[t]{\linewidth}\raggedright
-1. Public\\
-2. Subsidized private\\
-3. Private\strut
-\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
-2659 (42.4\%)\\
-3169 (50.5\%)\\
-444 ( 7.1\%)\strut
-\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
-6272\\
-(100.0\%)\strut
-\end{minipage} \\
 Socioeconomic level of school &
@@
 \end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
-720 (11.5\%)\\
-2282 (36.4\%)\\
-1383 (22.1\%)\\
-1309 (20.9\%)\\
-578 ( 9.2\%)\strut
-\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
-6272\\
+49 (21.1\%)\\
+92 (39.7\%)\\
+43 (18.5\%)\\
+29 (12.5\%)\\
+19 ( 8.2\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+232\\
 (100.0\%)\strut
 \end{minipage} \\
+SIMCE score achievement by school &
+\begin{minipage}[t]{\linewidth}\raggedright
+1. Low\\
+2. Medium\\
+3. High\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+108 (46.6\%)\\
+69 (29.7\%)\\
+55 (23.7\%)\strut
+\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
+232\\
+(100.0\%)\strut
+\end{minipage} \\
 \end{longtable}
 
-\subsubsection{Methods}\label{methods-1}
-
-The data has a hierarchical structure of students nested in schools, so
-the model estimation is performed in a multilevel (random effects)
-framework. This modeling approach lets us correctly estimate individual
-and contextual effects in a single model. We estimate cumulative link
-mixed models for the ordinal dependent variables, whereas we use linear
-mixed effects models for the average index of inequality justification.
+\textbf{\emph{Controls}}
+
+A series of control variables are included. At the individual level, an
+index of access to technology is constructed based on the number of
+computers, tablets, and cell phones in the household, as well as
+internet connectivity. At the contextual level, we use the
+administrative dependence of schools, classified as ``Public,''
+``Subsidized Private,'' or ``Private,'' and the proportion of parents
+with university or postgraduate education at each school.
+
+\subsection{Methods}\label{methods}
+
+Given the hierarchical structure of the data, with students nested
+within schools, model estimation is performed within a multilevel
+framework. These models are appropriate for capturing both individual
+and contextual effects in a single analysis, allowing for the estimation
+of fixed effects between groups and random effects that vary from one
+group to another \citep{bell_fixed_2019, hox_multilevel_2010}.
+Cumulative link mixed models were employed for the ordinal dependent
+variables, while linear mixed-effects models were applied for the
+average market justice preference index.
 
@@
 \href{https://doi.org/10.17605/OSF.IO/UFSDV}{link}. The statistical
-analysis of this research was performed using the free software R
-version 4.1.3.
-
-\subsection{Analysis}\label{analysis}
-
-\subsubsection{Descriptive analysis}\label{descriptive-analysis}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{paper_files/figure-pdf/unnamed-chunk-11-1.pdf}
+analysis for this research was conducted using the \emph{lme4} package
+in R version 4.1.3.
+
+\section{Results}\label{results}
+
+\subsection{Descriptive}\label{descriptive}
+
+Figure~\ref{fig-marketjustice} illustrates the distribution of responses
+to the three items on market justice preferences (healthcare, pensions,
+and education), indicating that most respondents tend to disagree or
+strongly disagree with the idea that it is just for people with higher
+incomes to have access to better services in these areas. The strongest
+opposition is observed in the healthcare sector, as the majority of
+respondents (66.4\%) are against the idea that higher-income individuals
+should have access to better healthcare services. Similarly, regarding
+the concept of justice related to access to better pensions based on
+income, the level of disagreement reaches 63.0\%. This decreases
+significantly in the field of education, as the percentage of
+respondents who disagree or strongly disagree with those with higher
+incomes having access to better education decreases to 58.8\%. Only a
+small proportion agree that access to healthcare should be conditional
+on income (23.6\%), and a minority strongly agree (10\%), as is the case
+with pensions. However, this increases in the case of education, being
+the area where there is a significant proportion of respondents who
+agree or strongly agree with access based on market justice (41.2\%).
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{template-mdpi_files/figure-pdf/fig-marketjustice-1.pdf}
 
@@
 
-\caption{Market justice preferences in education, health and pensions by
-social and school meritocracy}
+\caption{\label{fig-marketjustice}Market justice preferences in health,
+pensions and education}
 
@@
 
-Figure 3 shows a series of graphs depicting the association between the
-variables of market justice preferences - in education, health, and
-pensions - and the variables of meritocratic perception at school
-(effort and talent) and in society (effort, talent, and deservingness)
-(see conceptual diagram en figure X). On the left, we observe the social
-meritocracy diagrams, while on the right the school meritocracy diagrams
-are shown. For the three variables of perception of meritocracy in
-society the relationship is clear, since the average of market justice
-preferences increases the more there is agreement that people are
-rewarded for their effort, merit, and talent. This relationship needs to
-be clarified in the case of the variables of perception of meritocracy
-at school. To the extent that there is more agreement that the
-perception of talent is essential for obtaining good grades, the average
-of market justice preferences increases, but this relationship is not as
-clear as with the variables of meritocracy in society. In addition, the
-graph does not show a clear trend in the relationship between the
-perception that effort is essential to obtain good grades and market
+Regarding the perception of meritocracy, Figure~\ref{fig-meritocracy}
+displays the frequency distribution of five items related to the
+dimensions of school and society. At the school level, there is a strong
+belief that both effort and talent are important for achieving good
+grades, with especially high agreement for effort. Specifically, while
+78.6\% agree or strongly agree on the importance of talent for obtaining
+good grades, this proportion rises to 96.8\% for effort, indicating that
+respondents value effort more than talent in this context.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{template-mdpi_files/figure-pdf/fig-meritocracy-1.pdf}
+
+}
+
+\caption{\label{fig-meritocracy}Social and school meritocracy}
+
+\end{figure}%
+
+At the societal level, Figure~\ref{fig-meritocracy} shows that a
+majority of respondents agree or strongly agree that individuals are
+rewarded for their talents (63.7\%) and efforts (62\%). Nevertheless, a
+notable proportion perceives that both effort and talent are not
+adequately rewarded in society, with 38.1\% and 36.3\% respectively. The
+perception of whether people get what they deserve in society is more
+divided; while a slim majority (56.5\% combining agree and strongly
+agree) agrees, a significant minority (43.6\%) disagrees, indicating
+that this is the most contentious item among respondents. There is
+general agreement that people are rewarded for talent and effort in
+society, but the consensus is weaker compared to school meritocracy.
+
+Figure~\ref{fig-bivariate} shows a series of graphs depicting the
+association between the variables of market justice preferences - in
+education, health, and pensions - and the variables of meritocratic
+perception at school (effort and talent) and in society (effort, talent,
+and deservingness) (see conceptual diagram in
+Figure~\ref{fig-hypotheses}). In general we observe a positive
+association between meritocracy and market justice preferences for the
+societal questions, but this is not as straightforward for the school
+items, particularly for the one referred to effort.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{template-mdpi_files/figure-pdf/fig-bivariate-1.pdf}
+
+}
+
+\caption{\label{fig-bivariate}Market justice preferences in education,
+health and pensions by social and school meritocracy}
+
+\end{figure}%
+
+\subsection{Multivariate}\label{multivariate}
+
+This section presents the results of the multilevel models. We begin by
+showing the results of the cumulative link mixed models in
+Table~\ref{tbl-ordinal-reg} for the ordinal dependent variables of
+justice in health, pensions, and education separately in order to
+compare the effect of the predictors. Then, we present the results of
+the linear mixed-effects models for the average index of preference for
+market justice in Table~\ref{tbl-lineal-reg}\footnote{The table shows
+  the effects of individual and contextual-level independent variables
+  related to our hypotheses; however, the models also incorporate the
+  control variables for each of the three dependent variables separately
+  and for the average index (complete models are available in the
+  \hyperref[appendix]{Supplementary Material}).}.
+
+For the health item, the intraclass correlation (ICC) obtained for the
+null model (not shown) indicates that 5\% of the total variance of this
+item is associated with specific characteristics of the schools, while
+it is 3\% for the pensions item and 4\% for the education item
+\citep[ICC based on][p.15]{hox_multilevel_2010}. For the average index
+of market justice preferences, the ICC reaches 4\%, indicating that
+overall, only a small percentage of the total variance in students'
+market justice preferences is associated with school characteristics,
+thus limiting the possibility of finding effects at this second level of
+analysis.
+
+\begin{table}
+
+\caption{\label{tbl-ordinal-reg}Cumulative link multilevel models of
+differential acces justificacion of health, pensions, and education}
+
+\centering{
+
+\begin{center}
+\scalebox{0.8}{
+\begin{threeparttable}
+\begin{tabular}{l c c c c c c}
+\toprule
+ & \multicolumn{2}{c}{Health} & \multicolumn{2}{c}{Pensions} & \multicolumn{2}{c}{Education} \\
+\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
+ & Model 1 & Model 2 & Model 1 & Model 2 & Model 1 & Model 2 \\
+\midrule
+School talent                             & $0.23^{***}$  & $0.22^{***}$  & $0.17^{***}$  & $0.16^{***}$  & $0.25^{***}$  & $0.25^{***}$  \\
+                                          & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+School effort                             & $-0.26^{***}$ & $-0.26^{***}$ & $-0.26^{***}$ & $-0.25^{***}$ & $-0.23^{***}$ & $-0.23^{***}$ \\
+                                          & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
+Social talent                             & $0.16^{***}$  & $0.16^{***}$  & $0.12^{**}$   & $0.13^{**}$   & $0.14^{**}$   & $0.15^{***}$  \\
+                                          & $(0.05)$      & $(0.05)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
+Social effort                             & $0.11^{*}$    & $0.11^{*}$    & $0.27^{***}$  & $0.27^{***}$  & $0.15^{**}$   & $0.15^{**}$   \\
+                                          & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      \\
+Deservingness                             & $0.52^{***}$  & $0.51^{***}$  & $0.38^{***}$  & $0.37^{***}$  & $0.44^{***}$  & $0.43^{***}$  \\
+                                          & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      \\
+Parental education (Ref.= Non-university) &               &               &               &               &               &               \\
+                                          &               &               &               &               &               &               \\
+\quad University or posgraduate           & $0.05$        & $0.08$        & $0.14$        & $0.12$        & $0.09$        & $0.07$        \\
+                                          & $(0.08)$      & $(0.08)$      & $(0.08)$      & $(0.08)$      & $(0.08)$      & $(0.08)$      \\
+\quad Missing                             & $0.10$        & $0.11$        & $0.12^{*}$    & $0.11$        & $0.13^{*}$    & $0.12^{*}$    \\
+                                          & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      \\
+More than 25 books (Ref.= Less than 25)   & $-0.16^{**}$  & $-0.13^{*}$   & $-0.13^{*}$   & $-0.13^{*}$   & $-0.07$       & $-0.07$       \\
+                                          & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      \\
+Socioeconomic level (Ref.= Low)           &               &               &               &               &               &               \\
+                                          &               &               &               &               &               &               \\
+\quad SES Medium                          &               & $0.08$        &               & $0.02$        &               & $0.08$        \\
+                                          &               & $(0.10)$      &               & $(0.10)$      &               & $(0.10)$      \\
+\quad SES High                            &               & $0.71^{*}$    &               & $0.72^{**}$   &               & $0.93^{***}$  \\
+                                          &               & $(0.28)$      &               & $(0.26)$      &               & $(0.28)$      \\
+Achievement score (Ref.= Low)             &               &               &               &               &               &               \\
+                                          &               &               &               &               &               &               \\
+\quad Simce Medium                        &               & $-0.29^{***}$ &               & $-0.19^{*}$   &               & $-0.16^{*}$   \\
+                                          &               & $(0.08)$      &               & $(0.08)$      &               & $(0.08)$      \\
+\quad Simce High                          &               & $-0.56^{***}$ &               & $-0.53^{***}$ &               & $-0.46^{***}$ \\
+                                          &               & $(0.10)$      &               & $(0.09)$      &               & $(0.10)$      \\
+\midrule
+Controls                                  & Yes           & Yes           & Yes           & Yes           & Yes           & Yes           \\
+BIC                                       & $12747.58$    & $12761.47$    & $13103.81$    & $13112.87$    & $13329.33$    & $13352.18$    \\
+Num.obs                                   & $5144$        & $5144$        & $5177$        & $5177$        & $5159$        & $5159$        \\
+Num. groups: Schools                      & $231$         & $231$         & $232$         & $232$         & $231$         & $231$         \\
+Var: Schools (Intercept)                  & $0.11$        & $0.05$        & $0.09$        & $0.04$        & $0.10$        & $0.05$        \\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[flushleft]
+\scriptsize{\item Note: Cells contain regression coefficients with standard errors in parentheses.
+Control variables are included.
+$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$ \\ \item Source: own elaboration based on First Study of Civic Education 2017.}
+\end{tablenotes}
+\end{threeparttable}
+}
+\caption{}
+\label{table:coefficients}
+\end{center}
+
+}
+
+\end{table}%
+
+Model 1 in Table~\ref{tbl-ordinal-reg} for the three dependent variables
+of justice in differential access to health, pensions, and education
+demonstrates a similar trend regarding the effect of the perception of
+meritocracy on these items, with all effects being statistically
+significant (\emph{p} \textless{} 0.05). Regarding the variables of
+social meritocracy, the three variables---talent, effort, and
+deservingness---show that as these increase, the justification for
+differentiated access to pensions, education, and health also increases.
+In the context of school meritocracy, the effects are mixed: as
+perceived school talent increases, the justification for differentiated
+access to these services increases; conversely, as perceived school
+effort increases, the justification decreases, holding all other
+variables constant.
+
+Additionally, Model 1 examines family status through parents'
+educational level and the number of books at home. For health, pensions,
+and education, a negative but statistically non-significant association
+with parents' education is observed. However, having more than 25 books
+at home shows a negative and statistically significant effect on market
+justice in health and pensions, suggesting less justification for
+differentiated access compared to those with fewer books.
+
+Model 2 in Table~\ref{tbl-ordinal-reg} incorporates school-level
+variables. The results show that high socioeconomic level schools show
+greater preferences for market justice in health, pensions, and
+education compared to low socioeconomic level schools, with
+statistically significant results (\emph{p} \textless{} 0.05). Moreover,
+and contrary to our hypothesis, schools with higher academic performance
+show negative and statistically significant effects across the three
+items.
+
+\begin{table}
+
+\caption{\label{tbl-lineal-reg}Linear mixed-effects models for
+meritocracy perception and market justice preferences}
+
+\centering{
+
+\begin{center}
+\scalebox{0.85}{
+\begin{threeparttable}
+\begin{tabular}{l c c c c}
+\toprule
+ & Model 1 & Model 2 & Model 3 & Model 4 \\
+\midrule
+Constant                                  & $1.31^{***}$  & $1.30^{***}$  & $1.36^{***}$  & $1.39^{***}$  \\
+                                          & $(0.09)$      & $(0.10)$      & $(0.10)$      & $(0.10)$      \\
+School talent                             & $0.11^{***}$  & $0.11^{***}$  & $0.10^{***}$  & $0.10^{***}$  \\
+                                          & $(0.01)$      & $(0.01)$      & $(0.01)$      & $(0.01)$      \\
+School effort                             & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ \\
+                                          & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Social talent                             & $0.07^{***}$  & $0.07^{***}$  & $0.07^{***}$  & $0.07^{***}$  \\
+                                          & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Social effort                             & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  \\
+                                          & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Deservingness                             & $0.20^{***}$  & $0.20^{***}$  & $0.20^{***}$  & $0.20^{***}$  \\
+                                          & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Parental education (Ref.= Non-university) &               &               &               &               \\
+                                          &               &               &               &               \\
+\quad University or posgraduate           &               & $0.05$        & $0.05$        & $0.05$        \\
+                                          &               & $(0.03)$      & $(0.04)$      & $(0.04)$      \\
+\quad Missing                             &               & $0.06^{*}$    & $0.06^{*}$    & $0.06^{*}$    \\
+                                          &               & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+More than 25 books (Ref.= Less than 25)   &               & $-0.05$       & $-0.05$       & $-0.04$       \\
+                                          &               & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+Socioeconomic level (Ref.= Low)           &               &               &               &               \\
+                                          &               &               &               &               \\
+\quad SES Medium                          &               &               & $-0.01$       & $0.01$        \\
+                                          &               &               & $(0.05)$      & $(0.05)$      \\
+\quad SES High                            &               &               & $0.31^{*}$    & $0.37^{**}$   \\
+                                          &               &               & $(0.14)$      & $(0.13)$      \\
+Achievement score (Ref.= Low)             &               &               &               &               \\
+                                          &               &               &               &               \\
+\quad Simce Medium                        &               &               &               & $-0.11^{**}$  \\
+                                          &               &               &               & $(0.04)$      \\
+\quad Simce High                          &               &               &               & $-0.26^{***}$ \\
+                                          &               &               &               & $(0.04)$      \\
+\midrule
+Controls                                  & Yes           & Yes           & Yes           & Yes           \\
+Deviance                                  & $12368.01$    & $12359.40$    & $12344.10$    & $12310.90$    \\
+Deviance Test (p)                         & $0.00$        & $0.00$        & $0.00$        & $0.00$        \\
+BIC                                       & $12517.46$    & $12550.41$    & $12595.11$    & $12589.48$    \\
+Num.obs                                   & $5047$        & $5047$        & $5047$        & $5047$        \\
+Num. groups: Schools                      & $231$         & $231$         & $231$         & $231$         \\
+Var: Schools (Intercept)                  & $0.02$        & $0.02$        & $0.02$        & $0.01$        \\
+Var: Residual                             & $0.66$        & $0.66$        & $0.67$        & $0.66$        \\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[flushleft]
+\scriptsize{\item Note: Cells contain regression coefficients with standard errors in parentheses.
+Control variables are included.
+$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$ \\ \item Source: own elaboration based on First Study of Civic Education 2017.}
+\end{tablenotes}
+\end{threeparttable}
+}
+\caption{}
+\label{table:coefficients}
+\end{center}
+
+}
+
+\end{table}%
+
+Table~\ref{tbl-lineal-reg} presents the results of linear mixed-effects
+models for the market justice preference index. Model 1 shows that
+higher perceived societal meritocracy, through effort, talent, and
+deservingness, is positively and significantly associated with market
+justice preferences. School meritocracy perceptions yield mixed effects:
+perceived talent increases market justice preference, while perceived
+effort decreases it, both statistically significant (\emph{p}
+\textless{} 0.001). Model 2 includes family status, finding no
+statistically significant effects from parents' education or the number
+of books at home. Regarding school characteristics, Models 3 and 4
+introduce socioeconomic level and academic performance. The results of
+Model 3 suggest a statistically significant relationship between school
+socioeconomic levels and market justice preferences, specifically those
+schools with high socioeconomic levels (\(\beta\) = 0.37) compared to
+low socioeconomic levels. However, Model 4 indicates that schools with
+medium (\(\beta\) = -0.11) and high (\(\beta\) = -0.26) academic
+performance present lower market justice preferences compared to
+low-performing schools, with these effects being statistically
+significant when controlling for other variables.
+
+\begin{table}
+
+\caption{\label{tbl-interact}Interactions effects}
+
+\centering{
+
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
+\begin{tabu} to \linewidth {>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
+\toprule
+\multicolumn{1}{c}{\textbf{ }} & \multicolumn{5}{c}{\textbf{Market Justice Preferences}} \\
+\cmidrule(l{3pt}r{3pt}){2-6}
+Variable & School talent & School effort & Social talent & Social effort & Deservingness\\
+\midrule
+\textbf{University or Postgraduate} & -0.034 & -0.012 & -0.073. & -0.079* & -0.118**\\
+\cmidrule{1-6}
+\textbf{More than 25 books} & -0.015 & -0.046 & -0.029 & -0.081** & -0.035\\
+\cmidrule{1-6}
+\textbf{School SES high} & 0.004 & 0.001 & -0.057. & -0.063* & -0.08**\\
+\cmidrule{1-6}
+\textbf{Simce Medium} & 0.001 & 0.032 & -0.037 & -0.041 & -0.082*\\
+\cmidrule{1-6}
+\textbf{Simce High} & -0.016 & 0.008 & -0.043 & -0.063. & -0.087*\\
+\bottomrule
+\multicolumn{6}{l}{\textsuperscript{} Note: ***p < 0.001, **p < 0.01, *p < 0.05. Source: own elaboration based on First Study of Civic}\\
+\multicolumn{6}{l}{Education 2017.}\\
+\end{tabu}}
+
+}
+
+\end{table}%
+
+Table~\ref{tbl-interact} displays the interaction terms related to our
+hypotheses. The results suggest that students' family background
+moderates the relationship between their meritocratic perceptions and
+their justification of market allocation in social services.
+Specifically, the relationship between social effort and market justice
+preferences is less positive for students whose parents have university
+or postgraduate education. Lower-status students (measured by parental
+education) prefer more market justice when they strongly adhere to
+meritocratic perceptions of deservingness. A similar moderating effect
+is seen with family cultural capital: the relationship between effort
+and market justice preferences is less positive for students with more
+than 25 books at home.
+
+At the school level, we observe that school status and students'
+meritocratic beliefs moderate their preferences for market justice.
+High-status schools show less preference for market justice when their
+students, on average, have a greater perception of meritocracy, as
+measured by effort, deservingness, and talent. Conversely, low-achieving
+schools prefer more market justice when their students have stronger
+meritocratic beliefs. Figure~\ref{fig-interaction} illustrates this
+moderating effect of school achievement on the relationship between
+deservingness and market justice. No moderating effects were observed at
+the school level for students' meritocratic beliefs regarding effort or
+talent in achieving good grades.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{template-mdpi_files/figure-pdf/fig-interaction-1.pdf}
+
+}
+
+\caption{\label{fig-interaction}Interaction between deservingness and
+justification of inequality by school SIMCE achievement}
+
+\end{figure}%
+
+\section{Discussion}\label{discussion}
+
+The previous analyses aimed to explore the relationship between
+perceptions of meritocracy and preferences for market justice among
+students in Chilean society. Several models estimated how students'
+beliefs about the rewards for talent and effort in both societal and
+school contexts influence their attitudes toward access to social
+services like health, education, and pensions based on pay capacity.
+Additionally, the models examined the impact of students' socioeconomic
+status and school achievement on these preferences, assessing whether
+perceptions of meritocracy moderate these effects. The findings offer
+insights into the complex interplay between meritocratic beliefs,
+socioeconomic factors, and attitudes toward market justice criteria.
+
+Hypothesis \(H_{1a}\) posited that students who perceive greater
+meritocracy in society would show stronger preferences for market
+justice. The results support this hypothesis, indicating a positive
+relationship between the perception that talent and effort are rewarded
+in society and the preferences for market justice in access to services
+such as health, education, and pensions. This finding aligns with
+previous research suggesting that individuals who believe in a
+meritocratic system are more likely to justify social inequalities
+\citep{batruch_belief_2022, wiederkehr_belief_2015}. It is noteworthy
+that among the items measuring meritocratic perception, consistently the
+one about deservingness shows a larger effect when compared to the ones
+about talent and effort. Although the three items correlate, there is
+something distinctive about deservingness that probably encompasses
+other aspects beyond meritocracy, an aspect to which we will come back
+in the conclusions. Moving now to hypothesis \(H_{1b}\), this is simply
+a variation of 1a, replacing the perception of meritocracy at society by
+the one at school. In this case we have only two items of meritocratic
+perception - unfortunately deservingness was not measured at school
+level - and the direction of the effects is not as straightforward as in
+the case of societal meritocracy in \(H_{1a}\): positive for talent, but
+negative for effort. Still, one of the limitation for the comparison
+between perception of meritocracy and society is the difference of the
+items' phrasing. Whereas for society the item refers to the extent to
+which efforts are rewarded, at school is about the importance of effort,
+whereby something that is important does not necessarily mean that it is
+enough to be rewarded. Furthermore, the school-effort item is the one
+with the lowest variation among all meritocratic items (more than 95\%
+agree or strongly agree), and with the lowest effect size.
+
+Hypothesis \(H_2\) and hypothesis \(H_3\) propose that socioeconomic
+status could influence students' preferences, but the results show
+different patterns. On the one hand, \(H_2\) proposed that students from
+higher socioeconomic status families would show stronger preferences for
+market justice, which is not supported by the results. Although students
+from families with higher educational levels and more economic resources
+tend to prefer less market justice, this relationship is not
+statistically significant. This might be due to the influence of other
+contextual and cultural factors that modulate attitudes toward market
+justice \citep{kohn_social_1963, kohn_class_1969, almas_fairness_2017}.
+Notably, socioeconomic status measured by parental educational level and
+the number of books at home do not explain variations in market justice
+preferences, suggesting that other variables, such as market justice
+preferences of parents, could play an essential role in the attitudinal
+socialization process. On the other hand, Hypothesis \(H_3\) held that
+students from schools with higher socioeconomic status would show
+stronger preferences for market justice. The results confirm this
+hypothesis since students from high-status schools tend to prefer more
+market justice, compared to low-status schools. This finding, in line
+with the existing literature
+\citep{jonsson_institutional_2015, jost_attitudinal_2000}, does allow us
+to affirm the hypothesis about the role of the school environment,
+particularly socioeconomic segregation, in influencing attitudes towards
+market justice and the acceptance of inequalities. A possible
+explanation is the role of the context at the school level, which could
+have a more relevant role than the family socioeconomic status, since
+the highly stratified nature of the Chilean education system reinforces
+socioeconomic divisions and socializes students into different normative
+frameworks.
+
+Hypothesis \(H_4\) held that students from schools with higher school
+achievement would show stronger preferences for market justice,
+following the argument that such schools could foster meritocratic
+attitudes by showing a larger promotion of talent and effort in order to
+get better results. Nevertheless, the results show quite the opposite:
+schools with better academic results (as measured by a national
+standardized test) show in average less preferences for market justice.
+This finding is one of the most strongest and consistent throughout the
+models and it is relevant given the centrality of school performance in
+educational institutions. One possible explanation of this link is that
+students in high-achieving schools may be more exposed to discussions
+and a deeper understanding of inequality and social justice, leading to
+a critical perspective toward the weight of market mechanisms in the
+access to social services.
+
+Hypotheses \(H_5\) and \(H_6\) proposed that socioeconomic status would
+moderate the effect of meritocratic perceptions on market justice
+preferences. Hypothesis \(H_5\) refers to family socioeconomic status,
+whereby the results show a negative interaction with two of the three
+perceptual variables at the societal level. This means that the positive
+impact of the perception of social meritocracy on market justice
+preferences had a lower effect on students with parents with higher
+education and also on those with more books at home. Something similar
+occurs with hypothesis \(H_6\), where now school-level status moderates
+the effect of meritocracy variables, which in multilevel framework is
+referred as to cross-level interaction. The negative interaction in this
+context means that in schools of low socioeconomic levels, the positive
+effect of the social meritocracy variables on market justice preferences
+is stronger. These relations indicate that, although those who perceive
+society as meritocratic have greater preferences towards market justice,
+this influence is lower among those with a higher socioeconomic level.
+This finding highlights the complex interaction between socioeconomic
+status, school context, meritocratic beliefs, and preferences toward
+market justice. The moderating role of meritocratic perceptions suggests
+that students' views on social justice are shaped by their socioeconomic
+background and beliefs about how society rewards effort and talent. This
+underlines the importance of addressing structural inequalities and
+individual beliefs in educational and social policies.
+
+Hypothesis \(H7\) posited that the academic performance of schools could
+moderate the effect of meritocratic perceptions on preferences for
+market justice. This hypothesis is based on the argument that the level
+of academic performance in schools entails a differentiated promotion of
+meritocratic perceptions to achieve better results, which would, in
+turn, reinforce or diminish preferences for market justice. The results
+of the interactions between levels suggest that, indeed, in
+higher-performing schools, the effect of meritocratic perceptions on the
+justification of service provision based on market principles is weaker
+compared to lower-performing schools. Similarly to the effect of
+academic performance in \(H4\), a possible explanation for this
+moderation is that students in high-performing schools may be more
+exposed to knowledge about the effects of inequality and social justice.
+This exposure translates into an enlightening effect of education on
+inequality, which can mitigate the impact of meritocracy on market
 justice preferences.
 
-\subsubsection{Cumulative link mixed models for the Justification of
-inequality in education, health and
-pensions}\label{cumulative-link-mixed-models-for-the-justification-of-inequality-in-education-health-and-pensions}
-
-Three Cumulative link mixed models were estimated for the ordinal
-dependent variables of justice in differential access to pensions,
-education, and health according to income. Figure 4 shows the estimation
-of this regression model containing all the variables used in the study
-for the three dependent variables separately. However, this figure shows
-only the effect of meritocracy variables on society and school as
-independent variables. Complete models are available in the appendix.
-
-Figure 4 shows that for the three dependent variables of justice in
-differential access to pensions, education, and health, the trend is
-similar. In the context of school meritocracy, the effects are mixed. As
-school talent increases, the justification for differentiated access to
-pensions, education, and health increases; on the contrary, as the
-school effort variable increases, the justification for differentiated
-access to pensions, education, and health decreases, keeping the rest of
-the independent variables constant. Regarding the variables of social
-meritocracy, the three variables of talent, effort, and deservingness
-show that as these increase, the justification for differentiated access
-to pensions, education, and health also increases.
-
-\subsubsection{Multilevel regression models for market justice
-preferences}\label{multilevel-regression-models-for-market-justice-preferences}
-
-Table 4 shows the results of the multilevel estimation for justice
-market preferences. For this variable, the intraclass correlation
-obtained shows that the variation between schools corresponds to 4\% of
-the variation of students' preferences. This means that there is low
-variance between schools and therefore limits the possibilities of
-finding effects at the aggregate level.
-
-Model 1 introduces social meritocratic variables: effort (whether
-efforts are rewarded), deservingness (people get what they deserve), and
-social talent (intelligence and skills are rewarded in society). In line
-with our hypotheses, the perception of a meritocratic society is
-positively related to the justification of inequality. Model 2 shows a
-mixed picture: while those perceiving that talent is rewarded also
-justify the inequality, the perception that effort is rewarded at school
-is negatively related to justification of inequality. Family background
-variables in Model 3 reveal that education and technology access are not
-related to justification of inequality, whereby we observe a negative
-impact of family cultural capital as measured by the number of books at
-home. While school socio-structural variables added in Model 4 show no
-significant effects, average achievement scores in the SIMCE test depict
-a negative relationship with the dependent variable, meaning that
-students that attend schools with better achievement scores on average
-justify less inequality.
-
-In relation to model fit, when comparing the deviance with a model
-without predictors (null model), all the models have a statistically
-significant difference, with model 5 having the lowest deviance.
-According to Raudenbush and Bryk's (2002) estimate of R2, the level 1
-variance of model 5 is 0.11 and the level 2 variance is 0.72. The total
-variance of model 5 according to Snijders and Bosker (2012) is 0.14.
-
-\subsubsection{Interactions effects}\label{interactions-effects}
-
-The interaction terms in Table 6 suggest that students' family
-background moderates the relationship between their meritocratic
-perceptions in Chile and their justification of inequality. The
-direction of these effects confirms our initial prediction. Thus, in
-line with our Hypothesis 3, the relationship between social effort and
-justification of inequality becomes less positive for those students
-whose parents achieved university or postgraduate education. That is,
-lower-status students (measured by parental education) justify more
-inequality when they adhere more strongly to deservingnessmeritocratic
-perceptions in Chile. We depict this moderating effect in Figure 1. This
-result confirms the enlightenment thesis (see above). The same
-moderating effect is observed for students' family cultural capital: the
-relationship between effort (but not deservingness nor social talent)
-and justification of inequality becomes less positive for students with
-more than 25 books at home. Interestingly, family cultural capital also
-moderates the relationship between meritocratic beliefs at school (as
-measured by students' belief that effort is important to get good
-grades) and justification of inequality.
-
-At the school level, we also observe moderating effects of school status
-and students' meritocratic beliefs over their justification of
-inequality. Thus, in line with our Hypothesis 5, high-status schools
-justify less inequality when, on average, their students have a greater
-perception of meritocracy in Chile, as measured by the three indicators
-we used (i.e., effort, deservingness, and talent). Finally, as we stated
-in Hypothesis 7, lowhigh-achieving schools justify more inequality when
-their students have, on average, stronger meritocratic beliefs in Chile.
-In Figure 25 we depict this moderating effect of school achievement for
-the relationship between deservingness and justification of inequality.
-We did not observe these moderating effects at the school level for
-students' meritocratic beliefs in the school (i.e., the idea that effort
-or talent are important to get good grades).
-
-\subsection{Conclusion}\label{conclusion}
-
-\begin{itemize}
-\tightlist
-\item
-  Efecto ilustrador de la educación? (esfuerzo -\textgreater{}
-  redistribución)
-\end{itemize}
-
-\subsection{Appendix}\label{appendix}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.4496}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5504}}@{}}
-\toprule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-\textbf{English translation} & \textbf{Original Spanish} \\
-It is just that in Chile people who cna pay have a better education for
-their children & Es justo que en Chile las personas que puedan pagar
-tengan una mejor educación para sus hijos. \\
-It is just that in Chile people with higher incomes can have better
-pensions than people with low incomes & Es justo que en Chile las
-personas con mayores ingresos puedan tener mejores pensiones que las
-personas de ingresos más bajos \\
-It is just that in Chile people with higher incomes can access better
-health services than people with low incomes & Es justo que en Chile las
-personas con mayores ingresos puedan acceder a una mejor atención de
-salud que las personas con ingresos más bajos \\
-\end{longtable}
-
-\phantomsection\label{refs}
-\begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-acemoglu_obedience_2021}
-Acemoglu, Daron. 2021. {``Obedience in the {Labor Market} and {Social
-Mobility}: {A Socio-Economic Approach}.''} w29125. Cambridge, MA:
-National Bureau of Economic Research.
-\url{https://doi.org/10.3386/w29125}.
-
-\bibitem[\citeproctext]{ref-almas_fairness_2017}
-Almås, Ingvild, Alexander W Cappelen, Kjell G Salvanes, Erik Ø Sørensen,
-and Bertil Tungodden. 2017. {``Fairness and Family Background.''}
-\emph{Politics, Philosophy \& Economics} 16 (2): 117--31.
-\url{https://doi.org/10.1177/1470594X15618966}.
-
-\bibitem[\citeproctext]{ref-almas_fairness_2010}
-Almås, Ingvild, Alexander W. Cappelen, Erik Ø. Sørensen, and Bertil
-Tungodden. 2010. {``Fairness and the {Development} of {Inequality
-Acceptance}.''} \emph{Science} 328 (5982): 1176--78.
-\url{https://doi.org/10.1126/science.1187300}.
-
-\bibitem[\citeproctext]{ref-almas_cutthroat_2020}
-Almås, Ingvild, Alexander W. Cappelen, and Bertil Tungodden. 2020.
-{``Cutthroat {Capitalism} Versus {Cuddly Socialism}: {Are Americans More
-Meritocratic} and {Efficiency-Seeking} Than {Scandinavians}?''}
-\emph{Journal of Political Economy} 128 (5): 1753--88.
-\url{https://doi.org/10.1086/705551}.
-
-\bibitem[\citeproctext]{ref-barr_effect_2020}
-Barr, Abigail, and Luis Miller. 2020. {``The Effect of Education, Income
-Inequality and Merit on Inequality Acceptance.''} \emph{Journal of
-Economic Psychology} 80 (May): 102276.
-\url{https://doi.org/10.1016/j.joep.2020.102276}.
-
-\bibitem[\citeproctext]{ref-batruch_belief_2022}
-Batruch, Anatolia, Jolanda Jetten, Herman Van de Werfhorst, Céline
-Darnon, and Fabrizio Butera. 2022. {``Belief in {School Meritocracy} and
-the {Legitimization} of {Social} and {Income Inequality}.''}
-\emph{Social Psychological and Personality Science}, August,
-194855062211110. \url{https://doi.org/10.1177/19485506221111017}.
-
-\bibitem[\citeproctext]{ref-boltanski_new_2005}
-Boltanski, Luc, and Eve Chiapello. 2005. \emph{The New Spirit of
-Capitalism}. London ; New York: Verso.
-
-\bibitem[\citeproctext]{ref-bourdieu_reproduction_1990}
-Bourdieu, Pierre, and Jean Claude Passeron. 1990. \emph{Reproduction in
-{Education}, {Society} and {Culture}}. Second Edition. Sage Publications
-Ltd.
-
-\bibitem[\citeproctext]{ref-bucca_merit_2016}
-Bucca, Mauricio. 2016. {``Merit and Blame in Unequal Societies:
-{Explaining Latin Americans}' Beliefs about Wealth and Poverty.''}
-\emph{Research in Social Stratification and Mobility} 44 (June):
-98--112. \url{https://doi.org/10.1016/j.rssm.2016.02.005}.
-
-\bibitem[\citeproctext]{ref-castillo_legitimacy_2011}
-Castillo, Juan Carlos. 2011. {``Legitimacy of {Inequality} in a {Highly
-Unequal Context}: {Evidence} from the {Chilean Case}.''} \emph{Social
-Justice Research} 24 (4): 314--40.
-\url{https://doi.org/10.1007/s11211-011-0144-5}.
-
-\bibitem[\citeproctext]{ref-castillo_meritocracia_2019}
-Castillo, Juan Carlos, Alex Torres, Jorge Atria, and Luis Maldonado.
-2019. {``Meritocracia y Desigualdad Econ{ó}mica: {Percepciones},
-Preferencias e Implicancias.''} \emph{Revista Internacional de
-Sociolog{í}a} 77 (1): 117.
-\url{https://doi.org/10.3989/ris.2019.77.1.17.114}.
-
-\bibitem[\citeproctext]{ref-centeno_arc_2012}
-Centeno, Miguel A., and Joseph N. Cohen. 2012. {``The {Arc} of
-{Neoliberalism}.''} \emph{Annual Review of Sociology} 38 (1): 317--40.
-\url{https://doi.org/10.1146/annurev-soc-081309-150235}.
-
-\bibitem[\citeproctext]{ref-chafel_schooling_1997}
-Chafel, Judith A. 1997. {``Schooling, the {Hidden Curriculum}, and
-{Children}'s {Conceptions} of {Poverty}.''} \emph{Social Policy Report}
-11 (1): 1--28. \url{https://doi.org/10.1002/j.2379-3988.1997.tb00004.x}.
-
-\bibitem[\citeproctext]{ref-chong_mystery_2008}
-Chong, Alberto, and Hugo Ñopo. 2008. {``The {Mystery} of
-{Discrimination} in {Latin America}.''} \emph{Econom{í}a} 8 (2):
-79--107. \url{https://doi.org/10.1353/eco.0.0005}.
-
-\bibitem[\citeproctext]{ref-choudhury_social_2006}
-Choudhury, Suparna, Sarah-Jayne Blakemore, and Tony Charman. 2006.
-{``Social Cognitive Development During Adolescence.''} \emph{Social
-Cognitive and Affective Neuroscience} 1 (3): 165--74.
-\url{https://doi.org/10.1093/scan/nsl024}.
-
-\bibitem[\citeproctext]{ref-darnon_where_2018}
-Darnon, Céline, Virginie Wiederkehr, Benoît Dompnier, and Delphine
-Martinot. 2018. {``{`{Where} There Is a Will, There Is a Way'}: {Belief}
-in School Meritocracy and the Social-Class Achievement Gap.''}
-\emph{British Journal of Social Psychology} 57 (1): 250--62.
-\url{https://doi.org/10.1111/bjso.12214}.
-
-\bibitem[\citeproctext]{ref-durante_preferences_2009}
-Durante, Ruben, and Louis G. Putterman. 2009. {``Preferences for
-{Redistribution} and {Perception} of {Fairness}: {An Experimental
-Study}.''} \emph{SSRN Electronic Journal}.
-\url{https://doi.org/10.2139/ssrn.1004573}.
-
-\bibitem[\citeproctext]{ref-duru-bellat_who_2012}
-Duru-Bellat, Marie, and Elise Tenret. 2012. {``Who's for {Meritocracy}?
-{Individual} and {Contextual Variations} in the {Faith}.''}
-\emph{Comparative Education Review} 56 (2): 223--47.
-\url{https://doi.org/10.1086/661290}.
-
-\bibitem[\citeproctext]{ref-elenbaas_unfairness_2019}
-Elenbaas, Laura. 2019. {``Against Unfairness: {Young} Children's
-Judgments about Merit, Equity, and Equality.''} \emph{Journal of
-Experimental Child Psychology} 186 (October): 73--82.
-\url{https://doi.org/10.1016/j.jecp.2019.05.009}.
-
-\bibitem[\citeproctext]{ref-engelmann_children_2019}
-Engelmann, Jan M., and Michael Tomasello. 2019. {``Children's {Sense} of
-{Fairness} as {Equal Respect}.''} \emph{Trends in Cognitive Sciences} 23
-(6): 454--63. \url{https://doi.org/10.1016/j.tics.2019.03.001}.
-
-\bibitem[\citeproctext]{ref-erivwo_meritocracy_2021}
-Erivwo, Aroesiri, Elizabeth Varghese, Anthony Mathai, and Tasmia Afrin.
-2021. {``Meritocracy in the {Educational System},''} May.
-\url{https://doi.org/10.5281/ZENODO.4740695}.
-
-\bibitem[\citeproctext]{ref-evans_strong_2018}
-Evans, M. D. R., and Jonathan Kelley. 2018. {``Strong {Welfare States Do
-Not Intensify Public Support} for {Income Redistribution}, but {Even
-Reduce It} Among the {Prosperous}: {A Multilevel Analysis} of {Public
-Opinion} in 30 {Countries}.''} \emph{Societies} 8 (4): 105.
-\url{https://doi.org/10.3390/soc8040105}.
-
-\bibitem[\citeproctext]{ref-fernandez_positive_2013}
-Fernandez, J. J., and A. M. Jaime-Castillo. 2013. {``Positive or
-{Negative Policy Feedbacks}? {Explaining Popular Attitudes Towards
-Pragmatic Pension Policy Reforms}.''} \emph{European Sociological
-Review} 29 (4): 803--15. \url{https://doi.org/10.1093/esr/jcs059}.
-
-\bibitem[\citeproctext]{ref-frank_performance_2015}
-Frank, Douglas H., Klaus Wertenbroch, and William W. Maddux. 2015.
-{``Performance Pay or Redistribution? {Cultural} Differences in
-Just-World Beliefs and Preferences for Wage Inequality.''}
-\emph{Organizational Behavior and Human Decision Processes} 130
-(September): 160--70. \url{https://doi.org/10.1016/j.obhdp.2015.04.001}.
-
-\bibitem[\citeproctext]{ref-garcia-sanchez_attitudes_2020}
-García-Sánchez, Efraín, Danny Osborne, Guillermo B. Willis, and Rosa
-Rodríguez-Bailón. 2020. {``Attitudes Towards Redistribution and the
-Interplay Between Perceptions and Beliefs about Inequality.''}
-\emph{British Journal of Social Psychology} 59 (1): 111--36.
-\url{https://doi.org/10.1111/bjso.12326}.
-
-\bibitem[\citeproctext]{ref-goldthorpe_myth_2003}
-Goldthorpe, John. 2003. {``The Myth of Education-Based Meritocracy.''}
-\emph{New Economy} 10 (4): 234--39.
-\url{https://doi.org/10.1046/j.1468-0041.2003.00324.x}.
-
-\bibitem[\citeproctext]{ref-goudeau_hidden_2017}
-Goudeau, Sébastien, and Jean-Claude Croizet. 2017. {``Hidden
-{Advantages} and {Disadvantages} of {Social Class}: {How Classroom
-Settings Reproduce Social Inequality} by {Staging Unfair Comparison}.''}
-\emph{Psychological Science} 28 (2): 162--70.
-\url{https://doi.org/10.1177/0956797616676600}.
-
-\bibitem[\citeproctext]{ref-greitemeyer_subjective_2016}
-Greitemeyer, Tobias, and Christina Sagioglou. 2016. {``Subjective
-Socioeconomic Status Causes Aggression: {A} Test of the Theory of Social
-Deprivation.''} \emph{Journal of Personality and Social Psychology} 111
-(2): 178--94. \url{https://doi.org/10.1037/pspi0000058}.
-
-\bibitem[\citeproctext]{ref-harvey_breve_2015}
-Harvey, David. 2015. \emph{{Breve historia del neoliberalismo}}. Madrid
-(Espa{ñ}a): Ediciones Akal.
-
-\bibitem[\citeproctext]{ref-homan_being_2017}
-Homan, Patricia, Lauren Valentino, and Emi Weed. 2017. {``Being and
-{Becoming Poor}: {How Cultural Schemas Shape Beliefs About Poverty}.''}
-\emph{Social Forces} 95 (3): 1023--48.
-\url{https://doi.org/10.1093/sf/sox007}.
-
-\bibitem[\citeproctext]{ref-huppert_development_2019}
-Huppert, Elizabeth, Jason M. Cowell, Yawei Cheng, Carlos
-Contreras-Ibáñez, Natalia Gomez-Sicard, Maria Luz Gonzalez-Gadea, David
-Huepe, et al. 2019. {``The Development of Children's Preferences for
-Equality and Equity Across 13 Individualistic and Collectivist
-Cultures.''} \emph{Developmental Science} 22 (2): e12729.
-\url{https://doi.org/10.1111/desc.12729}.
-
-\bibitem[\citeproctext]{ref-hvidberg_social_2023}
-Hvidberg, Kristoffer B, Claus T Kreiner, and Stefanie Stantcheva. 2023.
-{``Social {Positions} and {Fairness Views} on {Inequality}.''}
-\emph{Review of Economic Studies} 90 (6): 3083--3118.
-\url{https://doi.org/10.1093/restud/rdad019}.
-
-\bibitem[\citeproctext]{ref-iacoviello_collectivism_2019}
-Iacoviello, Vincenzo, and Fabio Lorenzi-Cioldi. 2019. {``Collectivism
-and {Individualism} in {Status Hierarchies}: {Socialization} and {Social
-Identity Explanations}.''} \emph{International Review of Social
-Psychology} 32 (1): 15. \url{https://doi.org/10.5334/irsp.285}.
-
-\bibitem[\citeproctext]{ref-jack_no_2016}
-Jack, Anthony Abraham. 2016. {``({No}) {Harm} in {Asking}: {Class},
-{Acquired Cultural Capital}, and {Academic Engagement} at an {Elite
-University}.''} \emph{Sociology of Education} 89 (1): 1--19.
-\url{https://doi.org/10.1177/0038040715614913}.
-
-\bibitem[\citeproctext]{ref-jasso_how_1999}
-Jasso, Guillermina. 1999. {``How {Much Injustice} Is {There} in the
-{World}? {Two New Justice Indexes}.''} \emph{American Sociological
-Review} 64 (1): 133--68.
-\url{https://doi.org/10.1177/000312249906400110}.
-
-\bibitem[\citeproctext]{ref-jensen_worlds_2008}
-Jensen, Carsten. 2008. {``Worlds of Welfare Services and Transfers.''}
-\emph{Journal of European Social Policy} 18 (2): 151--62.
-\url{https://doi.org/10.1177/0958928707087591}.
-
-\bibitem[\citeproctext]{ref-jonsson_institutional_2015}
-Jonsson, Anna-Carin, and Dennis Beach. 2015. {``Institutional
-Discrimination: {Stereotypes} and Social Reproduction of {`Class'} in
-the {Swedish} Upper-Secondary School.''} \emph{Social Psychology of
-Education} 18 (4): 703--17.
-\url{https://doi.org/10.1007/s11218-014-9279-1}.
-
-\bibitem[\citeproctext]{ref-jost_attitudinal_2000}
-Jost, John T., and Diana Burgess. 2000. {``Attitudinal {Ambivalence} and
-the {Conflict} Between {Group} and {System Justification Motives} in
-{Low Status Groups}.''} \emph{Personality and Social Psychology
-Bulletin} 26 (3): 293--305.
-\url{https://doi.org/10.1177/0146167200265003}.
-
-\bibitem[\citeproctext]{ref-khan_privilege_2011}
-Khan, Shamus. 2011. \emph{Privilege: {The Making} of an {Adolescent
-Elite} at {St}. {Paul}'s {School}}. Princeton, N.J.: Princeton
-University Press.
-
-\bibitem[\citeproctext]{ref-kluegel_beliefs_1987}
-Kluegel, James R., and Eliot R. Smith. 1987. \emph{Beliefs about
-Inequality: {Americans}' Views of What Is and What Ought to Be}. London
-New York: Routledge, Taylor \& Francis Group.
-
-\bibitem[\citeproctext]{ref-kohn_social_1963}
-Kohn, Melvin L. 1963. {``Social {Class} and {Parent-Child
-Relationships}: {An Interpretation}.''} \emph{American Journal of
-Sociology} 68 (4): 471--80. \url{https://doi.org/10.1086/223403}.
-
-\bibitem[\citeproctext]{ref-kohn_class_1969}
-Kohn, Melvin L., and Carmi Schooler. 1969. {``Class, {Occupation}, and
-{Orientation}.''} \emph{American Sociological Review} 34 (5): 659.
-\url{https://doi.org/10.2307/2092303}.
-
-\bibitem[\citeproctext]{ref-kosse_prosociality_2020}
-Kosse, Fabian, and Michela M. Tincani. 2020. {``Prosociality Predicts
-Labor Market Success Around the World.''} \emph{Nature Communications}
-11 (1): 5298. \url{https://doi.org/10.1038/s41467-020-19007-1}.
-
-\bibitem[\citeproctext]{ref-lampert_meritocratic_2013}
-Lampert, Khen. 2013. \emph{Meritocratic {Education} and {Social
-Worthlessness}}. London: Palgrave Macmillan UK.
-\url{https://doi.org/10.1057/9781137324894}.
-
-\bibitem[\citeproctext]{ref-lane_market_1986}
-Lane, Robert E. 1986. {``Market {Justice}, {Political Justice}.''}
-\emph{American Political Science Review} 80 (2): 383--402.
-\url{https://doi.org/10.2307/1958264}.
-
-\bibitem[\citeproctext]{ref-lindh_public_2015}
-Lindh, Arvid. 2015. {``Public {Opinion} Against {Markets}? {Attitudes}
-Towards {Market Distribution} of {Social Services} -- {A Comparison} of
-17 {Countries}.''} \emph{Social Policy \& Administration} 49 (7):
-887--910. \url{https://doi.org/10.1111/spol.12105}.
-
-\bibitem[\citeproctext]{ref-mcauliffe_developmental_2017}
-McAuliffe, Katherine, Peter R. Blake, Nikolaus Steinbeis, and Felix
-Warneken. 2017. {``The Developmental Foundations of Human Fairness.''}
-\emph{Nature Human Behaviour} 1 (2): 0042.
-\url{https://doi.org/10.1038/s41562-016-0042}.
-
-\bibitem[\citeproctext]{ref-mijs_stratified_2016}
-Mijs, Jonathan. 2016. {``Stratified {Failure}: {Educational
-Stratification} and {Students}' {Attributions} of {Their Mathematics
-Performance} in 24 {Countries}.''} \emph{Sociology of Education} 89 (2):
-137--53. \url{https://doi.org/10.1177/0038040716636434}.
-
-\bibitem[\citeproctext]{ref-mijs_paradox_2021}
----------. 2021. {``The Paradox of Inequality: Income Inequality and
-Belief in Meritocracy Go Hand in Hand.''} \emph{Socio-Economic Review}
-19 (1): 7--35. \url{https://doi.org/10.1093/ser/mwy051}.
-
-\bibitem[\citeproctext]{ref-mikus_children_2020}
-Mikus, Karoline, Nicole Tieben, and Pia S. Schober. 2020. {``Children's
-Conversion of Cultural Capital into Educational Success: The Symbolic
-and Skill-Generating Functions of Cultural Capital.''} \emph{British
-Journal of Sociology of Education} 41 (2): 197--217.
-\url{https://doi.org/10.1080/01425692.2019.1677454}.
-
-\bibitem[\citeproctext]{ref-mishra_subjective_2015}
-Mishra, Sandeep, and R. Nicholas Carleton. 2015. {``Subjective Relative
-Deprivation Is Associated with Poorer Physical and Mental Health.''}
-\emph{Social Science \& Medicine} 147 (December): 144--49.
-\url{https://doi.org/10.1016/j.socscimed.2015.10.030}.
-
-\bibitem[\citeproctext]{ref-molyneux_neoliberal_2008}
-Molyneux, Maxine. 2008. {``The {`{Neoliberal Turn}'} and the {New Social
-Policy} in {Latin America}: {How Neoliberal}, {How New}?''}
-\emph{Development and Change} 39 (5): 775--97.
-\url{https://doi.org/10.1111/j.1467-7660.2008.00505.x}.
-
-\bibitem[\citeproctext]{ref-morris_representing_2022}
-Morris, Jennifer, John Reilly, Sergey Paltsev, Andrei Sokolov, and
-Kenneth Cox. 2022. {``Representing {Socio}-{Economic Uncertainty} in
-{Human System Models}.''} \emph{Earth's Future} 10 (4): e2021EF002239.
-\url{https://doi.org/10.1029/2021EF002239}.
-
-\bibitem[\citeproctext]{ref-pierson_increasing_2000}
-Pierson, Paul. 2000. {``Increasing {Returns}, {Path Dependence}, and the
-{Study} of {Politics}.''} \emph{American Political Science Review} 94
-(2): 251--67. \url{https://doi.org/10.2307/2586011}.
-
-\bibitem[\citeproctext]{ref-power_deprivationprotest_2018}
-Power, Séamus A. 2018. {``The {Deprivation-Protest Paradox}: {How} the
-{Perception} of {Unfair Economic Inequality Leads} to {Civic Unrest}.''}
-\emph{Current Anthropology} 59 (6): 765--89.
-\url{https://doi.org/10.1086/700679}.
-
-\bibitem[\citeproctext]{ref-power_relative_2020}
-Power, Séamus A, Thomas Madsen, and Thomas A Morton. 2020. {``Relative
-Deprivation and Revolt: Current and Future Directions.''} \emph{Current
-Opinion in Psychology} 35 (October): 119--24.
-\url{https://doi.org/10.1016/j.copsyc.2020.06.010}.
-
-\bibitem[\citeproctext]{ref-resh_sense_2010}
-Resh, Nura. 2010. {``Sense of Justice about Grades in School: Is It
-Stratified Like Academic Achievement?''} \emph{Social Psychology of
-Education} 13 (3): 313--29.
-\url{https://doi.org/10.1007/s11218-010-9117-z}.
-
-\bibitem[\citeproctext]{ref-resh_sense_2014}
-Resh, Nura, and Clara Sabbagh. 2014. {``Sense of Justice in School and
-Civic Attitudes.''} \emph{Social Psychology of Education} 17 (1):
-51--72. \url{https://doi.org/10.1007/s11218-013-9240-8}.
-
-\bibitem[\citeproctext]{ref-resh_sense_2017}
----------. 2017. {``Sense of Justice in School and Civic Behavior.''}
-\emph{Social Psychology of Education} 20 (2): 387--409.
-\url{https://doi.org/10.1007/s11218-017-9375-0}.
-
-\bibitem[\citeproctext]{ref-salamon_marketization_1993}
-Salamon, Lester M. 1993. {``The {Marketization} of {Welfare}: {Changing
-Nonprofit} and {For-Profit Roles} in the {American Welfare State}.''}
-\emph{Social Service Review} 67 (1): 16--39.
-\url{https://doi.org/10.1086/603963}.
-
-\bibitem[\citeproctext]{ref-salgado_inequality_2023}
-Salgado, Mauricio, and Javier Castillo. 2023. {``Inequality and
-{Stratification} in {Latin America}.''} In \emph{The {Oxford Handbook}
-of {Social Stratification}}, edited by Markus Gangl, Lucinda Platt,
-Javier G. Polavieja, and Herman G. van de Werfhorst, 0. 2023: Oxford
-University Press.
-
-\bibitem[\citeproctext]{ref-sandel_tyranny_2020}
-Sandel, Michael J. 2020. \emph{The Tyranny of Merit: {What}'s Become of
-the Common Good?} First edition. New York: {Farrar, Straus and Giroux}.
-
-\bibitem[\citeproctext]{ref-schunk_fairness_2023}
-Schunk, Daniel, and Isabell Zipperle. 2023. {``Fairness and Inequality
-Acceptance in Children and Adolescents: {A} Survey on Behaviors in
-Economic Experiments''} 37 (5): 1715--42.
-\url{https://doi.org/10.1111/joes.12553}.
-
-\bibitem[\citeproctext]{ref-shariff_income_2016}
-Shariff, Azim F., Dylan Wiwad, and Lara B. Aknin. 2016. {``Income
-{Mobility Breeds Tolerance} for {Income Inequality}: {Cross-National}
-and {Experimental Evidence}.''} \emph{Perspectives on Psychological
-Science} 11 (3): 373--80.
-\url{https://doi.org/10.1177/1745691616635596}.
-
-\bibitem[\citeproctext]{ref-sigelman_development_1991}
-Sigelman, Carol K., and Kara A. Waitzman. 1991. {``The {Development} of
-{Distributive Justice Orientations}: {Contextual Influences} on
-{Children}'s {Resource Allocations}.''} \emph{Child Development} 62 (6):
-1367. \url{https://doi.org/10.2307/1130812}.
-
-\bibitem[\citeproctext]{ref-sivesind_changing_2017}
-Sivesind, Karl Henrik. 2017. {``The {Changing Roles} of {For-Profit} and
-{Nonprofit Welfare Provision} in {Norway}, {Sweden}, and {Denmark}.''}
-In \emph{Promoting {Active Citizenship}}, edited by Karl Henrik Sivesind
-and Jo Saglie, 33--74. Cham: Springer International Publishing.
-\url{https://doi.org/10.1007/978-3-319-55381-8_2}.
-
-\bibitem[\citeproctext]{ref-smith_relative_2012}
-Smith, Heather J., Thomas F. Pettigrew, Gina M. Pippin, and Silvana
-Bialosiewicz. 2012. {``Relative {Deprivation}: {A Theoretical} and
-{Meta-Analytic Review}.''} \emph{Personality and Social Psychology
-Review} 16 (3): 203--32. \url{https://doi.org/10.1177/1088868311430825}.
-
-\bibitem[\citeproctext]{ref-soler-martinez_concerns_2023}
-Soler-Martínez, Francisco Miguel, Efraín García-Sánchez, and Guillermo
-B. Willis. 2023. {``Concerns {About Inequality} in {Health}, {Education}
-and {Income Jointly Predict Collective Actions}.''} \emph{Revista
-Latinoamericana de Psicolog{í}a} 55: 99.
-\url{https://doi.org/10.14349/rlp.2023.v55.12}.
-
-\bibitem[\citeproctext]{ref-stoy_worlds_2014}
-Stoy, Volquart. 2014. {``Worlds of {Welfare Services}: {From Discovery}
-to {Exploration}.''} \emph{Social Policy \& Administration} 48 (3):
-343--60. \url{https://doi.org/10.1111/spol.12006}.
-
-\bibitem[\citeproctext]{ref-streeck_citizens_2012}
-Streeck, Wolfgang. 2012. {``Citizens as {Customers}: {Considerations} on
-the {New Politics} of {Consumption}.''} \emph{New Left Review} 76.
-
-\bibitem[\citeproctext]{ref-torche_intergenerational_2014}
-Torche, Florencia. 2014. {``Intergenerational {Mobility} and
-{Inequality}: {The Latin American Case}.''} \emph{Annual Review of
-Sociology} 40 (1): 619--42.
-\url{https://doi.org/10.1146/annurev-soc-071811-145521}.
-
-\bibitem[\citeproctext]{ref-vonhippel_test_2019}
-Von Hippel, Paul, and Caitlin Hamrock. 2019. {``Do {Test Score Gaps
-Grow} Before, During, or Between the {School Years}? {Measurement
-Artifacts} and {What We Can Know} in {Spite} of {Them}.''}
-\emph{Sociological Science} 6: 43--80.
-\url{https://doi.org/10.15195/v6.a3}.
-
-\bibitem[\citeproctext]{ref-weaver_paths_2010}
-Weaver, Kent. 2010. {``Paths and {Forks} or {Chutes} and {Ladders}?:
-{Negative Feedbacks} and {Policy Regime Change}.''} \emph{Journal of
-Public Policy} 30 (2): 137--62.
-\url{https://doi.org/10.1017/S0143814X10000061}.
-
-\bibitem[\citeproctext]{ref-wegener_dominant_1995}
-Wegener, Bernd, and Stefan Liebig. 1995. {``Dominant {Ideologies} and
-the {Variation} of {Distributive Justice Norms}: {A Comparison} of
-{East} and {West Germany}, and the {United States}.''} In \emph{Social
-{Justice} and {Political Change}: {Public Opinion} in {Capitalist} and
-{Post-Communist States}}, edited by James R. Kluegel, David S. Mason,
-and Bernd Wegener, 239--59. New York: Walter de Gruyter.
-
-\bibitem[\citeproctext]{ref-wiederkehr_belief_2015}
-Wiederkehr, Virginie, Virginie Bonnot, Silvia Krauth-Gruber, and Céline
-Darnon. 2015. {``Belief in School Meritocracy as a System-Justifying
-Tool for Low Status Students.''} \emph{Frontiers in Psychology} 6.
-
-\bibitem[\citeproctext]{ref-wynn_not_2018}
-Wynn, Karen, Paul Bloom, Ashley Jordan, Julia Marshall, and Mark
-Sheskin. 2018. {``Not {Noble Savages After All}: {Limits} to {Early
-Altruism}.''} \emph{Current Directions in Psychological Science} 27 (1):
-3--8. \url{https://doi.org/10.1177/0963721417734875}.
-
-\bibitem[\citeproctext]{ref-young_rise_1958}
-Young, Michael. 1958. \emph{The Rise of the Meritocracy}. New Brunswick,
-N.J., U.S.A: Transaction Publishers.
-
-\end{CSLReferences}
-
-\end{document}
+\section{Conclusion}\label{conclusion}
+
+Our research aimed to examine the relationship between the perception of
+meritocracy and the justification of market justice preferences among
+eighth-grade students in Chile. From our knowledge this is the first
+study addressing market justice preferences at school level, finding
+that students exhibit large preferences for market justice (about one
+third agree or strongly agree with them), a share that is quite larger
+when compared with evidence in adult population. Perception of
+meritocracy is also high, particularly when it comes to the reward of
+effort, with an striking eighty six percent of agreement. Regarding
+these descriptive indicators, a a first message coming from this study
+is that there is something particular in school-age when it comes to
+perceptions and preferences about inequality. This can relate to
+development aspects
+\citep[\citet{rizzo_children_2020}]{kim_socioeconomic_2020} but it could
+be something specific to the Chilean case, which requires further
+comparative as well as longitudinal research in the area. In any case,
+teaching about the origins and characteristics of social inequality is
+not part of the chilean school curriculum, calling for a more dedicated
+consideration of these topics in areas such as citizensip education as
+well as in history and social science education.
+
+The core of this paper was the association between meritocratic
+perceptions and preferences for market justice. In general, we find that
+those who perceive that effort and talent/intelligence are rewarded, are
+more willing to agree with that richer individuals can have better
+health, education, and pensions. This is not the first time that
+meritocratic perceptions (and also beliefs) have been related with the
+legitimation of social inequalities
+\citep{mijs_paradox_2019, darnon_where_2018}, but the consideration of
+school level population as well as market justice preferences allows
+adding evidence regarding the role of meritocracy regarding inequality
+beliefs: legitimation of market criteria in social services seems to
+begin at school age, and what students think of meritocracy is linked to
+it. Although there are still several open questions, as for instance
+differences between meritocratic perceptions in school and in the
+society at large, this raises up the relevance of what is done (or not)
+in schools in order to challenge the meritocratic ideals.
+
+The consideration of the school context opened some caveats in this
+research. We found contrary evidence for the association between status,
+school performance, and market justice preferences: high status and high
+achievement at the school level are associated with less market justice
+preferences. This is opposed to the rational-interest argument that
+suggests that those doing better - both economically and academically-
+would demand less redistribution \citep[\citet{runst_does_2018},
+\citet{bullock_education_2021}]{gonthier_parallel_2017} and therefore
+are expected to challenge market mechanisms. A possible explanation is
+that education contributes to forming a critical view of markets in
+redistribution, nevertheless, further research in this area is needed to
+assert the specific mechanisms that might be playing a role. All in all,
+from this study we know that school contexts are not trivial in this
+regard, calling for larger attention to school-age research on the
+formation of inequality preferences and beliefs.
+
+%\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+The following statements should be used ``Conceptualization, X.X.
+and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
+and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
+All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
+Authorship must be limited to those who have contributed substantially to the work~reported.}
+
+\authorcontributions{Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: \getfirst{Juan Carlos}.\getfirst{Castillo}.:   Conceptualization, Supervision, Methodology, Writing
+- review and
+editing; \getfirst{Mauricio}.\getfirst{Salgado}.:   Conceptualization, Methodology, Writing
+- review and
+editing; \getfirst{Kevin}.\getfirst{Carrasco}.:   Methodology, formal
+analysis, software, visualization, Writing - review and
+editing; \getfirst{Andreas}.\getfirst{Laffert}.:   Methodology, formal
+analysis, software, visualization, Writing - review and editing}
+
+%\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
+Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
+
+\funding{The authors acknowledge the support of ANID/FONDECYT grant
+1210847 and the Center for Conflict and Social Cohesion Studies (COES)
+grant ANID/FONDAP 1513009.}
+
+%\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study.
+You might choose to exclude this statement if the study did not require ethical approval.
+Please note that the Editorial Office might ask you for further information.
+Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans.
+OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals.
+OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.}
+
+%\informedconsent{Any research article describing a study involving humans should contain this statement.
+Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.
+You might also choose to exclude this statement if the study did not involve humans.
+Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves).
+Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}
+
+%\dataavailability{We encourage all authors of articles published in MDPI journals to share their research data.
+In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study.
+Where no new data were created, or where data is unavailable due to privacy or ethical restrictions, a statement is still required.
+Suggested Data Availability Statements are available in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}.}
+
+\dataavailability{The original data presented in the study are openly
+available in
+\url{https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/}}
+
+% Only for journal Nursing Reports
+%\publicinvolvement{Please describe how the public (patients, consumers, carers) were involved in the research.
+Consider reporting against the GRIPP2 (Guidance for Reporting Involvement of Patients and the Public) checklist.
+If the public were not involved in any aspect of the research add: ``No public involvement in any aspect of this research''.}
+
+% Only for journal Nursing Reports
+%\guidelinesstandards{Please add a statement indicating which reporting guideline was used when drafting the report.
+For example, ``This manuscript was drafted against the XXX (the full name of reporting guidelines and citation) for XXX (type of research) research''.
+A complete list of reporting guidelines can be accessed via the equator network: \url{https://www.equator-network.org/}.}
+
+%\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
+This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
+
+%\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflicts of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
+Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript; or in the decision to publish the results must be declared in this section.
+If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the results''.}
+
+\conflictsofinterest{The authors declare no conflict of interest.}
+
+\begin{adjustwidth}{-\extralength}{0cm}
+%\printendnotes[custom] % Un-comment to print a list of endnotes
+
+\reftitle{References}
+
+  \bibliography{input/bib/just-ineq-merit.bib}
+
+\PublishersNote{}
+\end{adjustwidth}\end{document}

### 2024-08-06 07-supplementary-material.qmd
Archivos submit
Push formato
Co-Authored-By: Kevin Carrasco <67479225+Kevin-carrasco@users.noreply.github.com>

@@
----
-author: "Equipo EDUMER"
-bibliography: "input/bib/just-ineq-merit.bib"
-csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
----
-
-# Supplementary Material
+# Supplementary material
 
@@
 
-# Translated items for market justice preferences
+## Translated items for market justice preferences
 
@@
 
-# Complete multilevel models
+## Complete multilevel models
 
@@
 #| tbl-cap: "Linear mixed-effects models for meritocracy perception and market justice preferences"
-#| label: tbl-lineal-reg
+#| label: tbl-lineal-reg-complete
 ccoef <- list(

### 2024-08-06 paper-blinded.qmd
Push formato
Co-Authored-By: Kevin Carrasco <67479225+Kevin-carrasco@users.noreply.github.com>

@@
 format:
-  hikmah-pdf: 
-    indent: false
-  hikmah-manuscript-pdf: 
-    indent: false
+  mdpi-pdf:
+    keep-tex: true
+mdpi_journal: behavsci
+article_type: article
+article_status: submit
+#author:
+#  - name: Juan Carlos Castillo
+#    corresponding: true
+#    orcid: 0000-0003-1265-7854
+#    email: juancastillov@uchile.cl
+#    role:
+#      - Conceptualization
+#      - Supervision
+#      - Methodology
+#      - Writing - review and editing
+#    affiliations:
+#      - name: "Universidad de Chile"
+#        department: Department of Sociology
+#        id: 1
+#        country: Chile
+#  - name: Mauricio Salgado
+#    orcid: 0000-0002-8108-6154
+#    role:
+#      - Conceptualization
+#      - Methodology
+#      - Writing - review and editing    
+#    affiliations:
+#      - name: Centro de Estudios Públicos
+#        id: 2
+#        department: CEP
+#        country: Chile
+#  - name: Kevin Carrasco
+#    orcid: 0000-0002-9348-8361
+#    role:
+#      - Methodology
+#      - formal analysis
+#      - software
+#      - visualization
+#      - Writing - review and editing
+#    affiliations:
+#      - name: Center for Social Conflict and Cohesion Studies
+#        id: 3
+#        department: COES
+#        country: Chile
+#  - name: Andreas Laffert
+#    orcid: 0000-0002-9008-2454
+#    role:
+#      - Methodology
+#      - formal analysis
+#      - software
+#      - visualization
+#      - Writing - review and editing  
+#    affiliations:
+#      - name: Universidad Católica de Chile
+#        id: 4
+#        department: Institute of Sociology
+#        country: Chile
+fundingstatement: | 
+  The authors acknowledge the support of ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) grant ANID/FONDAP 1513009.
+dataavailability: |
+  The original data presented in the study are openly available in <https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/>
+  Access to the data, code, and analysis presented in the study are openly available in <https://github.com/educacion-meritocracia/market-justice-preferences-school>
+conflictsofinterest: |
+  The authors declare no conflict of interest.
 abstract: |
@@
 These results suggest that meritocratic beliefs and market justice preferences begin early and are influenced by the school environment.
-keywords: 
-  - market justice 
-  - meritocracy
-  - socialization
-  - family
-  - schools
-  - Chile
+keywords: [market justice, meritocracy, socialization, family, schools, Chile]
 bibliography: "input/bib/just-ineq-merit.bib"
-csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
-link-citations: true
-linestretch: 1.15       
-mainfont: Times New Roman
-fontsize: 12pt          
-geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
-papersize: letter
-colorlinks: true
-fig-height: 4
-fig-width: 7.5
-header-includes:
-  - \usepackage{threeparttable}
 ---
@@
 {{< include 06-conclusion.qmd >}}
+
+{{< include 07-supplementary-material.qmd >}}

### 2024-08-07 cover_letter.qmd
Update submit mdpi FINAL
@juancarloscastillo

@@
 
+## Acknowledgments
+
+The authors acknowledge the support of ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) grant ANID/FONDAP 1513009.
+
+## Authors' Contributions
+
+Author roles were classified using the Contributor Role Taxonomy [CRediT](https://credit.niso.org/) as follows: Juan Carlos Castillo: Conceptualization, Supervision, Methodology, Writing - review and editing; Mauricio Salgado: Conceptualization, Methodology, Writing - review and editing; Kevin Carrasco: Methodology, formal analysis, software, visualization, Writing - review and editing; Andreas Laffert: Methodology, formal analysis, software, visualization, Writing - review and editing.
+
+## Conflict of Interest
+
+The authors declare no conflict of interest.
+
+## Funding
+
+This research received funding from ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) through ANID/FONDAP grant 1513009.
+
+## Data Avaibility Statement
+
+The original data presented in the study are openly available in <https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/>
+
+Access to the data, code, and analysis presented in the study are openly available in <https://github.com/educacion-meritocracia/market-justice-preferences-school>.
+
 [^*]: Corresponding author (juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) - Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile).

### 2024-08-07 paper-blinded.qmd
Update submit mdpi FINAL
@juancarloscastillo

--- 2024-07-22 paper-blinded.qmd
@@
     indent: false
+    keep-tex: true
   hikmah-manuscript-pdf: 
     indent: false
+    keep-tex: true
 abstract: |
@@
 ---
+{{< pagebreak >}}
 
@@
 {{< include 06-conclusion.qmd >}}
+
+{{< pagebreak >}}
+
+{{< include 07-supplementary-material.qmd >}}

### 2024-08-07 submission/MDPI/cover_letter.tex
Update submit mdpi FINAL
@juancarloscastillo

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  letterpaper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% -----------------------
% CUSTOM PREAMBLE STUFF
% -----------------------

% -----------------
% Typography tweaks
% -----------------
% Indent size
\setlength{\parindent}{1pc}  % 1p0

% Fix widows and orphans
\usepackage[all,defaultlines=2]{nowidow}

% List things
\usepackage{enumitem}
% Same document-level indentation for ordered and ordered lists
\setlist[1]{labelindent=\parindent}
\setlist[itemize]{leftmargin=*}
\setlist[enumerate]{leftmargin=*}

% Wrap definition list terms
% https://tex.stackexchange.com/a/9763/11851
\setlist[description]{style=unboxed}

% For better TOCs
\usepackage{tocloft}

% Remove left margin in lists inside longtables
% https://tex.stackexchange.com/a/378190/11851
\AtBeginEnvironment{longtable}{\setlist[itemize]{nosep, wide=0pt, leftmargin=*, before=\vspace*{-\baselineskip}, after=\vspace*{-\baselineskip}}}

% Allow for /singlespacing and /doublespacing
\usepackage{setspace}

% -----------------
% Title block stuff
% -----------------

% Abstract
\usepackage[overload]{textcase}
\usepackage[runin]{abstract}
\renewcommand{\abstractnamefont}{\sffamily\footnotesize\bfseries\MakeUppercase}
\renewcommand{\abstracttextfont}{\sffamily\small}
\setlength{\absleftindent}{\parindent * 2}
\setlength{\absrightindent}{\parindent * 2}
\abslabeldelim{\quad}
\setlength{\abstitleskip}{-\parindent}

% Keywords
\newenvironment{keywords}
{\vskip -3em \hspace{\parindent}\small\sffamily{\sffamily\footnotesize\bfseries\MakeUppercase{Keywords}}\quad}
{\vskip 3em}

% Title
\usepackage{titling}
\setlength{\droptitle}{3em}
\pretitle{\par\vskip 5em \begin{flushleft}\LARGE\sffamily\bfseries}
\posttitle{\par\end{flushleft}\vskip 0.75em}

% Authors
%
% PHEW this is complicated for a number of reasons!
%
% When using \and with multiple authors, the article class in LaTeX wraps each 
% author block in a tabluar environment with a hardcoded center alignment. It's 
% possible to use \preauthor{} to start tabulars with a left alignment {l}, but 
% that only applies to the first author because the others all use \and with the 
% hardcoded {c}. But we can override the \and command and add our own {l}
%
% (See https://github.com/rstudio/rmarkdown/issues/1716#issuecomment-560601691 
% for an example of redefining \and to just be \\)
%
% That's all great, except tabulars have some amount of default horizontal 
% padding, which makes left-aligned author blocks not actuall get fully 
% left-aligned on the page. We can set the horizontal padding for the column to 
% 0, but it requires some wonky syntax: {@{\hspace{0em}}l@{}}
\renewcommand{\and}{\end{tabular} \hskip 3em \begin{tabular}[t]{@{\hspace{0em}}l@{}}}
\preauthor{\begin{flushleft}
           \lineskip 1.5em 
           \begin{tabular}[t]{@{\hspace{0em}}l@{}}}
\postauthor{\end{tabular}\par\end{flushleft}}

% Omit the date since the \published command does that
\predate{}
\postdate{}

% Command for a note at the top of the first page describing the publication
% status of the paper.
\newcommand{\published}[1]{%
   \gdef\puB{#1}}
   \newcommand{\puB}{}
   \renewcommand{\maketitlehooka}{%
       \par\noindent\footnotesize\sffamily \puB}

% ------------------
% Section headings
% ------------------
\usepackage{titlesec}
\titleformat*{\section}{\Large\sffamily\bfseries\raggedright}
\titleformat*{\subsection}{\large\sffamily\bfseries\raggedright}
\titleformat*{\subsubsection}{\normalsize\sffamily\bfseries\raggedright}
\titleformat*{\paragraph}{\small\sffamily\bfseries\raggedright}

% \titlespacing{<command>}{<left>}{<before-sep>}{<after-sep>}
% Starred version removes indentation in following paragraph
\titlespacing*{\section}{0em}{2em}{0.1em}
\titlespacing*{\subsection}{0em}{1.25em}{0.1em}
\titlespacing*{\subsubsection}{0em}{0.75em}{0em}

% -----------
% Footnotes
% -----------
% NB: footmisc has to come after setspace and biblatex because of conflicts
\usepackage[bottom, flushmargin]{footmisc}
\renewcommand*{\footnotelayout}{\footnotesize}

\addtolength{\skip\footins}{10pt}    % vertical space between rule and main text
\setlength{\footnotesep}{5pt}  % vertical space between footnotes

% ----------
% Captions
% ----------
\usepackage[font={small,sf}, labelfont={small,sf,bf}]{caption}

% --------
% Macros
% --------
% pandoc will not convert text within \begin{} XXX \end{} to Markdown and will
% treat it as regular TeX. Because of this, it's impossible to do stuff like
% this:

% \begin{landscape}
%
% | One | Two   |
% |-----+-------|
% | my  | table |
% | is  | nice  |
%
% \end{landscape}
%
% Since it'll render like: | One | Two | |—–+——-| | my | table | | is | nice |
% 
% BUT, from this http://stackoverflow.com/a/41945462/120898 we can get around
% this by creating new commands for \begin and \end, like this:
\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

% \blandscape
%
% | One | Two   |
% |-----+-------|
% | my  | table |
% | is  | nice  |
%
% \elandscape

% Same thing, but for generic groups
% But can't use \bgroup and \egroup because those are built-in TeX things
\newcommand{\stgroup}{\begingroup}
\newcommand{\fingroup}{\endgroup}

% ---------------------------
% END CUSTOM PREAMBLE STUFF
% ---------------------------
\usepackage{threeparttable}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The Socialization of Meritocracy and Market Justice Preferences at School},
  colorlinks=true,
  linkcolor={DarkSlateBlue},
  filecolor={Maroon},
  citecolor={DarkSlateBlue},
  urlcolor={DarkSlateBlue},
  pdfcreator={LaTeX via pandoc}}

% -----------------------
% END-OF-PREAMBLE STUFF
% -----------------------

% ---------------------- 
% Title block elements
% ---------------------- 
\usepackage{orcidlink}  % Create automatic ORCID icons/links

\title{The Socialization of Meritocracy and Market Justice Preferences
at School}

\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\vskip 0.25em \large #1 \par}}{}{}
}
\makeatother
\subtitle{Cover Letter}

\author{
}

\date{}

% Typeset URLs in the same font as their parent environment
%
% This has to come at the end of the preamble, after any biblatex stuff because 
% some biblatex styles (like APA) define their own \urlstyle{}
\usepackage{url}
\urlstyle{same}

% ---------------------------
% END END-OF-PREAMBLE STUFF
% ---------------------------
\begin{document}
% ---------------
% TITLE SECTION
% ---------------
\published{\textbf{}}

\maketitle

% -------------------
% END TITLE SECTION
% -------------------

\setstretch{1.15}
Previous research has shown that schools justify student performance
differences based on meritocratic ideals. However, the link between
school meritocracy and students' justification of inequalities has
received little attention. This paper examines how school and family
socialization influence students' beliefs about meritocracy and
market-based justifications for access to health, pensions, and
education in Chile. Using data from the First Study of Civic Education
in Chile, we employ a multilevel modeling strategy to analyze 5,047
students and parents nested within 231 schools. Our findings reveal that
a significant proportion of Chilean students agree with market justice
principles, at a higher rate than adults. Most students endorse
meritocratic views, particularly the reward of effort. Multilevel
estimations show strong associations between meritocratic beliefs and
market justice preferences: students who believe in the reward of effort
and talent are more likely to justify inequalities based on financial
capacity. At the school level, market justice preferences are higher in
high-status schools but lower in schools with better achievement.
Additionally, the conditional influence of meritocracy beliefs weakens
in schools with higher socioeconomic status and performance. These
results suggest that meritocratic beliefs and market justice preferences
begin early and are influenced by the school environment.

\subsection{Authors Identification}\label{authors-identification}

Juan Carlos Castillo, Department of Sociology, Universidad de Chile,
Santiago de Chile\footnote{Corresponding author (juancastillov@uchile.cl
  -
  \href{https://juancarloscastillo.github.io/jc-castillo/}{jc-castillo.com}
  - Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile).}.\\
ORCID: 0000-0003-1265-7854

Mauricio Salgado, School of Sociology, Universidad Andrés Bello.\\
ORCID: 0000-0002-8108-6154

Kevin Carrasco, Center for Social Conflict and Cohesion Studies.\\
ORCID: 0000-0002-9348-8361

Andreas Laffert, Institute of Sociology, Pontificia Universidad Católica
de Chile.\\
ORCID: 0000-0002-9008-2454

\subsection{Exclusive Submission
Statement}\label{exclusive-submission-statement}

The authors confirm that this article has been submitted exclusively to
this journal.

\subsection{Acknowledgments}\label{acknowledgments}

The authors acknowledge the support of ANID/FONDECYT grant 1210847 and
the Center for Conflict and Social Cohesion Studies (COES) grant
ANID/FONDAP 1513009.

\subsection{Authors' Contributions}\label{authors-contributions}

Author roles were classified using the Contributor Role Taxonomy
\href{https://credit.niso.org/}{CRediT} as follows: Juan Carlos
Castillo: Conceptualization, Supervision, Methodology, Writing - review
and editing; Mauricio Salgado: Conceptualization, Methodology, Writing -
review and editing; Kevin Carrasco: Methodology, formal analysis,
software, visualization, Writing - review and editing; Andreas Laffert:
Methodology, formal analysis, software, visualization, Writing - review
and editing.

\subsection{Conflict of Interest}\label{conflict-of-interest}

The authors declare no conflict of interest.

\subsection{Funding}\label{funding}

This research received funding from ANID/FONDECYT grant 1210847 and the
Center for Conflict and Social Cohesion Studies (COES) through
ANID/FONDAP grant 1513009.

\subsection{Data Avaibility Statement}\label{data-avaibility-statement}

The original data presented in the study are openly available in
\url{https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/}

Access to the data, code, and analysis presented in the study are openly
available in
\url{https://github.com/educacion-meritocracia/market-justice-preferences-school}.

\end{document}

### 2024-08-07 submission/MDPI/paper-blinded.tex
Update submit mdpi FINAL
@juancarloscastillo

--- 2024-08-05 papelera/paper.tex
@@
 \documentclass[
-    behavsci,
-    article,
-    submit,
-moreauthors
-]{mdpi}
+  12pt,
+  letterpaper,
+]{article}
 
 \usepackage{amsmath,amssymb}
+\usepackage{setspace}
 \usepackage{iftex}
@@
     % xetex/luatex font selection
+  \setmainfont[]{Times New Roman}
 \fi
@@
 \usepackage{xcolor}
+\usepackage[left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm]{geometry}
 \setlength{\emergencystretch}{3em} % prevent overfull lines
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-% Make \paragraph and \subparagraph free-standing
-\makeatletter
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}{
-    \@ifstar
-      \xxxParagraphStar
-      \xxxParagraphNoStar
-  }
-  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
-  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}{
-    \@ifstar
-      \xxxSubParagraphStar
-      \xxxSubParagraphNoStar
-  }
-  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
-  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-\makeatother
 
@@
 \makeatother
-
-% TODO: Add custom LaTeX header directives here
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
+  \fi
+  % set entry spacing
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+% -----------------------
+% CUSTOM PREAMBLE STUFF
+% -----------------------
+
+% -----------------
+% Typography tweaks
+% -----------------
+% Indent size
+\setlength{\parindent}{1pc}  % 1p0
+
+% Fix widows and orphans
+\usepackage[all,defaultlines=2]{nowidow}
+
+% List things
+\usepackage{enumitem}
+% Same document-level indentation for ordered and ordered lists
+\setlist[1]{labelindent=\parindent}
+\setlist[itemize]{leftmargin=*}
+\setlist[enumerate]{leftmargin=*}
+
+% Wrap definition list terms
+% https://tex.stackexchange.com/a/9763/11851
+\setlist[description]{style=unboxed}
+
+% For better TOCs
+\usepackage{tocloft}
+
+% Remove left margin in lists inside longtables
+% https://tex.stackexchange.com/a/378190/11851
+\AtBeginEnvironment{longtable}{\setlist[itemize]{nosep, wide=0pt, leftmargin=*, before=\vspace*{-\baselineskip}, after=\vspace*{-\baselineskip}}}
+
+% Allow for /singlespacing and /doublespacing
+\usepackage{setspace}
+
+% -----------------
+% Title block stuff
+% -----------------
+
+% Abstract
+\usepackage[overload]{textcase}
+\usepackage[runin]{abstract}
+\renewcommand{\abstractnamefont}{\sffamily\footnotesize\bfseries\MakeUppercase}
+\renewcommand{\abstracttextfont}{\sffamily\small}
+\setlength{\absleftindent}{\parindent * 2}
+\setlength{\absrightindent}{\parindent * 2}
+\abslabeldelim{\quad}
+\setlength{\abstitleskip}{-\parindent}
+
+% Keywords
+\newenvironment{keywords}
+{\vskip -3em \hspace{\parindent}\small\sffamily{\sffamily\footnotesize\bfseries\MakeUppercase{Keywords}}\quad}
+{\vskip 3em}
+
+% Title
+\usepackage{titling}
+\setlength{\droptitle}{3em}
+\pretitle{\par\vskip 5em \begin{flushleft}\LARGE\sffamily\bfseries}
+\posttitle{\par\end{flushleft}\vskip 0.75em}
+
+% Authors
+%
+% PHEW this is complicated for a number of reasons!
+%
+% When using \and with multiple authors, the article class in LaTeX wraps each 
+% author block in a tabluar environment with a hardcoded center alignment. It's 
+% possible to use \preauthor{} to start tabulars with a left alignment {l}, but 
+% that only applies to the first author because the others all use \and with the 
+% hardcoded {c}. But we can override the \and command and add our own {l}
+%
+% (See https://github.com/rstudio/rmarkdown/issues/1716#issuecomment-560601691 
+% for an example of redefining \and to just be \\)
+%
+% That's all great, except tabulars have some amount of default horizontal 
+% padding, which makes left-aligned author blocks not actuall get fully 
+% left-aligned on the page. We can set the horizontal padding for the column to 
+% 0, but it requires some wonky syntax: {@{\hspace{0em}}l@{}}
+\renewcommand{\and}{\end{tabular} \hskip 3em \begin{tabular}[t]{@{\hspace{0em}}l@{}}}
+\preauthor{\begin{flushleft}
+           \lineskip 1.5em 
+           \begin{tabular}[t]{@{\hspace{0em}}l@{}}}
+\postauthor{\end{tabular}\par\end{flushleft}}
+
+% Omit the date since the \published command does that
+\predate{}
+\postdate{}
+
+% Command for a note at the top of the first page describing the publication
+% status of the paper.
+\newcommand{\published}[1]{%
+   \gdef\puB{#1}}
+   \newcommand{\puB}{}
+   \renewcommand{\maketitlehooka}{%
+       \par\noindent\footnotesize\sffamily \puB}
+
+% ------------------
+% Section headings
+% ------------------
+\usepackage{titlesec}
+\titleformat*{\section}{\Large\sffamily\bfseries\raggedright}
+\titleformat*{\subsection}{\large\sffamily\bfseries\raggedright}
+\titleformat*{\subsubsection}{\normalsize\sffamily\bfseries\raggedright}
+\titleformat*{\paragraph}{\small\sffamily\bfseries\raggedright}
+
+% \titlespacing{<command>}{<left>}{<before-sep>}{<after-sep>}
+% Starred version removes indentation in following paragraph
+\titlespacing*{\section}{0em}{2em}{0.1em}
+\titlespacing*{\subsection}{0em}{1.25em}{0.1em}
+\titlespacing*{\subsubsection}{0em}{0.75em}{0em}
+
+% -----------
+% Footnotes
+% -----------
+% NB: footmisc has to come after setspace and biblatex because of conflicts
+\usepackage[bottom, flushmargin]{footmisc}
+\renewcommand*{\footnotelayout}{\footnotesize}
+
+\addtolength{\skip\footins}{10pt}    % vertical space between rule and main text
+\setlength{\footnotesep}{5pt}  % vertical space between footnotes
+
+% ----------
+% Captions
+% ----------
+\usepackage[font={small,sf}, labelfont={small,sf,bf}]{caption}
+
+% --------
+% Macros
+% --------
+% pandoc will not convert text within \begin{} XXX \end{} to Markdown and will
+% treat it as regular TeX. Because of this, it's impossible to do stuff like
+% this:
+
+% \begin{landscape}
+%
+% | One | Two   |
+% |-----+-------|
+% | my  | table |
+% | is  | nice  |
+%
+% \end{landscape}
+%
+% Since it'll render like: | One | Two | |—–+——-| | my | table | | is | nice |
+% 
+% BUT, from this http://stackoverflow.com/a/41945462/120898 we can get around
+% this by creating new commands for \begin and \end, like this:
+\usepackage{pdflscape}
+\newcommand{\blandscape}{\begin{landscape}}
+\newcommand{\elandscape}{\end{landscape}}
+
+% \blandscape
+%
+% | One | Two   |
+% |-----+-------|
+% | my  | table |
+% | is  | nice  |
+%
+% \elandscape
+
+% Same thing, but for generic groups
+% But can't use \bgroup and \egroup because those are built-in TeX things
+\newcommand{\stgroup}{\begingroup}
+\newcommand{\fingroup}{\endgroup}
+
+% ---------------------------
+% END CUSTOM PREAMBLE STUFF
+% ---------------------------
 \usepackage{booktabs}
@@
 \usepackage{threeparttable}
-\usepackage{graphicx}
-\includegraphics{pics/logo-mdpi.pdf}
 \makeatletter
@@
 \fi
-\usepackage[]{natbib}
-\bibliographystyle{plainnat}
 \usepackage{bookmark}
@@
 justice, meritocracy, socialization, family, schools, Chile},
-  pdfcreator={LaTeX via pandoc},
-  unicode=true,
-	bookmarksopen={true},
-	pdffitwindow=true, 
   colorlinks=true,
-	linkcolor=bluecite, 
-	citecolor=bluecite, 
-	urlcolor=bluecite, 
-	hyperfootnotes=true, 
-	pdfstartview={FitH},
-	pdfpagemode= UseNone}
-
-\firstpage{1} 
-\makeatletter 
-\setcounter{page}{\@firstpage} 
-\makeatother
-\pubvolume{1}
-\issuenum{1}
-\articlenumber{0}
-\pubyear{\the\year{}}
-\copyrightyear{\the\year{}}
-%\externaleditor{Academic Editor: Firstname Lastname}
-\datereceived{ } 
-\daterevised{ } % Comment out if no revised date
-\dateaccepted{ } 
-\datepublished{ } 
-%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
-%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
-\hreflink{https://doi.org/} % If needed use \linebreak
-%\doinum{}
-%\pdfoutput=1 % Uncommented for upload to arXiv.org
-%\CorrStatement{yes}  % For updates
-
-\Title{ The socialization of meritocracy and market justice preferences
-at school }
-
-% MDPI internal command: Title for citation in the left column
-\TitleCitation{ The socialization of meritocracy and market justice
-preferences at school }
-
-% Author Orchid ID: enter ID or remove command
-%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
-%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
-
-% Authors, for the paper (add full first names)
-%\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{2,\ddagger}$ and Firstname Lastname $^{2,}$*}
-
-\Author{Juan Carlos Castillo$^{aff-1}$, 
-Mauricio Salgado$^{aff-2}$, 
-Kevin Carrasco$^{aff-3}$~and~Andreas Laffert$^{aff-4}$
+  linkcolor={DarkSlateBlue},
+  filecolor={Maroon},
+  citecolor={DarkSlateBlue},
+  urlcolor={DarkSlateBlue},
+  pdfcreator={LaTeX via pandoc}}
+
+% -----------------------
+% END-OF-PREAMBLE STUFF
+% -----------------------
+
+% ---------------------- 
+% Title block elements
+% ---------------------- 
+\usepackage{orcidlink}  % Create automatic ORCID icons/links
+
+\title{The socialization of meritocracy and market justice preferences
+at school}
+
+\author{
 }
-%\longauthorlist{yes}
-
-% MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{
-  Juan Carlos Castillo,
-  Mauricio Salgado,
-  Kevin Carrasco,
-  Andreas Laffert}
-
-% MDPI internal command: Authors, for citation in the left column
-%\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
-%\usepackage{xstring}
-\newcommand\getfirst[1]{\getfirstaux#1\relax} \def\getfirstaux#1#2\relax{#1}
-\newcommand{\authorformatted}{Castillo, \getfirst{Juan Carlos}.; Salgado, \getfirst{Mauricio}.; Carrasco, \getfirst{Kevin}.; Laffert, \getfirst{Andreas}.}
-\AuthorCitation{\authorformatted}
-
-% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
-
-% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
-\address {%
-$^{aff-1}$ \quad , Department of Sociology, Universidad de Chile\\
-$^{aff-2}$ \quad , Centro de Estudios Públicos, Chile\\
-$^{aff-3}$ \quad , Center for Social Conflict and Cohesion Studies\\
-$^{aff-4}$ \quad , Institute of Sociology, Universidad Católica de
-Chile\\
-}
-
-%\address{
-%$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
-%$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
-
-% Contact information of the corresponding author
-%% see: https://quarto.org/docs/journals/authors.html
-\corres{Correspondence: juancastillov@uchile.cl}
-
-% Current address and/or shared authorship
-%\firstnote{Current address: Affiliation.}  % Current address should not be the same as any items in the Affiliation section.
-%\secondnote{These authors contributed equally to this work.}
-% The commands \thirdnote{} till \eighthnote{} are available for further notes
-
-%\simplesumm{} % Simple summary
-
-%\conference{} % An extended version of a conference paper
-
-% Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{ Previous research has shown that schools justify student
-performance differences based on meritocratic ideals. However, the link
-between school meritocracy and students' justification of inequalities
-has received little attention. This paper examines how school and family
+
+\date{}
+
+% Typeset URLs in the same font as their parent environment
+%
+% This has to come at the end of the preamble, after any biblatex stuff because 
+% some biblatex styles (like APA) define their own \urlstyle{}
+\usepackage{url}
+\urlstyle{same}
+
+% ---------------------------
+% END END-OF-PREAMBLE STUFF
+% ---------------------------
+\begin{document}
+% ---------------
+% TITLE SECTION
+% ---------------
+\published{\textbf{}}
+
+\maketitle
+
+\begin{abstract}
+Previous research has shown that schools justify student performance
+differences based on meritocratic ideals. However, the link between
+school meritocracy and students' justification of inequalities has
+received little attention. This paper examines how school and family
 socialization influence students' beliefs about meritocracy and
@@
 results suggest that meritocratic beliefs and market justice preferences
-begin early and are influenced by the school environment. }
-
-% Keywords
-\keyword{market
-justice; meritocracy; socialization; family; schools; Chile}
-
-\begin{document}
-
+begin early and are influenced by the school environment.
+\end{abstract}
+\vskip 3em
+
+\begin{keywords}
+\def\sep{;\ }
+market
+justice\sep meritocracy\sep socialization\sep family\sep schools\sep 
+Chile
+\end{keywords}
+
+\begin{ideuno}
+
+Front Matter
+
+\end{ideuno}
+% -------------------
+% END TITLE SECTION
+% -------------------
+
+\newpage{}
+
+\begin{idedos}
+
+Manuscript
+\end{idedos}
+
+\setstretch{1.15}
 \section{Introduction}\label{introduction}
@@
 at the school level represents a threat to the promise of education and
-a meritocratic system \citep{bourdieu_reproduction_1990}. A large part
-of the research in this field at an international level has addressed
-the extent to which the social origin of students affects their academic
-results and their life opportunities \citep{vonhippel_test_2019},
-confirming that schools have severe difficulties in closing the
-socio-economic and cultural gaps of origin. Besides this socioeconomic
-perspective on school opportunities, recent research has addressed to
-what extent inequalities in the school context are also influencing
-students' perceptions, beliefs, and attitudes: Are social inequalities
-even perceived at the school context? Are they rejected by the students,
-particularly those who are worst-off in socioeconomic terms? Or, Is
-there evidence at the school level that social inequalities are
-tolerated and even justified?
-\citep{batruch_belief_2022, wiederkehr_belief_2015}.
+a meritocratic system
+{[}\citeproc{ref-bourdieu_reproduction_1990}{1}{]}. A large part of the
+research in this field at an international level has addressed the
+extent to which the social origin of students affects their academic
+results and their life opportunities
+{[}\citeproc{ref-vonhippel_test_2019}{2}{]}, confirming that schools
+have severe difficulties in closing the socio-economic and cultural gaps
+of origin. Besides this socioeconomic perspective on school
+opportunities, recent research has addressed to what extent inequalities
+in the school context are also influencing students' perceptions,
+beliefs, and attitudes: Are social inequalities even perceived at the
+school context? Are they rejected by the students, particularly those
+who are worst-off in socioeconomic terms? Or, Is there evidence at the
+school level that social inequalities are tolerated and even justified?
+{[}\citeproc{ref-batruch_belief_2022}{3},\citeproc{ref-wiederkehr_belief_2015}{4}{]}.
 
@@
 based on their abilities, knowledge, and achievements
-\citep{young_rise_1958}. It is often seen as a way to create equal
-opportunities and fairness, as individuals can rise to positions of
-power and influence based on their own merit rather than their
-background or connections. However, some argue that meritocracy can
-actually lead to tolerating or even justifying social inequalities, as
-it can create a hierarchy where those who already have resources and
+{[}\citeproc{ref-young_rise_1958}{5}{]}. It is often seen as a way to
+create equal opportunities and fairness, as individuals can rise to
+positions of power and influence based on their own merit rather than
+their background or connections. However, some argue that meritocracy
+can actually lead to tolerating or even justifying social inequalities,
+as it can create a hierarchy where those who already have resources and
 advantages are more likely to succeed
-\citep{sandel_tyranny_2020, mcnamee_meritocracy_2004}. In this regard, a
-great deal of academic research about meritocracy delves into the
-assessment of to what extent rewards and privileges in society are
-related to merit, emphasizing the so-called unfulfillable promise of
-meritocracy \citep{mijs_stratified_2016}.
+{[}\citeproc{ref-sandel_tyranny_2020}{6},\citeproc{ref-mcnamee_meritocracy_2004}{7}{]}.
+In this regard, a great deal of academic research about meritocracy
+delves into the assessment of to what extent rewards and privileges in
+society are related to merit, emphasizing the so-called unfulfillable
+promise of meritocracy {[}\citeproc{ref-mijs_stratified_2016}{8}{]}.
 
@@
 payment capacity, referred to as market justice
-\citep{lane_market_1986, lindh_bringing_2023}. Most of the research in
-this realm has only considered adults, leaving aside the study of how
-beliefs in this field develop at student age as well as the impact of
-the school context and the family as the main socialization agencies
-\citep{batruch_belief_2022}. Regarding schools, the way in which they
-deal with unequal conditions of origin has been linked to the
-\emph{hidden curriculum} \citep{chafel_schooling_1997}, whereby students
-learn about distributive norms in society and mechanisms of
-justification of social differences.
+{[}\citeproc{ref-lane_market_1986}{9},\citeproc{ref-lindh_bringing_2023}{10}{]}.
+Most of the research in this realm has only considered adults, leaving
+aside the study of how beliefs in this field develop at student age as
+well as the impact of the school context and the family as the main
+socialization agencies {[}\citeproc{ref-batruch_belief_2022}{3}{]}.
+Regarding schools, the way in which they deal with unequal conditions of
+origin has been linked to the \emph{hidden curriculum}
+{[}\citeproc{ref-chafel_schooling_1997}{11}{]}, whereby students learn
+about distributive norms in society and mechanisms of justification of
+social differences.
 
@@
 justification of economic inequalities in the adult population
-\citep{batruch_belief_2022, wiederkehr_belief_2015}, the central
-hypothesis guiding this research is that school-age students with a
-higher perception of meritocracy - both at school and at the societal
-level - will show a larger market justice preferences, as individual
-achievement would be seen as appropriately rewarded and social
-mechanisms for correcting inequalities as less necessary
-\citep{batruch_belief_2022}. We focus on the student-age population as
-we point out that it is possible to track down the origin of
-meritocratic beliefs (and their consequences) to early socialization
+{[}\citeproc{ref-batruch_belief_2022}{3},\citeproc{ref-wiederkehr_belief_2015}{4}{]},
+the central hypothesis guiding this research is that school-age students
+with a higher perception of meritocracy - both at school and at the
+societal level - will show a larger market justice preferences, as
+individual achievement would be seen as appropriately rewarded and
+social mechanisms for correcting inequalities as less necessary
+{[}\citeproc{ref-batruch_belief_2022}{3}{]}. We focus on the student-age
+population as we point out that it is possible to track down the origin
+of meritocratic beliefs (and their consequences) to early socialization
 processes. To this regard, we take into account the family and the
@@
 economic inequality, which stands out in Latin America and among OECD
-countries \citep{flores_top_2020}. In Chile, the poorest 50\% captures
-only 10\% of the total income and has negative wealth, while the richest
-1\% receives almost 27\% of the income and holds 49.6\% of the wealth
-\citep{chancel_world_2022}. Much of this inequality has been attributed
-to the deep neoliberal reforms that institutionalized privatization and
-commodification of various economic sectors. Such reforms where
-introduced during the dictatorship (1973-1989) and expanded in democracy
-through concessions, demand credits, and specific regulatory frameworks
-\citep{ffrench-davis_reformas_2018}. This shift in economic policy has
-allowed the unprecedented emergence of markets in areas related to
-social reproduction, such as health, pensions, and education, with
-provision and access managed by private entities and segmented by
-individual payment capacity, heavily reliant on State subsidies
-\citep{boccardo_30_2020}. In health, although the majority of the
-population uses the public insurance system (78.9\%), 15.3\% are served
-by private insurers \citep{observatoriosocial_estadisticas_2024}. The
+countries {[}\citeproc{ref-flores_top_2020}{12}{]}. In Chile, the
+poorest 50\% captures only 10\% of the total income and has negative
+wealth, while the richest 1\% receives almost 27\% of the income and
+holds 49.6\% of the wealth {[}\citeproc{ref-chancel_world_2022}{13}{]}.
+Much of this inequality has been attributed to the deep neoliberal
+reforms that institutionalized privatization and commodification of
+various economic sectors. Such reforms where introduced during the
+dictatorship (1973-1989) and expanded in democracy through concessions,
+demand credits, and specific regulatory frameworks
+{[}\citeproc{ref-ffrench-davis_reformas_2018}{14}{]}. This shift in
+economic policy has allowed the unprecedented emergence of markets in
+areas related to social reproduction, such as health, pensions, and
+education, with provision and access managed by private entities and
+segmented by individual payment capacity, heavily reliant on State
+subsidies {[}\citeproc{ref-boccardo_30_2020}{15}{]}. In health, although
+the majority of the population uses the public insurance system
+(78.9\%), 15.3\% are served by private insurers
+{[}\citeproc{ref-observatoriosocial_estadisticas_2024}{16}{]}. The
 pension system is based on individual capitalization, with mandatory
@@
 financial market, currently involving 11 million contributors
-\citep{superintendenciadepensiones_estadisticas_2024}. In education,
-30.6\% of school enrollment is in public schools, 54.0\% in
-state-subsidized (voucher) private schools, and 9.3\% in fully private
-schools, generally attracting higher-income groups
-\citep{ministeriodeeducacion_resumen_2023}.
+{[}\citeproc{ref-superintendenciadepensiones_estadisticas_2024}{17}{]}.
+In education, 30.6\% of school enrollment is in public schools, 54.0\%
+in state-subsidized (voucher) private schools, and 9.3\% in fully
+private schools, generally attracting higher-income groups
+{[}\citeproc{ref-ministeriodeeducacion_resumen_2023}{18}{]}.
 
@@
 based on prices and individuals' ability to pay
-\citep{boltanski_new_2005, lane_market_1986, streeck_citizens_2012}.
+{[}\citeproc{ref-lane_market_1986}{9},\citeproc{ref-boltanski_new_2005}{19},\citeproc{ref-streeck_citizens_2012}{20}{]}.
 Research on social stratification beliefs, which explore individual
-perceptions of who deserves what and why \citep{kluegel_beliefs_1987},
-highlights that people's explanations and justifications of social
-inequality are closely tied to their judgments of deservingness. The
-influence of ideologies \citep{wegener_dominant_1995} and cultural
-schemas \citep{homan_being_2017} is pivotal in shaping these
+perceptions of who deserves what and why
+{[}\citeproc{ref-kluegel_beliefs_1987}{21}{]}, highlights that people's
+explanations and justifications of social inequality are closely tied to
+their judgments of deservingness. The influence of ideologies
+{[}\citeproc{ref-wegener_dominant_1995}{22}{]} and cultural schemas
+{[}\citeproc{ref-homan_being_2017}{23}{]} is pivotal in shaping these
 explanations by offering symbolic representations that frame societal
@@
 the literature
-\citep{castillo_legitimacy_2011, evans_family_2010, jasso_how_1999, shariff_income_2016},
+{[}\citeproc{ref-castillo_legitimacy_2011}{24}--\citeproc{ref-shariff_income_2016}{27}{]},
 there has been less examination of public beliefs about which life
 domains should be governed by market relations
-\citep{lindh_bringing_2023} and even less about children's acceptance or
-rejection of these market principles. This oversight is notable given
-the extensive encroachment of market logic into public goods, welfare
-policy, and social services over the past five decades
-\citep{centeno_arc_2012, harvey_breve_2015}, affecting areas such as
-pensions, health services, and education.
+{[}\citeproc{ref-lindh_bringing_2023}{10}{]} and even less about
+children's acceptance or rejection of these market principles. This
+oversight is notable given the extensive encroachment of market logic
+into public goods, welfare policy, and social services over the past
+five decades
+{[}\citeproc{ref-centeno_arc_2012}{28},\citeproc{ref-harvey_breve_2015}{29}{]},
+affecting areas such as pensions, health services, and education.
 
@@
 management of social services across nations
-\citep{jensen_worlds_2008, stoy_worlds_2014}. Nordic countries, for
-example, predominantly employ public agencies to produce and provide
-social services, funding these through collective taxation and offering
-them in kind to the majority of citizens. This system prioritizes social
-justice, placing it above market mechanisms in accessing services. In
-contrast, other countries rely more heavily on for-profit entities and
-private funding, where service distribution depends mainly on individual
-financial capacity to pay user fees, highlighting the influence of
-market justice in service allocation. The trend toward marketization of
-welfare services has been growing since the 1980s
-\citep{salamon_marketization_1993}, and this shift is increasingly
-evident even in countries where market solutions have traditionally had
-a minor role in social policy \citep{sivesind_changing_2017}. The
-expansion of marketization has been related to a larger justification of
-market mechanisms, whereby societies with larger private spending on
-services show larger market justice preferences
-\citep{lindh_public_2015}.
+{[}\citeproc{ref-jensen_worlds_2008}{30},\citeproc{ref-stoy_worlds_2014}{31}{]}.
+Nordic countries, for example, predominantly employ public agencies to
+produce and provide social services, funding these through collective
+taxation and offering them in kind to the majority of citizens. This
+system prioritizes social justice, placing it above market mechanisms in
+accessing services. In contrast, other countries rely more heavily on
+for-profit entities and private funding, where service distribution
+depends mainly on individual financial capacity to pay user fees,
+highlighting the influence of market justice in service allocation. The
+trend toward marketization of welfare services has been growing since
+the 1980s {[}\citeproc{ref-salamon_marketization_1993}{32}{]}, and this
+shift is increasingly evident even in countries where market solutions
+have traditionally had a minor role in social policy
+{[}\citeproc{ref-sivesind_changing_2017}{33}{]}. The expansion of
+marketization has been related to a larger justification of market
+mechanisms, whereby societies with larger private spending on services
+show larger market justice preferences
+{[}\citeproc{ref-lindh_public_2015}{34}{]}.
 
@@
 stimulating a sense of deserving more than one gets''
-\citeyearpar[p.~384]{lane_market_1986}. Contrary to the evidence that
+{[}\citeproc{ref-lane_market_1986}{9}{]}. Contrary to the evidence that
 unequal distribution produces feelings of dissatisfaction, anger, and
 resentment that might motivate forms of collective action
-\citep{greitemeyer_subjective_2016, mishra_subjective_2015, smith_relative_2012, power_deprivationprotest_2018},
+{[}\citeproc{ref-greitemeyer_subjective_2016}{35}--\citeproc{ref-power_deprivationprotest_2018}{38}{]},
 Lane pointed out that in market settings, social comparisons are more
@@
 distributive principles that characterize welfare states, such as need
-and equality (see \citep{wilson_role_2003}).
+and equality (see {[}\citeproc{ref-wilson_role_2003}{39}{]}).
 
@@
 that varies across the region
-\citep{bucca_merit_2016, chong_mystery_2008, torche_intergenerational_2014, salgado_inequality_2023}.
+{[}\citeproc{ref-bucca_merit_2016}{40}--\citeproc{ref-salgado_inequality_2023}{43}{]}.
 The reliance on private welfare providers and widespread user fees
-\citep{molyneux_neoliberal_2008} adds complexity to this context, as
-reflected in surveys conducted by the Center for Public Studies (CEP).
-According to this data, 35.9\% prefer private health insurance, and 63\%
-would prefer private education
-\citep{centrodeestudiospublicos_estudio_2024}. Yet, research on
-children's justification in this area remains limited, highlighting a
-significant gap in understanding how younger generations view
-market-based access to welfare and whether these views are associated
-with their meritocratic beliefs.
+{[}\citeproc{ref-molyneux_neoliberal_2008}{44}{]} adds complexity to
+this context, as reflected in surveys conducted by the Center for Public
+Studies (CEP). According to this data, 35.9\% prefer private health
+insurance, and 63\% would prefer private education
+{[}\citeproc{ref-centrodeestudiospublicos_estudio_2024}{45}{]}. Yet,
+research on children's justification in this area remains limited,
+highlighting a significant gap in understanding how younger generations
+view market-based access to welfare and whether these views are
+associated with their meritocratic beliefs.
 
@@
 The original definition of merit is a combination of effort and talent
-\citep{young_rise_1958}, and a meritocracy is a distributive system
-where merit is the main criterion for allocating valuable goods and
-rewards. From a sociological perspective, meritocracy has been used in
-research on social mobility to characterize societies with low mobility
-that threaten the meritocratic ideal \citep{goldthorpe_myth_2003}. More
-recently, sociology and social psychology research has attended to
-subjective aspects related to the support for meritocratic principles in
-different societies, such as beliefs in meritocracy
-\citep{castillo_multidimensional_2023, mijs_unfulfillable_2016, mijs_paradox_2019}.
+{[}\citeproc{ref-young_rise_1958}{5}{]}, and a meritocracy is a
+distributive system where merit is the main criterion for allocating
+valuable goods and rewards. From a sociological perspective, meritocracy
+has been used in research on social mobility to characterize societies
+with low mobility that threaten the meritocratic ideal
+{[}\citeproc{ref-goldthorpe_myth_2003}{46}{]}. More recently, sociology
+and social psychology research has attended to subjective aspects
+related to the support for meritocratic principles in different
+societies, such as beliefs in meritocracy
+{[}\citeproc{ref-castillo_multidimensional_2023}{47}--\citeproc{ref-mijs_paradox_2019}{49}{]}.
 
 Meritocratic beliefs can cover two types of subjective processes:
-preferences and perceptions \citep{castillo_multidimensional_2023}.
-While meritocratic preferences refer to a justification of distribution
-based on merit criteria (effort and talent), the perception of
-meritocracy refers to how individuals view and understand the concept of
-meritocracy in their society
-\citep{duru-bellat_who_2012, castillo_meritocracia_2019}. The perception
-can vary greatly depending on individual experiences, as well as social,
-economic, and cultural background. Some people may see meritocracy as a
-fair and just system that allows anyone to succeed based on their
-abilities and hard work. In contrast, others may view it as a myth or a
-cover for existing power dynamics and inequality, serving to maintain
-and even reinforce inequality
-\citep{lampert_meritocratic_2013, mijs_paradox_2019}. Some studies have
-analyzed how those with greater privileges believe more in meritocracy
-\citep{reynolds_perceptions_2014}, how greater economic inequality
-increases meritocratic beliefs \citep{mijs_paradox_2019}, and how larger
-inequality decreases it \citep{morris_representing_2022}.
+preferences and perceptions
+{[}\citeproc{ref-castillo_multidimensional_2023}{47}{]}. While
+meritocratic preferences refer to a justification of distribution based
+on merit criteria (effort and talent), the perception of meritocracy
+refers to how individuals view and understand the concept of meritocracy
+in their society
+{[}\citeproc{ref-duru-bellat_who_2012}{50},\citeproc{ref-castillo_meritocracia_2019}{51}{]}.
+The perception can vary greatly depending on individual experiences, as
+well as social, economic, and cultural background. Some people may see
+meritocracy as a fair and just system that allows anyone to succeed
+based on their abilities and hard work. In contrast, others may view it
+as a myth or a cover for existing power dynamics and inequality, serving
+to maintain and even reinforce inequality
+{[}\citeproc{ref-mijs_paradox_2019}{49},\citeproc{ref-lampert_meritocratic_2013}{52}{]}.
+Some studies have analyzed how those with greater privileges believe
+more in meritocracy {[}\citeproc{ref-reynolds_perceptions_2014}{53}{]},
+how greater economic inequality increases meritocratic beliefs
+{[}\citeproc{ref-mijs_paradox_2019}{49}{]}, and how larger inequality
+decreases it {[}\citeproc{ref-morris_representing_2022}{54}{]}.
 
@@
 less support for redistributive compensation systems
-\citep{frank_performance_2015}, as individual achievement would be seen
-as rewarded and social policies as less necessary.
-\citet{almas_cutthroat_2020} found that in the US the highly educated
-accept inequality significantly more than the less educated because they
-perceive inequality as justifiable owing to differences in productivity
-(i.e., merit). \citet{barr_effect_2020} found that in unequal societies,
-the highly educated accept inequality more than the low educated.
-Conversely, individuals tend to support redistribution when they believe
-that the disadvantaged lack the opportunities to succeed
-\citep{evans_strong_2018}.
+{[}\citeproc{ref-frank_performance_2015}{55}{]}, as individual
+achievement would be seen as rewarded and social policies as less
+necessary. {[}\citeproc{ref-almas_cutthroat_2020}{56}{]} found that in
+the US the highly educated accept inequality significantly more than the
+less educated because they perceive inequality as justifiable owing to
+differences in productivity (i.e., merit).
+{[}\citeproc{ref-barr_effect_2020}{57}{]} found that in unequal
+societies, the highly educated accept inequality more than the low
+educated. Conversely, individuals tend to support redistribution when
+they believe that the disadvantaged lack the opportunities to succeed
+{[}\citeproc{ref-evans_strong_2018}{58}{]}.
 
@@
 higher-class families because the dominant group's culture shapes
-educational institutions \citep{bourdieu_reproduction_1990}. Middle- and
+educational institutions
+{[}\citeproc{ref-bourdieu_reproduction_1990}{1}{]}. Middle- and
 upper-class students are better equipped to face academic challenges and
 are more familiar with academic expectations
-\citep{mikus_children_2019}. Such familiarity represents cultural
-capital in educational contexts because higher-status students come to
-school ready to meet these expectations and reap the benefits
-\citep{jack_no_2016, khan_privilege_2011}. Conversely, lower-status
-children lacking cultural capital must catch up while experiencing
-inequitable comparisons \citep{goudeau_hidden_2017}. Additionally,
-academic achievement is treated as the outcome of dispositional factors
-(e.g., pupils' efforts and talents or lack of them) rather than the
-result of differential access to critical resources. Due to the
-meritocratic frame schools encourage, both low- and high-status
-individuals tend to believe that success or failure is not due to the
-family background but rather to differences in efforts and talents
-\citep{darnon_where_2018}. In this sense, we believe that the perception
-of meritocracy can influence students' judgments about market justice
-preferences, leading to our first hypothesis:
+{[}\citeproc{ref-mikus_children_2019}{59}{]}. Such familiarity
+represents cultural capital in educational contexts because
+higher-status students come to school ready to meet these expectations
+and reap the benefits
+{[}\citeproc{ref-jack_no_2016}{60},\citeproc{ref-khan_privilege_2011}{61}{]}.
+Conversely, lower-status children lacking cultural capital must catch up
+while experiencing inequitable comparisons
+{[}\citeproc{ref-goudeau_hidden_2017}{62}{]}. Additionally, academic
+achievement is treated as the outcome of dispositional factors (e.g.,
+pupils' efforts and talents or lack of them) rather than the result of
+differential access to critical resources. Due to the meritocratic frame
+schools encourage, both low- and high-status individuals tend to believe
+that success or failure is not due to the family background but rather
+to differences in efforts and talents
+{[}\citeproc{ref-darnon_where_2018}{63}{]}. In this sense, we believe
+that the perception of meritocracy can influence students' judgments
+about market justice preferences, leading to our first hypothesis:
 
@@
 specifically to the school context. For instance,
-\citet{resh_sense_2010} find that perception of justice in grades has a
-positive effect on liberal democratic orientation, and trust in people
-and formal institutions \citep{resh_sense_2014}. Attending to this
-evidence, we differentiate in this study between meritocratic perception
-in society at large, and meritocratic perception at school, proposing
-the next hypothesis:
+{[}\citeproc{ref-resh_sense_2010}{64}{]} find that perception of justice
+in grades has a positive effect on liberal democratic orientation, and
+trust in people and formal institutions
+{[}\citeproc{ref-resh_sense_2014}{65}{]}. Attending to this evidence, we
+differentiate in this study between meritocratic perception in society
+at large, and meritocratic perception at school, proposing the next
+hypothesis:
 
@@
 consideration, while working-class parents promote deference, obedience,
-and conformity to authority \citep{kohn_social_1963, kohn_class_1969}.
+and conformity to authority
+{[}\citeproc{ref-kohn_social_1963}{66},\citeproc{ref-kohn_class_1969}{67}{]}.
 Although parents from all social backgrounds encourage individualism in
 their children, this shared norm translates into different forms in high
-and low social classes \citeyearpar{kohn_class_1969}.
-\citet{acemoglu_obedience_2021} claimed that the values families impart
-to their children interact with social mobility. On the one hand,
-children from privileged families are socialized to adopt a clear
-conception of individualism that highlights their internal states,
+and low social classes {[}\citeproc{ref-kohn_class_1969}{67}{]}.
+{[}\citeproc{ref-acemoglu_obedience_2021}{68}{]} claimed that the values
+families impart to their children interact with social mobility. On the
+one hand, children from privileged families are socialized to adopt a
+clear conception of individualism that highlights their internal states,
 independence, and idiosyncrasies. In contrast, children from
@@
 overcome collective impediments on the path to upward mobility
-\citep{iacoviello_collectivism_2019}. In the same line,
-\citet{almas_fairness_2017} found that adolescents from
+{[}\citeproc{ref-iacoviello_collectivism_2019}{69}{]}. In the same line,
+{[}\citeproc{ref-almas_fairness_2017}{70}{]} found that adolescents from
 low-socioeconomic-status families are likelier to have an egalitarian
@@
 significantly influences children's and adolescents' views on inequality
-and deservingness. For example, \citet{jonsson_institutional_2015} study
-revealed that higher-status adolescents in Sweden tend to perpetuate
-social class stereotypes while describing the vocational and academic
-tracks. Academic track students are depicted as wealthy, intelligent,
-ambitious, and diligent, while vocational track students are
-characterized as poor, unambitious, unintelligent, and lackadaisical.
-These stereotypes help individuals maintain a sense of superiority over
-others and legitimize the prevailing social hierarchies and economic
-disparities \citep{jost_attitudinal_2000}. In this line, and besides
-status, we also expect that schools that exhibit average better
+and deservingness. For example,
+{[}\citeproc{ref-jonsson_institutional_2015}{71}{]} study revealed that
+higher-status adolescents in Sweden tend to perpetuate social class
+stereotypes while describing the vocational and academic tracks.
+Academic track students are depicted as wealthy, intelligent, ambitious,
+and diligent, while vocational track students are characterized as poor,
+unambitious, unintelligent, and lackadaisical. These stereotypes help
+individuals maintain a sense of superiority over others and legitimize
+the prevailing social hierarchies and economic disparities
+{[}\citeproc{ref-jost_attitudinal_2000}{72}{]}. In this line, and
+besides status, we also expect that schools that exhibit average better
 performance scores in standardized large-scale assessments would also
@@
 exhibited repetitive and careless response patterns
-\citep{gottfried_autocorrelation_2022}. Additionally, we utilized data
-from the System for Measuring the Quality of Education (SIMCE) conducted
-by the Ministry of Education in 2017, which provides school-level
-information such as administrative dependence, socioeconomic
-classification, and results obtained in standardized mathematics and
-language tests. After processing the variables and removing missing
-cases, the final database for this study includes a two-level stratified
-sample composed of 5,047 students and parents (level 1) nested within
-231 schools (level 2).
+{[}\citeproc{ref-gottfried_autocorrelation_2022}{73}{]}. Additionally,
+we utilized data from the System for Measuring the Quality of Education
+(SIMCE) conducted by the Ministry of Education in 2017, which provides
+school-level information such as administrative dependence,
+socioeconomic classification, and results obtained in standardized
+mathematics and language tests. After processing the variables and
+removing missing cases, the final database for this study includes a
+two-level stratified sample composed of 5,047 students and parents
+(level 1) nested within 231 schools (level 2).
 
@@
 of fixed effects between groups and random effects that vary from one
-group to another \citep{bell_fixed_2019, hox_multilevel_2010}.
+group to another
+{[}\citeproc{ref-bell_fixed_2019}{74},\citeproc{ref-hox_multilevel_2010}{75}{]}.
 Cumulative link mixed models were employed for the ordinal dependent
@@
 
-\includegraphics{paper_files/figure-pdf/fig-marketjustice-1.pdf}
+\includegraphics{paper-blinded_files/figure-pdf/fig-marketjustice-1.pdf}
 
@@
 
-\includegraphics{paper_files/figure-pdf/fig-meritocracy-1.pdf}
+\includegraphics{paper-blinded_files/figure-pdf/fig-meritocracy-1.pdf}
 
@@
 
-\includegraphics{paper_files/figure-pdf/fig-bivariate-1.pdf}
+\includegraphics{paper-blinded_files/figure-pdf/fig-bivariate-1.pdf}
 
@@
 item is associated with specific characteristics of the schools, while
-it is 3\% for the pensions item and 4\% for the education item
-\citep[ICC based on][p.15]{hox_multilevel_2010}. For the average index
-of market justice preferences, the ICC reaches 4\%, indicating that
-overall, only a small percentage of the total variance in students'
+it is 3\% for the pensions item and 4\% for the education item {[}ICC
+based on \citeproc{ref-hox_multilevel_2010}{75}{]}. For the average
+index of market justice preferences, the ICC reaches 4\%, indicating
+that overall, only a small percentage of the total variance in students'
 market justice preferences is associated with school characteristics,
@@
 
-\includegraphics{paper_files/figure-pdf/fig-interaction-1.pdf}
+\includegraphics{paper-blinded_files/figure-pdf/fig-interaction-1.pdf}
 
@@
 meritocratic system are more likely to justify social inequalities
-\citep{batruch_belief_2022, wiederkehr_belief_2015}. It is noteworthy
-that among the items measuring meritocratic perception, consistently the
-one about deservingness shows a larger effect when compared to the ones
-about talent and effort. Although the three items correlate, there is
-something distinctive about deservingness that probably encompasses
-other aspects beyond meritocracy, an aspect to which we will come back
-in the conclusions. Moving now to hypothesis \(H_{1b}\), this is simply
-a variation of 1a, replacing the perception of meritocracy at society by
-the one at school. In this case we have only two items of meritocratic
-perception - unfortunately deservingness was not measured at school
-level - and the direction of the effects is not as straightforward as in
-the case of societal meritocracy in \(H_{1a}\): positive for talent, but
-negative for effort. Still, one of the limitation for the comparison
-between perception of meritocracy and society is the difference of the
-items' phrasing. Whereas for society the item refers to the extent to
-which efforts are rewarded, at school is about the importance of effort,
-whereby something that is important does not necessarily mean that it is
-enough to be rewarded. Furthermore, the school-effort item is the one
-with the lowest variation among all meritocratic items (more than 95\%
-agree or strongly agree), and with the lowest effect size.
+{[}\citeproc{ref-batruch_belief_2022}{3},\citeproc{ref-wiederkehr_belief_2015}{4}{]}.
+It is noteworthy that among the items measuring meritocratic perception,
+consistently the one about deservingness shows a larger effect when
+compared to the ones about talent and effort. Although the three items
+correlate, there is something distinctive about deservingness that
+probably encompasses other aspects beyond meritocracy, an aspect to
+which we will come back in the conclusions. Moving now to hypothesis
+\(H_{1b}\), this is simply a variation of 1a, replacing the perception
+of meritocracy at society by the one at school. In this case we have
+only two items of meritocratic perception - unfortunately deservingness
+was not measured at school level - and the direction of the effects is
+not as straightforward as in the case of societal meritocracy in
+\(H_{1a}\): positive for talent, but negative for effort. Still, one of
+the limitation for the comparison between perception of meritocracy and
+society is the difference of the items' phrasing. Whereas for society
+the item refers to the extent to which efforts are rewarded, at school
+is about the importance of effort, whereby something that is important
+does not necessarily mean that it is enough to be rewarded. Furthermore,
+the school-effort item is the one with the lowest variation among all
+meritocratic items (more than 95\% agree or strongly agree), and with
+the lowest effect size.
 
@@
 contextual and cultural factors that modulate attitudes toward market
-justice \citep{kohn_social_1963, kohn_class_1969, almas_fairness_2017}.
+justice
+{[}\citeproc{ref-kohn_social_1963}{66},\citeproc{ref-kohn_class_1969}{67},\citeproc{ref-almas_fairness_2017}{70}{]}.
 Notably, socioeconomic status measured by parental educational level and
@@
 with the existing literature
-\citep{jonsson_institutional_2015, jost_attitudinal_2000}, does allow us
-to affirm the hypothesis about the role of the school environment,
-particularly socioeconomic segregation, in influencing attitudes towards
-market justice and the acceptance of inequalities. A possible
-explanation is the role of the context at the school level, which could
-have a more relevant role than the family socioeconomic status, since
-the highly stratified nature of the Chilean education system reinforces
-socioeconomic divisions and socializes students into different normative
-frameworks.
+{[}\citeproc{ref-jonsson_institutional_2015}{71},\citeproc{ref-jost_attitudinal_2000}{72}{]},
+does allow us to affirm the hypothesis about the role of the school
+environment, particularly socioeconomic segregation, in influencing
+attitudes towards market justice and the acceptance of inequalities. A
+possible explanation is the role of the context at the school level,
+which could have a more relevant role than the family socioeconomic
+status, since the highly stratified nature of the Chilean education
+system reinforces socioeconomic divisions and socializes students into
+different normative frameworks.
 
@@
 perceptions and preferences about inequality. This can relate to
-development aspects
-\citep[\citet{rizzo_children_2020}]{kim_socioeconomic_2020} but it could
-be something specific to the Chilean case, which requires further
+development aspects {[}\citeproc{ref-rizzo_children_2020}{77}{]} but it
+could be something specific to the Chilean case, which requires further
 comparative as well as longitudinal research in the area. In any case,
@@
 legitimation of social inequalities
-\citep{mijs_paradox_2019, darnon_where_2018}, but the consideration of
-school level population as well as market justice preferences allows
-adding evidence regarding the role of meritocracy regarding inequality
-beliefs: legitimation of market criteria in social services seems to
-begin at school age, and what students think of meritocracy is linked to
-it. Although there are still several open questions, as for instance
-differences between meritocratic perceptions in school and in the
-society at large, this raises up the relevance of what is done (or not)
-in schools in order to challenge the meritocratic ideals.
+{[}\citeproc{ref-mijs_paradox_2019}{49},\citeproc{ref-darnon_where_2018}{63}{]},
+but the consideration of school level population as well as market
+justice preferences allows adding evidence regarding the role of
+meritocracy regarding inequality beliefs: legitimation of market
+criteria in social services seems to begin at school age, and what
+students think of meritocracy is linked to it. Although there are still
+several open questions, as for instance differences between meritocratic
+perceptions in school and in the society at large, this raises up the
+relevance of what is done (or not) in schools in order to challenge the
+meritocratic ideals.
 
@@
 suggests that those doing better - both economically and academically-
-would demand less redistribution \citep[\citet{runst_does_2018},
-\citet{bullock_education_2021}]{gonthier_parallel_2017} and therefore
-are expected to challenge market mechanisms. A possible explanation is
-that education contributes to forming a critical view of markets in
+would demand less redistribution
+{[}\citeproc{ref-bullock_education_2021}{80}{]} and therefore are
+expected to challenge market mechanisms. A possible explanation is that
+education contributes to forming a critical view of markets in
 redistribution, nevertheless, further research in this area is needed to
@@
 
-%\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
-The following statements should be used ``Conceptualization, X.X.
-and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
-and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
-All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
-Authorship must be limited to those who have contributed substantially to the work~reported.}
-
+\newpage{}
+
+\begin{idetres}
+
+Back Matter
+\end{idetres}
+
+\section*{References}\label{bibliography}
+\addcontentsline{toc}{section}{References}
+
+\phantomsection\label{refs}
+\begin{CSLReferences}{0}{0}
+\bibitem[\citeproctext]{ref-bourdieu_reproduction_1990}
+\CSLLeftMargin{1. }%
+\CSLRightInline{Bourdieu, P.; Passeron, J.C. \emph{Reproduction in
+{Education}, {Society} and {Culture}}; Second Edition.; Sage
+Publications Ltd, 1990; ISBN 0-8039-8320-4.}
+
+\bibitem[\citeproctext]{ref-vonhippel_test_2019}
+\CSLLeftMargin{2. }%
+\CSLRightInline{Von Hippel, P.; Hamrock, C. Do {Test Score Gaps Grow}
+Before, During, or Between the {School Years}? {Measurement Artifacts}
+and {What We Can Know} in {Spite} of {Them}. \emph{Sociological Science}
+\textbf{2019}, \emph{6}, 43--80,
+doi:\href{https://doi.org/10.15195/v6.a3}{10.15195/v6.a3}.}
+
+\bibitem[\citeproctext]{ref-batruch_belief_2022}
+\CSLLeftMargin{3. }%
+\CSLRightInline{Batruch, A.; Jetten, J.; Van de Werfhorst, H.; Darnon,
+C.; Butera, F. Belief in {School Meritocracy} and the {Legitimization}
+of {Social} and {Income Inequality}. \emph{Social Psychological and
+Personality Science} \textbf{2022}, 194855062211110,
+doi:\href{https://doi.org/10.1177/19485506221111017}{10.1177/19485506221111017}.}
+
+\bibitem[\citeproctext]{ref-wiederkehr_belief_2015}
+\CSLLeftMargin{4. }%
+\CSLRightInline{Wiederkehr, V.; Bonnot, V.; Krauth-Gruber, S.; Darnon,
+C. Belief in School Meritocracy as a System-Justifying Tool for Low
+Status Students. \emph{Frontiers in Psychology} \textbf{2015},
+\emph{6}.}
+
+\bibitem[\citeproctext]{ref-young_rise_1958}
+\CSLLeftMargin{5. }%
+\CSLRightInline{Young, M. \emph{The Rise of the Meritocracy};
+Transaction Publishers: New Brunswick, N.J., U.S.A, 1958;}
+
+\bibitem[\citeproctext]{ref-sandel_tyranny_2020}
+\CSLLeftMargin{6. }%
+\CSLRightInline{Sandel, M.J. \emph{The Tyranny of Merit: {What}'s Become
+of the Common Good?}; First edition.; {Farrar, Straus and Giroux}: New
+York, 2020; ISBN 978-0-374-28998-0.}
+
+\bibitem[\citeproctext]{ref-mcnamee_meritocracy_2004}
+\CSLLeftMargin{7. }%
+\CSLRightInline{McNamee, S.; Miller, R. \emph{The Meritocracy Myth};
+2004th ed.; Rowman \& Littlefield: Lanham Md., 2004; ISBN
+978-0-7425-1055-5.}
+
+\bibitem[\citeproctext]{ref-mijs_stratified_2016}
+\CSLLeftMargin{8. }%
+\CSLRightInline{Mijs, J. Stratified {Failure}: {Educational
+Stratification} and {Students}' {Attributions} of {Their Mathematics
+Performance} in 24 {Countries}. \emph{Sociology of Education}
+\textbf{2016}, \emph{89}, 137--153,
+doi:\href{https://doi.org/10.1177/0038040716636434}{10.1177/0038040716636434}.}
+
+\bibitem[\citeproctext]{ref-lane_market_1986}
+\CSLLeftMargin{9. }%
+\CSLRightInline{Lane, R.E. Market {Justice}, {Political Justice}.
+\emph{American Political Science Review} \textbf{1986}, \emph{80},
+383--402, doi:\href{https://doi.org/10.2307/1958264}{10.2307/1958264}.}
+
+\bibitem[\citeproctext]{ref-lindh_bringing_2023}
+\CSLLeftMargin{10. }%
+\CSLRightInline{Lindh, A.; McCall, L. Bringing the Market in: An
+Expanded Framework for Understanding Popular Responses to Economic
+Inequality. \emph{Socio-Economic Review} \textbf{2023}, \emph{21},
+1035--1055,
+doi:\href{https://doi.org/10.1093/ser/mwac018}{10.1093/ser/mwac018}.}
+
+\bibitem[\citeproctext]{ref-chafel_schooling_1997}
+\CSLLeftMargin{11. }%
+\CSLRightInline{Chafel, J.A. Schooling, the {Hidden Curriculum}, and
+{Children}'s {Conceptions} of {Poverty}. \emph{Social Policy Report}
+\textbf{1997}, \emph{11}, 1--28,
+doi:\href{https://doi.org/10.1002/j.2379-3988.1997.tb00004.x}{10.1002/j.2379-3988.1997.tb00004.x}.}
+
+\bibitem[\citeproctext]{ref-flores_top_2020}
+\CSLLeftMargin{12. }%
+\CSLRightInline{Flores, I.; Sanhueza, C.; Atria, J.; Mayer, R. Top
+{Incomes} in {Chile}: {A Historical Perspective} on {Income Inequality},
+1964--2017. \emph{Review of Income and Wealth} \textbf{2020}, \emph{66},
+850--874,
+doi:\href{https://doi.org/10.1111/roiw.12441}{10.1111/roiw.12441}.}
+
+\bibitem[\citeproctext]{ref-chancel_world_2022}
+\CSLLeftMargin{13. }%
+\CSLRightInline{Chancel, L.; Piketty, T.; Saez, E.; Zucman, G. World
+Inequality Report 2022 2022.}
+
+\bibitem[\citeproctext]{ref-ffrench-davis_reformas_2018}
+\CSLLeftMargin{14. }%
+\CSLRightInline{Ffrench-Davis, R. \emph{{Reformas econ{ó}micas en Chile
+1973-2017}}; TAURUS, 2018; ISBN 978-956-9635-21-2.}
+
+\bibitem[\citeproctext]{ref-boccardo_30_2020}
+\CSLLeftMargin{15. }%
+\CSLRightInline{Boccardo, G. \emph{30 a{ñ}os de Privatizaciones En
+{Chile}: Lo Que La Pandemia Revel{ó}}; Nodo XXI.; Santiago, 2020;}
+
+\bibitem[\citeproctext]{ref-observatoriosocial_estadisticas_2024}
+\CSLLeftMargin{16. }%
+\CSLRightInline{Social, O. \emph{Estad{í}sticas {Encuesta Casen} 2022,
+Secci{ó}n {Salud}}; Observatorio social, 2024;}
+
+\bibitem[\citeproctext]{ref-superintendenciadepensiones_estadisticas_2024}
+\CSLLeftMargin{17. }%
+\CSLRightInline{Superintendencia de Pensiones \emph{Estad{í}sticas de
+{Afiliados} En {AFP} En {Sistema} de {Pensiones}. {Abril} 2024};
+Superintendencia de Pensiones, 2024;}
+
+\bibitem[\citeproctext]{ref-ministeriodeeducacion_resumen_2023}
+\CSLLeftMargin{18. }%
+\CSLRightInline{Ministerio de Educación \emph{Resumen de Matr{í}cula Por
+Establecimiento Educacional}; Datos abiertos Ministerio de Educaci{ó}n,
+2023;}
+
+\bibitem[\citeproctext]{ref-boltanski_new_2005}
+\CSLLeftMargin{19. }%
+\CSLRightInline{Boltanski, L.; Chiapello, E. \emph{The New Spirit of
+Capitalism}; Verso: London ; New York, 2005; ISBN 978-1-85984-554-7.}
+
+\bibitem[\citeproctext]{ref-streeck_citizens_2012}
+\CSLLeftMargin{20. }%
+\CSLRightInline{Streeck, W. Citizens as {Customers}: {Considerations} on
+the {New Politics} of {Consumption}. \emph{New Left Review}
+\textbf{2012}, \emph{76}.}
+
+\bibitem[\citeproctext]{ref-kluegel_beliefs_1987}
+\CSLLeftMargin{21. }%
+\CSLRightInline{Kluegel, J.R.; Smith, E.R. \emph{Beliefs about
+{Inequality}: {Americans}' {Views} of {What} Is and {What Ought} to Be};
+Routledge: New York, 1987; ISBN 978-0-202-30327-7.}
+
+\bibitem[\citeproctext]{ref-wegener_dominant_1995}
+\CSLLeftMargin{22. }%
+\CSLRightInline{Wegener, B.; Liebig, S. Dominant {Ideologies} and the
+{Variation} of {Distributive Justice Norms}: {A Comparison} of {East}
+and {West Germany}, and the {United States}. In \emph{Social {Justice}
+and {Political Change}: {Public Opinion} in {Capitalist} and
+{Post-Communist States}}; Kluegel, J.R., Mason, D.S., Wegener, B., Eds.;
+Walter de Gruyter: New York, 1995; pp. 239--259 ISBN 978-3-11-086894-4.}
+
+\bibitem[\citeproctext]{ref-homan_being_2017}
+\CSLLeftMargin{23. }%
+\CSLRightInline{Homan, P.; Valentino, L.; Weed, E. Being and {Becoming
+Poor}: {How Cultural Schemas Shape Beliefs About Poverty}. \emph{Social
+Forces} \textbf{2017}, \emph{95}, 1023--1048,
+doi:\href{https://doi.org/10.1093/sf/sox007}{10.1093/sf/sox007}.}
+
+\bibitem[\citeproctext]{ref-castillo_legitimacy_2011}
+\CSLLeftMargin{24. }%
+\CSLRightInline{Castillo, J.C. Legitimacy of {Inequality} in a {Highly
+Unequal Context}: {Evidence} from the {Chilean Case}. \emph{Social
+Justice Research} \textbf{2011}, \emph{24}, 314--340,
+doi:\href{https://doi.org/10.1007/s11211-011-0144-5}{10.1007/s11211-011-0144-5}.}
+
+\bibitem[\citeproctext]{ref-evans_family_2010}
+\CSLLeftMargin{25. }%
+\CSLRightInline{Evans, M.; Kelley, J.; Sikora, J.; Treiman, D. Family
+Scholarly Culture and Educational Success: {Books} and Schooling in 27
+Nations. \emph{Research in Social Stratification and Mobility}
+\textbf{2010}, \emph{28}, 171--197,
+doi:\href{https://doi.org/10.1016/j.rssm.2010.01.002}{10.1016/j.rssm.2010.01.002}.}
+
+\bibitem[\citeproctext]{ref-jasso_how_1999}
+\CSLLeftMargin{26. }%
+\CSLRightInline{Jasso, G. How {Much Injustice} Is {There} in the
+{World}? {Two New Justice Indexes}. \emph{American Sociological Review}
+\textbf{1999}, \emph{64}, 133--168,
+doi:\href{https://doi.org/10.1177/000312249906400110}{10.1177/000312249906400110}.}
+
+\bibitem[\citeproctext]{ref-shariff_income_2016}
+\CSLLeftMargin{27. }%
+\CSLRightInline{Shariff, A.F.; Wiwad, D.; Aknin, L.B. Income {Mobility
+Breeds Tolerance} for {Income Inequality}: {Cross-National} and
+{Experimental Evidence}. \emph{Perspectives on Psychological Science}
+\textbf{2016}, \emph{11}, 373--380,
+doi:\href{https://doi.org/10.1177/1745691616635596}{10.1177/1745691616635596}.}
+
+\bibitem[\citeproctext]{ref-centeno_arc_2012}
+\CSLLeftMargin{28. }%
+\CSLRightInline{Centeno, M.A.; Cohen, J.N. The {Arc} of {Neoliberalism}.
+\emph{Annual Review of Sociology} \textbf{2012}, \emph{38}, 317--340,
+doi:\href{https://doi.org/10.1146/annurev-soc-081309-150235}{10.1146/annurev-soc-081309-150235}.}
+
+\bibitem[\citeproctext]{ref-harvey_breve_2015}
+\CSLLeftMargin{29. }%
+\CSLRightInline{Harvey, D. \emph{{Breve historia del neoliberalismo}};
+Ediciones Akal: Madrid (Espa{ñ}a), 2015; ISBN 978-84-460-2517-7.}
+
+\bibitem[\citeproctext]{ref-jensen_worlds_2008}
+\CSLLeftMargin{30. }%
+\CSLRightInline{Jensen, C. Worlds of Welfare Services and Transfers.
+\emph{Journal of European Social Policy} \textbf{2008}, \emph{18},
+151--162,
+doi:\href{https://doi.org/10.1177/0958928707087591}{10.1177/0958928707087591}.}
+
+\bibitem[\citeproctext]{ref-stoy_worlds_2014}
+\CSLLeftMargin{31. }%
+\CSLRightInline{Stoy, V. Worlds of {Welfare Services}: {From Discovery}
+to {Exploration}. \emph{Social Policy \& Administration} \textbf{2014},
+\emph{48}, 343--360,
+doi:\href{https://doi.org/10.1111/spol.12006}{10.1111/spol.12006}.}
+
+\bibitem[\citeproctext]{ref-salamon_marketization_1993}
+\CSLLeftMargin{32. }%
+\CSLRightInline{Salamon, L.M. The {Marketization} of {Welfare}:
+{Changing Nonprofit} and {For-Profit Roles} in the {American Welfare
+State}. \emph{Social Service Review} \textbf{1993}, \emph{67}, 16--39,
+doi:\href{https://doi.org/10.1086/603963}{10.1086/603963}.}
+
+\bibitem[\citeproctext]{ref-sivesind_changing_2017}
+\CSLLeftMargin{33. }%
+\CSLRightInline{Sivesind, K.H.
+\href{https://doi.org/10.1007/978-3-319-55381-8_2}{The {Changing Roles}
+of {For-Profit} and {Nonprofit Welfare Provision} in {Norway}, {Sweden},
+and {Denmark}}. In \emph{Promoting {Active Citizenship}}; Sivesind,
+K.H., Saglie, J., Eds.; Springer International Publishing: Cham, 2017;
+pp. 33--74 ISBN 978-3-319-55380-1 978-3-319-55381-8.}
+
+\bibitem[\citeproctext]{ref-lindh_public_2015}
+\CSLLeftMargin{34. }%
+\CSLRightInline{Lindh, A. Public {Opinion} Against {Markets}?
+{Attitudes} Towards {Market Distribution} of {Social Services} -- {A
+Comparison} of 17 {Countries}. \emph{Social Policy \& Administration}
+\textbf{2015}, \emph{49}, 887--910,
+doi:\href{https://doi.org/10.1111/spol.12105}{10.1111/spol.12105}.}
+
+\bibitem[\citeproctext]{ref-greitemeyer_subjective_2016}
+\CSLLeftMargin{35. }%
+\CSLRightInline{Greitemeyer, T.; Sagioglou, C. Subjective Socioeconomic
+Status Causes Aggression: {A} Test of the Theory of Social Deprivation.
+\emph{Journal of Personality and Social Psychology} \textbf{2016},
+\emph{111}, 178--194,
+doi:\href{https://doi.org/10.1037/pspi0000058}{10.1037/pspi0000058}.}
+
+\bibitem[\citeproctext]{ref-mishra_subjective_2015}
+\CSLLeftMargin{36. }%
+\CSLRightInline{Mishra, S.; Carleton, R.N. Subjective Relative
+Deprivation Is Associated with Poorer Physical and Mental Health.
+\emph{Social Science \& Medicine} \textbf{2015}, \emph{147}, 144--149,
+doi:\href{https://doi.org/10.1016/j.socscimed.2015.10.030}{10.1016/j.socscimed.2015.10.030}.}
+
+\bibitem[\citeproctext]{ref-smith_relative_2012}
+\CSLLeftMargin{37. }%
+\CSLRightInline{Smith, H.J.; Pettigrew, T.F.; Pippin, G.M.;
+Bialosiewicz, S. Relative {Deprivation}: {A Theoretical} and
+{Meta-Analytic Review}. \emph{Personality and Social Psychology Review}
+\textbf{2012}, \emph{16}, 203--232,
+doi:\href{https://doi.org/10.1177/1088868311430825}{10.1177/1088868311430825}.}
+
+\bibitem[\citeproctext]{ref-power_deprivationprotest_2018}
+\CSLLeftMargin{38. }%
+\CSLRightInline{Power, S.A. The {Deprivation-Protest Paradox}: {How} the
+{Perception} of {Unfair Economic Inequality Leads} to {Civic Unrest}.
+\emph{Current Anthropology} \textbf{2018}, \emph{59}, 765--789,
+doi:\href{https://doi.org/10.1086/700679}{10.1086/700679}.}
+
+\bibitem[\citeproctext]{ref-wilson_role_2003}
+\CSLLeftMargin{39. }%
+\CSLRightInline{Wilson, C. The {Role} of a {Merit Principle} in
+{Distributive Justice}. \emph{The Journal of Ethics} \textbf{2003},
+\emph{7}, 277--314,
+doi:\href{https://doi.org/10.1023/A:1024667228488}{10.1023/A:1024667228488}.}
+
+\bibitem[\citeproctext]{ref-bucca_merit_2016}
+\CSLLeftMargin{40. }%
+\CSLRightInline{Bucca, M. Merit and Blame in Unequal Societies:
+{Explaining Latin Americans}' Beliefs about Wealth and Poverty.
+\emph{Research in Social Stratification and Mobility} \textbf{2016},
+\emph{44}, 98--112,
+doi:\href{https://doi.org/10.1016/j.rssm.2016.02.005}{10.1016/j.rssm.2016.02.005}.}
+
+\bibitem[\citeproctext]{ref-chong_mystery_2008}
+\CSLLeftMargin{41. }%
+\CSLRightInline{Chong, A.; Ñopo, H. The {Mystery} of {Discrimination} in
+{Latin America}. \emph{Econom{í}a} \textbf{2008}, \emph{8}, 79--107,
+doi:\href{https://doi.org/10.1353/eco.0.0005}{10.1353/eco.0.0005}.}
+
+\bibitem[\citeproctext]{ref-torche_intergenerational_2014}
+\CSLLeftMargin{42. }%
+\CSLRightInline{Torche, F. Intergenerational {Mobility} and
+{Inequality}: {The Latin American Case}. \emph{Annual Review of
+Sociology} \textbf{2014}, \emph{40}, 619--642,
+doi:\href{https://doi.org/10.1146/annurev-soc-071811-145521}{10.1146/annurev-soc-071811-145521}.}
+
+\bibitem[\citeproctext]{ref-salgado_inequality_2023}
+\CSLLeftMargin{43. }%
+\CSLRightInline{Salgado, M.; Castillo, J. Inequality and
+{Stratification} in {Latin America}. In \emph{The {Oxford Handbook} of
+{Social Stratification}}; Gangl, M., Platt, L., Polavieja, J.G., van de
+Werfhorst, H.G., Eds.; Oxford University Press: 2023, 2023; p. 0 ISBN
+978-0-19-753948-4.}
+
+\bibitem[\citeproctext]{ref-molyneux_neoliberal_2008}
+\CSLLeftMargin{44. }%
+\CSLRightInline{Molyneux, M. The {``{Neoliberal Turn}''} and the {New
+Social Policy} in {Latin America}: {How Neoliberal}, {How New}?
+\emph{Development and Change} \textbf{2008}, \emph{39}, 775--797,
+doi:\href{https://doi.org/10.1111/j.1467-7660.2008.00505.x}{10.1111/j.1467-7660.2008.00505.x}.}
+
+\bibitem[\citeproctext]{ref-centrodeestudiospublicos_estudio_2024}
+\CSLLeftMargin{45. }%
+\CSLRightInline{Centro de Estudios Públicos \emph{Estudio {Nacional} de
+{Opini{ó}n P{ú}blica Base Consolidada}}; Centro de Estudios P{ú}blicos,
+2024;}
+
+\bibitem[\citeproctext]{ref-goldthorpe_myth_2003}
+\CSLLeftMargin{46. }%
+\CSLRightInline{Goldthorpe, J. The Myth of Education-Based Meritocracy.
+\emph{New Economy} \textbf{2003}, \emph{10}, 234--239,
+doi:\href{https://doi.org/10.1046/j.1468-0041.2003.00324.x}{10.1046/j.1468-0041.2003.00324.x}.}
+
+\bibitem[\citeproctext]{ref-castillo_multidimensional_2023}
+\CSLLeftMargin{47. }%
+\CSLRightInline{Castillo, J.C.; Iturra, J.; Maldonado, L.; Atria, J.;
+Meneses, F. A {Multidimensional Approach} for {Measuring Meritocratic
+Beliefs}: {Advantages}, {Limitations} and {Alternatives} to the {ISSP
+Social Inequality Survey}. \emph{International Journal of Sociology}
+\textbf{2023}, \emph{53}, 448--472,
+doi:\href{https://doi.org/10.1080/00207659.2023.2274712}{10.1080/00207659.2023.2274712}.}
+
+\bibitem[\citeproctext]{ref-mijs_unfulfillable_2016}
+\CSLLeftMargin{48. }%
+\CSLRightInline{Mijs, J. The {Unfulfillable Promise} of {Meritocracy}:
+{Three Lessons} and {Their Implications} for {Justice} in {Education}.
+\emph{Social Justice Research} \textbf{2016}, \emph{29}, 14--34,
+doi:\href{https://doi.org/10.1007/s11211-014-0228-0}{10.1007/s11211-014-0228-0}.}
+
+\bibitem[\citeproctext]{ref-mijs_paradox_2019}
+\CSLLeftMargin{49. }%
+\CSLRightInline{Mijs, J. The Paradox of Inequality: Income Inequality
+and Belief in Meritocracy Go Hand in Hand. \emph{Socio-Economic Review}
+\textbf{2019}, \emph{19}, 7--35,
+doi:\href{https://doi.org/10.1093/ser/mwy051}{10.1093/ser/mwy051}.}
+
+\bibitem[\citeproctext]{ref-duru-bellat_who_2012}
+\CSLLeftMargin{50. }%
+\CSLRightInline{Duru-Bellat, M.; Tenret, E. Who's for {Meritocracy}?
+{Individual} and {Contextual Variations} in the {Faith}.
+\emph{Comparative Education Review} \textbf{2012}, \emph{56}, 223--247,
+doi:\href{https://doi.org/10.1086/661290}{10.1086/661290}.}
+
+\bibitem[\citeproctext]{ref-castillo_meritocracia_2019}
+\CSLLeftMargin{51. }%
+\CSLRightInline{Castillo, J.C.; Torres, A.; Atria, J.; Maldonado, L.
+Meritocracia y Desigualdad Econ{ó}mica: {Percepciones}, Preferencias e
+Implicancias. \emph{Revista Internacional de Sociolog{í}a}
+\textbf{2019}, \emph{77}, 117,
+doi:\href{https://doi.org/10.3989/ris.2019.77.1.17.114}{10.3989/ris.2019.77.1.17.114}.}
+
+\bibitem[\citeproctext]{ref-lampert_meritocratic_2013}
+\CSLLeftMargin{52. }%
+\CSLRightInline{Lampert, K.
+\emph{\href{https://doi.org/10.1057/9781137324894}{Meritocratic
+{Education} and {Social Worthlessness}}}; Palgrave Macmillan UK: London,
+2013; ISBN 978-1-349-45908-7 978-1-137-32489-4.}
+
+\bibitem[\citeproctext]{ref-reynolds_perceptions_2014}
+\CSLLeftMargin{53. }%
+\CSLRightInline{Reynolds, J.; Xian, H. Perceptions of Meritocracy in the
+Land of Opportunity. \emph{Research in Social Stratification and
+Mobility} \textbf{2014}, \emph{36}, 121--137,
+doi:\href{https://doi.org/10.1016/j.rssm.2014.03.001}{10.1016/j.rssm.2014.03.001}.}
+
+\bibitem[\citeproctext]{ref-morris_representing_2022}
+\CSLLeftMargin{54. }%
+\CSLRightInline{Morris, J.; Reilly, J.; Paltsev, S.; Sokolov, A.; Cox,
+K. Representing {Socio}-{Economic Uncertainty} in {Human System Models}.
+\emph{Earth's Future} \textbf{2022}, \emph{10}, e2021EF002239,
+doi:\href{https://doi.org/10.1029/2021EF002239}{10.1029/2021EF002239}.}
+
+\bibitem[\citeproctext]{ref-frank_performance_2015}
+\CSLLeftMargin{55. }%
+\CSLRightInline{Frank, D.H.; Wertenbroch, K.; Maddux, W.W. Performance
+Pay or Redistribution? {Cultural} Differences in Just-World Beliefs and
+Preferences for Wage Inequality. \emph{Organizational Behavior and Human
+Decision Processes} \textbf{2015}, \emph{130}, 160--170,
+doi:\href{https://doi.org/10.1016/j.obhdp.2015.04.001}{10.1016/j.obhdp.2015.04.001}.}
+
+\bibitem[\citeproctext]{ref-almas_cutthroat_2020}
+\CSLLeftMargin{56. }%
+\CSLRightInline{Almås, I.; Cappelen, A.W.; Tungodden, B. Cutthroat
+{Capitalism} Versus {Cuddly Socialism}: {Are Americans More
+Meritocratic} and {Efficiency-Seeking} Than {Scandinavians}?
+\emph{Journal of Political Economy} \textbf{2020}, \emph{128},
+1753--1788, doi:\href{https://doi.org/10.1086/705551}{10.1086/705551}.}
+
+\bibitem[\citeproctext]{ref-barr_effect_2020}
+\CSLLeftMargin{57. }%
+\CSLRightInline{Barr, A.; Miller, L. The Effect of Education, Income
+Inequality and Merit on Inequality Acceptance. \emph{Journal of Economic
+Psychology} \textbf{2020}, \emph{80}, 102276,
+doi:\href{https://doi.org/10.1016/j.joep.2020.102276}{10.1016/j.joep.2020.102276}.}
+
+\bibitem[\citeproctext]{ref-evans_strong_2018}
+\CSLLeftMargin{58. }%
+\CSLRightInline{Evans, M.D.R.; Kelley, J. Strong {Welfare States Do Not
+Intensify Public Support} for {Income Redistribution}, but {Even Reduce
+It} Among the {Prosperous}: {A Multilevel Analysis} of {Public Opinion}
+in 30 {Countries}. \emph{Societies} \textbf{2018}, \emph{8}, 105,
+doi:\href{https://doi.org/10.3390/soc8040105}{10.3390/soc8040105}.}
+
+\bibitem[\citeproctext]{ref-mikus_children_2019}
+\CSLLeftMargin{59. }%
+\CSLRightInline{Mikus, K.; Tieben, N.; Schober, P.S. Children's
+Conversion of Cultural Capital into Educational Success: The Symbolic
+and Skill-Generating Functions of Cultural Capital. \emph{British
+Journal of Sociology of Education} \textbf{2019}, \emph{41}, 197--2017,
+doi:\href{https://doi.org/10.1080/01425692.2019.1677454}{10.1080/01425692.2019.1677454}.}
+
+\bibitem[\citeproctext]{ref-jack_no_2016}
+\CSLLeftMargin{60. }%
+\CSLRightInline{Jack, A.A. ({No}) {Harm} in {Asking}: {Class}, {Acquired
+Cultural Capital}, and {Academic Engagement} at an {Elite University}.
+\emph{Sociology of Education} \textbf{2016}, \emph{89}, 1--19,
+doi:\href{https://doi.org/10.1177/0038040715614913}{10.1177/0038040715614913}.}
+
+\bibitem[\citeproctext]{ref-khan_privilege_2011}
+\CSLLeftMargin{61. }%
+\CSLRightInline{Khan, S. \emph{Privilege: {The Making} of an {Adolescent
+Elite} at {St}. {Paul}'s {School}}; Princeton University Press:
+Princeton, N.J., 2011;}
+
+\bibitem[\citeproctext]{ref-goudeau_hidden_2017}
+\CSLLeftMargin{62. }%
+\CSLRightInline{Goudeau, S.; Croizet, J.-C. Hidden {Advantages} and
+{Disadvantages} of {Social Class}: {How Classroom Settings Reproduce
+Social Inequality} by {Staging Unfair Comparison}. \emph{Psychological
+Science} \textbf{2017}, \emph{28}, 162--170,
+doi:\href{https://doi.org/10.1177/0956797616676600}{10.1177/0956797616676600}.}
+
+\bibitem[\citeproctext]{ref-darnon_where_2018}
+\CSLLeftMargin{63. }%
+\CSLRightInline{Darnon, C.; Wiederkehr, V.; Dompnier, B.; Martinot, D.
+{``{Where} There Is a Will, There Is a Way''}: {Belief} in School
+Meritocracy and the Social-Class Achievement Gap. \emph{British Journal
+of Social Psychology} \textbf{2018}, \emph{57}, 250--262,
+doi:\href{https://doi.org/10.1111/bjso.12214}{10.1111/bjso.12214}.}
+
+\bibitem[\citeproctext]{ref-resh_sense_2010}
+\CSLLeftMargin{64. }%
+\CSLRightInline{Resh, N. Sense of Justice about Grades in School: Is It
+Stratified Like Academic Achievement? \emph{Social Psychology of
+Education} \textbf{2010}, \emph{13}, 313--329,
+doi:\href{https://doi.org/10.1007/s11218-010-9117-z}{10.1007/s11218-010-9117-z}.}
+
+\bibitem[\citeproctext]{ref-resh_sense_2014}
+\CSLLeftMargin{65. }%
+\CSLRightInline{Resh, N.; Sabbagh, C. Sense of Justice in School and
+Civic Attitudes. \emph{Social Psychology of Education} \textbf{2014},
+\emph{17}, 51--72,
+doi:\href{https://doi.org/10.1007/s11218-013-9240-8}{10.1007/s11218-013-9240-8}.}
+
+\bibitem[\citeproctext]{ref-kohn_social_1963}
+\CSLLeftMargin{66. }%
+\CSLRightInline{Kohn, M.L. Social {Class} and {Parent-Child
+Relationships}: {An Interpretation}. \emph{American Journal of
+Sociology} \textbf{1963}, \emph{68}, 471--480,
+doi:\href{https://doi.org/10.1086/223403}{10.1086/223403}.}
+
+\bibitem[\citeproctext]{ref-kohn_class_1969}
+\CSLLeftMargin{67. }%
+\CSLRightInline{Kohn, M.L.; Schooler, C. Class, {Occupation}, and
+{Orientation}. \emph{American Sociological Review} \textbf{1969},
+\emph{34}, 659--678,
+doi:\href{https://doi.org/10.2307/2092303}{10.2307/2092303}.}
+
+\bibitem[\citeproctext]{ref-acemoglu_obedience_2021}
+\CSLLeftMargin{68. }%
+\CSLRightInline{Acemoglu, D.
+\emph{\href{https://doi.org/10.3386/w29125}{Obedience in the {Labor
+Market} and {Social Mobility}: {A Socio-Economic Approach}}}; National
+Bureau of Economic Research: Cambridge, MA, 2021; p. w29125;}
+
+\bibitem[\citeproctext]{ref-iacoviello_collectivism_2019}
+\CSLLeftMargin{69. }%
+\CSLRightInline{Iacoviello, V.; Lorenzi-Cioldi, F. Collectivism and
+{Individualism} in {Status Hierarchies}: {Socialization} and {Social
+Identity Explanations}. \emph{International Review of Social Psychology}
+\textbf{2019}, \emph{32}, 15,
+doi:\href{https://doi.org/10.5334/irsp.285}{10.5334/irsp.285}.}
+
+\bibitem[\citeproctext]{ref-almas_fairness_2017}
+\CSLLeftMargin{70. }%
+\CSLRightInline{Almås, I.; Cappelen, A.W.; Salvanes, K.G.; Sørensen,
+E.Ø.; Tungodden, B. Fairness and Family Background. \emph{Politics,
+Philosophy \& Economics} \textbf{2017}, \emph{16}, 117--131,
+doi:\href{https://doi.org/10.1177/1470594x15618966}{10.1177/1470594x15618966}.}
+
+\bibitem[\citeproctext]{ref-jonsson_institutional_2015}
+\CSLLeftMargin{71. }%
+\CSLRightInline{Jonsson, A.-C.; Beach, D. Institutional Discrimination:
+{Stereotypes} and Social Reproduction of {``Class''} in the {Swedish}
+Upper-Secondary School. \emph{Social Psychology of Education}
+\textbf{2015}, \emph{18}, 703--717,
+doi:\href{https://doi.org/10.1007/s11218-014-9279-1}{10.1007/s11218-014-9279-1}.}
+
+\bibitem[\citeproctext]{ref-jost_attitudinal_2000}
+\CSLLeftMargin{72. }%
+\CSLRightInline{Jost, J.T.; Burgess, D. Attitudinal {Ambivalence} and
+the {Conflict} Between {Group} and {System Justification Motives} in
+{Low Status Groups}. \emph{Personality and Social Psychology Bulletin}
+\textbf{2000}, \emph{26}, 293--305,
+doi:\href{https://doi.org/10.1177/0146167200265003}{10.1177/0146167200265003}.}
+
+\bibitem[\citeproctext]{ref-gottfried_autocorrelation_2022}
+\CSLLeftMargin{73. }%
+\CSLRightInline{Gottfried, J.; Ježek, S.; Králová, M.; Řiháček, T.
+Autocorrelation {Screening}: {A Potentially Efficient Method} for
+{Detecting Repetitive Response Patterns} in {Questionnaire Data}.
+\textbf{2022},
+doi:\href{https://doi.org/10.7275/VYXB-GT24}{10.7275/VYXB-GT24}.}
+
+\bibitem[\citeproctext]{ref-bell_fixed_2019}
+\CSLLeftMargin{74. }%
+\CSLRightInline{Bell, A.; Fairbrother, M.; Jones, K. Fixed and Random
+Effects Models: Making an Informed Choice. \emph{Quality \& Quantity}
+\textbf{2019}, \emph{53}, 1051--1074,
+doi:\href{https://doi.org/10.1007/s11135-018-0802-x}{10.1007/s11135-018-0802-x}.}
+
+\bibitem[\citeproctext]{ref-hox_multilevel_2010}
+\CSLLeftMargin{75. }%
+\CSLRightInline{Hox, J.; Moerbeek, M.; Schoot, R. van de
+\emph{\href{https://doi.org/10.4324/9780203852279}{Multilevel
+{Analysis}: {Techniques} and {Applications}, {Second Edition}}}; 2nd
+ed.; Routledge: New York, 2010; ISBN 978-0-203-85227-9.}
+
+\bibitem[\citeproctext]{ref-kim_socioeconomic_2020}
+\CSLLeftMargin{76. }%
+\CSLRightInline{Kim, C.-Y.; Gewirtz, S. Socioeconomic Inequality and
+Adjustments in Children's Perceptions of Their Agency as They Age in
+{South Korea}. \emph{Children \& Society} \textbf{2020}, \emph{34},
+371--391,
+doi:\href{https://doi.org/10.1111/chso.12372}{10.1111/chso.12372}.}
+
+\bibitem[\citeproctext]{ref-rizzo_children_2020}
+\CSLLeftMargin{77. }%
+\CSLRightInline{Rizzo, M.T.; Killen, M. Children's Evaluations of
+Individually and Structurally Based Inequalities: {The} Role of Status.
+\emph{Developmental Psychology} \textbf{2020}, \emph{56}, 2223--2235,
+doi:\href{https://doi.org/10.1037/dev0001118}{10.1037/dev0001118}.}
+
+\bibitem[\citeproctext]{ref-gonthier_parallel_2017}
+\CSLLeftMargin{78. }%
+\CSLRightInline{Gonthier, F. Parallel Publics? {Support} for Income
+Redistribution in Times of Economic Crisis. \emph{European Journal of
+Political Research} \textbf{2017}, \emph{56}, 92--114,
+doi:\href{https://doi.org/10.1111/1475-6765.12168}{10.1111/1475-6765.12168}.}
+
+\bibitem[\citeproctext]{ref-runst_does_2018}
+\CSLLeftMargin{79. }%
+\CSLRightInline{Runst, P. Does {Immigration Affect Demand} for
+{Redistribution}? -- {An Experimental Design}. \emph{German Economic
+Review} \textbf{2018}, \emph{19}, 383--400,
+doi:\href{https://doi.org/10.1111/geer.12133}{10.1111/geer.12133}.}
+
+\bibitem[\citeproctext]{ref-bullock_education_2021}
+\CSLLeftMargin{80. }%
+\CSLRightInline{Bullock, J.G. Education and {Attitudes} Toward
+{Redistribution} in the {United States}. \emph{British Journal of
+Political Science} \textbf{2021}, \emph{51}, 1230--1250,
+doi:\href{https://doi.org/10.1017/S0007123419000504}{10.1017/S0007123419000504}.}
+
+\end{CSLReferences}
+
+\newpage{}
+
+\section{Supplementary material}\label{supplementary-material}
+
+\subsection{Complete multilevel
+models}\label{complete-multilevel-models}
+
+\begin{table}
+
+\caption{\label{tbl-ordinal-health}Cumulative Link Multilevel Models for
+Justifying Differential Access to Health Care}
+
+\centering{
+
+\begin{center}
+\scalebox{0.7}{
+\begin{threeparttable}
+\begin{tabular}{l c c c c c c}
+\toprule
+ & Model 0 & Model 1 & Model 2 & Model 3 & Model 4 & Model 5 \\
+\midrule
+Strongly Disagree|Disagree                             & $-0.55^{***}$ & $1.23^{***}$  & $1.20^{***}$  & $1.24^{***}$  & $1.05^{***}$  & $1.04^{***}$  \\
+                                                       & $(0.04)$      & $(0.20)$      & $(0.20)$      & $(0.22)$      & $(0.23)$      & $(0.23)$      \\
+Disagree|Agree                                         & $0.68^{***}$  & $2.54^{***}$  & $2.52^{***}$  & $2.55^{***}$  & $2.37^{***}$  & $2.36^{***}$  \\
+                                                       & $(0.04)$      & $(0.20)$      & $(0.20)$      & $(0.22)$      & $(0.24)$      & $(0.24)$      \\
+Agree|Strongly Agree                                   & $2.25^{***}$  & $4.20^{***}$  & $4.18^{***}$  & $4.22^{***}$  & $4.03^{***}$  & $4.02^{***}$  \\
+                                                       & $(0.06)$      & $(0.21)$      & $(0.21)$      & $(0.23)$      & $(0.24)$      & $(0.24)$      \\
+School talent                                          &               & $0.23^{***}$  & $0.23^{***}$  & $0.23^{***}$  & $0.22^{***}$  & $0.22^{***}$  \\
+                                                       &               & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+School effort                                          &               & $-0.27^{***}$ & $-0.26^{***}$ & $-0.26^{***}$ & $-0.26^{***}$ & $-0.26^{***}$ \\
+                                                       &               & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
+Social talent                                          &               & $0.16^{***}$  & $0.16^{***}$  & $0.16^{***}$  & $0.16^{***}$  & $0.16^{***}$  \\
+                                                       &               & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      \\
+Social effort                                          &               & $0.11^{*}$    & $0.11^{*}$    & $0.11^{*}$    & $0.11^{*}$    & $0.11^{*}$    \\
+                                                       &               & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      \\
+Deservingness                                          &               & $0.52^{***}$  & $0.52^{***}$  & $0.52^{***}$  & $0.51^{***}$  & $0.51^{***}$  \\
+                                                       &               & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      \\
+Parental education (Ref.= Non-university)              &               &               &               &               &               &               \\
+                                                       &               &               &               &               &               &               \\
+\quad University or posgraduate                        &               &               & $0.05$        & $0.05$        & $0.06$        & $0.08$        \\
+                                                       &               &               & $(0.08)$      & $(0.08)$      & $(0.08)$      & $(0.08)$      \\
+\quad Missing                                          &               &               & $0.10$        & $0.10$        & $0.10$        & $0.11$        \\
+                                                       &               &               & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      \\
+More than 25 books (Ref.= Less than 25)                &               &               & $-0.16^{**}$  & $-0.16^{**}$  & $-0.14^{*}$   & $-0.13^{*}$   \\
+                                                       &               &               & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      \\
+Technology acces                                       &               &               &               & $0.00$        & $0.00$        & $0.00$        \\
+                                                       &               &               &               & $(0.01)$      & $(0.01)$      & $(0.01)$      \\
+No internet connectivity (Ref.= Internet connectivity) &               &               &               & $0.03$        & $0.01$        & $-0.00$       \\
+                                                       &               &               &               & $(0.06)$      & $(0.07)$      & $(0.07)$      \\
+Socioeconomic level (Ref.= Low)                        &               &               &               &               &               &               \\
+                                                       &               &               &               &               &               &               \\
+\quad SES Medium                                       &               &               &               &               & $0.04$        & $0.08$        \\
+                                                       &               &               &               &               & $(0.10)$      & $(0.10)$      \\
+\quad SES High                                         &               &               &               &               & $0.48^{**}$   & $0.71^{*}$    \\
+                                                       &               &               &               &               & $(0.16)$      & $(0.28)$      \\
+Achievement score (Ref.= Low)                          &               &               &               &               &               &               \\
+                                                       &               &               &               &               &               &               \\
+\quad Simce Medium                                     &               &               &               &               & $-0.31^{***}$ & $-0.29^{***}$ \\
+                                                       &               &               &               &               & $(0.08)$      & $(0.08)$      \\
+\quad Simce High                                       &               &               &               &               & $-0.61^{***}$ & $-0.56^{***}$ \\
+                                                       &               &               &               &               & $(0.09)$      & $(0.10)$      \\
+Prop. university level at school                       &               &               &               &               &               & $-0.34$       \\
+                                                       &               &               &               &               &               & $(0.36)$      \\
+School dependence (Ref.= Public)                       &               &               &               &               &               &               \\
+                                                       &               &               &               &               &               &               \\
+\quad Subsidized Private                               &               &               &               &               &               & $-0.05$       \\
+                                                       &               &               &               &               &               & $(0.08)$      \\
+\quad Private                                          &               &               &               &               &               & $-0.14$       \\
+                                                       &               &               &               &               &               & $(0.27)$      \\
+\midrule
+Log Likelihood                                         & $-6581.84$    & $-6314.90$    & $-6309.98$    & $-6309.70$    & $-6287.72$    & $-6286.73$    \\
+AIC                                                    & $13171.69$    & $12647.80$    & $12643.95$    & $12649.39$    & $12613.44$    & $12617.47$    \\
+BIC                                                    & $13197.87$    & $12706.71$    & $12722.50$    & $12747.58$    & $12737.80$    & $12761.47$    \\
+Num.obs                                                & $5144$        & $5144$        & $5144$        & $5144$        & $5144$        & $5144$        \\
+Num. groups: Schools                                   & $231$         & $231$         & $231$         & $231$         & $231$         & $231$         \\
+Var: Schools (Intercept)                               & $0.17$        & $0.11$        & $0.11$        & $0.11$        & $0.05$        & $0.05$        \\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[flushleft]
+\scriptsize{\item Note: Cells contain regression coefficients with standard errors in parentheses.
+$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$ \\ \item Source: own elaboration based on First Study of Civic Education 2017.}
+\end{tablenotes}
+\end{threeparttable}
 }
-
-%\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
-Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
-
-%\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study.
-You might choose to exclude this statement if the study did not require ethical approval.
-Please note that the Editorial Office might ask you for further information.
-Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans.
-OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals.
-OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.}
-
-%\informedconsent{Any research article describing a study involving humans should contain this statement.
-Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.
-You might also choose to exclude this statement if the study did not involve humans.
-Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves).
-Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}
-
-%\dataavailability{We encourage all authors of articles published in MDPI journals to share their research data.
-In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study.
-Where no new data were created, or where data is unavailable due to privacy or ethical restrictions, a statement is still required.
-Suggested Data Availability Statements are available in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}.}
-
-% Only for journal Nursing Reports
-%\publicinvolvement{Please describe how the public (patients, consumers, carers) were involved in the research.
-Consider reporting against the GRIPP2 (Guidance for Reporting Involvement of Patients and the Public) checklist.
-If the public were not involved in any aspect of the research add: ``No public involvement in any aspect of this research''.}
-
-% Only for journal Nursing Reports
-%\guidelinesstandards{Please add a statement indicating which reporting guideline was used when drafting the report.
-For example, ``This manuscript was drafted against the XXX (the full name of reporting guidelines and citation) for XXX (type of research) research''.
-A complete list of reporting guidelines can be accessed via the equator network: \url{https://www.equator-network.org/}.}
-
-%\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
-This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
-
-%\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflicts of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
-Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript; or in the decision to publish the results must be declared in this section.
-If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the results''.}
-
-\begin{adjustwidth}{-\extralength}{0cm}
-%\printendnotes[custom] % Un-comment to print a list of endnotes
-
-\reftitle{References}
-
-  \bibliography{input/bib/just-ineq-merit.bib}
-
-\PublishersNote{}
-\end{adjustwidth}\end{document}
+\caption{}
+\label{table:coefficients}
+\end{center}
+
+}
+
+\end{table}%
+
+\begin{table}
+
+\caption{\label{tbl-ordinal-pension}Cumulative Link Multilevel Models
+for Justifying Differential Access to Pensions}
+
+\centering{
+
+\begin{center}
+\scalebox{0.7}{
+\begin{threeparttable}
+\begin{tabular}{l c c c c c c}
+\toprule
+ & Model 0 & Model 1 & Model 2 & Model 3 & Model 4 & Model 5 \\
+\midrule
+Strongly Disagree|Disagree                             & $-0.89^{***}$ & $0.69^{***}$  & $0.71^{***}$  & $0.85^{***}$  & $0.67^{**}$   & $0.67^{**}$   \\
+                                                       & $(0.04)$      & $(0.20)$      & $(0.20)$      & $(0.22)$      & $(0.23)$      & $(0.23)$      \\
+Disagree|Agree                                         & $0.52^{***}$  & $2.18^{***}$  & $2.21^{***}$  & $2.35^{***}$  & $2.17^{***}$  & $2.17^{***}$  \\
+                                                       & $(0.04)$      & $(0.20)$      & $(0.20)$      & $(0.22)$      & $(0.23)$      & $(0.23)$      \\
+Agree|Strongly Agree                                   & $2.25^{***}$  & $4.01^{***}$  & $4.03^{***}$  & $4.18^{***}$  & $4.00^{***}$  & $4.00^{***}$  \\
+                                                       & $(0.05)$      & $(0.21)$      & $(0.21)$      & $(0.22)$      & $(0.24)$      & $(0.24)$      \\
+School talent                                          &               & $0.17^{***}$  & $0.17^{***}$  & $0.17^{***}$  & $0.16^{***}$  & $0.16^{***}$  \\
+                                                       &               & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+School effort                                          &               & $-0.25^{***}$ & $-0.25^{***}$ & $-0.26^{***}$ & $-0.25^{***}$ & $-0.25^{***}$ \\
+                                                       &               & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
+Social talent                                          &               & $0.13^{**}$   & $0.12^{**}$   & $0.12^{**}$   & $0.13^{**}$   & $0.13^{**}$   \\
+                                                       &               & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
+Social effort                                          &               & $0.27^{***}$  & $0.27^{***}$  & $0.27^{***}$  & $0.27^{***}$  & $0.27^{***}$  \\
+                                                       &               & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      \\
+Deservingness                                          &               & $0.38^{***}$  & $0.38^{***}$  & $0.38^{***}$  & $0.37^{***}$  & $0.37^{***}$  \\
+                                                       &               & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      \\
+Parental education (Ref.= Non-university)              &               &               &               &               &               &               \\
+                                                       &               &               &               &               &               &               \\
+\quad University or posgraduate                        &               &               & $0.16^{*}$    & $0.14$        & $0.12$        & $0.12$        \\
+                                                       &               &               & $(0.08)$      & $(0.08)$      & $(0.08)$      & $(0.08)$      \\
+\quad Missing                                          &               &               & $0.13^{*}$    & $0.12^{*}$    & $0.11$        & $0.11$        \\
+                                                       &               &               & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      \\
+More than 25 books (Ref.= Less than 25)                &               &               & $-0.12^{*}$   & $-0.13^{*}$   & $-0.13^{*}$   & $-0.13^{*}$   \\
+                                                       &               &               & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      \\
+Technology acces                                       &               &               &               & $0.02^{*}$    & $0.02$        & $0.02$        \\
+                                                       &               &               &               & $(0.01)$      & $(0.01)$      & $(0.01)$      \\
+No internet connectivity (Ref.= Internet connectivity) &               &               &               & $-0.01$       & $-0.03$       & $-0.03$       \\
+                                                       &               &               &               & $(0.06)$      & $(0.06)$      & $(0.06)$      \\
+Socioeconomic level (Ref.= Low)                        &               &               &               &               &               &               \\
+                                                       &               &               &               &               &               &               \\
+\quad SES Medium                                       &               &               &               &               & $0.00$        & $0.02$        \\
+                                                       &               &               &               &               & $(0.10)$      & $(0.10)$      \\
+\quad SES High                                         &               &               &               &               & $0.69^{***}$  & $0.72^{**}$   \\
+                                                       &               &               &               &               & $(0.15)$      & $(0.26)$      \\
+Achievement score (Ref.= Low)                          &               &               &               &               &               &               \\
+                                                       &               &               &               &               &               &               \\
+\quad Simce Medium                                     &               &               &               &               & $-0.21^{**}$  & $-0.19^{*}$   \\
+                                                       &               &               &               &               & $(0.07)$      & $(0.08)$      \\
+\quad Simce High                                       &               &               &               &               & $-0.53^{***}$ & $-0.53^{***}$ \\
+                                                       &               &               &               &               & $(0.08)$      & $(0.09)$      \\
+Prop. university level at school                       &               &               &               &               &               & $0.22$        \\
+                                                       &               &               &               &               &               & $(0.34)$      \\
+School dependence (Ref.= Public)                       &               &               &               &               &               &               \\
+                                                       &               &               &               &               &               &               \\
+\quad Subsidized Private                               &               &               &               &               &               & $-0.09$       \\
+                                                       &               &               &               &               &               & $(0.07)$      \\
+\quad Private                                          &               &               &               &               &               & $-0.15$       \\
+                                                       &               &               &               &               &               & $(0.25)$      \\
+\midrule
+Log Likelihood                                         & $-6737.05$    & $-6494.86$    & $-6490.03$    & $-6487.76$    & $-6463.24$    & $-6462.36$    \\
+AIC                                                    & $13482.11$    & $13007.72$    & $13004.05$    & $13005.53$    & $12964.49$    & $12968.73$    \\
+BIC                                                    & $13508.31$    & $13066.68$    & $13082.68$    & $13103.81$    & $13088.98$    & $13112.87$    \\
+Num.obs                                                & $5177$        & $5177$        & $5177$        & $5177$        & $5177$        & $5177$        \\
+Num. groups: Schools                                   & $232$         & $232$         & $232$         & $232$         & $232$         & $232$         \\
+Var: Schools (Intercept)                               & $0.11$        & $0.09$        & $0.09$        & $0.09$        & $0.04$        & $0.04$        \\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[flushleft]
+\scriptsize{\item Note: Cells contain regression coefficients with standard errors in parentheses.
+$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$ \\ \item Source: own elaboration based on First Study of Civic Education 2017.}
+\end{tablenotes}
+\end{threeparttable}
+}
+\caption{}
+\label{table:coefficients}
+\end{center}
+
+}
+
+\end{table}%
+
+\begin{table}
+
+\caption{\label{tbl-ordinal-educ}Cumulative Link Multilevel Models for
+Justifying Differential Access to Education}
+
+\centering{
+
+\begin{center}
+\scalebox{0.7}{
+\begin{threeparttable}
+\begin{tabular}{l c c c c c c}
+\toprule
+ & Model 0 & Model 1 & Model 2 & Model 3 & Model 4 & Model 5 \\
+\midrule
+Strongly Disagree|Disagree                             & $-0.93^{***}$ & $0.84^{***}$  & $0.87^{***}$  & $0.82^{***}$  & $0.75^{**}$   & $0.75^{**}$   \\
+                                                       & $(0.04)$      & $(0.20)$      & $(0.20)$      & $(0.22)$      & $(0.23)$      & $(0.23)$      \\
+Disagree|Agree                                         & $0.34^{***}$  & $2.19^{***}$  & $2.23^{***}$  & $2.18^{***}$  & $2.11^{***}$  & $2.11^{***}$  \\
+                                                       & $(0.04)$      & $(0.20)$      & $(0.20)$      & $(0.22)$      & $(0.23)$      & $(0.23)$      \\
+Agree|Strongly Agree                                   & $1.98^{***}$  & $3.93^{***}$  & $3.96^{***}$  & $3.92^{***}$  & $3.85^{***}$  & $3.85^{***}$  \\
+                                                       & $(0.05)$      & $(0.20)$      & $(0.21)$      & $(0.22)$      & $(0.24)$      & $(0.24)$      \\
+School talent                                          &               & $0.25^{***}$  & $0.25^{***}$  & $0.25^{***}$  & $0.25^{***}$  & $0.25^{***}$  \\
+                                                       &               & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+School effort                                          &               & $-0.24^{***}$ & $-0.23^{***}$ & $-0.23^{***}$ & $-0.23^{***}$ & $-0.23^{***}$ \\
+                                                       &               & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
+Social talent                                          &               & $0.14^{**}$   & $0.14^{**}$   & $0.14^{**}$   & $0.15^{***}$  & $0.15^{***}$  \\
+                                                       &               & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
+Social effort                                          &               & $0.15^{**}$   & $0.15^{**}$   & $0.15^{**}$   & $0.15^{**}$   & $0.15^{**}$   \\
+                                                       &               & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      \\
+Deservingness                                          &               & $0.44^{***}$  & $0.44^{***}$  & $0.44^{***}$  & $0.43^{***}$  & $0.43^{***}$  \\
+                                                       &               & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      \\
+Parental education (Ref.= Non-university)              &               &               &               &               &               &               \\
+                                                       &               &               &               &               &               &               \\
+\quad University or posgraduate                        &               &               & $0.09$        & $0.09$        & $0.06$        & $0.07$        \\
+                                                       &               &               & $(0.08)$      & $(0.08)$      & $(0.08)$      & $(0.08)$      \\
+\quad Missing                                          &               &               & $0.13^{*}$    & $0.13^{*}$    & $0.12^{*}$    & $0.12^{*}$    \\
+                                                       &               &               & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      \\
+More than 25 books (Ref.= Less than 25)                &               &               & $-0.07$       & $-0.07$       & $-0.07$       & $-0.07$       \\
+                                                       &               &               & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      \\
+Technology acces                                       &               &               &               & $-0.00$       & $-0.01$       & $-0.01$       \\
+                                                       &               &               &               & $(0.01)$      & $(0.01)$      & $(0.01)$      \\
+No internet connectivity (Ref.= Internet connectivity) &               &               &               & $-0.04$       & $-0.04$       & $-0.05$       \\
+                                                       &               &               &               & $(0.06)$      & $(0.06)$      & $(0.06)$      \\
+Socioeconomic level (Ref.= Low)                        &               &               &               &               &               &               \\
+                                                       &               &               &               &               &               &               \\
+\quad SES Medium                                       &               &               &               &               & $0.08$        & $0.08$        \\
+                                                       &               &               &               &               & $(0.10)$      & $(0.10)$      \\
+\quad SES High                                         &               &               &               &               & $0.75^{***}$  & $0.93^{***}$  \\
+                                                       &               &               &               &               & $(0.16)$      & $(0.28)$      \\
+Achievement score (Ref.= Low)                          &               &               &               &               &               &               \\
+                                                       &               &               &               &               &               &               \\
+\quad Simce Medium                                     &               &               &               &               & $-0.16^{*}$   & $-0.16^{*}$   \\
+                                                       &               &               &               &               & $(0.08)$      & $(0.08)$      \\
+\quad Simce High                                       &               &               &               &               & $-0.47^{***}$ & $-0.46^{***}$ \\
+                                                       &               &               &               &               & $(0.09)$      & $(0.10)$      \\
+Prop. university level at school                       &               &               &               &               &               & $-0.17$       \\
+                                                       &               &               &               &               &               & $(0.35)$      \\
+School dependence (Ref.= Public)                       &               &               &               &               &               &               \\
+                                                       &               &               &               &               &               &               \\
+\quad Subsidized Private                               &               &               &               &               &               & $0.04$        \\
+                                                       &               &               &               &               &               & $(0.08)$      \\
+\quad Private                                          &               &               &               &               &               & $-0.14$       \\
+                                                       &               &               &               &               &               & $(0.27)$      \\
+\midrule
+Log Likelihood                                         & $-6846.16$    & $-6603.77$    & $-6600.80$    & $-6600.55$    & $-6582.51$    & $-6582.06$    \\
+AIC                                                    & $13700.33$    & $13225.55$    & $13225.60$    & $13231.10$    & $13203.03$    & $13208.12$    \\
+BIC                                                    & $13726.52$    & $13284.48$    & $13304.18$    & $13329.33$    & $13327.45$    & $13352.18$    \\
+Num.obs                                                & $5159$        & $5159$        & $5159$        & $5159$        & $5159$        & $5159$        \\
+Num. groups: Schools                                   & $231$         & $231$         & $231$         & $231$         & $231$         & $231$         \\
+Var: Schools (Intercept)                               & $0.13$        & $0.10$        & $0.10$        & $0.10$        & $0.05$        & $0.05$        \\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[flushleft]
+\scriptsize{\item Note: Cells contain regression coefficients with standard errors in parentheses.
+$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$ \\ \item Source: own elaboration based on First Study of Civic Education 2017.}
+\end{tablenotes}
+\end{threeparttable}
+}
+\caption{}
+\label{table:coefficients}
+\end{center}
+
+}
+
+\end{table}%
+
+\begin{table}
+
+\caption{\label{tbl-lineal-reg-complete}Linear mixed-effects models for
+meritocracy perception and market justice preferences}
+
+\centering{
+
+\begin{center}
+\scalebox{0.7}{
+\begin{threeparttable}
+\begin{tabular}{l c c c c c c}
+\toprule
+ & Model 0 & Model 1 & Model 2 & Model 3 & Model 4 & Model 5 \\
+\midrule
+Constant                                               & $2.17^{***}$ & $1.34^{***}$  & $1.33^{***}$  & $1.30^{***}$  & $1.38^{***}$  & $1.39^{***}$  \\
+                                                       & $(0.02)$     & $(0.09)$      & $(0.09)$      & $(0.10)$      & $(0.10)$      & $(0.10)$      \\
+School talent                                          &              & $0.11^{***}$  & $0.11^{***}$  & $0.11^{***}$  & $0.10^{***}$  & $0.10^{***}$  \\
+                                                       &              & $(0.01)$      & $(0.01)$      & $(0.01)$      & $(0.01)$      & $(0.01)$      \\
+School effort                                          &              & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ \\
+                                                       &              & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Social talent                                          &              & $0.07^{***}$  & $0.07^{***}$  & $0.07^{***}$  & $0.07^{***}$  & $0.07^{***}$  \\
+                                                       &              & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Social effort                                          &              & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  \\
+                                                       &              & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Deservingness                                          &              & $0.20^{***}$  & $0.20^{***}$  & $0.20^{***}$  & $0.20^{***}$  & $0.20^{***}$  \\
+                                                       &              & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Parental education (Ref.= Non-university)              &              &               &               &               &               &               \\
+                                                       &              &               &               &               &               &               \\
+\quad University or posgraduate                        &              &               & $0.05$        & $0.05$        & $0.04$        & $0.05$        \\
+                                                       &              &               & $(0.03)$      & $(0.03)$      & $(0.04)$      & $(0.04)$      \\
+\quad Missing                                          &              &               & $0.06^{*}$    & $0.06^{*}$    & $0.06^{*}$    & $0.06^{*}$    \\
+                                                       &              &               & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+More than 25 books (Ref.= Less than 25)                &              &               & $-0.05$       & $-0.05$       & $-0.04$       & $-0.04$       \\
+                                                       &              &               & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+Technology acces                                       &              &               &               & $0.00$        & $0.00$        & $0.00$        \\
+                                                       &              &               &               & $(0.01)$      & $(0.01)$      & $(0.01)$      \\
+No internet connectivity (Ref.= Internet connectivity) &              &               &               & $0.00$        & $-0.01$       & $-0.01$       \\
+                                                       &              &               &               & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+Socioeconomic level (Ref.= Low)                        &              &               &               &               &               &               \\
+                                                       &              &               &               &               &               &               \\
+\quad SES Medium                                       &              &               &               &               & $0.01$        & $0.01$        \\
+                                                       &              &               &               &               & $(0.04)$      & $(0.05)$      \\
+\quad SES High                                         &              &               &               &               & $0.29^{***}$  & $0.37^{**}$   \\
+                                                       &              &               &               &               & $(0.07)$      & $(0.13)$      \\
+Achievement score (Ref.= Low)                          &              &               &               &               &               &               \\
+                                                       &              &               &               &               &               &               \\
+\quad Simce Medium                                     &              &               &               &               & $-0.12^{***}$ & $-0.11^{**}$  \\
+                                                       &              &               &               &               & $(0.04)$      & $(0.04)$      \\
+\quad Simce High                                       &              &               &               &               & $-0.27^{***}$ & $-0.26^{***}$ \\
+                                                       &              &               &               &               & $(0.04)$      & $(0.04)$      \\
+School dependence (Ref.= Public)                       &              &               &               &               &               &               \\
+                                                       &              &               &               &               &               &               \\
+\quad Prop. university level at school                 &              &               &               &               &               & $-0.05$       \\
+                                                       &              &               &               &               &               & $(0.16)$      \\
+\quad Subsidized Private                               &              &               &               &               &               & $-0.01$       \\
+                                                       &              &               &               &               &               & $(0.04)$      \\
+\quad Private                                          &              &               &               &               &               & $-0.08$       \\
+                                                       &              &               &               &               &               & $(0.13)$      \\
+\midrule
+AIC                                                    & $12967.40$   & $12422.86$    & $12436.31$    & $12459.04$    & $12437.84$    & $12452.43$    \\
+BIC                                                    & $12986.98$   & $12475.07$    & $12508.10$    & $12550.41$    & $12555.32$    & $12589.48$    \\
+Log Likelihood                                         & $-6480.70$   & $-6203.43$    & $-6207.16$    & $-6215.52$    & $-6200.92$    & $-6205.21$    \\
+Num.obs                                                & $5047$       & $5047$        & $5047$        & $5047$        & $5047$        & $5047$        \\
+Num. groups: Schools                                   & $231$        & $231$         & $231$         & $231$         & $231$         & $231$         \\
+Var: Schools (Intercept)                               & $0.03$       & $0.02$        & $0.02$        & $0.02$        & $0.01$        & $0.01$        \\
+Var: Residual                                          & $0.74$       & $0.66$        & $0.66$        & $0.66$        & $0.66$        & $0.66$        \\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[flushleft]
+\scriptsize{\item Note: Cells contain regression coefficients with standard errors in parentheses.
+$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$ \\ \item Source: own elaboration based on First Study of Civic Education 2017.}
+\end{tablenotes}
+\end{threeparttable}
+}
+\caption{}
+\label{table:coefficients}
+\end{center}
+
+}
+
+\end{table}%
+
+\begin{table}
+
+\caption{\label{tbl-interac-educparent}Linear mixed-effects models for
+meritocracy perception, parental education and market justice
+preferences}
+
+\centering{
+
+\begin{center}
+\scalebox{0.7}{
+\begin{threeparttable}
+\begin{tabular}{l c c c c c}
+\toprule
+ & Model 1 & Model 2 & Model 3 & Model 4 & Model 5 \\
+\midrule
+Constant                                            & $1.35^{***}$  & $1.31^{***}$  & $1.34^{***}$  & $1.37^{***}$  & $1.38^{***}$  \\
+                                                    & $(0.10)$      & $(0.10)$      & $(0.10)$      & $(0.12)$      & $(0.10)$      \\
+School talent                                       & $0.10^{***}$  & $0.10^{***}$  & $0.10^{***}$  & $0.10^{***}$  & $0.11^{***}$  \\
+                                                    & $(0.01)$      & $(0.01)$      & $(0.01)$      & $(0.01)$      & $(0.02)$      \\
+School effort                                       & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ \\
+                                                    & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.03)$      & $(0.02)$      \\
+Social talent                                       & $0.07^{***}$  & $0.08^{***}$  & $0.09^{***}$  & $0.08^{***}$  & $0.08^{***}$  \\
+                                                    & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Social effort                                       & $0.09^{***}$  & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  \\
+                                                    & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Deservingness                                       & $0.20^{***}$  & $0.22^{***}$  & $0.20^{***}$  & $0.20^{***}$  & $0.20^{***}$  \\
+                                                    & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Parental education (Ref.= Non-university)           &               &               &               &               &               \\
+                                                    &               &               &               &               &               \\
+\quad University or posgraduate                     & $0.26^{*}$    & $0.35^{***}$  & $0.25^{*}$    & $0.10$        & $0.15$        \\
+                                                    & $(0.11)$      & $(0.10)$      & $(0.11)$      & $(0.23)$      & $(0.12)$      \\
+\quad Missing                                       & $0.03$        & $0.13$        & $0.07$        & $0.08$        & $0.03$        \\
+                                                    & $(0.09)$      & $(0.09)$      & $(0.09)$      & $(0.16)$      & $(0.10)$      \\
+More than 25 books (Ref.= Less than 25)             & $-0.04$       & $-0.04$       & $-0.04$       & $-0.04$       & $-0.04$       \\
+                                                    & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+SES High (Ref.= Low)                                & $-0.06$       & $-0.06$       & $-0.06$       & $-0.06$       & $-0.06$       \\
+                                                    & $(0.06)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      \\
+Achievement score (Ref.= Low)                       &               &               &               &               &               \\
+                                                    &               &               &               &               &               \\
+\quad Simce Medium                                  & $-0.11^{**}$  & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ \\
+                                                    & $(0.04)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+\quad Simce High                                    & $-0.24^{***}$ & $-0.24^{***}$ & $-0.24^{***}$ & $-0.24^{***}$ & $-0.24^{***}$ \\
+                                                    & $(0.05)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
+Parental education x Meritocracy                    &               &               &               &               &               \\
+                                                    &               &               &               &               &               \\
+\quad University or posgraduate * Social effort     & $-0.08^{*}$   &               &               &               &               \\
+                                                    & $(0.04)$      &               &               &               &               \\
+\quad Missing * Social effort                       & $0.01$        &               &               &               &               \\
+                                                    & $(0.03)$      &               &               &               &               \\
+\quad University or posgraduate * Deservingness     &               & $-0.12^{**}$  &               &               &               \\
+                                                    &               & $(0.04)$      &               &               &               \\
+\quad Missing * Deservingness                       &               & $-0.02$       &               &               &               \\
+                                                    &               & $(0.03)$      &               &               &               \\
+\quad University or posgraduate * Social talent     &               &               & $-0.07$       &               &               \\
+                                                    &               &               & $(0.04)$      &               &               \\
+\quad Missing * Social talent                       &               &               & $0.00$        &               &               \\
+                                                    &               &               & $(0.03)$      &               &               \\
+\quad University or posgraduate * School effort     &               &               &               & $-0.01$       &               \\
+                                                    &               &               &               & $(0.06)$      &               \\
+\quad Missing * School effort                       &               &               &               & $-0.00$       &               \\
+                                                    &               &               &               & $(0.04)$      &               \\
+\quad University or posgraduate * School talent     &               &               &               &               & $-0.03$       \\
+                                                    &               &               &               &               & $(0.04)$      \\
+\quad Missing * School talent                       &               &               &               &               & $0.01$        \\
+                                                    &               &               &               &               & $(0.03)$      \\
+\midrule
+Controls                                            & Yes           & Yes           & Yes           & Yes           & Yes           \\
+AIC                                                 & $12472.82$    & $12480.94$    & $12486.54$    & $12489.10$    & $12489.36$    \\
+BIC                                                 & $12649.03$    & $12657.15$    & $12662.76$    & $12665.31$    & $12665.57$    \\
+Log Likelihood                                      & $-6209.41$    & $-6213.47$    & $-6216.27$    & $-6217.55$    & $-6217.68$    \\
+Num.obs                                             & $5047$        & $5047$        & $5047$        & $5047$        & $5047$        \\
+Num. groups: Schools                                & $231$         & $231$         & $231$         & $231$         & $231$         \\
+Var: Schools (Intercept)                            & $0.01$        & $0.00$        & $0.00$        & $0.00$        & $0.00$        \\
+Var: Schools University or Postgraduate             & $0.00$        & $0.01$        & $0.02$        & $0.02$        & $0.02$        \\
+Var: Schools Missing                                & $0.00$        & $0.02$        & $0.02$        & $0.02$        & $0.02$        \\
+Cov: Schools (Intercept) University or Postgraduate & $0.00$        & $0.00$        & $0.00$        & $0.00$        & $0.00$        \\
+Cov: Schools (Intercept) Missing                    & $0.00$        & $0.00$        & $0.00$        & $0.00$        & $0.00$        \\
+Cov: Schools University or Postgraduate, Missing    & $0.00$        & $0.02$        & $0.02$        & $0.02$        & $0.02$        \\
+Var: Residual                                       & $0.66$        & $0.67$        & $0.67$        & $0.67$        & $0.67$        \\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[flushleft]
+\scriptsize{\item Note: Cells contain regression coefficients with standard errors in parentheses.
+Control variables are included.
+$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$ \\ \item Source: own elaboration based on First Study of Civic Education 2017.}
+\end{tablenotes}
+\end{threeparttable}
+}
+\caption{}
+\label{table:coefficients}
+\end{center}
+
+}
+
+\end{table}%
+
+\begin{table}
+
+\caption{\label{tbl-interac-books}Linear mixed-effects models for
+meritocracy perception, books at home and market justice preferences}
+
+\centering{
+
+\begin{center}
+\scalebox{0.7}{
+\begin{threeparttable}
+\begin{tabular}{l c c c c c}
+\toprule
+ & Model 1 & Model 2 & Model 3 & Model 4 & Model 5 \\
+\midrule
+Constant                                    & $1.29^{***}$  & $1.34^{***}$  & $1.34^{***}$  & $1.31^{***}$  & $1.35^{***}$  \\
+                                            & $(0.10)$      & $(0.10)$      & $(0.10)$      & $(0.11)$      & $(0.10)$      \\
+School talent                               & $0.10^{***}$  & $0.10^{***}$  & $0.10^{***}$  & $0.10^{***}$  & $0.11^{***}$  \\
+                                            & $(0.01)$      & $(0.01)$      & $(0.01)$      & $(0.01)$      & $(0.02)$      \\
+School effort                               & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ & $-0.10^{***}$ & $-0.12^{***}$ \\
+                                            & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Social talent                               & $0.07^{***}$  & $0.07^{***}$  & $0.09^{***}$  & $0.07^{***}$  & $0.07^{***}$  \\
+                                            & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Social effort                               & $0.11^{***}$  & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  \\
+                                            & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Deservingness                               & $0.20^{***}$  & $0.21^{***}$  & $0.20^{***}$  & $0.20^{***}$  & $0.20^{***}$  \\
+                                            & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Parental education (Ref.= Non-university)   &               &               &               &               &               \\
+                                            &               &               &               &               &               \\
+\quad University or posgraduate             & $0.05$        & $0.05$        & $0.05$        & $0.05$        & $0.05$        \\
+                                            & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
+\quad Missing                               & $0.06^{*}$    & $0.07^{*}$    & $0.07^{*}$    & $0.07^{*}$    & $0.07^{*}$    \\
+                                            & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+More than 25 books (Ref.= Less than 25)     & $0.17^{*}$    & $0.05$        & $0.03$        & $0.12$        & $-0.00$       \\
+                                            & $(0.08)$      & $(0.08)$      & $(0.08)$      & $(0.15)$      & $(0.09)$      \\
+SES High (Ref.= Low)                        & $-0.06$       & $-0.06$       & $-0.06$       & $-0.06$       & $-0.06$       \\
+                                            & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      \\
+Achievement score (Ref.= Low)               &               &               &               &               &               \\
+                                            &               &               &               &               &               \\
+\quad Simce Medium                          & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ \\
+                                            & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+\quad Simce High                            & $-0.24^{***}$ & $-0.24^{***}$ & $-0.24^{***}$ & $-0.24^{***}$ & $-0.24^{***}$ \\
+                                            & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
+Books at home x Meritocracy                 &               &               &               &               &               \\
+                                            &               &               &               &               &               \\
+\quad More than 25 books * Social effort    & $-0.08^{**}$  &               &               &               &               \\
+                                            & $(0.03)$      &               &               &               &               \\
+\quad More than 25 books * Deservingness    &               & $-0.04$       &               &               &               \\
+                                            &               & $(0.03)$      &               &               &               \\
+\quad More than 25 books * Social talent    &               &               & $-0.03$       &               &               \\
+                                            &               &               & $(0.03)$      &               &               \\
+\quad More than 25 books * School effort    &               &               &               & $-0.05$       &               \\
+                                            &               &               &               & $(0.04)$      &               \\
+\quad More than 25 books * School talent    &               &               &               &               & $-0.01$       \\
+                                            &               &               &               &               & $(0.03)$      \\
+\midrule
+Controls                                    & Yes           & Yes           & Yes           & Yes           & Yes           \\
+AIC                                         & $12462.52$    & $12469.09$    & $12469.61$    & $12468.65$    & $12470.35$    \\
+BIC                                         & $12612.63$    & $12619.20$    & $12619.72$    & $12618.76$    & $12620.46$    \\
+Log Likelihood                              & $-6208.26$    & $-6211.54$    & $-6211.81$    & $-6211.32$    & $-6212.18$    \\
+Num.obs                                     & $5047$        & $5047$        & $5047$        & $5047$        & $5047$        \\
+Num. groups: Schools                        & $231$         & $231$         & $231$         & $231$         & $231$         \\
+Var: Schools (Intercept)                    & $0.00$        & $0.00$        & $0.00$        & $0.00$        & $0.00$        \\
+Var: Schools More than 25 books             & $0.03$        & $0.03$        & $0.03$        & $0.03$        & $0.03$        \\
+Cov: Schools (Intercept) More than 25 books & $0.00$        & $0.00$        & $0.00$        & $0.00$        & $0.00$        \\
+Var: Residual                               & $0.67$        & $0.67$        & $0.67$        & $0.67$        & $0.67$        \\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[flushleft]
+\scriptsize{\item Note: Cells contain regression coefficients with standard errors in parentheses.
+Control variables are included.
+$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$ \\ \item Source: own elaboration based on First Study of Civic Education 2017.}
+\end{tablenotes}
+\end{threeparttable}
+}
+\caption{}
+\label{table:coefficients}
+\end{center}
+
+}
+
+\end{table}%
+
+\begin{table}
+
+\caption{\label{tbl-interac-schoolses}Linear mixed-effects models for
+meritocracy perception, school socioeconomic status and market justice
+preferences}
+
+\centering{
+
+\begin{center}
+\scalebox{0.7}{
+\begin{threeparttable}
+\begin{tabular}{l c c c c c}
+\toprule
+ & Model 1 & Model 2 & Model 3 & Model 4 & Model 5 \\
+\midrule
+Constant                                  & $1.32^{***}$  & $1.31^{***}$  & $1.32^{***}$  & $1.37^{***}$  & $1.38^{***}$  \\
+                                          & $(0.10)$      & $(0.10)$      & $(0.10)$      & $(0.11)$      & $(0.10)$      \\
+School talent                             & $0.10^{***}$  & $0.10^{***}$  & $0.10^{***}$  & $0.10^{***}$  & $0.10^{***}$  \\
+                                          & $(0.01)$      & $(0.01)$      & $(0.01)$      & $(0.01)$      & $(0.02)$      \\
+School effort                             & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ & $-0.12^{***}$ \\
+                                          & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Social talent                             & $0.07^{***}$  & $0.07^{***}$  & $0.09^{***}$  & $0.07^{***}$  & $0.07^{***}$  \\
+                                          & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Social effort                             & $0.10^{***}$  & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  \\
+                                          & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Deservingness                             & $0.20^{***}$  & $0.22^{***}$  & $0.20^{***}$  & $0.20^{***}$  & $0.20^{***}$  \\
+                                          & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Parental education (Ref.= Non-university) &               &               &               &               &               \\
+                                          &               &               &               &               &               \\
+\quad University or posgraduate           & $0.05$        & $0.05$        & $0.05$        & $0.05$        & $0.05$        \\
+                                          & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
+\quad Missing                             & $0.06^{*}$    & $0.07^{*}$    & $0.07^{*}$    & $0.07^{*}$    & $0.07^{*}$    \\
+                                          & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+More than 25 books (Ref.= Less than 25)   & $-0.04$       & $-0.04$       & $-0.04$       & $-0.04$       & $-0.04$       \\
+                                          & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+SES High (Ref.= Low)                      & $0.10$        & $0.13$        & $0.08$        & $-0.07$       & $-0.08$       \\
+                                          & $(0.10)$      & $(0.09)$      & $(0.10)$      & $(0.17)$      & $(0.10)$      \\
+Achievement score (Ref.= Low)             &               &               &               &               &               \\
+                                          &               &               &               &               &               \\
+\quad Simce Medium                        & $-0.11^{**}$  & $-0.11^{**}$  & $-0.11^{**}$  & $-0.11^{**}$  & $-0.11^{**}$  \\
+                                          & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
+\quad Simce High                          & $-0.25^{***}$ & $-0.25^{***}$ & $-0.24^{***}$ & $-0.24^{***}$ & $-0.24^{***}$ \\
+                                          & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      & $(0.05)$      \\
+School SES x Meritocracy                  &               &               &               &               &               \\
+                                          &               &               &               &               &               \\
+\quad SES High * Social effort            & $-0.06^{*}$   &               &               &               &               \\
+                                          & $(0.03)$      &               &               &               &               \\
+\quad SES High * Deservingness            &               & $-0.08^{**}$  &               &               &               \\
+                                          &               & $(0.03)$      &               &               &               \\
+\quad SES High * Social talent            &               &               & $-0.06$       &               &               \\
+                                          &               &               & $(0.03)$      &               &               \\
+\quad SES High * School effort            &               &               &               & $0.00$        &               \\
+                                          &               &               &               & $(0.04)$      &               \\
+\quad SES High * School talent            &               &               &               &               & $0.00$        \\
+                                          &               &               &               &               & $(0.03)$      \\
+\midrule
+Controls                                  & Yes           & Yes           & Yes           & Yes           & Yes           \\
+AIC                                       & $12461.41$    & $12458.82$    & $12462.31$    & $12465.17$    & $12465.95$    \\
+BIC                                       & $12611.52$    & $12608.93$    & $12612.42$    & $12615.28$    & $12616.06$    \\
+Log Likelihood                            & $-6207.71$    & $-6206.41$    & $-6208.15$    & $-6209.59$    & $-6209.97$    \\
+Num.obs                                   & $5047$        & $5047$        & $5047$        & $5047$        & $5047$        \\
+Num. groups: Schools                      & $231$         & $231$         & $231$         & $231$         & $231$         \\
+Var: Schools (Intercept)                  & $0.02$        & $0.02$        & $0.02$        & $0.02$        & $0.02$        \\
+Var: Schools SES High                     & $0.02$        & $0.02$        & $0.02$        & $0.01$        & $0.02$        \\
+Cov: Schools (Intercept) SES High         & $-0.01$       & $-0.01$       & $-0.01$       & $-0.01$       & $-0.01$       \\
+Var: Residual                             & $0.66$        & $0.66$        & $0.66$        & $0.66$        & $0.66$        \\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[flushleft]
+\scriptsize{\item Note: Cells contain regression coefficients with standard errors in parentheses.
+Control variables are included.
+$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$ \\ \item Source: own elaboration based on First Study of Civic Education 2017.}
+\end{tablenotes}
+\end{threeparttable}
+}
+\caption{}
+\label{table:coefficients}
+\end{center}
+
+}
+
+\end{table}%
+
+\begin{table}
+
+\caption{\label{tbl-interac-schoolperfo}Linear mixed-effects models for
+meritocracy perception, school perfomance status and market justice
+preferences}
+
+\centering{
+
+\begin{center}
+\scalebox{0.7}{
+\begin{threeparttable}
+\begin{tabular}{l c c c c c}
+\toprule
+ & Model 1 & Model 2 & Model 3 & Model 4 & Model 5 \\
+\midrule
+Constant                                  & $1.28^{***}$  & $1.22^{***}$  & $1.30^{***}$  & $1.43^{***}$  & $1.37^{***}$  \\
+                                          & $(0.11)$      & $(0.11)$      & $(0.11)$      & $(0.14)$      & $(0.11)$      \\
+School talent                             & $0.10^{***}$  & $0.10^{***}$  & $0.10^{***}$  & $0.10^{***}$  & $0.11^{***}$  \\
+                                          & $(0.01)$      & $(0.01)$      & $(0.01)$      & $(0.01)$      & $(0.03)$      \\
+School effort                             & $-0.11^{***}$ & $-0.11^{***}$ & $-0.11^{***}$ & $-0.13^{***}$ & $-0.12^{***}$ \\
+                                          & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.03)$      & $(0.02)$      \\
+Social talent                             & $0.07^{***}$  & $0.07^{***}$  & $0.10^{***}$  & $0.07^{***}$  & $0.07^{***}$  \\
+                                          & $(0.02)$      & $(0.02)$      & $(0.03)$      & $(0.02)$      & $(0.02)$      \\
+Social effort                             & $0.12^{***}$  & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  & $0.08^{***}$  \\
+                                          & $(0.03)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Deservingness                             & $0.20^{***}$  & $0.25^{***}$  & $0.20^{***}$  & $0.20^{***}$  & $0.20^{***}$  \\
+                                          & $(0.02)$      & $(0.03)$      & $(0.02)$      & $(0.02)$      & $(0.02)$      \\
+Parental education (Ref.= Non-university) &               &               &               &               &               \\
+                                          &               &               &               &               &               \\
+\quad University or posgraduate           & $0.05$        & $0.05$        & $0.05$        & $0.05$        & $0.05$        \\
+                                          & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      & $(0.04)$      \\
+\quad Missing                             & $0.06^{*}$    & $0.07^{*}$    & $0.06^{*}$    & $0.07^{*}$    & $0.07^{*}$    \\
+                                          & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+More than 25 books (Ref.= Less than 25)   & $-0.04$       & $-0.04$       & $-0.04$       & $-0.04$       & $-0.04$       \\
+                                          & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      & $(0.03)$      \\
+SES High (Ref.= Low)                      & $-0.06$       & $-0.06$       & $-0.06$       & $-0.05$       & $-0.05$       \\
+                                          & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      & $(0.06)$      \\
+Achievement score (Ref.= Low)             &               &               &               &               &               \\
+                                          &               &               &               &               &               \\
+\quad Simce Medium                        & $-0.00$       & $0.10$        & $-0.01$       & $-0.23$       & $-0.12$       \\
+                                          & $(0.10)$      & $(0.10)$      & $(0.10)$      & $(0.17)$      & $(0.11)$      \\
+\quad Simce High                          & $-0.07$       & $-0.01$       & $-0.12$       & $-0.27$       & $-0.19$       \\
+                                          & $(0.10)$      & $(0.10)$      & $(0.11)$      & $(0.19)$      & $(0.12)$      \\
+School perfomance (SIMCE) x Meritocracy   &               &               &               &               &               \\
+                                          &               &               &               &               &               \\
+\quad Simce Medium * Social effort        & $-0.04$       &               &               &               &               \\
+                                          & $(0.03)$      &               &               &               &               \\
+\quad Simce High * Social effort          & $-0.06$       &               &               &               &               \\
+                                          & $(0.03)$      &               &               &               &               \\
+\quad Simce Medium * Deservingness        &               & $-0.08^{*}$   &               &               &               \\
+                                          &               & $(0.03)$      &               &               &               \\
+\quad Simce High * Deservingness          &               & $-0.09^{*}$   &               &               &               \\
+                                          &               & $(0.03)$      &               &               &               \\
+\quad Simce Medium * Social talent        &               &               & $-0.04$       &               &               \\
+                                          &               &               & $(0.03)$      &               &               \\
+\quad Simce High * Social talent          &               &               & $-0.04$       &               &               \\
+                                          &               &               & $(0.03)$      &               &               \\
+\quad Simce Medium * School effort        &               &               &               & $0.03$        &               \\
+                                          &               &               &               & $(0.05)$      &               \\
+\quad Simce High * School effort          &               &               &               & $0.01$        &               \\
+                                          &               &               &               & $(0.05)$      &               \\
+\quad Simce Medium * School talent        &               &               &               &               & $0.00$        \\
+                                          &               &               &               &               & $(0.03)$      \\
+\quad Simce High * School talent          &               &               &               &               & $-0.02$       \\
+                                          &               &               &               &               & $(0.04)$      \\
+\midrule
+Controls                                  & Yes           & Yes           & Yes           & Yes           & Yes           \\
+AIC                                       & $12473.25$    & $12468.62$    & $12474.86$    & $12474.76$    & $12476.32$    \\
+BIC                                       & $12649.46$    & $12644.84$    & $12651.08$    & $12650.98$    & $12652.54$    \\
+Log Likelihood                            & $-6209.62$    & $-6207.31$    & $-6210.43$    & $-6210.38$    & $-6211.16$    \\
+Num.obs                                   & $5047$        & $5047$        & $5047$        & $5047$        & $5047$        \\
+Num. groups: Schools                      & $231$         & $231$         & $231$         & $231$         & $231$         \\
+Var: Schools (Intercept)                  & $0.01$        & $0.01$        & $0.01$        & $0.01$        & $0.01$        \\
+Var: Schools Simce Medium                 & $0.00$        & $0.00$        & $0.00$        & $0.00$        & $0.00$        \\
+Var: Schools Simce High                   & $0.07$        & $0.07$        & $0.07$        & $0.07$        & $0.07$        \\
+Cov: Schools (Intercept) Simce Medium     & $-0.00$       & $-0.00$       & $-0.00$       & $-0.00$       & $-0.00$       \\
+Cov: Schools (Intercept) Simce High       & $-0.03$       & $-0.03$       & $-0.03$       & $-0.03$       & $-0.03$       \\
+Cov: Schools Simce Medium, Simce High     & $0.01$        & $0.01$        & $0.01$        & $0.01$        & $0.01$        \\
+Var: Residual                             & $0.66$        & $0.66$        & $0.66$        & $0.66$        & $0.66$        \\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[flushleft]
+\scriptsize{\item Note: Cells contain regression coefficients with standard errors in parentheses.
+Control variables are included.
+$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$ \\ \item Source: own elaboration based on First Study of Civic Education 2017.}
+\end{tablenotes}
+\end{threeparttable}
+}
+\caption{}
+\label{table:coefficients}
+\end{center}
+
+}
+
+\end{table}%
+
+\end{document}

### 2024-10-02 01-introduction.qmd
template mdpi
agregue un nuevo qmd para el template mdpi
RR-societies, abstract modification in paper-blinded.qmd and reviewed introduction.qmd

@@
 In particular, we focus on justifying inequalities in health, education and pensions, which are traditional social policy areas where access to better services could be justified based on payment capacity, referred to as market justice [@lane_market_1986; @lindh_bringing_2023].
-Most of the research in this realm has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies [@batruch_belief_2022].
-Regarding schools, the way in which they deal with unequal conditions of origin has been linked to the *hidden curriculum* [@chafel_schooling_1997], whereby students learn about distributive norms in society and mechanisms of justification of social differences.
+Market justice preferences refers to a belief in a market-based distributive principle, where goods and rewards are allocated according to individual merit, effort and contributions in a free-market economy (Lindh, 2015).
 
-Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population [@batruch_belief_2022; @wiederkehr_belief_2015], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger market justice preferences, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary [@batruch_belief_2022].
-We focus on the student-age population as we point out that it is possible to track down the origin of meritocratic beliefs (and their consequences) to early socialization processes.
-To this regard, we take into account the family and the school as two main socialization agencies that play a significant role in the socialization of cultural beliefs by transmitting cultural norms, values, and expectations to young people.
+Most of the research on meritocracy and justice preferences has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies [@batruch_belief_2022].
+Studying beliefs about meritocracy and market justice preferences at school age is crucial because these early beliefs can shape individuals' attitudes towards inequality and social justice later in life.
+As Castillo (2019) argues, meritocratic ideals "serve as a principle that legitimizes the distribution of goods and rewards based on individual talent and effort," making it essential to understand how these beliefs develop during adolescence.
+Furthermore, early exposure to these beliefs may influence broader perceptions of fairness, as Reynolds (2014) highlights, "beliefs about the stratification system influence people's judgments about the fairness of inequality," suggesting that such attitudes formed at a young age can have lasting impacts.
+Additionally, García-Sierra (2023) emphasizes that the effects of holding meritocratic beliefs are not uniform, as these ideologies may reinforce socioeconomic disparities depending on students' social backgrounds.
 
@@
 In education, 30.6% of school enrollment is in public schools, 54.0% in state-subsidized (voucher) private schools, and 9.3% in fully private schools, generally attracting higher-income groups [@ministeriodeeducacion_resumen_2023].
+
+Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population [@batruch_belief_2022; @wiederkehr_belief_2015], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger market justice preferences, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary [@batruch_belief_2022].
+We focus on the student-age population as we point out that it is possible to track down the origin of meritocratic beliefs (and their consequences) to early socialization processes.
+To this regard, we take into account the family and the school as two main socialization agencies that play a significant role in the socialization of cultural beliefs by transmitting cultural norms, values, and expectations to young people.

### 2024-10-02 paper-blinded.qmd
RR-societies, abstract modification in paper-blinded.qmd and reviewed introduction.qmd

@@
 abstract: |
-  Previous research has shown that schools justify student performance differences based on meritocratic ideals.
-However, the link between school meritocracy and students’ justification of inequalities has received little attention.
-This paper examines how school and family socialization influence students’ beliefs about meritocracy and market-based justifications for access to health, pensions, and education in Chile.
-Using data from the First Study of Civic Education in Chile, we employ a multilevel modeling strategy to analyze 5,047 students and parents nested within 231 schools.
-Our findings reveal that a significant proportion of Chilean students agree with market justice principles, at a higher rate than adults.
-Most students endorse meritocratic views, particularly the reward of effort.
-Multilevel estimations show strong associations between meritocratic beliefs and market justice preferences: students who believe in the reward of effort and talent are more likely to justify inequalities based on financial capacity.
-At the school level, market justice preferences are higher in high-status schools but lower in schools with better achievement.
-Additionally, the conditional influence of meritocracy beliefs weakens in schools with higher socioeconomic status and performance.
-These results suggest that meritocratic beliefs and market justice preferences begin early and are influenced by the school environment.
+  Previous research has shown that schools often justify student performance differences using meritocratic ideals.
+One potential consequence of such ideals is the legitimization of outcome inequalities across various spheres, including those traditionally associated with equality and redistribution.
+In this study, we argue that the promotion of meritocratic values during school-age can shape students' beliefs about meritocracy and influence their views on market-based access to health, pensions, and education.
+Using data from the 2017 National Study of Civic Education in Chile, which includes 5,047 8th-grade students from 231 schools, we estimated a series of multilevel models (lme4 library, R version 4.1.3) to test our hypotheses.
+Our findings show that a significant proportion of Chilean students agree with market justice principles—more so than adults.
+Most students endorse meritocratic views, particularly the notion that effort should be rewarded, which strongly correlates with market justice preferences: students who believe in meritocracy are more likely to justify inequalities based on financial capacity.
+At the school level, market justice preferences are higher in high-status schools but lower in schools with higher academic achievement.
+Furthermore, the conditional influence of meritocratic beliefs diminishes in schools with higher socioeconomic status and performance levels.
+These results suggest that the association between meritocratic beliefs and market justice preferences is already established at school age and is shaped by the school environment.
 keywords: 

### 2024-10-02 03-methods.qmd
template mdpi
agregue un nuevo qmd para el template mdpi
Update submit mdpi FINAL
@juancarloscastillo
coments methods & analisis

@@
 ---
-author: "Equipo Edumer"
+author: "Equipo EDUMER"
 bibliography: "input/bib/just-ineq-merit.bib"
-csl: "input/bib/apa6.csl"
+csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
 ---
@@
 
-The First Study of Civic Education in Chile, conducted by the Agency for Quality Education of the Ministry of Education in 2017, serves as the main data source.
-This database is composed by a civic knowledge test score and a series of items batteries measuring different aspects of citizenship.
-The target population includes 8th-grade students in 242 schools nationwide, as well as their parents and teachers.
-Initially, the database contains 8,589 student observations and 6,770 parent observations.
-To ensure higher data quality and considering the survey's unit of analysis, we removed 171 student cases and 79 parent cases that exhibited repetitive and careless response patterns [@gottfried_autocorrelation_2022].
+The First Study of Civic Education in Chile, conducted by the Agency for Quality Education of the Ministry of Education in 2017, is the primary data source.
+This database comprises a civic knowledge test score and a series of items that measure different aspects of citizenship.
+The target population includes 8th-grade students in 242 schools nationwide.
+Initially, the database contains 8,589 student observations.To ensure higher data quality and considering the survey's unit of analysis, we removed 171 student cases that exhibited repetitive and careless response patterns [@gottfried_autocorrelation_2022].
 Additionally, we utilized data from the System for Measuring the Quality of Education (SIMCE) conducted by the Ministry of Education in 2017, which provides school-level information such as administrative dependence, socioeconomic classification, and results obtained in standardized mathematics and language tests.
-After processing the variables and removing missing cases, the final database for this study includes a two-level stratified sample composed of 5,047 students and parents (level 1) nested within 231 schools (level 2).
+After processing the variables and removing missing cases, the final database for this study includes a two-level stratified sample composed of 5,047 students (level 1) nested within 231 schools (level 2).
 
@@
 **Market Justice Preferences**: The dependent variables in this study are market justice preferences.
-This construct is measured through three variables that address the degree of justification regarding whether access to social services in pensions, education, and health should be income conditional.
-Justification of inequality in health is measured by the item: "Is it fair in Chile that people with higher incomes can access better healthcare than those with lower incomes?" The same question is asked for education and pensions.
+This construct is measured through three variables that address the degree of justification regarding whether access to social services in pensions, education, and health should be conditional on income.
+The Justification of inequality in health is measured by the following item: "Is it fair in Chile that people with higher incomes can access better healthcare than those with lower incomes?" The same question is asked for education and pensions.
 In all cases, respondents indicate their preferences on a Likert scale ranging from "completely disagree" (1) to "completely agree" (4).
-Additionally, we include a summarized indicator of "market justice preferences", measured by an average index across these items (α = 0.86), with values ranging from 1 to 4, where higher values represent greater preferences for market justice (see @tbl-desc-dependientes).
-We analyzed these items independently as well as by the average index.
+Additionally, we include a summarized indicator of "market justice preferences", measured by an average index across these items (α = 0.86), with values ranging from 1 to 4, where higher values represent stronger preferences for market justice (see @tbl-desc-dependientes).
+We analyzed these items independently and by the average index.
 
@@
 Socioeconomic status is assessed using the Ministry of Education's classification, measured as an ordinal item with five categories ranging from "low" (1) to "high" (5).
-Academic performance is measured using the school's results in the SIMCE (System of Measurement of Educational Quality) standardized tests, which are administered yearly at different educational levels in areas such as language and mathematics.
+Academic performance is measured using the school's results in the SIMCE (System of Measurement of Educational Quality) standardized tests, administered yearly at different educational levels in language and mathematics.
 These results are categorized as "low," "medium," and "high." The contextual level items used, response categories, and their frequencies are detailed in @tbl-desc-school.
@@
 
-The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open Science (OSF), the access to the document is available at this [link](https://doi.org/10.17605/OSF.IO/UFSDV).
+The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open Science (OSF), and access to the document is available at this [link](https://doi.org/10.17605/OSF.IO/UFSDV).
 The statistical analysis for this research was conducted using the *lme4* package in R version 4.1.3.

### 2024-10-02 04-analisis.qmd
template mdpi
agregue un nuevo qmd para el template mdpi
coments methods & analisis

@@
 ---
-author: "Equipo Edumer"
+author: "Equipo EDUMER"
 bibliography: "input/bib/just-ineq-merit.bib"
@@
 
-Model 1 in @tbl-ordinal-reg for the three dependent variables of justice in differential access to health, pensions, and education demonstrates a similar trend regarding the effect of the perception of meritocracy on these items, with all effects being statistically significant (*p* \< 0.05).
-Regarding the variables of social meritocracy, the three variables---talent, effort, and deservingness---show that as these increase, the justification for differentiated access to pensions, education, and health also increases.
-In the context of school meritocracy, the effects are mixed: as perceived school talent increases, the justification for differentiated access to these services increases; conversely, as perceived school effort increases, the justification decreases, holding all other variables constant.
+Model 1 in @tbl-ordinal-reg for the three dependent variables of justice in differential access to health, pensions, and education demonstrates a similar trend regarding the effect of the perception of meritocracy on these items, with all effects being statistically significant (*p* < 0.05).
+Regarding the variables of social meritocracy, the three variables—talent ($\beta$ = 0.16), effort ($\beta$ = 0.11), and deservingness ($\beta$ = 0.51)—show that as these increase, the justification for differentiated access to pensions, education, and health also increases.
+In the context of school meritocracy, the effects are mixed: as perceived school talent increases, the justification for differentiated access to these services increases ($\beta$ = 0.23); conversely, as perceived school effort increases, the justification decreases ($\beta$ = -0.26), holding all other variables constant.
 
-Additionally, Model 1 examines family status through parents' educational level and the number of books at home.
-For health, pensions, and education, a negative but statistically non-significant association with parents' education is observed.
-However, having more than 25 books at home shows a negative and statistically significant effect on market justice in health and pensions, suggesting less justification for differentiated access compared to those with fewer books.
+Additionally, Model 1 examines family status through parents’ educational level and the number of books at home.
+For health, pensions, and education, a negative but statistically non-significant association with parents’ education is observed.
+However, having more than 25 books at home shows a negative and statistically significant effect on market justice in health and pensions ($\beta$ = -0.13; *p* < 0.001), suggesting less justification for differentiated access compared to those with fewer books.
 
 Model 2 in @tbl-ordinal-reg incorporates school-level variables.
-The results show that high socioeconomic level schools show greater preferences for market justice in health, pensions, and education compared to low socioeconomic level schools, with statistically significant results (*p* \< 0.05).
-Moreover, and contrary to our hypothesis, schools with higher academic performance show negative and statistically significant effects across the three items.
+The results show that high socioeconomic level schools show greater preferences for market justice in health, pensions, and education compared to low socioeconomic level schools, with statistically significant results ($\beta$ = 0.71; *p* < 0.05).
+Moreover, and contrary to our hypothesis, schools with higher academic performance show negative and statistically significant effects across the three items ($\beta$ = -0.56; *p* < 0.001).
 
@@
 
-@tbl-lineal-reg presents the results of linear mixed-effects models for the market justice preference index.
-Model 1 shows that higher perceived societal meritocracy, through effort, talent, and deservingness, is positively and significantly associated with market justice preferences.
-School meritocracy perceptions yield mixed effects: perceived talent increases market justice preference, while perceived effort decreases it, both statistically significant (*p* \< 0.001).
-Model 2 includes family status, finding no statistically significant effects from parents' education or the number of books at home.
+@tbl-lineal-reg ppresents the results of linear mixed-effects models for the market justice preference index.
+Model 1 shows that higher perceived societal meritocracy, through effort ($\beta$ = 0.08), talent ($\beta$ = 0.07), and deservingness ($\beta$ = 0.2), is positively and significantly associated with market justice preferences (*p* < 0.001).
+School meritocracy perceptions yield mixed effects: perceived talent increases market justice preference ($\beta$ = 0.1), while perceived effort decreases it ($\beta$ = -0.12), both statistically significant (*p* < 0.001).
+Model 2 includes family status, finding no statistically significant effects from parents’ education or the number of books at home.
 Regarding school characteristics, Models 3 and 4 introduce socioeconomic level and academic performance.
-The results of Model 3 suggest a statistically significant relationship between school socioeconomic levels and market justice preferences, specifically those schools with high socioeconomic levels ($\beta$ = 0.37) compared to low socioeconomic levels.
-However, Model 4 indicates that schools with medium ($\beta$ = -0.11) and high ($\beta$ = -0.26) academic performance present lower market justice preferences compared to low-performing schools, with these effects being statistically significant when controlling for other variables.
+The results of Model 3 suggest a statistically significant relationship between school socioeconomic levels and market justice preferences, specifically those schools with high socioeconomic levels ($\beta$ = 0.37; *p* < 0.01) compared to low socioeconomic levels.
+However, Model 4 indicates that schools with medium ($\beta$ = -0.11; *p* < 0.01) and high ($\beta$ = -0.26; *p* < 0.001) academic performance present lower market justice preferences compared to low-performing schools, with these effects being statistically significant when controlling for other variables.
 
@@
 @tbl-interact displays the interaction terms related to our hypotheses.
-The results suggest that students' family background moderates the relationship between their meritocratic perceptions and their justification of market allocation in social services.
-Specifically, the relationship between social effort and market justice preferences is less positive for students whose parents have university or postgraduate education.
-Lower-status students (measured by parental education) prefer more market justice when they strongly adhere to meritocratic perceptions of deservingness.
-A similar moderating effect is seen with family cultural capital: the relationship between effort and market justice preferences is less positive for students with more than 25 books at home.
+The results suggest that students’ family background moderates the relationship between their meritocratic perceptions and their justification of market allocation in social services.
+Specifically, the relationship between social effort and market justice preferences is less positive for students whose parents have university or postgraduate education ($\beta$ = -0.079; *p* < 0.05).
+Lower-status students (measured by parental education) prefer more market justice when they strongly adhere to meritocratic perceptions of deservingness ($\beta$ = -0.118; *p* < 0.01).
+A similar moderating effect is seen with family cultural capital: the relationship between effort and market justice preferences is less positive for students with more than 25 books at home ($\beta$ = -0.081; *p* < 0.01).
 

### 2024-10-02 01-introduction.qmd
Update revised version MPI Submit

@@
 In particular, we focus on justifying inequalities in health, education and pensions, which are traditional social policy areas where access to better services could be justified based on payment capacity, referred to as market justice [@lane_market_1986; @lindh_bringing_2023].
-Market justice preferences refers to a belief in a market-based distributive principle, where goods and rewards are allocated according to individual merit, effort and contributions in a free-market economy (Lindh, 2015).
+Market justice preferences refers to a belief in a market-based distributive principle, where goods and rewards are allocated according to individual merit, effort and contributions in a free-market economy [@lindh_public_2015].
 
@@
 Studying beliefs about meritocracy and market justice preferences at school age is crucial because these early beliefs can shape individuals' attitudes towards inequality and social justice later in life.
-As Castillo (2019) argues, meritocratic ideals "serve as a principle that legitimizes the distribution of goods and rewards based on individual talent and effort," making it essential to understand how these beliefs develop during adolescence.
-Furthermore, early exposure to these beliefs may influence broader perceptions of fairness, as Reynolds (2014) highlights, "beliefs about the stratification system influence people's judgments about the fairness of inequality," suggesting that such attitudes formed at a young age can have lasting impacts.
-Additionally, García-Sierra (2023) emphasizes that the effects of holding meritocratic beliefs are not uniform, as these ideologies may reinforce socioeconomic disparities depending on students' social backgrounds.
+As Castillo et al.
+[-@castillo_meritocracia_2019] argues, meritocratic ideals serve as a principle that legitimizes the distribution of goods and rewards based on individual talent and effort, making it essential to understand how these beliefs develop during adolescence.
+Furthermore, early exposure to these beliefs may influence broader perceptions of fairness, as Reynolds and Xian [-@reynolds_perceptions_2014, p.134] highlights, beliefs about the stratification system influence people's judgments about the fairness of inequality, suggesting that such attitudes formed at a young age can have lasting impacts.
+Additionally, García-Sierra [-@garcia-sierra_dark_2023] emphasizes that the effects of holding meritocratic beliefs are not uniform, as these ideologies may reinforce socioeconomic disparities depending on students' social backgrounds.
 

### 2024-10-02 cover_letter.qmd
Update revised version MPI Submit

@@
 title: "The Socialization of Meritocracy and Market Justice Preferences at School"
-subtitle: "Cover Letter"
+subtitle: "Cover Letter [Revised Version]"
 format:
@@
 
-Previous research has shown that schools justify student performance differences based on meritocratic ideals.
-However, the link between school meritocracy and students’ justification of inequalities has received little attention.
-This paper examines how school and family socialization influence students’ beliefs about meritocracy and market-based justifications for access to health, pensions, and education in Chile.
-Using data from the First Study of Civic Education in Chile, we employ a multilevel modeling strategy to analyze 5,047 students and parents nested within 231 schools.
-Our findings reveal that a significant proportion of Chilean students agree with market justice principles, at a higher rate than adults.
-Most students endorse meritocratic views, particularly the reward of effort.
-Multilevel estimations show strong associations between meritocratic beliefs and market justice preferences: students who believe in the reward of effort and talent are more likely to justify inequalities based on financial capacity.
-At the school level, market justice preferences are higher in high-status schools but lower in schools with better achievement.
-Additionally, the conditional influence of meritocracy beliefs weakens in schools with higher socioeconomic status and performance.
-These results suggest that meritocratic beliefs and market justice preferences begin early and are influenced by the school environment.
+Previous research has shown that schools often justify student performance differences using meritocratic ideals.
+One potential consequence of such ideals is the legitimization of outcome inequalities across various spheres, including those traditionally associated with equality and redistribution.
+In this study, we argue that the promotion of meritocratic values during school-age can shape students' beliefs about meritocracy and influence their views on market-based access to health, pensions, and education.
+Using data from the 2017 National Study of Civic Education in Chile, which includes 5,047 8th-grade students from 231 schools, we estimated a series of multilevel models (lme4 library, R version 4.1.3) to test our hypotheses.
+Our findings show that a significant proportion of Chilean students agree with market justice principles—more so than adults.
+Most students endorse meritocratic views, particularly the notion that effort should be rewarded, which strongly correlates with market justice preferences: students who believe in meritocracy are more likely to justify inequalities based on financial capacity.
+At the school level, market justice preferences are higher in high-status schools but lower in schools with higher academic achievement.
+Furthermore, the conditional influence of meritocratic beliefs diminishes in schools with higher socioeconomic status and performance levels.
+These results suggest that the association between meritocratic beliefs and market justice preferences is already established at school age and is shaped by the school environment.
 
@@
 
-Mauricio Salgado, School of Sociology, Universidad Andrés Bello.  
+Mauricio Salgado, School of Sociology, Universidad Andrés Bello and Centro de Estudios Públicos (CEP).
 ORCID: 0000-0002-8108-6154

### 2024-10-02 cover_letter.tex
Update revised version MPI Submit

--- 2024-08-07 submission/MDPI/cover_letter.tex
@@
 \makeatother
-\subtitle{Cover Letter}
+\subtitle{Cover Letter {[}Revised Version{]}}
 
@@
 \setstretch{1.15}
-Previous research has shown that schools justify student performance
-differences based on meritocratic ideals. However, the link between
-school meritocracy and students' justification of inequalities has
-received little attention. This paper examines how school and family
-socialization influence students' beliefs about meritocracy and
-market-based justifications for access to health, pensions, and
-education in Chile. Using data from the First Study of Civic Education
-in Chile, we employ a multilevel modeling strategy to analyze 5,047
-students and parents nested within 231 schools. Our findings reveal that
+Previous research has shown that schools often justify student
+performance differences using meritocratic ideals. One potential
+consequence of such ideals is the legitimization of outcome inequalities
+across various spheres, including those traditionally associated with
+equality and redistribution. In this study, we argue that the promotion
+of meritocratic values during school-age can shape students' beliefs
+about meritocracy and influence their views on market-based access to
+health, pensions, and education. Using data from the 2017 National Study
+of Civic Education in Chile, which includes 5,047 8th-grade students
+from 231 schools, we estimated a series of multilevel models (lme4
+library, R version 4.1.3) to test our hypotheses. Our findings show that
 a significant proportion of Chilean students agree with market justice
-principles, at a higher rate than adults. Most students endorse
-meritocratic views, particularly the reward of effort. Multilevel
-estimations show strong associations between meritocratic beliefs and
-market justice preferences: students who believe in the reward of effort
-and talent are more likely to justify inequalities based on financial
-capacity. At the school level, market justice preferences are higher in
-high-status schools but lower in schools with better achievement.
-Additionally, the conditional influence of meritocracy beliefs weakens
-in schools with higher socioeconomic status and performance. These
-results suggest that meritocratic beliefs and market justice preferences
-begin early and are influenced by the school environment.
+principles---more so than adults. Most students endorse meritocratic
+views, particularly the notion that effort should be rewarded, which
+strongly correlates with market justice preferences: students who
+believe in meritocracy are more likely to justify inequalities based on
+financial capacity. At the school level, market justice preferences are
+higher in high-status schools but lower in schools with higher academic
+achievement. Furthermore, the conditional influence of meritocratic
+beliefs diminishes in schools with higher socioeconomic status and
+performance levels. These results suggest that the association between
+meritocratic beliefs and market justice preferences is already
+established at school age and is shaped by the school environment.
 
@@
 
-Mauricio Salgado, School of Sociology, Universidad Andrés Bello.\\
-ORCID: 0000-0002-8108-6154
+Mauricio Salgado, School of Sociology, Universidad Andrés Bello and
+Centro de Estudios Públicos (CEP). ORCID: 0000-0002-8108-6154
 

### 2024-10-02 paper-blinded.qmd
Update revised version MPI Submit

@@
   hikmah-pdf: 
-    indent: false
-    keep-tex: true
-  hikmah-manuscript-pdf: 
     indent: false
@@
 colorlinks: true
+number-sections: true
 fig-height: 4

### 2024-10-02 paper-blinded.tex
Update revised version MPI Submit

--- 2024-08-07 submission/MDPI/paper-blinded.tex
@@
 \setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\setcounter{secnumdepth}{5}
 
@@
 \begin{abstract}
-Previous research has shown that schools justify student performance
-differences based on meritocratic ideals. However, the link between
-school meritocracy and students' justification of inequalities has
-received little attention. This paper examines how school and family
-socialization influence students' beliefs about meritocracy and
-market-based justifications for access to health, pensions, and
-education in Chile. Using data from the First Study of Civic Education
-in Chile, we employ a multilevel modeling strategy to analyze 5,047
-students and parents nested within 231 schools. Our findings reveal that
+Previous research has shown that schools often justify student
+performance differences using meritocratic ideals. One potential
+consequence of such ideals is the legitimization of outcome inequalities
+across various spheres, including those traditionally associated with
+equality and redistribution. In this study, we argue that the promotion
+of meritocratic values during school-age can shape students' beliefs
+about meritocracy and influence their views on market-based access to
+health, pensions, and education. Using data from the 2017 National Study
+of Civic Education in Chile, which includes 5,047 8th-grade students
+from 231 schools, we estimated a series of multilevel models (lme4
+library, R version 4.1.3) to test our hypotheses. Our findings show that
 a significant proportion of Chilean students agree with market justice
-principles, at a higher rate than adults. Most students endorse
-meritocratic views, particularly the reward of effort. Multilevel
-estimations show strong associations between meritocratic beliefs and
-market justice preferences: students who believe in the reward of effort
-and talent are more likely to justify inequalities based on financial
-capacity. At the school level, market justice preferences are higher in
-high-status schools but lower in schools with better achievement.
-Additionally, the conditional influence of meritocracy beliefs weakens
-in schools with higher socioeconomic status and performance. These
-results suggest that meritocratic beliefs and market justice preferences
-begin early and are influenced by the school environment.
+principles---more so than adults. Most students endorse meritocratic
+views, particularly the notion that effort should be rewarded, which
+strongly correlates with market justice preferences: students who
+believe in meritocracy are more likely to justify inequalities based on
+financial capacity. At the school level, market justice preferences are
+higher in high-status schools but lower in schools with higher academic
+achievement. Furthermore, the conditional influence of meritocratic
+beliefs diminishes in schools with higher socioeconomic status and
+performance levels. These results suggest that the association between
+meritocratic beliefs and market justice preferences is already
+established at school age and is shaped by the school environment.
 \end{abstract}
@@
 
-\begin{ideuno}
-
-Front Matter
-
-\end{ideuno}
 % -------------------
@@
 
+\setstretch{1.15}
 \newpage{}
 
-\begin{idedos}
-
-Manuscript
-\end{idedos}
-
-\setstretch{1.15}
 \section{Introduction}\label{introduction}
@@
 {[}\citeproc{ref-lane_market_1986}{9},\citeproc{ref-lindh_bringing_2023}{10}{]}.
-Most of the research in this realm has only considered adults, leaving
-aside the study of how beliefs in this field develop at student age as
-well as the impact of the school context and the family as the main
-socialization agencies {[}\citeproc{ref-batruch_belief_2022}{3}{]}.
-Regarding schools, the way in which they deal with unequal conditions of
-origin has been linked to the \emph{hidden curriculum}
-{[}\citeproc{ref-chafel_schooling_1997}{11}{]}, whereby students learn
-about distributive norms in society and mechanisms of justification of
-social differences.
+Market justice preferences refers to a belief in a market-based
+distributive principle, where goods and rewards are allocated according
+to individual merit, effort and contributions in a free-market economy
+{[}\citeproc{ref-lindh_public_2015}{11}{]}.
+
+Most of the research on meritocracy and justice preferences has only
+considered adults, leaving aside the study of how beliefs in this field
+develop at student age as well as the impact of the school context and
+the family as the main socialization agencies
+{[}\citeproc{ref-batruch_belief_2022}{3}{]}. Studying beliefs about
+meritocracy and market justice preferences at school age is crucial
+because these early beliefs can shape individuals' attitudes towards
+inequality and social justice later in life. As Castillo et al.
+{[}\citeproc{ref-castillo_meritocracia_2019}{12}{]} argues, meritocratic
+ideals serve as a principle that legitimizes the distribution of goods
+and rewards based on individual talent and effort, making it essential
+to understand how these beliefs develop during adolescence. Furthermore,
+early exposure to these beliefs may influence broader perceptions of
+fairness, as Reynolds and Xian
+{[}\citeproc{ref-reynolds_perceptions_2014}{13}{]} highlights, beliefs
+about the stratification system influence people's judgments about the
+fairness of inequality, suggesting that such attitudes formed at a young
+age can have lasting impacts. Additionally, García-Sierra
+{[}\citeproc{ref-garcia-sierra_dark_2023}{14}{]} emphasizes that the
+effects of holding meritocratic beliefs are not uniform, as these
+ideologies may reinforce socioeconomic disparities depending on
+students' social backgrounds.
+
+The Chilean case is particularly intriguing for studying market justice
+preferences. This country is characterized by acute and persistent
+economic inequality, which stands out in Latin America and among OECD
+countries {[}\citeproc{ref-flores_top_2020}{15}{]}. In Chile, the
+poorest 50\% captures only 10\% of the total income and has negative
+wealth, while the richest 1\% receives almost 27\% of the income and
+holds 49.6\% of the wealth {[}\citeproc{ref-chancel_world_2022}{16}{]}.
+Much of this inequality has been attributed to the deep neoliberal
+reforms that institutionalized privatization and commodification of
+various economic sectors. Such reforms where introduced during the
+dictatorship (1973-1989) and expanded in democracy through concessions,
+demand credits, and specific regulatory frameworks
+{[}\citeproc{ref-ffrench-davis_reformas_2018}{17}{]}. This shift in
+economic policy has allowed the unprecedented emergence of markets in
+areas related to social reproduction, such as health, pensions, and
+education, with provision and access managed by private entities and
+segmented by individual payment capacity, heavily reliant on State
+subsidies {[}\citeproc{ref-boccardo_30_2020}{18}{]}. In health, although
+the majority of the population uses the public insurance system
+(78.9\%), 15.3\% are served by private insurers
+{[}\citeproc{ref-observatoriosocial_estadisticas_2024}{19}{]}. The
+pension system is based on individual capitalization, with mandatory
+contributions managed by private administrators investing in the
+financial market, currently involving 11 million contributors
+{[}\citeproc{ref-superintendenciadepensiones_estadisticas_2024}{20}{]}.
+In education, 30.6\% of school enrollment is in public schools, 54.0\%
+in state-subsidized (voucher) private schools, and 9.3\% in fully
+private schools, generally attracting higher-income groups
+{[}\citeproc{ref-ministeriodeeducacion_resumen_2023}{21}{]}.
 
@@
 
-The Chilean case is particularly intriguing for studying market justice
-preferences. This country is characterized by acute and persistent
-economic inequality, which stands out in Latin America and among OECD
-countries {[}\citeproc{ref-flores_top_2020}{12}{]}. In Chile, the
-poorest 50\% captures only 10\% of the total income and has negative
-wealth, while the richest 1\% receives almost 27\% of the income and
-holds 49.6\% of the wealth {[}\citeproc{ref-chancel_world_2022}{13}{]}.
-Much of this inequality has been attributed to the deep neoliberal
-reforms that institutionalized privatization and commodification of
-various economic sectors. Such reforms where introduced during the
-dictatorship (1973-1989) and expanded in democracy through concessions,
-demand credits, and specific regulatory frameworks
-{[}\citeproc{ref-ffrench-davis_reformas_2018}{14}{]}. This shift in
-economic policy has allowed the unprecedented emergence of markets in
-areas related to social reproduction, such as health, pensions, and
-education, with provision and access managed by private entities and
-segmented by individual payment capacity, heavily reliant on State
-subsidies {[}\citeproc{ref-boccardo_30_2020}{15}{]}. In health, although
-the majority of the population uses the public insurance system
-(78.9\%), 15.3\% are served by private insurers
-{[}\citeproc{ref-observatoriosocial_estadisticas_2024}{16}{]}. The
-pension system is based on individual capitalization, with mandatory
-contributions managed by private administrators investing in the
-financial market, currently involving 11 million contributors
-{[}\citeproc{ref-superintendenciadepensiones_estadisticas_2024}{17}{]}.
-In education, 30.6\% of school enrollment is in public schools, 54.0\%
-in state-subsidized (voucher) private schools, and 9.3\% in fully
-private schools, generally attracting higher-income groups
-{[}\citeproc{ref-ministeriodeeducacion_resumen_2023}{18}{]}.
-
 \subsection{Justification of inequality and market
@@
 based on prices and individuals' ability to pay
-{[}\citeproc{ref-lane_market_1986}{9},\citeproc{ref-boltanski_new_2005}{19},\citeproc{ref-streeck_citizens_2012}{20}{]}.
+{[}\citeproc{ref-lane_market_1986}{9},\citeproc{ref-boltanski_new_2005}{22},\citeproc{ref-streeck_citizens_2012}{23}{]}.
 Research on social stratification beliefs, which explore individual
 perceptions of who deserves what and why
-{[}\citeproc{ref-kluegel_beliefs_1987}{21}{]}, highlights that people's
+{[}\citeproc{ref-kluegel_beliefs_1987}{24}{]}, highlights that people's
 explanations and justifications of social inequality are closely tied to
 their judgments of deservingness. The influence of ideologies
-{[}\citeproc{ref-wegener_dominant_1995}{22}{]} and cultural schemas
-{[}\citeproc{ref-homan_being_2017}{23}{]} is pivotal in shaping these
+{[}\citeproc{ref-wegener_dominant_1995}{25}{]} and cultural schemas
+{[}\citeproc{ref-homan_being_2017}{26}{]} is pivotal in shaping these
 explanations by offering symbolic representations that frame societal
@@
 the literature
-{[}\citeproc{ref-castillo_legitimacy_2011}{24}--\citeproc{ref-shariff_income_2016}{27}{]},
+{[}\citeproc{ref-castillo_legitimacy_2011}{27}--\citeproc{ref-shariff_income_2016}{30}{]},
 there has been less examination of public beliefs about which life
@@
 five decades
-{[}\citeproc{ref-centeno_arc_2012}{28},\citeproc{ref-harvey_breve_2015}{29}{]},
+{[}\citeproc{ref-centeno_arc_2012}{31},\citeproc{ref-harvey_breve_2015}{32}{]},
 affecting areas such as pensions, health services, and education.
@@
 management of social services across nations
-{[}\citeproc{ref-jensen_worlds_2008}{30},\citeproc{ref-stoy_worlds_2014}{31}{]}.
+{[}\citeproc{ref-jensen_worlds_2008}{33},\citeproc{ref-stoy_worlds_2014}{34}{]}.
 Nordic countries, for example, predominantly employ public agencies to
@@
 trend toward marketization of welfare services has been growing since
-the 1980s {[}\citeproc{ref-salamon_marketization_1993}{32}{]}, and this
+the 1980s {[}\citeproc{ref-salamon_marketization_1993}{35}{]}, and this
 shift is increasingly evident even in countries where market solutions
 have traditionally had a minor role in social policy
-{[}\citeproc{ref-sivesind_changing_2017}{33}{]}. The expansion of
+{[}\citeproc{ref-sivesind_changing_2017}{36}{]}. The expansion of
 marketization has been related to a larger justification of market
@@
 show larger market justice preferences
-{[}\citeproc{ref-lindh_public_2015}{34}{]}.
+{[}\citeproc{ref-lindh_public_2015}{11}{]}.
 
@@
 resentment that might motivate forms of collective action
-{[}\citeproc{ref-greitemeyer_subjective_2016}{35}--\citeproc{ref-power_deprivationprotest_2018}{38}{]},
+{[}\citeproc{ref-greitemeyer_subjective_2016}{37}--\citeproc{ref-power_deprivationprotest_2018}{40}{]},
 Lane pointed out that in market settings, social comparisons are more
@@
 distributive principles that characterize welfare states, such as need
-and equality (see {[}\citeproc{ref-wilson_role_2003}{39}{]}).
+and equality (see {[}\citeproc{ref-wilson_role_2003}{41}{]}).
 
@@
 that varies across the region
-{[}\citeproc{ref-bucca_merit_2016}{40}--\citeproc{ref-salgado_inequality_2023}{43}{]}.
+{[}\citeproc{ref-bucca_merit_2016}{42}--\citeproc{ref-salgado_inequality_2023}{45}{]}.
 The reliance on private welfare providers and widespread user fees
-{[}\citeproc{ref-molyneux_neoliberal_2008}{44}{]} adds complexity to
+{[}\citeproc{ref-molyneux_neoliberal_2008}{46}{]} adds complexity to
 this context, as reflected in surveys conducted by the Center for Public
@@
 insurance, and 63\% would prefer private education
-{[}\citeproc{ref-centrodeestudiospublicos_estudio_2024}{45}{]}. Yet,
+{[}\citeproc{ref-centrodeestudiospublicos_estudio_2024}{47}{]}. Yet,
 research on children's justification in this area remains limited,
@@
 with low mobility that threaten the meritocratic ideal
-{[}\citeproc{ref-goldthorpe_myth_2003}{46}{]}. More recently, sociology
+{[}\citeproc{ref-goldthorpe_myth_2003}{48}{]}. More recently, sociology
 and social psychology research has attended to subjective aspects
@@
 societies, such as beliefs in meritocracy
-{[}\citeproc{ref-castillo_multidimensional_2023}{47}--\citeproc{ref-mijs_paradox_2019}{49}{]}.
+{[}\citeproc{ref-castillo_multidimensional_2023}{49}--\citeproc{ref-mijs_paradox_2019}{51}{]}.
 
@@
 preferences and perceptions
-{[}\citeproc{ref-castillo_multidimensional_2023}{47}{]}. While
+{[}\citeproc{ref-castillo_multidimensional_2023}{49}{]}. While
 meritocratic preferences refer to a justification of distribution based
@@
 in their society
-{[}\citeproc{ref-duru-bellat_who_2012}{50},\citeproc{ref-castillo_meritocracia_2019}{51}{]}.
+{[}\citeproc{ref-castillo_meritocracia_2019}{12},\citeproc{ref-duru-bellat_who_2012}{52}{]}.
 The perception can vary greatly depending on individual experiences, as
@@
 to maintain and even reinforce inequality
-{[}\citeproc{ref-mijs_paradox_2019}{49},\citeproc{ref-lampert_meritocratic_2013}{52}{]}.
+{[}\citeproc{ref-mijs_paradox_2019}{51},\citeproc{ref-lampert_meritocratic_2013}{53}{]}.
 Some studies have analyzed how those with greater privileges believe
-more in meritocracy {[}\citeproc{ref-reynolds_perceptions_2014}{53}{]},
+more in meritocracy {[}\citeproc{ref-reynolds_perceptions_2014}{13}{]},
 how greater economic inequality increases meritocratic beliefs
-{[}\citeproc{ref-mijs_paradox_2019}{49}{]}, and how larger inequality
+{[}\citeproc{ref-mijs_paradox_2019}{51}{]}, and how larger inequality
 decreases it {[}\citeproc{ref-morris_representing_2022}{54}{]}.
@@
 
-\(H_{1a}\): Students who perceive that there is more meritocracy in
-society will show larger market justice preferences
+\(H_{1a}\): Students who perceive greater meritocracy in society will
+exhibit stronger preferences for market justice.
 
@@
 
-\(H_{1b}\): Students who perceive that there is more meritocracy at
-school will show larger market justice preferences
+\(H_{1b}\): Students who perceive greater meritocracy in their schools
+will exhibit stronger preferences for market justice.
 
@@
 
-\(H_2\): Students from families of higher social status will show larger
-market justice preferences.
+\(H_2\): Students from higher social status families will exhibit
+stronger preferences for market justice.
 
@@
 
-\(H_3\): Students from schools of higher social status will show larger
-market justice preferences.
-
-\(H_4\): Students from schools with an average better performance in
-standardized achievement tests will show larger market justice
-preferences.
+\(H_3\): Students from higher social status schools will show stronger
+preferences for market justice.
+
+\(H_4\): Students from schools with higher average performance on
+standardized achievement tests will show stronger preferences for market
+justice.
 
@@
 
-\(H_5\): The family's social status will moderate the effect of
-meritocratic perception on market justice preferences.
-
-\(H_6\): The school's status will moderate the effect of meritocratic
-perceptions on market justice preferences.
-
-\(H_7\): The school's performance in standardized achievement tests will
-moderate the effect of meritocratic perceptions on market justice
-preferences.
+\(H_5\): The family's social status will moderate the relationship
+between perceptions of meritocracy and preferences for market justice.
+
+\(H_6\): The school's status will moderate the relationship between
+perceptions of meritocracy and preferences for market justice.
+
+\(H_7\): The school's performance on standardized achievement tests will
+moderate the relationship between perceptions of meritocracy and
+preferences for market justice.
 
@@
 
-\section{Data, variables and method}\label{data-variables-and-method}
+\section{Data, Variables and Methods}\label{data-variables-and-methods}
 
@@
 The First Study of Civic Education in Chile, conducted by the Agency for
-Quality Education of the Ministry of Education in 2017, serves as the
-main data source. This database is composed by a civic knowledge test
-score and a series of items batteries measuring different aspects of
-citizenship. The target population includes 8th-grade students in 242
-schools nationwide, as well as their parents and teachers. Initially,
-the database contains 8,589 student observations and 6,770 parent
-observations. To ensure higher data quality and considering the survey's
-unit of analysis, we removed 171 student cases and 79 parent cases that
-exhibited repetitive and careless response patterns
+Quality Education of the Ministry of Education in 2017, is the primary
+data source. This database comprises a civic knowledge test score and a
+series of items that measure different aspects of citizenship. The
+target population includes 8th-grade students in 242 schools nationwide.
+Initially, the database contains 8,589 student observations.To ensure
+higher data quality and considering the survey's unit of analysis, we
+removed 171 student cases that exhibited repetitive and careless
+response patterns
 {[}\citeproc{ref-gottfried_autocorrelation_2022}{73}{]}. Additionally,
@@
 removing missing cases, the final database for this study includes a
-two-level stratified sample composed of 5,047 students and parents
-(level 1) nested within 231 schools (level 2).
+two-level stratified sample composed of 5,047 students (level 1) nested
+within 231 schools (level 2).
 
@@
 whether access to social services in pensions, education, and health
-should be income conditional. Justification of inequality in health is
-measured by the item: ``Is it fair in Chile that people with higher
-incomes can access better healthcare than those with lower incomes?''
-The same question is asked for education and pensions. In all cases,
-respondents indicate their preferences on a Likert scale ranging from
-``completely disagree'' (1) to ``completely agree'' (4). Additionally,
-we include a summarized indicator of ``market justice preferences'',
-measured by an average index across these items (α = 0.86), with values
-ranging from 1 to 4, where higher values represent greater preferences
-for market justice (see Table~\ref{tbl-desc-dependientes}). We analyzed
-these items independently as well as by the average index.
+should be conditional on income. The Justification of inequality in
+health is measured by the following item: ``Is it fair in Chile that
+people with higher incomes can access better healthcare than those with
+lower incomes?'' The same question is asked for education and pensions.
+In all cases, respondents indicate their preferences on a Likert scale
+ranging from ``completely disagree'' (1) to ``completely agree'' (4).
+Additionally, we include a summarized indicator of ``market justice
+preferences'', measured by an average index across these items (α =
+0.86), with values ranging from 1 to 4, where higher values represent
+stronger preferences for market justice (see
+Table~\ref{tbl-desc-dependientes}). We analyzed these items
+independently and by the average index.
 
@@
 performance is measured using the school's results in the SIMCE (System
-of Measurement of Educational Quality) standardized tests, which are
-administered yearly at different educational levels in areas such as
-language and mathematics. These results are categorized as ``low,''
-``medium,'' and ``high.'' The contextual level items used, response
-categories, and their frequencies are detailed in
-Table~\ref{tbl-desc-school}.
+of Measurement of Educational Quality) standardized tests, administered
+yearly at different educational levels in language and mathematics.
+These results are categorized as ``low,'' ``medium,'' and ``high.'' The
+contextual level items used, response categories, and their frequencies
+are detailed in Table~\ref{tbl-desc-school}.
 
@@
 The hypotheses of this research were pre-registered in the Open Science
-Framework platform of the Center for Open Science (OSF), the access to
+Framework platform of the Center for Open Science (OSF), and access to
 the document is available at this
@@
 
-\subsection{Descriptive}\label{descriptive}
+\subsection{Descriptive statistics}\label{descriptive-statistics}
 
@@
 
-\subsection{Multivariate}\label{multivariate}
+\subsection{Multivariate analysis}\label{multivariate-analysis}
 
@@
 significant (\emph{p} \textless{} 0.05). Regarding the variables of
-social meritocracy, the three variables---talent, effort, and
-deservingness---show that as these increase, the justification for
-differentiated access to pensions, education, and health also increases.
-In the context of school meritocracy, the effects are mixed: as
-perceived school talent increases, the justification for differentiated
-access to these services increases; conversely, as perceived school
-effort increases, the justification decreases, holding all other
-variables constant.
+social meritocracy, the three variables---talent (\(\beta\) = 0.16),
+effort (\(\beta\) = 0.11), and deservingness (\(\beta\) = 0.51)---show
+that as these increase, the justification for differentiated access to
+pensions, education, and health also increases. In the context of school
+meritocracy, the effects are mixed: as perceived school talent
+increases, the justification for differentiated access to these services
+increases (\(\beta\) = 0.23); conversely, as perceived school effort
+increases, the justification decreases (\(\beta\) = -0.26), holding all
+other variables constant.
 
@@
 at home shows a negative and statistically significant effect on market
-justice in health and pensions, suggesting less justification for
-differentiated access compared to those with fewer books.
+justice in health and pensions (\(\beta\) = -0.13; \emph{p} \textless{}
+0.001), suggesting less justification for differentiated access compared
+to those with fewer books.
 
@@
 education compared to low socioeconomic level schools, with
-statistically significant results (\emph{p} \textless{} 0.05). Moreover,
-and contrary to our hypothesis, schools with higher academic performance
-show negative and statistically significant effects across the three
-items.
+statistically significant results (\(\beta\) = 0.71; \emph{p}
+\textless{} 0.05). Moreover, and contrary to our hypothesis, schools
+with higher academic performance show negative and statistically
+significant effects across the three items (\(\beta\) = -0.56; \emph{p}
+\textless{} 0.001).
 
@@
 
-Table~\ref{tbl-lineal-reg} presents the results of linear mixed-effects
+Table~\ref{tbl-lineal-reg} ppresents the results of linear mixed-effects
 models for the market justice preference index. Model 1 shows that
-higher perceived societal meritocracy, through effort, talent, and
-deservingness, is positively and significantly associated with market
-justice preferences. School meritocracy perceptions yield mixed effects:
-perceived talent increases market justice preference, while perceived
-effort decreases it, both statistically significant (\emph{p}
+higher perceived societal meritocracy, through effort (\(\beta\) =
+0.08), talent (\(\beta\) = 0.07), and deservingness (\(\beta\) = 0.2),
+is positively and significantly associated with market justice
+preferences (\emph{p} \textless{} 0.001). School meritocracy perceptions
+yield mixed effects: perceived talent increases market justice
+preference (\(\beta\) = 0.1), while perceived effort decreases it
+(\(\beta\) = -0.12), both statistically significant (\emph{p}
 \textless{} 0.001). Model 2 includes family status, finding no
@@
 socioeconomic levels and market justice preferences, specifically those
-schools with high socioeconomic levels (\(\beta\) = 0.37) compared to
-low socioeconomic levels. However, Model 4 indicates that schools with
-medium (\(\beta\) = -0.11) and high (\(\beta\) = -0.26) academic
-performance present lower market justice preferences compared to
-low-performing schools, with these effects being statistically
-significant when controlling for other variables.
+schools with high socioeconomic levels (\(\beta\) = 0.37; \emph{p}
+\textless{} 0.01) compared to low socioeconomic levels. However, Model 4
+indicates that schools with medium (\(\beta\) = -0.11; \emph{p}
+\textless{} 0.01) and high (\(\beta\) = -0.26; \emph{p} \textless{}
+0.001) academic performance present lower market justice preferences
+compared to low-performing schools, with these effects being
+statistically significant when controlling for other variables.
 
@@
 preferences is less positive for students whose parents have university
-or postgraduate education. Lower-status students (measured by parental
-education) prefer more market justice when they strongly adhere to
-meritocratic perceptions of deservingness. A similar moderating effect
-is seen with family cultural capital: the relationship between effort
-and market justice preferences is less positive for students with more
-than 25 books at home.
+or postgraduate education (\(\beta\) = -0.079; \emph{p} \textless{}
+0.05). Lower-status students (measured by parental education) prefer
+more market justice when they strongly adhere to meritocratic
+perceptions of deservingness (\(\beta\) = -0.118; \emph{p} \textless{}
+0.01). A similar moderating effect is seen with family cultural capital:
+the relationship between effort and market justice preferences is less
+positive for students with more than 25 books at home (\(\beta\) =
+-0.081; \emph{p} \textless{} 0.01).
 
@@
 legitimation of social inequalities
-{[}\citeproc{ref-mijs_paradox_2019}{49},\citeproc{ref-darnon_where_2018}{63}{]},
+{[}\citeproc{ref-mijs_paradox_2019}{51},\citeproc{ref-darnon_where_2018}{63}{]},
 but the consideration of school level population as well as market
@@
 formation of inequality preferences and beliefs.
-
-\newpage{}
-
-\begin{idetres}
-
-Back Matter
-\end{idetres}
-
-\section*{References}\label{bibliography}
-\addcontentsline{toc}{section}{References}
-
-\phantomsection\label{refs}
-\begin{CSLReferences}{0}{0}
-\bibitem[\citeproctext]{ref-bourdieu_reproduction_1990}
-\CSLLeftMargin{1. }%
-\CSLRightInline{Bourdieu, P.; Passeron, J.C. \emph{Reproduction in
-{Education}, {Society} and {Culture}}; Second Edition.; Sage
-Publications Ltd, 1990; ISBN 0-8039-8320-4.}
-
-\bibitem[\citeproctext]{ref-vonhippel_test_2019}
-\CSLLeftMargin{2. }%
-\CSLRightInline{Von Hippel, P.; Hamrock, C. Do {Test Score Gaps Grow}
-Before, During, or Between the {School Years}? {Measurement Artifacts}
-and {What We Can Know} in {Spite} of {Them}. \emph{Sociological Science}
-\textbf{2019}, \emph{6}, 43--80,
-doi:\href{https://doi.org/10.15195/v6.a3}{10.15195/v6.a3}.}
-
-\bibitem[\citeproctext]{ref-batruch_belief_2022}
-\CSLLeftMargin{3. }%
-\CSLRightInline{Batruch, A.; Jetten, J.; Van de Werfhorst, H.; Darnon,
-C.; Butera, F. Belief in {School Meritocracy} and the {Legitimization}
-of {Social} and {Income Inequality}. \emph{Social Psychological and
-Personality Science} \textbf{2022}, 194855062211110,
-doi:\href{https://doi.org/10.1177/19485506221111017}{10.1177/19485506221111017}.}
-
-\bibitem[\citeproctext]{ref-wiederkehr_belief_2015}
-\CSLLeftMargin{4. }%
-\CSLRightInline{Wiederkehr, V.; Bonnot, V.; Krauth-Gruber, S.; Darnon,
-C. Belief in School Meritocracy as a System-Justifying Tool for Low
-Status Students. \emph{Frontiers in Psychology} \textbf{2015},
-\emph{6}.}
-
-\bibitem[\citeproctext]{ref-young_rise_1958}
-\CSLLeftMargin{5. }%
-\CSLRightInline{Young, M. \emph{The Rise of the Meritocracy};
-Transaction Publishers: New Brunswick, N.J., U.S.A, 1958;}
-
-\bibitem[\citeproctext]{ref-sandel_tyranny_2020}
-\CSLLeftMargin{6. }%
-\CSLRightInline{Sandel, M.J. \emph{The Tyranny of Merit: {What}'s Become
-of the Common Good?}; First edition.; {Farrar, Straus and Giroux}: New
-York, 2020; ISBN 978-0-374-28998-0.}
-
-\bibitem[\citeproctext]{ref-mcnamee_meritocracy_2004}
-\CSLLeftMargin{7. }%
-\CSLRightInline{McNamee, S.; Miller, R. \emph{The Meritocracy Myth};
-2004th ed.; Rowman \& Littlefield: Lanham Md., 2004; ISBN
-978-0-7425-1055-5.}
-
-\bibitem[\citeproctext]{ref-mijs_stratified_2016}
-\CSLLeftMargin{8. }%
-\CSLRightInline{Mijs, J. Stratified {Failure}: {Educational
-Stratification} and {Students}' {Attributions} of {Their Mathematics
-Performance} in 24 {Countries}. \emph{Sociology of Education}
-\textbf{2016}, \emph{89}, 137--153,
-doi:\href{https://doi.org/10.1177/0038040716636434}{10.1177/0038040716636434}.}
-
-\bibitem[\citeproctext]{ref-lane_market_1986}
-\CSLLeftMargin{9. }%
-\CSLRightInline{Lane, R.E. Market {Justice}, {Political Justice}.
-\emph{American Political Science Review} \textbf{1986}, \emph{80},
-383--402, doi:\href{https://doi.org/10.2307/1958264}{10.2307/1958264}.}
-
-\bibitem[\citeproctext]{ref-lindh_bringing_2023}
-\CSLLeftMargin{10. }%
-\CSLRightInline{Lindh, A.; McCall, L. Bringing the Market in: An
-Expanded Framework for Understanding Popular Responses to Economic
-Inequality. \emph{Socio-Economic Review} \textbf{2023}, \emph{21},
-1035--1055,
-doi:\href{https://doi.org/10.1093/ser/mwac018}{10.1093/ser/mwac018}.}
-
-\bibitem[\citeproctext]{ref-chafel_schooling_1997}
-\CSLLeftMargin{11. }%
-\CSLRightInline{Chafel, J.A. Schooling, the {Hidden Curriculum}, and
-{Children}'s {Conceptions} of {Poverty}. \emph{Social Policy Report}
-\textbf{1997}, \emph{11}, 1--28,
-doi:\href{https://doi.org/10.1002/j.2379-3988.1997.tb00004.x}{10.1002/j.2379-3988.1997.tb00004.x}.}
-
-\bibitem[\citeproctext]{ref-flores_top_2020}
-\CSLLeftMargin{12. }%
-\CSLRightInline{Flores, I.; Sanhueza, C.; Atria, J.; Mayer, R. Top
-{Incomes} in {Chile}: {A Historical Perspective} on {Income Inequality},
-1964--2017. \emph{Review of Income and Wealth} \textbf{2020}, \emph{66},
-850--874,
-doi:\href{https://doi.org/10.1111/roiw.12441}{10.1111/roiw.12441}.}
-
-\bibitem[\citeproctext]{ref-chancel_world_2022}
-\CSLLeftMargin{13. }%
-\CSLRightInline{Chancel, L.; Piketty, T.; Saez, E.; Zucman, G. World
-Inequality Report 2022 2022.}
-
-\bibitem[\citeproctext]{ref-ffrench-davis_reformas_2018}
-\CSLLeftMargin{14. }%
-\CSLRightInline{Ffrench-Davis, R. \emph{{Reformas econ{ó}micas en Chile
-1973-2017}}; TAURUS, 2018; ISBN 978-956-9635-21-2.}
-
-\bibitem[\citeproctext]{ref-boccardo_30_2020}
-\CSLLeftMargin{15. }%
-\CSLRightInline{Boccardo, G. \emph{30 a{ñ}os de Privatizaciones En
-{Chile}: Lo Que La Pandemia Revel{ó}}; Nodo XXI.; Santiago, 2020;}
-
-\bibitem[\citeproctext]{ref-observatoriosocial_estadisticas_2024}
-\CSLLeftMargin{16. }%
-\CSLRightInline{Social, O. \emph{Estad{í}sticas {Encuesta Casen} 2022,
-Secci{ó}n {Salud}}; Observatorio social, 2024;}
-
-\bibitem[\citeproctext]{ref-superintendenciadepensiones_estadisticas_2024}
-\CSLLeftMargin{17. }%
-\CSLRightInline{Superintendencia de Pensiones \emph{Estad{í}sticas de
-{Afiliados} En {AFP} En {Sistema} de {Pensiones}. {Abril} 2024};
-Superintendencia de Pensiones, 2024;}
-
-\bibitem[\citeproctext]{ref-ministeriodeeducacion_resumen_2023}
-\CSLLeftMargin{18. }%
-\CSLRightInline{Ministerio de Educación \emph{Resumen de Matr{í}cula Por
-Establecimiento Educacional}; Datos abiertos Ministerio de Educaci{ó}n,
-2023;}
-
-\bibitem[\citeproctext]{ref-boltanski_new_2005}
-\CSLLeftMargin{19. }%
-\CSLRightInline{Boltanski, L.; Chiapello, E. \emph{The New Spirit of
-Capitalism}; Verso: London ; New York, 2005; ISBN 978-1-85984-554-7.}
-
-\bibitem[\citeproctext]{ref-streeck_citizens_2012}
-\CSLLeftMargin{20. }%
-\CSLRightInline{Streeck, W. Citizens as {Customers}: {Considerations} on
-the {New Politics} of {Consumption}. \emph{New Left Review}
-\textbf{2012}, \emph{76}.}
-
-\bibitem[\citeproctext]{ref-kluegel_beliefs_1987}
-\CSLLeftMargin{21. }%
-\CSLRightInline{Kluegel, J.R.; Smith, E.R. \emph{Beliefs about
-{Inequality}: {Americans}' {Views} of {What} Is and {What Ought} to Be};
-Routledge: New York, 1987; ISBN 978-0-202-30327-7.}
-
-\bibitem[\citeproctext]{ref-wegener_dominant_1995}
-\CSLLeftMargin{22. }%
-\CSLRightInline{Wegener, B.; Liebig, S. Dominant {Ideologies} and the
-{Variation} of {Distributive Justice Norms}: {A Comparison} of {East}
-and {West Germany}, and the {United States}. In \emph{Social {Justice}
-and {Political Change}: {Public Opinion} in {Capitalist} and
-{Post-Communist States}}; Kluegel, J.R., Mason, D.S., Wegener, B., Eds.;
-Walter de Gruyter: New York, 1995; pp. 239--259 ISBN 978-3-11-086894-4.}
-
-\bibitem[\citeproctext]{ref-homan_being_2017}
-\CSLLeftMargin{23. }%
-\CSLRightInline{Homan, P.; Valentino, L.; Weed, E. Being and {Becoming
-Poor}: {How Cultural Schemas Shape Beliefs About Poverty}. \emph{Social
-Forces} \textbf{2017}, \emph{95}, 1023--1048,
-doi:\href{https://doi.org/10.1093/sf/sox007}{10.1093/sf/sox007}.}
-
-\bibitem[\citeproctext]{ref-castillo_legitimacy_2011}
-\CSLLeftMargin{24. }%
-\CSLRightInline{Castillo, J.C. Legitimacy of {Inequality} in a {Highly
-Unequal Context}: {Evidence} from the {Chilean Case}. \emph{Social
-Justice Research} \textbf{2011}, \emph{24}, 314--340,
-doi:\href{https://doi.org/10.1007/s11211-011-0144-5}{10.1007/s11211-011-0144-5}.}
-
-\bibitem[\citeproctext]{ref-evans_family_2010}
-\CSLLeftMargin{25. }%
-\CSLRightInline{Evans, M.; Kelley, J.; Sikora, J.; Treiman, D. Family
-Scholarly Culture and Educational Success: {Books} and Schooling in 27
-Nations. \emph{Research in Social Stratification and Mobility}
-\textbf{2010}, \emph{28}, 171--197,
-doi:\href{https://doi.org/10.1016/j.rssm.2010.01.002}{10.1016/j.rssm.2010.01.002}.}
-
-\bibitem[\citeproctext]{ref-jasso_how_1999}
-\CSLLeftMargin{26. }%
-\CSLRightInline{Jasso, G. How {Much Injustice} Is {There} in the
-{World}? {Two New Justice Indexes}. \emph{American Sociological Review}
-\textbf{1999}, \emph{64}, 133--168,
-doi:\href{https://doi.org/10.1177/000312249906400110}{10.1177/000312249906400110}.}
-
-\bibitem[\citeproctext]{ref-shariff_income_2016}
-\CSLLeftMargin{27. }%
-\CSLRightInline{Shariff, A.F.; Wiwad, D.; Aknin, L.B. Income {Mobility
-Breeds Tolerance} for {Income Inequality}: {Cross-National} and
-{Experimental Evidence}. \emph{Perspectives on Psychological Science}
-\textbf{2016}, \emph{11}, 373--380,
-doi:\href{https://doi.org/10.1177/1745691616635596}{10.1177/1745691616635596}.}
-
-\bibitem[\citeproctext]{ref-centeno_arc_2012}
-\CSLLeftMargin{28. }%
-\CSLRightInline{Centeno, M.A.; Cohen, J.N. The {Arc} of {Neoliberalism}.
-\emph{Annual Review of Sociology} \textbf{2012}, \emph{38}, 317--340,
-doi:\href{https://doi.org/10.1146/annurev-soc-081309-150235}{10.1146/annurev-soc-081309-150235}.}
-
-\bibitem[\citeproctext]{ref-harvey_breve_2015}
-\CSLLeftMargin{29. }%
-\CSLRightInline{Harvey, D. \emph{{Breve historia del neoliberalismo}};
-Ediciones Akal: Madrid (Espa{ñ}a), 2015; ISBN 978-84-460-2517-7.}
-
-\bibitem[\citeproctext]{ref-jensen_worlds_2008}
-\CSLLeftMargin{30. }%
-\CSLRightInline{Jensen, C. Worlds of Welfare Services and Transfers.
-\emph{Journal of European Social Policy} \textbf{2008}, \emph{18},
-151--162,
-doi:\href{https://doi.org/10.1177/0958928707087591}{10.1177/0958928707087591}.}
-
-\bibitem[\citeproctext]{ref-stoy_worlds_2014}
-\CSLLeftMargin{31. }%
-\CSLRightInline{Stoy, V. Worlds of {Welfare Services}: {From Discovery}
-to {Exploration}. \emph{Social Policy \& Administration} \textbf{2014},
-\emph{48}, 343--360,
-doi:\href{https://doi.org/10.1111/spol.12006}{10.1111/spol.12006}.}
-
-\bibitem[\citeproctext]{ref-salamon_marketization_1993}
-\CSLLeftMargin{32. }%
-\CSLRightInline{Salamon, L.M. The {Marketization} of {Welfare}:
-{Changing Nonprofit} and {For-Profit Roles} in the {American Welfare
-State}. \emph{Social Service Review} \textbf{1993}, \emph{67}, 16--39,
-doi:\href{https://doi.org/10.1086/603963}{10.1086/603963}.}
-
-\bibitem[\citeproctext]{ref-sivesind_changing_2017}
-\CSLLeftMargin{33. }%
-\CSLRightInline{Sivesind, K.H.
-\href{https://doi.org/10.1007/978-3-319-55381-8_2}{The {Changing Roles}
-of {For-Profit} and {Nonprofit Welfare Provision} in {Norway}, {Sweden},
-and {Denmark}}. In \emph{Promoting {Active Citizenship}}; Sivesind,
-K.H., Saglie, J., Eds.; Springer International Publishing: Cham, 2017;
-pp. 33--74 ISBN 978-3-319-55380-1 978-3-319-55381-8.}
-
-\bibitem[\citeproctext]{ref-lindh_public_2015}
-\CSLLeftMargin{34. }%
-\CSLRightInline{Lindh, A. Public {Opinion} Against {Markets}?
-{Attitudes} Towards {Market Distribution} of {Social Services} -- {A
-Comparison} of 17 {Countries}. \emph{Social Policy \& Administration}
-\textbf{2015}, \emph{49}, 887--910,
-doi:\href{https://doi.org/10.1111/spol.12105}{10.1111/spol.12105}.}
-
-\bibitem[\citeproctext]{ref-greitemeyer_subjective_2016}
-\CSLLeftMargin{35. }%
-\CSLRightInline{Greitemeyer, T.; Sagioglou, C. Subjective Socioeconomic
-Status Causes Aggression: {A} Test of the Theory of Social Deprivation.
-\emph{Journal of Personality and Social Psychology} \textbf{2016},
-\emph{111}, 178--194,
-doi:\href{https://doi.org/10.1037/pspi0000058}{10.1037/pspi0000058}.}
-
-\bibitem[\citeproctext]{ref-mishra_subjective_2015}
-\CSLLeftMargin{36. }%
-\CSLRightInline{Mishra, S.; Carleton, R.N. Subjective Relative
-Deprivation Is Associated with Poorer Physical and Mental Health.
-\emph{Social Science \& Medicine} \textbf{2015}, \emph{147}, 144--149,
-doi:\href{https://doi.org/10.1016/j.socscimed.2015.10.030}{10.1016/j.socscimed.2015.10.030}.}
-
-\bibitem[\citeproctext]{ref-smith_relative_2012}
-\CSLLeftMargin{37. }%
-\CSLRightInline{Smith, H.J.; Pettigrew, T.F.; Pippin, G.M.;
-Bialosiewicz, S. Relative {Deprivation}: {A Theoretical} and
-{Meta-Analytic Review}. \emph{Personality and Social Psychology Review}
-\textbf{2012}, \emph{16}, 203--232,
-doi:\href{https://doi.org/10.1177/1088868311430825}{10.1177/1088868311430825}.}
-
-\bibitem[\citeproctext]{ref-power_deprivationprotest_2018}
-\CSLLeftMargin{38. }%
-\CSLRightInline{Power, S.A. The {Deprivation-Protest Paradox}: {How} the
-{Perception} of {Unfair Economic Inequality Leads} to {Civic Unrest}.
-\emph{Current Anthropology} \textbf{2018}, \emph{59}, 765--789,
-doi:\href{https://doi.org/10.1086/700679}{10.1086/700679}.}
-
-\bibitem[\citeproctext]{ref-wilson_role_2003}
-\CSLLeftMargin{39. }%
-\CSLRightInline{Wilson, C. The {Role} of a {Merit Principle} in
-{Distributive Justice}. \emph{The Journal of Ethics} \textbf{2003},
-\emph{7}, 277--314,
-doi:\href{https://doi.org/10.1023/A:1024667228488}{10.1023/A:1024667228488}.}
-
-\bibitem[\citeproctext]{ref-bucca_merit_2016}
-\CSLLeftMargin{40. }%
-\CSLRightInline{Bucca, M. Merit and Blame in Unequal Societies:
-{Explaining Latin Americans}' Beliefs about Wealth and Poverty.
-\emph{Research in Social Stratification and Mobility} \textbf{2016},
-\emph{44}, 98--112,
-doi:\href{https://doi.org/10.1016/j.rssm.2016.02.005}{10.1016/j.rssm.2016.02.005}.}
-
-\bibitem[\citeproctext]{ref-chong_mystery_2008}
-\CSLLeftMargin{41. }%
-\CSLRightInline{Chong, A.; Ñopo, H. The {Mystery} of {Discrimination} in
-{Latin America}. \emph{Econom{í}a} \textbf{2008}, \emph{8}, 79--107,
-doi:\href{https://doi.org/10.1353/eco.0.0005}{10.1353/eco.0.0005}.}
-
-\bibitem[\citeproctext]{ref-torche_intergenerational_2014}
-\CSLLeftMargin{42. }%
-\CSLRightInline{Torche, F. Intergenerational {Mobility} and
-{Inequality}: {The Latin American Case}. \emph{Annual Review of
-Sociology} \textbf{2014}, \emph{40}, 619--642,
-doi:\href{https://doi.org/10.1146/annurev-soc-071811-145521}{10.1146/annurev-soc-071811-145521}.}
-
-\bibitem[\citeproctext]{ref-salgado_inequality_2023}
-\CSLLeftMargin{43. }%
-\CSLRightInline{Salgado, M.; Castillo, J. Inequality and
-{Stratification} in {Latin America}. In \emph{The {Oxford Handbook} of
-{Social Stratification}}; Gangl, M., Platt, L., Polavieja, J.G., van de
-Werfhorst, H.G., Eds.; Oxford University Press: 2023, 2023; p. 0 ISBN
-978-0-19-753948-4.}
-
-\bibitem[\citeproctext]{ref-molyneux_neoliberal_2008}
-\CSLLeftMargin{44. }%
-\CSLRightInline{Molyneux, M. The {``{Neoliberal Turn}''} and the {New
-Social Policy} in {Latin America}: {How Neoliberal}, {How New}?
-\emph{Development and Change} \textbf{2008}, \emph{39}, 775--797,
-doi:\href{https://doi.org/10.1111/j.1467-7660.2008.00505.x}{10.1111/j.1467-7660.2008.00505.x}.}
-
-\bibitem[\citeproctext]{ref-centrodeestudiospublicos_estudio_2024}
-\CSLLeftMargin{45. }%
-\CSLRightInline{Centro de Estudios Públicos \emph{Estudio {Nacional} de
-{Opini{ó}n P{ú}blica Base Consolidada}}; Centro de Estudios P{ú}blicos,
-2024;}
-
-\bibitem[\citeproctext]{ref-goldthorpe_myth_2003}
-\CSLLeftMargin{46. }%
-\CSLRightInline{Goldthorpe, J. The Myth of Education-Based Meritocracy.
-\emph{New Economy} \textbf{2003}, \emph{10}, 234--239,
-doi:\href{https://doi.org/10.1046/j.1468-0041.2003.00324.x}{10.1046/j.1468-0041.2003.00324.x}.}
-
-\bibitem[\citeproctext]{ref-castillo_multidimensional_2023}
-\CSLLeftMargin{47. }%
-\CSLRightInline{Castillo, J.C.; Iturra, J.; Maldonado, L.; Atria, J.;
-Meneses, F. A {Multidimensional Approach} for {Measuring Meritocratic
-Beliefs}: {Advantages}, {Limitations} and {Alternatives} to the {ISSP
-Social Inequality Survey}. \emph{International Journal of Sociology}
-\textbf{2023}, \emph{53}, 448--472,
-doi:\href{https://doi.org/10.1080/00207659.2023.2274712}{10.1080/00207659.2023.2274712}.}
-
-\bibitem[\citeproctext]{ref-mijs_unfulfillable_2016}
-\CSLLeftMargin{48. }%
-\CSLRightInline{Mijs, J. The {Unfulfillable Promise} of {Meritocracy}:
-{Three Lessons} and {Their Implications} for {Justice} in {Education}.
-\emph{Social Justice Research} \textbf{2016}, \emph{29}, 14--34,
-doi:\href{https://doi.org/10.1007/s11211-014-0228-0}{10.1007/s11211-014-0228-0}.}
-
-\bibitem[\citeproctext]{ref-mijs_paradox_2019}
-\CSLLeftMargin{49. }%
-\CSLRightInline{Mijs, J. The Paradox of Inequality: Income Inequality
-and Belief in Meritocracy Go Hand in Hand. \emph{Socio-Economic Review}
-\textbf{2019}, \emph{19}, 7--35,
-doi:\href{https://doi.org/10.1093/ser/mwy051}{10.1093/ser/mwy051}.}
-
-\bibitem[\citeproctext]{ref-duru-bellat_who_2012}
-\CSLLeftMargin{50. }%
-\CSLRightInline{Duru-Bellat, M.; Tenret, E. Who's for {Meritocracy}?
-{Individual} and {Contextual Variations} in the {Faith}.
-\emph{Comparative Education Review} \textbf{2012}, \emph{56}, 223--247,
-doi:\href{https://doi.org/10.1086/661290}{10.1086/661290}.}
-
-\bibitem[\citeproctext]{ref-castillo_meritocracia_2019}
-\CSLLeftMargin{51. }%
-\CSLRightInline{Castillo, J.C.; Torres, A.; Atria, J.; Maldonado, L.
-Meritocracia y Desigualdad Econ{ó}mica: {Percepciones}, Preferencias e
-Implicancias. \emph{Revista Internacional de Sociolog{í}a}
-\textbf{2019}, \emph{77}, 117,
-doi:\href{https://doi.org/10.3989/ris.2019.77.1.17.114}{10.3989/ris.2019.77.1.17.114}.}
-
-\bibitem[\citeproctext]{ref-lampert_meritocratic_2013}
-\CSLLeftMargin{52. }%
-\CSLRightInline{Lampert, K.
-\emph{\href{https://doi.org/10.1057/9781137324894}{Meritocratic
-{Education} and {Social Worthlessness}}}; Palgrave Macmillan UK: London,
-2013; ISBN 978-1-349-45908-7 978-1-137-32489-4.}
-
-\bibitem[\citeproctext]{ref-reynolds_perceptions_2014}
-\CSLLeftMargin{53. }%
-\CSLRightInline{Reynolds, J.; Xian, H. Perceptions of Meritocracy in the
-Land of Opportunity. \emph{Research in Social Stratification and
-Mobility} \textbf{2014}, \emph{36}, 121--137,
-doi:\href{https://doi.org/10.1016/j.rssm.2014.03.001}{10.1016/j.rssm.2014.03.001}.}
-
-\bibitem[\citeproctext]{ref-morris_representing_2022}
-\CSLLeftMargin{54. }%
-\CSLRightInline{Morris, J.; Reilly, J.; Paltsev, S.; Sokolov, A.; Cox,
-K. Representing {Socio}-{Economic Uncertainty} in {Human System Models}.
-\emph{Earth's Future} \textbf{2022}, \emph{10}, e2021EF002239,
-doi:\href{https://doi.org/10.1029/2021EF002239}{10.1029/2021EF002239}.}
-
-\bibitem[\citeproctext]{ref-frank_performance_2015}
-\CSLLeftMargin{55. }%
-\CSLRightInline{Frank, D.H.; Wertenbroch, K.; Maddux, W.W. Performance
-Pay or Redistribution? {Cultural} Differences in Just-World Beliefs and
-Preferences for Wage Inequality. \emph{Organizational Behavior and Human
-Decision Processes} \textbf{2015}, \emph{130}, 160--170,
-doi:\href{https://doi.org/10.1016/j.obhdp.2015.04.001}{10.1016/j.obhdp.2015.04.001}.}
-
-\bibitem[\citeproctext]{ref-almas_cutthroat_2020}
-\CSLLeftMargin{56. }%
-\CSLRightInline{Almås, I.; Cappelen, A.W.; Tungodden, B. Cutthroat
-{Capitalism} Versus {Cuddly Socialism}: {Are Americans More
-Meritocratic} and {Efficiency-Seeking} Than {Scandinavians}?
-\emph{Journal of Political Economy} \textbf{2020}, \emph{128},
-1753--1788, doi:\href{https://doi.org/10.1086/705551}{10.1086/705551}.}
-
-\bibitem[\citeproctext]{ref-barr_effect_2020}
-\CSLLeftMargin{57. }%
-\CSLRightInline{Barr, A.; Miller, L. The Effect of Education, Income
-Inequality and Merit on Inequality Acceptance. \emph{Journal of Economic
-Psychology} \textbf{2020}, \emph{80}, 102276,
-doi:\href{https://doi.org/10.1016/j.joep.2020.102276}{10.1016/j.joep.2020.102276}.}
-
-\bibitem[\citeproctext]{ref-evans_strong_2018}
-\CSLLeftMargin{58. }%
-\CSLRightInline{Evans, M.D.R.; Kelley, J. Strong {Welfare States Do Not
-Intensify Public Support} for {Income Redistribution}, but {Even Reduce
-It} Among the {Prosperous}: {A Multilevel Analysis} of {Public Opinion}
-in 30 {Countries}. \emph{Societies} \textbf{2018}, \emph{8}, 105,
-doi:\href{https://doi.org/10.3390/soc8040105}{10.3390/soc8040105}.}
-
-\bibitem[\citeproctext]{ref-mikus_children_2019}
-\CSLLeftMargin{59. }%
-\CSLRightInline{Mikus, K.; Tieben, N.; Schober, P.S. Children's
-Conversion of Cultural Capital into Educational Success: The Symbolic
-and Skill-Generating Functions of Cultural Capital. \emph{British
-Journal of Sociology of Education} \textbf{2019}, \emph{41}, 197--2017,
-doi:\href{https://doi.org/10.1080/01425692.2019.1677454}{10.1080/01425692.2019.1677454}.}
-
-\bibitem[\citeproctext]{ref-jack_no_2016}
-\CSLLeftMargin{60. }%
-\CSLRightInline{Jack, A.A. ({No}) {Harm} in {Asking}: {Class}, {Acquired
-Cultural Capital}, and {Academic Engagement} at an {Elite University}.
-\emph{Sociology of Education} \textbf{2016}, \emph{89}, 1--19,
-doi:\href{https://doi.org/10.1177/0038040715614913}{10.1177/0038040715614913}.}
-
-\bibitem[\citeproctext]{ref-khan_privilege_2011}
-\CSLLeftMargin{61. }%
-\CSLRightInline{Khan, S. \emph{Privilege: {The Making} of an {Adolescent
-Elite} at {St}. {Paul}'s {School}}; Princeton University Press:
-Princeton, N.J., 2011;}
-
-\bibitem[\citeproctext]{ref-goudeau_hidden_2017}
-\CSLLeftMargin{62. }%
-\CSLRightInline{Goudeau, S.; Croizet, J.-C. Hidden {Advantages} and
-{Disadvantages} of {Social Class}: {How Classroom Settings Reproduce
-Social Inequality} by {Staging Unfair Comparison}. \emph{Psychological
-Science} \textbf{2017}, \emph{28}, 162--170,
-doi:\href{https://doi.org/10.1177/0956797616676600}{10.1177/0956797616676600}.}
-
-\bibitem[\citeproctext]{ref-darnon_where_2018}
-\CSLLeftMargin{63. }%
-\CSLRightInline{Darnon, C.; Wiederkehr, V.; Dompnier, B.; Martinot, D.
-{``{Where} There Is a Will, There Is a Way''}: {Belief} in School
-Meritocracy and the Social-Class Achievement Gap. \emph{British Journal
-of Social Psychology} \textbf{2018}, \emph{57}, 250--262,
-doi:\href{https://doi.org/10.1111/bjso.12214}{10.1111/bjso.12214}.}
-
-\bibitem[\citeproctext]{ref-resh_sense_2010}
-\CSLLeftMargin{64. }%
-\CSLRightInline{Resh, N. Sense of Justice about Grades in School: Is It
-Stratified Like Academic Achievement? \emph{Social Psychology of
-Education} \textbf{2010}, \emph{13}, 313--329,
-doi:\href{https://doi.org/10.1007/s11218-010-9117-z}{10.1007/s11218-010-9117-z}.}
-
-\bibitem[\citeproctext]{ref-resh_sense_2014}
-\CSLLeftMargin{65. }%
-\CSLRightInline{Resh, N.; Sabbagh, C. Sense of Justice in School and
-Civic Attitudes. \emph{Social Psychology of Education} \textbf{2014},
-\emph{17}, 51--72,
-doi:\href{https://doi.org/10.1007/s11218-013-9240-8}{10.1007/s11218-013-9240-8}.}
-
-\bibitem[\citeproctext]{ref-kohn_social_1963}
-\CSLLeftMargin{66. }%
-\CSLRightInline{Kohn, M.L. Social {Class} and {Parent-Child
-Relationships}: {An Interpretation}. \emph{American Journal of
-Sociology} \textbf{1963}, \emph{68}, 471--480,
-doi:\href{https://doi.org/10.1086/223403}{10.1086/223403}.}
-
-\bibitem[\citeproctext]{ref-kohn_class_1969}
-\CSLLeftMargin{67. }%
-\CSLRightInline{Kohn, M.L.; Schooler, C. Class, {Occupation}, and
-{Orientation}. \emph{American Sociological Review} \textbf{1969},
-\emph{34}, 659--678,
-doi:\href{https://doi.org/10.2307/2092303}{10.2307/2092303}.}
-
-\bibitem[\citeproctext]{ref-acemoglu_obedience_2021}
-\CSLLeftMargin{68. }%
-\CSLRightInline{Acemoglu, D.
-\emph{\href{https://doi.org/10.3386/w29125}{Obedience in the {Labor
-Market} and {Social Mobility}: {A Socio-Economic Approach}}}; National
-Bureau of Economic Research: Cambridge, MA, 2021; p. w29125;}
-
-\bibitem[\citeproctext]{ref-iacoviello_collectivism_2019}
-\CSLLeftMargin{69. }%
-\CSLRightInline{Iacoviello, V.; Lorenzi-Cioldi, F. Collectivism and
-{Individualism} in {Status Hierarchies}: {Socialization} and {Social
-Identity Explanations}. \emph{International Review of Social Psychology}
-\textbf{2019}, \emph{32}, 15,
-doi:\href{https://doi.org/10.5334/irsp.285}{10.5334/irsp.285}.}
-
-\bibitem[\citeproctext]{ref-almas_fairness_2017}
-\CSLLeftMargin{70. }%
-\CSLRightInline{Almås, I.; Cappelen, A.W.; Salvanes, K.G.; Sørensen,
-E.Ø.; Tungodden, B. Fairness and Family Background. \emph{Politics,
-Philosophy \& Economics} \textbf{2017}, \emph{16}, 117--131,
-doi:\href{https://doi.org/10.1177/1470594x15618966}{10.1177/1470594x15618966}.}
-
-\bibitem[\citeproctext]{ref-jonsson_institutional_2015}
-\CSLLeftMargin{71. }%
-\CSLRightInline{Jonsson, A.-C.; Beach, D. Institutional Discrimination:
-{Stereotypes} and Social Reproduction of {``Class''} in the {Swedish}
-Upper-Secondary School. \emph{Social Psychology of Education}
-\textbf{2015}, \emph{18}, 703--717,
-doi:\href{https://doi.org/10.1007/s11218-014-9279-1}{10.1007/s11218-014-9279-1}.}
-
-\bibitem[\citeproctext]{ref-jost_attitudinal_2000}
-\CSLLeftMargin{72. }%
-\CSLRightInline{Jost, J.T.; Burgess, D. Attitudinal {Ambivalence} and
-the {Conflict} Between {Group} and {System Justification Motives} in
-{Low Status Groups}. \emph{Personality and Social Psychology Bulletin}
-\textbf{2000}, \emph{26}, 293--305,
-doi:\href{https://doi.org/10.1177/0146167200265003}{10.1177/0146167200265003}.}
-
-\bibitem[\citeproctext]{ref-gottfried_autocorrelation_2022}
-\CSLLeftMargin{73. }%
-\CSLRightInline{Gottfried, J.; Ježek, S.; Králová, M.; Řiháček, T.
-Autocorrelation {Screening}: {A Potentially Efficient Method} for
-{Detecting Repetitive Response Patterns} in {Questionnaire Data}.
-\textbf{2022},
-doi:\href{https://doi.org/10.7275/VYXB-GT24}{10.7275/VYXB-GT24}.}
-
-\bibitem[\citeproctext]{ref-bell_fixed_2019}
-\CSLLeftMargin{74. }%
-\CSLRightInline{Bell, A.; Fairbrother, M.; Jones, K. Fixed and Random
-Effects Models: Making an Informed Choice. \emph{Quality \& Quantity}
-\textbf{2019}, \emph{53}, 1051--1074,
-doi:\href{https://doi.org/10.1007/s11135-018-0802-x}{10.1007/s11135-018-0802-x}.}
-
-\bibitem[\citeproctext]{ref-hox_multilevel_2010}
-\CSLLeftMargin{75. }%
-\CSLRightInline{Hox, J.; Moerbeek, M.; Schoot, R. van de
-\emph{\href{https://doi.org/10.4324/9780203852279}{Multilevel
-{Analysis}: {Techniques} and {Applications}, {Second Edition}}}; 2nd
-ed.; Routledge: New York, 2010; ISBN 978-0-203-85227-9.}
-
-\bibitem[\citeproctext]{ref-kim_socioeconomic_2020}
-\CSLLeftMargin{76. }%
-\CSLRightInline{Kim, C.-Y.; Gewirtz, S. Socioeconomic Inequality and
-Adjustments in Children's Perceptions of Their Agency as They Age in
-{South Korea}. \emph{Children \& Society} \textbf{2020}, \emph{34},
-371--391,
-doi:\href{https://doi.org/10.1111/chso.12372}{10.1111/chso.12372}.}
-
-\bibitem[\citeproctext]{ref-rizzo_children_2020}
-\CSLLeftMargin{77. }%
-\CSLRightInline{Rizzo, M.T.; Killen, M. Children's Evaluations of
-Individually and Structurally Based Inequalities: {The} Role of Status.
-\emph{Developmental Psychology} \textbf{2020}, \emph{56}, 2223--2235,
-doi:\href{https://doi.org/10.1037/dev0001118}{10.1037/dev0001118}.}
-
-\bibitem[\citeproctext]{ref-gonthier_parallel_2017}
-\CSLLeftMargin{78. }%
-\CSLRightInline{Gonthier, F. Parallel Publics? {Support} for Income
-Redistribution in Times of Economic Crisis. \emph{European Journal of
-Political Research} \textbf{2017}, \emph{56}, 92--114,
-doi:\href{https://doi.org/10.1111/1475-6765.12168}{10.1111/1475-6765.12168}.}
-
-\bibitem[\citeproctext]{ref-runst_does_2018}
-\CSLLeftMargin{79. }%
-\CSLRightInline{Runst, P. Does {Immigration Affect Demand} for
-{Redistribution}? -- {An Experimental Design}. \emph{German Economic
-Review} \textbf{2018}, \emph{19}, 383--400,
-doi:\href{https://doi.org/10.1111/geer.12133}{10.1111/geer.12133}.}
-
-\bibitem[\citeproctext]{ref-bullock_education_2021}
-\CSLLeftMargin{80. }%
-\CSLRightInline{Bullock, J.G. Education and {Attitudes} Toward
-{Redistribution} in the {United States}. \emph{British Journal of
-Political Science} \textbf{2021}, \emph{51}, 1230--1250,
-doi:\href{https://doi.org/10.1017/S0007123419000504}{10.1017/S0007123419000504}.}
-
-\end{CSLReferences}
 
@@
 
+\section*{References}\label{bibliography}
+\addcontentsline{toc}{section}{References}
+
+\phantomsection\label{refs}
+\begin{CSLReferences}{0}{0}
+\bibitem[\citeproctext]{ref-bourdieu_reproduction_1990}
+\CSLLeftMargin{1. }%
+\CSLRightInline{Bourdieu, P.; Passeron, J.C. \emph{Reproduction in
+{Education}, {Society} and {Culture}}; Second Edition.; Sage
+Publications Ltd, 1990; ISBN 0-8039-8320-4.}
+
+\bibitem[\citeproctext]{ref-vonhippel_test_2019}
+\CSLLeftMargin{2. }%
+\CSLRightInline{Von Hippel, P.; Hamrock, C. Do {Test Score Gaps Grow}
+Before, During, or Between the {School Years}? {Measurement Artifacts}
+and {What We Can Know} in {Spite} of {Them}. \emph{Sociological Science}
+\textbf{2019}, \emph{6}, 43--80,
+doi:\href{https://doi.org/10.15195/v6.a3}{10.15195/v6.a3}.}
+
+\bibitem[\citeproctext]{ref-batruch_belief_2022}
+\CSLLeftMargin{3. }%
+\CSLRightInline{Batruch, A.; Jetten, J.; Van de Werfhorst, H.; Darnon,
+C.; Butera, F. Belief in {School Meritocracy} and the {Legitimization}
+of {Social} and {Income Inequality}. \emph{Social Psychological and
+Personality Science} \textbf{2022}, 194855062211110,
+doi:\href{https://doi.org/10.1177/19485506221111017}{10.1177/19485506221111017}.}
+
+\bibitem[\citeproctext]{ref-wiederkehr_belief_2015}
+\CSLLeftMargin{4. }%
+\CSLRightInline{Wiederkehr, V.; Bonnot, V.; Krauth-Gruber, S.; Darnon,
+C. Belief in School Meritocracy as a System-Justifying Tool for Low
+Status Students. \emph{Frontiers in Psychology} \textbf{2015},
+\emph{6}.}
+
+\bibitem[\citeproctext]{ref-young_rise_1958}
+\CSLLeftMargin{5. }%
+\CSLRightInline{Young, M. \emph{The Rise of the Meritocracy};
+Transaction Publishers: New Brunswick, N.J., U.S.A, 1958;}
+
+\bibitem[\citeproctext]{ref-sandel_tyranny_2020}
+\CSLLeftMargin{6. }%
+\CSLRightInline{Sandel, M.J. \emph{The Tyranny of Merit: {What}'s Become
+of the Common Good?}; First edition.; {Farrar, Straus and Giroux}: New
+York, 2020; ISBN 978-0-374-28998-0.}
+
+\bibitem[\citeproctext]{ref-mcnamee_meritocracy_2004}
+\CSLLeftMargin{7. }%
+\CSLRightInline{McNamee, S.; Miller, R. \emph{The Meritocracy Myth};
+2004th ed.; Rowman \& Littlefield: Lanham Md., 2004; ISBN
+978-0-7425-1055-5.}
+
+\bibitem[\citeproctext]{ref-mijs_stratified_2016}
+\CSLLeftMargin{8. }%
+\CSLRightInline{Mijs, J. Stratified {Failure}: {Educational
+Stratification} and {Students}' {Attributions} of {Their Mathematics
+Performance} in 24 {Countries}. \emph{Sociology of Education}
+\textbf{2016}, \emph{89}, 137--153,
+doi:\href{https://doi.org/10.1177/0038040716636434}{10.1177/0038040716636434}.}
+
+\bibitem[\citeproctext]{ref-lane_market_1986}
+\CSLLeftMargin{9. }%
+\CSLRightInline{Lane, R.E. Market {Justice}, {Political Justice}.
+\emph{American Political Science Review} \textbf{1986}, \emph{80},
+383--402, doi:\href{https://doi.org/10.2307/1958264}{10.2307/1958264}.}
+
+\bibitem[\citeproctext]{ref-lindh_bringing_2023}
+\CSLLeftMargin{10. }%
+\CSLRightInline{Lindh, A.; McCall, L. Bringing the Market in: An
+Expanded Framework for Understanding Popular Responses to Economic
+Inequality. \emph{Socio-Economic Review} \textbf{2023}, \emph{21},
+1035--1055,
+doi:\href{https://doi.org/10.1093/ser/mwac018}{10.1093/ser/mwac018}.}
+
+\bibitem[\citeproctext]{ref-lindh_public_2015}
+\CSLLeftMargin{11. }%
+\CSLRightInline{Lindh, A. Public {Opinion} Against {Markets}?
+{Attitudes} Towards {Market Distribution} of {Social Services} -- {A
+Comparison} of 17 {Countries}. \emph{Social Policy \& Administration}
+\textbf{2015}, \emph{49}, 887--910,
+doi:\href{https://doi.org/10.1111/spol.12105}{10.1111/spol.12105}.}
+
+\bibitem[\citeproctext]{ref-castillo_meritocracia_2019}
+\CSLLeftMargin{12. }%
+\CSLRightInline{Castillo, J.C.; Torres, A.; Atria, J.; Maldonado, L.
+Meritocracia y Desigualdad Econ{ó}mica: {Percepciones}, Preferencias e
+Implicancias. \emph{Revista Internacional de Sociolog{í}a}
+\textbf{2019}, \emph{77}, 117,
+doi:\href{https://doi.org/10.3989/ris.2019.77.1.17.114}{10.3989/ris.2019.77.1.17.114}.}
+
+\bibitem[\citeproctext]{ref-reynolds_perceptions_2014}
+\CSLLeftMargin{13. }%
+\CSLRightInline{Reynolds, J.; Xian, H. Perceptions of Meritocracy in the
+Land of Opportunity. \emph{Research in Social Stratification and
+Mobility} \textbf{2014}, \emph{36}, 121--137,
+doi:\href{https://doi.org/10.1016/j.rssm.2014.03.001}{10.1016/j.rssm.2014.03.001}.}
+
+\bibitem[\citeproctext]{ref-garcia-sierra_dark_2023}
+\CSLLeftMargin{14. }%
+\CSLRightInline{García-Sierra, A. The Dark Side of Meritocratic Beliefs:
+{Is} Believing in Meritocracy Detrimental to Individuals from Low
+Socioeconomic Backgrounds? \emph{Social Justice Research} \textbf{2023},
+\emph{36}, 385--409,
+doi:\href{https://doi.org/10.1007/s11211-023-00413-x}{10.1007/s11211-023-00413-x}.}
+
+\bibitem[\citeproctext]{ref-flores_top_2020}
+\CSLLeftMargin{15. }%
+\CSLRightInline{Flores, I.; Sanhueza, C.; Atria, J.; Mayer, R. Top
+{Incomes} in {Chile}: {A Historical Perspective} on {Income Inequality},
+1964--2017. \emph{Review of Income and Wealth} \textbf{2020}, \emph{66},
+850--874,
+doi:\href{https://doi.org/10.1111/roiw.12441}{10.1111/roiw.12441}.}
+
+\bibitem[\citeproctext]{ref-chancel_world_2022}
+\CSLLeftMargin{16. }%
+\CSLRightInline{Chancel, L.; Piketty, T.; Saez, E.; Zucman, G. World
+Inequality Report 2022 2022.}
+
+\bibitem[\citeproctext]{ref-ffrench-davis_reformas_2018}
+\CSLLeftMargin{17. }%
+\CSLRightInline{Ffrench-Davis, R. \emph{{Reformas econ{ó}micas en Chile
+1973-2017}}; TAURUS, 2018; ISBN 978-956-9635-21-2.}
+
+\bibitem[\citeproctext]{ref-boccardo_30_2020}
+\CSLLeftMargin{18. }%
+\CSLRightInline{Boccardo, G. \emph{30 a{ñ}os de Privatizaciones En
+{Chile}: Lo Que La Pandemia Revel{ó}}; Nodo XXI.; Santiago, 2020;}
+
+\bibitem[\citeproctext]{ref-observatoriosocial_estadisticas_2024}
+\CSLLeftMargin{19. }%
+\CSLRightInline{Social, O. \emph{Estad{í}sticas {Encuesta Casen} 2022,
+Secci{ó}n {Salud}}; Observatorio social, 2024;}
+
+\bibitem[\citeproctext]{ref-superintendenciadepensiones_estadisticas_2024}
+\CSLLeftMargin{20. }%
+\CSLRightInline{Superintendencia de Pensiones \emph{Estad{í}sticas de
+{Afiliados} En {AFP} En {Sistema} de {Pensiones}. {Abril} 2024};
+Superintendencia de Pensiones, 2024;}
+
+\bibitem[\citeproctext]{ref-ministeriodeeducacion_resumen_2023}
+\CSLLeftMargin{21. }%
+\CSLRightInline{Ministerio de Educación \emph{Resumen de Matr{í}cula Por
+Establecimiento Educacional}; Datos abiertos Ministerio de Educaci{ó}n,
+2023;}
+
+\bibitem[\citeproctext]{ref-boltanski_new_2005}
+\CSLLeftMargin{22. }%
+\CSLRightInline{Boltanski, L.; Chiapello, E. \emph{The New Spirit of
+Capitalism}; Verso: London ; New York, 2005; ISBN 978-1-85984-554-7.}
+
+\bibitem[\citeproctext]{ref-streeck_citizens_2012}
+\CSLLeftMargin{23. }%
+\CSLRightInline{Streeck, W. Citizens as {Customers}: {Considerations} on
+the {New Politics} of {Consumption}. \emph{New Left Review}
+\textbf{2012}, \emph{76}.}
+
+\bibitem[\citeproctext]{ref-kluegel_beliefs_1987}
+\CSLLeftMargin{24. }%
+\CSLRightInline{Kluegel, J.R.; Smith, E.R. \emph{Beliefs about
+{Inequality}: {Americans}' {Views} of {What} Is and {What Ought} to Be};
+Routledge: New York, 1987; ISBN 978-0-202-30327-7.}
+
+\bibitem[\citeproctext]{ref-wegener_dominant_1995}
+\CSLLeftMargin{25. }%
+\CSLRightInline{Wegener, B.; Liebig, S. Dominant {Ideologies} and the
+{Variation} of {Distributive Justice Norms}: {A Comparison} of {East}
+and {West Germany}, and the {United States}. In \emph{Social {Justice}
+and {Political Change}: {Public Opinion} in {Capitalist} and
+{Post-Communist States}}; Kluegel, J.R., Mason, D.S., Wegener, B., Eds.;
+Walter de Gruyter: New York, 1995; pp. 239--259 ISBN 978-3-11-086894-4.}
+
+\bibitem[\citeproctext]{ref-homan_being_2017}
+\CSLLeftMargin{26. }%
+\CSLRightInline{Homan, P.; Valentino, L.; Weed, E. Being and {Becoming
+Poor}: {How Cultural Schemas Shape Beliefs About Poverty}. \emph{Social
+Forces} \textbf{2017}, \emph{95}, 1023--1048,
+doi:\href{https://doi.org/10.1093/sf/sox007}{10.1093/sf/sox007}.}
+
+\bibitem[\citeproctext]{ref-castillo_legitimacy_2011}
+\CSLLeftMargin{27. }%
+\CSLRightInline{Castillo, J.C. Legitimacy of {Inequality} in a {Highly
+Unequal Context}: {Evidence} from the {Chilean Case}. \emph{Social
+Justice Research} \textbf{2011}, \emph{24}, 314--340,
+doi:\href{https://doi.org/10.1007/s11211-011-0144-5}{10.1007/s11211-011-0144-5}.}
+
+\bibitem[\citeproctext]{ref-evans_family_2010}
+\CSLLeftMargin{28. }%
+\CSLRightInline{Evans, M.; Kelley, J.; Sikora, J.; Treiman, D. Family
+Scholarly Culture and Educational Success: {Books} and Schooling in 27
+Nations. \emph{Research in Social Stratification and Mobility}
+\textbf{2010}, \emph{28}, 171--197,
+doi:\href{https://doi.org/10.1016/j.rssm.2010.01.002}{10.1016/j.rssm.2010.01.002}.}
+
+\bibitem[\citeproctext]{ref-jasso_how_1999}
+\CSLLeftMargin{29. }%
+\CSLRightInline{Jasso, G. How {Much Injustice} Is {There} in the
+{World}? {Two New Justice Indexes}. \emph{American Sociological Review}
+\textbf{1999}, \emph{64}, 133--168,
+doi:\href{https://doi.org/10.1177/000312249906400110}{10.1177/000312249906400110}.}
+
+\bibitem[\citeproctext]{ref-shariff_income_2016}
+\CSLLeftMargin{30. }%
+\CSLRightInline{Shariff, A.F.; Wiwad, D.; Aknin, L.B. Income {Mobility
+Breeds Tolerance} for {Income Inequality}: {Cross-National} and
+{Experimental Evidence}. \emph{Perspectives on Psychological Science}
+\textbf{2016}, \emph{11}, 373--380,
+doi:\href{https://doi.org/10.1177/1745691616635596}{10.1177/1745691616635596}.}
+
+\bibitem[\citeproctext]{ref-centeno_arc_2012}
+\CSLLeftMargin{31. }%
+\CSLRightInline{Centeno, M.A.; Cohen, J.N. The {Arc} of {Neoliberalism}.
+\emph{Annual Review of Sociology} \textbf{2012}, \emph{38}, 317--340,
+doi:\href{https://doi.org/10.1146/annurev-soc-081309-150235}{10.1146/annurev-soc-081309-150235}.}
+
+\bibitem[\citeproctext]{ref-harvey_breve_2015}
+\CSLLeftMargin{32. }%
+\CSLRightInline{Harvey, D. \emph{{Breve historia del neoliberalismo}};
+Ediciones Akal: Madrid (Espa{ñ}a), 2015; ISBN 978-84-460-2517-7.}
+
+\bibitem[\citeproctext]{ref-jensen_worlds_2008}
+\CSLLeftMargin{33. }%
+\CSLRightInline{Jensen, C. Worlds of Welfare Services and Transfers.
+\emph{Journal of European Social Policy} \textbf{2008}, \emph{18},
+151--162,
+doi:\href{https://doi.org/10.1177/0958928707087591}{10.1177/0958928707087591}.}
+
+\bibitem[\citeproctext]{ref-stoy_worlds_2014}
+\CSLLeftMargin{34. }%
+\CSLRightInline{Stoy, V. Worlds of {Welfare Services}: {From Discovery}
+to {Exploration}. \emph{Social Policy \& Administration} \textbf{2014},
+\emph{48}, 343--360,
+doi:\href{https://doi.org/10.1111/spol.12006}{10.1111/spol.12006}.}
+
+\bibitem[\citeproctext]{ref-salamon_marketization_1993}
+\CSLLeftMargin{35. }%
+\CSLRightInline{Salamon, L.M. The {Marketization} of {Welfare}:
+{Changing Nonprofit} and {For-Profit Roles} in the {American Welfare
+State}. \emph{Social Service Review} \textbf{1993}, \emph{67}, 16--39,
+doi:\href{https://doi.org/10.1086/603963}{10.1086/603963}.}
+
+\bibitem[\citeproctext]{ref-sivesind_changing_2017}
+\CSLLeftMargin{36. }%
+\CSLRightInline{Sivesind, K.H.
+\href{https://doi.org/10.1007/978-3-319-55381-8_2}{The {Changing Roles}
+of {For-Profit} and {Nonprofit Welfare Provision} in {Norway}, {Sweden},
+and {Denmark}}. In \emph{Promoting {Active Citizenship}}; Sivesind,
+K.H., Saglie, J., Eds.; Springer International Publishing: Cham, 2017;
+pp. 33--74 ISBN 978-3-319-55380-1 978-3-319-55381-8.}
+
+\bibitem[\citeproctext]{ref-greitemeyer_subjective_2016}
+\CSLLeftMargin{37. }%
+\CSLRightInline{Greitemeyer, T.; Sagioglou, C. Subjective Socioeconomic
+Status Causes Aggression: {A} Test of the Theory of Social Deprivation.
+\emph{Journal of Personality and Social Psychology} \textbf{2016},
+\emph{111}, 178--194,
+doi:\href{https://doi.org/10.1037/pspi0000058}{10.1037/pspi0000058}.}
+
+\bibitem[\citeproctext]{ref-mishra_subjective_2015}
+\CSLLeftMargin{38. }%
+\CSLRightInline{Mishra, S.; Carleton, R.N. Subjective Relative
+Deprivation Is Associated with Poorer Physical and Mental Health.
+\emph{Social Science \& Medicine} \textbf{2015}, \emph{147}, 144--149,
+doi:\href{https://doi.org/10.1016/j.socscimed.2015.10.030}{10.1016/j.socscimed.2015.10.030}.}
+
+\bibitem[\citeproctext]{ref-smith_relative_2012}
+\CSLLeftMargin{39. }%
+\CSLRightInline{Smith, H.J.; Pettigrew, T.F.; Pippin, G.M.;
+Bialosiewicz, S. Relative {Deprivation}: {A Theoretical} and
+{Meta-Analytic Review}. \emph{Personality and Social Psychology Review}
+\textbf{2012}, \emph{16}, 203--232,
+doi:\href{https://doi.org/10.1177/1088868311430825}{10.1177/1088868311430825}.}
+
+\bibitem[\citeproctext]{ref-power_deprivationprotest_2018}
+\CSLLeftMargin{40. }%
+\CSLRightInline{Power, S.A. The {Deprivation-Protest Paradox}: {How} the
+{Perception} of {Unfair Economic Inequality Leads} to {Civic Unrest}.
+\emph{Current Anthropology} \textbf{2018}, \emph{59}, 765--789,
+doi:\href{https://doi.org/10.1086/700679}{10.1086/700679}.}
+
+\bibitem[\citeproctext]{ref-wilson_role_2003}
+\CSLLeftMargin{41. }%
+\CSLRightInline{Wilson, C. The {Role} of a {Merit Principle} in
+{Distributive Justice}. \emph{The Journal of Ethics} \textbf{2003},
+\emph{7}, 277--314,
+doi:\href{https://doi.org/10.1023/A:1024667228488}{10.1023/A:1024667228488}.}
+
+\bibitem[\citeproctext]{ref-bucca_merit_2016}
+\CSLLeftMargin{42. }%
+\CSLRightInline{Bucca, M. Merit and Blame in Unequal Societies:
+{Explaining Latin Americans}' Beliefs about Wealth and Poverty.
+\emph{Research in Social Stratification and Mobility} \textbf{2016},
+\emph{44}, 98--112,
+doi:\href{https://doi.org/10.1016/j.rssm.2016.02.005}{10.1016/j.rssm.2016.02.005}.}
+
+\bibitem[\citeproctext]{ref-chong_mystery_2008}
+\CSLLeftMargin{43. }%
+\CSLRightInline{Chong, A.; Ñopo, H. The {Mystery} of {Discrimination} in
+{Latin America}. \emph{Econom{í}a} \textbf{2008}, \emph{8}, 79--107,
+doi:\href{https://doi.org/10.1353/eco.0.0005}{10.1353/eco.0.0005}.}
+
+\bibitem[\citeproctext]{ref-torche_intergenerational_2014}
+\CSLLeftMargin{44. }%
+\CSLRightInline{Torche, F. Intergenerational {Mobility} and
+{Inequality}: {The Latin American Case}. \emph{Annual Review of
+Sociology} \textbf{2014}, \emph{40}, 619--642,
+doi:\href{https://doi.org/10.1146/annurev-soc-071811-145521}{10.1146/annurev-soc-071811-145521}.}
+
+\bibitem[\citeproctext]{ref-salgado_inequality_2023}
+\CSLLeftMargin{45. }%
+\CSLRightInline{Salgado, M.; Castillo, J. Inequality and
+{Stratification} in {Latin America}. In \emph{The {Oxford Handbook} of
+{Social Stratification}}; Gangl, M., Platt, L., Polavieja, J.G., van de
+Werfhorst, H.G., Eds.; Oxford University Press: 2023, 2023; p. 0 ISBN
+978-0-19-753948-4.}
+
+\bibitem[\citeproctext]{ref-molyneux_neoliberal_2008}
+\CSLLeftMargin{46. }%
+\CSLRightInline{Molyneux, M. The {``{Neoliberal Turn}''} and the {New
+Social Policy} in {Latin America}: {How Neoliberal}, {How New}?
+\emph{Development and Change} \textbf{2008}, \emph{39}, 775--797,
+doi:\href{https://doi.org/10.1111/j.1467-7660.2008.00505.x}{10.1111/j.1467-7660.2008.00505.x}.}
+
+\bibitem[\citeproctext]{ref-centrodeestudiospublicos_estudio_2024}
+\CSLLeftMargin{47. }%
+\CSLRightInline{Centro de Estudios Públicos \emph{Estudio {Nacional} de
+{Opini{ó}n P{ú}blica Base Consolidada}}; Centro de Estudios P{ú}blicos,
+2024;}
+
+\bibitem[\citeproctext]{ref-goldthorpe_myth_2003}
+\CSLLeftMargin{48. }%
+\CSLRightInline{Goldthorpe, J. The Myth of Education-Based Meritocracy.
+\emph{New Economy} \textbf{2003}, \emph{10}, 234--239,
+doi:\href{https://doi.org/10.1046/j.1468-0041.2003.00324.x}{10.1046/j.1468-0041.2003.00324.x}.}
+
+\bibitem[\citeproctext]{ref-castillo_multidimensional_2023}
+\CSLLeftMargin{49. }%
+\CSLRightInline{Castillo, J.C.; Iturra, J.; Maldonado, L.; Atria, J.;
+Meneses, F. A {Multidimensional Approach} for {Measuring Meritocratic
+Beliefs}: {Advantages}, {Limitations} and {Alternatives} to the {ISSP
+Social Inequality Survey}. \emph{International Journal of Sociology}
+\textbf{2023}, \emph{53}, 448--472,
+doi:\href{https://doi.org/10.1080/00207659.2023.2274712}{10.1080/00207659.2023.2274712}.}
+
+\bibitem[\citeproctext]{ref-mijs_unfulfillable_2016}
+\CSLLeftMargin{50. }%
+\CSLRightInline{Mijs, J. The {Unfulfillable Promise} of {Meritocracy}:
+{Three Lessons} and {Their Implications} for {Justice} in {Education}.
+\emph{Social Justice Research} \textbf{2016}, \emph{29}, 14--34,
+doi:\href{https://doi.org/10.1007/s11211-014-0228-0}{10.1007/s11211-014-0228-0}.}
+
+\bibitem[\citeproctext]{ref-mijs_paradox_2019}
+\CSLLeftMargin{51. }%
+\CSLRightInline{Mijs, J. The Paradox of Inequality: Income Inequality
+and Belief in Meritocracy Go Hand in Hand. \emph{Socio-Economic Review}
+\textbf{2019}, \emph{19}, 7--35,
+doi:\href{https://doi.org/10.1093/ser/mwy051}{10.1093/ser/mwy051}.}
+
+\bibitem[\citeproctext]{ref-duru-bellat_who_2012}
+\CSLLeftMargin{52. }%
+\CSLRightInline{Duru-Bellat, M.; Tenret, E. Who's for {Meritocracy}?
+{Individual} and {Contextual Variations} in the {Faith}.
+\emph{Comparative Education Review} \textbf{2012}, \emph{56}, 223--247,
+doi:\href{https://doi.org/10.1086/661290}{10.1086/661290}.}
+
+\bibitem[\citeproctext]{ref-lampert_meritocratic_2013}
+\CSLLeftMargin{53. }%
+\CSLRightInline{Lampert, K.
+\emph{\href{https://doi.org/10.1057/9781137324894}{Meritocratic
+{Education} and {Social Worthlessness}}}; Palgrave Macmillan UK: London,
+2013; ISBN 978-1-349-45908-7 978-1-137-32489-4.}
+
+\bibitem[\citeproctext]{ref-morris_representing_2022}
+\CSLLeftMargin{54. }%
+\CSLRightInline{Morris, J.; Reilly, J.; Paltsev, S.; Sokolov, A.; Cox,
+K. Representing {Socio}-{Economic Uncertainty} in {Human System Models}.
+\emph{Earth's Future} \textbf{2022}, \emph{10}, e2021EF002239,
+doi:\href{https://doi.org/10.1029/2021EF002239}{10.1029/2021EF002239}.}
+
+\bibitem[\citeproctext]{ref-frank_performance_2015}
+\CSLLeftMargin{55. }%
+\CSLRightInline{Frank, D.H.; Wertenbroch, K.; Maddux, W.W. Performance
+Pay or Redistribution? {Cultural} Differences in Just-World Beliefs and
+Preferences for Wage Inequality. \emph{Organizational Behavior and Human
+Decision Processes} \textbf{2015}, \emph{130}, 160--170,
+doi:\href{https://doi.org/10.1016/j.obhdp.2015.04.001}{10.1016/j.obhdp.2015.04.001}.}
+
+\bibitem[\citeproctext]{ref-almas_cutthroat_2020}
+\CSLLeftMargin{56. }%
+\CSLRightInline{Almås, I.; Cappelen, A.W.; Tungodden, B. Cutthroat
+{Capitalism} Versus {Cuddly Socialism}: {Are Americans More
+Meritocratic} and {Efficiency-Seeking} Than {Scandinavians}?
+\emph{Journal of Political Economy} \textbf{2020}, \emph{128},
+1753--1788, doi:\href{https://doi.org/10.1086/705551}{10.1086/705551}.}
+
+\bibitem[\citeproctext]{ref-barr_effect_2020}
+\CSLLeftMargin{57. }%
+\CSLRightInline{Barr, A.; Miller, L. The Effect of Education, Income
+Inequality and Merit on Inequality Acceptance. \emph{Journal of Economic
+Psychology} \textbf{2020}, \emph{80}, 102276,
+doi:\href{https://doi.org/10.1016/j.joep.2020.102276}{10.1016/j.joep.2020.102276}.}
+
+\bibitem[\citeproctext]{ref-evans_strong_2018}
+\CSLLeftMargin{58. }%
+\CSLRightInline{Evans, M.D.R.; Kelley, J. Strong {Welfare States Do Not
+Intensify Public Support} for {Income Redistribution}, but {Even Reduce
+It} Among the {Prosperous}: {A Multilevel Analysis} of {Public Opinion}
+in 30 {Countries}. \emph{Societies} \textbf{2018}, \emph{8}, 105,
+doi:\href{https://doi.org/10.3390/soc8040105}{10.3390/soc8040105}.}
+
+\bibitem[\citeproctext]{ref-mikus_children_2019}
+\CSLLeftMargin{59. }%
+\CSLRightInline{Mikus, K.; Tieben, N.; Schober, P.S. Children's
+Conversion of Cultural Capital into Educational Success: The Symbolic
+and Skill-Generating Functions of Cultural Capital. \emph{British
+Journal of Sociology of Education} \textbf{2019}, \emph{41}, 197--2017,
+doi:\href{https://doi.org/10.1080/01425692.2019.1677454}{10.1080/01425692.2019.1677454}.}
+
+\bibitem[\citeproctext]{ref-jack_no_2016}
+\CSLLeftMargin{60. }%
+\CSLRightInline{Jack, A.A. ({No}) {Harm} in {Asking}: {Class}, {Acquired
+Cultural Capital}, and {Academic Engagement} at an {Elite University}.
+\emph{Sociology of Education} \textbf{2016}, \emph{89}, 1--19,
+doi:\href{https://doi.org/10.1177/0038040715614913}{10.1177/0038040715614913}.}
+
+\bibitem[\citeproctext]{ref-khan_privilege_2011}
+\CSLLeftMargin{61. }%
+\CSLRightInline{Khan, S. \emph{Privilege: {The Making} of an {Adolescent
+Elite} at {St}. {Paul}'s {School}}; Princeton University Press:
+Princeton, N.J., 2011;}
+
+\bibitem[\citeproctext]{ref-goudeau_hidden_2017}
+\CSLLeftMargin{62. }%
+\CSLRightInline{Goudeau, S.; Croizet, J.-C. Hidden {Advantages} and
+{Disadvantages} of {Social Class}: {How Classroom Settings Reproduce
+Social Inequality} by {Staging Unfair Comparison}. \emph{Psychological
+Science} \textbf{2017}, \emph{28}, 162--170,
+doi:\href{https://doi.org/10.1177/0956797616676600}{10.1177/0956797616676600}.}
+
+\bibitem[\citeproctext]{ref-darnon_where_2018}
+\CSLLeftMargin{63. }%
+\CSLRightInline{Darnon, C.; Wiederkehr, V.; Dompnier, B.; Martinot, D.
+{``{Where} There Is a Will, There Is a Way''}: {Belief} in School
+Meritocracy and the Social-Class Achievement Gap. \emph{British Journal
+of Social Psychology} \textbf{2018}, \emph{57}, 250--262,
+doi:\href{https://doi.org/10.1111/bjso.12214}{10.1111/bjso.12214}.}
+
+\bibitem[\citeproctext]{ref-resh_sense_2010}
+\CSLLeftMargin{64. }%
+\CSLRightInline{Resh, N. Sense of Justice about Grades in School: Is It
+Stratified Like Academic Achievement? \emph{Social Psychology of
+Education} \textbf{2010}, \emph{13}, 313--329,
+doi:\href{https://doi.org/10.1007/s11218-010-9117-z}{10.1007/s11218-010-9117-z}.}
+
+\bibitem[\citeproctext]{ref-resh_sense_2014}
+\CSLLeftMargin{65. }%
+\CSLRightInline{Resh, N.; Sabbagh, C. Sense of Justice in School and
+Civic Attitudes. \emph{Social Psychology of Education} \textbf{2014},
+\emph{17}, 51--72,
+doi:\href{https://doi.org/10.1007/s11218-013-9240-8}{10.1007/s11218-013-9240-8}.}
+
+\bibitem[\citeproctext]{ref-kohn_social_1963}
+\CSLLeftMargin{66. }%
+\CSLRightInline{Kohn, M.L. Social {Class} and {Parent-Child
+Relationships}: {An Interpretation}. \emph{American Journal of
+Sociology} \textbf{1963}, \emph{68}, 471--480,
+doi:\href{https://doi.org/10.1086/223403}{10.1086/223403}.}
+
+\bibitem[\citeproctext]{ref-kohn_class_1969}
+\CSLLeftMargin{67. }%
+\CSLRightInline{Kohn, M.L.; Schooler, C. Class, {Occupation}, and
+{Orientation}. \emph{American Sociological Review} \textbf{1969},
+\emph{34}, 659--678,
+doi:\href{https://doi.org/10.2307/2092303}{10.2307/2092303}.}
+
+\bibitem[\citeproctext]{ref-acemoglu_obedience_2021}
+\CSLLeftMargin{68. }%
+\CSLRightInline{Acemoglu, D.
+\emph{\href{https://doi.org/10.3386/w29125}{Obedience in the {Labor
+Market} and {Social Mobility}: {A Socio-Economic Approach}}}; National
+Bureau of Economic Research: Cambridge, MA, 2021; p. w29125;}
+
+\bibitem[\citeproctext]{ref-iacoviello_collectivism_2019}
+\CSLLeftMargin{69. }%
+\CSLRightInline{Iacoviello, V.; Lorenzi-Cioldi, F. Collectivism and
+{Individualism} in {Status Hierarchies}: {Socialization} and {Social
+Identity Explanations}. \emph{International Review of Social Psychology}
+\textbf{2019}, \emph{32}, 15,
+doi:\href{https://doi.org/10.5334/irsp.285}{10.5334/irsp.285}.}
+
+\bibitem[\citeproctext]{ref-almas_fairness_2017}
+\CSLLeftMargin{70. }%
+\CSLRightInline{Almås, I.; Cappelen, A.W.; Salvanes, K.G.; Sørensen,
+E.Ø.; Tungodden, B. Fairness and Family Background. \emph{Politics,
+Philosophy \& Economics} \textbf{2017}, \emph{16}, 117--131,
+doi:\href{https://doi.org/10.1177/1470594x15618966}{10.1177/1470594x15618966}.}
+
+\bibitem[\citeproctext]{ref-jonsson_institutional_2015}
+\CSLLeftMargin{71. }%
+\CSLRightInline{Jonsson, A.-C.; Beach, D. Institutional Discrimination:
+{Stereotypes} and Social Reproduction of {``Class''} in the {Swedish}
+Upper-Secondary School. \emph{Social Psychology of Education}
+\textbf{2015}, \emph{18}, 703--717,
+doi:\href{https://doi.org/10.1007/s11218-014-9279-1}{10.1007/s11218-014-9279-1}.}
+
+\bibitem[\citeproctext]{ref-jost_attitudinal_2000}
+\CSLLeftMargin{72. }%
+\CSLRightInline{Jost, J.T.; Burgess, D. Attitudinal {Ambivalence} and
+the {Conflict} Between {Group} and {System Justification Motives} in
+{Low Status Groups}. \emph{Personality and Social Psychology Bulletin}
+\textbf{2000}, \emph{26}, 293--305,
+doi:\href{https://doi.org/10.1177/0146167200265003}{10.1177/0146167200265003}.}
+
+\bibitem[\citeproctext]{ref-gottfried_autocorrelation_2022}
+\CSLLeftMargin{73. }%
+\CSLRightInline{Gottfried, J.; Ježek, S.; Králová, M.; Řiháček, T.
+Autocorrelation {Screening}: {A Potentially Efficient Method} for
+{Detecting Repetitive Response Patterns} in {Questionnaire Data}.
+\textbf{2022},
+doi:\href{https://doi.org/10.7275/VYXB-GT24}{10.7275/VYXB-GT24}.}
+
+\bibitem[\citeproctext]{ref-bell_fixed_2019}
+\CSLLeftMargin{74. }%
+\CSLRightInline{Bell, A.; Fairbrother, M.; Jones, K. Fixed and Random
+Effects Models: Making an Informed Choice. \emph{Quality \& Quantity}
+\textbf{2019}, \emph{53}, 1051--1074,
+doi:\href{https://doi.org/10.1007/s11135-018-0802-x}{10.1007/s11135-018-0802-x}.}
+
+\bibitem[\citeproctext]{ref-hox_multilevel_2010}
+\CSLLeftMargin{75. }%
+\CSLRightInline{Hox, J.; Moerbeek, M.; Schoot, R. van de
+\emph{\href{https://doi.org/10.4324/9780203852279}{Multilevel
+{Analysis}: {Techniques} and {Applications}, {Second Edition}}}; 2nd
+ed.; Routledge: New York, 2010; ISBN 978-0-203-85227-9.}
+
+\bibitem[\citeproctext]{ref-kim_socioeconomic_2020}
+\CSLLeftMargin{76. }%
+\CSLRightInline{Kim, C.-Y.; Gewirtz, S. Socioeconomic Inequality and
+Adjustments in Children's Perceptions of Their Agency as They Age in
+{South Korea}. \emph{Children \& Society} \textbf{2020}, \emph{34},
+371--391,
+doi:\href{https://doi.org/10.1111/chso.12372}{10.1111/chso.12372}.}
+
+\bibitem[\citeproctext]{ref-rizzo_children_2020}
+\CSLLeftMargin{77. }%
+\CSLRightInline{Rizzo, M.T.; Killen, M. Children's Evaluations of
+Individually and Structurally Based Inequalities: {The} Role of Status.
+\emph{Developmental Psychology} \textbf{2020}, \emph{56}, 2223--2235,
+doi:\href{https://doi.org/10.1037/dev0001118}{10.1037/dev0001118}.}
+
+\bibitem[\citeproctext]{ref-gonthier_parallel_2017}
+\CSLLeftMargin{78. }%
+\CSLRightInline{Gonthier, F. Parallel Publics? {Support} for Income
+Redistribution in Times of Economic Crisis. \emph{European Journal of
+Political Research} \textbf{2017}, \emph{56}, 92--114,
+doi:\href{https://doi.org/10.1111/1475-6765.12168}{10.1111/1475-6765.12168}.}
+
+\bibitem[\citeproctext]{ref-runst_does_2018}
+\CSLLeftMargin{79. }%
+\CSLRightInline{Runst, P. Does {Immigration Affect Demand} for
+{Redistribution}? -- {An Experimental Design}. \emph{German Economic
+Review} \textbf{2018}, \emph{19}, 383--400,
+doi:\href{https://doi.org/10.1111/geer.12133}{10.1111/geer.12133}.}
+
+\bibitem[\citeproctext]{ref-bullock_education_2021}
+\CSLLeftMargin{80. }%
+\CSLRightInline{Bullock, J.G. Education and {Attitudes} Toward
+{Redistribution} in the {United States}. \emph{British Journal of
+Political Science} \textbf{2021}, \emph{51}, 1230--1250,
+doi:\href{https://doi.org/10.1017/S0007123419000504}{10.1017/S0007123419000504}.}
+
+\end{CSLReferences}
+
 \end{document}

### 2024-10-02 submission/MDPI/paper-blinded.tex
Update revised version MPI Submit

@@
 \end{ideuno}
+
 % -------------------
@@
 
-\bibitem[\citeproctext]{ref-chafel_schooling_1997}
+\bibitem[\citeproctext]{ref-lindh_public_2015}
 \CSLLeftMargin{11. }%
-\CSLRightInline{Chafel, J.A. Schooling, the {Hidden Curriculum}, and
-{Children}'s {Conceptions} of {Poverty}. \emph{Social Policy Report}
-\textbf{1997}, \emph{11}, 1--28,
-doi:\href{https://doi.org/10.1002/j.2379-3988.1997.tb00004.x}{10.1002/j.2379-3988.1997.tb00004.x}.}
+\CSLRightInline{Lindh, A. Public {Opinion} Against {Markets}?
+{Attitudes} Towards {Market Distribution} of {Social Services} -- {A
+Comparison} of 17 {Countries}. \emph{Social Policy \& Administration}
+\textbf{2015}, \emph{49}, 887--910,
+doi:\href{https://doi.org/10.1111/spol.12105}{10.1111/spol.12105}.}
+
+\bibitem[\citeproctext]{ref-castillo_meritocracia_2019}
+\CSLLeftMargin{12. }%
+\CSLRightInline{Castillo, J.C.; Torres, A.; Atria, J.; Maldonado, L.
+Meritocracia y Desigualdad Econ{ó}mica: {Percepciones}, Preferencias e
+Implicancias. \emph{Revista Internacional de Sociolog{í}a}
+\textbf{2019}, \emph{77}, 117,
+doi:\href{https://doi.org/10.3989/ris.2019.77.1.17.114}{10.3989/ris.2019.77.1.17.114}.}
+
+\bibitem[\citeproctext]{ref-reynolds_perceptions_2014}
+\CSLLeftMargin{13. }%
+\CSLRightInline{Reynolds, J.; Xian, H. Perceptions of Meritocracy in the
+Land of Opportunity. \emph{Research in Social Stratification and
+Mobility} \textbf{2014}, \emph{36}, 121--137,
+doi:\href{https://doi.org/10.1016/j.rssm.2014.03.001}{10.1016/j.rssm.2014.03.001}.}
+
+\bibitem[\citeproctext]{ref-garcia-sierra_dark_2023}
+\CSLLeftMargin{14. }%
+\CSLRightInline{García-Sierra, A. The Dark Side of Meritocratic Beliefs:
+{Is} Believing in Meritocracy Detrimental to Individuals from Low
+Socioeconomic Backgrounds? \emph{Social Justice Research} \textbf{2023},
+\emph{36}, 385--409,
+doi:\href{https://doi.org/10.1007/s11211-023-00413-x}{10.1007/s11211-023-00413-x}.}
 
 \bibitem[\citeproctext]{ref-flores_top_2020}
-\CSLLeftMargin{12. }%
+\CSLLeftMargin{15. }%
 \CSLRightInline{Flores, I.; Sanhueza, C.; Atria, J.; Mayer, R. Top
@@
 \bibitem[\citeproctext]{ref-chancel_world_2022}
-\CSLLeftMargin{13. }%
+\CSLLeftMargin{16. }%
 \CSLRightInline{Chancel, L.; Piketty, T.; Saez, E.; Zucman, G. World
@@
 \bibitem[\citeproctext]{ref-ffrench-davis_reformas_2018}
-\CSLLeftMargin{14. }%
+\CSLLeftMargin{17. }%
 \CSLRightInline{Ffrench-Davis, R. \emph{{Reformas econ{ó}micas en Chile
@@
 \bibitem[\citeproctext]{ref-boccardo_30_2020}
-\CSLLeftMargin{15. }%
+\CSLLeftMargin{18. }%
 \CSLRightInline{Boccardo, G. \emph{30 a{ñ}os de Privatizaciones En
@@
 \bibitem[\citeproctext]{ref-observatoriosocial_estadisticas_2024}
-\CSLLeftMargin{16. }%
+\CSLLeftMargin{19. }%
 \CSLRightInline{Social, O. \emph{Estad{í}sticas {Encuesta Casen} 2022,
@@
 \bibitem[\citeproctext]{ref-superintendenciadepensiones_estadisticas_2024}
-\CSLLeftMargin{17. }%
+\CSLLeftMargin{20. }%
 \CSLRightInline{Superintendencia de Pensiones \emph{Estad{í}sticas de
@@
 \bibitem[\citeproctext]{ref-ministeriodeeducacion_resumen_2023}
-\CSLLeftMargin{18. }%
+\CSLLeftMargin{21. }%
 \CSLRightInline{Ministerio de Educación \emph{Resumen de Matr{í}cula Por
@@
 \bibitem[\citeproctext]{ref-boltanski_new_2005}
-\CSLLeftMargin{19. }%
+\CSLLeftMargin{22. }%
 \CSLRightInline{Boltanski, L.; Chiapello, E. \emph{The New Spirit of
@@
 \bibitem[\citeproctext]{ref-streeck_citizens_2012}
-\CSLLeftMargin{20. }%
+\CSLLeftMargin{23. }%
 \CSLRightInline{Streeck, W. Citizens as {Customers}: {Considerations} on
@@
 \bibitem[\citeproctext]{ref-kluegel_beliefs_1987}
-\CSLLeftMargin{21. }%
+\CSLLeftMargin{24. }%
 \CSLRightInline{Kluegel, J.R.; Smith, E.R. \emph{Beliefs about
@@
 \bibitem[\citeproctext]{ref-wegener_dominant_1995}
-\CSLLeftMargin{22. }%
+\CSLLeftMargin{25. }%
 \CSLRightInline{Wegener, B.; Liebig, S. Dominant {Ideologies} and the
@@
 \bibitem[\citeproctext]{ref-homan_being_2017}
-\CSLLeftMargin{23. }%
+\CSLLeftMargin{26. }%
 \CSLRightInline{Homan, P.; Valentino, L.; Weed, E. Being and {Becoming
@@
 \bibitem[\citeproctext]{ref-castillo_legitimacy_2011}
-\CSLLeftMargin{24. }%
+\CSLLeftMargin{27. }%
 \CSLRightInline{Castillo, J.C. Legitimacy of {Inequality} in a {Highly
@@
 \bibitem[\citeproctext]{ref-evans_family_2010}
-\CSLLeftMargin{25. }%
+\CSLLeftMargin{28. }%
 \CSLRightInline{Evans, M.; Kelley, J.; Sikora, J.; Treiman, D. Family
@@
 \bibitem[\citeproctext]{ref-jasso_how_1999}
-\CSLLeftMargin{26. }%
+\CSLLeftMargin{29. }%
 \CSLRightInline{Jasso, G. How {Much Injustice} Is {There} in the
@@
 \bibitem[\citeproctext]{ref-shariff_income_2016}
-\CSLLeftMargin{27. }%
+\CSLLeftMargin{30. }%
 \CSLRightInline{Shariff, A.F.; Wiwad, D.; Aknin, L.B. Income {Mobility
@@
 \bibitem[\citeproctext]{ref-centeno_arc_2012}
-\CSLLeftMargin{28. }%
+\CSLLeftMargin{31. }%
 \CSLRightInline{Centeno, M.A.; Cohen, J.N. The {Arc} of {Neoliberalism}.
@@
 \bibitem[\citeproctext]{ref-harvey_breve_2015}
-\CSLLeftMargin{29. }%
+\CSLLeftMargin{32. }%
 \CSLRightInline{Harvey, D. \emph{{Breve historia del neoliberalismo}};
@@
 \bibitem[\citeproctext]{ref-jensen_worlds_2008}
-\CSLLeftMargin{30. }%
+\CSLLeftMargin{33. }%
 \CSLRightInline{Jensen, C. Worlds of Welfare Services and Transfers.
@@
 \bibitem[\citeproctext]{ref-stoy_worlds_2014}
-\CSLLeftMargin{31. }%
+\CSLLeftMargin{34. }%
 \CSLRightInline{Stoy, V. Worlds of {Welfare Services}: {From Discovery}
@@
 \bibitem[\citeproctext]{ref-salamon_marketization_1993}
-\CSLLeftMargin{32. }%
+\CSLLeftMargin{35. }%
 \CSLRightInline{Salamon, L.M. The {Marketization} of {Welfare}:
@@
 \bibitem[\citeproctext]{ref-sivesind_changing_2017}
-\CSLLeftMargin{33. }%
+\CSLLeftMargin{36. }%
 \CSLRightInline{Sivesind, K.H.
@@
 
-\bibitem[\citeproctext]{ref-lindh_public_2015}
-\CSLLeftMargin{34. }%
-\CSLRightInline{Lindh, A. Public {Opinion} Against {Markets}?
-{Attitudes} Towards {Market Distribution} of {Social Services} -- {A
-Comparison} of 17 {Countries}. \emph{Social Policy \& Administration}
-\textbf{2015}, \emph{49}, 887--910,
-doi:\href{https://doi.org/10.1111/spol.12105}{10.1111/spol.12105}.}
-
 \bibitem[\citeproctext]{ref-greitemeyer_subjective_2016}
-\CSLLeftMargin{35. }%
+\CSLLeftMargin{37. }%
 \CSLRightInline{Greitemeyer, T.; Sagioglou, C. Subjective Socioeconomic
@@
 \bibitem[\citeproctext]{ref-mishra_subjective_2015}
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{Mishra, S.; Carleton, R.N. Subjective Relative
@@
 \bibitem[\citeproctext]{ref-smith_relative_2012}
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{Smith, H.J.; Pettigrew, T.F.; Pippin, G.M.;
@@
 \bibitem[\citeproctext]{ref-power_deprivationprotest_2018}
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{Power, S.A. The {Deprivation-Protest Paradox}: {How} the
@@
 \bibitem[\citeproctext]{ref-wilson_role_2003}
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{Wilson, C. The {Role} of a {Merit Principle} in
@@
 \bibitem[\citeproctext]{ref-bucca_merit_2016}
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{42. }%
 \CSLRightInline{Bucca, M. Merit and Blame in Unequal Societies:
@@
 \bibitem[\citeproctext]{ref-chong_mystery_2008}
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{Chong, A.; Ñopo, H. The {Mystery} of {Discrimination} in
@@
 \bibitem[\citeproctext]{ref-torche_intergenerational_2014}
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{Torche, F. Intergenerational {Mobility} and
@@
 \bibitem[\citeproctext]{ref-salgado_inequality_2023}
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Salgado, M.; Castillo, J. Inequality and
@@
 \bibitem[\citeproctext]{ref-molyneux_neoliberal_2008}
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{Molyneux, M. The {``{Neoliberal Turn}''} and the {New
@@
 \bibitem[\citeproctext]{ref-centrodeestudiospublicos_estudio_2024}
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{Centro de Estudios Públicos \emph{Estudio {Nacional} de
@@
 \bibitem[\citeproctext]{ref-goldthorpe_myth_2003}
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Goldthorpe, J. The Myth of Education-Based Meritocracy.
@@
 \bibitem[\citeproctext]{ref-castillo_multidimensional_2023}
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{49. }%
 \CSLRightInline{Castillo, J.C.; Iturra, J.; Maldonado, L.; Atria, J.;
@@
 \bibitem[\citeproctext]{ref-mijs_unfulfillable_2016}
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{50. }%
 \CSLRightInline{Mijs, J. The {Unfulfillable Promise} of {Meritocracy}:
@@
 \bibitem[\citeproctext]{ref-mijs_paradox_2019}
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{51. }%
 \CSLRightInline{Mijs, J. The Paradox of Inequality: Income Inequality
@@
 \bibitem[\citeproctext]{ref-duru-bellat_who_2012}
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{52. }%
 \CSLRightInline{Duru-Bellat, M.; Tenret, E. Who's for {Meritocracy}?
@@
 
-\bibitem[\citeproctext]{ref-castillo_meritocracia_2019}
-\CSLLeftMargin{51. }%
-\CSLRightInline{Castillo, J.C.; Torres, A.; Atria, J.; Maldonado, L.
-Meritocracia y Desigualdad Econ{ó}mica: {Percepciones}, Preferencias e
-Implicancias. \emph{Revista Internacional de Sociolog{í}a}
-\textbf{2019}, \emph{77}, 117,
-doi:\href{https://doi.org/10.3989/ris.2019.77.1.17.114}{10.3989/ris.2019.77.1.17.114}.}
-
 \bibitem[\citeproctext]{ref-lampert_meritocratic_2013}
-\CSLLeftMargin{52. }%
+\CSLLeftMargin{53. }%
 \CSLRightInline{Lampert, K.
@@
 2013; ISBN 978-1-349-45908-7 978-1-137-32489-4.}
-
-\bibitem[\citeproctext]{ref-reynolds_perceptions_2014}
-\CSLLeftMargin{53. }%
-\CSLRightInline{Reynolds, J.; Xian, H. Perceptions of Meritocracy in the
-Land of Opportunity. \emph{Research in Social Stratification and
-Mobility} \textbf{2014}, \emph{36}, 121--137,
-doi:\href{https://doi.org/10.1016/j.rssm.2014.03.001}{10.1016/j.rssm.2014.03.001}.}
 
@@
 \end{CSLReferences}
-
-\newpage{}
 
